### 2015-05-22 participant-table.tex
Initial proposal template cribbed from David

% Generated by bin/config-participant-table on Fri Jan 17 21:40:42 EST 2014
\newlength{\longestpientry}
\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Hasan Abbasi} (\emph{habbasi@ornl.gov})}}}
\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Karsten Schwan} (\emph{schwan@cc.gatech.edu})}}}
\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Allen Malony} (\emph{malony@cs.uoregon.edu})}}}
\begin{tabularx}{\textwidth}{l>{\raggedright\arraybackslash}X}
\emph{Institution} & \makebox[\longestpientry][l]{\emph{Principal Investigator (Email)}} \quad \emph{Additional Senior Personnel}
\\
\textbf{\gatech} & \makebox[\longestpientry][l]{\textbf{Karsten Schwan} (\emph{schwan@cc.gatech.edu})}\hspace*{1em}\\
\textbf{\oregon} & \makebox[\longestpientry][l]{\textbf{Allen Malony} (\emph{malony@cs.uoregon.edu})}\hspace*{1em}\\
\textbf{\ornl} & \makebox[\longestpientry][l]{\textbf{Hasan Abbasi} (\emph{habbasi@ornl.gov})}\hspace*{1em}\
\end{tabularx}

### 2015-06-01 abstract.tex

% Special formatting for the abstract
\titleformat{\chapter}[block]
%{\filcenter\normalfont\huge\bfseries}{}{0pt}{\huge}
{\filcenter\normalfont\LARGE\bfseries}{}{0pt}{\LARGE}
%{\filcenter\normalfont\Large\bfseries}{}{0pt}{\Large}
%{\filcenter\normalfont\Large\bfseries}{}{0pt}{\large}
\titlespacing*{\chapter}{0pt}{-0.35in}{0pt} %%% * kill first para indentation

\chapter*{\MakeUppercase{\proposaltitle}}
\phantomsection
\addstarredchapter{Abstract}
%%\adjustmtc
\markboth{Abstract}{}
\label{sec:abstract}

% Restore standard formatting for chapter titles
\titleformat{\chapter}[block]
{\normalfont\Huge\bfseries}{\thechapter.}{10pt}{\Huge}
\titlespacing*{\chapter}{0pt}{-0.25in}{0pt} %%% * kill first para indentation

%% \ifthenelse{\equal{\proposalcontent}{abstract}}{ % Standalone abstract-only
%%    \typeout{output-controls: empty pagestyle for standalone abstract}
%%    \thispagestyle{empty}
%% }{}

% Vertical spacing is careful controlled in order to maximize space
% available while maintaining a reasonable experience.

\vspace{-\abovedisplayskip}
\medskip

% Lead PI slug is centered internally
\textbf{\input{leadpi-slug}}

\vspace{-\belowdisplayskip}
\vspace{-\abovedisplayskip}
\smallskip

\begin{center}
\input{participant-table}
\end{center}

\vspace{-\belowdisplayskip}
\vspace{-\abovedisplayskip}
\medskip

\begin{center}
\textbf{ABSTRACT}
\end{center}

\vspace{-\belowdisplayskip}

% Even with the space manipulation above, we need to cheat a little
\enlargethispage{2\baselineskip}

\noindent
Stuff
\vfill

### 2015-06-01 narrative/intro.tex

We propose to explore the development of a performance engineering data pipeline that combines extraction of fine grained monitoring data, collation and curation of this data into a usable distributed object store, and in-situ analysis and visualization of the data to enable performance understanding at extreme scale for application developers and scientific users. Additionally we will explore the characterization of end-to-end pipelines for typical scientific workflows to better identify key optimization points and explore both application and system level optimization models for these data-centric tasks. 
The inherent complexity of extreme scale systems, combining heterogeneous processing architectures, deep memory hierarchy and severe power constraints, provides a challenge for performance engineering…%motivation

### 2015-06-03 LOI/inc2.tex
LOI ... begin

\documentclass[11pt,letterpaper]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pagestyle{plain}                                                      %%
%%%%%%%%%% EXACT 1in MARGINS %%%%%%%                                   %%
\setlength{\textwidth}{6.5in}     %%                                   %%
\setlength{\oddsidemargin}{0in}   %% (It is recommended that you       %%
\setlength{\evensidemargin}{0in}  %%  not change these parameters,     %%
\setlength{\textheight}{9.0in}    %%  at the risk of having your       %%
\setlength{\topmargin}{0in}       %%  proposal dismissed on the basis  %%
\setlength{\headheight}{0in}      %%  of incorrect formatting!!!)      %%
\setlength{\headsep}{0in}         %%                                   %%
\setlength{\footskip}{.5in}       %%                                   %%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[pdftex]{graphicx}
\usepackage{color}
\usepackage{url}
\usepackage{tabularx}
\usepackage{tikz}

% From PPoPP
\usepackage{amsmath}
\usepackage{amsfonts}
%\usepackage{graphicx}
%\usepackage{xspace}
\usepackage{verbatim}
%\usepackage{listings}
\usepackage{multirow}
\usepackage{subfigure}
\usepackage{pdfpages}
%% % Tweak spacing to fit in page limit if needed
%% % ============================================
%% % gap between text and figs/tables:
%% \addtolength\textfloatsep{-0.2in}
%% % gap between figs/tables and other figs/tables
%% \addtolength\floatsep{-0.1in}
%% % gap between figure and caption
%% \addtolength\abovecaptionskip{-0.05in}
%% \addtolength\intextsep{0in}

\hyphenation{ }

\setlength{\parindent}{0.5cm}

\newif\ifdraft
% comment out the next line to turn off comments
%\drafttrue

\ifdraft
  \definecolor{darkgreen}{rgb}{0,0.5,0}
  \newcommand{\manish}[1]{ {\it \color{red} \{#1 -Tahsin\}}}
  \newcommand{\hasan}[1]{ {\textcolor{blue} { #1 -Hasan }}}
  \definecolor{orange}{rgb}{0.7,0.5,0.0}
  \newcommand{\klasky}[1]{{\textcolor{orange}{ #1 -Scott }}}
  % Red star denotes items that need further work or discussion
  \newcommand{\TODO}[1]{{\textcolor{red}{ TO DO: #1 }}}
\else
  \newcommand{\klasky}[1]{}
  \newcommand{\hasan}[1]{}
\fi

%\newcommand{\ititle}{\textsc{\textbf{HESK}}}
\newcommand{\ititle}{\textsc{HESK}}
\newcommand{\insitu}{\textit{in situ }}

\let\oldenumerate\enumerate
\renewcommand{\enumerate}{
      \oldenumerate
      \setlength{\itemsep}{1pt}
      \setlength{\parskip}{0pt}
      \setlength{\parsep}{0pt}
}

% A definition we do not want the reader to forget
\newcommand{\defn}[1] {\textbf{\textit{#1}}}

\definecolor{teal}{rgb}{0.06,0.3,0.3}
\definecolor{maroon}{rgb}{0.5,0.0,0.25}
\definecolor{darkblue}{rgb}{0.0,0.2,0.75}
\definecolor{darkred}{rgb}{0.7,0.0,0.0}
\definecolor{darkgreen2}{rgb}{0,0.35,0}

\newcommand*\circled[1]{\tikz[baseline=(char.base)]{
  \node[shape=circle,draw,inner sep=2pt] (char) {#1};}}

\begin{document}
\begin{comment}
\noindent
\textbf{Pre-proposal Cover Sheet}
\newline
\vskip .1in
Storage Systems and Input/Output for Extreme Scale Science 
Scale 2 (LAB 14-1043)

\vskip .3in

\noindent
\textbf{Project title:}
\vskip .1in
Hierarchal Extreme Scale Knowledge Management

\vskip .3in

\noindent
\textbf{Principal investigator:}
\vskip .1in

Scott Klasky klasky@ornl.gov

\vskip .2in

\noindent
\textbf{Co-principal investigator:}
\vskip .1in

Manish Parashar, Rutgers, parashar@rutgers.edu, 732-445-5388
Carlson Malzahn, UCSC
Jay Lofstead, Sandia

\vskip .2in

\noindent
\textbf{Senior Personnel:}
\vskip .1in
Hasan Abbasi, ORNL
Mark Ainsworth, ORNL
Matthew Curry, Sandia
Qing Liu, ORNL
 Lee Ward, Sandia

\begin{tabular}{| l| r| r| r| r| }
\hline
  \emph{Institution} & \emph{Year 1} & \emph{Year 2} & \emph{Year 3} \\
\hline
Oak Ridge National Laboratory & \$350,000 & \$350,000 & \$350,000	\\
\hline
  Argonne National Laboratory & \$350,000 & \$350,000 & \$350,000 \\
\hline
  Lawrence Berkeley Laboratory & \$300,000 & \$300,000 & \$300,000\\
\hline
  Brookhaven National Laboratory & \$150,000 & \$150,000 & \$150,000 \\
\hline
  Rutgers University & \$175,000 & \$175,000 & \$175,000 \\
\hline
  Stony Brook University & \$175,000 & \$175,000 & \$175,000 \\
\hline
  Total & \$1,500,000 & \$1,500,000 & \$1,500,000 \\
\hline
\end{tabular}
\newpage

\end{comment}
\includepdf[pages={1}]{ornl-cover.pdf}
\pagebreak

\author{
  \IEEEauthorblockN {
    PI: Scott Klasky, % \IEEEauthorrefmark{1},
    Co-PI Manish Parashar % \IEEEauthorrefmark{1}
  }
  \IEEEauthorblockA {
    % \IEEEauthorrefmark{1}
    Mathematics and Computer Science Division,
    Argonne National Laboratory,
    Argonne, IL, USA}
}

%\maketitle

% Heilmeyer Points: from http://www.csee.umbc.edu/~finin/home/heilmeyerCatechism.html
% 
% \begin{enumerate}

% \item What is the problem, why is it hard?

% \item How is it solved today?
%  Fig A: much

% \item What is the new technical idea; why can we succeed now?
%  Integration of proven components into an architecure that solves
%  problems. New research will resolve remaining technical hurdles.

% \item What is the impact if successful?
%  Users and developers will be able to ...

% \item How will the program be organized?
%  Participants and focus areas

% \item How will intermediate results be generated?
%  Toolkit snapshots and application

% \item How will you measure progress?
%  Provide metrics

% \item What will it cost?
%  Cover sheet

% \end{enumerate}
%

\underline{\textbf{Executive Summary:}} Exascale scientific discovery will
introduce more complex hardware, and many simulations will be bottlenecked 
without sufficient new research into manging and storing the large data which will be
produced during the simulation, and analzyed for months after the simulations.

Our goal is to explore and address
the multi-tier challenges that are faced by scientists in creating and
managing and storing their data  to expedite insights into
mission critical scientific processes in exascale computing.
%
We propose a research program that aims to explore a hiearchical organization and storage infstraturure which will facilitiate 
how we data can be written and read efficiently. In order to help us explore our many research issues, we will build upon the
sucess of our middleware system, ADIOS, our multi-tier storage system, Sirocco, and our distributed storage system, Cepth.

%
We will also study ??
%
Finally, we will, jointly with application partners, study the behvior of how data gets written and read from the storage layers, and
what key inishgts can be used in the next generation LCF systems, along with future exascale systems.

\underline{\textbf{Key Technical Contribution:}} The primary contribution of
our research are insights into how to build a hiearchical organization and storage
of massive scientific data sets generated from exascale computations along with 
the necessary information needed for advanced data validation and exploration from 
experimental and observational data. 
We are fundamentally trying to address the fundmantal questions : 1)  How can we 
place and mangage massive scienfici data across all of the tiers of the storage and memory
hiearchy? 2) What are the proper semantics needed in order to help bring knowledge from the applications
to the middleware layer. 3) What information from the storage layer can be exposed to the middleweare such
that the placement of information can be agreed upon from what the application requires and what the
system resources are available.

Our approach will allow users to ``plug-in'' techniques to classify data, not as bytes but as motifs, where
we can understand relationships between objects (into data models) and relationships of data in variables.
This will allow us to adapt data from a variable into the various storage tiers. Rather than focus on simple data
compression, we will incoroarte a multitude of techniques to reduce data on the faster storage tiers, and keep the less
reduced information on the lower tiers.

We wish to support additional data access modes as well. First, within a given
accuracy bounds, offer a mode for recomputation from data stored in a ``fast''
tier to reduce data latencies. Second, exploratory analysis operations
frequently entail an overall data set view followed by targeted data
exploration based on identified features. We will offer support for a
configurable data access mode to support these sorts of analysis accesses. An
``overview'' access mode that gives a quick, approximate within error bounds,
data view that can guide feature selection offering rapid coarse-grained data
exploration without requiring loading the complete, detailed data set from
storage. Based on the granularity requested, the accuracy and size of the data
returned can be adjusted. At the most extreme setting, the original data can
be retrieved at the time cost of moving the potentially huge data quantity.
Third, to ensure available storage for subsequent operations, we will offer
automatic data migration based on user annotations for required data lifetimes
using monitoring and learning techniques. Unlike existing approaches, this will
be tempered both by the user annotations and through learned access patterns.
While past access patterns may not indicate future access because the
simulation run purpose may have changed, we are focused on scalability where
runs are subsequently larger as the simulation prepares for a capability run.
By learning from the output and access patterns during this run sequence, we
can accurate decide how to place and organize data for the critical capability
runs. Fourth, we anticipiate storing mutliple data copies, each compressed in
different ways according to the underlying media, some of these copies will
disappear based on storage pressures, but data persistence will be maintained
according to user specifications. Assuming a relatively low latency cache layer
before a tape system, we can offer exporatory data access reserving pulling
data from tape to just the data required. This will save scientists time and
make data stored on tape usable without long delays.

Our research efforts will be heavily focused on the need to, in a coordinated
manner, adapt data and metadata retention policies to the dynamic
resource balancing that will need to take place between the application,
OS/R, and hardware.

\underline{\textbf{Significance and Impact:}}
The success of this project will  provide insights into how to build middleware and storage layers which can interact
well with each other, and take user-provided hints. Today data is reduced by application scientist who have 
limited information on what the storage layer can provide. They often make comprimises based on this limited knowledge
and either tune their output for writing or reading performance. This data then gets moved to other locations, and much of
the tuning is lost when the data is read back during their post processing. Furthermore, there is a limited set of operations
which users will be able to stage to other staging nodes for real-tiime-reduction and visualization.

We will identify,   through concrete application evaluation, the requirements for highly 
usable and scalable middleware and storage layers

\underline{\textbf{Research Plan:}} 
Our research strategy is to study

\emph{Adaptive Data Management.}

\emph{Proxy Analyses and Workflow Proxies.}

%\includepdf[pages={1}]{COI.pdf}

\end{document}

%%% End:

### 2015-06-04 LOI/inc2.tex
setting up the LOI

@@
 \end{comment}
-\includepdf[pages={1}]{ornl-cover.pdf}
-\pagebreak
+%\includepdf[pages={1}]{ornl-cover.pdf}
+%\pagebreak
 
@@
 
-\underline{\textbf{Executive Summary:}} Exascale scientific discovery will
+{\bf Title of Pre-application:} Hierarchal Extreme Scale Knowledge Management \par
+{\bf Principal Investigator:} Scott Klasky, Group Leader, ORNL, 854-241-9980, klasky@ornl.gov \par
+{\bf Funding Opportunity Announcement Number: DE-FOA-0001338} \par
+{\bf List of all co-PIs and Key/Senior Personnel} \par
+Hasan Abbasi, Oak Ridge National Laboratory \par
+Mark Ainsworth, Oak Ridge National Laboratory JFA, Brown University \par
+Matthew Curry, Sandia National Laboratory\par
+Jay Lofstead, Sandia National Laboratory \par
+Carlos Malzahn, U. Cal. Santa Cruz \par
+Manish Parashar, Rutgers University, Oak Ridge National Laboratory \par
+Lee Ward, Sandia National Laboratory \par
+\vskip 0.2 in
+
+\underline{\textbf{Objective:}} Exascale scientific discovery will
 introduce more complex hardware, and many simulations will be bottlenecked 
@@
 
-\underline{\textbf{Key Technical Contribution:}} The primary contribution of
+\underline{\textbf{Key Technical Approach:}} The primary contribution of
 our research are insights into how to build a hiearchical organization and storage
@@
 
-\underline{\textbf{Significance and Impact:}}
 The success of this project will  provide insights into how to build middleware and storage layers which can interact
@@
 
-\underline{\textbf{Research Plan:}} 
-Our research strategy is to study
-
-\emph{Adaptive Data Management.}
-
-\emph{Proxy Analyses and Workflow Proxies.}
-
 %\includepdf[pages={1}]{COI.pdf}

### 2015-06-04 LOI/inc2.tex
LOI..

--- 2015-06-03 LOI/inc2.tex
@@
 
-\underline{\textbf{Executive Summary:}} Exascale scientific discovery will
+{\bf Title of Pre-application:} Hierarchal Extreme Scale Knowledge Management \par
+{\bf Principal Investigator:} Scott Klasky, Group Leader, ORNL, 854-241-9980, klasky@ornl.gov \par
+{\bf Funding Opportunity Announcement Number: DE-FOA-0001338} \par
+{\bf List of all co-PIs and Key/Senior Personnel} \par
+Hasan Abbasi, Oak Ridge National Laboratory \par
+Mark Ainsworth, Oak Ridge National Laboratory JFA, Brown University \par
+Matthew Curry, Sandia National Laboratory\par
+Qing Liu, Oak Ridge National Laboratory \par
+Jay Lofstead, Sandia National Laboratory \par
+Kimmy Mu, Oak Ridge National Laboratory \par
+Carlos Malzahn, U. Cal. Santa Cruz \par
+Manish Parashar, Rutgers University, Oak Ridge National Laboratory \par
+ Sudharshan Vazhkudai, Oak Ridge National Laboratory \par
+Lee Ward, Sandia National Laboratory \par
+\vskip 0.2 in
+
+\underline{\textbf{Objectives:}} Exascale scientific discovery will
 introduce more complex hardware, and many simulations will be bottlenecked 
@@
 produced during the simulation, and analzyed for months after the simulations.
-
 Our goal is to explore and address
@@
 sucess of our middleware system, ADIOS, our multi-tier storage system, Sirocco, and our distributed storage system, Cepth.
+We will also take advantage of our staging system, Data Spaces, and our Comon communication Interface (CCI) and build on all of
+our sucessful software used in petascale simulations to understand the new directions which need to be understood to fully develop
+the next generation I/O and Storage System for exascale and beyond systems.
 
 %
-We will also study ??
+We will apply a technique which will re-organize and possibly reduce the total amount of information based on user intentions along
+with direct feedback from the storage and I/O system to allow data to be optimized not just for output, but also for many of the complex
+reading patterns. Since data analysis and visualization will be performed in both an in situ and post processing environment, 
+we will continue to use our I/O and storage abstractions to help enable applications to take full advantage of the hardware and softare 
+environments. Because data volumes are strictly outstripping available bandwidths, compression will
+be required. Rather than focusing solely on lossless compression, we will
+incorporate lossy compression with the ability to annotate data allowing
+optimized algorithm selection based on the numerical methods employed in the
+source simulation. We plan to epxlore this tradeoff of storage capacity and
+bandwidth for computation to determine how to offer the user options that can
+meet their needs with desireable performance characteristics.
 %
-Finally, we will, jointly with application partners, study the behvior of how data gets written and read from the storage layers, and
-what key inishgts can be used in the next generation LCF systems, along with future exascale systems.
+Our objective is to reduce the time to knowledge, which means that we will also support additional access modes, beyond the traditional 
+Checkpoint/Restart methods, and understand that all data has errors associated with it and we will ensure that these become encoded to allow
+for additional optimizations. For example, all scientific simulations contain approximations, and all measurements from obervations and experiments 
+also contain errors. These errors can help guide how information can be presented to users, by allowing us to ask for information within a given 
+accuracy bound, and also allow us to offer a mode for recomputation vs. data storage and reterival.  
+%
+{\bf: please list other things from UCSC, OLCF}.
+%
 
-\underline{\textbf{Key Technical Contribution:}} The primary contribution of
-our research are insights into how to build a hiearchical organization and storage
-of massive scientific data sets generated from exascale computations along with 
-the necessary information needed for advanced data validation and exploration from 
-experimental and observational data. 
-We are fundamentally trying to address the fundmantal questions : 1)  How can we 
-place and mangage massive scienfici data across all of the tiers of the storage and memory
-hiearchy? 2) What are the proper semantics needed in order to help bring knowledge from the applications
-to the middleware layer. 3) What information from the storage layer can be exposed to the middleweare such
-that the placement of information can be agreed upon from what the application requires and what the
-system resources are available.
+\underline{\textbf{Key Technical Approach:}}
+%We are fundamentally trying to address the fundmantal questions : 1)  How can we 
+%place and mangage massive scienfici data across all of the tiers of the storage and memory
+%hiearchy? 2) What are the proper semantics needed in order to help bring knowledge from the applications
+%to the middleware layer. 3) What information from the storage layer can be exposed to the middleweare such
+%that the placement of information can be agreed upon from what the application requires and what the
+%system resources are available.
 

### 2015-06-04 bios/MABio.tex
Added COI information for Mark

\documentclass[11pt]{article}
%\usepackage[utf8]{inputenc}
\usepackage[margin=1.0in]{geometry}
%opening
\title{Mark Ainsworth}
\author{}
\date{}

\begin{document}

\maketitle

\flushleft{\sc\bf Professional Preparation}
       PhD Mathematics (1989). Durham University, UK. 1986-1989;
       BSc (Hons) Mathematics, {\em First class.} Durham University,
       UK. 1983-1986.

\flushleft{\bf Selected Appointments and Honours}
     2012-Present \emph{Professor of Applied Mathematics}, Brown
         University;
     2004-2012 \emph{The 1825 Chair Professor of Mathematics},
         Strathclyde University, UK;
     1998-2004 \emph{Professor of Applied Mathematics}, Strathclyde
         University, UK;
	 \emph{LMS Whitehead Prize (London Math Society)},  
	 \emph{J.L. Lions Prize for Computational Mathematics},
	 \emph{SIAM Fellow},
	 \emph{Fellow Royal Society of Edinburgh},
	 \emph{Fellow Inst. Math. Applic.}.

  \nocite{Ainsworth+Andriamaro+etal:Bernstein-Bezier}
  \nocite{Ainsworth+Rankin:Guaranteed}
  \nocite{Ainsworth:synthesis}
  \nocite{Ainsworth+Oden:book}
  \nocite{Ainsworth+Rankin:Robust}
  \nocite{Ainsworth+Rankin:Quantities}
  \nocite{Ainsworth+Ma:Non-uniform}
  \nocite{Ainsworth+Vejchodsky:Fully}
  \nocite{Ainsworth+Senior:Aspects}
  \nocite{Ainsworth:framework}

\flushleft{\bf Synergistic Activities}
        Director of NAIS, 2010-2012. NAIS is a UK based research
            centre supported by \pounds 4.5M from EPSRC under contract
            EP/G036136/1 {\em Numerical Algorithms and Intelligent Software
            for the Evolving HPC Platform}. 
        Organiser of meeting on {\em Scientific Computing at the
            Exascale}, 11th October 2011, London sponsored by London
            Mathematical Society;
        Membership of over 10 editorial boards including {\em SIAM Journal on Scientific
            Computing} (2001-present) {\em SIAM Journal on Numerical Analysis}
            (2012-present);
        Chair of {\em Biennial Conference on Numerical Analysis} held
            at Strathclyde University, June 2011, June 2013;
	Plenary speaker at {\em ENUMATH European Conference on Numerical Mathematics},
            September 2015, Turkey.

\renewcommand{\refname}{\small\bf Selected Recent Publications}
%\bibliography{$HOME/Research/Papers/tools/mybib}
%\bibliographystyle{siam}
\begin{thebibliography}{10}

\bibitem{Ainsworth:synthesis}
{\sc M.~Ainsworth}, {\em A synthesis of a posteriori error estimation
  techniques for conforming, non-conforming and discontinuous {G}alerkin finite
  element methods}, Contemp. Math., 383 (2005), pp.~1--14.

\bibitem{Ainsworth:framework}
\leavevmode\vrule height 2pt depth -1.6pt width 23pt, {\em A framework for
  obtaining guaranteed error bounds for finite element approximations}, J.
  Comput. Appl. Math., 234 (2010), pp.~2618 -- 2632.

\bibitem{Ainsworth+Andriamaro+etal:Bernstein-Bezier}
{\sc M.~Ainsworth, G.~Andriamaro, and O.~Davydov}, {\em Bernstein-{B}\'ezier
  {FEM} and optimal assembly algorithms}, SIAM J. Sci. Comp.,  (2011),
  pp.~3087--3109.

\bibitem{Ainsworth+Ma:Non-uniform}
{\sc M.~Ainsworth and X.~Ma}, {\em Non-uniform order mixed {FEM} approximation:
  implementation, post-processing, computable error bound and adaptivity}, J.
  Comput. Phys., 231 (2012), pp.~436--453.

\bibitem{Ainsworth+Oden:book}
{\sc M.~Ainsworth and J.~Oden}, {\em A posteriori error estimation in finite
  element analysis}, Pure and Applied Mathematics, Wiley-Interscience, John
  Wiley \& Sons, New York, 2000.

\bibitem{Ainsworth+Rankin:Robust}
{\sc M.~Ainsworth and R.~Rankin}, {\em Robust a posteriori error estimation for
  the nonconforming {F}ortin-{S}oulie finite element approximation}, Math.
  Comp., 77 (2008), pp.~1917--1939.

\bibitem{Ainsworth+Rankin:Guaranteed}
\leavevmode\vrule height 2pt depth -1.6pt width 23pt, {\em Guaranteed
  computable error bounds for conforming and nonconforming finite element
  analyses in planar elasticity}, Internat. J. Numer. Methods Engrg., 82
  (2010), pp.~1114--1157.

\bibitem{Ainsworth+Rankin:Quantities}
\leavevmode\vrule height 2pt depth -1.6pt width 23pt, {\em Guaranteed
  computable bounds on quantities of interest in finite element computations},
  Internat. J. Numer. Methods Engrg., 89 (2012), pp.~1605--1634.

\bibitem{Ainsworth+Senior:Aspects}
{\sc M.~Ainsworth and B.~Senior}, {\em Aspects of an adaptive $hp$-finite
  element method: Adaptive strategy, conforming approximation and efficient
  solvers}, Comput. Methods Appl. Mech. Engrg., 150 (1997), pp.~65--87.

\bibitem{Ainsworth+Vejchodsky:Fully}
{\sc M.~Ainsworth and T.~Vejchodsk{\'y}}, {\em Fully computable robust a
  posteriori error bounds for singularly perturbed reaction-diffusion
  problems}, Numer. Math., 119 (2011), pp.~219--243.

\end{thebibliography}

\flushleft{\bf Recent Collaborators and Co-Editors}
Allendes, Alejandro (Universidad T\'ecnica Federico Santa
Mar\'ia, Chile); 
Andriamaro, Gaelle (Strathclyde University, UK); 
Barrenechea, Gabriel R. (Strathclyde University, UK); 
Davydov, Oleg (Giessen University, Germany); 
Guzman, Johnny (Brown University); 
Karniadakis, George (Brown University); 
Ma, Xinhui (Abertay University, UK); 
Mihai, L. Angela (Cardiff University, UK); 
Rankin, Richard (Rice University, US); 
Sayas, Francisco-Javier (University of Delaware); 
Wajid, Hafiz Abdul (COMSATS Institute of Technology, Pakistan).

\flushleft{\bf Graduate and Postdoctoral Advisors and Advisees}
Graduate Advisor: Craig, Alan W. (European Commission, Brussels); Post-doctoral
Advisor: Oden, John T. (University of Texas at Austin, US); Graduate and
Postdoctoral Advisors and Advisees: Allendes, Alejandro (Universidad T\'ecnica
Federico Santa Mar\'ia, Chile); Andriamaro, Gaelle (Strathclyde University, UK);
Ma, Xinhui (Abertay University, UK); Mihai, L. Angela (Cardiff University, UK);
Rankin, Richard (Rice University, US); Wajid, Hafiz Abdul (COMSATS Institute of
Technology, Pakistan).
\end{document}

### 2015-06-04 LOI/inc2.tex
--

--- 2015-06-03 LOI/inc2.tex
@@
 introduce more complex hardware, and many simulations will be bottlenecked 
-without sufficient new research into manging and storing the large data which will be
+without sufficient new research into managing and storing the large data which will be
 produced during the simulation, and analzyed for months after the simulations.
@@
 the multi-tier challenges that are faced by scientists in creating and
-managing and storing their data  to expedite insights into
-mission critical scientific processes in exascale computing.
+managing and storing their data  to expedite insights xsinto
+mission critical scientific processes in exascale compuxsting.
 %

### 2015-06-04 LOI/inc2.tex
reworded the stack paragraph.
Comment:
I am not sure why we need to go into
such detail so quickly in the pre-proposal into what seem to be very
disparate set of software tools.

--- 2015-06-04 LOI/inc2.tex
@@
 \underline{\textbf{Objectives:}} Exascale scientific discovery will
-introduce more complex hardware, and many simulations will be bottlenecked 
-without sufficient new research into manging and storing the large data which will be
+%introduce more complex hardware, and many simulations will 
+be bottlenecked 
+without sufficient new research into managing and storing the large data which will be
 produced during the simulation, and analzyed for months after the simulations.
-Our goal is to explore and address
-the multi-tier challenges that are faced by scientists in creating and
-managing and storing their data  to expedite insights into
-mission critical scientific processes in exascale computing.
+Our goal is to investigate a unique approach to addressing this challenge in
+a multi-tier storage platform, expediting insights into mission critical
+scientific processes. 
+% the multi-tier challenges that are faced by scientists in creating and
+% managing and storing their data  to expedite insights xsinto
+% mission critical scientific processes in exascale compuxsting.
 %
-We propose a research program that aims to explore a hiearchical organization and storage infstraturure which will facilitiate 
-how we data can be written and read efficiently. In order to help us explore our many research issues, we will build upon the
-sucess of our middleware system, ADIOS, our multi-tier storage system, Sirocco, and our distributed storage system, Cepth.
-We will also take advantage of our staging system, Data Spaces, and our Comon communication Interface (CCI) and build on all of
-our sucessful software used in petascale simulations to understand the new directions which need to be understood to fully develop
-the next generation I/O and Storage System for exascale and beyond systems.
+We propose a research program that aims to explore a hiearchical
+organization and storage infstraturure which will facilitiate how data can
+be written to and read from a complex heterogenous storage hierarchy
+efficiently. In order to explore the many research challenges in this
+domain, we will build upon the success of our middleware system, ADIOS, our
+multi-tier storage system, Sirocco, and our distributed storage system,
+Ceph.  We will also leverage data movement and staging technologies, Data
+Spaces and the Common Communication Interface (CCI), as part of a software
+stack used successfully on current leadership class machines. Our experience
+with this stack will guide us in understanding the new directions for
+research in storage systems and I/O for exascale and beyond systems.
+
+% We will also take advantage of our staging system, Data Spaces, and the
+% network abstraction layer, the Common communication Interface (CCI),
+% and build on all of
+% our sucessful software used in petascale simulations to understand the new directions which need to be understood to fully develop
+% the next generation I/O and Storage System for exascale and beyond systems.
 

### 2015-06-04 LOI/inc2.tex
reordered the software stack and proposed work paragraphs

@@
 be written to and read from a complex heterogenous storage hierarchy
-efficiently. In order to explore the many research challenges in this
+efficiently.  
+%
+We will apply a technique which will re-organize and possibly
+reduce the total amount of information based on user intentions along with
+direct feedback from the storage and I/O system to allow data to be
+optimized not just for output, but also for many of the complex reading
+patterns. Since data analysis and visualization will be performed in both an
+in situ and post processing environment, we will continue to use our I/O and
+storage abstractions to help enable applications to take full advantage of
+the hardware and softare environments. Because data volumes are strictly
+outstripping available bandwidths, compression will be required. Rather than
+focusing solely on lossless compression, we will incorporate lossy
+compression with the ability to annotate data allowing optimized algorithm
+selection based on the numerical methods employed in the source
+simulation. We plan to epxlore this tradeoff of storage capacity and
+bandwidth for computation to determine how to offer the user options that
+can meet their needs with desireable performance characteristics.
+
+In order to explore the many research challenges in this
 domain, we will build upon the success of our middleware system, ADIOS, our
@@
 %
-We will apply a technique which will re-organize and possibly reduce the total amount of information based on user intentions along
-with direct feedback from the storage and I/O system to allow data to be optimized not just for output, but also for many of the complex
-reading patterns. Since data analysis and visualization will be performed in both an in situ and post processing environment, 
-we will continue to use our I/O and storage abstractions to help enable applications to take full advantage of the hardware and softare 
-environments. Because data volumes are strictly outstripping available bandwidths, compression will
-be required. Rather than focusing solely on lossless compression, we will
-incorporate lossy compression with the ability to annotate data allowing
-optimized algorithm selection based on the numerical methods employed in the
-source simulation. We plan to epxlore this tradeoff of storage capacity and
-bandwidth for computation to determine how to offer the user options that can
-meet their needs with desireable performance characteristics.
+
 %
-Our objective is to reduce the time to knowledge, which means that we will also support additional access modes, beyond the traditional 
-Checkpoint/Restart methods, and understand that all data has errors associated with it and we will ensure that these become encoded to allow
-for additional optimizations. For example, all scientific simulations contain approximations, and all measurements from obervations and experiments 
-also contain errors. These errors can help guide how information can be presented to users, by allowing us to ask for information within a given 
-accuracy bound, and also allow us to offer a mode for recomputation vs. data storage and reterival.  
+Our objective here is to reduce the time to knowledge, an end-to-end metric
+relevant to the scientific discovery process. Beyond the traditional high
+volume I/O pattern of checkpoint/restart, we will address the challenges
+posed by additional data access patterns in the knowledge gathering
+process. With a deeper insight into the scientific process we will encode
+and utilize errors and accuracy as an optimization parameter. For example, 
+%
+% We will explore access to
+% large data sets as part of this process, w
+
+% which means that we will
+% also support additional access modes, beyond the traditional
+% Checkpoint/Restart methods, and understand that all data has errors
+% associated with it and we will ensure that these become encoded to allow for
+% additional optimizations. 
+For example, all scientific simulations contain
+approximations, and all measurements from obervations and experiments also
+contain errors. These errors can help guide how information can be presented
+to users, by allowing us to ask for information within a given accuracy
+bound, and also allow us to offer a mode for recomputation vs. data storage
+and reterival.
+
 %

### 2015-06-04 LOI/inc2.tex
minor changes

--- 2015-06-04 LOI/inc2.tex
@@
 Carlson Malzahn, UCSC
-Jay Lofstead, Sandia
+Jay Lofstead, Sandia, gflofst@sandia.gov, 505-284-5803
 
@@
 \hline
-  Argonne National Laboratory & \$350,000 & \$350,000 & \$350,000 \\
-\hline
-  Lawrence Berkeley Laboratory & \$300,000 & \$300,000 & \$300,000\\
-\hline
-  Brookhaven National Laboratory & \$150,000 & \$150,000 & \$150,000 \\
-\hline
   Rutgers University & \$175,000 & \$175,000 & \$175,000 \\
 \hline
-  Stony Brook University & \$175,000 & \$175,000 & \$175,000 \\
+  University of California Santa Cruz & \$160,000 & \$160,000 & \$160,000 \\
 \hline
-  Total & \$1,500,000 & \$1,500,000 & \$1,500,000 \\
+  Sandia National Laboratories & \$360,000 & \$360,000 & \$360,000 \\
+\hline
+  Total & \$1,250,000 & \$1,250,000 & \$1,250,000 \\
 \hline
@@
 introduce more complex hardware, and many simulations will be bottlenecked 
-without sufficient new research into managing and storing the large data which will be
+without sufficient new research into manging and storing the large data which will be
 produced during the simulation, and analzyed for months after the simulations.
@@
 the multi-tier challenges that are faced by scientists in creating and
-managing and storing their data  to expedite insights xsinto
-mission critical scientific processes in exascale compuxsting.
+managing and storing their data  to expedite insights into
+mission critical scientific processes in exascale computing.
 %
@@
 how we data can be written and read efficiently. In order to help us explore our many research issues, we will build upon the
-sucess of our middleware system, ADIOS, our multi-tier storage system, Sirocco, and our distributed storage system, Cepth.
+sucess of our middleware system, ADIOS, our multi-tier storage system, Sirocco, and our distributed storage system, Ceph.
 We will also take advantage of our staging system, Data Spaces, and our Comon communication Interface (CCI) and build on all of
@@
 optimized algorithm selection based on the numerical methods employed in the
-source simulation. We plan to epxlore this tradeoff of storage capacity and
+source simulation. We plan to explore this tradeoff of storage capacity and
 bandwidth for computation to determine how to offer the user options that can
@@
 well with each other, and take user-provided hints. Today data is reduced by application scientist who have 
-limited information on what the storage layer can provide. They often make comprimises based on this limited knowledge
+limited information on what the storage layer can provide. They often make compromises based on this limited knowledge
 and either tune their output for writing or reading performance. This data then gets moved to other locations, and much of

### 2015-06-05 LOI/inc2.tex
Initial edits -- a work in progress

@@
 %introduce more complex hardware, and many simulations will 
-be bottlenecked 
-without sufficient new research into managing and storing the large data which will be
-produced during the simulation, and analzyed for months after the simulations.
-Our goal is to investigate a unique approach to addressing this challenge in
-a multi-tier storage platform, expediting insights into mission critical
-scientific processes. 
+be severely bottlenecked without sufficient new 
+ research into managing and storing the large amounts of 
+data that will be produced during the simulation, and analyzed for months after the simulations.
+%Our goal in this project is to investigate a unique approach to addressing this data management 
+%and storage challenge for multi-tier exascale storage architectures, thereby expediting insights into mission critical
+%scientific processes. 
+Our goal in this project is to address this data management and storage challenge. Specifically we plan
+ to research and develop a unique application-aware hierarchy data organization that can effectively map 
+ to  multi-tier exascale storage architectures, and can facilitate how data can be efficiently written to and 
+ read from a complex heterogeneous storage hierarchy  
+ thereby expediting insights into mission critical scientific processes. 
 % the multi-tier challenges that are faced by scientists in creating and
@@
 %
-We propose a research program that aims to explore a hiearchical
-organization and storage infstraturure which will facilitiate how data can
-be written to and read from a complex heterogenous storage hierarchy
-efficiently.  
+%Specifically, we propose a research program that aims to explore a hierarchical
+%organization and storage infrastructure that will facilitate how data can
+%be written to and read from a complex heterogenous storage hierarchy
+%efficiently.  
 %
-We will apply a technique which will re-organize and possibly
-reduce the total amount of information based on user intentions along with
-direct feedback from the storage and I/O system to allow data to be
-optimized not just for output, but also for many of the complex reading
-patterns. Since data analysis and visualization will be performed in both an
-in situ and post processing environment, we will continue to use our I/O and
-storage abstractions to help enable applications to take full advantage of
-the hardware and softare environments. Because data volumes are strictly
-outstripping available bandwidths, compression will be required. Rather than
-focusing solely on lossless compression, we will incorporate lossy
-compression with the ability to annotate data allowing optimized algorithm
-selection based on the numerical methods employed in the source
-simulation. We plan to epxlore this tradeoff of storage capacity and
-bandwidth for computation to determine how to offer the user options that
-can meet their needs with desireable performance characteristics.
+Our approach is to explicitly and implicitly capture user/application intentions 
+and use this information, along with direct feedback from the storage and I/O 
+system, to dynamically structure, organize, process and map the data so that 
+it can most effectively support complex I/O patterns. Furthermore, we will do 
+this with an awareness of data usage semantics and with an understanding 
+of current and future hardware and software technologies. 
+%re-organize and possibly reduce the total amount of information 
+%Since data analysis and visualization will be performed in both an
+%in situ and post processing environment, we will continue to use our I/O and
+%storage abstractions to help enable applications to take full advantage of
+%the hardware and software environments. 
+For example, as data volumes are quickly outgrowing available bandwidths, 
+compression is becoming essential. Rather than focusing solely on lossless 
+compression, we will incorporate lossy compression with the ability to annotate 
+data allowing optimized algorithm selection based on the numerical methods 
+employed in the source simulation. We plan to explore this tradeoff of storage 
+capacity and bandwidth for computation to determine how to offer the user 
+options that can meet their needs with desirable performance characteristics.
 
-In order to explore the many research challenges in this
-domain, we will build upon the success of our middleware system, ADIOS, our
-multi-tier storage system, Sirocco, and our distributed storage system,
-Ceph.  We will also leverage data movement and staging technologies, Data
-Spaces and the Common Communication Interface (CCI), as part of a software
-stack used successfully on current leadership class machines. Our experience
-with this stack will guide us in understanding the new directions for
-research in storage systems and I/O for exascale and beyond systems.
+In order to explore the many research challenges in this domain, we will build 
+upon the success of our middleware system, ADIOS, our multi-tier storage 
+system, Sirocco, and our distributed storage system, Ceph.  We will also leverage 
+data staging and data movement technologies, DataSpaces and the Common 
+Communication Interface (CCI), as part of a software stack. 
+%used successfully on current leadership class machines. 
+%Our experience with this stack will guide us in understanding the new directions for
+%research in storage systems and I/O for exascale and beyond systems.
 
@@
 For example, all scientific simulations contain
-approximations, and all measurements from obervations and experiments also
+approximations, and all measurements from observations and experiments also
 contain errors. These errors can help guide how information can be presented
 to users, by allowing us to ask for information within a given accuracy
-bound, and also allow us to offer a mode for recomputation vs. data storage
+bound, and also allow us to offer a mode for re-computation vs. data storage
 and reterival.
@@
 This will allow us to adapt data from a variable into the various storage tiers. Rather than focus on simple data
-compression, we will incoroarte a multitude of techniques to reduce data on the faster storage tiers, and keep the less
+compression, we will incorporate a multitude of techniques to reduce data on the faster storage tiers, and keep the less
 reduced information on the lower tiers.
@@
 We wish to support additional data access modes as well. First, within a given
-accuracy bounds, offer a mode for recomputation from data stored in a ``fast''
+accuracy bounds, offer a mode for re-computation from data stored in a ``fast''
 tier to reduce data latencies. Second, exploratory analysis operations
@@
 can accurate decide how to place and organize data for the critical capability
-runs. Fourth, we anticipiate storing mutliple data copies, each compressed in
+runs. Fourth, we anticipate storing multiple data copies, each compressed in
 different ways according to the underlying media, some of these copies will
@@
 according to user specifications. Assuming a relatively low latency cache layer
-before a tape system, we can offer exporatory data access reserving pulling
+before a tape system, we can offer exploratory data access reserving pulling
 data from tape to just the data required. This will save scientists time and
@@
 the tuning is lost when the data is read back during their post processing. Furthermore, there is a limited set of operations
-which users will be able to stage to other staging nodes for real-tiime-reduction and visualization.
+which users will be able to stage to other staging nodes for real-time-reduction and visualization.
 

### 2015-06-05 LOI/inc2.tex
Releasing lock

@@
 % \item What is the new technical idea; why can we succeed now?
-%  Integration of proven components into an architecure that solves
+%  Integration of proven components into an architecture that solves
 %  problems. New research will resolve remaining technical hurdles.
@@
 volume I/O pattern of checkpoint/restart, we will address the challenges
-posed by additional data access patterns in the knowledge gathering
-process. With a deeper insight into the scientific process we will encode
-and utilize errors and accuracy as an optimization parameter. For example, 
+posed by other essential data access patterns in the knowledge gathering
+process. Through a deeper insight into the scientific process we will encode
+and utilize accuracy and errors as optimization parameters. 
 %
@@
 % additional optimizations. 
-For example, all scientific simulations contain
-approximations, and all measurements from observations and experiments also
-contain errors. These errors can help guide how information can be presented
-to users, by allowing us to ask for information within a given accuracy
-bound, and also allow us to offer a mode for re-computation vs. data storage
+For example, scientific simulations contain approximations, as do measurements 
+from observations and experiments, and depending on the goal, these approximations 
+are acceptable. We can leverage this observation to optimize the presentation of data
+to the user and to implement various tradeoffs. For example, users can ask for information 
+within a given accuracy bound, allowing us to offer a mode for re-computation vs. data storage
 and reterival.
@@
 %system resources are available. 
+Our overall technical approach is based on an application-aware runtime realization of tradeoffs in 
+data-representation, data placement and data access. For example, we will allow users to ``plug-in'' 
+techniques to classify data, not as bytes but as motifs, allowing us to understand relationships 
+between objects (based on data models), the relationships of data and variables, as well as data
+access and transport patterns.  This will allow us 
+to adaptively map data from a variable onto the various storage tiers. Rather than focus on simple data
+compression, we will incorporate a multitude of techniques to reduce data on the faster storage tiers, and keep the less
+reduced information in the lower storage tiers.
 
-Our approach will allow users to ``plug-in'' techniques to classify data, not as bytes but as motifs, where
-we can understand relationships between objects (into data models) and relationships of data in variables.
-This will allow us to adapt data from a variable into the various storage tiers. Rather than focus on simple data
-compression, we will incorporate a multitude of techniques to reduce data on the faster storage tiers, and keep the less
-reduced information on the lower tiers.
-
-We wish to support additional data access modes as well. First, within a given
+We will support additional data access modes as well. First, within a given
 accuracy bounds, offer a mode for re-computation from data stored in a ``fast''
@@
 be retrieved at the time cost of moving the potentially huge data quantity.
+
 Third, to ensure available storage for subsequent operations, we will offer
@@
 
-The success of this project will  provide insights into how to build middleware and storage layers which can interact
-well with each other, and take user-provided hints. Today data is reduced by application scientist who have 
+The success of this project will  provide insights into how to build autonomic middleware and storage layers which can interact
+well with each other, and take user-provided hints. Today, data is reduced by application scientist who have 
 limited information on what the storage layer can provide. They often make compromises based on this limited knowledge
@@
 
-We will identify,   through concrete application evaluation, the requirements for highly 
-usable and scalable middleware and storage layers
+%We will identify,  through concrete application evaluation, the requirements for highly usable and scalable middleware and storage layers
 

### 2015-06-05 LOI/inc2-old.tex
Major overhaul based on the Carlos, Matthew, and Jay conversation

--- 2015-06-05 LOI/inc2.tex
@@
 Matthew Curry, Sandia National Laboratory\par
-Qing Liu, Oak Ridge National Laboratory \par
+Qing Gary Liu, Oak Ridge National Laboratory \par
 Jay Lofstead, Sandia National Laboratory \par
@@
  research into managing and storing the large amounts of 
-data that will be produced during the simulation, and analyzed for months after the simulations.
+data that will be produced during the simulation, and analyzed for months afterwards.
 %Our goal in this project is to investigate a unique approach to addressing this data management 
@@
 %scientific processes. 
-Our goal in this project is to address this data management and storage challenge. Specifically we plan
- to research and develop a unique application-aware hierarchy data organization that can effectively map 
- to  multi-tier exascale storage architectures, and can facilitate how data can be efficiently written to and 
- read from a complex heterogeneous storage hierarchy  
- thereby expediting insights into mission critical scientific processes. 
+Our goal in this project is to address the associated I/O and storage challenges on the 
+future storage landscape, and expedite insights into mission critical
+scientific processes. To this end, we plan
+ to research and develop a unique application-aware data management scheme that
+can (1) effectively map data to multi-tier exascale storage architectures, 
+(2) facilitate how data can be efficiently written to and 
+ read from a complex heterogeneous storage hierarchy, and (3) reduce and re-generate
+data on-the-fly without compromising the quality of science. The project
+will assemble a team with strong expertise in I/O middleware (ORNL, Rutgers), file system (SNL, UCSC)
+and storage (UCSC), and connect and coordinate
+these key storage components in a seamless fashion. 
 % the multi-tier challenges that are faced by scientists in creating and
@@
 Our approach is to explicitly and implicitly capture user/application intentions 
-and use this information, along with direct feedback from the storage and I/O 
+and use this information, along with direct feedback from the I/O and storage
 system, to dynamically structure, organize, process and map the data so that 
-it can most effectively support complex I/O patterns. Furthermore, we will do 
-this with an awareness of data usage semantics and with an understanding 
+it can most effectively support complex I/O patterns. This will be done
+with an awareness of data usage semantics and an understanding 
 of current and future hardware and software technologies. 
@@
 %the hardware and software environments. 
-For example, as data volumes are quickly outgrowing available bandwidths, 
-compression is becoming essential. Rather than focusing solely on lossless 
+On the other hand, data reduction is becoming increasingly important, as a result of 
+the disparirty between data volume/frequency and storage bandwdith. 
+Rather than focusing solely on lossless 
 compression, we will incorporate lossy compression with the ability to annotate 
-data allowing optimized algorithm selection based on the numerical methods 
-employed in the source simulation. We plan to explore this tradeoff of storage 
+data, allowing for optimized algorithm selection based on the numerical accuracy
+required by the simulation and analysis. We plan to explore this tradeoff of storage 
 capacity and bandwidth for computation to determine how to offer the user 
@@
 
-In order to explore the many research challenges in this domain, we will build 
-upon the success of our middleware system, ADIOS, our multi-tier storage 
-system, Sirocco, and our distributed storage system, Ceph.  We will also leverage 
-data staging and data movement technologies, DataSpaces and the Common 
-Communication Interface (CCI), as part of a software stack. 
-%used successfully on current leadership class machines. 
+%In order to explore the many research challenges in this domain, we will build 
+%upon the success of our middleware system, ADIOS, our multi-tier storage 
+%system, Sirocco, and our distributed storage system, Ceph.  We will also leverage 
+%data staging and data movement technologies, DataSpaces and the Common 
+%Communication Interface (CCI), as part of the I/O stack. 
+%%used successfully on current leadership class machines. 
 %Our experience with this stack will guide us in understanding the new directions for
@@
 Our overall technical approach is based on an application-aware runtime realization of tradeoffs in 
-data-representation, data placement and data access. For example, we will allow users to ``plug-in'' 
-techniques to classify data, not as bytes but as motifs, allowing us to understand relationships 
-between objects (based on data models), the relationships of data and variables, as well as data
-access and transport patterns.  This will allow us 
-to adaptively map data from a variable onto the various storage tiers. Rather than focus on simple data
-compression, we will incorporate a multitude of techniques to reduce data on the faster storage tiers, and keep the less
+data representation, data placement and data access. We will allow users to ``plug-in'' 
+their knowledge about the data, not as bytes but as motifs, allowing the I/O and storage system
+to understand user intentions and the relationships between data objects, as well as data
+access and transport patterns. This will facilitate
+efficient mapping of data from user space, such as a multi-dimensional array, onto various storage tiers. 
+Rather than focus on simple data
+compression, we will incorporate a multitude of techniques to reduce data on the faster yet smaller
+storage tiers, and keep the less
 reduced information in the lower storage tiers.

### 2015-06-05 LOI/inc2.tex
Major overhaul based on the Carlos, Matthew, and Jay conversation

@@
 
-\underline{\textbf{Objectives:}} Exascale scientific discovery will
-%introduce more complex hardware, and many simulations will 
-be severely bottlenecked without sufficient new 
- research into managing and storing the large amounts of 
-data that will be produced during the simulation, and analyzed for months after the simulations.
-%Our goal in this project is to investigate a unique approach to addressing this data management 
-%and storage challenge for multi-tier exascale storage architectures, thereby expediting insights into mission critical
-%scientific processes. 
-Our goal in this project is to address this data management and storage challenge. Specifically we plan
- to research and develop a unique application-aware hierarchy data organization that can effectively map 
- to  multi-tier exascale storage architectures, and can facilitate how data can be efficiently written to and 
- read from a complex heterogeneous storage hierarchy  
- thereby expediting insights into mission critical scientific processes. 
-% the multi-tier challenges that are faced by scientists in creating and
-% managing and storing their data  to expedite insights xsinto
-% mission critical scientific processes in exascale compuxsting.
-%
-%Specifically, we propose a research program that aims to explore a hierarchical
-%organization and storage infrastructure that will facilitate how data can
-%be written to and read from a complex heterogenous storage hierarchy
-%efficiently.  
-%
-Our approach is to explicitly and implicitly capture user/application intentions 
-and use this information, along with direct feedback from the storage and I/O 
-system, to dynamically structure, organize, process and map the data so that 
-it can most effectively support complex I/O patterns. Furthermore, we will do 
-this with an awareness of data usage semantics and with an understanding 
-of current and future hardware and software technologies. 
-%re-organize and possibly reduce the total amount of information 
-%Since data analysis and visualization will be performed in both an
-%in situ and post processing environment, we will continue to use our I/O and
-%storage abstractions to help enable applications to take full advantage of
-%the hardware and software environments. 
-For example, as data volumes are quickly outgrowing available bandwidths, 
-compression is becoming essential. Rather than focusing solely on lossless 
-compression, we will incorporate lossy compression with the ability to annotate 
-data allowing optimized algorithm selection based on the numerical methods 
-employed in the source simulation. We plan to explore this tradeoff of storage 
-capacity and bandwidth for computation to determine how to offer the user 
-options that can meet their needs with desirable performance characteristics.
+\underline{\textbf{Objectives:}}
+Exascale scientific discovery will be severely bottlenecked without sufficient
+new research into managing and storing the large amounts of data that will be
+produced during the simulation, and analyzed for months afterwards.  Our goal
+in this project is to address the associated I/O and storage challenges on the
+future storage landscape, and expedite insights into mission critical
+scientific processes. To that end, we will build on the capabilities offered
+by the Sirocco peer-to-peer file system under development at Sandia and the
+object storage and annotation expertise of UC Santa Cruz to develop a unique
+application aware data management scheme. It will build on the existing
+multi-tier storage aware approach offered by Sirroco to incorporate new
+functionality and extended object metadata and APIs for 1) matching available
+storage bandwidth across the storage hierarchy with data importance employing
+application aware data compression (both lossy with specified error bounds and
+lossless), 2) a selectable performance/quality tradeoff when reading data, and
+3) incorporate reader prioritization for data annotation, placement, and data
+quality when writing data.  The major challenge in this project is metadata
+management connecting different object quality levels and connecting the user
+with the best possible outcomes given the current system state. The new user
+APIs required to interact with this rich metadata system will drive effective
+use of the entire middleware and storage stack. The project is compriesed of a
+team with strong expertise in I/O middleware (ORNL, Rutgers), file system (SNL,
+UCSC) and storage (UCSC), and connect and coordinate these key storage
+components in a seamless fashion.
 
-In order to explore the many research challenges in this domain, we will build 
-upon the success of our middleware system, ADIOS, our multi-tier storage 
-system, Sirocco, and our distributed storage system, Ceph.  We will also leverage 
-data staging and data movement technologies, DataSpaces and the Common 
-Communication Interface (CCI), as part of a software stack. 
-%used successfully on current leadership class machines. 
-%Our experience with this stack will guide us in understanding the new directions for
-%research in storage systems and I/O for exascale and beyond systems.
+Our approach can most easily be understood by considering a road navigation
+program. Similar to the navigation program, we will offer a coarse grained,
+fast overview of a data set with progressively more detail in areas of interest
+noted by data features and less data in areas without. These progressively
+more detailed data views require more storage space and time to retrieve. By
+incorporating middleware informed of both how to identify relevant data
+features and the intended use locations, it will be able to selectively store
+data at different quality levels, with bounded errors in the case of lossy
+compression, and allow a user to select what data to retrieve by specifying
+both an acceptable retrieval timeframe and quality required. If the request
+cannot be fulfilled, the storage system can negotiate with the user providing
+information about the quality and timeframes available for a subsequent
+request.
 
-% We will also take advantage of our staging system, Data Spaces, and the
-% network abstraction layer, the Common communication Interface (CCI),
-% and build on all of
-% our sucessful software used in petascale simulations to understand the new directions which need to be understood to fully develop
-% the next generation I/O and Storage System for exascale and beyond systems.
+Our objective here is to reduce the time to knowledge, an end-to-end metric
+relevant to the scientific discovery process. Beyond the traditional high
+volume I/O pattern of checkpoint/restart, we will address the challenges posed
+by other essential data access patterns in the knowledge gathering process.
+Through a deeper insight into the scientific process we will encode and utilize
+accuracy and errors as optimization parameters.
+
+For example, scientific simulations contain approximations, as do measurements
+from observations and experiments, and depending on the goal, these
+approximations are acceptable. We can leverage this observation to optimize the
+presentation of data to the user and to implement various tradeoffs. For
+example, users can ask for information within a given accuracy bound, allowing
+us to offer a mode for re-computation vs. data storage and reterival.
 
 %
-
-%
-Our objective here is to reduce the time to knowledge, an end-to-end metric
-relevant to the scientific discovery process. Beyond the traditional high
-volume I/O pattern of checkpoint/restart, we will address the challenges
-posed by other essential data access patterns in the knowledge gathering
-process. Through a deeper insight into the scientific process we will encode
-and utilize accuracy and errors as optimization parameters. 
-%
-% We will explore access to
-% large data sets as part of this process, w
-
-% which means that we will
-% also support additional access modes, beyond the traditional
-% Checkpoint/Restart methods, and understand that all data has errors
-% associated with it and we will ensure that these become encoded to allow for
-% additional optimizations. 
-For example, scientific simulations contain approximations, as do measurements 
-from observations and experiments, and depending on the goal, these approximations 
-are acceptable. We can leverage this observation to optimize the presentation of data
-to the user and to implement various tradeoffs. For example, users can ask for information 
-within a given accuracy bound, allowing us to offer a mode for re-computation vs. data storage
-and reterival.
-
-%
-{\bf: please list other things from UCSC, OLCF}.
+{\bf: please list other things from OLCF}.
 %
@@
 \underline{\textbf{Key Technical Approach:}}
-%We are fundamentally trying to address the fundmantal questions : 1)  How can we 
-%place and mangage massive scienfici data across all of the tiers of the storage and memory
-%hiearchy? 2) What are the proper semantics needed in order to help bring knowledge from the applications
-%to the middleware layer. 3) What information from the storage layer can be exposed to the middleweare such
-%that the placement of information can be agreed upon from what the application requires and what the
-%system resources are available. 
-Our overall technical approach is based on an application-aware runtime realization of tradeoffs in 
-data-representation, data placement and data access. For example, we will allow users to ``plug-in'' 
-techniques to classify data, not as bytes but as motifs, allowing us to understand relationships 
-between objects (based on data models), the relationships of data and variables, as well as data
-access and transport patterns.  This will allow us 
-to adaptively map data from a variable onto the various storage tiers. Rather than focus on simple data
-compression, we will incorporate a multitude of techniques to reduce data on the faster storage tiers, and keep the less
-reduced information in the lower storage tiers.
+Our overall technical approach is based on an application-aware runtime
+realization of tradeoffs in data representation, data placement and data
+access. We will allow users to ``plug-in'' their knowledge about the data, not
+as bytes but as motifs, allowing the I/O and storage system to understand user
+intentions and the relationships between data objects, as well as data access
+and transport patterns. This will facilitate efficient mapping of data from
+user space, such as a multi-dimensional array, onto various storage tiers.
+Rather than focus on simple data compression, we will incorporate a multitude
+of techniques to reduce data on the faster yet smaller storage tiers, and keep
+the less reduced information in the lower storage tiers.
 
-We will support additional data access modes as well. First, within a given
-accuracy bounds, offer a mode for re-computation from data stored in a ``fast''
-tier to reduce data latencies. Second, exploratory analysis operations
-frequently entail an overall data set view followed by targeted data
+We will support additional data access modes as well. \textbf{First}, within a
+given accuracy bounds, offer a mode for re-computation from data stored in a
+``fast'' tier to reduce data latencies. \textbf{Second}, exploratory analysis
+operations frequently entail an overall data set view followed by targeted data
 exploration based on identified features. We will offer support for a
@@
 be retrieved at the time cost of moving the potentially huge data quantity.
-
-Third, to ensure available storage for subsequent operations, we will offer
-automatic data migration based on user annotations for required data lifetimes
-using monitoring and learning techniques. Unlike existing approaches, this will
-be tempered both by the user annotations and through learned access patterns.
-While past access patterns may not indicate future access because the
-simulation run purpose may have changed, we are focused on scalability where
-runs are subsequently larger as the simulation prepares for a capability run.
-By learning from the output and access patterns during this run sequence, we
-can accurate decide how to place and organize data for the critical capability
-runs. Fourth, we anticipate storing multiple data copies, each compressed in
-different ways according to the underlying media, some of these copies will
-disappear based on storage pressures, but data persistence will be maintained
-according to user specifications. Assuming a relatively low latency cache layer
-before a tape system, we can offer exploratory data access reserving pulling
-data from tape to just the data required. This will save scientists time and
-make data stored on tape usable without long delays.
+\textbf{Third}, to ensure available storage for subsequent operations, we will
+offer automatic data migration based on user annotations for required data
+lifetimes using monitoring and learning techniques. Unlike existing approaches,
+this will be tempered both by the user annotations and through learned access
+patterns.  While past access patterns may not indicate future access because
+the simulation run purpose may have changed, we are focused on scalability
+where runs are subsequently larger as the simulation prepares for a capability
+run.  By learning from the output and access patterns during this run sequence,
+we can accurate decide how to place and organize data for the critical
+capability runs. Fourth, we anticipate storing multiple data copies, each
+compressed in different ways according to the underlying media, some of these
+copies will disappear based on storage pressures, but data persistence will be
+maintained according to user specifications. Assuming a relatively low latency
+cache layer before a tape system, we can offer exploratory data access
+reserving pulling data from tape to just the data required. This will save
+scientists time and make data stored on tape usable without long delays.
 
 Our research efforts will be heavily focused on the need to, in a coordinated
-manner, adapt data and metadata retention policies to the dynamic
-resource balancing that will need to take place between the application,
-OS/R, and hardware.
+manner, adapt data and metadata retention policies to the dynamic resource
+balancing that will need to take place between the application, OS/R, and
+hardware.
 
-The success of this project will  provide insights into how to build autonomic middleware and storage layers which can interact
-well with each other, and take user-provided hints. Today, data is reduced by application scientist who have 
-limited information on what the storage layer can provide. They often make compromises based on this limited knowledge
-and either tune their output for writing or reading performance. This data then gets moved to other locations, and much of
-the tuning is lost when the data is read back during their post processing. Furthermore, there is a limited set of operations
-which users will be able to stage to other staging nodes for real-time-reduction and visualization.
-
-%We will identify,  through concrete application evaluation, the requirements for highly usable and scalable middleware and storage layers
+The success of this project will  provide insights into how to build autonomic
+middleware and storage layers which can interact well with each other, and take
+user-provided hints. Today, data is reduced by application scientist who have
+limited information on what the storage layer can provide. They often make
+compromises based on this limited knowledge and either tune their output for
+writing or reading performance. This data then gets moved to other locations,
+and much of the tuning is lost when the data is read back during their post
+processing. Furthermore, there is a limited set of operations which users will
+be able to stage to other staging nodes for real-time-reduction and
+visualization.
 

### 2015-06-06 LOI/inc2.tex
My edits.

@@
 scientific processes. To that end, we will build on the capabilities offered
-by the Sirocco peer-to-peer file system under development at Sandia and the
+by ADIOS that provides I/O abstractions, 
+the Sirocco peer-to-peer file system under development at Sandia and the
 object storage and annotation expertise of UC Santa Cruz to develop a unique
-application aware data management scheme. It will build on the existing
-multi-tier storage aware approach offered by Sirroco to incorporate new
-functionality and extended object metadata and APIs for 1) matching available
-storage bandwidth across the storage hierarchy with data importance employing
-application aware data compression (both lossy with specified error bounds and
+application-aware data management scheme. It will center upon
+a multi-tier storage-aware approach and incorporate new
+functionalities and extended object metadata and APIs for 1) matching
+storage capability across the storage hierarchy with data importance employing
+application-aware data compression (both lossy with specified error bounds and
 lossless), 2) a selectable performance/quality tradeoff when reading data, and
+the ability to recompute data if needed, and
 3) incorporate reader prioritization for data annotation, placement, and data
 quality when writing data.  The major challenge in this project is metadata
-management connecting different object quality levels and connecting the user
-with the best possible outcomes given the current system state. The new user
+management connecting different object quality levels and enabling the user
+with the best possible outcomes given the user intent and the storage system state. The new user
 APIs required to interact with this rich metadata system will drive effective
@@
 fast overview of a data set with progressively more detail in areas of interest
-noted by data features and less data in areas without. These progressively
+such as those contain features, and less detail in areas without. These progressively
 more detailed data views require more storage space and time to retrieve. By
-incorporating middleware informed of both how to identify relevant data
-features and the intended use locations, it will be able to selectively store
+incorporating middleware informed of either the criteria for identifying areas of interest
+or the intended use locations, it will be able to selectively store
 data at different quality levels, with bounded errors in the case of lossy
 compression, and allow a user to select what data to retrieve by specifying
-both an acceptable retrieval timeframe and quality required. If the request
-cannot be fulfilled, the storage system can negotiate with the user providing
-information about the quality and timeframes available for a subsequent
-request.
+both an acceptable retrieval timeframe and error bound. 
+%If the request
+%cannot be fulfilled, the storage system can negotiate with the user providing
+%information about the quality and timeframes available for a subsequent
+%request.
+%%%%%%%%%%%% I am not sure if we should do this type of expert system in the project.
+The user may also specify guidlines that are funneled through I/O middle and 
+advise how to react if the timeframe and error bound cannot be accepted.
 
@@
 realization of tradeoffs in data representation, data placement and data
-access. We will allow users to ``plug-in'' their knowledge about the data, not
-as bytes but as motifs, allowing the I/O and storage system to understand user
+access. We will allow users to be able to ``plug-in'' their knowledge 
+about the data, not as bytes but as motifs, thus allowing the I/O and 
+storage system to understand user
 intentions and the relationships between data objects, as well as data access
 and transport patterns. This will facilitate efficient mapping of data from
-user space, such as a multi-dimensional array, onto various storage tiers.
-Rather than focus on simple data compression, we will incorporate a multitude
-of techniques to reduce data on the faster yet smaller storage tiers, and keep
-the less reduced information in the lower storage tiers.
+user space onto various storage tiers, and application-guided
+data reductions.
 
-We will support additional data access modes as well. \textbf{First}, within a
-given accuracy bounds, offer a mode for re-computation from data stored in a
-``fast'' tier to reduce data latencies. \textbf{Second}, exploratory analysis
+The I/O and storage system produced from the project will
+enable highly flexible data access modes that are common in data analytics. 
+\textbf{First}, data re-computation is allowed to re-generate
+data with higher accuracy, if the associated computing and I/O cost is lower
+than retrieving the target data from a ``slow'' tier. 
+\textbf{Second}, exploratory analysis
 operations frequently entail an overall data set view followed by targeted data
-exploration based on identified features. We will offer support for a
-configurable data access mode to support these sorts of analysis accesses. An
-``overview'' access mode that gives a quick, approximate within error bounds,
+exploration on extracted features. 
+In particular, the ``overview'' access mode gives a quick but approximate
 data view that can guide feature selection offering rapid coarse-grained data
-exploration without requiring loading the complete, detailed data set from
+exploration without requiring loading the high-accuracy data from
 storage. Based on the granularity requested, the accuracy and size of the data
-returned can be adjusted. At the most extreme setting, the original data can
+returned can be adjusted. In the most extreme setting, the original data can
 be retrieved at the time cost of moving the potentially huge data quantity.

### 2015-06-06 LOI/inc2.tex
Worked on the first part.

@@
   \definecolor{darkgreen}{rgb}{0,0.5,0}
-  \newcommand{\manish}[1]{ {\it \color{red} \{#1 -Tahsin\}}}
+  \newcommand{\manish}[1]{ {\it \color{red} \{#1 -Manish}}}
   \newcommand{\hasan}[1]{ {\textcolor{blue} { #1 -Hasan }}}
@@
 produced during the simulation, and analyzed for months afterwards.  Our goal
-in this project is to address the associated I/O and storage challenges on the
-future storage landscape, and expedite insights into mission critical
-scientific processes. To that end, we will build on the capabilities offered
-by ADIOS that provides I/O abstractions, 
-the Sirocco peer-to-peer file system under development at Sandia and the
-object storage and annotation expertise of UC Santa Cruz to develop a unique
-application-aware data management scheme. It will center upon
-a multi-tier storage-aware approach and incorporate new
-functionalities and extended object metadata and APIs for 1) matching
-storage capability across the storage hierarchy with data importance employing
-application-aware data compression (both lossy with specified error bounds and
-lossless), 2) a selectable performance/quality tradeoff when reading data, and
-the ability to recompute data if needed, and
-3) incorporate reader prioritization for data annotation, placement, and data
-quality when writing data.  The major challenge in this project is metadata
-management connecting different object quality levels and enabling the user
-with the best possible outcomes given the user intent and the storage system state. The new user
-APIs required to interact with this rich metadata system will drive effective
-use of the entire middleware and storage stack. The project is compriesed of a
-team with strong expertise in I/O middleware (ORNL, Rutgers), file system (SNL,
+in this project is to address associated I/O and storage challenges in the context 
+of current and emerging storage landscapes, and expedite insights into mission 
+critical scientific processes. To that end, we will build on the capabilities offered
+by ADIOS and DataSpaces that provide I/O abstractions and services, the Sirocco 
+peer-to-peer file system under development at Sandia and the object storage 
+and annotation expertise of UC Santa Cruz, to develop a application and 
+multi-tier storage aware data management solution. Our solutions will provide
+novel functionalities and APIs for 
+% It will center upon a multi-tier storage-aware approach and incorporate new
+%functionalities, extended object metadata and APIs for 
+1) specifying, at the application level, data annotations than enable the qualification
+ of the relative importance and utility of data objects, and enable them to be 
+ mapped at runtime to appropriate storage capabilities across the storage hierarchy;
+2) specifying selectable performance/quality/cost tradeoffs from both application and system perspectives; and
+3) evaluating and executing these tradeoffs at runtime, in an autonomic manner,
+during data placement and movement, using techniques such as application-aware 
+data compression (both lossy with specified error bounds and lossless), re-computation, 
+reader prioritization, etc.
+%2) a selectable performance/quality tradeoff when reading data, and
+%the ability to recompute data if needed, and
+%3) incorporate reader prioritization for data annotation, placement, and data
+%quality when writing data.  
+The key challenge in this project is defining and maintaining the metadata
+ connecting different object quality and utility levels, and using this metadata 
+ to most appropriate manage the placement and access of data object based on 
+ user/application intent/contraints and storage system state.
+% to best enabling the user with the best possible outcomes 
+ %given the user intent and the storage system state. 
+The new user APIs required to interact with this rich metadata system will drive effective
+use of the entire middleware and storage stack. 
+The project is comprised of a team with strong expertise in I/O middleware (ORNL, Rutgers), file system (SNL,
 UCSC) and storage (UCSC), and connect and coordinate these key storage
@@
 
+\manish{I believe this analogy needs to be developed further -- the point we are trying to make here, i.e., autonomic cross layer management, negotiation, are not completely obvious.}
 Our approach can most easily be understood by considering a road navigation
@@
 %%%%%%%%%%%% I am not sure if we should do this type of expert system in the project.
-The user may also specify guidlines that are funneled through I/O middle and 
-advise how to react if the timeframe and error bound cannot be accepted.
+The user may also specify guidelines that are funneled through the I/O middleware and 
+provide advice on how to react if the timeframe and error bound cannot be accepted.
 
 Our objective here is to reduce the time to knowledge, an end-to-end metric
-relevant to the scientific discovery process. Beyond the traditional high
-volume I/O pattern of checkpoint/restart, we will address the challenges posed
+relevant to the scientific discovery process. We will explore beyond the traditional high
+volume I/O pattern of checkpoint/restart, and will address the challenges posed
 by other essential data access patterns in the knowledge gathering process.
@@
 accuracy and errors as optimization parameters. 
-
 For example, scientific simulations contain approximations, as do measurements
@@
 approximations are acceptable. We can leverage this observation to optimize the
-presentation of data to the user and to implement various tradeoffs. For
-example, users can ask for information within a given accuracy bound, allowing
-us to offer a mode for re-computation vs. data storage and reterival.
+presentation of data to the user and to implement various tradeoffs  -- users can 
+ask for information within a given accuracy bound, allowing
+us to offer a mode for re-computation vs. data storage and retrieval.
 
@@
 we can accurate decide how to place and organize data for the critical
-capability runs. Fourth, we anticipate storing multiple data copies, each
+capability runs. 
+\textbf{Fourth}, we anticipate storing multiple data copies, each
 compressed in different ways according to the underlying media, some of these

### 2015-06-06 LOI/inc2.tex
no message

@@
 2) specifying selectable performance/quality/cost tradeoffs from both application and system perspectives; and
-3) evaluating and executing these tradeoffs at runtime, in an autonomic manner,
-during data placement and movement, using techniques such as application-aware 
-data compression (both lossy with specified error bounds and lossless), re-computation, 
-reader prioritization, etc.
+3) evaluating and executing these tradeoffs at runtime during data placement and movement, 
+in an autonomic manner using models, heuristics and learning and leveraging  
+techniques such as application-aware data compression (both lossy with specified error bounds 
+and lossless), re-computation,  reader prioritization, etc.
 %2) a selectable performance/quality tradeoff when reading data, and
@@
 The I/O and storage system produced from the project will
-enable highly flexible data access modes that are common in data analytics. 
+enable highly flexible data access modes that are common in scientific 
+applications and associated analysis. 
 \textbf{First}, data re-computation is allowed to re-generate
@@
 \textbf{Third}, to ensure available storage for subsequent operations, we will
-offer automatic data migration based on user annotations for required data
-lifetimes using monitoring and learning techniques. Unlike existing approaches,
-this will be tempered both by the user annotations and through learned access
-patterns.  While past access patterns may not indicate future access because
+offer automatic data migration.
+%based on user annotations for required data and using monitoring and learning techniques. 
+ Unlike existing approaches,
+this will be driven both by the user annotations and through learned access
+patterns in an autonomic and crosslayer manner.  While past access patterns may not indicate future access because
 the simulation run purpose may have changed, we are focused on scalability

### 2015-06-06 LOI/inc2.tex
One big comment that people should look through. Otherwise mostly some focus changes. I am not a fan of the analogy with GPS navigation, more so with what Carlos had said about logsitics. Also added some text about the data life-time as a metric for guiding purges.

@@
 \usepackage{amsfonts}
-%\usepackage{graphicx}
-%\usepackage{xspace}
+%  \usepackage{graphicx}
+  %  \usepackage{xspace}
 \usepackage{verbatim}
-%\usepackage{listings}
+%  \usepackage{listings}
 \usepackage{multirow}
@@
 % comment out the next line to turn off comments
-%\drafttrue
+% \drafttrue
+\drafttrue
 
@@
 \definecolor{darkgreen}{rgb}{0,0.5,0}
-  \newcommand{\manish}[1]{ {\it \color{red} \{#1 -Manish}}}
-  \newcommand{\hasan}[1]{ {\textcolor{blue} { #1 -Hasan }}}
+\newcommand{\manish}[1]{{\it \color{red} #1 -Manish}}
+\newcommand{\hasan}[1]{{\it \color{darkgreen} #1 -Hasan }}
 \definecolor{orange}{rgb}{0.7,0.5,0.0}
-  \newcommand{\klasky}[1]{{\textcolor{orange}{ #1 -Scott }}}
+\newcommand{\klasky}[1]{{\textcolor{orange} #1 -Scott }}
+  % Red star denotes items that need further work or discussio
+\newcommand{\TODO}[1]{\textcolor{red}{ TO DO: #1 }}
+\newcommand{\ALT}[1]{{\color{red} #1 - {\bf Alternate Text}}}
+\else
+\definecolor{darkgreen}{rgb}{0,0.5,0}
+\newcommand{\manish}[1]{}
+\definecolor{orange}{rgb}{0.7,0.5,0.0}
 % Red star denotes items that need further work or discussion
-  \newcommand{\TODO}[1]{{\textcolor{red}{ TO DO: #1 }}}
-\else
+\newcommand{\TODO}[1]{}
 \newcommand{\klasky}[1]{}
 \newcommand{\hasan}[1]{}
+\newcommand{\ALT}[1]{}}
 \fi
 
-%\newcommand{\ititle}{\textsc{\textbf{HESK}}}
+%  \newcommand{\ititle}{\textsc{\textbf{HESK}}}
 \newcommand{\ititle}{\textsc{HESK}}
@@
 \end{comment}
-%\includepdf[pages={1}]{ornl-cover.pdf}
-%\pagebreak
+%  \includepdf[pages={1}]{ornl-cover.pdf}
+  %  \pagebreak
 
@@
 
-%\maketitle
+%  \maketitle
 
@@
 
-\underline{\textbf{Objectives:}}
-Exascale scientific discovery will be severely bottlenecked without sufficient
-new research into managing and storing the large amounts of data that will be
-produced during the simulation, and analyzed for months afterwards.  Our goal
-in this project is to address associated I/O and storage challenges in the context 
-of current and emerging storage landscapes, and expedite insights into mission 
-critical scientific processes. To that end, we will build on the capabilities offered
-by ADIOS and DataSpaces that provide I/O abstractions and services, the Sirocco 
-peer-to-peer file system under development at Sandia and the object storage 
-and annotation expertise of UC Santa Cruz, to develop a application and 
-multi-tier storage aware data management solution. Our solutions will provide
-novel functionalities and APIs for 
+\underline{\textbf{Objectives:}} Exascale scientific discovery will be
+severely bottlenecked without sufficient new research into managing and
+storing the large amounts of data that will be produced during the
+simulation, and analyzed for months afterwards.  Our goal in this project is
+to address the associated I/O and storage challenges in the context of
+current ancd emerging storage landscapes, and expedite insights into mission
+critical scientific processes. To that end, we will build on the
+capabilities offered by ADIOS and DataSpaces that provide I/O abstractions
+and services, the Sirocco peer-to-peer file system under development at
+Sandia and the object storage and annotation expertise of UC Santa Cruz, to
+explore application and multi-tier storage aware data management
+solutions. Our solutions will provide novel functionalities and APIs for
 % It will center upon a multi-tier storage-aware approach and incorporate new
-%functionalities, extended object metadata and APIs for 
-1) specifying, at the application level, data annotations than enable the qualification
- of the relative importance and utility of data objects, and enable them to be 
- mapped at runtime to appropriate storage capabilities across the storage hierarchy;
-2) specifying selectable performance/quality/cost tradeoffs from both application and system perspectives; and
-3) evaluating and executing these tradeoffs at runtime during data placement and movement, 
-in an autonomic manner using models, heuristics and learning and leveraging  
-techniques such as application-aware data compression (both lossy with specified error bounds 
-and lossless), re-computation,  reader prioritization, etc.
-%2) a selectable performance/quality tradeoff when reading data, and
-%the ability to recompute data if needed, and
-%3) incorporate reader prioritization for data annotation, placement, and data
-%quality when writing data.  
-The key challenge in this project is defining and maintaining the metadata
+% functionalities, extended object metadata and APIs for
+
+%
+1) specifying, at the application level, data annotations that enable the
+qualification of the relative importance and utility of data objects, and
+enable them to be mapped at runtime to appropriate storage capabilities
+across the storage hierarchy;
+%
+2) specifying selectable performance/quality/cost tradeoffs from both the
+application and system perspectives; and
+%
+3) evaluating these tradeoffs at runtime during data placement and movement,
+and executing the resultant policies in an autonomic system using models,
+heuristics and continous learning,
+%
+4) Leveraging techniques such as application-aware data compression (both
+lossy with specified error bounds and lossless), re-computation at the
+requisite level of accuracy, and I/O prioritization to enforce these
+policies.
+%  2) a selectable performance/quality tradeoff when reading data, and
+  %  the ability to recompute data if needed, and
+  %  3) incorporate reader prioritization for data annotation, placement, and data
+  %  quality when writing data.  
+A key challenge in this project is defining and maintaining the metadata
 connecting different object quality and utility levels, and using this metadata 
- to most appropriate manage the placement and access of data object based on 
+to most appropriately manage the placement and access of data object based on 
 user/application intent/contraints and storage system state.
 % to best enabling the user with the best possible outcomes 
- %given the user intent and the storage system state. 
-The new user APIs required to interact with this rich metadata system will drive effective
-use of the entire middleware and storage stack. 
-The project is comprised of a team with strong expertise in I/O middleware (ORNL, Rutgers), file system (SNL,
-UCSC) and storage (UCSC), and connect and coordinate these key storage
-components in a seamless fashion.
+% given the user intent and the storage system state. 
+The new user APIs required to interact with this rich metadata system will
+drive effective use of the entire middleware and storage stack.  The project
+is comprised of a team with strong expertise in I/O middleware (ORNL,
+Rutgers), file system (SNL, UCSC) and storage (UCSC), and connect and
+coordinate these key storage components in a seamless fashion.
 
-\manish{I believe this analogy needs to be developed further -- the point we are trying to make here, i.e., autonomic cross layer management, negotiation, are not completely obvious.}
+\manish{I believe this analogy needs to be developed further -- the point we
+  are trying to make here, i.e., autonomic cross layer management,
+  negotiation, are not completely obvious.}
+\hasan{I am not a fan of this analogy either. I liked the walmart logistics
+  one better, it more correctly represented the biggest challenge we have,
+  viz. negotiating in a complex system where objects should be placed. The
+  partition of data is an addendum to the organization (thus moving us away
+  from simple logistical solutions. I am going add an alternate analogy as a
+  comparison}
+
+\ALT{Our approach can be most easily understood by considering a traditional
+  supermarket logistics problem. Given a limited amount of shelf space,
+  fluctuating user demands and priorities, and the need to optimize profit at
+  a market wide level instead of per product, the logistical problem combines
+  predictions about the variants and the indicated constraints to find a
+  solution. Self space is itself not a homogenous, with different locations
+  providing greater rewards to the placed items. Furthermore, haphazard
+  organization of items produces suboptimal results because it would ignore
+  locality in purchasing. Backroom and warehouses represent tiers of storage
+  where objects can be stored in bulk, but have a substantial overhead (and
+  thus lower utility) in access. Similar to this scenario, the data management
+  problem is a complex logistics problem. Shelf space is analagous to high
+  performance/lower capacity storage. Locality is an important consideration,
+  not just for single data objects but also through the identification of
+  domain meaningful relationships between objects. The tiers of the storage
+  hierarchy provide lower cost for data storage, but also increase the
+  overhead for access.}
+
 Our approach can most easily be understood by considering a road navigation
 program. Similar to the navigation program, we will offer a coarse grained,
-fast overview of a data set with progressively more detail in areas of interest
-such as those contain features, and less detail in areas without. These progressively
-more detailed data views require more storage space and time to retrieve. By
-incorporating middleware informed of either the criteria for identifying areas of interest
-or the intended use locations, it will be able to selectively store
-data at different quality levels, with bounded errors in the case of lossy
-compression, and allow a user to select what data to retrieve by specifying
-both an acceptable retrieval timeframe and error bound. 
-%If the request
-%cannot be fulfilled, the storage system can negotiate with the user providing
-%information about the quality and timeframes available for a subsequent
-%request.
+fast overview of a data set with progressively more detail in areas of
+interest such as those contain features, and less detail in areas
+without. These progressively more detailed data views require more storage
+space and time to retrieve. By incorporating middleware informed of either
+the criteria for identifying areas of interest or the intended use
+locations, it will be able to selectively store data at different quality
+levels, with bounded errors in the case of lossy compression, and allow a
+user to select what data to retrieve by specifying both an acceptable
+retrieval timeframe and error bound.
+%  If the request
+  %   cannot be fulfilled, the storage system can negotiate with the user providing
+   %    information about the quality and timeframes available for a subsequent
+    %     request.
      %%%%%%%%%%%%                 I am not sure if we should do this type of expert system in the project.
@@
 
-Our objective here is to reduce the time to knowledge, an end-to-end metric
-relevant to the scientific discovery process. We will explore beyond the traditional high
-volume I/O pattern of checkpoint/restart, and will address the challenges posed
-by other essential data access patterns in the knowledge gathering process.
-Through a deeper insight into the scientific process we will encode and utilize
-accuracy and errors as optimization parameters. 
-For example, scientific simulations contain approximations, as do measurements
-from observations and experiments, and depending on the goal, these
-approximations are acceptable. We can leverage this observation to optimize the
-presentation of data to the user and to implement various tradeoffs  -- users can 
-ask for information within a given accuracy bound, allowing
-us to offer a mode for re-computation vs. data storage and retrieval.
+\hasan{Since we are laying the foundation for a real storage system we can't
+  be solving the problem for a single application. We will eventually have
+  to consider things like fairness, availability and this is also why we
+  want to tie in with the resource management system} 
+%
+Our objective here is to reduce the time to knowledge, not just for a single
+application or workflow, but for the entire workload in the system. This is
+an end-to-end metric relevant to the scientific discovery process. We will
+explore beyond the traditional high volume I/O pattern of
+checkpoint/restart, and will address the challenges posed by other essential
+data access patterns in the knowledge gathering process.  Through a deeper
+insight into the scientific process we will encode and utilize accuracy and
+errors as optimization parameters.  For example, scientific simulations
+contain approximations, as do measurements from observations and
+experiments, and depending on the goal, these approximations are
+acceptable. We can leverage this observation to optimize the presentation of
+data to the user and to implement various tradeoffs -- users can ask for
+information within a given accuracy bound, allowing us to offer a mode for
+re-computation vs. data storage and retrieval.
 
 %  
-{\bf: please list other things from OLCF}.
+\TODO{\bf: please list other things from OLCF}
 %
@@
 realization of tradeoffs in data representation, data placement and data
-access. We will allow users to be able to ``plug-in'' their knowledge 
-about the data, not as bytes but as motifs, thus allowing the I/O and 
+access. We will allow users to be able to plug-in their knowledge 
+about the data, represented not as bytes but as motifs, allowing the I/O and 
 storage system to understand user
-intentions and the relationships between data objects, as well as data access
+intentions, the relationships between data objects, and also data access
 and transport patterns. This will facilitate efficient mapping of data from
@@
 
-The I/O and storage system produced from the project will
-enable highly flexible data access modes that are common in scientific 
-applications and associated analysis. 
-\textbf{First}, data re-computation is allowed to re-generate
-data with higher accuracy, if the associated computing and I/O cost is lower
-than retrieving the target data from a ``slow'' tier. 
-\textbf{Second}, exploratory analysis
-operations frequently entail an overall data set view followed by targeted data
-exploration on extracted features. 
-In particular, the ``overview'' access mode gives a quick but approximate
-data view that can guide feature selection offering rapid coarse-grained data
-exploration without requiring loading the high-accuracy data from
-storage. Based on the granularity requested, the accuracy and size of the data
-returned can be adjusted. In the most extreme setting, the original data can
-be retrieved at the time cost of moving the potentially huge data quantity.
-\textbf{Third}, to ensure available storage for subsequent operations, we will
-offer automatic data migration.
-%based on user annotations for required data and using monitoring and learning techniques. 
- Unlike existing approaches,
-this will be driven both by the user annotations and through learned access
-patterns in an autonomic and crosslayer manner.  While past access patterns may not indicate future access because
-the simulation run purpose may have changed, we are focused on scalability
-where runs are subsequently larger as the simulation prepares for a capability
-run.  By learning from the output and access patterns during this run sequence,
-we can accurate decide how to place and organize data for the critical
-capability runs. 
-\textbf{Fourth}, we anticipate storing multiple data copies, each
-compressed in different ways according to the underlying media, some of these
-copies will disappear based on storage pressures, but data persistence will be
-maintained according to user specifications. Assuming a relatively low latency
-cache layer before a tape system, we can offer exploratory data access
-reserving pulling data from tape to just the data required. This will save
-scientists time and make data stored on tape usable without long delays.
+The I/O and storage system prototype produced from the project will enable
+highly flexible data access modes that are common in scientific applications
+and associated analysis.
+%
+\hasan{Can we actually take some lower accuracy representation of data (low
+  granularity MR for example, and compute more stuff to provide a higher
+  accuracy representation? I am thinking of the work with Varis Carey with
+  storing boundary state instead of full volume and then using those
+  boundaries to fill in blanks over temporally changing data}
+%
+\textbf{First}, data re-computation is allowed to re-generate data with
+customizable accuracy, if the associated computing and I/O cost is lower
+than retrieving the target data from a {\em slow} tier.
+%
+\textbf{Second}, exploratory analysis operations frequently entail an
+overall data set view followed by targeted data exploration on extracted
+features.  In particular, the {\em overview} access mode gives a quick but
+approximate data view that can guide feature selection offering rapid
+coarse-grained data exploration without requiring loading the high-accuracy
+data from storage. Based on the granularity requested, the accuracy and size
+of the data returned can be adjusted. In the most extreme setting, the
+original data can be retrieved at the time cost of moving the potentially
+huge data quantity.
+%
+\textbf{Third}, to ensure available storage for subsequent operations, we
+will offer automatic data migration. based on user annotations for required data and using monitoring and learning techniques.  Unlike existing approaches, this will be
+driven both by the user annotations and through learned access patterns in
+an autonomic and crosslayer manner.  While past access patterns may not
+indicate future access because the simulation run purpose may have changed,
+we are focused on scalability where subsequent runs are larger as the
+simulation prepares for a capability run.  By learning from the output and
+access patterns during this run sequence, we can accurately decide how to
+place and organize data for the critical capability runs.
+%
+\textbf{Fourth}, we anticipate storing multiple data copies, each compressed
+in different ways according to the underlying media, some of these copies
+will disappear based on storage pressures, but data persistence will be
+maintained according to user specifications. Assuming a relatively low
+latency cache layer before a tape system, we can offer exploratory data
+access reserving pulling data from tape to just the data required. This will
+save scientists time and make data stored on tape usable without long
+delays. Additionally, we will incorporate data life time, at sub-object
+granularity, as a user described metric to allow the storage system to
+intelligently make decisions on purging and evicting data.
 
-Our research efforts will be heavily focused on the need to, in a coordinated
-manner, adapt data and metadata retention policies to the dynamic resource
-balancing that will need to take place between the application, OS/R, and
-hardware.
+Our research efforts will be heavily focused on the need to, in a
+coordinated manner, adapt data and metadata retention policies to the
+dynamic resource balancing that will need to take place between the
+application, OS/R, and hardware.
 
-The success of this project will  provide insights into how to build autonomic
-middleware and storage layers which can interact well with each other, and take
-user-provided hints. Today, data is reduced by application scientist who have
-limited information on what the storage layer can provide. They often make
-compromises based on this limited knowledge and either tune their output for
-writing or reading performance. This data then gets moved to other locations,
-and much of the tuning is lost when the data is read back during their post
-processing. Furthermore, there is a limited set of operations which users will
-be able to stage to other staging nodes for real-time-reduction and
-visualization.
+The success of this project will provide insights into how to build
+autonomic middleware and storage layers which can interact well with each
+other, and take user-provided hints. Today, data is reduced by application
+scientist who have limited information on what the storage layer can
+provide. They often make compromises based on this limited knowledge and
+either tune their output for writing or reading performance. This data then
+gets moved to other locations, and much of the tuning is lost when the data
+is read back during their post processing. Furthermore, there is a limited
+set of operations which users will be able to stage to other staging nodes
+for real-time-reduction and visualization.
 
-%\includepdf[pages={1}]{COI.pdf}
+%  \includepdf[pages={1}]{COI.pdf}
 
@@
 %%% End:
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2015-06-07 LOI/inc2.tex
minor changes to wording

@@
 % functionalities, extended object metadata and APIs for 
-
 %
@@
 application or workflow, but for the entire workload in the system. This is
-an end-to-end metric relevant to the scientific discovery process. We will
-explore beyond the traditional high volume I/O pattern of
+an end-to-end system wide metric relevant to the scientific discovery
+process. We will explore beyond the traditional high volume I/O pattern of
 checkpoint/restart, and will address the challenges posed by other essential
@@
 data to the user and to implement various tradeoffs -- users can ask for
-information within a given accuracy bound, allowing us to offer a mode for
+information within a given accuracy bound, allowing us to offer a tradeoff for
 re-computation vs. data storage and retrieval.
@@
 %  
-\TODO{\bf: please list other things from OLCF}
+% \TODO{\bf: please list other things from OLCF}
 %

### 2015-06-07 LOI/inc2.tex
Quick pass.

@@
 to address the associated I/O and storage challenges in the context of
-current ancd emerging storage landscapes, and expedite insights into mission
+current and emerging storage landscapes, and expedite insights into mission
 critical scientific processes. To that end, we will build on the
@@
 and executing the resultant policies in an autonomic system using models,
-heuristics and continous learning,
+heuristics and continuous learning,
 %
-4) Leveraging techniques such as application-aware data compression (both
+4) leveraging techniques such as application-aware data compression (both
 lossy with specified error bounds and lossless), re-computation at the
@@
 to most appropriately manage the placement and access of data object based on 
-user/application intent/contraints and storage system state.
+user/application intent/constraints and storage system state.
 % to best enabling the user with the best possible outcomes 
@@
 The new user APIs required to interact with this rich metadata system will
-drive effective use of the entire middleware and storage stack.  The project
-is comprised of a team with strong expertise in I/O middleware (ORNL,
+drive effective use of the entire middleware and storage stack.  This project
+brings together a unique team with strong expertise in I/O middleware (ORNL,
 Rutgers), file system (SNL, UCSC) and storage (UCSC), and connect and
@@
   thus lower utility) in access. Similar to this scenario, the data management
-  problem is a complex logistics problem. Shelf space is analagous to high
+  problem is a complex logistics problem. Shelf space is analogous to high
   performance/lower capacity storage. Locality is an important consideration,
@@
 
-Our approach can most easily be understood by considering a road navigation
-program. Similar to the navigation program, we will offer a coarse grained,
-fast overview of a data set with progressively more detail in areas of
-interest such as those contain features, and less detail in areas
+\manish{I have tried to rework the text below, but I am still not comfortable with it. I have left the original text commented.
+
+%Our approach can most easily be understood by considering a road navigation
+%program. Similar to the navigation program, 
+Incorporating awareness of the data content and user intent (to the extent possible) into
+our decision making and data management process is integral to our approach. For example, 
+we will offer a coarse grained, quick overview of a data set with progressively more detail in 
+areas of interest such as those containing features, and less detail in areas
 without. These progressively more detailed data views require more storage
-space and time to retrieve. By incorporating middleware informed of either
-the criteria for identifying areas of interest or the intended use
-locations, it will be able to selectively store data at different quality
-levels, with bounded errors in the case of lossy compression, and allow a
-user to select what data to retrieve by specifying both an acceptable
+space and time to retrieve. By defining mechanisms for applications to provide this 
+knowledge and incorporating its awareness into our middleware 
+%so that it is informed of either the criteria for identifying areas of interest or the intended use locations, 
+it will be able to selectively store data at different quality levels, with bounded errors in the case of 
+lossy compression, and allow a user to select what data to retrieve by specifying both an acceptable
 retrieval timeframe and error bound.

### 2015-06-08 LOI/inc2.tex
finally getting my changes in. More to come!

@@
   %   
-
-{\bf Title of Pre-application:} Hierarchal Extreme Scale Knowledge Management \par
+\setlength{\parindent}{0cm}
+{\bf Title of Pre-application:}  Extreme Scale Hierarchal Knowledge Management \par
 {\bf Principal Investigator:} Scott Klasky, Group Leader, ORNL, 854-241-9980, klasky@ornl.gov \par
 {\bf Funding Opportunity Announcement Number: DE-FOA-0001338} \par
-{\bf List of all co-PIs and Key/Senior Personnel} \par
-Hasan Abbasi, Oak Ridge National Laboratory \par
-Mark Ainsworth, Oak Ridge National Laboratory JFA, Brown University \par
-Matthew Curry, Sandia National Laboratory\par
-Qing Gary Liu, Oak Ridge National Laboratory \par
-Jay Lofstead, Sandia National Laboratory \par
-Kimmy Mu, Oak Ridge National Laboratory \par
-Carlos Malzahn, U. Cal. Santa Cruz \par
-Manish Parashar, Rutgers University, Oak Ridge National Laboratory \par
-Sudharshan Vazhkudai, Oak Ridge National Laboratory \par
-Lee Ward, Sandia National Laboratory \par
-\vskip 0.2 in
+ \vskip 0.1 in
+
+\begin{tabular} {ll}
+ \multicolumn{2}{l}{\bf Co-PIs and Key/Senior Personnel}\\
+Hasan Abbasi, ORNL  & Mark Ainsworth, ORNL JFA, Brown University \\
+Matthew Curry, SNL & Qing  Liu, ORNL \\
+Jay Lofstead, ORNL & Kimmy Mu, ORNL \\
+Carlos Malzahn, U. Cal. Santa Cruz & Manish Parashar, Rutgers University, ORNL \\
+Feiyi Wang, ORNL & Lee Ward, SNL
+\end{tabular}
+
+\setlength{\parindent}{0.5cm}
+
+\vskip 0.1 in
 
@@
 explore application and multi-tier storage aware data management
-solutions. Our solutions will provide novel functionalities and APIs for
-% It will center upon a multi-tier storage-aware approach and incorporate new
-% functionalities, extended object metadata and APIs for 
-%
+solutions.
+
+Our solutions will provide novel functionalities and APIs for
+\setlength{\parindent}{0cm}
 1) specifying, at the application level, data annotations that enable the
@@
 2) specifying selectable performance/quality/cost tradeoffs from both the
-application and system perspectives; and
+application and system perspectives;
 %
@@
 and executing the resultant policies in an autonomic system using models,
-heuristics and continuous learning,
+heuristics and continuous learning; and
 %
@@
   %  quality when writing data.
+
+  \setlength{\parindent}{0.5cm}
+  Our objective here is to reduce the time to knowledge, not just for a single
+application or workflow, but for the entire workload in the system. This is
+an end-to-end system wide metric relevant to the scientific discovery
+process. We will explore beyond the traditional high volume I/O pattern of
+checkpoint/restart, and will address the challenges posed by other essential
+data access patterns in the knowledge gathering process.  Through a deeper
+insight into the scientific process we will encode and utilize accuracy and
+errors as optimization parameters.  
+Finally, we will take the knowledge from the storage system to provide vital feedback to the middleware 
+so that the best possible decisions can be autonomically  between the user intentions and
+the available system resources. Our overall goal is to ensure that optimizations can be made across
+1) a single application, 2) an ensemble of applications, and 3) the entire suite of applications which are running on the system.
+
+  \setlength{\parindent}{0.5cm}
 A key challenge in this project is defining and maintaining the metadata
@@
 
-\manish{I believe this analogy needs to be developed further -- the point we
-  are trying to make here, i.e., autonomic cross layer management,
-  negotiation, are not completely obvious.}
-\hasan{I am not a fan of this analogy either. I liked the walmart logistics
-  one better, it more correctly represented the biggest challenge we have,
-  viz. negotiating in a complex system where objects should be placed. The
-  partition of data is an addendum to the organization (thus moving us away
-  from simple logistical solutions. I am going add an alternate analogy as a
-  comparison}
+%
+Our objective here is to reduce the time to knowledge, not just for a single
+application or workflow, but for the entire workload in the system. This is
+an end-to-end system wide metric relevant to the scientific discovery
+process. We will explore beyond the traditional high volume I/O pattern of
+checkpoint/restart, and will address the challenges posed by other essential
+data access patterns in the knowledge gathering process.  Through a deeper
+insight into the scientific process we will encode and utilize accuracy and
+errors as optimization parameters.
 
-\ALT{Our approach can be most easily understood by considering a traditional
+Ultimately, we will take the knowledge from the storage system to provide vital feedback to the middleware 
+so that the best possible decisions can be autonomically  between the user intentions and
+the available system resources. Our overall goal is to ensure that optimizations can be made across
+1) a single application, 2) an ensemble of applications, and 3) the entire suite of applications which are running on the system.
+Many of our prototypes will eventually be incorporated into the Sirocco peer-to-peer file system,
+and linked to the ADIOS middleware system. We will test our techniques on the Titan and Summitt systems
+at ORNL, along with the current and next generation machines with many of our application partner simulations
+such as XGC1, GTC, QMCPack, and SpecFM3D. The test will ensure that many of our research
+ideas can reduce the time to knowledge from these simulations and their corresponding analysis and visualization
+routines.
+
+ % For example, scientific simulations
+%contain approximations, as do measurements from observations and
+%experiments, and depending on the goal, these approximations are
+%acceptable. We can leverage this observation to optimize the presentation of
+%data to the user and to implement various tradeoffs -- users can ask for
+%information within a given accuracy bound, allowing us to offer a tradeoff for
+%re-computation vs. data storage and retrieval.
+
+%  
+% \TODO{\bf: please list other things from OLCF}
+%
+
+\underline{\textbf{Key Technical Approach:}}
+
+Our overall technical approach is based on an application-aware runtime
+realization of tradeoffs in data representation, data placement and data
+access. We will allow users to be able to plug-in their knowledge 
+about the data, represented not as bytes but as motifs, allowing the I/O and 
+storage system to understand user
+intentions, the relationships between data objects, and also data access
+and transport patterns. This will facilitate efficient mapping of data from
+user space onto various storage tiers, and application-guided
+data reductions.
+
+Our approach can be most easily understood by considering a traditional
   supermarket logistics problem. Given a limited amount of shelf space,
@@
   hierarchy provide lower cost for data storage, but also increase the
-  overhead for access.}
+  overhead for access.
 
-\manish{I have tried to rework the text below, but I am still not comfortable with it. I have left the original text commented.
+%\manish{I believe this analogy needs to be developed further -- the point we
+%  are trying to make here, i.e., autonomic cross layer management,
+%  negotiation, are not completely obvious.}
+%\hasan{I am not a fan of this analogy either. I liked the walmart logistics
+%  one better, it more correctly represented the biggest challenge we have,
+%  viz. negotiating in a complex system where objects should be placed. The
+%  partition of data is an addendum to the organization (thus moving us away
+%  from simple logistical solutions. I am going add an alternate analogy as a
+%  comparison}
+%
+%\manish{I have tried to rework the text below, but I am still not comfortable with it. I have left the original text commented.}
 
-%Our approach can most easily be understood by considering a road navigation
-%program. Similar to the navigation program, 
-Incorporating awareness of the data content and user intent (to the extent possible) into
+Incorporating awareness of the data content and user intent  into
 our decision making and data management process is integral to our approach. For example, 
@@
 knowledge and incorporating its awareness into our middleware 
-%so that it is informed of either the criteria for identifying areas of interest or the intended use locations, 
 it will be able to selectively store data at different quality levels, with bounded errors in the case of 
@@
 retrieval timeframe and error bound.
-%  If the request
-  %   cannot be fulfilled, the storage system can negotiate with the user providing
-   %    information about the quality and timeframes available for a subsequent
-    %     request.
-     %%%%%%%%%%%%                 I am not sure if we should do this type of expert system in the project.
-The user may also specify guidelines that are funneled through the I/O middleware and 
-provide advice on how to react if the timeframe and error bound cannot be accepted.
 
-\hasan{Since we are laying the foundation for a real storage system we can't
-  be solving the problem for a single application. We will eventually have
-  to consider things like fairness, availability and this is also why we
-  want to tie in with the resource management system} 
-%
-Our objective here is to reduce the time to knowledge, not just for a single
-application or workflow, but for the entire workload in the system. This is
-an end-to-end system wide metric relevant to the scientific discovery
-process. We will explore beyond the traditional high volume I/O pattern of
-checkpoint/restart, and will address the challenges posed by other essential
-data access patterns in the knowledge gathering process.  Through a deeper
-insight into the scientific process we will encode and utilize accuracy and
-errors as optimization parameters.  For example, scientific simulations
-contain approximations, as do measurements from observations and
-experiments, and depending on the goal, these approximations are
-acceptable. We can leverage this observation to optimize the presentation of
-data to the user and to implement various tradeoffs -- users can ask for
-information within a given accuracy bound, allowing us to offer a tradeoff for
-re-computation vs. data storage and retrieval.
+%The user may also specify guidelines that are funneled through the I/O middleware and 
+%provide advice on how to react if the timeframe and error bound cannot be accepted.
 
-%  
-% \TODO{\bf: please list other things from OLCF}
-%
-
-\underline{\textbf{Key Technical Approach:}}
-Our overall technical approach is based on an application-aware runtime
-realization of tradeoffs in data representation, data placement and data
-access. We will allow users to be able to plug-in their knowledge 
-about the data, represented not as bytes but as motifs, allowing the I/O and 
-storage system to understand user
-intentions, the relationships between data objects, and also data access
-and transport patterns. This will facilitate efficient mapping of data from
-user space onto various storage tiers, and application-guided
-data reductions.
+%\hasan{Since we are laying the foundation for a real storage system we can't
+%  be solving the problem for a single application. We will eventually have
+%  to consider things like fairness, availability and this is also why we
+%  want to tie in with the resource management system}
 
@@
 %
-\hasan{Can we actually take some lower accuracy representation of data (low
-  granularity MR for example, and compute more stuff to provide a higher
-  accuracy representation? I am thinking of the work with Varis Carey with
-  storing boundary state instead of full volume and then using those
-  boundaries to fill in blanks over temporally changing data}
+%\hasan{Can we actually take some lower accuracy representation of data (low
+%  granularity MR for example, and compute more stuff to provide a higher
+%  accuracy representation? I am thinking of the work with Varis Carey with
+%  storing boundary state instead of full volume and then using those
+%  boundaries to fill in blanks over temporally changing data}
 %

### 2015-06-08 LOI/inc2.tex
testing this
more changes... trying to get down to 2 pages. removed analogy since it doesn't add value to a pre proposal

@@
 {\bf Funding Opportunity Announcement Number: DE-FOA-0001338} \par
- \vskip 0.1 in
-
+ \vskip 0.0 in
+\
 \begin{tabular} {ll}
@@
 capabilities offered by ADIOS and DataSpaces that provide I/O abstractions
-and services, the Sirocco peer-to-peer file system under development at
+and services, the Sirocco peer-to-peer file system at
 Sandia and the object storage and annotation expertise of UC Santa Cruz, to
 explore application and multi-tier storage aware data management
-solutions.
+solutions.    This project
+brings together a  team with strong expertise in I/O middleware (ORNL,
+Rutgers), file system (SNL, UCSC) and storage (UCSC), and connect and
+coordinate these key storage components in a seamless fashion.
 
@@
 %
-4) leveraging techniques such as application-aware data compression (both
-lossy with specified error bounds and lossless), re-computation at the
+4) leveraging techniques such as application-aware data compression, re-computation at the
 requisite level of accuracy, and I/O prioritization to enforce these
@@
 an end-to-end system wide metric relevant to the scientific discovery
-process. We will explore beyond the traditional high volume I/O pattern of
+process. We will explore beyond the traditional I/O pattern of
 checkpoint/restart, and will address the challenges posed by other essential
@@
   \setlength{\parindent}{0.5cm}
-A key challenge in this project is defining and maintaining the metadata
-connecting different object quality and utility levels, and using this metadata 
-to most appropriately manage the placement and access of data object based on 
-user/application intent/constraints and storage system state.
-% to best enabling the user with the best possible outcomes 
-% given the user intent and the storage system state. 
-The new user APIs required to interact with this rich metadata system will
-drive effective use of the entire middleware and storage stack.  This project
-brings together a unique team with strong expertise in I/O middleware (ORNL,
-Rutgers), file system (SNL, UCSC) and storage (UCSC), and connect and
-coordinate these key storage components in a seamless fashion.
 
@@
 errors as optimization parameters. 
-
 Ultimately, we will take the knowledge from the storage system to provide vital feedback to the middleware 
 so that the best possible decisions can be autonomically  between the user intentions and
-the available system resources. Our overall goal is to ensure that optimizations can be made across
-1) a single application, 2) an ensemble of applications, and 3) the entire suite of applications which are running on the system.
-Many of our prototypes will eventually be incorporated into the Sirocco peer-to-peer file system,
-and linked to the ADIOS middleware system. We will test our techniques on the Titan and Summitt systems
-at ORNL, along with the current and next generation machines with many of our application partner simulations
-such as XGC1, GTC, QMCPack, and SpecFM3D. The test will ensure that many of our research
-ideas can reduce the time to knowledge from these simulations and their corresponding analysis and visualization
-routines.
+the available system resources.  
+We will test our prototypes  on current and future DOE system with many of today's applications, including the
+s XGC1, GTC, QMCPack, and SpecFM3D simulations.
 
@@
 realization of tradeoffs in data representation, data placement and data
-access. We will allow users to be able to plug-in their knowledge 
+access. We will allow users plug-in their knowledge 
 about the data, represented not as bytes but as motifs, allowing the I/O and 
@@
 
-Our approach can be most easily understood by considering a traditional
-  supermarket logistics problem. Given a limited amount of shelf space,
-  fluctuating user demands and priorities, and the need to optimize profit at
-  a market wide level instead of per product, the logistical problem combines
-  predictions about the variants and the indicated constraints to find a
-  solution. Self space is itself not a homogenous, with different locations
-  providing greater rewards to the placed items. Furthermore, haphazard
-  organization of items produces suboptimal results because it would ignore
-  locality in purchasing. Backroom and warehouses represent tiers of storage
-  where objects can be stored in bulk, but have a substantial overhead (and
-  thus lower utility) in access. Similar to this scenario, the data management
-  problem is a complex logistics problem. Shelf space is analogous to high
-  performance/lower capacity storage. Locality is an important consideration,
-  not just for single data objects but also through the identification of
-  domain meaningful relationships between objects. The tiers of the storage
-  hierarchy provide lower cost for data storage, but also increase the
-  overhead for access.
-
-%\manish{I believe this analogy needs to be developed further -- the point we
-%  are trying to make here, i.e., autonomic cross layer management,
-%  negotiation, are not completely obvious.}
-%\hasan{I am not a fan of this analogy either. I liked the walmart logistics
-%  one better, it more correctly represented the biggest challenge we have,
-%  viz. negotiating in a complex system where objects should be placed. The
-%  partition of data is an addendum to the organization (thus moving us away
-%  from simple logistical solutions. I am going add an alternate analogy as a
-%  comparison}
-%
-%\manish{I have tried to rework the text below, but I am still not comfortable with it. I have left the original text commented.}
+%Our approach can be most easily understood by considering a traditional
+%  supermarket logistics problem. Given a limited amount of shelf space,
+%  fluctuating user demands and priorities, and the need to optimize profit at
+%  a market wide level instead of per product, the logistical problem combines
+%  predictions about the variants and the indicated constraints to find a
+%  solution. Self space is itself not a homogenous, with different locations
+%  providing greater rewards to the placed items. Furthermore, haphazard
+%  organization of items produces suboptimal results because it would ignore
+%  locality in purchasing. Backroom and warehouses represent tiers of storage
+%  where objects can be stored in bulk, but have a substantial overhead (and
+%  thus lower utility) in access. Similar to this scenario, the data management
+%  problem is a complex logistics problem. Shelf space is analogous to high
+%  performance/lower capacity storage. Locality is an important consideration,
+%  not just for single data objects but also through the identification of
+%  domain meaningful relationships between objects. The tiers of the storage
+%  hierarchy provide lower cost for data storage, but also increase the
+%  overhead for access.
 
@@
 space and time to retrieve. By defining mechanisms for applications to provide this 
-knowledge and incorporating its awareness into our middleware 
-it will be able to selectively store data at different quality levels, with bounded errors in the case of 
-lossy compression, and allow a user to select what data to retrieve by specifying both an acceptable
-retrieval timeframe and error bound.
+knowledge and incorporating its awareness into our middleware  and storage allow
+it will be able to selectively store data at different  levels, and allow a user to select what data to retrieve by specifying both an acceptable  timeframe and accuracy.
 
@@
 
+A key challenge in this project is defining and maintaining the metadata
+connecting different object quality and utility levels, and using this metadata 
+to most appropriately manage the placement and access of data object based on 
+user/application intent/constraints and storage system state.
+The new user APIs required to interact with this rich metadata system will
+drive effective use of the entire middleware and storage stack.
+
 Our research efforts will be heavily focused on the need to, in a

### 2015-06-08 LOI/inc2.tex
Reducing it further. We must ensure that this is as much about storage as it is about the middleware.

@@
 
-The I/O and storage system prototype produced from the project will enable
+The I/O and storage system prototype will enable
 highly flexible data access modes that are common in scientific applications
@@
 \textbf{Third}, to ensure available storage for subsequent operations, we
-will offer automatic data migration. based on user annotations for required data and using monitoring and learning techniques.  Unlike existing approaches, this will be
+will offer automatic data migration, based on user annotations for required data and using monitoring and learning techniques.  Unlike existing approaches, this will be
 driven both by the user annotations and through learned access patterns in
@@
 latency cache layer before a tape system, we can offer exploratory data
-access reserving pulling data from tape to just the data required. This will
-save scientists time and make data stored on tape usable without long
-delays. Additionally, we will incorporate data life time, at sub-object
+access reserving pulling data from tape to just the data required. 
+ Additionally, we will incorporate data life time, at sub-object
 granularity, as a user described metric to allow the storage system to
@@
 user/application intent/constraints and storage system state.
-The new user APIs required to interact with this rich metadata system will
-drive effective use of the entire middleware and storage stack.
-
+%The new user APIs required to interact with this rich metadata system will
+%drive effective use of the entire middleware and storage stack.
 Our research efforts will be heavily focused on the need to, in a
@@
 autonomic middleware and storage layers which can interact well with each
-other, and take user-provided hints. Today, data is reduced by application
-scientist who have limited information on what the storage layer can
-provide. They often make compromises based on this limited knowledge and
-either tune their output for writing or reading performance. This data then
-gets moved to other locations, and much of the tuning is lost when the data
-is read back during their post processing. Furthermore, there is a limited
-set of operations which users will be able to stage to other staging nodes
-for real-time-reduction and visualization.
+other, and take user-provided hints.  This will be managed against the needs of
+individual simulations as well as what's happening on the entire system. This
+will reduce the time to knowledge by allowing user and system knowledge into
+the SSIO software layers.
+
+%Today, data is reduced by application
+%scientist who have limited information on what the storage layer can
+%provide.
+
+%They often make compromises based on this limited knowledge and
+%either tune their output for writing or reading performance. This data then
+%gets moved to other locations, and much of the tuning is lost when the data
+%is read back during their post processing. Furthermore, there is a limited
+%set of operations which users will be able to stage to other staging nodes
+%for real-time-reduction and visualization.
 

### 2015-06-08 LOI/inc2.tex
still need to finesse this better
Refocused technical approach to a storage focus.

@@
 %
-\textbf{First}, data re-computation is allowed to re-generate data with
-customizable accuracy, if the associated computing and I/O cost is lower
-than retrieving the target data from a {\em slow} tier.
+\textbf{First}, the storage system will offer the ability to directly store
+different portions of a single dataset--even a single variable at differnt
+levels of the storage hierarchy and apply different data transformation
+operations to each portion. For example, full data can be queued to tape while
+a highly compressed version intended for high-level analytical views can be
+stored in NVM or even RAM within the storage hierarchy. The storage system will
+support the metadata and plug-ins reuired to support this sort of data storage
+approach.
+%data re-computation is allowed to re-generate data with
+%customizable accuracy, if the associated computing and I/O cost is lower
+%than retrieving the target data from a {\em slow} tier.
 %
-\textbf{Second}, exploratory analysis operations frequently entail an
-overall data set view followed by targeted data exploration on extracted
-features.  In particular, the {\em overview} access mode gives a quick but
-approximate data view that can guide feature selection offering rapid
-coarse-grained data exploration without requiring loading the high-accuracy
-data from storage. Based on the granularity requested, the accuracy and size
-of the data returned can be adjusted. In the most extreme setting, the
-original data can be retrieved at the time cost of moving the potentially
-huge data quantity.
+\textbf{Second}, new storage access APIs will be developed that understand the
+notion of a time and data quality for requested data. The storage system,
+through the knowledge of both the data quality stored in differeht storage
+hierarchy tiers and the approximate time to retrieve data from the different
+tiers, the user can manage the tradeoff between data precision and time to
+retrieve. This storage system support will be managed through middleware
+insulating the user as much as possible from the new APIs. The storage system
+must also support plug-ins to potentially decompress or expand data to the
+original size. Both lossless and lossy with error bounds style storage is
+assumed. The time factor must take into account how long this operation takes
+for servicing the user request.
+%exploratory analysis operations frequently entail an
+%overall data set view followed by targeted data exploration on extracted
+%features.  In particular, the {\em overview} access mode gives a quick but
+%approximate data view that can guide feature selection offering rapid
+%coarse-grained data exploration without requiring loading the high-accuracy
+%data from storage. Based on the granularity requested, the accuracy and size
+%of the data returned can be adjusted. In the most extreme setting, the
+%original data can be retrieved at the time cost of moving the potentially
+%huge data quantity.
 %
-\textbf{Third}, to ensure available storage for subsequent operations, we
-will offer automatic data migration, based on user annotations for required data and using monitoring and learning techniques.  Unlike existing approaches, this will be
-driven both by the user annotations and through learned access patterns in
-an autonomic and crosslayer manner.  While past access patterns may not
-indicate future access because the simulation run purpose may have changed,
-we are focused on scalability where subsequent runs are larger as the
-simulation prepares for a capability run.  By learning from the output and
-access patterns during this run sequence, we can accurately decide how to
-place and organize data for the critical capability runs.
+\textbf{Third}, the storage system will offer annotations within the metadata
+and support for both predictive and reactive data placement and migration.
+While past access patterns may not indicate future access because the
+simulation run purpose may have changed, we are focused on scalability where
+subsequent runs are larger as the simulation prepares for a capability run.  By
+learning from the output and access patterns during this run sequence, we can
+accurately decide how to place and organize data for the critical capability
+runs. The reactive mechanisms will consider the space and performance
+characteristics, as well as the requested data fidelity, to determine where and
+how to pull and store data sets. In some cases, a full fidelity will be pulled
+from tape to disk and no further to preserve 100\% data fidelity. In others, a
+lossy-compressed within error bounds version may be generated and stored in
+local node RAM for very fast access with an acceptable error.
+%to ensure available storage for subsequent operations, we
+%will offer automatic data migration, based on user annotations for required
+%data and using monitoring and learning techniques.  Unlike existing
+%approaches, this will be driven both by the user annotations and through
+%learned access patterns in
+%an autonomic and crosslayer manner.
 %
-\textbf{Fourth}, we anticipate storing multiple data copies, each compressed
-in different ways according to the underlying media, some of these copies
-will disappear based on storage pressures, but data persistence will be
-maintained according to user specifications. Assuming a relatively low
-latency cache layer before a tape system, we can offer exploratory data
-access reserving pulling data from tape to just the data required. 
- Additionally, we will incorporate data life time, at sub-object
-granularity, as a user described metric to allow the storage system to
-intelligently make decisions on purging and evicting data.
+\textbf{Fourth}, to effectively manage data storage capacities, the storage
+system's capability for storing multiple, potentially different data quality
+versions of the same data will be leveraged to offer eviction and migration. The main challenge of this approach is to keep in mind the intentional placement
+decisions made to optimize future data access while ensuring proper system
+operation by not exhausting any tier inappropriately. If a single application
+has been allocated the entire machine for a capability run, completely
+consuming resources on the machine is expected. For smaller runs, these
+allocations must be balanced to support the diverse workload.
+%we anticipate storing multiple data copies, each compressed
+%in different ways according to the underlying media, some of these copies
+%will disappear based on storage pressures, but data persistence will be
+%maintained according to user specifications. Assuming a relatively low
+%latency cache layer before a tape system, we can offer exploratory data
+%access reserving pulling data from tape to just the data required. 
+% Additionally, we will incorporate data life time, at sub-object
+%granularity, as a user described metric to allow the storage system to
+%intelligently make decisions on purging and evicting data.
 

### 2015-06-08 LOI/inc2.tex
A pass on the entire LOI.
Commented out some text that look repeated.

@@
 brings together a  team with strong expertise in I/O middleware (ORNL,
-Rutgers), file system (SNL, UCSC) and storage (UCSC), and connect and
-coordinate these key storage components in a seamless fashion.
+Rutgers), file system (SNL, UCSC) and storage (UCSC), and connects and
+coordinates these key storage components in a seamless fashion.
 
@@
   \setlength{\parindent}{0.5cm}
-  Our objective here is to reduce the time to knowledge, not just for a single
-application or workflow, but for the entire workload in the system. This is
-an end-to-end system wide metric relevant to the scientific discovery
-process. We will explore beyond the traditional I/O pattern of
-checkpoint/restart, and will address the challenges posed by other essential
-data access patterns in the knowledge gathering process.  Through a deeper
-insight into the scientific process we will encode and utilize accuracy and
-errors as optimization parameters.  
-Finally, we will take the knowledge from the storage system to provide vital feedback to the middleware 
-so that the best possible decisions can be autonomically  between the user intentions and
-the available system resources. Our overall goal is to ensure that optimizations can be made across
-1) a single application, 2) an ensemble of applications, and 3) the entire suite of applications which are running on the system.
+
+ %% Manish: Commented this since it seems to be replicated below.  
+%  Our objective here is to reduce the time to knowledge, not just for a single
+%application or workflow, but for the entire workload in the system. This is
+%an end-to-end system wide metric relevant to the scientific discovery
+%process. We will explore beyond the traditional I/O pattern of
+%checkpoint/restart, and will address the challenges posed by other essential
+%data access patterns in the knowledge gathering process.  Through a deeper
+%insight into the scientific process we will encode and utilize accuracy and
+%errors as optimization parameters.  
+%Finally, we will take the knowledge from the storage system to provide vital feedback to the middleware 
+%so that the best possible decisions can be autonomically  between the user intentions and
+%the available system resources. Our overall goal is to ensure that optimizations can be made across
+%1) a single application, 2) an ensemble of applications, and 3) the entire suite of applications which are running on the system.
 
@@
 realization of tradeoffs in data representation, data placement and data
-access. We will allow users plug-in their knowledge 
+access. We will allow users to plug-in their knowledge 
 about the data, represented not as bytes but as motifs, allowing the I/O and 
@@
 user space onto various storage tiers, and application-guided
-data reductions.
+data reductions/transformations.
 
@@
 space and time to retrieve. By defining mechanisms for applications to provide this 
-knowledge and incorporating its awareness into our middleware  and storage allow
-it will be able to selectively store data at different  levels, and allow a user to select what data to retrieve by specifying both an acceptable  timeframe and accuracy.
+knowledge and incorporating its awareness into the middleware and storage,
+we will be able to selectively store data at different levels, and allow a user to select 
+not only what data to retrieve but also specify an acceptable timeframe and accuracy.
 
@@
 
-The I/O and storage system prototype will enable
+Our proposed I/O and storage system prototype will enable
 highly flexible data access modes that are common in scientific applications
@@
 \textbf{First}, the storage system will offer the ability to directly store
-different portions of a single dataset--even a single variable at differnt
+different portions of a single dataset, even a single variable, at different
 levels of the storage hierarchy and apply different data transformation
@@
 stored in NVM or even RAM within the storage hierarchy. The storage system will
-support the metadata and plug-ins reuired to support this sort of data storage
+support the metadata and plug-ins required to support this data storage
 approach.
@@
 \textbf{Second}, new storage access APIs will be developed that understand the
-notion of a time and data quality for requested data. The storage system,
-through the knowledge of both the data quality stored in differeht storage
+notion of time and data quality for requested data. The storage system,
+through knowledge of both the data quality stored in different storage
 hierarchy tiers and the approximate time to retrieve data from the different
-tiers, the user can manage the tradeoff between data precision and time to
-retrieve. This storage system support will be managed through middleware
+tiers, can manage (with the user in the loop) the tradeoff between data precision and time to
+retrieve. This storage system support will be executed and managed through our middleware, 
 insulating the user as much as possible from the new APIs. The storage system
-must also support plug-ins to potentially decompress or expand data to the
+must also support plug-ins, for example, to potentially decompress or expand data to the
 original size. Both lossless and lossy with error bounds style storage is
@@
 While past access patterns may not indicate future access because the
-simulation run purpose may have changed, we are focused on scalability where
-subsequent runs are larger as the simulation prepares for a capability run.  By
+simulation details may have changed, we are focused on scalability where
+subsequent runs are larger in scale as the simulation prepares for a capability run.  By
 learning from the output and access patterns during this run sequence, we can
@@
 characteristics, as well as the requested data fidelity, to determine where and
-how to pull and store data sets. In some cases, a full fidelity will be pulled
+how to pull and store data sets. In some cases, a full fidelity dataset may be pulled
 from tape to disk and no further to preserve 100\% data fidelity. In others, a
@@
 system's capability for storing multiple, potentially different data quality
-versions of the same data will be leveraged to offer eviction and migration. The main challenge of this approach is to keep in mind the intentional placement
+versions of the same data will be leveraged to guide data eviction and migration. 
+The main challenge of this approach is to be aware of intentional placement
 decisions made to optimize future data access while ensuring proper system
@@
 Our research efforts will be heavily focused on the need to, in a
-coordinated manner, adapt data and metadata retention policies to the
+coordinated manner, adapt data and metadata retention policies according to the 
 dynamic resource balancing that will need to take place between the
@@
 
-The success of this project will provide insights into how to build
-autonomic middleware and storage layers which can interact well with each
-other, and take user-provided hints.  This will be managed against the needs of
-individual simulations as well as what's happening on the entire system. This
-will reduce the time to knowledge by allowing user and system knowledge into
-the SSIO software layers.
+The success of this project will provide insights into how to build extreme scale
+autonomic middleware and storage layers that can interact effectively with each
+other, and bring the user in the loop through user-provided hints, allowing user 
+and system knowledge to be incorporated into the SSIO software layers.
+Storage resources will be managed against the needs of individual simulations 
+as well as what's happening on the entire system. The impact of such a solution 
+will be a significant reduction in the time to knowledge, and the overall increase in 
+the effectiveness and utility of exascale simulations.
 

### 2015-06-08 LOI/inc2.tex
minor edits.

@@
 storage system to understand user
-intentions, the relationships between data objects, and also data access
+intentions, the relationships between data objects, as well as data access
 and transport patterns. This will facilitate efficient mapping of data from
@@
 Incorporating awareness of the data content and user intent into
-our decision making and data management process is integral to our approach. For example, 
+storage and data management process is integral to our approach. For example, 
 we will offer a coarse grained, quick overview of a data set with progressively more detail in 
-areas of interest such as those containing features, and less detail in areas
+areas of interest such as those with features, and less detail in areas
 without. These progressively more detailed data views require more storage
-space and time to retrieve. By defining mechanisms for applications to provide this 
+space and time to retrieve if left unmanaged. By defining mechanisms for applications to provide this 
 knowledge and incorporating its awareness into the middleware and storage,
@@
 \textbf{First}, the storage system will offer the ability to directly store
-different portions of a single dataset, even a single variable, at different
-levels of the storage hierarchy and apply different data transformation
-operations to each portion. For example, full data can be queued to tape while
+different versions of a single variable to different
+levels of the storage hierarchy and allow each version to be processed 
+differently. For example, full data can be queued to tape while
 a highly compressed version intended for high-level analytical views can be
-stored in NVM or even RAM within the storage hierarchy. The storage system will
-support the metadata and plug-ins required to support this data storage
-approach.
+stored in non-volatile memory (NVM) or even RAM within the storage hierarchy. 
+The storage system will support the metadata and plug-ins required to support
+this data storage approach.
 %data re-computation is allowed to re-generate data with
@@
 %
-\textbf{Second}, new storage access APIs will be developed that understand the
+\textbf{Second}, new storage access APIs will be developed that incorporates the
 notion of time and data quality for requested data. The storage system,
-through knowledge of both the data quality stored in different storage
+through the knowledge of both data quality stored in different storage
 hierarchy tiers and the approximate time to retrieve data from the different
-tiers, can manage (with the user in the loop) the tradeoff between data precision and time to
-retrieve. This storage system support will be executed and managed through our middleware, 
+tiers, can manage (with the user in the loop) the tradeoff between error bound and the time to
+retrieve data. This storage system support will be executed and managed through our middleware, 
 insulating the user as much as possible from the new APIs. The storage system

### 2015-06-08 LOI/COI/COI-lofstead.tex
added all of the kinds of relationship data

  \item Hasan Abassi, Oak Ridge National Laboratory, collaborator
  \item Andrea Arpaci-Dusseah, UW Madison, co-author
  \item Remzi Arpaci-Dusseau, UW Madison, co-author
  \item Roselyne Barreto, University of Chicago, co-author
  \item John Bent, EMC Corporation, co-author
  \item Michael Booth, Oracle, co-author
  \item Franck Cappello, Argonne National Laboratory, collaborator
  \item Jacqueline Chen, Sandia National Laboratories, co-author
  \item Julian Cummings, California Institute of Technology, co-author
  \item Jai Dayal, Georgia Institute of Technologies, collaborator
  \item Ciprian Docan, Rutgers, co-author
  \item Stephane Ethier, Princenton Univeristy, co-author
  \item Ada Gavrilovska, Georgia Institute of Technology, co-author
  \item Garth Gibson, Carnegie Mellon University, co-author
  \item Ray Grout, National Renewal Energy Laboratory, co-author
  \item Gregory Jean-Baptiste, Florida International University, co-author
  \item Ivo Jimenez, University of California Santa Cruz, co-author
  \item Scott Klasky, Oak Ridge National Laboratory, co-author
  \item Todd Kordenbrock, Hewlett-Packard, collaborator
  \item Seung-Hoe Ku, New York University, co-author
  \item Qing Liu, Oak Ridge National Laboratory, co-author
  \item Mike Lang, Los Alamos National Laboratory, collaborator
  \item Jack Lange, University of Pittsburgh, collaborator
  \item Jeremy Logan, Oak Ridge National Laboratory, co-author
  \item Xiaosong Ma, North Carolina State University, co-author
  \item Carlos Maltzahn, University of California Santa Cruz, collaborator
  \item Kathryn Mohror, Lawrence Livermore National Laboratory, collaborator
  \item Adam Moody, Lawrence Livermore National Laboratory, collaborator
  \item Ron Oldfield, Sandia National Laboratoriey, collaborator
  \item Manish Parashar, Rutgers, co-author
  \item Tom Peterka, Argonne National Laboratory, collaborator
  \item Norbert Podhorzski, Oak Ridge National Laboratory, co-author
  \item Milo Polte, WibiData, co-author
  \item Charles Reiss, University of California Berkeley, co-author
  \item Karsten Schwan, Georgia Institute of Technology, PhD advisor
  \item Arie Shoshani, Lawrence Berkeley National Laboratory, co-author
  \item Alexander Sim, Lawrence Berkeley National Laboratory, co-author
  \item Gregory Sjaardema, Sandia National Laboratories, co-author
  \item Sagar Thapaliya, University of Alabama, Birmingham, collaborator
  \item Yuan Tian, Oak Ridge National Laboratory, co-author
  \item Yandong Wang, Auburn University, co-author
  \item Meghan Wingate McClelland, Xyratex, co-author
  \item Weikuan Yu, Auburn University, co-author
  \item Fang Zheng, IBM, co-author

### 2015-06-08 LOI/COI/COI-lofstead.tex
fixed typo
removed extraneous text

@@
-  \item Hasan Abassi, Oak Ridge National Laboratory, collaborator
-  \item Andrea Arpaci-Dusseah, UW Madison, co-author
-  \item Remzi Arpaci-Dusseau, UW Madison, co-author
-  \item Roselyne Barreto, University of Chicago, co-author
-  \item John Bent, EMC Corporation, co-author
-  \item Michael Booth, Oracle, co-author
-  \item Franck Cappello, Argonne National Laboratory, collaborator
-  \item Jacqueline Chen, Sandia National Laboratories, co-author
-  \item Julian Cummings, California Institute of Technology, co-author
-  \item Jai Dayal, Georgia Institute of Technologies, collaborator
-  \item Ciprian Docan, Rutgers, co-author
-  \item Stephane Ethier, Princenton Univeristy, co-author
-  \item Ada Gavrilovska, Georgia Institute of Technology, co-author
-  \item Garth Gibson, Carnegie Mellon University, co-author
-  \item Ray Grout, National Renewal Energy Laboratory, co-author
-  \item Gregory Jean-Baptiste, Florida International University, co-author
-  \item Ivo Jimenez, University of California Santa Cruz, co-author
-  \item Scott Klasky, Oak Ridge National Laboratory, co-author
-  \item Todd Kordenbrock, Hewlett-Packard, collaborator
-  \item Seung-Hoe Ku, New York University, co-author
-  \item Qing Liu, Oak Ridge National Laboratory, co-author
-  \item Mike Lang, Los Alamos National Laboratory, collaborator
-  \item Jack Lange, University of Pittsburgh, collaborator
-  \item Jeremy Logan, Oak Ridge National Laboratory, co-author
-  \item Xiaosong Ma, North Carolina State University, co-author
-  \item Carlos Maltzahn, University of California Santa Cruz, collaborator
-  \item Kathryn Mohror, Lawrence Livermore National Laboratory, collaborator
-  \item Adam Moody, Lawrence Livermore National Laboratory, collaborator
-  \item Ron Oldfield, Sandia National Laboratoriey, collaborator
-  \item Manish Parashar, Rutgers, co-author
-  \item Tom Peterka, Argonne National Laboratory, collaborator
-  \item Norbert Podhorzski, Oak Ridge National Laboratory, co-author
-  \item Milo Polte, WibiData, co-author
-  \item Charles Reiss, University of California Berkeley, co-author
-  \item Karsten Schwan, Georgia Institute of Technology, PhD advisor
-  \item Arie Shoshani, Lawrence Berkeley National Laboratory, co-author
-  \item Alexander Sim, Lawrence Berkeley National Laboratory, co-author
-  \item Gregory Sjaardema, Sandia National Laboratories, co-author
-  \item Sagar Thapaliya, University of Alabama, Birmingham, collaborator
-  \item Yuan Tian, Oak Ridge National Laboratory, co-author
-  \item Yandong Wang, Auburn University, co-author
-  \item Meghan Wingate McClelland, Xyratex, co-author
-  \item Weikuan Yu, Auburn University, co-author
-  \item Fang Zheng, IBM, co-author
+Hasan Abassi, Oak Ridge National Laboratory, collaborator
+Andrea Arpaci-Dusseau, UW Madison, co-author
+Remzi Arpaci-Dusseau, UW Madison, co-author
+Roselyne Barreto, University of Chicago, co-author
+John Bent, EMC Corporation, co-author
+Michael Booth, Oracle, co-author
+Franck Cappello, Argonne National Laboratory, collaborator
+Jacqueline Chen, Sandia National Laboratories, co-author
+Julian Cummings, California Institute of Technology, co-author
+Jai Dayal, Georgia Institute of Technologies, collaborator
+Ciprian Docan, Rutgers, co-author
+Stephane Ethier, Princenton Univeristy, co-author
+Ada Gavrilovska, Georgia Institute of Technology, co-author
+Garth Gibson, Carnegie Mellon University, co-author
+Ray Grout, National Renewal Energy Laboratory, co-author
+Gregory Jean-Baptiste, Florida International University, co-author
+Ivo Jimenez, University of California Santa Cruz, co-author
+Scott Klasky, Oak Ridge National Laboratory, co-author
+Todd Kordenbrock, Hewlett-Packard, collaborator
+Seung-Hoe Ku, New York University, co-author
+Qing Liu, Oak Ridge National Laboratory, co-author
+Mike Lang, Los Alamos National Laboratory, collaborator
+Jack Lange, University of Pittsburgh, collaborator
+Jeremy Logan, Oak Ridge National Laboratory, co-author
+Xiaosong Ma, North Carolina State University, co-author
+Carlos Maltzahn, University of California Santa Cruz, collaborator
+Kathryn Mohror, Lawrence Livermore National Laboratory, collaborator
+Adam Moody, Lawrence Livermore National Laboratory, collaborator
+Ron Oldfield, Sandia National Laboratoriey, collaborator
+Manish Parashar, Rutgers, co-author
+Tom Peterka, Argonne National Laboratory, collaborator
+Norbert Podhorzski, Oak Ridge National Laboratory, co-author
+Milo Polte, WibiData, co-author
+Charles Reiss, University of California Berkeley, co-author
+Karsten Schwan, Georgia Institute of Technology, PhD advisor
+Arie Shoshani, Lawrence Berkeley National Laboratory, co-author
+Alexander Sim, Lawrence Berkeley National Laboratory, co-author
+Gregory Sjaardema, Sandia National Laboratories, co-author
+Sagar Thapaliya, University of Alabama, Birmingham, collaborator
+Yuan Tian, Oak Ridge National Laboratory, co-author
+Yandong Wang, Auburn University, co-author
+Meghan Wingate McClelland, Xyratex, co-author
+Weikuan Yu, Auburn University, co-author
+Fang Zheng, IBM, co-author

### 2015-06-08 LOI/inc2.tex
Reorganization of the introduction a bit. Also merged in the research challenges into the technical areas. I am not super happy with this but perhaps someone can modify for readability.

@@
 Jay Lofstead, ORNL & Kimmy Mu, ORNL \\
-Carlos Malzahn, U. Cal. Santa Cruz & Manish Parashar, Rutgers University, ORNL \\
+Carlos Malzahn, U. Cal. Santa Cruz & Manish Parashar, Rutgers University, ORNL JFA \\
 Feiyi Wang, ORNL & Lee Ward, SNL
@@
 critical scientific processes. To that end, we will build on the
-capabilities offered by ADIOS and DataSpaces that provide I/O abstractions
+capabilities offered by ADIOS and DataSpaces at ORNL that provide I/O abstractions
 and services, the Sirocco peer-to-peer file system at
@@
 solutions.    This project
-brings together a  team with strong expertise in I/O middleware (ORNL,
-Rutgers), file system (SNL, UCSC) and storage (UCSC), and connects and
+brings together a  team with strong expertise in I/O middleware (ORNL), file system (SNL, UCSC) and storage (UCSC), and connects and
 coordinates these key storage components in a seamless fashion.
+
+Our objective here is to reduce the time to knowledge, not just for a single
+application or workflow, but for the entire workload in the system. This is
+an end-to-end system wide metric relevant to the scientific discovery
+process. We will explore beyond the traditional high volume I/O pattern of
+checkpoint/restart, and will address the challenges posed by other essential
+data access patterns in the knowledge gathering process.  Through a deeper
+insight into the scientific process we will encode and utilize accuracy and
+errors as optimization parameters. 
+Ultimately, we will take the knowledge from the storage system to provide vital feedback to the middleware 
+so that the best possible decisions can be autonomically  between the user intentions and
+the available system resources.  
+We will test our prototypes  on current and future DOE system with many of today's applications, including the
+s XGC1, GTC, QMCPack, and SpecFM3D simulations.
 
@@
 \setlength{\parindent}{0cm}
+%
 1) specifying, at the application level, data annotations that enable the
-qualification of the relative importance and utility of data objects, and
-enable them to be mapped at runtime to appropriate storage capabilities
-across the storage hierarchy;
+quantification of the relative importance and utility of data objects and
+enable partitioning of objects across the storage hierarch;
 %
@@
 %
-4) leveraging techniques such as application-aware data compression, re-computation at the
-requisite level of accuracy, and I/O prioritization to enforce these
+4) leveraging techniques such as application-aware data compression, 
+%re-computation at the requisite level of accuracy, 
+and I/O prioritization to enforce these
 policies.
@@
 
-%
-Our objective here is to reduce the time to knowledge, not just for a single
-application or workflow, but for the entire workload in the system. This is
-an end-to-end system wide metric relevant to the scientific discovery
-process. We will explore beyond the traditional high volume I/O pattern of
-checkpoint/restart, and will address the challenges posed by other essential
-data access patterns in the knowledge gathering process.  Through a deeper
-insight into the scientific process we will encode and utilize accuracy and
-errors as optimization parameters. 
-Ultimately, we will take the knowledge from the storage system to provide vital feedback to the middleware 
-so that the best possible decisions can be autonomically  between the user intentions and
-the available system resources.  
-We will test our prototypes  on current and future DOE system with many of today's applications, including the
-s XGC1, GTC, QMCPack, and SpecFM3D simulations.
-
  % For example, scientific simulations
@@
 
-Our overall technical approach is based on an application-aware runtime
-realization of tradeoffs in data representation, data placement and data
-access. We will allow users to plug-in their knowledge 
+Our technical approach is based on the insight that storage system awareness
+of application data requirements provides a powerful tool for guiding key
+extreme scale challenges such as data access method,  data representation in
+storage, and data
+placement within the hierarchy.
+%Our overall technical approach is based on an application-aware runtime
+%realization of tradeoffs in data representation, data placement and data
+%access. 
+Towards this goal, we will allow users, both the developers of applications
+and science users running these applications, 
+ to plug-in their knowledge 
 about the data, represented not as bytes but as motifs, allowing the I/O and 
@@
 and transport patterns. This will facilitate efficient mapping of data from
-user space onto various storage tiers, and application-guided
-data reductions/transformations.
-
+user space onto multiple storage tiers, and enable application-guided
+data reductions/transformations to address capacity and bandwidth
+bottlenecks. 
+%
 %Our approach can be most easily understood by considering a traditional
@@
 %  overhead for access.
-
-Incorporating awareness of the data content and user intent into
-storage and data management process is integral to our approach. For example, 
+%
+ For example, based on user and application insights into how the data is used,
 we will offer a coarse grained, quick overview of a data set with progressively more detail in 
@@
 
-Our proposed I/O and storage system prototype will enable
-highly flexible data access modes that are common in scientific applications
-and associated analysis.
+% Our proposed I/O and storage system prototype will enable
+% highly flexible data access modes that are common in scientific applications
+% and associated analysis.
 %
@@
 %
-\textbf{First}, the storage system will offer the ability to directly store
+The proposed work will address the following technical areas;
+\\\textbf{First}, the storage system will offer the ability to directly store
 different versions of a single variable to different
@@
 a highly compressed version intended for high-level analytical views can be
-stored in non-volatile memory (NVM) or even RAM within the storage hierarchy. 
+stored in non-volatile memory (NVM) or even RAM within the storage
+hierarchy. 
+The key challenge here would be defining and maintaining the metadata
+connecting differrent object quality and utility levels, 
 The storage system will support the metadata and plug-ins required to support
@@
 insulating the user as much as possible from the new APIs. The storage system
-must also support plug-ins, for example, to potentially decompress or expand data to the
+will also support plug-ins, for example, to potentially decompress or expand data to the
 original size. Both lossless and lossy with error bounds style storage is
@@
 characteristics, as well as the requested data fidelity, to determine where and
-how to pull and store data sets. In some cases, a full fidelity dataset may be pulled
-from tape to disk and no further to preserve 100\% data fidelity. In others, a
-lossy-compressed within error bounds version may be generated and stored in
-local node RAM for very fast access with an acceptable error.
+how to pull and store data sets. The challenge here is to utilize this
+metadata to manage the placement and access of data object based on the
+defined constraints as well as storage system state.
+
+% In some cases, a full fidelity dataset may be pulled
+% from tape to disk and no further to preserve 100\% data fidelity. In others, a
+% lossy-compressed within error bounds version may be generated and stored in
+% local node RAM for very fast access with an acceptable error.
 %to ensure available storage for subsequent operations, we
@@
 allocations must be balanced to support the diverse workload. 
+Our research efforts will be heavily focused on the need to, in a
+coordinated manner, adapt data and metadata retention policies according to the 
+dynamic resource balancing that will need to take place between the
+application, OS/R, and hardware.
+
 %we anticipate storing multiple data copies, each compressed
@@
 
-A key challenge in this project is defining and maintaining the metadata
-connecting different object quality and utility levels, and using this metadata 
-to most appropriately manage the placement and access of data object based on 
-user/application intent/constraints and storage system state.
-%The new user APIs required to interact with this rich metadata system will
-%drive effective use of the entire middleware and storage stack.
-Our research efforts will be heavily focused on the need to, in a
-coordinated manner, adapt data and metadata retention policies according to the 
-dynamic resource balancing that will need to take place between the
-application, OS/R, and hardware.
+% A key challenge in this project is defining and maintaining the metadata
+% connecting different object quality and utility levels, and using this metadata 
+% to most appropriately manage the placement and access of data object based on 
+% user/application intent/constraints and storage system state.
+% %The new user APIs required to interact with this rich metadata system will
+% %drive effective use of the entire middleware and storage stack.
+% Our research efforts will be heavily focused on the need to, in a
+% coordinated manner, adapt data and metadata retention policies according to the 
+% dynamic resource balancing that will need to take place between the
+% application, OS/R, and hardware.
 

### 2015-06-09 LOI/inc2.tex
some edits on the first part.

@@
 Our objective here is to reduce the time to knowledge, not just for a single
-application or workflow, but for the entire workload in the system. This is
-an end-to-end system wide metric relevant to the scientific discovery
-process. We will explore beyond the traditional high volume I/O pattern of
+application, but for the entire workload in a multi-user environment,
+and allow for selectable data quality, by adjusting its accuracy and error, 
+in order to meet the time or resource constraint. 
+We will explore beyond the traditional high volume I/O pattern of
 checkpoint/restart, and will address the challenges posed by other essential
-data access patterns in the knowledge gathering process.  Through a deeper
-insight into the scientific process we will encode and utilize accuracy and
-errors as optimization parameters. 
+data access patterns in the knowledge gathering process.
 Ultimately, we will take the knowledge from the storage system to provide vital feedback to the middleware 
-so that the best possible decisions can be autonomically  between the user intentions and
+so that the best possible decisions can be autonomically made between the user intentions and
 the available system resources.  

### 2015-06-09 LOI/inc2.tex
Fixed a typo.

@@
 solutions.    This project
-brings together a  team with strong expertise in I/O middleware (ORNL), file system (SNL, UCSC) and storage (UCSC), and connects and
+brings together a  team with strong expertise in I/O middleware (ORNL, Rutgers), file system (SNL, UCSC) and storage (UCSC), and connects and
 coordinates these key storage components in a seamless fashion.
@@
 The key challenge here would be defining and maintaining the metadata
-connecting differrent object quality and utility levels, 
+connecting different object quality and utility levels, 
 The storage system will support the metadata and plug-ins required to support

### 2015-06-09 LOI/inc2.tex
added a couple of sentences about 'negotiation' stuff. Didn't write exactly about 'negotiation' - The word sounds to me there is a handshake going on.

--- 2015-06-09 LOI/inc2.tex
@@
 Our objective here is to reduce the time to knowledge, not just for a single
-application, but for the entire workload in a multi-user environment,
-and allow for selectable data quality, by adjusting its accuracy and error, 
+application, but for the entire workload in a multi-user environment, where
+the storage is shared among users.
+We achieve this goal by allowing selectable data quality, by trading its accuracy and error
 in order to meet the time or resource constraint. 
@@
 not only what data to retrieve but also specify an acceptable timeframe and accuracy.
+On the other hand, the storage system layer may also dynamically 
+adjust the level of data that will be retrieved
+for an individual application, if that's allowed, to maximize the overall system efficiency and 
+provide fairness among applications.
 

### 2015-06-09 LOI/inc2.tex
Get this down to 2 pages.

@@
 current and emerging storage landscapes, and expedite insights into mission
-critical scientific processes. To that end, we will build on the
+critical scientific processes; which is associated with {\bf theme two} of the FOA. To that end, we will build on the
 capabilities offered by ADIOS and DataSpaces at ORNL that provide I/O abstractions
@@
 in order to meet the time or resource constraint. 
-We will explore beyond the traditional high volume I/O pattern of
-checkpoint/restart, and will address the challenges posed by other essential
+We will explore beyond 
+%the traditional high volume I/O pattern of
+checkpoint/restart I/O, and will address the challenges posed by 
+%other essential
+key
 data access patterns in the knowledge gathering process.
@@
 
-Our solutions will provide novel functionalities and APIs for
+Our solutions will provide  new functionalities and APIs for
 \setlength{\parindent}{0cm}
@@
 quantification of the relative importance and utility of data objects and
-enable partitioning of objects across the storage hierarch;
+enable partitioning  across the storage hierarchy;
 %
@@
 
-Our technical approach is based on the insight that storage system awareness
+Our  approach is based on the insight that storage system awareness
 of application data requirements provides a powerful tool for guiding key
-extreme scale challenges such as data access method,  data representation in
+I/O challenges such as data access,  data representation in
 storage, and data
@@
 %access. 
-Towards this goal, we will allow users, both the developers of applications
-and science users running these applications, 
+Towards this goal, we will allow both the developers of applications
+and users running these applications, 
  to plug-in their knowledge 
-about the data, represented not as bytes but as motifs, allowing the I/O and 
-storage system to understand user
-intentions, the relationships between data objects, as well as data access
+about their data, represented not as bytes but as motifs, allowing the I/O and 
+storage system to understand their
+intentions, the relationships between the data objects, as well as access
 and transport patterns. This will facilitate efficient mapping of data from
-user space onto multiple storage tiers, and enable application-guided
+the user space onto multiple storage tiers, and enable application-guided
 data reductions/transformations to address capacity and bandwidth
@@
 %
- For example, based on user and application insights into how the data is used,
+ For example, based on insights into how the data is used,
 we will offer a coarse grained, quick overview of a data set with progressively more detail in 
@@
 without. These progressively more detailed data views require more storage
-space and time to retrieve if left unmanaged. By defining mechanisms for applications to provide this 
+space and time to retrieve if left unmanaged. By defining mechanisms to provide this 
 knowledge and incorporating its awareness into the middleware and storage,
@@
 not only what data to retrieve but also specify an acceptable timeframe and accuracy.
-On the other hand, the storage system layer may also dynamically 
+The storage system layer may also dynamically 
 adjust the level of data that will be retrieved
-for an individual application, if that's allowed, to maximize the overall system efficiency and 
+for an individual application,  to maximize the overall system efficiency and 
 provide fairness among applications.
-
 %The user may also specify guidelines that are funneled through the I/O middleware and 
 %provide advice on how to react if the timeframe and error bound cannot be accepted.
-
 %\hasan{Since we are laying the foundation for a real storage system we can't
@@
 %  want to tie in with the resource management system} 
-
 % Our proposed I/O and storage system prototype will enable
@@
 %
-The proposed work will address the following technical areas;
+The proposed work will address the following technical areas:
 \\\textbf{First}, the storage system will offer the ability to directly store
-different versions of a single variable to different
-levels of the storage hierarchy and allow each version to be processed 
+different portions of a single variable to different
+levels of the hierarchy and allow each version to be processed 
 differently. For example, full data can be queued to tape while
 a highly compressed version intended for high-level analytical views can be
-stored in non-volatile memory (NVM) or even RAM within the storage
+stored in non-volatile memory (NVM)  within the storage
 hierarchy. 
-The key challenge here would be defining and maintaining the metadata
-connecting different object quality and utility levels, 
+The key challenge would be defining and maintaining the metadata
+connecting different object quality and utility levels. 
 The storage system will support the metadata and plug-ins required to support
-this data storage approach.   
+this approach.   
 %data re-computation is allowed to re-generate data with
@@
 \textbf{Second}, new storage access APIs will be developed that incorporates the
-notion of time and data quality for requested data. The storage system,
-through the knowledge of both data quality stored in different storage
-hierarchy tiers and the approximate time to retrieve data from the different
-tiers, can manage (with the user in the loop) the tradeoff between error bound and the time to
-retrieve data. This storage system support will be executed and managed through our middleware, 
+notion of time and data quality. The storage system,
+through the knowledge of both the data quality stored in different storage
+hierarchy tiers and the approximate time to retrieve it from the different
+tiers, can manage the tradeoff between error bound and the time to
+retrieve the data. This storage system support will be executed and managed through our middleware, 
 insulating the user as much as possible from the new APIs. The storage system
@@
 defined constraints as well as storage system state. 
-
 % In some cases, a full fidelity dataset may be pulled
@@
 The main challenge of this approach is to be aware of intentional placement
-decisions made to optimize future data access while ensuring proper system
+decisions made to optimize future  access while ensuring proper system
 operation by not exhausting any tier inappropriately. If a single application
-has been allocated the entire machine for a capability run, completely
+is allocated on the entire machine, completely
 consuming resources on the machine is expected. For smaller runs, these

### 2015-06-09 LOI/inc2.tex
include COI....
new title... Please help get a better title...

@@
 \setlength{\parindent}{0cm}
-{\bf Title of Pre-application:}  Extreme Scale Hierarchal Knowledge Management \par
+{\bf Title of Pre-application:}  Hierarchal Storage Management  for Extreme Scale Science\par
 {\bf Principal Investigator:} Scott Klasky, Group Leader, ORNL, 854-241-9980, klasky@ornl.gov \par
@@
 
-%  \includepdf[pages={1}]{COI.pdf}
-
+  \includepdf[pages={1}]{COI/COI-all.pdf}
+  \includepdf[pages={2}]{COI/COI-all.pdf}
 \end{document}

### 2015-06-09 LOI/inc2.tex
title
new title

@@
   \vskip .1in
-  Hierarchal Extreme Scale Knowledge Management
+ Science-driven Data Management for multi-tier storage
 
@@
 \setlength{\parindent}{0cm} 
-{\bf Title of Pre-application:}  Hierarchal Storage Management  for Extreme Scale Science\par
+{\bf Title of Pre-application:}  Science-driven Data Management for multi-tier storage\par
 {\bf Principal Investigator:} Scott Klasky, Group Leader, ORNL, 854-241-9980, klasky@ornl.gov \par

### 2015-06-09 LOI/inc2.tex
final version to submit to our office. You can still comment, but....

@@
 \setlength{\parindent}{0cm} 
-{\bf Title of Pre-application:}  Science-driven Data Management for multi-tier storage\par
+{\bf Title of Pre-application:}  Science-driven Data Management for multi-tiered storage\par
 {\bf Principal Investigator:} Scott Klasky, Group Leader, ORNL, 854-241-9980, klasky@ornl.gov \par
@@
  \vskip 0.0 in
-\
+\hskip -0.08in
 \begin{tabular} {ll}

### 2015-06-10 LOI/inc2.tex
Moved up the description to the first paragraph and commented out the other description.

@@
 storing the large amounts of data that will be produced during the
-simulation, and analyzed for months afterwards.  Our goal in this project is
+simulation, and analyzed for months afterwards. In this project we will
+demonstrate novel techniques to  facilitate efficient mapping of data from
+the user space onto multiple storage tiers, and enable application-guided
+data reductions/transformations to address capacity and bandwidth
+bottlenecks. 
+Our goal in this project is
 to address the associated I/O and storage challenges in the context of
@@
 intentions, the relationships between the data objects, as well as access
-and transport patterns. This will facilitate efficient mapping of data from
-the user space onto multiple storage tiers, and enable application-guided
-data reductions/transformations to address capacity and bandwidth
-bottlenecks. 
+and transport patterns. 
+% This will facilitate efficient mapping of data from
+% the user space onto multiple storage tiers, and enable application-guided
+% data reductions/transformations to address capacity and bandwidth
+% bottlenecks. 
 %

### 2015-06-10 LOI/inc2.tex
Fixed the header for the pre-proposal

@@
 {\bf Principal Investigator:} Scott Klasky, Group Leader, ORNL, 854-241-9980, klasky@ornl.gov \par
-{\bf Funding Opportunity Announcement Number: DE-FOA-0001338} \par
+{\bf Laboratory Announcement Number: LAB 15-1338, Funding/year: \$1.25M} \par
  \vskip 0.0 in

### 2015-06-10 LOI/inc2.tex
Fitting in 2 page again
add the variable parittioning thing that scott mentioned

@@
 simulation, and analyzed for months afterwards. In this project we will
-demonstrate novel techniques to  facilitate efficient mapping of data from
+demonstrate novel techniques to  facilitate efficient mapping of data
+objects, even partitioning individual variables, from
 the user space onto multiple storage tiers, and enable application-guided
@@
 \underline{\textbf{Key Technical Approach:}}
-
 Our  approach is based on the insight that storage system awareness

### 2015-06-11 LOI/inc2.tex
changed my affiliation since I am not looking to change jobs :-)

@@
 Matthew Curry, SNL & Qing  Liu, ORNL \\
-Jay Lofstead, ORNL & Kimmy Mu, ORNL \\
+Jay Lofstead, SNL & Kimmy Mu, ORNL \\
 Carlos Malzahn, U. Cal. Santa Cruz & Manish Parashar, Rutgers University, ORNL JFA \\

### 2015-06-17 narrative/deliverables.tex
added some outline material and instructions in the introduction

\section*{Deliverables}

\begin{itemize}
\item plug-in architecture for Sirocco
\item storage APIs to manipulate storing and retrieving data to specific tiers
\item metadata management for data stored in multiple tiers using different compression
\item metadata management for deeper data knowledge
\item middleware hooks to manage new storage APIs
\item example lossy and lossless compression plug-ins to demonstrate the effectiveness of this approach
\end{itemize}

### 2015-06-17 narrative/intro.tex
added some outline material and instructions in the introduction

@@
-We propose to explore the development of a performance engineering data pipeline that combines extraction of fine grained monitoring data, collation and curation of this data into a usable distributed object store, and in-situ analysis and visualization of the data to enable performance understanding at extreme scale for application developers and scientific users.
-Additionally we will explore the characterization of end-to-end pipelines for typical scientific workflows to better identify key optimization points and explore both application and system level optimization models for these data-centric tasks.
-The inherent complexity of extreme scale systems, combining heterogeneous processing architectures, deep memory hierarchy and severe power constraints, provides a challenge for performance engineering…%motivation
+\section*{Introduction}
+
+\begin{itemize}
+
+\item describe data size problems for getting scientific results out
+\item describe annotating and differentiating data storage according to both
+use and inherent data qualities (e.g., contains desireable features)
+\item describe idea of using heavy compression for less interesting areas and less or no compression for more interesting areas.
+\item describe inherent errors in scientific calculations
+\item describe lossy compression within error bounds
+\item describe how incorporating plug-ins for application aware data compression opportunities to aid data storage
+\item describe Sirocco's current state and future designs as a solid foundation on which to work
+\item describe how these new features can enhance a system like Sirocco to reduce data intensity
+\item describe metadata challenges that we will also have to face.
+
+\end{itemize}
+
+every section should include:
+
+Problem and background, one or more solution approaches, and investigation goals/questions.
+
+Technical Areas
+
+1. Application Intentions.\\
+2. Metadata Management (middleware and storage system)\\
+3. Pluggable infrastructure for data transformation (storage system)\\
+4. Intelligent data mapping to storage/memory hierarchies (middleware and storage system)\\
+5. Predictable Performance and resolution tradeoffs\\
+6. Data re-generation. (middleware & storage)\\
+7. Learning Motifs and data re-organization.\\

### 2015-06-17 narrative/management.tex
added some outline material and instructions in the introduction

\section*{Management}

\begin{itemize}

\item We need to manage things

\end{itemize}

### 2015-06-17 narrative/proposed-work.tex
added some outline material and instructions in the introduction

\section*{Proposed Work}

\begin{itemize}

\item Add plug-in architecture to Sirocco to support selective data compression
\item Add specific tier destinations command to Sirocco (assuming that the caching mechansism cannot achieve the desired effects).
\item Develop profiling system to determine how a data set is used during preparation runs prior to a capability run to determine how to optimize data placement for capability run analytics.
\item

\end{itemize}

### 2015-06-17 participant-table.tex
added some outline material and instructions in the introduction

@@
 \setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Hasan Abbasi} (\emph{habbasi@ornl.gov})}}}
-\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Karsten Schwan} (\emph{schwan@cc.gatech.edu})}}}
-\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Allen Malony} (\emph{malony@cs.uoregon.edu})}}}
+\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Mark Ainsworth} (\emph{schwan@cc.gatech.edu})}}}
+\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Matthew Curry} (\emph{malony@cs.uoregon.edu})}}}
+\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Scott Klasky} (\emph{malony@cs.uoregon.edu})}}}
+\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Qing Liu} (\emph{malony@cs.uoregon.edu})}}}
+\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Gerald Lofstead} (\emph{malony@cs.uoregon.edu})}}}
+\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Carlos Maltzahn} (\emph{malony@cs.uoregon.edu})}}}
+\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Kimmy Mu} (\emph{malony@cs.uoregon.edu})}}}
+\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Manish Parashar} (\emph{malony@cs.uoregon.edu})}}}
+\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{H. Lee Ward} (\emph{malony@cs.uoregon.edu})}}}
 \begin{tabularx}{\textwidth}{l>{\raggedright\arraybackslash}X}

### 2015-06-18 narrative/intro.tex
Some initial intro text

@@
-\section*{Introduction}
-
 \begin{itemize}
@@
 7. Learning Motifs and data re-organization.\\
+
+\section*{Introduction}
+
+Exascale scientific discovery will be severely bottlenecked without
+sufficient new research into managing and storing the large amounts of data
+that will be produced during the simulation, and analyzed for months
+afterwards.  
+%
+In this project we will demonstrate novel techniques to
+facilitate efficient mapping of data objects, even partitioning individual
+variables, from the user space onto multiple storage tiers, and enable
+application-guided data reductions/transformations to address capacity and
+bandwidth bottlenecks, while constraining the error to be within user
+provided bounds.
+%
+We will address the associated I/O and storage challenges in the context of
+current and emerging storage landscapes, and expedite insights into critical
+scientific processes, demonstrating the validity of our approach in key DOE
+domains.
+
+The metric we are most intrested in optimizing is time to knowledge. Current
+approaches to addressing the I/O bottleneck fall into two broad
+categories. Parallel file system approaches that optimize the throughput for
+an entire system, and I/O middleware approaches that optimize the
+performance of a single application. Both approaches have seen success but
+are unlikely to overcome the major obstactles in reaching exascale. Instead
+of trying to optimize throughput, we will seek to reduce the time to
+knowledge. This is the most significant metric for scientific applications
+where the desired outcome is not storing data, but rather executing a
+knowledge extraction process on the data. Moreover, we aim to perform the
+optimization not for a single application, but rather for workload of a
+multi-user multi-application system environment.
+
+Our approach leverages the expected characteristics of exascale storage
+hardware. The storage layer will be partitioned into multiple heterogenous
+tiers with vastly different performance characteristics. This difference
+between layers will be further exacerbated by the constraints on capacity
+and data lifetime within a layer. Tape archival storage will still maintain
+data long term, but access to this data will be orders of magnitude slower
+than the next layer and naively accessing data from archival storage will
+greatly impact productivity.
+
+We will achieve reduced time to knowledge using a combination of techniques;
+\begin{enumerate}
+\item Data annotations specified at the application level to quantify the
+  relative important, utility and lifetime of data objects;
+\item Partitioning of data objects across the storage hierarchy utilizing
+  the additional knowledge embedded in the annotations;
+\item Evaluating the tradeoffs at runtime to guide data placement, movement,
+  and migration across storage layers using models, heuristics and continuous
+  learning;
+\item Utilize the additional knowledge available about the data to perform
+  application-aware data compression and I/O prioritization;
+\end{enumerate}

### 2015-06-18 participant-table.tex
Added people in the personnel.config and institutions.config file. Everyone should check their email/name/spell/affiliations in the file to make sure its correct.

--- 2015-05-22 participant-table.tex
@@
-% Generated by bin/config-participant-table on Fri Jan 17 21:40:42 EST 2014
+% Generated by bin/config-participant-table on Thu Jun 18 15:07:03 EDT 2015
 \newlength{\longestpientry}
-\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Hasan Abbasi} (\emph{habbasi@ornl.gov})}}}
-\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Karsten Schwan} (\emph{schwan@cc.gatech.edu})}}}
-\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Allen Malony} (\emph{malony@cs.uoregon.edu})}}}
+\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Scott Klasky} (\emph{klasky@ornl.gov})}}}
+\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Gerald Lofstead} (\emph{glofstead@sandia.gov})}}}
+\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Carlos Maltzahn} (\emph{carlos@uci.edu})}}}
 \begin{tabularx}{\textwidth}{l>{\raggedright\arraybackslash}X}
@@
 \\
-\textbf{\gatech} & \makebox[\longestpientry][l]{\textbf{Karsten Schwan} (\emph{schwan@cc.gatech.edu})}\hspace*{1em}\\
-\textbf{\oregon} & \makebox[\longestpientry][l]{\textbf{Allen Malony} (\emph{malony@cs.uoregon.edu})}\hspace*{1em}\\
-\textbf{\ornl} & \makebox[\longestpientry][l]{\textbf{Hasan Abbasi} (\emph{habbasi@ornl.gov})}\hspace*{1em}\
+\textbf{\ornl} & \makebox[\longestpientry][l]{\textbf{Scott Klasky} (\emph{klasky@ornl.gov})}\hspace*{1em}Hasan Abbasi, Kimmy Mu, Manish Parashar\\
+\textbf{\snl} & \makebox[\longestpientry][l]{\textbf{Gerald Lofstead} (\emph{glofstead@sandia.gov})}\hspace*{1em}Matthew Curry, Lee Ward\\
+\textbf{\uci} & \makebox[\longestpientry][l]{\textbf{Carlos Maltzahn} (\emph{carlos@uci.edu})}\hspace*{1em}\
 \end{tabularx}

### 2015-06-20 narrative/intro.tex
Added people in the personnel.config and institutions.config file. Everyone should check their email/name/spell/affiliations in the file to make sure its correct.
Minor updates

@@
 3. Pluggable infrastructure for data transformation (storage system)\\
-4. Intelligent data mapping to storage/memory hierarchies (middleware and storage system)\\
+4. Intelligent (autonomic, application-aware) data mapping to storage/memory hierarchies (middleware and storage system)\\
 5. Predictable Performance and resolution tradeoffs\\
-6. Data re-generation. (middleware & storage)\\
+6. Data re-generation. (middleware \& storage)\\
 7. Learning Motifs and data re-organization.\\
@@
 
-The metric we are most intrested in optimizing is time to knowledge. Current
+The metric we are most interested in optimizing is time to knowledge. Current
 approaches to addressing the I/O bottleneck fall into two broad
@@
 performance of a single application. Both approaches have seen success but
-are unlikely to overcome the major obstactles in reaching exascale. Instead
+are unlikely to overcome the major obstacles in reaching exascale. Instead
 of trying to optimize throughput, we will seek to reduce the time to
@@
 \end{enumerate}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "../proposal"
+%%% End:

### 2015-06-22 facilities/snl.tex
added snl facilities and lofstead candp

\facilitiessection{Sandia National Laboratories}{snl}

\section*{Sandia National Laboratories Computer Science Research Institute}

The mission of
the Computer Science Research Institute (CSRI) is to conduct collaborative
research and development with external researchers (particularly university
faculty) in the areas of computer science and mathematics to build Sandia’s and
DOE’s modeling and simulation capabilities. The CSRI is funded by the
University Programs element within the Advanced Simulation and Computing (ASC)
program is managed on behalf of Sandia by the Computers, Computation,
Mathematics and Information Center.

The CSRI has established research in the areas of computer and computational
science critical to the laboratories’ and DOE’s mission. They work to identify
external research collaboration that will impact DOE objectives and provide
funds for work through university researchers (faculty and students) working at
Sandia.  The CSRI projects respond to the need for computer science and applied
mathematics research, with a portfolio of exploratory R\&D activities, which
will provide computational-enabling technologies for Sandia code development
and applications.

The CSRI provides a mechanism by which university researchers
learn about problems in computer and computational science at DOE laboratories,
conduct leading-edge research, interact with scientists and engineers at the
laboratories, and help transfer the results of their research to programs at
the laboratories. The CSRI also enables the laboratories to maintain and expand
the computer science and mathematics expertise required for DOE projects and
initiatives to be successful. The CSRI is a focal point, both physically and in
terms of research collaborations, for university researchers, students, and
laboratory staff engaged in computer and computational science, modeling, and
simulation. This project will leverage ongoing CSRI activities to foster
additional collaborations and interactions with academic, industry, and lab
researchers as appropriate.  The CSRI is housed in a 34,500 ft2 facility that
opened in 2006.  It houses approximately 190 researchers comprising 140
full-time staff and 50 visitors and includes conference rooms and collaborative
work areas. The CSRI facility has significant local computing power with high-
bandwidth network connection to large-scale Sandia computing platforms.

The general Sandia computing environment contains a large number of diverse systems from PC’s to workstations, several different Linux cluster systems and supercomputers. This includes the 264 TF Red Sky system (22,528 2.93 GHz Intel Nehalem X5570 cores with InfiniBand) and the 38 TF Glory system (4,352 2.2 GHz AMD Opteron cores with InfiniBand). Sandia CSRI researchers on this project also have access to the 1.374 PF ACES Cielo platform at Los Alamos (Cray XE6 with 143,104 2.4 GHz AMD Opteron cores).
In addition to these production machines, Sandia CSRI system software researchers have access to several experimental architecture technology development testbed platforms, including an Intel MIC architecture testbed (42 nodes connected with QDR InfiniBand, each node contains two six-core 3.46 GHz Intel Westmere processors and two 30-core 1.05 GHz Knights Ferry co-processors), an AMD Fusion testbed (104 nodes connected with QDR InfiniBand, each node contains one Llano Fusion APU consisting of a four-core 2.9 GHz AMD K10 and a 600 MHz 400-core Radeon HD 6550D and a 256-GB Micron MLC flash drive), a Convey testbed (one Convey GV HC-1EX board containing a 2.13 GHz quad-core Intel Nehalem, 4 Xilinx Vertex6 LX760 FPGA co-processors, and 8 Xilinx FGPA programmable memory controllers), and a Tilera streaming mulit-core testbed (four Tile-Gx8036 tiles, each containing 36 1.2 GHz cores).

### 2015-06-22 other-support/lofstead-cp.tex
added snl facilities and lofstead candp

\documentclass[10pt]{article}
\usepackage[english]{babel}
\usepackage{cite}
\usepackage{url}
\usepackage{times}
\usepackage{graphicx}
\usepackage[sectionbib]{bibunits}
\pagestyle{empty}

% Set page geometry
\usepackage{geometry}
\geometry{verbose,letterpaper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

\begin{document}

\begin{center}
{\Large Current and Pending Support for Gerald F. Lofstead}
\end{center}

Dr. Lofstead is a senior member of the technical staff in the Scalable
System Software group at Sandia National Laboratories (SNL). The following
table identifies current and pending support. This funding provides for effort,
travel, and resources. The effort levels will be adjusted if pending proposals
are funded, and as current projects finish.

\section*{Current Support}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \= Scalable I/O Services\\
Source of Support           \>: \> NNSA Advanced Simulation and Computing Program\\
Award Amount                \>: \> \$1000K/year to SNL\\
Period Covered              \>: \> 10/2014 -- 9/2014\\
Investigator Commitment     \>: \> 0.2 FTE
\end{tabbing}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \= ATDM Data Management Services\\
Source of Support           \>: \> NNSA Advanced Simulation and Computing Program\\
Award Amount                \>: \> \$9800K/year to SNL\\
Period Covered              \>: \> 10/2014 -- 9/2018\\
Investigator Commitment     \>: \> 0.25 FTE
\end{tabbing}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \=  HOBBES: OS and Runtime Support For Application Composition\\
Source of Support           \>: \> 2013 DOE Exascale Operating Systems and Runtimes\\
Award Amount                \>: \> \$340K/year to SNL\\
Period Covered              \>: \> 8/2013 -- 7/2016\\
Investigator Commitment     \>: \> 0.35 FTE
\end{tabbing}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \= High-Performance Decoupling of Tightly-Coupled Data Flows\\
Source of Support           \>: \> DOE Scientific Data Management, Analysis and Visualization at Extreme Scale 2\\
Award Amount                \>: \> \$200K/year to SNL\\
Period Covered              \>: \> 8/2014 -- 7/2017\\
Investigator Commitment     \>: \> 0.2 FTE
\end{tabbing}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Pending Proposals (in addition to this)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Pending Proposals (in addition to this)}

None

%\begin{tabbing}
%Project Title \hspace{5.5em}\=: \= Code Coupling for Scalability and Robust Data Analytics\\
%Source of Support           \>: \> Sandia National Laboratories LDRD \\
%Award Amount                \>: \> \$600K/year to SNL\\
%Period Covered              \>: \> 8/2014 -- 7/2017\\
%Investigator Commitment     \>: \> 0.5 FTE
%\end{tabbing}

\end{document}

### 2015-06-22 participant-table.tex
added snl facilities and lofstead candp

@@
 \setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Gerald Lofstead} (\emph{glofstead@sandia.gov})}}}
-\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Carlos Maltzahn} (\emph{carlos@uci.edu})}}}
+\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Carlos Maltzahn} (\emph{carlosm@soe.ucsc.edu})}}}
 \begin{tabularx}{\textwidth}{l>{\raggedright\arraybackslash}X}
@@
 \textbf{\snl} & \makebox[\longestpientry][l]{\textbf{Gerald Lofstead} (\emph{glofstead@sandia.gov})}\hspace*{1em}Matthew Curry, Lee Ward\\
-\textbf{\uci} & \makebox[\longestpientry][l]{\textbf{Carlos Maltzahn} (\emph{carlos@uci.edu})}\hspace*{1em}\
+\textbf{\ucsc} & \makebox[\longestpientry][l]{\textbf{Carlos Maltzahn} (\emph{carlosm@soe.ucsc.edu})}\hspace*{1em}\
 \end{tabularx}

### 2015-06-22 narrative/intro.tex
added some clarifying questions for discussion

@@
 
+We are aiming to spread an output across the vertical layers of the storage
+hierarchy simultaneously.  When data should migrate to a higher tier, what
+happens to the existing version?
+
+Is there a canonical copy that is the originally written version?
+
+Is there annotation about this (I think so) attached to that part of the variable?
+
+At the tape layer, we see having a lossy compressed version just above the tape
+layer used as a directory/index and then the ``full'' resolution version on
+tape only retrieving once the user accepts the time/quality tradeoff. This is
+going to require some serious language to describe. If we have 100s PB of tape
+space, we'll need 1s PB of disk even at a 99\% data reduction. That is
+non-trivial.
+
+What happens when that data is retrieved from tape? Is a third copy made in an
+appropriately sized/quality level version for the tier it is requested to be
+pulled to? What if it is just the index version? Or the on tape version? Does
+it have a TTL because of the pain retrieving it from tape?
+
+Do we specify a target tier or just allow the system to place it in a place
+that makes sense? Given the potential space limitations, I think this is pretty
+critical because it could cause evictions or other insufficient space actions
+that are unintended. I can see us specifying that when you ask for data to be
+pulled at a particular quality level to a particular performance tier that it
+is a best effort with the use specifying if compromise or just failure is the
+result if the request cannot be fulfilled.  Making this interaction make sense
+is going to require some reasonable thought and testing. We'll probably need to
+run it past apps people to get their feedback too as I don't think we are able
+to give a solid answer without some broad input.
+
+How does the migration to ``data lakes''/``campaign storage'' work? How does the
+migration to tape ultimately happen? I think we need to incorporate some
+explicit staging commands to move data both up and down the stack along with
+some variability in placement based on data features and current system state.
+
+Given the storage scarcity, particularly for NVM, I think we need to have a
+solid story here as part of the proposal. Yes, there are questions that have to
+be answered to build this still, but we need to have some pretty solid clarity
+where possible. I don't feel like I can explain it well enough at what I
+believe to be a proper clarity level at this point.
+
 %%% Local Variables:

### 2015-06-25 abstract.tex
writing assignments

@@
 \noindent
-Stuff
+Outline and writing assignments:
+later the abstract will go here.
+\begin{enumerate}
+\item Introduction - Scott - Hasan - Gary - Jay 4 pages
+   \begin{enumerate}
+      \item resource drivers
+      \item application drivers
+      \item overall example which we will use for the test
+      \item  need to answer the questions what we are doing, research  , prototype, aim, and what is new
+   \end{enumerate}
+\item Enabling Technologies - Jay - 1 page
+   \begin{enumerate}
+   	\item Sirocco
+   	\item ADIOS
+   	\item DataSpaces
+   	\item Anything else
+   \end{enumerate}
+ \item {Proposed Research and Methods} - 14 pages
+    \begin{enumerate}
+       \item Application interface - Hasan, Manish, Gary, Kimmy, Jay, Mark
+          \begin{enumerate}
+            \item APIS - Hasan
+            \item description of data - Gary
+            \item collection of data reduction techniques - Mark
+            \item data re-organization specification - Scott, Mark, Kimmy
+            \item QoS - Hasan
+            \item execution policies - Manish
+            \item Initial placement and movement  - Gary
+            \item policies - Jay
+          \end{enumerate}
+      \item Storage System - Jay, Carlos, Manish, Lee, Matthew, Feyi
+      \begin{enumerate}
+         \item migration - Lee
+         \item resource management -Jay
+         \item QoS - Carlos 
+         \item Discovery - Jay
+         \item Eviction metrics - Carlos
+         \item Naming Service - Manish
+         \item Estimation of Times - Feyi
+      \end{enumerate}
+    \end{enumerate}
+\item Evaluation - 1 pages
+\item Timetable of Activities - 1/2 page
+\item Project Management Plan - 1 page
+\item Project Objective  1/2 page
+\end{enumerate}
+
 \vfill

### 2015-06-26 narrative/intro.tex
adding missing files and editing the intro

@@
 scientific processes, demonstrating the validity of our approach in key DOE
-domains.
+domains.  Our techniques will be to research novel techniques into a Scalable 
+Storage Software Infrastructure by integrating services from the middle-ware layer which will
+talk to the applications, with the Storage system. The negotation between these layers will
+be a fundamental service which we will create in this project to ensure that users will be able
+to save `'the best'' amount of data in the different storage tiers.
 

### 2015-06-26 bios/curry.tex
Add bio

\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
% Max 3 pages

\begin{document}
\section{Matthew L. Curry, Ph.D.}

Matthew L. Curry is a Senior Member of Technical Staff at Sandia
National Laboratories. His primary areas of research include
extreme-scale file and storage systems, application of erasure coding
algorithms and other data safety algorithms to storage, heterogeneous
computing, and characteristics of low-level storage devices. He has
given invited lectures at EMC, the University of Alabama at
Birmingham, and Auburn University.

\subsection{Program Committees}
\begin{itemize}
\item The 7th Parallel Data Storage Workshop, held in conjunction with
  Supercomputing 2012.
\item The 28th IEEE Conference on Massive Data Storage (MSST 2012)
\item The 4th International Workshop on Frontiers of Heterogeneous
  Computing (FHC 2013)
\end{itemize}

\subsection{Reviewing Activities}
He performs several reviews a year for the following journals:
\begin{itemize}
\item ACM Transactions on Storage
\item ACM Transactions on Parallel Computing
\item Concurrency and Computation, Practice and Experience
\end{itemize}

\subsection{Education and Training}
\begin{itemize}
\item Ph.D., Computer and Information Sciences\\ University of Alabama at Birmingham, December 2010
\item M.S., Computer and Information Sciences\\ University of Alabama at Birmingham, May 2008
\item B.S., Computer and Information Sciences\\ University of Alabama at Birmingham, December 2004
\end{itemize}

\subsection{Research and Professional Experience}
\begin{itemize}
\item {\bf Senior Member of Technical Staff, Scalable System Software,
  Sandia National Laboratories}, 2010-Present.\\ Performing research
  and development activities associated with high-performance storage
  and operating systems for massively parallel computing systems.
\item {\bf Systems Analyst, Department of Computer and Information
  Sciences, University of Alabama at Birmingham}, January 2009 - May
  2009.\\ Supported research software development activities for NSF
  Award CNS-0821497, ``MRI: Development of a GPU-Enabled Integrated
  Storage Computation Architecture and System.''
\item {\bf Student Intern, Sandia National Laboratories}, Summers
  2005-2008 and May 2009-December 2010.\\ Performed research and
  development activities associated with high-performance storage and
  operating systems for massively parallel computing systems.
\item {\bf Teaching and Research Assistant, Department of Computer and
  Information Sciences, University of Alabama at Birmingham}, during
  spring and fall semesters beginning January 2005 and ending May
  2009.\\ Performed research, taught courses (Operating Systems,
  Introduction to C++), gave guest lectures (Parallel Computing,
  Software Engineering), graded, and conducted labs.
\end{itemize}

\subsection{Publications}

\begin{enumerate}
\item {\sl Gibraltar: A Reed-Solomon Coding Library for Storage Applications
  on Programmable Graphics Processors}.  Matthew L. Curry, Anthony
Skjellum, H. Lee Ward, Ron Brightwell (2011). Concurrency and
Computation, Practice and Experience, Volume 23, Issue 18 (pages
2477-2495).
\item {\sl GPUstore: Harnessing GPU Computing for Storage Systems in
  the OS Kernel}. Weibin Sun, Robert Ricci, Matthew L Curry. The 5th
  Annual International Systems and Storage Conference (SYSTOR 2012).
\item {\sl A Lightweight, GPU-Based Software RAID System}. Matthew
  L. Curry, H. Lee Ward, Anthony Skjellum, Ron
  Brightwell. International Conference on Parallel Processing (ICPP)
  2010.
\item {\sl Accelerating Reed-Solomon coding in RAID systems with
  GPUs}. Matthew L. Curry, Anthony Skjellum, H. Lee Ward, Ron
  Brightwell. IEEE International Parallel and Distributed Processing
  Symposium (IPDPS) 2008.
\item {\sl Power Use of Disk Subsystems in Supercomputers}.  Matthew
  L. Curry, H. Lee Ward, Gary Grider, Jill Gemmill, Jay Harris, David
  Martinez (2011). Proceedings, 6th Parallel Data Storage Workshop,
  Seattle, Washington.
\item {\sl Failing in Place for Low-Serviceability Infrastructure
  Using High-Parity GPU-Based RAID}. Matthew L. Curry, H. Lee Ward,
  Anthony Skjellum, High Performance Embedded Computing Workshop
  (HPEC) 2010.
\item {\sl Using GPUs to Enable Highly Reliable Embedded Storage}.
  Matthew L. Curry, Anthony Skjellum, Lee Ward, Ron Brightwell. HPEC
  2008. September 24, 2008.
\item {\sl Arbitrary Dimension Reed-Solomon Coding and Decoding for
  Extended RAID on GPUs}.  Matthew L. Curry, H. Lee Ward, Anthony
  Skjellum, Ron Brightwell. 3rd Petascale Data Storage Workshop held
  in conjunction with SC08, Austin, TX. Nov. 17, 2008.
\item {\sl Using the Sirocco File System for High-Bandwidth
  Checkpoints}. Matthew L. Curry, Ruth Klundt, H. Lee Ward. Technical
  Report. SAND2012-1087.
\item {\sl A Highly Reliable GPU-based RAID System}. Matthew
  L. Curry. Doctoral thesis, University of Alabama at Birmingham,
  2010, 141 pages. Available online:
  \url{http://www.mhsl.uab.edu/dt/2010p/curry.pdf}.
\end{enumerate}

\subsection{Synergistic Activities}
\begin{itemize}
\item Member of the Board of Directors for the New Mexico Supercomputing Challenge (\url{http://www.supercomputingchallenge.org/about/}.
\item Sandia POC for the DOE Storage Fast-Forward 2 effort.
\end{itemize}
\subsection{Identification of Potential Conflicts of Interest or Bias in Selection of Reviewers}
\begin{itemize}
\item Purushotham V. Bangalore; University of Alabama at Birmingham (collaborator)
\item Ron Brightwell; Sandia National Laboratories (co-author, collaborator)
\item Phil Carns; Argonne National Laboratory (co-author, collaborator)
\item Geoff Danielson; Sandia National Laboratories (collaborator)
\item Jill Gemmill; Clemson University (co-author)
\item Kevin Harms; Argonne National Laboratory (co-author)
\item Jay Harris; Clemson University (co-author)
\item Dries Kimpe; Argonne National Laboratory (co-author, collaborator)
\item Ruth Klundt; Sandia National Laboratories (co-author)
\item Thomas Kroeger; Sandia National Laboratories (collaborator)
\item Darrell Long; University of California-Santa Cruz (collaborator)
\item Carlos Maltzahn; University of California-Santa Cruz (co-author, collaborator)
\item Ethan Miller; University of California-Santa Cruz (collaborator)
\item Ron Oldfield; Sandia National Laboratories (collaborator)
\item David R. Resnick; Sandia National Laboratories (co-author)
\item Robert Ricci; University of Utah (co-author)
\item Robert Ross; Argonne National Laboratory (co-author, collaborator)
\item Haiying Shen; Clemson University (collaborator)
\item Anthony Skjellum; Auburn University (co-author, collaborator, Ph.D. advisor)
\item Weibin Sun; Google (co-author)
\item Zhiwei Sun; Data Direct Networks (co-author, collaborator)
\item Lee Ward; Sandia National Laboratories (co-author, collaborator)
\end{itemize}

\end{document}

### 2015-06-26 narrative/intro.tex
started to add a story and research challenges in the introduction so everyone can get started writing.

@@
 \section*{Introduction}
+
+We will examine the research challenges from the perspective of the (1) user of the system who is attempting to run an exascale simulation,
+(2) the Storage System and I/O layer which needs to negotiate amongst all of the users, and finally (3) the user of the system who is attempting to
+understand the data which was produced from their collection of simulations.
+
+From the perspective of the simulation scientist, they want to be able to have knowledge about how much time they will 
+put into writing their data, and how much storage space they will be able to get during the lifetime of their simulations. Furthermore, they
+would like the ability to get a certain amount of Quality of Service such that they can then make decisions when the expected bandwidth, for example
+is less than what they desire. The user will then place in a certain set of rules which the system can make autonomic decisions to help decide what
+should be done. In our simple scenario, a user wants to perhaps write 1 PB of data every hour for their checkpoint/restart, and they which to write 500TB of data
+every 15 minutes for their analysis and visualization results which have been reduced by using in situ reduction techniques. At this stage, they will ask the system
+to write their data, and they would like to get an estimated time which they can then figure out, through a series of rules, whether they want to write out all of the
+data or if they want to write none of the data but rather wait for a later time to write or write a reduced amount of data, along with some code container to allow them 
+to regenerate the data with a certain level of accuracy. Furthermore, they realize that the analysis data will be saved on either the parallel file system, or the campaign storage,
+a longer term storage area with lower bandwidth than the parallel file system, or if some of the data will just go to tape. Each piece of data the user will specify the
+time they will want to keep the data round. We want the system to give the users a certain amount of currency in terms of bandwidth, storage space on each level, and latency expectations.
+These notions will be fuzzy but they will allow the user to make ad-hoc decisions to figure out what needs to be saved.
+
+Since the data will be a collection of objects placed to the storage system, some of the objects will be broken up into multiple pieces by ``plug-ins'' to the system
+which will allow data to be characterized in terms of the most important pieces, and the next level of details. We can think of this as something similar to an Adaptive Mesh Refinement scheme
+which keeps track of the places where there are perhaps steep gradients (large errors on the coarser views), and smaller errors.  These then allows us to make decisions about which 
+sub-objects to save (all of the data, which will then go to different layers) or just some of the sub-objects?  The research questions are numerous: How does the user specify their intentions? 
+How does the user prioritize the different sub-objects of the data? Where does the data go into the storage hierarchy? Does the data get replicated in the lower tiers of the hierarchy? 
+When the user specifies the time they want the data saved, how does this get into the storage system?
+
+From the perspective of the storage system, it needs to manage the data from a given user request amongst all of the request and try to optimize the entire system. Today the problem is that
+when a user writes there is a tremendous amount of interference from other users on the system who may be writing or reading at the same time. By maiming a certain level of quality of service
+the system must potentially lock out users with lower currency they want to place in the reading or writing of their data.  The storage system must also automatically move data from one tier of the
+storage to another, without effecting the quality of service that it guarantees at the time of questioning. The storage system must also look at ways to organize data amongst all of the tiers of
+storage in one consistent way.
+
+From the perspective of the users who want to read in the data from their simulations, and then operate on this data and possibly write data, we
+see that in general these users will be running on much less compute resources than the users who are running the simulation. One of the
+important aspects of this class of user is that they want a reduction of latency and they require a certain amount of quality of service as well. 
+When someone is doing interactive analysis and expects that the data they read in is about 10s, but it terms into 1,000s, the user typically
+tends to either wait another day for doing their analysis, or gets frustrated and reads much less data than they really need to. Furthermore
+the user tends to read in just a small sample of the data without much knowledge if they are missing important pieces of the data.
+
+Our research into the storage system and middle ware layers must address many of these challenges including
+\begin{enumerate}
+\item  How do we describe the user intentions at the API layer and have this communicated down to the middle-ware and storage layers?
+\item How do we allow users to define their user defined compression techniques and have this data read back from the storage system layers? Where do we execute the decompression techniques
+\item How do we evaluate the tradeoffs at runtime to guide data placement, including the movement of data across the network, across the
+different storage tiers? 
+\item Can we use different forms of learning techniques as daemons on the system to perform data migration from one storage tier to another?
+For example, if we see that a user is looking at one time slice after another for a certain object in their dataset which is in the slower tiers of 
+the hierarchy, do we automatically propane up the data which has not been requested in the hope that this will be requested data?
+\item Can we understand the true need of campaign storage and understand the possible impact of the campaign storage if we run a 
+hadoop-like file-system instead of a lustre/gpfs file system?
+\item Can we limit the amount of data duplication? Data space is going to be very constrained on the exascale system so we must ensure that
+minimal copies are made of data, and when data is duplicated, they are removed through a garbage collection routine.
+\item Can we build a model to give us the time estimates during the reading and writing phases, so that rules can be in place from the user
+perspective to make adaptive decisions. For example, if the file system says that reading will take 3 months, users can they place in rules to
+then read in a subset of data, and they can understand which data can be read quickly and which pieces will take more time.
+\item Can we understand how to place code in the system which will allow data-regeneration to take place.
+
+\end{enumerate}
 

### 2015-06-26 narrative/intro.tex
more questions

@@
 \item Can we understand how to place code in the system which will allow data-regeneration to take place. 
+\item How does the storage system do a better job in managing request from all of the users on a LCF than today? We realize that today
+users who have a better middlware system can often lock other users from getting high performance when they are running. If we have 
+the concept of currency which is eventually used in the same extent as node-hours, then users will have to be able to think about how much
+storage and how much bandwidth they can choose. One question that needs to be understood is if there are times when the system sees
+that there are very few  storage system resources being used, then the lucky users can get the bandwidth cheaper than at times of heavy usage.
+How can we enforce this?
+\item How do we manage all of the metdata not only from the principle objects, but from the sub-objects? How well will this scale when we have
+users who can potentially create billions of objects from their simulation?
 

### 2015-06-26 abstract.tex
Minor update to the outline.
abstract update

@@
           \begin{enumerate}
-            \item APIS - Hasan
+            \item APIS - Kimmy  (rules/policy, user decision of data compression, quick feedback from system monitoring, QoS)
             \item description of data - Gary
             \item collection of data reduction techniques - Mark
-            \item data re-organization specification - Scott, Mark, Kimmy
+            \item data re-organization specification - Scott, Mark
             \item QoS - Hasan
-            \item execution policies - Manish
-            \item Initial placement and movement  - Gary
+            \item data placement and movement (anticipatory and on-demand)  - Gary/Manish
             \item policies - Jay
@@
          \item Eviction metrics - Carlos
-         \item Naming Service - Manish
+         \item Naming Service (not completely sure what this is?) - Manish
          \item Estimation of Times - Feyi

### 2015-06-26 narrative/intro.tex
update small intro section
added overall project structure statement in the intro

--- 2015-06-26 narrative/intro.tex
@@
 
+The Lightweight File Systems project from Sandia sought to strip down a file
+system to the minimum features and encourage layering our auxiliary services to
+add features for a particular use. Sirocco is the storage mechanism for LWFS.
+We are proposing to combine the apps knowledge from ORNL with the SNL system
+and the expertise gaps represented in the additional people to investigate how
+to make a data, workflow, and application aware storage system.
+
 Exascale scientific discovery will be severely bottlenecked without
@@
 scientific processes, demonstrating the validity of our approach in key DOE
-domains.  Our techniques will be to research novel techniques into a Scalable 
-Storage Software Infrastructure by integrating services from the middle-ware layer which will
-talk to the applications, with the Storage system. The negotation between these layers will
-be a fundamental service which we will create in this project to ensure that users will be able
-to save `'the best'' amount of data in the different storage tiers.
+domains.
 

### 2015-06-26 facilities/facilities-rutgers.tex

\begin{center}
\section*{Facilities, Equipment and Other Resources at Rutgers University}
\end{center}

\noindent The proposed research will be conducted at the {\bf Rutgers Discovery Informatics Institute  (RDI$^2$)} (see \url{rdi2.rutgers.edu}). RDI$^2$ was created in 2012 as New Jersey's Center for Advanced Computation, an institute with the overarching goal of establishing a comprehensive and internationally competitive Computational and Data-Enabled Science and Engineering (CDS\&E) effort at Rutgers University that will nurture the fundamental integration of research, education, and infrastructure.

RDI$^2$ also hosts the Cloud and Autonomic Computing Center (CAC), a national research center that was established in January 2008 between Rutgers University, The University of Florida, and The University of Arizona.  The overall goal of the center is to combine resources from these universities, private companies, and federal government to make all kinds of computer systems and applications -- from humble desktop computers to complex air traffic control systems and scientific and engineering applications -- more reliable, more secure, and more efficient.

The CAC center was funded by the Industry/University Cooperative Research Center program of the National Science Foundation, by its members from industry and government, and by university matching funds. CAC members are afforded access to leading edge developments in autonomic computing and to knowledge accumulated by academic researchers and other industry partners. Per NSF guidelines, industry and government contributions in the form of annual CAC memberships, coupled with baseline funds from NSF and university matching funds, directly support the Center's expenses for personnel, equipment, travel, and supplies.

RDI$^2$ provides the Rutgers community with advanced computational resources as well as support, training and consulting services -- details can be found at \url{http://rdi2.rutgers.edu/}.

RDI$^2$ is home to Excalibur, Rutgers' IBM BlueGene/P HPC system installation.  Excalibur consists of two full racks of IBM BlueGene/P.  Each rack has 1024 quad-core processors, with each processor having 2 GB of dedicated RAM (2048 total processors, 8192 total cores, 4 TB total memory).  Excalibur is served by a 228 TB GPFS filesystem, with 1.2 TB of faster SSD storage.  Network interconnect to the storage subsystem is through multiple 10 Gigabit Ethernet links.  Excalibur is a University-wide resource.  All researchers and students at Rutgers can get access to Excalibur using a startup allocation; larger allocations are available via a proposal process. Jobs are scheduled round-the-clock so as to maximize utilization of the system.

Other resources available at RDI$^2$/CAC include a 32 node Dell M610 blade cluster, consisting of two Dell M1000E Modular Blade Enclosures, necessary interconnect/management infrastructure, and a supervisory node. Each enclosure is maximally configured with sixteen blades, each blade having two Intel Xeon E5504 Nehalem family quad-core processors at 2.0 GHz, forming an eight core node. Each node has 24 GB RAM and 73 GB of local disk storage (10,000 RPM), twelve nodes have an additional 1 TB of local storage. The network infrastructure is comprised of an integrated 16-port Mellanox InfiniBand switch within each blade chassis, each switch linked to the switch in the other chassis. All blades have Mellanox Quad-Data-Rate (QDR) InfiniBand interface cards. There is also an integrated (redundant) 1 Gigabit Ethernet within each chassis, with two pairs of 10 Gigabit uplink capabilities in each chassis. In the aggregate, the cluster system consists of 32 nodes, 256 cores, 768 GB memory and ~14.5 TB disk capacity, with a 20 Gigabit InfiniBand network and two 1 Gigabit Ethernet networks.

RDI$^2$ has recently deployed a NSF funded research instrument for ``Computational And data-enabled Platform for Energy efficiency Research (CAPER)''. 
This is an eight-node cluster based on Super Micro Computer's SYS-4027GR-TRT system, which is capable of housing concurrently, in one node: (i) eight general-purpose graphical processing units (GPGPU), or eight Intel many-integrated-core (MIC) coprocessors -- or any eight-card combination of the two;(ii) two Fusion-io IoDrive-2 non-volatile random access memory (NVRAM) storage cards, or two LSI 24-port disk controllers -- or any two-card combination of the two; (iii) a single-port fourteen-data-rate (FDR) InfiniBand networking card; and (iv) up to 48 hard disk drives (HDD), or solid-state drives (SSD). The initial configuration will tentatively feature servers two Intel Xeon Ivybridge E5-2650v2 (16 cores/32 hardware threads), 128GB of DRAM, 1TB of Flash-based NVRAM, 2TB of SSD and 4TB of hard disk, one Intel Xeon Phi 7120P, one NVIDIA K40, and Infiniband FDR and 10G Ethernet network connectivity. This hardware is unprecedented in its flexibility and adaptability.  The platform will be instrumented with both coarse- and fine-grained power metering. The focus of the research in this proposal is a perfect fit to this hardware platform, and vice-versa.

RDI$^2$ has established a distributed computational infrastructure composed of clusters at each site that are linked together. Current facilities include multiple clusters, multiple compute servers and GPU resources. These systems are housed in a dedicated machine room that has been recently upgraded to increase cooling capabilities as well as UPS capabilities.

The state-of-the-art RDI$^2$ machine room features proper refrigeration and necessary security. In order to optimize cooling, the machine room follows a variant of the (accepted best practice) host aisle/cold aisle cabinet layout. The cold air is pushed from ceiling to the front of the racks. The exhaust air is routed directly into computer room air conditioner (CRAC) units.

RDI$^2$ with support from the State of New Jersey (see  \url{http://nj.gov/governor/news/news/552013/approved/20130429i.html}) is working on establishing a balanced advanced research computational and data environment composed of large-scale high-end compute engines and clusters, as well as significant co-located storage with embedded analytics capabilities. The platform will provide researchers, students, industry and government across the state with access to ACI capabilities for research and instruction.

This project will be deployed in three phases. The first phase provides an Inbiniband-based high-performance computing cluster with the following configuration: 120 baseline compute nodes, 18 medium memory, 2 large memory nodes, 24 single NVidia GPU based nodes, 2 multi NVidia GPU nodes, 2 Intel Xeon Phi nodes, along with 2 head nodes, 3 login nodes and 2 service nodes. All the compute is supported with an Infiniband FDR compute fabric (2:1 oversubscribed) a Gigabit Ethernet management fabric, a 10 Gigabit network for connection to public networks. The baseline and medium memory nodes are built leveraging the Dell M1000e blade environment high density, integrated switching and ease of deployment. The nodes themselves are M630 half height blades with dual Intel Xeon E5-2680v3 2.5GHz, 12-Core, 120W processors, 256G of 2133 MHz RAM (10.7GB RAM per core), single 300G 10K hard drive, dual 10 Gigabit Ethernet ports and a Mellanox FDR Infiniband adapter. The rest of nodes are in a rack-based configuration with the same basic configuration and specialized capabilities (i.e., NVidia GPU, Intel Xeon Phi or large memory size). The first phase also includes a scalable high-performance parallel file system based on GPFS (General Parallel File System), which provides 1PB of usable storage capacity.

The second phase is focused on a self-contained, flexible modular data center, connected with high-speed networking. The modular data center will be able to house up to 17 racks and will provide at least 350KW of electrical power for computer systems.

The third phase will provide a supercomputing system, which will deliver a peak performance of hundreds of Teraflops, making this system one of the top academic computing systems in the nation. This system will be featured with tens of thousands of CPU cores, a low-latency high-bandwidth network and innovative non-volatile memory solutions.

### 2015-06-26 narrative/intro.tex

--- 2015-06-26 narrative/intro.tex
@@
 \section*{Introduction}
+
+The Lightweight File Systems project from Sandia sought to strip down a file
+system to the minimum features and encourage layering our auxiliary services to
+add features for a particular use. Sirocco is the storage mechanism for LWFS.
+We are proposing to combine the apps knowledge from ORNL with the SNL system
+and the expertise gaps represented in the additional people to investigate how
+to make a data, workflow, and application aware storage system.
 

### 2015-06-26 narrative/enabling_tech.tex
added enabling tech section

\section*{Enabling Technologies}

\begin{itemize}

\item Sirocco
\item ADIOS
\item DataSpaces
\item Anything Else

\end{itemize}

### 2015-06-26 narrative/intro.tex

@@
   use and inherent data qualities (e.g., contains desireable features)
-\item describe idea of using heavy compression for less interesting areas and less or no compression for more interesting areas.
+\item describe idea of using heavy compression for less interesting areas
+  and less or no compression for more interesting areas.
 \item describe inherent errors in scientific calculations
 \item describe lossy compression within error bounds
-\item describe how incorporating plug-ins for application aware data compression opportunities to aid data storage
-\item describe Sirocco's current state and future designs as a solid foundation on which to work
-\item describe how these new features can enhance a system like Sirocco to reduce data intensity
+\item describe how incorporating plug-ins for application aware data
+  compression opportunities to aid data storage
+\item describe Sirocco's current state and future designs as a solid
+  foundation on which to work
+\item describe how these new features can enhance a system like Sirocco to
+  reduce data intensity
 \item describe metadata challenges that we will also have to face.
@@
 
-Problem and background, one or more solution approaches, and investigation goals/questions.
+Problem and background, one or more solution approaches, and investigation
+goals/questions.
 
@@
 
-We will examine the research challenges from the perspective of the (1) user of the system who is attempting to run an exascale simulation,
-(2) the Storage System and I/O layer which needs to negotiate amongst all of the users, and finally (3) the user of the system who is attempting to
-understand the data which was produced from their collection of simulations.
+We will examine the research challenges from the perspective of the
+%
+(1) user of the system who is attempting to run an exascale simulation,
+%
+(2) the Storage System and I/O layer which needs to negotiate amongst all of
+the users, and finally
+%
+(3) the user of the system who is attempting to understand the data which
+was produced from their collection of simulations.
 
-From the perspective of the simulation scientist, they want to be able to have knowledge about how much time they will 
-put into writing their data, and how much storage space they will be able to get during the lifetime of their simulations. Furthermore, they
-would like the ability to get a certain amount of Quality of Service such that they can then make decisions when the expected bandwidth, for example
-is less than what they desire. The user will then place in a certain set of rules which the system can make autonomic decisions to help decide what
-should be done. In our simple scenario, a user wants to perhaps write 1 PB of data every hour for their checkpoint/restart, and they which to write 500TB of data
-every 15 minutes for their analysis and visualization results which have been reduced by using in situ reduction techniques. At this stage, they will ask the system
-to write their data, and they would like to get an estimated time which they can then figure out, through a series of rules, whether they want to write out all of the
-data or if they want to write none of the data but rather wait for a later time to write or write a reduced amount of data, along with some code container to allow them 
-to regenerate the data with a certain level of accuracy. Furthermore, they realize that the analysis data will be saved on either the parallel file system, or the campaign storage,
-a longer term storage area with lower bandwidth than the parallel file system, or if some of the data will just go to tape. Each piece of data the user will specify the
-time they will want to keep the data round. We want the system to give the users a certain amount of currency in terms of bandwidth, storage space on each level, and latency expectations.
-These notions will be fuzzy but they will allow the user to make ad-hoc decisions to figure out what needs to be saved.
+From the perspective of the simulation scientist, they want to be able to
+have knowledge about how much time they will put into writing their data,
+and how much storage space they will be able to get during the lifetime of
+their simulations. Furthermore, they would like the ability to get a certain
+amount of Quality of Service such that they can then make decisions when the
+expected bandwidth, for example is less than what they desire. The user will
+then place in a certain set of rules which the system can make autonomic
+decisions to help decide what should be done. In our simple scenario, a user
+wants to perhaps write 1 PB of data every hour for their checkpoint/restart,
+and they which to write 500TB of data every 15 minutes for their analysis
+and visualization results which have been reduced by using in situ reduction
+techniques. At this stage, they will ask the system to write their data, and
+they would like to get an estimated time which they can then figure out,
+through a series of rules, whether they want to write out all of the data or
+if they want to write none of the data but rather wait for a later time to
+write or write a reduced amount of data, along with some code container to
+allow them to regenerate the data with a certain level of accuracy.
+Furthermore, they realize that the analysis data will be saved on either the
+parallel file system, or the campaign storage, a longer term storage area
+with lower bandwidth than the parallel file system, or if some of the data
+will just go to tape. Each piece of data the user will specify the time they
+will want to keep the data round. We want the system to give the users a
+certain amount of currency in terms of bandwidth, storage space on each
+level, and latency expectations. These notions will be fuzzy but they will
+allow the user to make ad-hoc decisions to figure out what needs to be
+saved.
 
-Since the data will be a collection of objects placed to the storage system, some of the objects will be broken up into multiple pieces by ``plug-ins'' to the system
-which will allow data to be characterized in terms of the most important pieces, and the next level of details. We can think of this as something similar to an Adaptive Mesh Refinement scheme
-which keeps track of the places where there are perhaps steep gradients (large errors on the coarser views), and smaller errors.  These then allows us to make decisions about which 
-sub-objects to save (all of the data, which will then go to different layers) or just some of the sub-objects?  The research questions are numerous: How does the user specify their intentions? 
-How does the user prioritize the different sub-objects of the data? Where does the data go into the storage hierarchy? Does the data get replicated in the lower tiers of the hierarchy? 
-When the user specifies the time they want the data saved, how does this get into the storage system?
+Since the data will be a collection of objects placed to the storage system,
+some of the objects will be broken up into multiple pieces by ``plug-ins''
+to the system which will allow data to be characterized in terms of the most
+important pieces, and the next level of details. We can think of this as
+something similar to an Adaptive Mesh Refinement scheme which keeps track of
+the places where there are perhaps steep gradients (large errors on the
+coarser views), and smaller errors. These then allows us to make decisions
+about which sub-objects to save (all of the data, which will then go to
+different layers) or just some of the sub-objects? The research questions
+are numerous: How does the user specify their intentions? How does the user
+prioritize the different sub-objects of the data? Where does the data go
+into the storage hierarchy? Does the data get replicated in the lower tiers
+of the hierarchy? When the user specifies the time they want the data saved,
+how does this get into the storage system?
 
-From the perspective of the storage system, it needs to manage the data from a given user request amongst all of the request and try to optimize the entire system. Today the problem is that
-when a user writes there is a tremendous amount of interference from other users on the system who may be writing or reading at the same time. By maiming a certain level of quality of service
-the system must potentially lock out users with lower currency they want to place in the reading or writing of their data.  The storage system must also automatically move data from one tier of the
-storage to another, without effecting the quality of service that it guarantees at the time of questioning. The storage system must also look at ways to organize data amongst all of the tiers of
-storage in one consistent way.
+From the perspective of the storage system, it needs to manage the data from
+a given user request amongst all of the request and try to optimize the
+entire system. Today the problem is that when a user writes there is a
+tremendous amount of interference from other users on the system who may be
+writing or reading at the same time. By maiming a certain level of quality
+of service the system must potentially lock out users with lower currency
+they want to place in the reading or writing of their data. The storage
+system must also automatically move data from one tier of the storage to
+another, without effecting the quality of service that it guarantees at the
+time of questioning. The storage system must also look at ways to organize
+data amongst all of the tiers of storage in one consistent way.
 
-From the perspective of the users who want to read in the data from their simulations, and then operate on this data and possibly write data, we
-see that in general these users will be running on much less compute resources than the users who are running the simulation. One of the
-important aspects of this class of user is that they want a reduction of latency and they require a certain amount of quality of service as well. 
-When someone is doing interactive analysis and expects that the data they read in is about 10s, but it terms into 1,000s, the user typically
-tends to either wait another day for doing their analysis, or gets frustrated and reads much less data than they really need to. Furthermore
-the user tends to read in just a small sample of the data without much knowledge if they are missing important pieces of the data.
+From the perspective of the users who want to read in the data from their
+simulations, and then operate on this data and possibly write data, we see
+that in general these users will be running on much less compute resources
+than the users who are running the simulation. One of the important aspects
+of this class of user is that they want a reduction of latency and they
+require a certain amount of quality of service as well. When someone is
+doing interactive analysis and expects that the data they read in is about
+10s, but it terms into 1,000s, the user typically tends to either wait
+another day for doing their analysis, or gets frustrated and reads much less
+data than they really need to. Furthermore the user tends to read in just a
+small sample of the data without much knowledge if they are missing
+important pieces of the data.
 
-Our research into the storage system and middle ware layers must address many of these challenges including
+Our research into the storage system and middle ware layers must address
+many of these challenges including
 \begin{enumerate}
-\item  How do we describe the user intentions at the API layer and have this communicated down to the middle-ware and storage layers?
-\item How do we allow users to define their user defined compression techniques and have this data read back from the storage system layers? Where do we execute the decompression techniques
-\item How do we evaluate the tradeoffs at runtime to guide data placement, including the movement of data across the network, across the
-different storage tiers? 
-\item Can we use different forms of learning techniques as daemons on the system to perform data migration from one storage tier to another?
-For example, if we see that a user is looking at one time slice after another for a certain object in their dataset which is in the slower tiers of 
-the hierarchy, do we automatically propane up the data which has not been requested in the hope that this will be requested data?
-\item Can we understand the true need of campaign storage and understand the possible impact of the campaign storage if we run a 
-hadoop-like file-system instead of a lustre/gpfs file system?
-\item Can we limit the amount of data duplication? Data space is going to be very constrained on the exascale system so we must ensure that
-minimal copies are made of data, and when data is duplicated, they are removed through a garbage collection routine.
-\item Can we build a model to give us the time estimates during the reading and writing phases, so that rules can be in place from the user
-perspective to make adaptive decisions. For example, if the file system says that reading will take 3 months, users can they place in rules to
-then read in a subset of data, and they can understand which data can be read quickly and which pieces will take more time.
-\item Can we understand how to place code in the system which will allow data-regeneration to take place. 
-\item How does the storage system do a better job in managing request from all of the users on a LCF than today? We realize that today
-users who have a better middlware system can often lock other users from getting high performance when they are running. If we have 
-the concept of currency which is eventually used in the same extent as node-hours, then users will have to be able to think about how much
-storage and how much bandwidth they can choose. One question that needs to be understood is if there are times when the system sees
-that there are very few  storage system resources being used, then the lucky users can get the bandwidth cheaper than at times of heavy usage.
-How can we enforce this?
-\item How do we manage all of the metdata not only from the principle objects, but from the sub-objects? How well will this scale when we have
-users who can potentially create billions of objects from their simulation?
-
+\item How do we describe the user intentions at the API layer and have this
+  communicated down to the middle-ware and storage layers?
+\item How do we allow users to define their user defined compression
+  techniques and have this data read back from the storage system layers?
+  Where do we execute the decompression techniques 
+\item How do we evaluate the tradeoffs at runtime to guide data placement,
+  including the movement of data across the network, across the different
+  storage tiers? (data movement and quality tradeoff)
+\item Can we use different forms of learning techniques as daemons on the
+  system to perform data migration from one storage tier to another? For
+  example, if we see that a user is looking at one time slice after another
+  for a certain object in their dataset which is in the slower tiers of the
+  hierarchy, do we automatically propane up the data which has not been
+  requested in the hope that this will be requested data? (prediction and
+  prefetching, user defined compression on movement)
+\item Can we understand the true need of campaign storage and understand the
+  possible impact of the campaign storage if we run a hadoop-like
+  file-system instead of a lustre/gpfs file system? (additional layers)
+\item Can we limit the amount of data duplication? Data space is going to be
+  very constrained on the exascale system so we must ensure that minimal
+  copies are made of data, and when data is duplicated, they are removed
+  through a garbage collection routine. (space management)
+\item Can we build a model to give us the time estimates during the reading
+  and writing phases, so that rules can be in place from the user
+  perspective to make adaptive decisions. For example, if the file system
+  says that reading will take 3 months, users can they place in rules to
+  then read in a subset of data, and they can understand which data can be
+  read quickly and which pieces will take more time. (estimation)
+\item Can we understand how to place code in the system which will allow
+  data-regeneration to take place. (regeneration)
+\item How does the storage system do a better job in managing request from
+  all of the users on a LCF than today? We realize that today users who have
+  a better middlware system can often lock other users from getting high
+  performance when they are running. If we have the concept of currency
+  which is eventually used in the same extent as node-hours, then users will
+  have to be able to think about how much storage and how much bandwidth
+  they can choose. One question that needs to be understood is if there are
+  times when the system sees that there are very few storage system
+  resources being used, then the lucky users can get the bandwidth cheaper
+  than at times of heavy usage. How can we enforce this? (Fairness)
+\item How do we manage all of the metdata not only from the principle
+  objects, but from the sub-objects? How well will this scale when we have
+  users who can potentially create billions of objects from their
+  simulation? (scalability and discovery)
 \end{enumerate}
@@
 %
-In this project we will demonstrate novel techniques to
-facilitate efficient mapping of data objects, even partitioning individual
-variables, from the user space onto multiple storage tiers, and enable
-application-guided data reductions/transformations to address capacity and
-bandwidth bottlenecks, while constraining the error to be within user
-provided bounds. 
+In this project we will demonstrate novel techniques to facilitate efficient
+mapping of data objects, even partitioning individual variables, from the
+user space onto multiple storage tiers, and enable application-guided data
+reductions/transformations to address capacity and bandwidth bottlenecks,
+while constraining the error to be within user provided bounds.
 %
@@
 domains. Our techniques will be to research novel techniques into a Scalable
-Storage Software Infrastructure by integrating services from the middle-ware layer which will
-talk to the applications, with the Storage system. The negotation between these layers will
-be a fundamental service which we will create in this project to ensure that users will be able
-to save `'the best'' amount of data in the different storage tiers.
+Storage Software Infrastructure by integrating services from the middle-ware
+layer which will talk to the applications, with the Storage system. The
+negotation between these layers will be a fundamental service which we will
+create in this project to ensure that users will be able to save `'the
+best'' amount of data in the different storage tiers.
 
-The metric we are most interested in optimizing is time to knowledge. Current
-approaches to addressing the I/O bottleneck fall into two broad
+The metric we are most interested in optimizing is time to knowledge.
+Current approaches to addressing the I/O bottleneck fall into two broad
 categories. Parallel file system approaches that optimize the throughput for
@@
 \item Evaluating the tradeoffs at runtime to guide data placement, movement,
-  and migration across storage layers using models, heuristics and continuous
-  learning;
+  and migration across storage layers using models, heuristics and
+  continuous learning;
 \item Utilize the additional knowledge available about the data to perform
@@
 
-Is there annotation about this (I think so) attached to that part of the variable?
+Is there annotation about this (I think so) attached to that part of the
+variable?
 
-At the tape layer, we see having a lossy compressed version just above the tape
-layer used as a directory/index and then the ``full'' resolution version on
-tape only retrieving once the user accepts the time/quality tradeoff. This is
-going to require some serious language to describe. If we have 100s PB of tape
-space, we'll need 1s PB of disk even at a 99\% data reduction. That is
-non-trivial.
+At the tape layer, we see having a lossy compressed version just above the
+tape layer used as a directory/index and then the ``full'' resolution
+version on tape only retrieving once the user accepts the time/quality
+tradeoff. This is going to require some serious language to describe. If we
+have 100s PB of tape space, we'll need 1s PB of disk even at a 99\% data
+reduction. That is non-trivial.
 
-What happens when that data is retrieved from tape? Is a third copy made in an
-appropriately sized/quality level version for the tier it is requested to be
-pulled to? What if it is just the index version? Or the on tape version? Does
-it have a TTL because of the pain retrieving it from tape?
+What happens when that data is retrieved from tape? Is a third copy made in
+an appropriately sized/quality level version for the tier it is requested to
+be pulled to? What if it is just the index version? Or the on tape version?
+Does it have a TTL because of the pain retrieving it from tape?
 
 Do we specify a target tier or just allow the system to place it in a place
-that makes sense? Given the potential space limitations, I think this is pretty
-critical because it could cause evictions or other insufficient space actions
-that are unintended. I can see us specifying that when you ask for data to be
-pulled at a particular quality level to a particular performance tier that it
-is a best effort with the use specifying if compromise or just failure is the
-result if the request cannot be fulfilled.  Making this interaction make sense
-is going to require some reasonable thought and testing. We'll probably need to
-run it past apps people to get their feedback too as I don't think we are able
-to give a solid answer without some broad input.
+that makes sense? Given the potential space limitations, I think this is
+pretty critical because it could cause evictions or other insufficient space
+actions that are unintended. I can see us specifying that when you ask for
+data to be pulled at a particular quality level to a particular performance
+tier that it is a best effort with the use specifying if compromise or just
+failure is the result if the request cannot be fulfilled. Making this
+interaction make sense is going to require some reasonable thought and
+testing. We'll probably need to run it past apps people to get their
+feedback too as I don't think we are able to give a solid answer without
+some broad input.
 
-How does the migration to ``data lakes''/``campaign storage'' work? How does the
-migration to tape ultimately happen? I think we need to incorporate some
+How does the migration to ``data lakes''/``campaign storage'' work? How does
+the migration to tape ultimately happen? I think we need to incorporate some
 explicit staging commands to move data both up and down the stack along with
-some variability in placement based on data features and current system state.
+some variability in placement based on data features and current system
+state.
 
 Given the storage scarcity, particularly for NVM, I think we need to have a
-solid story here as part of the proposal. Yes, there are questions that have to
-be answered to build this still, but we need to have some pretty solid clarity
-where possible. I don't feel like I can explain it well enough at what I
-believe to be a proper clarity level at this point.
+solid story here as part of the proposal. Yes, there are questions that have
+to be answered to build this still, but we need to have some pretty solid
+clarity where possible. I don't feel like I can explain it well enough at
+what I believe to be a proper clarity level at this point.
 

### 2015-06-26 narrative/intro.tex
added more details to the opening

--- 2015-06-26 narrative/intro.tex
@@
 
-The Lightweight File Systems project from Sandia sought to strip down a file
-system to the minimum features and encourage layering our auxiliary services to
-add features for a particular use. Sirocco is the storage mechanism for LWFS.
-We are proposing to combine the apps knowledge from ORNL with the SNL system
-and the expertise gaps represented in the additional people to investigate how
-to make a data, workflow, and application aware storage system.
+The Lightweight File Systems~\cite{lwfs} (LWFS) project from Sandia sought to
+demonstrate stripping down a file system to the bare minimum features required
+for any storage system. Then, additional functionality and semantics, such as
+POSIX-style directories and consistency requirements, are added in using
+auxiliary services separate from the stack or by layering providing a new
+interface to the storage system. The project demonstrated that using a core
+object storage system with integrated security providing authentication and
+autorization services, customized storage interfaces could be developed with
+exactly the overhead required for different applications. The second phase of
+LWFS, known as Sirocco~\cite{sirocco}, is investigating how to manage storage
+devices for efficient and scalable storage performance.  Sirocco has been in
+process for three years and has a usable storage infrastructure system. This
+proposal seeks to build upon this work to demonstrate an application, data, and
+usage aware storage system optimized for exascale composite applications and
+capability application runs. We are proposing to combine the applications
+knowledge from ORNL with the SNL LWFS system and storage system knowledge, the
+metadata and storage systems knowledge of UCSC, and the middleware expertise of
+Rutgers to investigate how to make a data, workflow, and application aware
+storage system.
 

### 2015-06-26 narrative/data-reorg.tex

\subsubsection{Data Reorganization Techniques}

Mark, Scott

### 2015-06-26 abstract.tex
new outline changes

@@
  \item {Proposed Research and Methods} - 14 pages
+          Proposed work
+          Execution path
+          Related work
+
     \begin{enumerate}
@@
             \item APIS - Kimmy  (rules/policy, user decision of data compression, quick feedback from system monitoring, QoS)
-            \item description of data - Gary
+            \item description of data - Gary `` how we describe the structure and semantic of the data, how will it be stored, application level importance, utility curve''
             \item collection of data reduction techniques - Mark
             \item data re-organization specification - Scott, Mark
-            \item QoS - Hasan
-            \item data placement and movement (anticipatory and on-demand)  - Gary/Manish
+            \item QoS - Hasan/Gary/Carlos - emission control, some guarantees on the estimates
+            \item data placement and movement (anticipatory and on-demand)  - Manish, Gary 
             \item policies - Jay
@@
          \item migration - Lee
+         \item Match organization of data with the medium characteristics - Carlos, Hasan
+         \item Eviction metrics - Carlos, Hasan
          \item resource management -Jay
          \item QoS - Carlos 
-         \item Discovery - Jay
-         \item Eviction metrics - Carlos
-         \item Naming Service (not completely sure what this is?) - Manish
+         \item Discovery - Jay - moves things based on pressure on the system, getting to some things take effort because it is at different places, understand what our different options in terms of data quality...
+ability to decide the trade-offs
+         \item Naming Service (not completely sure what this is?) - Jay
+          \item data placement and movement (anticipatory and on-demand) from a storage system perspective - Manish 
          \item Estimation of Times - Feyi
@@
     \end{enumerate}
-\item Evaluation - 1 pages
+\item Evaluation - 1 pages - Scott
 \item Timetable of Activities - 1/2 page
-\item Project Management Plan - 1 page
+\item Project Management Plan - 1 page - Scott
 \item Project Objective  1/2 page

### 2015-06-26 narrative/placement.tex

\subsubsection{Data Placement and Movement}

Gary / Manish

### 2015-06-27 narrative/enabling_tech.tex
Added enabling tech and policies information. Policies needs some additional thought. References need to be fixed for enabling tech still.

@@
 
-\begin{itemize}
+Significant existing work will serve as the foundation upon which this project
+will be performed.
 
-\item Sirocco
-\item ADIOS
-\item DataSpaces
-\item Anything Else
+At the lowest level is the Lightweight File Systems (LWFS) project~\cite{lwfs}.
+As mentioned above, LWFS's goal is to provide the simplest core services for
+a storage system and investigate adding functionality by using auxiliary
+services or layering an API on top that offers more functionality. For example,
+by eliminating consistency checks from the core system, applications already
+managing consistency can eliminate the redundant check enhancing performance.
 
-\end{itemize}
+The current LWFS project phase is investigating new ways for managing storage
+devices. This project, Sirocco, offers a peer-to-peer style data storage device
+mesh wtih data migrating based on use, required resilience, and available
+resources. For example, when writing data initially into Sirocco, required
+resilience characteristics are provided prompting Sirocco to ensure the data is
+stored on appropriate storage device(s) that meet the resilience requirement.
+When a particular device is full, Sirocco determines if it should migrate the
+data to a new location or if it is a copy beyond the resilience requirement. If
+this copy must still be protected, Sirocco will find a place of sufficient 
+resilience to move the data to before deleting the local copy. Further, when
+data is requested, a client can ask any Sirocco server and Sirocco will find
+the data returning it to the client. If desired, the canonical data location
+can be moved on client request to offer better performance for the client. In
+additional features, a stored object can have multiple data forks that can
+store different versions or aspects of the same data.  For this project, we
+will build on this Sirocco foundation for data placement, migration, and the
+version storage (via the fork mechanism).
+
+The Sirocco API is much more complex than a standard POSIX call prompting
+layering a more user friendly interface on top. The ADIOS~\cite{adios} API
+offers an API nearly as simple as POSIX, but with the ability to change the
+actual data transport mechanism without changing the source code. This affords
+writing to a single output API while shifting from writing to a POSIX file
+system, some middleware like DataSpaces~\cite{dataspaces}, or some other system
+like Sirocco all without changing the source code. ADIOS has proven effective
+for writing to POSIX file systems, data staging
+systems~\cite{predata,dataspaces,nessie}, workflow systems~\cite{flexpath}, and
+even nothing through the use of a NULL transport.
+
+At a similar level as ADIOS, but can be used in conjunction, are tools like
+DataSpaces~\cite{dataspaces}. They offer a way to request data set portions
+from a one component by the next componet in a data flow. In DataSpaces case,
+the data transfer does not have to correspond to the data from a single
+process.  Instead, DataSpaces will determine where the data is and issue
+requests to pull the data from the source to the sink in an on demand process.
+
+For this project, we plan to use the Sirocco data storage mechanism to
+offer managing the actual data storage on all of the different tiers. We'll add
+a data management layer as part of the file system to handle data discovery and
+metadata tasks. As a middleware layer, we will investigate offering differing
+data compression, fidelity, and storage placement options. Through this 
+combination, we will be able to deliver a complete end-to-end system.

### 2015-06-27 narrative/policies.tex
Added enabling tech and policies information. Policies needs some additional thought. References need to be fixed for enabling tech still.

\subsubsection{Policies}

One of the key differentiating features this project offers is driving storage
decisions through policies rather than continuing the current model of allowing
all application to compete, without restriction, for part or all of the storage
resources. This current approach leads to intereference
effects~\cite{lofstead:2010:adaptive} that can greatly impact IO performance
predictability. Further, with the introduction of additional tiers in the
storage stack, the performance variability will increase with greater
competiion for limited high performance resources. Unfettered, this competition
will yield lower performnce for all users than if users either restrict IO
resource uses to a disjoint subset each or even using lower raw performance
resources.

Our policies are driven from multiple different approaches. First, supporting
capability runs is a priority for capability machines. We propose to address
capability run performance by incorporating monitoring for prepartory runs at
smaller scale to profile the application output characteristics both from a
writing velocity and volume, but also for example read patterns for the data
analytics required to generate scientific insights from the raw data. By
discovering approximate data proportions that will likely be the key subsets
targeted by the simulation run, we can generate a policy to preserve a certain
storage portion for high fidelity data storagae and use slower or perhaps lower
fidelity or compressed data storage for less interesting data portions. We will
support spreading data appropriately across the storage hierarchy. In
particular, we will investigate how to place data across different layers to
meet the performance requirements for output while maintaining system
availability for other applications and offering the best possible performance
for the data analytics that will ultimately process this data.

Second, once data has been placed for future performance requirements, policies
must be able to address maintaining data on a priority basis in that location
to deliver on the placement optimization generated by the capability run
profilier.

Third, most future NVM devices have limited write endurance prompting
management to ensure fair use by all machine users. Technologies like
NAND-flash and Phase Change Memory have limited write endurance. By
incorporating policies about the proportion of write endurance a compute
allocation is entitled to, data placement decisions can be made to ensure fair
resource usage. While the offending application will suffer worse IO
performance, other applications that more carefully address their data
intensity on these limited devices will achieve higher overall performnce. Only
by instituting such policies can we encourage application users to adopt
technology to manage the limited resources.

We propose to investigate both offering a policy mechanism and how to implement
these kinds of policies in a way that addresses offering the shortest
end-to-end time for scientific insights.

### 2015-06-28 abstract.tex
making changes...

--- 2015-06-25 abstract.tex
@@
 \enlargethispage{2\baselineskip}
+Exascale scientific discovery will be
+severely bottlenecked without sufficient new research into managing and
+storing the large amounts of data that will be produced during the
+simulation, and analyzed for months afterwards. In this project we will
+demonstrate novel techniques to  facilitate efficient mapping of data
+objects, even partitioning individual variables, from
+the user space onto multiple storage tiers, and enable application-guided
+data reductions/transformations to address capacity and bandwidth
+bottlenecks. 
+Our goal in this project is
+to address the associated I/O and storage challenges in the context of
+current and emerging storage landscapes, and expedite insights into mission
+critical scientific processes; which is associated with {\bf theme two} of the FOA. To that end, we will build on the
+capabilities offered by ADIOS and DataSpaces at ORNL that provide I/O abstractions
+and services, the Sirocco peer-to-peer file system at
+Sandia and the object storage and annotation expertise of UC Santa Cruz, to
+explore application and multi-tier storage aware data management
+solutions.    This project
+brings together a  team with strong expertise in I/O middleware (ORNL, Rutgers), file system (SNL, UCSC) and storage (UCSC), and connects and
+coordinates these key storage components in a seamless fashion.
 
+Our objective here is to reduce the time to knowledge, not just for a single
+application, but for the entire workload in a multi-user environment, where
+the storage is shared among users.
+We achieve this goal by allowing selectable data quality, by trading its accuracy and error
+in order to meet the time or resource constraint. 
+We will explore beyond 
+checkpoint/restart I/O, and will address the challenges posed by 
+key
+data access patterns in the knowledge gathering process.
+Ultimately, we will take the knowledge from the storage system to provide vital feedback to the middleware 
+so that the best possible decisions can be autonomically made between the user intentions and
+the available system resources.  
+We will test our prototypes  on current and future DOE system with many of today's applications, including the
+s XGC1, GTC, QMCPack, and SpecFM3D simulations. 
+Our solutions will provide  new functionalities and APIs for
+1) specifying, at the application level, data annotations that enable the
+quantification of the relative importance and utility of data objects and
+enable partitioning  across the storage hierarchy;
+2) specifying selectable performance/quality/cost tradeoffs from both the
+application and system perspectives;
+3) evaluating these tradeoffs at runtime during data placement and movement,
+and executing the resultant policies in an autonomic system using models,
+heuristics and continuous learning; and
+4) leveraging techniques such as application-aware data compression, 
+and I/O prioritization to enforce these
+policies.
 \noindent
-Outline and writing assignments:
-later the abstract will go here.
-\begin{enumerate}
-\item Introduction - Scott - Hasan - Gary - Jay 4 pages
-   \begin{enumerate}
-      \item resource drivers
-      \item application drivers
-      \item overall example which we will use for the test
-      \item  need to answer the questions what we are doing, research  , prototype, aim, and what is new
-   \end{enumerate}
-\item Enabling Technologies - Jay - 1 page
-   \begin{enumerate}
-   	\item Sirocco
-   	\item ADIOS
-   	\item DataSpaces
-   	\item Anything else
-   \end{enumerate}
- \item {Proposed Research and Methods} - 14 pages
-    \begin{enumerate}
-       \item Application interface - Hasan, Manish, Gary, Kimmy, Jay, Mark
-          \begin{enumerate}
-            \item APIS - Hasan
-            \item description of data - Gary
-            \item collection of data reduction techniques - Mark
-            \item data re-organization specification - Scott, Mark, Kimmy
-            \item QoS - Hasan
-            \item execution policies - Manish
-            \item Initial placement and movement  - Gary
-            \item policies - Jay
-          \end{enumerate}
-      \item Storage System - Jay, Carlos, Manish, Lee, Matthew, Feyi
-      \begin{enumerate}
-         \item migration - Lee
-         \item resource management -Jay
-         \item QoS - Carlos 
-         \item Discovery - Jay
-         \item Eviction metrics - Carlos
-         \item Naming Service - Manish
-         \item Estimation of Times - Feyi
-      \end{enumerate}
-    \end{enumerate}
-\item Evaluation - 1 pages
-\item Timetable of Activities - 1/2 page
-\item Project Management Plan - 1 page
-\item Project Objective  1/2 page
-\end{enumerate}
 

### 2015-06-28 narrative/intro.tex
making changes...

@@
-\begin{itemize}
-
-\item describe data size problems for getting scientific results out
-\item describe annotating and differentiating data storage according to both
-  use and inherent data qualities (e.g., contains desireable features)
-\item describe idea of using heavy compression for less interesting areas
-  and less or no compression for more interesting areas.
-\item describe inherent errors in scientific calculations
-\item describe lossy compression within error bounds
-\item describe how incorporating plug-ins for application aware data
-  compression opportunities to aid data storage
-\item describe Sirocco's current state and future designs as a solid
-  foundation on which to work
-\item describe how these new features can enhance a system like Sirocco to
-  reduce data intensity
-\item describe metadata challenges that we will also have to face.
-
-\end{itemize}
-
-every section should include:
-
-Problem and background, one or more solution approaches, and investigation
-goals/questions.
-
-Technical Areas
-
-1. Application Intentions.\\
-2. Metadata Management (middleware and storage system)\\
-3. Pluggable infrastructure for data transformation (storage system)\\
-4. Intelligent (autonomic, application-aware) data mapping to storage/memory hierarchies (middleware and storage system)\\
-5. Predictable Performance and resolution tradeoffs\\
-6. Data re-generation. (middleware \& storage)\\
-7. Learning Motifs and data re-organization.\\
-
-\section*{Introduction}
-
+\section*{Introduction} {\color{red} Scott, Hasan, Gary, Jay - 4 pages}
+\label{sec:introduction}
 The Lightweight File Systems~\cite{lwfs} (LWFS) project from Sandia sought to

### 2015-06-28 narrative/intro.tex
working more on the introduction

@@
 \label{sec:introduction}
-The Lightweight File Systems~\cite{lwfs} (LWFS) project from Sandia sought to
+
+Exascale scientific discovery will be severely bottlenecked without
+sufficient new research into managing and storing the large amounts of data
+that will be produced during the simulation, and analyzed for months
+afterwards.
+%
+In this project we will demonstrate novel techniques to facilitate efficient
+mapping of data objects, even partitioning individual variables, from the
+user space onto multiple storage tiers, and enable application-guided data
+reductions/transformations to address capacity and bandwidth bottlenecks,
+while constraining the error to be within user provided bounds.
+%
+We will address the associated I/O and storage challenges in the context of
+current and emerging storage landscapes, and expedite insights into critical
+scientific processes, demonstrating the validity of our approach in key DOE
+domains. Our techniques will be to research novel techniques into a Scalable
+Storage Software Infrastructure by integrating services from the middle-ware
+layer which will talk to the applications, with the Storage system. The
+negotation between these layers will be a fundamental service which we will
+create in this project to ensure that users will be able to save `'the
+best'' amount of data in the different storage tiers.
+
+The metric we are most interested in optimizing is time to knowledge.
+Current approaches to addressing the I/O bottleneck fall into two broad
+categories. Parallel file system approaches that optimize the throughput for
+an entire system, and I/O middleware approaches that optimize the
+performance of a single application. Both approaches have seen success but
+are unlikely to overcome the major obstacles in reaching exascale. Instead
+of trying to optimize throughput, we will seek to reduce the time to
+knowledge. This is the most significant metric for scientific applications
+where the desired outcome is not storing data, but rather executing a
+knowledge extraction process on the data. Moreover, we aim to perform the
+optimization not for a single application, but rather for workload of a
+multi-user multi-application system environment.
+
+Our approach leverages the expected characteristics of exascale storage
+hardware. The storage layer will be partitioned into multiple heterogenous
+tiers with vastly different performance characteristics. This difference
+between layers will be further exacerbated by the constraints on capacity
+and data lifetime within a layer. Tape archival storage will still maintain
+data long term, but access to this data will be orders of magnitude slower
+than the next layer and naively accessing data from archival storage will
+greatly impact productivity.
+
+To address these challenges we will We are proposing to combine the applications
+knowledge from ORNL, from the ADIOS project, with the SNL LWFS system and storage system knowledge, the
+metadata and storage systems knowledge of UCSC, and the middleware expertise of
+Rutgers to investigate how to make a data, workflow, and application aware
+storage system.
+
+%ADIOS was developed with the understanding that we must
+%not only address the bottlenecks for current applications and
+%hardware platforms but also provide a path forward for the next
+%generation of applications and systems that would need to both
+%maximize bandwidth to the storage system and also support
+%transparently working around the storage system bandwidth
+%limitations with new techniques and tools. To support the
+%diverse operating modes of both using persistent storage and
+%other data storage and processing technology, we made a
+%great effort to provide a simplified interface to application
+%developers, offering a simple, portable and scalable way for
+%scientists to manage data that may need to be written, read
+%or processed during simulation runs. This required abstracting
+%away many decisions typically made in the application code
+%so that they may be configured externally.
+%In addition to this abstracted interface with external configuration
+%options, common services were incorporated to afford
+%optimizations beyond those for a single platform, such as
+%buffering, aggregation, subfiling, and chunking with options
+%to select each based on the data distribution characteristics
+%of the application. A variety of asynchronous I/O techniques
+%have been investigated and are being integrated with ADIOS.
+%A recognition of application complexity has led to new techniques
+%for testing I/O performance by extracting I/O patterns
+%from applications and automatically generating benchmark
+%codes. And finally, the march toward exascale has fueled the
+%need to consider additional features such as compression and
+%indexing to better cope with the expected tsunami of data.
+
+In this project we extend ADIOS with new APIs and novel techniques to talk to the
+storage back-end. We will also test our ideas on the next generation OLCF system along with
+the next generation NERSC system. One of the clear challenges of this work will be in accessing new APIs which
+will give the applications much more control and flexibility of their I/O.
+%
+In this project we will make use of the many inovations we discovered in 
+the Lightweight File Systems~\cite{lwfs} (LWFS) project, where Sandia sought to
 demonstrate stripping down a file system to the bare minimum features required
@@
 object storage system with integrated security providing authentication and
-autorization services, customized storage interfaces could be developed with
+authorization services, customized storage interfaces could be developed with
 exactly the overhead required for different applications. The second phase of
@@
 devices for efficient and scalable storage performance.  Sirocco has been in
-process for three years and has a usable storage infrastructure system. This
-proposal seeks to build upon this work to demonstrate an application, data, and
-usage aware storage system optimized for exascale composite applications and
-capability application runs. We are proposing to combine the applications
-knowledge from ORNL with the SNL LWFS system and storage system knowledge, the
-metadata and storage systems knowledge of UCSC, and the middleware expertise of
-Rutgers to investigate how to make a data, workflow, and application aware
-storage system.
+process for three years and has a usable storage infrastructure system.
+
+%  This
+%proposal seeks to build upon this work to demonstrate an application, data, and
+%usage aware storage system optimized for exascale composite applications and
+%capability application runs. We are proposing to combine the applications
+%knowledge from ORNL with the SNL LWFS system and storage system knowledge, the
+%metadata and storage systems knowledge of UCSC, and the middleware expertise of
+%Rutgers to investigate how to make a data, workflow, and application aware
+%storage system.
 
@@
 
+\subsection{Writing Perspective}
+\label{subsec:sim-perspective}
 From the perspective of the simulation scientist, they want to be able to
@@
 then place in a certain set of rules which the system can make autonomic
-decisions to help decide what should be done. In our simple scenario, a user
+decisions to help decide what should be done.
+%
+
+One of the key aspects of our approach is to allow the users to define the utility (importance) of their data;
+in other-words, for each data object, users need to come up with a way to define pieces of the object
+which have more utility (more importance) . There are many clear research challenges faced with this approach 
+including: 1) How do we make this simple enough so that users can do this, 2) can we come up with a model to
+allow the system to then make the autonomic decisions to save the data which has the highest utility, and 3) can
+we allow the back-end storage system to interact with this definition so that it can optimize not just over one application
+but across all of the applications.
+ %
+  In our simple scenario, a user
 wants to perhaps write 1 PB of data every hour for their checkpoint/restart,
@@
 with lower bandwidth than the parallel file system, or if some of the data
-will just go to tape. Each piece of data the user will specify the time they
-will want to keep the data round. We want the system to give the users a
-certain amount of currency in terms of bandwidth, storage space on each
-level, and latency expectations. These notions will be fuzzy but they will
-allow the user to make ad-hoc decisions to figure out what needs to be
-saved.
+will just go to tape. Each piece of data the user will specify the lifetime so that the 
+data with the most important utility is kept around for the longest period of time, but the data which has the least utility might
+be migrated to lower layers of the storage hierarchy faster.
 
@@
 
+\subsection{Storage System Perspective}
+\label{subsec:storage-perspective}
 From the perspective of the storage system, it needs to manage the data from
@@
 
+One of the clear research objectives of this project is to understand how the storage system can interact
+with the middleware (application-aware) piece so that the data with the highest utility is kept on the fastest 
+layers of the storage system for as long as specified. Clearly we need to maintain some sort of QoS such that
+we can give fuzzy-estimates of the time it will take to write/read data from the different layers, and clearly it is critical
+that the storage system can manage the data across all of the layers. We will use learning techniques to automatically
+migrate data sub-objects across the hierarchy as well as to eventually evict data which is large and has very low utility.
+
+\subsection{Reading  Perspective}
+\label{subsec:reading-perspective}
 From the perspective of the users who want to read in the data from their
@@
 \end{enumerate}
-
-Exascale scientific discovery will be severely bottlenecked without
-sufficient new research into managing and storing the large amounts of data
-that will be produced during the simulation, and analyzed for months
-afterwards.
 %
-In this project we will demonstrate novel techniques to facilitate efficient
-mapping of data objects, even partitioning individual variables, from the
-user space onto multiple storage tiers, and enable application-guided data
-reductions/transformations to address capacity and bandwidth bottlenecks,
-while constraining the error to be within user provided bounds.
+%Exascale scientific discovery will be severely bottlenecked without
+%sufficient new research into managing and storing the large amounts of data
+%that will be produced during the simulation, and analyzed for months
+%afterwards.
+%%
+%In this project we will demonstrate novel techniques to facilitate efficient
+%mapping of data objects, even partitioning individual variables, from the
+%user space onto multiple storage tiers, and enable application-guided data
+%reductions/transformations to address capacity and bandwidth bottlenecks,
+%while constraining the error to be within user provided bounds.
+%%
+%We will address the associated I/O and storage challenges in the context of
+%current and emerging storage landscapes, and expedite insights into critical
+%scientific processes, demonstrating the validity of our approach in key DOE
+%domains. Our techniques will be to research novel techniques into a Scalable
+%Storage Software Infrastructure by integrating services from the middle-ware
+%layer which will talk to the applications, with the Storage system. The
+%negotation between these layers will be a fundamental service which we will
+%create in this project to ensure that users will be able to save `'the
+%best'' amount of data in the different storage tiers.
 %
-We will address the associated I/O and storage challenges in the context of
-current and emerging storage landscapes, and expedite insights into critical
-scientific processes, demonstrating the validity of our approach in key DOE
-domains. Our techniques will be to research novel techniques into a Scalable
-Storage Software Infrastructure by integrating services from the middle-ware
-layer which will talk to the applications, with the Storage system. The
-negotation between these layers will be a fundamental service which we will
-create in this project to ensure that users will be able to save `'the
-best'' amount of data in the different storage tiers.
-
-The metric we are most interested in optimizing is time to knowledge.
-Current approaches to addressing the I/O bottleneck fall into two broad
-categories. Parallel file system approaches that optimize the throughput for
-an entire system, and I/O middleware approaches that optimize the
-performance of a single application. Both approaches have seen success but
-are unlikely to overcome the major obstacles in reaching exascale. Instead
-of trying to optimize throughput, we will seek to reduce the time to
-knowledge. This is the most significant metric for scientific applications
-where the desired outcome is not storing data, but rather executing a
-knowledge extraction process on the data. Moreover, we aim to perform the
-optimization not for a single application, but rather for workload of a
-multi-user multi-application system environment.
-
-Our approach leverages the expected characteristics of exascale storage
-hardware. The storage layer will be partitioned into multiple heterogenous
-tiers with vastly different performance characteristics. This difference
-between layers will be further exacerbated by the constraints on capacity
-and data lifetime within a layer. Tape archival storage will still maintain
-data long term, but access to this data will be orders of magnitude slower
-than the next layer and naively accessing data from archival storage will
-greatly impact productivity.
-
-We will achieve reduced time to knowledge using a combination of techniques;
-\begin{enumerate}
-\item Data annotations specified at the application level to quantify the
-  relative important, utility and lifetime of data objects;
-\item Partitioning of data objects across the storage hierarchy utilizing
-  the additional knowledge embedded in the annotations;
-\item Evaluating the tradeoffs at runtime to guide data placement, movement,
-  and migration across storage layers using models, heuristics and
-  continuous learning;
-\item Utilize the additional knowledge available about the data to perform
-  application-aware data compression and I/O prioritization;
-\end{enumerate}
-
-We are aiming to spread an output across the vertical layers of the storage
-hierarchy simultaneously. When data should migrate to a higher tier, what
-happens to the existing version?
-
-Is there a canonical copy that is the originally written version?
-
-Is there annotation about this (I think so) attached to that part of the
-variable?
-
-At the tape layer, we see having a lossy compressed version just above the
-tape layer used as a directory/index and then the ``full'' resolution
-version on tape only retrieving once the user accepts the time/quality
-tradeoff. This is going to require some serious language to describe. If we
-have 100s PB of tape space, we'll need 1s PB of disk even at a 99\% data
-reduction. That is non-trivial.
-
-What happens when that data is retrieved from tape? Is a third copy made in
-an appropriately sized/quality level version for the tier it is requested to
-be pulled to? What if it is just the index version? Or the on tape version?
-Does it have a TTL because of the pain retrieving it from tape?
-
-Do we specify a target tier or just allow the system to place it in a place
-that makes sense? Given the potential space limitations, I think this is
-pretty critical because it could cause evictions or other insufficient space
-actions that are unintended. I can see us specifying that when you ask for
-data to be pulled at a particular quality level to a particular performance
-tier that it is a best effort with the use specifying if compromise or just
-failure is the result if the request cannot be fulfilled. Making this
-interaction make sense is going to require some reasonable thought and
-testing. We'll probably need to run it past apps people to get their
-feedback too as I don't think we are able to give a solid answer without
-some broad input.
-
-How does the migration to ``data lakes''/``campaign storage'' work? How does
-the migration to tape ultimately happen? I think we need to incorporate some
-explicit staging commands to move data both up and down the stack along with
-some variability in placement based on data features and current system
-state.
-
-Given the storage scarcity, particularly for NVM, I think we need to have a
-solid story here as part of the proposal. Yes, there are questions that have
-to be answered to build this still, but we need to have some pretty solid
-clarity where possible. I don't feel like I can explain it well enough at
-what I believe to be a proper clarity level at this point.
+%The metric we are most interested in optimizing is time to knowledge.
+%Current approaches to addressing the I/O bottleneck fall into two broad
+%categories. Parallel file system approaches that optimize the throughput for
+%an entire system, and I/O middleware approaches that optimize the
+%performance of a single application. Both approaches have seen success but
+%are unlikely to overcome the major obstacles in reaching exascale. Instead
+%of trying to optimize throughput, we will seek to reduce the time to
+%knowledge. This is the most significant metric for scientific applications
+%where the desired outcome is not storing data, but rather executing a
+%knowledge extraction process on the data. Moreover, we aim to perform the
+%optimization not for a single application, but rather for workload of a
+%multi-user multi-application system environment.
+%
+%Our approach leverages the expected characteristics of exascale storage
+%hardware. The storage layer will be partitioned into multiple heterogenous
+%tiers with vastly different performance characteristics. This difference
+%between layers will be further exacerbated by the constraints on capacity
+%and data lifetime within a layer. Tape archival storage will still maintain
+%data long term, but access to this data will be orders of magnitude slower
+%than the next layer and naively accessing data from archival storage will
+%greatly impact productivity.
+%
+%We will achieve reduced time to knowledge using a combination of techniques;
+%\begin{enumerate}
+%\item Data annotations specified at the application level to quantify the
+%  relative important, utility and lifetime of data objects;
+%\item Partitioning of data objects across the storage hierarchy utilizing
+%  the additional knowledge embedded in the annotations;
+%\item Evaluating the tradeoffs at runtime to guide data placement, movement,
+%  and migration across storage layers using models, heuristics and
+%  continuous learning;
+%\item Utilize the additional knowledge available about the data to perform
+%  application-aware data compression and I/O prioritization;
+%\end{enumerate}
+%
+%We are aiming to spread an output across the vertical layers of the storage
+%hierarchy simultaneously. When data should migrate to a higher tier, what
+%happens to the existing version?
+%
+%Is there a canonical copy that is the originally written version?
+%
+%Is there annotation about this (I think so) attached to that part of the
+%variable?
+%
+%At the tape layer, we see having a lossy compressed version just above the
+%tape layer used as a directory/index and then the ``full'' resolution
+%version on tape only retrieving once the user accepts the time/quality
+%tradeoff. This is going to require some serious language to describe. If we
+%have 100s PB of tape space, we'll need 1s PB of disk even at a 99\% data
+%reduction. That is non-trivial.
+%
+%What happens when that data is retrieved from tape? Is a third copy made in
+%an appropriately sized/quality level version for the tier it is requested to
+%be pulled to? What if it is just the index version? Or the on tape version?
+%Does it have a TTL because of the pain retrieving it from tape?
+%
+%Do we specify a target tier or just allow the system to place it in a place
+%that makes sense? Given the potential space limitations, I think this is
+%pretty critical because it could cause evictions or other insufficient space
+%actions that are unintended. I can see us specifying that when you ask for
+%data to be pulled at a particular quality level to a particular performance
+%tier that it is a best effort with the use specifying if compromise or just
+%failure is the result if the request cannot be fulfilled. Making this
+%interaction make sense is going to require some reasonable thought and
+%testing. We'll probably need to run it past apps people to get their
+%feedback too as I don't think we are able to give a solid answer without
+%some broad input.
+%
+%How does the migration to ``data lakes''/``campaign storage'' work? How does
+%the migration to tape ultimately happen? I think we need to incorporate some
+%explicit staging commands to move data both up and down the stack along with
+%some variability in placement based on data features and current system
+%state.
+%
+%Given the storage scarcity, particularly for NVM, I think we need to have a
+%solid story here as part of the proposal. Yes, there are questions that have
+%to be answered to build this still, but we need to have some pretty solid
+%clarity where possible. I don't feel like I can explain it well enough at
+%what I believe to be a proper clarity level at this point.
 

### 2015-06-28 narrative/intro.tex
Significant edits to guide the understanding of the rest of the proposal in the beginning of the introduction. Key idea: user and storage system cooperating on storage goals.

@@
 %
-In this project we will demonstrate novel techniques to facilitate efficient
-mapping of data objects, even partitioning individual variables, from the
-user space onto multiple storage tiers, and enable application-guided data
-reductions/transformations to address capacity and bandwidth bottlenecks,
-while constraining the error to be within user provided bounds.
+In this project we will demonstrate a cooperative approach for storing data
+where the user and the storage system work together to achieve the best data
+quality and performance based on data features, storage tier characteristics,
+and system state. We will investigate novel techniques to facilitate efficient
+mapping of data objects, even partitioning individual variables, from the user
+space onto multiple storage tiers and enable application-guided data
+reductions/transformations to address capacity and bandwidth bottlenecks while
+constraining the error to be within user provided bounds.
 %
 We will address the associated I/O and storage challenges in the context of
-current and emerging storage landscapes, and expedite insights into critical
-scientific processes, demonstrating the validity of our approach in key DOE
-domains. Our techniques will be to research novel techniques into a Scalable
-Storage Software Infrastructure by integrating services from the middle-ware
-layer which will talk to the applications, with the Storage system. The
-negotation between these layers will be a fundamental service which we will
-create in this project to ensure that users will be able to save `'the
-best'' amount of data in the different storage tiers.
+current and emerging storage landscapes and expedite insights into critical
+scientific processes demonstrating the validity of our approach in key DOE
+domains. We will research novel techniques into a Scalable
+Storage Software Infrastructure integrating services from the middle-ware
+layer adapted to each application with Storage System features to support the
+novel data distribution both horizontally and vertically across the storage
+system. The negotation between these layers will be a fundamental service 
+to ensure that users will be able to save `'the best'' amount of data using the
+right storage technology available at request time across the entire storage
+stack.
 
@@
 
-To address these challenges we will We are proposing to combine the applications
-knowledge from ORNL, from the ADIOS project, with the SNL LWFS system and storage system knowledge, the
-metadata and storage systems knowledge of UCSC, and the middleware expertise of
-Rutgers to investigate how to make a data, workflow, and application aware
+To address these challenges we will We are proposing to combine the
+applications knowledge from ORNL, from the ADIOS project, with the SNL
+Lightweight File Systems project and storage system knowledge, the metadata
+management and storage systems knowledge of UCSC, and the middleware expertise
+of Rutgers to investigate how to make a data, workflow, and application aware
 storage system.
@@
 
-In this project we extend ADIOS with new APIs and novel techniques to talk to the
-storage back-end. We will also test our ideas on the next generation OLCF system along with
-the next generation NERSC system. One of the clear challenges of this work will be in accessing new APIs which
-will give the applications much more control and flexibility of their I/O.
+In this project we extend ADIOS~\cite{adios} with new APIs and novel techniques
+to talk to the storage back-end. The key idea is offering the ability to
+negotiate with the storage system to agree on the tradeoff between data quality
+and response time.  We will test our ideas on the next generation OLCF,
+SNL/LANL, and  NERSC systems.
 %
@@
 auxiliary services separate from the stack or by layering providing a new
-interface to the storage system. The project demonstrated that using a core
-object storage system with integrated security providing authentication and
-authorization services, customized storage interfaces could be developed with
-exactly the overhead required for different applications. The second phase of
-LWFS, known as Sirocco~\cite{sirocco}, is investigating how to manage storage
-devices for efficient and scalable storage performance.  Sirocco has been in
-process for three years and has a usable storage infrastructure system.
+interface to the storage system. The project's first phase demonstrated that
+using a core object storage system with integrated security providing
+authentication and authorization services, customized storage interfaces could
+be developed with exactly the overhead required for different applications. The
+second phase of LWFS, known as Sirocco~\cite{sirocco}, is investigating how to
+manage storage devices for efficient and scalable storage performance.  Sirocco
+has been in process for three years and has a usable storage infrastructure
+system.
 
@@
 amount of Quality of Service such that they can then make decisions when the
-expected bandwidth, for example is less than what they desire. The user will
+expected bandwidth, for example, is less than what they desire. The user will
 then place in a certain set of rules which the system can make autonomic
@@
 
-One of the key aspects of our approach is to allow the users to define the utility (importance) of their data;
-in other-words, for each data object, users need to come up with a way to define pieces of the object
-which have more utility (more importance) . There are many clear research challenges faced with this approach 
-including: 1) How do we make this simple enough so that users can do this, 2) can we come up with a model to
-allow the system to then make the autonomic decisions to save the data which has the highest utility, and 3) can
-we allow the back-end storage system to interact with this definition so that it can optimize not just over one application
-but across all of the applications.
+One of the key aspects of our approach is to allow the users to define the
+utility (importance) of their data; in other-words, for each data object, users
+need to come up with a way to define pieces of the object which have more
+utility (more importance) . There are many clear research challenges faced with
+this approach including: 1) How do we make this simple enough so that users can
+do this, 2) can we come up with a model to allow the system to then make the
+autonomic decisions to save the data which has the highest utility, and 3) can
+we allow the back-end storage system to interact with this definition so that
+it can optimize not just over one application but across all of the
+applications.
  %
-  In our simple scenario, a user
-wants to perhaps write 1 PB of data every hour for their checkpoint/restart,
-and they which to write 500TB of data every 15 minutes for their analysis
-and visualization results which have been reduced by using in situ reduction
-techniques. At this stage, they will ask the system to write their data, and
-they would like to get an estimated time which they can then figure out,
-through a series of rules, whether they want to write out all of the data or
-if they want to write none of the data but rather wait for a later time to
-write or write a reduced amount of data, along with some code container to
-allow them to regenerate the data with a certain level of accuracy.
-Furthermore, they realize that the analysis data will be saved on either the
-parallel file system, or the campaign storage, a longer term storage area
-with lower bandwidth than the parallel file system, or if some of the data
-will just go to tape. Each piece of data the user will specify the lifetime so that the 
-data with the most important utility is kept around for the longest period of time, but the data which has the least utility might
-be migrated to lower layers of the storage hierarchy faster.
+  In our simple scenario, a user wants to perhaps write 1 PB of data every hour
+for their checkpoint/restart, and they which to write 500TB of data every 15
+minutes for their analysis and visualization results which have been reduced by
+using in situ reduction techniques. At this stage, they will ask the system to
+write their data, and they would like to get an estimated time which they can
+then figure out, through a series of rules, whether they want to write out all
+of the data or if they want to write none of the data but rather wait for a
+later time to write or write a reduced amount of data, along with some code
+container to allow them to regenerate the data with a certain level of
+accuracy.  Furthermore, they realize that the analysis data will be saved on
+either the parallel file system, or the campaign storage, a longer term storage
+area with lower bandwidth than the parallel file system, or if some of the data
+will just go to tape. Each piece of data the user will specify the lifetime so
+that the data with the most important utility is kept around for the longest
+period of time, but the data which has the least utility might be migrated to
+lower layers of the storage hierarchy faster.
 
@@
 tremendous amount of interference from other users on the system who may be
-writing or reading at the same time. By maiming a certain level of quality
+writing or reading at the same time. By requiring a certain quality
 of service the system must potentially lock out users with lower currency
@@
 
-One of the clear research objectives of this project is to understand how the storage system can interact
-with the middleware (application-aware) piece so that the data with the highest utility is kept on the fastest 
-layers of the storage system for as long as specified. Clearly we need to maintain some sort of QoS such that
-we can give fuzzy-estimates of the time it will take to write/read data from the different layers, and clearly it is critical
-that the storage system can manage the data across all of the layers. We will use learning techniques to automatically
-migrate data sub-objects across the hierarchy as well as to eventually evict data which is large and has very low utility.
+One of the clear research objectives of this project is to understand how the
+storage system can interact with the middleware (application-aware) piece so
+that the data with the highest utility is kept on the fastest layers of the
+storage system for as long as specified. Clearly we need to maintain 
+QoS such that fuzzy-estimates of the time it will take to write/read data from
+the different layers informs any migration/eviction decisions, and clearly it
+is critical that the storage system can manage the data across all of the
+layers. We will use learning techniques to automatically migrate data
+sub-objects across the hierarchy as well as to eventually evict data which is
+large and has very low utility.
 
@@
 doing interactive analysis and expects that the data they read in is about
-10s, but it terms into 1,000s, the user typically tends to either wait
-another day for doing their analysis, or gets frustrated and reads much less
+10s, but it turns into 1,000s, the user typically tends to either wait
+another day for doing their analysis or gets frustrated and reads much less
 data than they really need to. Furthermore the user tends to read in just a

### 2015-06-28 coverpage.tex

%------------------------------------------------------------------------------
%                 P r o p o s a  l   C o v e r   P a g e
%------------------------------------------------------------------------------
\thispagestyle{empty}

\begin{center}
\textbf{\Large \baseproposaltitle}
\end{center}

\begin{center}
%\includegraphics[width=1in]{LOGO.gif}
\end{center}

\medskip

%% Application/Institution
\begin{tabbing}
\hspace*{10mm} \= \makebox[65mm][l]{\textbf{Applicant Institution:}} \= Oak Ridge National Laboratory \\
\>  \>PO BOX 2008 MS6057 Oak Ridge TN 37831-6057\\
\end{tabbing}

\medskip

%% Lead PI name, telephone number, email:
\begin{tabbing}
\hspace*{10mm} \= \makebox[65mm][l]{\textbf{Lead Principal Investigator:}} \=
\vrule height0.4pt depth 0pt width 2.5in \\
\> \> Scott KIasky,  Distinguished Scientist \\
\> \> Oak Ridge National Laboratory \\
\> \> Computer Science and Mathematics Division \\
\> \> Oak Ridge, TN 37831-6057 \\
\> \> (T) 865-241-9980 (Fax) 865-241-4811 \\
\> \> klasky@ornl.gov
\end{tabbing}

\medskip

%% Adminisrative Point of Contact name, telephone number, email:
\begin{tabbing}
\hspace*{10mm} \= \makebox[65mm][l]{\textbf{Administrative Point of Contact:}} \=
\vrule height0.4pt depth 0pt width 2.5in \\
\> \> Tammy Reed \\
\> \> Oak Ridge National Laboratory \\
\> \>Computer Science and Mathematics Division \\
\> \> Oak Ridge, TN 37831-6057 \\
\> \> (T) 865-574-1936 (Fax) 865-241-4811\\
\> \> tsr@ornl.gov
\end{tabbing}

\bigskip

\begin{tabbing}
\hspace*{10mm} \= \makebox[80mm][l]{\textbf{Program Announcement Number:}} \= LAB 15-1338 \\
\> \textbf{DOE/Office of Science Program Office:} \> Office of Advanced Scientific Computing Research\\
\> \textbf{DOE/Office of Science Program Office}  \> \\
\> \makebox[70mm][r]{\textbf{Technical Contact:}} \> Lucy Nowell\\
\end{tabbing}

\medskip

\pagebreak

\begin{tabular}{| l | l | l | r | r | r | r |} \hline
Role & Name & Institution & Year 1 (\$K) & Year 2 (\$K) &   Year 3 (\$K) & Total (\$K) \\
\hline
Lead-PI & Scott Klasky & ORNL & 660 & 660 & 660 & 1,980 \\
Lead PI & Jay Lofstead & Sandia  & 350 & 350 & 350 & 1,050 \\
Co-PI & Carlos  Maltzahn & UCSC  & 160 & 160 & 160 & 480 \\
Co-PI & Manish Parashar & Rutgers & 80 & 80 & 80 & 240 \\
\hline
{\bf Totals}  & & & {\bf 1,250} & {\bf 1,250} & {\bf 1,250} & {\bf 3,750}\\
\hline
\end{tabular}

\newpage

### 2015-06-28 narrative/management.tex

@@
 \section*{Management}
+\label{sec:management}
 
-\begin{itemize}
+Careful management will be required to ensure that the problems and delays that are inevitable in any research project
+do not unduly disrupt overall progress. To this end, we define a detailed workplan,
+establish a  management structure with defined oversight responsibilities, and establish a schedule of regular meetings and
+reviews to ensure progress.  We will use google docs to share documents, emails lists for informal communications and a 
+weekly phone call in the main research areas to discuss project tasks and milestones.
 
-\item We need to manage things
+Dr. Klasky will oversee the outcome of this project,  Dr. Abbasi (ORNL) will lead the Application Interface area, and Dr. Lofstead
+will lead the Storage System. This project is focused on the integration of application aware-middleware which communications to the storage system,
+and each of these leaders have a strong track record of tight-collaboration, both working with Dr. Klasky for their Ph.D. at the Georgia
+Institute of Technology.  Our structure includes an area lead for each major research area plus
+an integration lead responsible for coordinating periodic integration tests. The area leads plus the project PIs
+form the management team, which will meet weekly to review progress and revise plans as needed.
+We will also have a {\bf Science Advisory Board} consisting of one application Scientist, Dr. C. S. Chang (PPPL), one storage system expert from the 
+OLCF, Dr. Vazhkudai (ORNL) who we will advice us yearly on our project goals and outcomes and make adjustments accordingly. We will also have our
+lead Math Lead, Dr. Ainsworth, to work with the applied mathematics community to ensure that the hierarchical decomposition of data will work in theory
+for current and next generation exascale simulations.
 
-\end{itemize}
+We will conduct six-monthly {\bf integration fests}  to validate our ability to combine models, apply
+combined models to real problems, and cope with the challenges of real data. We apply this approach routinely
+in other projects and find that it helps to focus effort and identify problems early. The tasks of defining,
+scheduling, and running critical post hoc reviews of integration tests will be coordinated by the Integration Lead, Dr. Parashar. 
+These meetings will bring all hands together at least once a year.
+
+We will monitor and report {\bf success metrics} including: 
+research publications;
+our ability to predict and explain performance of I/O workloads on the current machines., in different environments;
+our ability to reduce data and charcterize data from an object into multiple bins, 
+our ability to integrate our routines into the Sirocco file system and test the feasibily on current and next generation LCF systems, and
+how tools lead to workflow performance improvements, for both reading and writing data from our exemplar applications which we will test.
+
+\begin{table}[t]
+\caption{S2E2 Integration Tests. We list for each the technology area involved and new technologies expected.}
+\centering \footnotesize
+\vspace{2ex}
+\label{tab:ifests}
+%\begin{tabular}{| p{2cm} | p{2.5cm} | p{5cm} | p{5cm} | }
+\begin{tabular}{| l | l |l |l |} 
+\hline
+{\bf \#} &  {\bf Focus} & {\bf New technologies to be introduced} \\\hline
+1 & TBD			 & TBD\ \ \\
+2 &TBD & TBD\\
+3 & TBD         & TBD \\
+4 & TBD                                    & TBD  \\
+5 & TBD                                    & TBD \\
+6 & TBD                                    & TBD  \\\hline
+\end{tabular}
+\vspace{-8ex}
+\end{table}

### 2015-06-28 narrative/discovery.tex
Put in a first pass at discovery, naming, and resource management. Still need to add deliverables.

\subsubsection{Discovery}

Hierarchical Storage Management (HSM) systems offer a strct caching approach
for managing different storage capacities trading off performance for capacity.
By maintaining a single namespace across all tiers, it is possible to list a
single directory view with files stored at different tiers. While this approach
to managing multi-tier storage works, it is far from ideal for scientific
simulations.

Our goal with this proposal is to offer a similar capability, but use a finer
granularity. Instead of a single file such as might be used to store an entire
timestep output for a simulation, we will demonstrate offering the same
capability at a subset of a single variable level. By shifting to a
finer-grained approach, we will enable more effective use of close/fast/small
storage tiers. Traiditional HSM stores an entire file on a tier making room if
insufficient space is available. With this shift to a partial variable
granularity, a 1 PB output with 500 GB of ``high interest'' data can limit this
costly tier ussage to just 500 GB greatly enhancing usability.

Tagging a variable subset as ``high interest'' requires intervention from the
application and/or middleware to determine what data meets this criteria. The
storage system itself simply needs to offer an ability to perform different
actions based on this information.

The challenge for discovery is that potentially, data will migrate from where
it is initially stored to a new location within the storage system. Sirocco
offers an ability to search for data that has moved as well as forcing a
particular resilience-based replica be the ``authoritative'' version. We will
investigate if the current Sirocco functionality is capable of supporting the
new operating modes we wish to offer. Initial expectations suggest having
bounded time guarantees for finding data are critical for offering the quality
of service guarantees we wish to offer. This new work will be an expansion of
Sirocco's currently planned features.

The other aspect of discovery is the negotiation between the user and the
storage system for a data quality/retrevial time tradeoff. The naming service
will work hand-in-hand with the discovery, data migration, and time estimation
services to offer the best possible options for data retrevial based on quality
of service requested.

### 2015-06-29 narrative/resource-mgmt.tex

\subsubsection{Resource Management}

Sirocco currently offers automatic resource management capabilities. However,
these features are strictly driven on resource capacity and data resilience
requirements. For this project, we seek to expand the resource management
capabilities to support additional metrics to guide resource management
decisions. Our basic goal is to offer data-centric metrics that help guide the
forced resource management decisions.

As stated in the introduction, we seek to make a storage stack that works with
the user. Towards that end, we must incorporate data attributes into this
process. For example, particularly for a capability run, the storage system
should give priority to data subset tagged with high importance to stay in fast
storage tiers. Enabling this priority will be driven by other job
characteristics such as job size or a manually set priority inserted by system
administrators. It should also store the highest quality data possible given
the performance requirement. Ideally, the storage system will store the raw
data at full fidelity. When this tier approaches capacity, data should migrate
not just based on age, but also based on this priority. This will help ensure
the shortest time to insight by making sure the most important machine runs
have priority to use system resources.

Existing efforts on scheduing resources~\cite{io-cop,dorier} attempt to offer
admission control to maximize storage performance. Unfortunately, these efforts 
are limited to participating applications and only applications from a single
platform. For this approach to be effective, the storage system must offer an
approach that handles clients from all connected platforms and does not
require modification to use new storage access APIs. By using a priority
tagging system, we will offer defaults suitable for all applications that can
be informed by simple extensions to the job scheduler or more advanced
management through new APIs.

What else do we need to talk about?

### 2015-06-29 narrative/naming-service.tex

\subsubsection{Naming Service}

A traditional POSIX naming service offers a hierarchical space consisting of
directories and files. LWFS and Sirocco have abandonded this as the default
model in favor of a container/object/fork/addrress tuple for identifying data.
LWFS demonstrated a POSIX-style namespace on the side kept in sync using a
transaction process like D2T~\cite{d2t} showing that this alternative approach
can support traditional POSIX API calls even though the underlying storage
system uses a different model.

For this project, we seek to offer a similar POSIX-style namespace for naming
objects, but store a rich metadata collection with each entry to handle data
migration, discovery, and writing and retrevial time estimations. Further, we
will also have to track data stored in different tiers, even at the variable
subset level. This additional metadata load will require a more complex naming
service that reacts to automatic data movement while caching current data
locations accelerating data access operations avoiding any data discovery
operations.

In addition to the basic naming operations, we will also need to incorporate
authorization capabilities. Sirocco currently integrates with a Kerberos
service for authentication and authorization. Given a capability ticket, a
user can access different objects as needed.

The main challenge of incorporating the additional, rich metadata will be
joined by the additional challenge of coordinating with the other storage
and application-layer services to offer the best access times possible for data
stored in the system.

### 2015-06-29 narrative/apis.tex

\subsubsection{APIs}

The new APIs will provide new functionality of specifying selectable performance/quality/- cost tradeoffs from both the application and system perspectives based upon the user guided rules/policy and runtime system monitoring status. It allows the middleware to make best possible decisions from the feedback of storage system knowledge, such that it will embed user intentions and the available system storage. The user would like the ability to get a certain amount of Quality of Service such that they can then make decisions when the expected bandwidth, for example is less than what they desire and will then place in a certain set of rules which the system can make autonomic decisions to help decide what should be done. We want the system to give the users a certain amount of currency in terms of bandwidth, storage space on each level, and latency expectations. These notions will be fuzzy but they will allow the user to make ad-hoc decisions to figure out what needs to be saved.

When user ask the system to write their data, and they would like to get an estimated time which they can then figure out, through a series of rules, whether they want to write out all of the data or wait for a later time to write or write a reduced amount of data. The decision is based upon if the system monitoring model could give us the estimate time during the reading and writing process, so that rules can be in place from the user perspective to make adaptive decisions.

If user decide to write or read data to or from hierarchical storage systems, they will rely on the API to send their intentions for inquiry and examine the system status, including how much data they would like to write/read, desired bandwidth, data compression method, etc, or the user can express their intension of writing data right now no matter what the traffic is now. After receiving the negotiation signal, the API return value will contain the estimate time of write/read data, potential bandwidth, concurrency, etc. User will adapt their decision from the system feedback and determine how much data they will write/read if the estimated time is too long for them.

The capability of APIs provide dynamic runtime information to user and make the system more transparent to them rather than a black box. The end user could analyze or visualize data in a more comfortable position, where they are able to envision how much longer they are going to wait for data.

### 2015-06-29 narrative/intro.tex
Polishing the intro

@@
 
-Exascale scientific discovery will be severely bottlenecked without
-sufficient new research into managing and storing the large amounts of data
-that will be produced during the simulation, and analyzed for months
+Avoiding impending bottlenecks in exascale scientific discovery will require
+new research into managing and storing the large amounts of data
+that will be produced by simulations and analyzed for months
 afterwards.
@@
 In this project we will demonstrate a cooperative approach for storing data
-where the user and the storage system work together to achieve the best data
-quality and performance based on data features, storage tier characteristics,
-and system state. We will investigate novel techniques to facilitate efficient
+where the user and the storage system work together to achieve the best possible data
+quality and performance given data features, storage tier characteristics,
+and current system state. We will investigate novel techniques to facilitate efficient
 mapping of data objects, even partitioning individual variables, from the user
@@
 current and emerging storage landscapes and expedite insights into critical
-scientific processes demonstrating the validity of our approach in key DOE
-domains. We will research novel techniques into a Scalable
-Storage Software Infrastructure integrating services from the middle-ware
-layer adapted to each application with Storage System features to support the
-novel data distribution both horizontally and vertically across the storage
-system. The negotation between these layers will be a fundamental service 
-to ensure that users will be able to save `'the best'' amount of data using the
-right storage technology available at request time across the entire storage
-stack.
+scientific processes, while demonstrating the validity of our approach in key DOE
+domains. We will research techniques for creating a Scalable
+Storage Software Infrastructure, integrating services from the middleware
+layer with Storage System features to support the
+novel strategy of distributing data both horizontally and vertically across the storage
+system. 
+The negotation between these layers will be provided as a fundamental service, 
+allowing users to easily save ``the best'' amount of data, automatically placed
+throughout the entire storage stack.
 
@@
 Current approaches to addressing the I/O bottleneck fall into two broad
-categories. Parallel file system approaches that optimize the throughput for
+categories: parallel file system approaches that optimize the throughput for
 an entire system, and I/O middleware approaches that optimize the
-performance of a single application. Both approaches have seen success but
-are unlikely to overcome the major obstacles in reaching exascale. Instead
+performance of a single application. Both approaches have been successful to date, but
+are unlikely to overcome the remaining obstacles in reaching exascale. Instead
 of trying to optimize throughput, we will seek to reduce the time to
@@
 knowledge extraction process on the data. Moreover, we aim to perform the
-optimization not for a single application, but rather for workload of a
+optimization not for a single application, but rather for workloads within a
 multi-user multi-application system environment.
 
-Our approach leverages the expected characteristics of exascale storage
+Our approach targets the expected characteristics of exascale storage
 hardware. The storage layer will be partitioned into multiple heterogenous
-tiers with vastly different performance characteristics. This difference
-between layers will be further exacerbated by the constraints on capacity
-and data lifetime within a layer. Tape archival storage will still maintain
+tiers with vastly different performance characteristics. The layers
+will be further differentiated by the constraints on capacity
+and data lifetime within each layer. Tape archival storage will still maintain
 data long term, but access to this data will be orders of magnitude slower
@@
 
-To address these challenges we will We are proposing to combine the
-applications knowledge from ORNL, from the ADIOS project, with the SNL
-Lightweight File Systems project and storage system knowledge, the metadata
+To address these challenges we will We are proposing to combine knowledge
+from a wide range of sources including ORNL applications, the ADIOS project,
+the SNL Lightweight File Systems project and storage system knowledge, the metadata
 management and storage systems knowledge of UCSC, and the middleware expertise
-of Rutgers to investigate how to make a data, workflow, and application aware
-storage system.
+of Rutgers. We will integrate this diverse expertise to investigate how to make
+a data, workflow, and application aware storage system.
 
@@
 
-In this project we extend ADIOS~\cite{adios} with new APIs and novel techniques
-to talk to the storage back-end. The key idea is offering the ability to
+In this project we will extend ADIOS~\cite{adios} with new APIs and novel techniques
+to coordinate with the storage back-end. The key idea is offering the ability to
 negotiate with the storage system to agree on the tradeoff between data quality

### 2015-06-29 narrative/migration.tex
added Lee's migration text.

\subsubsection{Migration}

One important research question in this proposal is whether and how a liberated
Hierarchical Storage Manager designed for online use, such as Sirocco, might be
employed in supporting the higher level concepts discussed previously.
Conceptually, in this effort, data is accessed by the middleware IO libraries
at various tiers in the lower-level storage hierarchy. While Sirocco does
autonomously group the distributed set of media into like groups and manage
movement of data between these groups the various motivations to do so include
both the traditional ones found in classical HSM (Hierarchical Storage
Management) as well as the novel.

Sirocco groups like media into aggregated volumes on a service node, and like
service nodes with like media into pseudo-tiers. It accomplishes both using
attributes such as latency to first byte, bandwidth to and from the media, and
some function representing the resilience capabilities of the attached media.
In addition to these static attributes, when grouping servers into
pseudo-tiers, Sirocco also may employ more dynamic attributes such as CPU load
and the rate-of-change of the media use.

Grouping into pseudo-tiers is not a formal, nor rigid, concept within Sirocco,
at present. In classic HSM the tiering is often a formally expressed concept,
embodied both in the architecture and the implementation. In Sirocco, though,
it is not directly expressed. Instead, a client or server that attempts to make
a decision relative to migration and staging uses the attributes discussed
above to constrain the choice of candidates. For instance, a server acting as
an off-node writeback cache for a compute node might only have access to
volatile dynamic RAM for use as storage media and a client might deposit data
that is marked for persistent media as a resilience constraint. At some point,
then, such a server must copy or move the data to another Sirocco server with
media matching, or exceeding, the resilience constraint. At that point in time,
all servers managing appropriate persistent media are candidates but
second-order attributes, such as latency and bandwidth of a candidate's managed
media would tend to dominate the ranking. Thus, the concept of a ``tier''
becomes an artifact of decision rather than a formal concept.

Further complication arises when a Sirocco server is forced to make choices for
capacity reasons. The Sirocco architecture dictates that a server function as a
victim cache. As media managed by a server fills the server will begin to
manage for capacity. In order to do so, currently, it need only verify or make
a copy of the data on another server that meets the resiliency constraint of
the held data prior to removing the local copy and in the current
implementation the server makes no attempt either to remember it ever had the
data or to notify others about what it is doing.  others that it has removed
the local copy or remember that it has

Taken together then, the above at least complicates the higher level research
questions discussed in this proposal, potentially even renders some
unachievable since a few involve determinism. We are forced, then, to consider
either or both mitigating management policies within and between Sirocco
servers as well as explicitly allowing clients to participate in capacity
management decisions, for a time at least, on the relevant Sirocco servers.

Our research intent involves short term and long term activities. In the short
term we will modify the relevant clients to provide hit and miss related data
and associated activities and explore an enhancement to our storage servers
that would provide some, hopefully tightly, bounded estimate about time to
first byte for a potential read operation. In the long term we hope that an
augmented client-server API providing usage hints will allow us to modify local
server policy algorithms to more effectively manage it's media and communicate
to other servers holding copies of the relevant data how they might best manage
theirs. In the event that this approach proves ineffective at meeting the
overall approach described in this proposal we will explore modifying Sirocco
servers to provide notifications about locations of the relevant data as they
migrate and stage as well as supporting client-directed pinning of location,
for a period at least. Client-directed pinning could introduce potentially
serious negative impacts to Sirocco's overarching goals though, since the
fundamental design of Sirocco assumes much about opaque and unrestricted data
movement, so this is contemplated only as a last resort.

### 2015-06-29 narrative/intro.tex
More polishing

@@
 
-To address these challenges we will We are proposing to combine knowledge
+To address these challenges we are proposing to combine knowledge
 from a wide range of sources including ORNL applications, the ADIOS project,
@@
 %
-In this project we will make use of the many inovations we discovered in 
-the Lightweight File Systems~\cite{lwfs} (LWFS) project, where Sandia sought to
+We plan to leverage the many inovations we discovered in 
+the Lightweight File Systems~\cite{lwfs} (LWFS) project.
+In this project, Sandia sought to
 demonstrate stripping down a file system to the bare minimum features required
 for any storage system. Then, additional functionality and semantics, such as
-POSIX-style directories and consistency requirements, are added in using
+POSIX-style directories and consistency requirements, were added in using
 auxiliary services separate from the stack or by layering providing a new
@@
 manage storage devices for efficient and scalable storage performance.  Sirocco
-has been in process for three years and has a usable storage infrastructure
-system.
+has been in process for three years and has developed a usable storage infrastructure.
 
-%  This
-%proposal seeks to build upon this work to demonstrate an application, data, and
-%usage aware storage system optimized for exascale composite applications and
-%capability application runs. We are proposing to combine the applications
-%knowledge from ORNL with the SNL LWFS system and storage system knowledge, the
-%metadata and storage systems knowledge of UCSC, and the middleware expertise of
-%Rutgers to investigate how to make a data, workflow, and application aware
-%storage system.
-
-We will examine the research challenges from the perspective of the
+We will examine the research challenges from the perspective of
 %
-(1) user of the system who is attempting to run an exascale simulation,
+(1) the user of the system who is attempting to run an exascale simulation,
 %
-(2) the Storage System and I/O layer which needs to negotiate amongst all of
+(2) the Storage System and I/O layer which needs to negotiate between all of
 the users, and finally
@@
 (3) the user of the system who is attempting to understand the data which
-was produced from their collection of simulations.
+was produced from a set of simulations.
 

### 2015-06-29 narrative/intro.tex
polishing intro

@@
 %
-(3) the user of the system who is attempting to understand the data which
-was produced from a set of simulations.
+(3) the user of the system who is attempting to understand the data
+produced by a set of simulations.
 
@@
 \label{subsec:sim-perspective}
-From the perspective of the simulation scientist, they want to be able to
-have knowledge about how much time they will put into writing their data,
+The simulation scientist ideally should be able to
+obtain an estimate of how much time will be required to write their data,
 and how much storage space they will be able to get during the lifetime of
-their simulations. Furthermore, they would like the ability to get a certain
-amount of Quality of Service such that they can then make decisions when the
-expected bandwidth, for example, is less than what they desire. The user will
-then place in a certain set of rules which the system can make autonomic
-decisions to help decide what should be done.
+their simulations. Furthermore, they would like the ability to get an assurance
+of Quality of Service such that they can react appropriately when the
+expected bandwidth, for example, is less than what they desire. Ideally, the user
+would specify a set of rules with which the system can make autonomic
+decisions to determine what should be done.
 %
@@
  %
-  In our simple scenario, a user wants to perhaps write 1 PB of data every hour
-for their checkpoint/restart, and they which to write 500TB of data every 15
-minutes for their analysis and visualization results which have been reduced by
+  In our simple scenario, a user wants to write 1 PB of data every hour
+for checkpoint/restart, and they wish to write 500TB of data every 15
+minutes to save analysis and visualization results which have already been reduced
 using in situ reduction techniques. At this stage, they will ask the system to
@@
 then figure out, through a series of rules, whether they want to write out all
-of the data or if they want to write none of the data but rather wait for a
-later time to write or write a reduced amount of data, along with some code
-container to allow them to regenerate the data with a certain level of
+of the data, postpone the write until a later time, or write a reduced amount
+of data, along with some code
+container to allow them to regenerate the data with a specified level of
 accuracy.  Furthermore, they realize that the analysis data will be saved on
-either the parallel file system, or the campaign storage, a longer term storage
-area with lower bandwidth than the parallel file system, or if some of the data
-will just go to tape. Each piece of data the user will specify the lifetime so
-that the data with the most important utility is kept around for the longest
+either the parallel file system, long-term tape storage, or on the campaign storage,
+which is a medium-term storage
+area with lower bandwidth than the parallel file system.
+For each piece of data, the user will specify the lifetime so
+that the data with the highest utility is kept around for the longest
 period of time, but the data which has the least utility might be migrated to
-lower layers of the storage hierarchy faster.
+lower layers of the storage hierarchy more quickly.
 
-Since the data will be a collection of objects placed to the storage system,
+Since the data will be a collection of objects placed on the storage system,
 some of the objects will be broken up into multiple pieces by ``plug-ins''

### 2015-06-29 qos-storage.tex
new file:   qos-storage.tex

\paragraph{Background:} ``Quality of Service'' (QoS) refers to the
properties of the performance qualtiy of a particular service.
Performance quality can be expressed in either relative or absolute
terms. Examples of relative performance qualities are ``fair'',
``proportional'', or ``priority/class-based'' (``the higher the
priority the better the service'' or ``1st class is better than
2nd''). The key advantage of relative performance qualities is that
they are easy to implement. The key disadvantage is the difficulty
of making any strong guarantees based on those implementations --
even relative guarantees are mired by the well-known effect of
priority inversion~\cite{lampson:cacm80}. Examples of absolute
performance qualities are ``rate-based'', ``soft real-time'', and
``hard real-time''. The key advantage of absolute performance
qualities is that it is principally easier to implement strong
guarantees: they enable an agreement between a task and a service,
also known as a \emph{reservation}, does not change meaning depending
on the service's state, as long as the service is able to isolate
performance between different tasks. The key disadvantage of absolute
performance qualities is that their implementation is non-trivial,
especially in large-scale systems.

\paragraph{Approach:} The goals of this proposal require applications
to demand guarantees in terms of absolute performance qualities.
The first key enabler for any service to make absolute guarantees
is \emph{performance isolation}, i.e. the ability to shield the
performance impact of one task from another. The most straight-forward
but least flexible strategy is to simply cap the latency of each
task, irregardless of its completeness (see for
example~\cite{decandia:sosp07}). This strategy is applicable when
incomplete tasks still have value and missing results are either
not important or can be easily retrieved in subsequent tasks. A
more generally applicable strategy is to control the admission of
tasks to the system: \emph{admission control} maintains a
\emph{utilization model} that can predict the utilization of system
resources given a new task, estimates the current utilization of
the system, and decides whether the task can be admitted without
overloading the system. Thus, admission control avoids system
overload conditions that lead to hard-to-model chaotic behaviors.
Consequently, utilization models can frequently be approximated by
automatically calibrated linear models (see for
example~\cite{skourtis:hpdc12}).

Another essential component of performance isolation is the
\emph{charging model} which determines which task(s) should be
charged how much for a request. Performance isolation is implemented
by the charging model by ensuring that performance of a task only
depends on its reservation and its workload behavior (but not on
other workloads in the system). Each task has a budget based on its
reservation and spends it based on the cost of each request determined
by the charging model. A charging model determines the cost based
on the utilization model and the interactions a request can have
with other requests (e.g. a random request to spinning media
interrupting a stream of sequential requests).

Performance isolation (including its utilization and charging models)
critically depends on \emph{workload-independent metrics}. Even
though throughput and latency are frequently the most meaningful
metrics for an application, they are not workload-independent and
therefore are inadequate for measuring the utilization of a device
across all workloads, unless one accepts worst-case estimates that
can lead to gross underutilization of resources (e.g. the throughput
of random and sequential workloads in spinning media differ by
orders of magnitudes). A common workload-independent metric is
\emph{time utilization}, i.e. the amount of time a resource is
utilized in a given time interval. Time utilization has the advantage
of having an always defined maximum and therefore makes resources
fully reservable. Even though a workload-independent metric might
at first not appear meaningful for an application, given a particular
workload an application can discover the relationship between its
workload-dependent metric and the workload-independent metric.
Because of performance isolation, the application has to discover
this relationship only once.

\paragraph{Related Work:}

\paragraph{Challenges:} The deep and heterogenous memory and storage
hierarchy we are assuming for this proposal complicates the
relationship between workload-dependent and workload-independent
metrics: the performance of a task can significantly differ depending
on what levels of the hierarchy are involved. Furthermore, an
important goal of the proposed project is to enable applications
to reason about the trade-off between resolution and latency, adding
yet another dimension to how tasks are mapped to resources.

\begin{itemized}

\item A key challenge of admission control is scalability: the
decision of whether to admit a task could potentially depend on
global knowledge of the current utilization of every single resource
in the system. Scalability therefore depends on whether admission
control is able to accurately map a new task to a relatively small
set of resources which can quickly provide up-to-date utilization.
One approach might involve pseudo-random mapping that also load-balances
as a side-effect.

\item To offer latency/latency trade-offs, the system must be able
to quickly generate a number of data production plans, involving
different parts of the hierarchy and different data resolutions.
It will then use the utilization model to estimate the latency for
each production plan and the resolutions they can provide. Here
again, pseudo-random selection could reduce the number of resources
that would be involved, thereby increasing scalability.

\item The complexity of a utilization model involving the entire
storage hieararchy with 100,000s of devices is potentially daunting.
However, the utilization model can be simplified by modeling classes
of devices as well as classes of requests. In particular, each
device could restrict access to its content via a set of well-defined
methods with known, absolute performance properties.

\item Flash devices become inherently unpredictable when reads and
writes are mixed on the same device because of garbage collection.
While write latency can be easily hidden using asynchronous I/O,
hiding read latency is more difficult. By separating reads and
writes for each device, read latency becomes predictable. The
challenge is to minimize duplication of data, at least on fast
layers and leverage redundancy across layers.

\end{itemized}

### 2015-06-30 narrative/placement.tex
Initial draft.

@@
 
-Gary / Manish
+The application workflows targeted in this proposal generate very large amounts of data, and this data must be processed and analyzed before potential insights from the simulations can be realized.
+In-situ/in-transit data processing using in-memory data-staging approaches have been effectively used to address data-related challenges and enable simulations workflows on current high end systems -- the key idea is to use compute and storage resources on the high-end system to process the data close to where it is generated.
+However, given current architectural trends, indicate that emerging systems will have increasing numbers of cores per node and correspondingly decreasing amounts of DDRAM memory per core as well as decreasing memory bandwidth.
+Architectural trends also indicate deeper memory hierarchies and multilevel storage structures.
+As a results, to address the increases data volumes and data processing requirements requires that these levels of storage be effectively utilized and exposes as a uniform staging abstraction that can be used by the application workflows.
+A key requirements is application-driven runtime mechanisms for dynamically managing data placement across the layers of the distributed storage hierarchy, coordinating data movement and data sharing between the components of the application workflow so as to maximize its utility to the application and reduce access costs.
+
+\paragraph{State of the art:} NAND flash-based devices are being increasingly used in HPC storage systems at different levels and for a variety of purposes.
+Several studies such as \cite{multitier}, \cite{sc10li} have investigated using compute node local SSDs as storage buffers, for example, to temporarily cache checkpoint data to support recovery from failures.
+Research efforts have also explored using deep memory devices for data analysis.
+Active Flash \cite{activeflash} proposes in-situ scientific data analysis by directly executing data analysis tasks on emerging storage devices.
+Minerva \cite{minerva} extends the conventional SSD architecture using a FPGA-based storage controller to offload data or I/O intensive application code to the SSD to accelerate data analysis.
+However, these solutions cannot deal with coupled application workflows, where the data exchange and access patterns can be both complex and dynamic.
+Furthermore, some of these solutions require hardware modifications or special access priorities to the HPC systems, which can limit their use to specific resources.
+
+\paragraph{Proposed research approach:} In this project we will utilizes deep memory hierarchies and multilevel storage structures to support staging-based in-transit data management and processing for data-intensive scientific workflows at extreme scale.
+Specifically, we will address the coordination and data access/sharing requirements of application workflows, as well as the  online data management requirements such as for example, in-transit online data analysis,  check-pointing,  data reduction.
+We will also deal with the complexities associated with data placement of data and movement in the multiple layers and components of the storage hierarchy.
+The key components of our research include:
+
+{\em Application hints:}  Application hints can provide additional information to the runtime based on the users' knowledge of the application workflow as well as on past experiences, for example, possible data access patterns, nature of regions of interest, and domain specific information.
+This application level information will be added to the specification of the workflow and can provide advisory inputs to efficiently help the runtime make smarter decisions about the placement and/or movement of data objects and the scheduling and execution of tasks, as well as to effectively manage trade-offs between for example, power consumption and time-to-solution.
+
+{\em Data utility:} We will formulate the concept of utility (from an application perspective) that is assigned data objects to quantify the benefits of placement and/or movement actions to application.
+These utilities may be based on the importance of the data objects, the frequency with which they are accessed, etc., and are used during runtime data management.
+For example, data objects will higher utility value may be placed closer (e.g., in faster memory) to the accessing application.
+Similarly, when it is time to evict data objects from a higher level storage layer, the decision about which object to evict may take the utility into account.
+
+{\em Autonomic in-transit data management:} Efficiently placing data objects vertically across multiple memory hierarchy layers and storage levels as well as horizontally across different storage nodes requires balancing competing objectives.
+In this project we will explore autonomic data management strategies that can  evaluate utility/cost tradeoffs and appropriately place/move data objects at runtime.
+The autonomic approach will leverage both user hints and information gathered at  runtime (e.g., runtime data access history, network topology, etc.) to place data object.
+
+{\em Runtime tracking and estimation:}
+Autonomic data placement and/or movement will leverage the information gathered at runtime together with predictive analysis requirements.
+This information may include runtime spatial-temporal data access patterns, physical network topology information, and etc., which will be tracked and gathered at runtime, and then be used to estimate and predict the behavior of coupled applications in the rest of simulation-time.
+A key research aspect here will be to anticipate accesses, for example using learning techniques or using application knowledge, and use this information to prefetch data objects.
+For example, we will identify and catalogue key phases and access patterns in the target workflows and then use it to characterize patterns at runtime and trigger appropriate place/movement actions.
+
+\paragraph{Preliminary work and results:} Our recent work \cite{tongipdps15} explored a two-tiered staging method that spans both DRAM and solid state disks (SSD).
+It allows us to support both code coupling and data management for data intensive simulation workflows in a loosely coupled manner.
+In addition, our application-aware adaptive data placement has demonstrated the effectiveness of using user provided access information as ``hints'', and runtime data access pattern history to improve data access efficiency and overall end-to-end execution.
+Besides, additional data placement optimization \cite{qiansc15} that leverages physical network topology along with data access patterns has also demonstrated lower data access costs and good scalability.
+Our work have been deployed and proven with real applications, e.g., coupled combustion simulations (DNS-LES) that are part of the ExaCT Exascale Co-design Center and coupled Fusion simulations (XGC0-XGCa) that are part of the EPSI SciDAC on current high-end systems such Titan at ORNL.

### 2015-06-30 narrative/intro.tex

@@
 One of the key aspects of our approach is to allow the users to define the
-utility (importance) of their data; in other-words, for each data object, users
-need to come up with a way to define pieces of the object which have more
-utility (more importance) . There are many clear research challenges faced with
+utility of their data; in other-words, the system should provide a
+mechanism such that, for each data object, a user can specify pieces of the object
+which have more utility (more importance).
+Clearly there are many research challenges faced with
 this approach including: 1) How do we make this simple enough so that users can
-do this, 2) can we come up with a model to allow the system to then make the
-autonomic decisions to save the data which has the highest utility, and 3) can
-we allow the back-end storage system to interact with this definition so that
-it can optimize not just over one application but across all of the
+do this, 2) Can we define a model to allow the system to make
+autonomic decisions to save the data which has the highest utility, and 3) Can
+we allow the back-end storage system to interact with this model so that
+it can optimize not just a single application but across all of the
 applications.
@@
 some of the objects will be broken up into multiple pieces by ``plug-ins''
-to the system which will allow data to be characterized in terms of the most
-important pieces, and the next level of details. We can think of this as
+to the system which will allow data to be characterized first in terms of the most
+important pieces, and then by subsequent levels of detail. We can think of this as
 something similar to an Adaptive Mesh Refinement scheme which keeps track of
-the places where there are perhaps steep gradients (large errors on the
-coarser views), and smaller errors. These then allows us to make decisions
+the places where there are features such as steep gradients, which contain large errors on the
+coarser views, as opposed to other areas with smaller errors. These distinctions allow us to make decisions
 about which sub-objects to save (all of the data, which will then go to
@@
 \label{subsec:storage-perspective}
-From the perspective of the storage system, it needs to manage the data from
-a given user request amongst all of the request and try to optimize the
-entire system. Today the problem is that when a user writes there is a
+From the perspective of the storage system, it is necessary to manage the data from
+a given user request amongst all of the requests from all users, and try to optimize the
+entire system accordingly. Today the problem is that when a user writes there is a
 tremendous amount of interference from other users on the system who may be

### 2015-06-30 narrative/data-desc.tex
START THE SECTION

\subsubsection{Description of Data}

One of the key advantages for a storage system to communicate through objects rather than 
bytes is that it gives a meaning and relationship of certain bytes placed together to form one object.
This permits the system to understand that all of the data in an object needs to be organized together,
and that commonly the entire object will be written and read. Data Models then constitute a collection
of objects which are then placed together via a relationship. In other words, we can think of a
simulation which contains a three dimensional mesh, and that the coordinates of the mesh, along
with the connectivity of the mesh will generally be accessed together. This also us to organize
the data in the storage system in such a way that will allow a user to easily to describe this 
collection of data together, and it will allow the user to describe relationships between the objects
in the data model.

The challenge which all users face today, and will face much more on exascale systems is that 
the users must prioritize which objects to save, and how often (in a simulation which produces more
data in time)  to save the information. As this data gets large,  users realize that these objects will
eventually move from a fast storage into either a slower form of storage, such as tape, or it will be
purged from the system.  We propose a new technique, described by a utility curve, where objects
can be cast into multiple buckets of data, and these buckets can be prioritized with one-another.
Once the data is re-organized, users will be able to use an API to describe the importance of this data.
This importance will allow users to specify: 1) the relationships of different objects with one another,
2) the expected lifetime of the object.

The relationship of the objects to one another can be done by

\paragraph{State of the art:} 
\paragraph{Proposed research approach:} 
We will first create an API where users can describe the utility based on their expectations of how
long they want each piece to live in the different storage areas
\paragraph{Preliminary work and results:}

### 2015-06-30 narrative/qos-app.tex
initial text on application level qos

\subsubsection{QOS in the Application}
%proposed idea
A key insight in this proposed project has been the increased interaction of
the application with the storage system. Towards this end, we propose to
address the quality of service requirements by porivding the application
with mechanisms to specify the quality of I/O service, interrogate the
storage system, and react to the responses. Guided by our past work in
ADIOS\cite{lofstead2008flexible} we propose to explore the design of two
mechanisms for this purpose. Firstly, we will explore the augmentation of
I/O applictation programing interfaces (I/O APIs) to allow applications to
both specify timing and quality information and also query the storage
system for timing estimates. Through this interface we expect the
application and user to gain insights into how long a single output or input
call will take given the required quality information, and then react to
these estimates by adjusting the quality or restricting the scope of the
data required. Likewise we will explore how an application can provide
timing information to the storage system to allow the storage system to make
optimization decisions to best meet the requirements from the application. 
%
Secondly, we will also explore external data annotations, such as those
provided by the configuration file in ADIOS. Through the use of these
external augmentation the user can provide insights to the storage system on
the relative value of the data, expected life time and performance
characteristics, as well as relationships between different data sets. With
this information the storage system can make optimizations specific to a use
case. We expect these augmentations to be particularly important for
eviction of data from a storage layer, and migration of data sets to a
different storage layer.

%execution

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../proposal"
%%% End:

### 2015-06-30 narrative/qos-app.tex
some more text on execution

@@
 %execution
+We will design an updated I/O API that will expand on the current POSIX I/O
+semantics by adding new information to the I/O request. We will explore the
+set of application level hints that can be easily provided to the storage
+system in order for the storage system to make the most appropriate
+optimization decisions. In particular, we expect the application to provide
+hints on the length of time before the output data is persistent and
+visible to users, and the expected lifetime of the data for data output
+calls. For read calls we expect the set of hints to combine latency and
+precision requirements. Additional hints will also be investigated.
+
+We will also develop a new type of querying system that allows application
+to ask the storage system about completion timing information, the response
+being an estimate provided by the storage system. We will explore how these
+query functions can be integrated into common applications with minimal code
+disruptions, and the set of policies that applications can use to respond to
+this new information. We will also explore how applications can query the
+storage system to gain insights on available space in different tiers and
+how applications can tailor their output data to match the available space.
+
+Finally we will study the use of an external data annotation system that can
+provide information to the storage system without requiring the application
+to be recompiled.
 

### 2015-06-30 narrative/qos-app.tex
initial challenges

@@
 how applications can tailor their output data to match the available space. 
-
 Finally we will study the use of an external data annotation system that can
@@
 to be recompiled.
+
+\paragraph{Challenges}
+The design of new APIs posses both technical and adoption challenges. Here
+we will only consider the technical challenges. The proposed set of APIs aim
+to expose broad system level environmental information to the application,
+and allow the application to adapt dynamically based on this information.
+One aspect of this challenge is the design of an interrogative API that
+enables a back-and-forth between the application and the storage system to
+enable the application to recieve enough information to make adaptation
+decisions.
 

### 2015-06-30 narrative/data-desc.tex

@@
 
-One of the key advantages for a storage system to communicate through objects rather than 
-bytes is that it gives a meaning and relationship of certain bytes placed together to form one object.
-This permits the system to understand that all of the data in an object needs to be organized together,
-and that commonly the entire object will be written and read. Data Models then constitute a collection
+One of the key advantages of representing data as \textit{objects}, rather than 
+bytes, is that it allows data semantics, user intentions, QoS requirements, and
+the relationship between user data to be readily captured and embedded with the 
+raw data. This description of data, as enabled by managing data as objects, bridges 
+the semantics gap between applictions, middleware, and storage systems,
+and allows the system to understand user-level data, execute QoS requirements, and
+optimize application or system performance.
+
+Data Models then constitute a collection
 of objects which are then placed together via a relationship. In other words, we can think of a

### 2015-06-30 narrative/intro.tex

@@
 system must also automatically move data from one tier of the storage to
-another, without effecting the quality of service that it guarantees at the
-time of questioning. The storage system must also look at ways to organize
-data amongst all of the tiers of storage in one consistent way.
+another, without affecting the quality of service that was offered at request
+time. The storage system must be able to organize
+data amongst all of the tiers of storage in a known, consistent way.
 
@@
 storage system for as long as specified. Clearly we need to maintain 
-QoS such that fuzzy-estimates of the time it will take to write/read data from
+QoS such that fuzzy estimates of the time it will take to write/read data from
 the different layers informs any migration/eviction decisions, and clearly it

### 2015-06-30 narrative/data-reorg.tex
start of the data-reorganization,. I will let Mark iterate a bit and continue tomorrow morning on this.

@@
 \subsubsection{Data Reorganization Techniques}
+One of the key element to get efficient data access during reading is the re-organization of the
+data to ensure that the data that will be read back will involve minimal seeks on the file system 
+along with aggregating similar data together to get optimal performance and placing the data
+across multiple disks to get high levels of concurrency during reading. We will create APIs to
+allow users to group similar data together in a data-model, and annotations (initially done in the
+ADIOS layer) to allow relationships of variables to be expressed. We will then investigate both
+system (general) data-re-organization techniques based on the utility of data and the relationships
+of the data, and application-specific routines (for our initial set of evaluation test cases).  Once
+we can generalize the application specific routines to a set of motifs, we can then allow this piece
+of code to propagate to the general routines. In all cases we will investigate re-organzing the data
+into an integer (most likely 8) bins, where they fall into the most likely accessed data to the least
+likely accessed data.
 
-Mark, Scott
+\paragraph{State of the art:}
+\paragraph{Proposed research approach:}
+General methods to re-organize the data fall into the following categories
+\begin{itemize}
+\item Precision based re-organization. This is where the most significant bytes of the data are
+all group together from each object. This data generally will have a higher utility then the data with
+the least significant bytes.
+
+\item Frequency-based re-organization. FFT, Wavelets.
+
+\item Histogram binning. Re-organize the data through histograms to ensure we have enough
+phase-space coverage. Where we have finner coverage we can move that data to other
+quantiles of the storage.
+
+\item Multi-resolution. Traditionally we can look at every other point, and have a good feeling that 
+this data will give a good estimate of the data.
+
+\item Region of Interest. Using error techniques similar to AMR, we can understand where there
+is data changing more rapidly and save this data at a higher resolution version than the smoother
+regions of space.
+{\bf Mark: can you fill these out much more and talk about the new techniques we will work on}.
+\end{itemize}
+\paragraph{Preliminary work and results:}

### 2015-06-30 narrative/intro.tex
start of the data-reorganization,. I will let Mark iterate a bit and continue tomorrow morning on this.

@@
 important pieces of the data.
-
+\subsection{Research Challenges to be address}
+\label{subsec:challenges}
 Our research into the storage system and middle ware layers must address

### 2015-07-01 narrative/eviction-metrics.tex
modified:   narrative/eviction-metrics.tex 	modified:   proposal.tex

\subsubsection{Eviction Metrics}

% The title might have to change...

There is a spectrum of approaches on how to leverage a memory and storage hierarchy: On one end of the spectrum is to copy or move data from one tier to another, the closer the tier to the hierarchy's top, the faster the access to the data it contains. On the other end of the spectrum is to use higher tiers exclusively for auxiliary data that to some extent represents the actual data on the lowest tier. Examples of auxiliary data are views, indices, lossy compressions, lower resolution data, and summary data (sometimes also known as metadata). There are a number of drivers that push

### 2015-07-01 narrative/eviction-metrics.tex
modified:   eviction-metrics.tex

@@
 
-There is a spectrum of approaches on how to leverage a memory and storage hierarchy: On one end of the spectrum is to copy or move data from one tier to another, the closer the tier to the hierarchy's top, the faster the access to the data it contains.
-On the other end of the spectrum is to use higher tiers exclusively for auxiliary data that to some extent represents the actual data on the lowest tier.
-Examples of auxiliary data are views, indices, lossy compressions, lower resolution data, and summary data (sometimes also known as metadata).
-There are a number of drivers that push
+\paragraph{Background:} There is a spectrum of approaches on how
+to leverage a memory and storage hierarchy: On one end of the
+spectrum is to copy or move data from one tier to another, the
+closer the tier to the hierarchy's top, the faster the access to
+the data it contains. On the other end of the spectrum is to use
+higher tiers exclusively for auxiliary data that to some extent
+represents the actual data on the lowest tier. Examples of auxiliary
+data are views, indices, lossy compressions, lower resolution data,
+and summary data (sometimes also known as metadata). There are a
+number of drivers that push approaches to one or the other extreme:
+(1) if a high tier has sufficient room for an application's working
+set, approaches that move the working set into that tier will perform
+well; Conversely, if a high tier very rarely captures working sets,
+it is better to fill it with auxiliary data that speeds up access
+to the working set captured by a lower tier; working sets only
+reduce capacity misses but the concept can be generalized to the
+likelihood of having the right data in place, i.e. proactively
+moving data to high tiers accessed in predictable patterns reduces
+compulsory misses as well; (2) if the application generates parallel
+read/write requests and strong consistency semantics are important,
+approaches that use auxiliary data to speed up access to persistent
+shared data might yield better performance than approaches that
+require coherence overhead to keep all copies on higher tiers
+consistent; (3) if the application's access patterns are mismatched
+with low tier access characteristics, a combination of approaches
+that use auxiliary data and move data between tiers in order to
+convert access patterns into a better match will perform well.
+
+% Any other drivers?
+
+\paragraph{Approach:} This spectrum enables interesting strategies
+to manage space pressure on high tiers: instead of just evicting
+pieces of data (blocks, pages, objects, files), one can free space
+by ``thinning'' data so that accesses within more limited views or
+lower resolutions can still occur without misses while requests
+outside of views or higher resolutions can leverage the information
+stored in auxiliary data. Conversely, data on a high tier can be
+``enriched'' (e.g. turned into a higher resolution) if there is
+value in doing so.
+
+\paragraph{Related Work:} Mention Stanford's Legion system.
+
+\paragraph{Challenge:} The decision on when to thin and when to
+enrich data on a particular level depends on \emph{access patterns}.
+Approaches that leveragee access patterns in application at national
+labs has shown to greatly improve performance and reduce the amount
+of overhead required for data management~\cite{he:hpdc13}. The
+challenge is for applications to communicate these access patterns
+to the I/O stack layers that manages tiers outside of the address
+space. In the proposed work we will develop an abstraction for
+\emph{locality} that allow applications or runtime profilers to
+describe locality properties of access patterns, and implement
+services that leverage locality properties to dynamically determine
+the value of thinning or enriching data at a particular tier.
+Consider the example of a road navigation system: the locality
+properties of each navigation client are based on physical constraints
+of speed, direction, likely resolution of map, and the fact that
+clients almost never go offroad.

### 2015-07-01 narrative/data-desc.tex

--- 2015-06-30 narrative/data-desc.tex
@@
 \subsubsection{Description of Data}
+In order for the middleware and storage layer to understand data,
+I/O intentions, and act in a smart way, description of data is needed,
+which is achieved by managing data as \textit{objects} in this project.
+An object is the smallest unit of data that consists of raw data and metadata,
+and a user space variable, such as \textit{temperature}, may consist of a collection
+of objects possibly with different accuracy or resolution.
+A key advantage of representing data as \textit{objects}, rather than 
+bytes, is that it allows data semantics, user intentions, QoS requirements, and
+the relationship between user data to be readily captured and embedded with the 
+raw data. As an example, retrieving a 3D field generated from a simulation and visualizing it
+requires the coordinates and connectivity data 1) to be accessed together 2) with low
+latency. This description of data, as enabled by managing data as objects, bridges 
+the semantics gap between applictions, middleware, and storage systems,
+and allows the system to understand user-level data, execute QoS requirements and policies, and
+optimize application and system performance.
 
-One of the key advantages for a storage system to communicate through objects rather than 
-bytes is that it gives a meaning and relationship of certain bytes placed together to form one object.
-This permits the system to understand that all of the data in an object needs to be organized together,
-and that commonly the entire object will be written and read. Data Models then constitute a collection
+Data models then constitute a collection
 of objects which are then placed together via a relationship. In other words, we can think of a

### 2015-07-01 narrative/infrastructure.tex
Added feiyi's text and graphics

There are two fundamental causes that can impede the I/O performance in an extreme-scale SSIO system. One is indirection, where layers of storage tiers are employed either for performance and/or for scalability reasons.  The direct consequence is that there could be multiple traversing paths from application end to the rest place of data, and more often than not, the I/O paths are not under control under any single authority. The other cause is the shared use of resources. The best effort I/O request/response nature and lack of QoS mechanisms imply that there is little guarantee in terms of expected performance.  Both indirection and shared use of resource contribute to a high probability of imbalanced use of resources thus the occurrence of congestion and degraded performance. Traditionally, there is a disconnection between storage infrastructure and rest of system software and applications: the infrastructure details are not exposed anywhere at all. One argument favoring this is to ensure a platform agnostic design.

As an example to demonstrate that this disconnection will hurt system performance: We launch 4096 processes with each process doing a single file I/O operation against half of the Spider II file system. The traces of those files are analyzed to examine the utilization distribution of different components. Figure (a), (b) and (c) shows the resource usage distribution for OSTs, OSSes, and LNETs, respectively. We observe that there exists a significant variation in usage across components of any given type (e.g., OST, OSS or LNET). For example, some OSTs are used more than 10 times while some others are never used (corresponding to zero frequency count). Similarly, OSSes and LNETs show significant imbalance in usage under the default placement strategy. Consequently, imbalanced resource utilization increases the contention at certain components more than others.

Given these insights, we advocate the idea that the infrastructure knowledge should find a way to relay to the upper layer for better and more effective use of resources.   We think this is more pertinent and critical given the recent development of multi-tier storage and storage component heterogeneity.  There needs to be a way for application/middleware layer to gain more exposure of storage system for more intelligent processing logic.  One prime example of such exposed knowledge can be request/response time. To most applications, this is a black box. Profiling it at the upper layer is neither efficient nor effective, as it doesn't reflect cross-layer characteristics. However, Most of storage layer does keep a detailed profiling of such information. We therefore envision and propose a histogram-based request/response profiling API that application and middleware layer can leverage and make more informative decisions.

\begin{figure}[tbh]
  \centering
  \includegraphics[width=\columnwidth]{graphics/infrastructure.pdf}\vspace{-1.2in}
  \caption{Resource usage distribution for OST(a), OSS(b) and LNETs(c). }
\end{figure}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../proposal"
%%% End:

### 2015-07-01 narrative/data-reduction.tex

\subsubsection{Data Reduction Techniques}

{\em The purpose of computing is insight, not numbers}~\cite{Hamming:book}.
Richard Hamming's remark, made over 50 years ago at the dawning of the age of
large scale scientific computation, is even more relevant today as we prepare
for scientific simulation at the exascale. Hamming was a pioneer in the fields
of coding theory, scientific computing and information theory who would
instantly sympathize with the central objective of this proposal--minimizing
the time to knowledge.

The tradiional approach to transforming results of scientific computation into
knowledge consists of storing the numerical output of a simulation on disk,
followed by a post-processing or analysis stage that might take place over
several ensuing weeks or months. The overarching objective is to gain insight,
or knowledge.  As already mentioned, the anticipated scale of computations
means that this workflow is simply no longer viable.

It is useful to classify scientific data into two basic categories: structured
data (in the sense that it satisfies a known or relatively simple model) or
unstructured data (in which information content follows no obvious or explicit
model).  Although scientific data generally contains random components (due to
finite precision, measurement and calibration effects etc.), scientific data is
never purely random. Broadly speaking, the path from data to knowledge consists
of extracting underlying models or patterns from the datasets, and interpreting
the resulting models.

Ideally, the scientist would perform the analysis stage {\em in situ} to
extract the relevant information and, in the unlikely event of this being
possible, would effectively circumvent the large data issue completely.  The
problem of course is that this is unlikely to be possible since, by their
nature, large scale simulations aim to discover new information that is often
hidden as higher order effects amongst the data deluge. In particular, this
means if data thinning or truncation is applied then one is in danger of
effectively throwing out the baby with the bath water. One cannot store the
entire data set in easily accessible source, due to sheer size, yet one cannot
reduce it prior to archival without risking losing the information one is
trying to extract.

Viewed in this way, the problem would appear intractable. However, the above
discussion fails to recognize that much of the data is redundant in the
information theoretic sense.  That is to say, the amount of information
contained in the data is often significantly less than the amount of data.  The
difficulty stems from the fact that one does not know about this redundancy in
advance.

We propose a fresh approach to tackle the data overload problem head-on through
using preprocessors in which information is not simply stored in the form of
raw data, but rather in the form of appropriate parameterized algorithms. A
priori information on the nature of the data is used to choose the basic
algorithm while the actual values of the data are used to tune the parameters
used in the reconstruction procedures. We shall refer to this paradigm as an
Auditor.

Scientific data is sometimes described as essentially incompressible owing to
noise and other random effects present in the data, stemming from the numerical
algorithm, software implementation, system specific behavior etc. The reasoning
behind the description lies in the mathematical fact that it is impossible to
achieve lossless compression on completely random data sets \cite{Gray:book}. Even
recent state of the art lossless compression approaches \cite{GomezCappello} are only
able to obtain 15\% improvement in lossless compression ratio while reducing
the time spent in compression of floating point data by one half in some cases.
Nevertheless, these limitations apply only if one ignores two key features:
Firstly, randomness in scientific data generally occurs only below some
threshold related to the accuracy of the measurement apparatus or the numerical
algorithm. Secondly, lossy compression is perfectly acceptable provided the
loss level falls below this threshold.

Data compression broadly consists of three basic steps: (1) Pre-processing to
extract as much structure as feasible. This step is variously referred to as
filtering, preconditioning, and pre- processing but in all cases, the objective
is to represent the data in a form that becomes more amenable to compression.
(2) Removal of redundancy such as omission of duplicates etc.  Finally, (3) the
resulting data is compressed using entropy coding. The level of compression
that can be achieved in the final step is limited by the Shannon information
content of the data \cite{Gray:book}, and many standard tools are available for
efficient and perfect entropy coding. The potential for compression through
removal of redundancy is generally minor. The main scope for achieving
significant compression (a compression rate of two or better say) therefore
lies in the pre-processing step.

Research in the construction of pre-processing techniques is still in its
infancy, especially in context of exascale simulations. Techniques that have
been advocated range from simple difference filtering, bit-sorting, prefix
transformation through to simply reordering the data.  What these, and other
catchall methods, fail to exploit is a priori information on the nature and
provenance of the data.

As we have already observed, the incorporation of a priori information of this
type is vital if one is to obtain significant data reduction. The use of an
Auditor as a pre-processor exploits the wide (and increasing) cost gap between
CPU operations versus storage-based operations. The preprocessor would ideally
be employed in situ and is used in a similar fashion to pre-processors in
standard compression approaches such as FPC \cite{BurtscherFPC} in which an XOR
operation against the previous data is used as a pre-processor. The fact that
even such a crude pre- processor is able to deliver reasonable results augurs
well: Even a relatively crude Auditor can be expected to significantly
out-perform the XOR and other similar approaches.

### 2015-07-01 narrative/naming-service.tex
Lots more details. Need to resolve refs still.

@@
 A traditional POSIX naming service offers a hierarchical space consisting of
-directories and files. LWFS and Sirocco have abandonded this as the default
-model in favor of a container/object/fork/addrress tuple for identifying data.
-LWFS demonstrated a POSIX-style namespace on the side kept in sync using a
-transaction process like D2T~\cite{d2t} showing that this alternative approach
-can support traditional POSIX API calls even though the underlying storage
-system uses a different model.
+directories and files. This structure has led to scalability problems because
+of the serialized access to a single source for creating and accessing files.
+Several efforts~\cite{giga+,pvfs,others} have worked to reduce this contention
+by doing things like reducing the serialized scope to a single directory or
+subtree. While these approaches help, they do not address this key scalability
+limitation.
 
-For this project, we seek to offer a similar POSIX-style namespace for naming
-objects, but store a rich metadata collection with each entry to handle data
-migration, discovery, and writing and retrevial time estimations. Further, we
-will also have to track data stored in different tiers, even at the variable
-subset level. This additional metadata load will require a more complex naming
-service that reacts to automatic data movement while caching current data
-locations accelerating data access operations avoiding any data discovery
-operations.
+Pure object stores, such as those popular in the big data
+domain~\cite{memcached,others} avoid this bottleneck by strictly offering an
+object ID with the application required to manage how this ID maps to something
+of interest. This approach of removing the metadata service from the system
+level completely can work well for scale out applications where data is created
+or consumed by a single process at a time rather than potentially O(1 million)
+processes all actively collectively for a single ``object''. To address this
+case, having some system integrated metadata services to associate names with
+these object is a preferable solution.
 
-In addition to the basic naming operations, we will also need to incorporate
-authorization capabilities. Sirocco currently integrates with a Kerberos
-service for authentication and authorization. Given a capability ticket, a
-user can access different objects as needed.
+LWFS and Sirocco have take an approach similar to the pure object stores, but
+with a focus on the HPC setting. They have abandonded a fully POSIX compliant
+metadata service as the default model in favor of a
+container/object/fork/addrress tuple for identifying data similar to those
+used for pure object stores. By having a service that addresses the object
+collections that comprise a single thing, such as a variable or timestep
+output, just enough metadata is maintained to make the storage system usable
+without additional heavy lifting by clients.  LWFS demonstrated a POSIX-style
+namespace on the side kept in sync using a transaction process like
+D2T~\cite{d2t} showing that this alternative approach can support traditional
+POSIX API calls even though the underlying storage system uses a different
+model.
+
+The challenge this proposal brings is in the additional metadata required to
+track all of the different pieces of a single variable across the storage
+hierarchy both horizontally and vertically and keeping this metadata in sync
+as the storage system (Sirocco) manages data according to user requests and
+system load. Further, the profiled performance for annotating data according to
+future read needs also must be tracked. Overall, since the end-to-end quality
+of service requirements must drive all data placement and movement decisions,
+the metadata and naming service must be rich enough, and low overhead enough,
+to support these operations as well as maintaining data accessibility.
+
+In addition to the basic naming and data tracking operations, we will also need
+to incorporate authorization capabilities. Sirocco currently integrates with a
+Kerberos service for authentication and authorization. Given a capability
+ticket, a user can access different objects as needed. This ticket structure
+offers protection services typically offered on POSIX directories and files,
+but can do it at the fork level instead. This allows a reduced quality data
+version to be available for the general users while the high-quality version
+would be limited to the data creator. This and other considerations for
+security must be incorporated into the entire naming and metadat service.
 
 The main challenge of incorporating the additional, rich metadata will be
-joined by the additional challenge of coordinating with the other storage
-and application-layer services to offer the best access times possible for data
-stored in the system.
+joined by the additional challenge of coordinating with the other storage and
+application-layer services to offer the best access times possible for data
+stored in the system. The developed metadata services that drive data
+compression and subsetting operations must have easy, consistent, and ideally
+non-blocking or locking access to this metadata service. We must investigate
+how to build such a metadata and naming service that also incorporates and
+maintains the additional metadata required to support our advanced
+functionality.

### 2015-07-01 narrative/data-desc.tex

@@
 In order for the middleware and storage layer to understand data,
-I/O intentions, and act in a smart way, description of data is needed,
-which is achieved by managing data as \textit{objects} in this project.
+I/O intentions, and act accordingly, describing data in a way that can
+be understood by the system is needed. In this project, this is achieved 
+by managing data as \textit{objects}.
 An object is the smallest unit of data that consists of raw data and metadata,
@@
 the relationship between user data to be readily captured and embedded with the 
-raw data. As an example, retrieving a 3D field generated from a simulation and visualizing it
-requires the coordinates and connectivity data 1) to be accessed together 2) with low
-latency. This description of data, as enabled by managing data as objects, bridges 
+raw data. As an example, retrieving a 3D field data generated from a simulation and then visualizing it
+requires the coordinates and connectivity variable 1) to be accessed together and 2) with low
+latency in order to achieve a good user experience. The description of data, 
+as enabled by managing data as objects, bridges 
 the semantics gap between applictions, middleware, and storage systems,
@@
 
-Data models then constitute a collection
-of objects which are then placed together via a relationship. In other words, we can think of a
-simulation which contains a three dimensional mesh, and that the coordinates of the mesh, along
-with the connectivity of the mesh will generally be accessed together. This also us to organize
-the data in the storage system in such a way that will allow a user to easily to describe this 
-collection of data together, and it will allow the user to describe relationships between the objects
-in the data model.
+The description of data is essentially metadata and it may include conventional attributes, such as 
+data type, size, dimensionality, relationship to other data, and etc. In this project, a key metadata attribute
+that enables QoS scheduling, data placement, and ???, is \textit{data utility}, which
+captures the priority of objects - it is to our belief that
+users must prioritize a small set of objects to be saved on 
+higher-level capacity-limited storage layer in exascale
+to avoid the slow access to large-capacity lower-level storage, such as tapes.
+We propose a new technique, described by a utility function, where objects
+can be cast into multiple buckets of data, and each bucket can be prioritized differently according
+to its utility value.
+Once the data is re-organized, users will be able to use API (section ??)
+to describe the importance of this data.
 
-The challenge which all users face today, and will face much more on exascale systems is that 
-the users must prioritize which objects to save, and how often (in a simulation which produces more
-data in time)  to save the information. As this data gets large,  users realize that these objects will
-eventually move from a fast storage into either a slower form of storage, such as tape, or it will be
-purged from the system.  We propose a new technique, described by a utility curve, where objects
-can be cast into multiple buckets of data, and these buckets can be prioritized with one-another.
-Once the data is re-organized, users will be able to use an API to describe the importance of this data.
-This importance will allow users to specify: 1) the relationships of different objects with one another,
-2) the expected lifetime of the object.
+\paragraph{State of the art:} ADIOS implements a binary-packed data format that allows
+data characteristics such as min, max and index to be wrapped around data chunks. A
+directy benefit is that each data chunk can be operated independently and I/O concurrenty
+can be maximized. We will build upon this capability and further augment the format
+to incorporate data utility.
 
-The relationship of the objects to one another can be done by
-
-\paragraph{State of the art:} 
 \paragraph{Proposed research approach:} 
-We will first create an API where users can describe the utility based on their expectations of how
-long they want each piece to live in the different storage areas
-\paragraph{Preliminary work and results:}
+We will design and deveop new APIs through which users can describe the utility of data
+based on their expectations.

### 2015-07-01 narrative/data-desc.tex
more edits on data description.

@@
 The description of data is essentially metadata and it may include conventional attributes, such as 
-data type, size, dimensionality, relationship to other data, and etc. In this project, a key metadata attribute
-that enables QoS scheduling, data placement, and ???, is \textit{data utility}, which
+data type, size, dimensionality, relationship to other data, and etc. In particular,
+relationship to other data objects can be captured by implementing \textit{soft links}
+and different types of relationships can be expressed by different types of soft links. 
+In this project, a key metadata attribute
+that enables QoS scheduling, data placement, and ???, is \textit{data utility value}, which
 captures the priority of objects - it is to our belief that
-users must prioritize a small set of objects to be saved on 
-higher-level capacity-limited storage layer in exascale
+for exascale science, users must prioritize a small set of objects to be saved on 
+higher-level capacity-limited storage layer
 to avoid the slow access to large-capacity lower-level storage, such as tapes.
-We propose a new technique, described by a utility function, where objects
+We propose a new technique, described by a utility function plugged in by a user, where objects
 can be cast into multiple buckets of data, and each bucket can be prioritized differently according
 to its utility value.
-Once the data is re-organized, users will be able to use API (section ??)
-to describe the importance of this data.
+Once this is done, the middleware system will construct the description of data, re-organize
+and place data to achieve the desired QoS goals and policies.
 
@@
 data characteristics such as min, max and index to be wrapped around data chunks. A
-directy benefit is that each data chunk can be operated independently and I/O concurrenty
+direct benefit is that each data chunk can be operated independently and I/O concurrenty
 can be maximized. We will build upon this capability and further augment the format
-to incorporate data utility.
+to express data utility. Recently DAMSEL \cite{damsel} provides a rich metadata representation
+and management layer that captures the relationship between data blocks for scientific applications.
+This allows application data to be mapped to storage system efficiently, without overburdening
+users to managing complex data model, such as AMR, from the user space. However, the aspect of 
+data importantance hasn't been explored, which we believe will be important for exascale storage solutions.
 

### 2015-07-01 narrative/data-reorg.tex
fixed the includes to work with emacs

@@
 we can generalize the application specific routines to a set of motifs, we can then allow this piece
-of code to propagate to the general routines. In all cases we will investigate re-organzing the data
+of code to propagate to the general routines. In all cases we will investigate re-organizing the data
 into an integer (most likely 8) bins, where they fall into the most likely accessed data to the least
-likely accessed data.
+likely accessed data. These methods will NOT be used for checkpoint restart data, since this
+data needs to be accessed exactly how it is originally in memory.  We generally see that these
+techniques are highly useful for analysis and  visualization data, where users can inherently place
+their intentions of the data (life time of the data for eviction policies, prioritization of data 
+importance).
 
 \paragraph{State of the art:}
+
 \paragraph{Proposed research approach:}
@@
 all group together from each object. This data generally will have a higher utility then the data with
-the least significant bytes.
+the least significant bytes.  This process has the following three steps which need to be
+efficiently implemented when data gets placed to the storage system and from the storage 
+system. 
+%
+First, the data needs to be re-arranged which involves memory operations and needs
+to be done in-situ. This operation involves no communication.  Data needs to be copied, so
+the memory requirements are increased. At most this requires two copies an individual dataset
+in memory, and we will investigate techniques to allocate and deallocate this memory if the user
+will specify that the data will be overwritten after it is written to the storage system. This is often
+the case for many of the data quantities written form the simulation, but there are many cases
+where, for example, we want to write all of the particles from a Particle In Cell (PIC) simulation. 
+Since the particles will be used later in the calculation, we need to duplicate the storage. Our
+observation with working with the XGC1, GTC, Warp, and PiconGPU simulations (all leadership
+class simulations) is that we can temporary increase the storage of the particles, and then 
+release them since the temporary arrays used in the calculations are often freed when a PIC
+iteration is finished.. 
+%
+The second step to give the metadata that allows the system to understand that an object is now contained in multiple bins of data. One of the challenges is the ability to bring together multiple bins
+of data together (the high priority data along with the medium priority data) efficiently when the 
+data is being read.  One research question is how we can manage the metadata efficiently.
+We propose to place this knowledge into the middleware which will communicate this to the 
+storage system.  This will give the storage system potential ways to increase the concurrency 
+when accessing this information. For example, we can place the highest priority bits of an object
+on the fastest storage system, and the next level of importance in the campaign storage. If a 
+user request precision to a level where we need to access both, concurrently we can be
+accessing this data, and moving the data to memory where it can be re-arranged. 
+%
+Since data will be placed in different bins along the hierarchy we will need to understand how
+to initially place this in the fastest storage, and then evict this from the fastest storage without
+effecting the performance of the simulation which is generating this data. Policies must be 
+placed to ensure that we do not create any internal interference during these operations. We
+will investigate doing this with ``meta-bots'' first introduced in the LWFS project and have 
+these bots move the data.{\bf probably belongs in a different section}
 
 \item Frequency-based re-organization. FFT, Wavelets.
+Another common technique to classify the importance of data is to first re-organize the data
+in frequency space and then specify that as the frequency increases the importance of data
+decreases. This allows us to place the lowest frequency pieces in the fastest storage, and
+the highest frequency either on the slowest storage tiers, or if the data sizes are prohibitively  
+costly, not even write out the highest frequency. This binning of data in frequency space can
+take as an input (storage bandwidth knowledge, storage size limitations, and user intentions
+of how long the majority of the data will be frequently accessed.
+
+These operations also require data transformations (for example an FFT and an inverse FFT)
+and then re-arranging the data in real-space once we re-organize this. The research questions
+we must address are: 1) what is the cost of re-organzing the data, 2) how do we
 
@@
 \paragraph{Preliminary work and results:}
+Text here
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "../proposal"
+%%% End:

### 2015-07-01 narrative/data-desc.tex
edits on description of data.

@@
 We will design and deveop new APIs through which users can describe the utility of data
-based on their expectations.
+based on their expectations. This can be done by plugging user level functions that will
+partition user level data into objects and be executed by the middleware to calculate the 
+utility value for each object. The utility value along with other metadata will be encapsulated
+by the middleware and utilized by the underlying storage system to management data placement.

### 2015-07-01 narrative/data-reorg.tex
minor changes.

@@
 in frequency space and then specify that as the frequency increases the importance of data
-decreases. This allows us to place the lowest frequency pieces in the fastest storage, and
+decreases.   Common technique which is used for streaming data, and data reduction techniques such as JPEG-2000~\cite{jpeg2000}
+is to  code streams which have regions of interest that offer several mechanisms to support spatial random access or region of interest access at varying degrees of granularity.
+It is possible to store different parts of the same data using different quality.
+
+This allows us to place the lowest frequency pieces in the fastest storage, and
 the highest frequency either on the slowest storage tiers, or if the data sizes are prohibitively  
@@
 
-\item Multi-resolution. Traditionally we can look at every other point, and have a good feeling that 
-this data will give a good estimate of the data.
+\item Multi-resolution.  
+A
 

### 2015-07-01 narrative/intro.tex
minor changes.
put in a use case form CS Chang

@@
-\section*{Introduction} {\color{red} Scott, Hasan, Gary, Jay - 4 pages}
+\section{Introduction} {\color{red} Scott, Hasan, Gary, Jay - 4 pages}
 \label{sec:introduction}
@@
 system. 
-The negotation between these layers will be provided as a fundamental service, 
+The negotiation between these layers will be provided as a fundamental service, 
 allowing users to easily save ``the best'' amount of data, automatically placed
@@
 greatly impact productivity.
+%
+The overarching idea is that data from applications can be characterized by different
+levels of importance and different level of importance over time. 
+Codes such as the XGC1 simulation from C.
+S.
+Chang (PPPL), which is one of largest users of Leadership Class Facilities (over 300M hours at ANL, NERSC, and ORNL) has launched a series of simulations which need to write out 100 PB of data on the Titan system.
+Due to resource limitations, the simulation will only write about 10 PB, but
+when our team considered not only the time it would take to write out 10 PB in 10 days of run, but also where we would store this, which had to be tape since it would take
+over half of the file system, we decided that we couldn't write this amount of data. In other words we had to make decisions of which data to save from the original 100 PB.
+The first thing that we had to discover is how much space we could have on the parallel file system at any given time, and how much time it would take to archive 1 PB of data,
+which turned out to be 10 days, and we could only store 2 PB of data at once. This quickly led us to 1) decide which pieces of data to save from the large amount of data; in 
+other words, we had to prioritize which were the most important pieces. We had to further breakdown the large data into two categories, data which we think we will touch after
+the simulation is complete ( after the 10 days of wallclock time was finished), and what we would do with this data (the access pattern of the large data). We then had to 
+discover application specific data reduction techniques, and encode that into our solution, on top of ADIOS. We then had to implement a crude discovery system to understand
+which pieces of data were stored on the parallel file system and which pieces moved over to PPPL, and which pieces were on tape.  This solution worked well for one day
+but after the second day more problems occurred, the variability of the time it took to write the data made the I/O prohibitively costly, because even though we had the
+majority of the OLCF, the file system was shared and sometimes the writing time took less 100 seconds, and at other times it was over 1,000 seconds. We had no method
+to monitor this time and then write less data when the time was going to be longer.   After the simulation is complete we also need to visualize this reduced data, and 
+an additional problem appeared; just to get a first look at the visualization data seemed like it could take over a month because some of the timesteps were on tape and other
+timesteps were on the file system. We had no method to just look at the data quickly and understand which regions of the simulation to focus on.
 

### 2015-07-02 narrative/naming-service.tex
merged naming and discovery

@@
-\subsubsection{Naming Service}
+\subsubsection{Naming and Discovery Service}
 
-A traditional POSIX naming service offers a hierarchical space consisting of
-directories and files. This structure has led to scalability problems because
-of the serialized access to a single source for creating and accessing files.
-Several efforts~\cite{giga+,pvfs,others} have worked to reduce this contention
-by doing things like reducing the serialized scope to a single directory or
-subtree. While these approaches help, they do not address this key scalability
-limitation.
+Given the well documented difficulties with POSIX-compliant naming and our
+potentially heavily decomposed objects, we must take a new approach to naming.
+Adding complexity is data migration based on system pressures. Our naming
+service must support sufficient information to discover or track all variables
+pieces across the entire storage hierarchy.
 
-Pure object stores, such as those popular in the big data
+Hierarchical Storage Management (HSM) systems offer a strct caching approach
+for managing different storage capacities trading off performance for capacity.
+By maintaining a single namespace across all tiers, it is possible to list a
+single directory view with files stored at different tiers. While this approach
+to managing multi-tier storage works for whole files that fit on single logial
+devices, it is far from ideal for scientific simulations.
+
+Our goal with this proposal is to offer a similar capability, but use a finer
+granularity and also manage discovering where the data may have migrated to
+based on system pressures. Instead of a single file such as might be used to
+store an entire timestep output for a simulation, we will demonstrate offering
+the same capability at a subset of a single variable level. By shifting to a
+finer-grained approach, we will enable more effective use of close/fast/small
+storage tiers and manage the impact when there is insufficient space forcing
+data to move. Traiditional HSM stores an entire file on a tier making room if
+insufficient space is available. With this shift to a partial variable
+granularity, a 1 PB output with 500 GB of ``high interest'' data can limit this
+costly tier ussage to just 500 GB greatly enhancing usability. This reduces
+pressure forcing data migration, but does not prevent it. We still need to be
+able to find the data should it move based on system pressure.
+
+The resulting naming and discovery service will attach the Sirocco
+contaier/object/fork/address tuple with a special attribute indicating that
+this record is of ``high interest'' strongly encouraging keeping the variable
+in close/fast/small storage as long as possible. In addition to maintaining
+this additional metadata, the application and/or middleware must provide this
+information. The storage system simply needs to act appropriately on this data
+based on the attribute.
+
+The challenge for discovery is that potentially, data will migrate from where
+it is initially stored to a new location within the storage system. Sirocco
+offers an ability to search for data that has moved as well as forcing a
+particular resilience-based replica be the ``authoritative'' version. We will
+investigate if the current Sirocco functionality is capable of supporting the
+new operating modes we wish to offer. Initial expectations suggest having
+bounded time guarantees for finding data are critical for offering the quality
+of service guarantees we wish to offer. Some potential approaches can build on
+CRUSH~\cite{ceph} from Ceph to offer a map of where to search for data
+sequentially. Given our multiple storage tiers, we would extend this using some
+search limiting mechanism indicating to shift to the next tier to continue
+searching because the requested data never made it to this location. This new
+work will be an expansion of Sirocco's currently planned features.
+
+The overriding theme for this proposal of providing a cooperative relationship
+between the user and the storage system extends into this area as well. In
+order for the user to negotiate with the storage system for data
+quality/retrieval time tradeoff, additional data is required. Providing this
+functionality requries the naming service be able to both return a list of data
+versions, but also some indication of data quality and estimated retrieval time
+for each. This estimate will be disucssed elsewhere in this proposal.
+
+Structurally, the traditional POSIX naming service offeriing a hierarchical
+space consisting of directories and files may be maintained for backwards
+compatibility. However, this view will not offer the same strict semantics
+POSIX requires. We must break these strict semantics to address scalability
+problems because of the serialized access to a single source for creating and
+accessing files.  Several efforts~\cite{giga+,pvfs,others} have worked to
+reduce this contention by doing things like reducing the serialized scope to a
+single directory or subtree. While these approaches help, they do not address
+this key scalability limitation. Instead, we ar willing to offer a short
+duration to eventually consistent POSIX view to offer the performance and
+consistency requirements an exascale application demands.
+
+While pure object stores, such as those popular in the big data
 domain~\cite{memcached,others} avoid this bottleneck by strictly offering an
@@
 case, having some system integrated metadata services to associate names with
-these object is a preferable solution.
+these object is a preferable solution. While it will not have the same
+synchronous consistency semantics as POSIX, we will offer something as close as
+possible to address legacy application and command-line style data access and
+migration to other storage types using these POSIX semantics.
 
-LWFS and Sirocco have take an approach similar to the pure object stores, but
-with a focus on the HPC setting. They have abandonded a fully POSIX compliant
-metadata service as the default model in favor of a
-container/object/fork/addrress tuple for identifying data similar to those
-used for pure object stores. By having a service that addresses the object
+We plan to address this scalability problem by continuing with the LWFS and
+Sirocco model. LWFS and Sirocco have take an approach similar to the pure
+object stores, but with a focus on the HPC setting. They have abandonded a
+fully POSIX compliant metadata service as the default model in favor of a
+container/object/fork/addrress tuple for identifying data similar to those used
+for pure object stores. By having a service that addresses the object
 collections that comprise a single thing, such as a variable or timestep
@@
 POSIX API calls even though the underlying storage system uses a different
-model.
-
-The challenge this proposal brings is in the additional metadata required to
-track all of the different pieces of a single variable across the storage
-hierarchy both horizontally and vertically and keeping this metadata in sync
-as the storage system (Sirocco) manages data according to user requests and
-system load. Further, the profiled performance for annotating data according to
-future read needs also must be tracked. Overall, since the end-to-end quality
-of service requirements must drive all data placement and movement decisions,
-the metadata and naming service must be rich enough, and low overhead enough,
-to support these operations as well as maintaining data accessibility.
+model. Because we are not requiring the entire storage space be addressible
+from a single tree root, we can offer multiple independent roots using the
+Sirocco object storage. We will investigate how to make the naming and
+discovery service scalable under the circumstances outlined above.
 
@@
 
+While strongly encouraging data to stay in a particular type of storage tier,
+ultimately it must be possible to flush this data to make room for other uses.
+Sirocco offers a ``flush'' operation that forcibly migrates data to the
+long-term resilience requirements freeing space in the scare resources. In
+many, if not most cases, this will move data towards or actually on to a device
+like tape intended for long-term storage.
+
 The main challenge of incorporating the additional, rich metadata will be
@@
 functionality.
+
+\begin{enumerate}
+
+\item Milestone 1: Demonstrate a metadata service capable of serving both POSIX
+clients and our clients, but without consideration for authentication, authorization, or data migration.
+
+\item Milestone 2: Demonstrate a time bounded search approach for finding data
+within the storage system to identify data the current location and retrieval
+performance characteristics for data since migration. This will update the
+cached information in the metadata service to short-circuit future requests for
+the data presuming that it does not migrate again.
+
+\item Milestone 3: Demonstrate incorporating A\&A and show scalability under
+both application load as well as storage system pressures showing that we can
+tolerate both loads and maintain quality of service.
+
+\end{enumerate}

### 2015-07-02 narrative/data-description.tex

\subsection{Description of Data}
In order for the middleware and storage layer to understand data,
I/O intentions, and act accordingly, describing data in a way that can
be understood by the system is needed. In this project, this is achieved 
by managing data as \textit{objects}.
An object is the smallest unit of data that consists of raw data and metadata,
and a user space variable, such as \textit{temperature}, may consist of a collection
of objects possibly with different accuracy or resolution.
A key advantage of representing data as \textit{objects}, rather than 
bytes, is that it allows data semantics, user intentions, QoS requirements, and
the relationship between user data to be readily captured and embedded with the 
raw data. As an example, retrieving a 3D field data generated from a simulation and then visualizing it
requires the coordinates and connectivity variable 1) to be accessed together and 2) with low
latency in order to achieve a good user experience. The description of data, 
as enabled by managing data as objects, bridges 
the semantics gap between applications, middleware, and storage systems,
and allows the system to understand user-level data, execute QoS requirements and policies, and
optimize application and system performance.

The description of data is essentially metadata and it may include conventional attributes, such as 
data type, size, dimensionality, relationship to other data, and etc. In particular,
relationship to other data objects can be captured by implementing \textit{soft links}
and different types of relationships can be expressed by different types of soft links. 
In this project, a key metadata attribute
that enables QoS scheduling, data placement, and ???, is \textit{data utility value}, which
captures the priority of objects - it is to our belief that
for exascale science, users must prioritize a small set of objects to be saved on 
higher-level capacity-limited storage layer
to avoid the slow access to large-capacity lower-level storage, such as tapes.
We propose a new technique, described by a utility function plugged in by a user, where objects
can be cast into multiple buckets of data, and each bucket can be prioritized differently according
to its utility value.
Once this is done, the middleware system will construct the description of data, re-organize
and place data to achieve the desired QoS goals and policies.

The user would like the ability to get a certain amount of Quality of Service 
such that they can then make decisions when the expected bandwidth, 
for example is less than what they desire and will then place in a certain set 
of rules which the system can make autonomic decisions to help decide what should be done.
A key insight in this proposed project has been the increased interaction of
the application with the storage system. Towards this end, we propose to
address the quality of service requirements by proivding the application
with mechanisms to specify the quality of I/O service, interrogate the
storage system, and react to the responses. Guided by our past work in
ADIOS \cite{lofstead2008flexible} we propose to explore the design of two
mechanisms for this purpose. Firstly, we will explore the augmentation of
I/O applicatition programing interfaces (I/O APIs) to allow applications to
both specify timing and quality information and also query the storage
system for timing estimates. Through this interface we expect the
application and user to gain insights into how long a single output or input
call will take given the required quality information, and then react to
these estimates by adjusting the quality or restricting the scope of the
data required. Likewise we will explore how an application can provide
timing information to the storage system to allow the storage system to make
optimization decisions to best meet the requirements from the application.

We will also explore external data annotations, such as those
provided by the configuration file in ADIOS. Through the use of these
external augmentation the user can provide insights to the storage system on
the relative value of the data, expected life time and performance
characteristics, as well as relationships between different data sets. With
this information the storage system can make optimizations specific to a use
case. We expect these augmentations to be particularly important for
eviction of data from a storage layer, and migration of data sets to a
different storage layer.

\paragraph{State of the art:} ADIOS implements a binary-packed data format that allows
data characteristics such as min, max and index to be wrapped around data chunks. A
direct benefit is that each data chunk can be operated independently and I/O concurrently
can be maximized. We will build upon this capability and further augment the format
to express data utility. Recently DAMSEL \cite{damsel} provides a rich metadata representation
and management layer that captures the relationship between data blocks for scientific applications.
This allows application data to be mapped to storage system efficiently, without overburdening
users to managing complex data model, such as AMR, from the user space. However, the aspect of 
data importance hasn't been explored, which we believe will be important for exascale storage solutions.

\paragraph{Proposed research approach:} 
We will design and develop new APIs through which users can describe the utility of data
based on their expectations. This can be done by plugging user level functions that will
partition user level data into objects and be executed by the middleware to calculate the 
utility value for each object. The utility value along with other metadata will be encapsulated
by the middleware and utilized by the underlying storage system to management data placement.

We will design an updated I/O API that will expand on the current POSIX I/O
semantics by adding new information to the I/O request. We will explore the
set of application level hints that can be easily provided to the storage
system in order for the storage system to make the most appropriate
optimization decisions. In particular, we expect the application to provide
hints on the length of time before the output data is persistent and
visible to users, and the expected lifetime of the data for data output
calls. For read calls we expect the set of hints to combine latency and
precision requirements. Additional hints will also be investigated.

We will also develop a new type of querying system that allows application
to ask the storage system about completion timing information, the response
being an estimate provided by the storage system. We will explore how these
query functions can be integrated into common applications with minimal code
disruptions, and the set of policies that applications can use to respond to
this new information. We will also explore how applications can query the
storage system to gain insights on available space in different tiers and
how applications can tailor their output data to match the available space. 
Finally we will study the use of an external data annotation system that can
provide information to the storage system without requiring the application
to be recompiled.

\paragraph{Challenges}
The design of new APIs posses both technical and adoption challenges. Here
we will only consider the technical challenges. The proposed set of APIs aim
to expose broad system level environmental information to the application,
and allow the application to adapt dynamically based on this information.
One aspect of this challenge is the design of an interrogative API that
enables a back-and-forth between the application and the storage system to
enable the application to receive enough information to make adaptation
decisions.

### 2015-07-02 narrative/qos-storage.tex
modified:   proposed-work.tex 	deleted:    qos-ss.tex 	renamed:    ../qos-storage.tex -> qos-storage.tex
modified:   qos-storage.tex
modified:   ../bib/maltzahn.bib 	modified:   infrastructure.tex 	deleted:    policies.tex 	modified:   proposed-work.tex 	modified:   qos-storage.tex 	deleted:    time-estimation.tex

--- 2015-06-29 qos-storage.tex
@@
-\paragraph{Background:} ``Quality of Service'' (QoS) refers to the
-properties of the performance qualtiy of a particular service.
+\subsubsection{QoS}
+
+\paragraph{Background:} One of the key differentiating features this project offers is driving storage
+decisions through quality of service (QoS) \emph{reservations} rather than continuing the current model of allowing
+all application to compete, without restriction, for part or all of the storage
+resources. This traditional approach leads to intereference
+effects~\cite{lofstead:2010:adaptive} that can greatly impact IO performance
+predictability. Further, with the introduction of additional tiers in the
+storage stack, the performance variability will increase with greater
+competiion for limited high performance resources. Unfettered, this competition
+will yield lower performance for all users than if applications attempt to reserve performance, thereby enabling the storage system to reason over its performance in terms of reservations and actual performance, accepting reservations that are within resource availability, rejecting those that would lead to overload.
+
+It is important to note that we use performance reservations to reduce interference and to avoid overload conditions.
+If the system uses reservations intelligently, its performance is predictable and efficient because resources are not wasted on interference.
+However, in cases where an appication arrives that needs to run more urgently than any of the other already running applications, administrators can suspend less urgent applications which will save their state and relinquish their performance reservations.
+Upon restart, these applications make new performance reservations.
+
+There are two fundamental causes that can impede the I/O performance in an extreme-scale SSIO system.
+One is indirection, where layers of storage tiers are employed either for performance and/or for scalability reasons.
+The direct consequence is that there could be multiple traversing paths from application end to the rest place of data, and more often than not, the I/O paths are not under control under any single authority.
+The other cause is the shared use of resources.
+The best effort I/O request/response nature and lack of QoS mechanisms imply that there is little guarantee in terms of expected performance.
+Both indirection and shared use of resource contribute to a high probability of imbalanced use of resources thus the occurrence of congestion and degraded performance.
+Traditionally, there is a disconnection between storage infrastructure and rest of system software and applications: the infrastructure details are not exposed anywhere at all.
+One argument favoring this is to ensure a platform agnostic design.
+
+As an example to demonstrate that this disconnection will hurt system performance: We launch 4096 processes with each process doing a single file I/O operation against half of the Spider II file system.
+The traces of those files are analyzed to examine the utilization distribution of different components.
+Figure (a), (b) and (c) shows the resource usage distribution for OSTs, OSSes, and LNETs, respectively.
+We observe that there exists a significant variation in usage across components of any given type (e.g., OST, OSS or LNET).
+For example, some OSTs are used more than 10 times while some others are never used (corresponding to zero frequency count).
+Similarly, OSSes and LNETs show significant imbalance in usage under the default placement strategy.
+Consequently, imbalanced resource utilization increases the contention at certain components more than others.
+
+Given these insights, we advocate the idea that the infrastructure knowledge should find a way to relay to the upper layer for better and more effective use of resources.
+We think this is more pertinent and critical given the recent development of multi-tier storage and storage component heterogeneity.
+There needs to be a way for application/middleware layer to gain more exposure of storage system for more intelligent processing logic.
+One prime example of such exposed knowledge can be request/response time.
+To most applications, this is a black box.
+Profiling it at the upper layer is neither efficient nor effective, as it doesn't reflect cross-layer characteristics.
+However, Most of storage layer does keep a detailed profiling of such information.
+We therefore envision and propose a histogram-based request/response profiling API that application and middleware layer can leverage and make more informative decisions.
+
+\begin{figure}[tbh]
+  \centering
+  \includegraphics[width=\columnwidth]{graphics/infrastructure.pdf}\vspace{-1.2in}
+  \caption{Resource usage distribution for OST(a), OSS(b) and LNETs(c). }
+\end{figure}
+
+% This is about finding the relationshiop of workload-dependent to workload-independent metrics. -- Carlos
+ %are driven from multiple different approaches. First, supporting
+%capability runs is a priority for capability machines. We propose to address
+%capability run performance by incorporating monitoring for prepartory runs at
+%smaller scale to profile the application output characteristics both from a
+%writing velocity and volume, but also for example read patterns for the data
+%analytics required to generate scientific insights from the raw data.
+
+% Not sure how to reuse this ... -- Carlos
+%By
+%discovering approximate data proportions that will likely be the key subsets
+%targeted by the simulation run, we can generate a policy to preserve a certain
+%storage portion for high fidelity data storagae and use slower or perhaps lower
+%fidelity or compressed data storage for less interesting data portions. We will
+%support spreading data appropriately across the storage hierarchy. In
+%particular, we will investigate how to place data across different layers to
+%meet the performance requirements for output while maintaining system
+%availability for other applications and offering the best possible performance
+%for the data analytics that will ultimately process this data.
+
+% Not sure how to reuse this... -- Carlos
+%Second, once data has been placed for future performance requirements, policies
+%must be able to address maintaining data on a priority basis in that location
+%to deliver on the placement optimization generated by the capability run
+%profilier.
+
+% This is a matter of metering, not performance reservations. -- Carlos
+%Third, most future NVM devices have limited write endurance prompting
+%management to ensure fair use by all machine users. Technologies like
+%NAND-flash and Phase Change Memory have limited write endurance. By
+%incorporating policies about the proportion of write endurance a compute
+%allocation is entitled to, data placement decisions can be made to ensure fair
+%resource usage. While the offending application will suffer worse IO
+%performance, other applications that more carefully address their data
+%intensity on these limited devices will achieve higher overall performnce. Only
+%by instituting such policies can we encourage application users to adopt
+%technology to manage the limited resources.
+
+%We propose to investigate both offering a policy mechanism and how to implement
+%these kinds of policies in a way that addresses offering the shortest
+%end-to-end time for scientific insights.
+
+``Quality of Service'' (QoS) refers to the
+properties of the performance qualtiy of a particular service, in this case the storage system.
 Performance quality can be expressed in either relative or absolute
-terms. Examples of relative performance qualities are ``fair'',
+terms. Examples of relative performance terms are ``fair'',
 ``proportional'', or ``priority/class-based'' (``the higher the
 priority the better the service'' or ``1st class is better than
-2nd''). The key advantage of relative performance qualities is that
+2nd''). The key advantage of relative performance terms is that
 they are easy to implement. The key disadvantage is the difficulty
-of making any strong guarantees based on those implementations --
+of making any guarantees based on those implementations --
 even relative guarantees are mired by the well-known effect of
 priority inversion~\cite{lampson:cacm80}. Examples of absolute
-performance qualities are ``rate-based'', ``soft real-time'', and
+performance terms are ``rate-based'', ``soft real-time'', and
 ``hard real-time''. The key advantage of absolute performance
-qualities is that it is principally easier to implement strong
-guarantees: they enable an agreement between a task and a service,
-also known as a \emph{reservation}, does not change meaning depending
-on the service's state, as long as the service is able to isolate
-performance between different tasks. The key disadvantage of absolute
-performance qualities is that their implementation is non-trivial,
+terms are that they enable the implementation of strong
+guarantees: absolute terms enable agreements between application tasks and storage that we call \emph{reservations} and that do not change meaning depending
+on the storage system's workload. The key challenge of absolute
+performance terms is that their implementation is non-trivial,
 especially in large-scale systems.
@@
 
+\subsubsection{QoS}
+
 \paragraph{Related Work:}
@@
 
-\begin{itemized}
+\begin{itemize}
 
@@
 
-\end{itemized}
+\end{itemize}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "../proposal"
+%%% End:

### 2015-07-02 narrative/naming-service.tex
saved some work. TBC.

@@
 Given the well documented difficulties with POSIX-compliant naming and our
-potentially heavily decomposed objects, we must take a new approach to naming.
-Adding complexity is data migration based on system pressures. Our naming
-service must support sufficient information to discover or track all variables
-pieces across the entire storage hierarchy.
+approach that decomposes data into potentially large number of objects, 
+we must take a new approach to naming.
+Data migration, as a result of system pressures or user/system policies,
+further adds to the complexity of managing naming and discovery. Our naming
+service must maintain sufficient metadata information, track objects
+as they move across storage layers, and facilitates data discovery.
 
-Hierarchical Storage Management (HSM) systems offer a strct caching approach
+Hierarchical Storage Management (HSM) systems \cite{hsm} offer a strict caching approach
 for managing different storage capacities trading off performance for capacity.
@@
 
-Our goal with this proposal is to offer a similar capability, but use a finer
-granularity and also manage discovering where the data may have migrated to
+Our goal with this proposal is to offer a similar capability but with a finer
+granularity leveraging the inner structure of scientific data, 
+and manage discovering efficiently, where the data may
+be split and placed on several storage layers, or may have migrated
 based on system pressures. Instead of a single file such as might be used to

### 2015-07-02 narrative/intro.tex
use case
modified:   bib/maltzahn.bib 	modified:   narrative/intro.tex

@@
 We plan to leverage the many inovations we discovered in 
-the Lightweight File Systems~\cite{lwfs} (LWFS) project.
-In this project, Sandia sought to
+the Lightweight File Systems~\cite{lwfs} (LWFS) project and the Ceph storage system project (ceph.com).
+In the LWFS project, Sandia sought to
 demonstrate stripping down a file system to the bare minimum features required
@@
 has been in process for three years and has developed a usable storage infrastructure.
+
+The Ceph project (ceph.com) started at UC Santa Cruz in 2004 and
+was initially funded by DOE/NNSA involving LLNL, LANL, and Sandia,
+with the goal to create an object-based parallel file system that
+addresses the well-known metadata service bottleneck of creating
+and accessing 1000s of files comprising a checkpoint~\cite{weil:osdi06}.
+However, after the Ceph kernel client was pulled into the Linux
+mainline in 2010, the focus of Ceph shifted to its object-based
+storage subsystem, also called Reliable Autonomous Distributed
+Object Store (RADOS), which now is the dominant storage system in
+OpenStack production-level deployments and was recently selected
+as the new backend for Flickr and Yahoo! Cloud Services. In this
+project we will leverage two of the key innovations of Ceph: (1)
+Controlled Replication Under Scalable Hashing (CRUSH)~\cite{weil:sc06}
+will help making resource identification and predictable performance
+scalable, and (2) extensible storage object classes~\cite{watkins:ucsctr15}
+will provide the main mechanisms of introducing new APIs,
+metadata vocabulary, and quality of service into the storage stack.
 

### 2015-07-02 narrative/data-reduction.tex
start to reshape the proposal

@@
-\subsubsection{Data Reduction Techniques}
-
-{\em The purpose of computing is insight, not numbers}~\cite{Hamming:book}.
-Richard Hamming's remark, made over 50 years ago at the dawning of the age of
-large scale scientific computation, is even more relevant today as we prepare
-for scientific simulation at the exascale. Hamming was a pioneer in the fields
-of coding theory, scientific computing and information theory who would
-instantly sympathize with the central objective of this proposal--minimizing
-the time to knowledge.
-
-The tradiional approach to transforming results of scientific computation into
-knowledge consists of storing the numerical output of a simulation on disk,
-followed by a post-processing or analysis stage that might take place over
-several ensuing weeks or months. The overarching objective is to gain insight,
-or knowledge.  As already mentioned, the anticipated scale of computations
-means that this workflow is simply no longer viable.
+\subsection{Data Refactoring}
 

### 2015-07-02 narrative/data-reorg.tex
start to reshape the proposal

@@
-\subsubsection{Data Reorganization Techniques}
+%\subsubsection{Data Reorganization Techniques}
 One of the key element to get efficient data access during reading is the re-organization of the

### 2015-07-02 narrative/intro.tex
start to reshape the proposal

@@
-\section{Introduction} {\color{red} Scott, Hasan, Gary, Jay - 4 pages}
+\section{Introduction} 
 \label{sec:introduction}
@@
 greatly impact productivity.
+%
+{\em The purpose of computing is insight, not numbers}~\cite{Hamming:book}.
+Richard Hamming's remark, made over 50 years ago at the dawning of the age of
+large scale scientific computation, is even more relevant today as we prepare
+for scientific simulation at the exascale. Hamming was a pioneer in the fields
+of coding theory, scientific computing and information theory who would
+instantly sympathize with the central objective of this proposal--minimizing
+the time to knowledge.  Our approach in this proposal is to enable researchers
+to place more knowledge into the SSIO system so that data can be recast from a bunch of bytes
+to information which the system can then place and optimize the time to gain knowledge from exascale simulations.
+We first motivate this by a problem our team has been working on in the past six months, and then recast this problem into
+a series of research questions and challenges which must be addressed in order to optimize the SSIO system for exascale science.
+
+The traditional approach to transforming results of scientific computation into
+knowledge consists of storing the numerical output of a simulation on disk,
+followed by a post-processing or analysis stage that might take place over
+several ensuing weeks or months. The overarching objective is to gain insight,
+or knowledge.  As already mentioned, the anticipated scale of computations
+means that this workflow is simply no longer viable.
 %
@@
 important pieces of the data.
-\subsection{Research Challenges to be address}
-\label{subsec:challenges}
-Our research into the storage system and middle ware layers must address
-many of these challenges including
-\begin{enumerate}
-\item How do we describe the user intentions at the API layer and have this
-  communicated down to the middle-ware and storage layers?
-\item How do we allow users to define their user defined compression
-  techniques and have this data read back from the storage system layers?
-  Where do we execute the decompression techniques 
-\item How do we evaluate the tradeoffs at runtime to guide data placement,
-  including the movement of data across the network, across the different
-  storage tiers? (data movement and quality tradeoff)
-\item Can we use different forms of learning techniques as daemons on the
-  system to perform data migration from one storage tier to another? For
-  example, if we see that a user is looking at one time slice after another
-  for a certain object in their dataset which is in the slower tiers of the
-  hierarchy, do we automatically propane up the data which has not been
-  requested in the hope that this will be requested data? (prediction and
-  prefetching, user defined compression on movement)
-\item Can we understand the true need of campaign storage and understand the
-  possible impact of the campaign storage if we run a hadoop-like
-  file-system instead of a lustre/gpfs file system? (additional layers)
-\item Can we limit the amount of data duplication? Data space is going to be
-  very constrained on the exascale system so we must ensure that minimal
-  copies are made of data, and when data is duplicated, they are removed
-  through a garbage collection routine. (space management)
-\item Can we build a model to give us the time estimates during the reading
-  and writing phases, so that rules can be in place from the user
-  perspective to make adaptive decisions. For example, if the file system
-  says that reading will take 3 months, users can they place in rules to
-  then read in a subset of data, and they can understand which data can be
-  read quickly and which pieces will take more time. (estimation)
-\item Can we understand how to place code in the system which will allow
-  data-regeneration to take place. (regeneration)
-\item How does the storage system do a better job in managing request from
-  all of the users on a LCF than today? We realize that today users who have
-  a better middlware system can often lock other users from getting high
-  performance when they are running. If we have the concept of currency
-  which is eventually used in the same extent as node-hours, then users will
-  have to be able to think about how much storage and how much bandwidth
-  they can choose. One question that needs to be understood is if there are
-  times when the system sees that there are very few storage system
-  resources being used, then the lucky users can get the bandwidth cheaper
-  than at times of heavy usage. How can we enforce this? (Fairness)
-\item How do we manage all of the metdata not only from the principle
-  objects, but from the sub-objects? How well will this scale when we have
-  users who can potentially create billions of objects from their
-  simulation? (scalability and discovery)
-\end{enumerate}
+
 %

### 2015-07-02 narrative/proposed-work.tex
Adding some sections
adding more sections
Added feiyi's text and graphics
fixed the includes to work with emacs
minor changes.
merge data description, qos app and API
merged naming and discovery
modified:   ../bib/maltzahn.bib 	modified:   infrastructure.tex 	deleted:    policies.tex 	modified:   proposed-work.tex 	modified:   qos-storage.tex 	deleted:    time-estimation.tex
start to reshape the proposal

@@
-\section*{Proposed Work}
-
-\begin{itemize}
+\section{Proposed Research and Methods}
+\label{subsec:challenges}
+Our research into the storage system and middle ware layers must address
+many of these challenges including
+\begin{enumerate}
+\item How do we describe the user intentions at the API layer and have this
+  communicated down to the middle-ware and storage layers?
+\item How do we allow users to define their user defined compression
+  techniques and have this data read back from the storage system layers?
+  Where do we execute the decompression techniques 
+\item How do we evaluate the tradeoffs at runtime to guide data placement,
+  including the movement of data across the network, across the different
+  storage tiers? (data movement and quality tradeoff)
+\item Can we use different forms of learning techniques as daemons on the
+  system to perform data migration from one storage tier to another? For
+  example, if we see that a user is looking at one time slice after another
+  for a certain object in their dataset which is in the slower tiers of the
+  hierarchy, do we automatically propane up the data which has not been
+  requested in the hope that this will be requested data? (prediction and
+  prefetching, user defined compression on movement)
+\item Can we understand the true need of campaign storage and understand the
+  possible impact of the campaign storage if we run a hadoop-like
+  file-system instead of a lustre/gpfs file system? (additional layers)
+\item Can we limit the amount of data duplication? Data space is going to be
+  very constrained on the exascale system so we must ensure that minimal
+  copies are made of data, and when data is duplicated, they are removed
+  through a garbage collection routine. (space management)
+\item Can we build a model to give us the time estimates during the reading
+  and writing phases, so that rules can be in place from the user
+  perspective to make adaptive decisions. For example, if the file system
+  says that reading will take 3 months, users can they place in rules to
+  then read in a subset of data, and they can understand which data can be
+  read quickly and which pieces will take more time. (estimation)
+\item Can we understand how to place code in the system which will allow
+  data-regeneration to take place. (regeneration)
+\item How does the storage system do a better job in managing request from
+  all of the users on a LCF than today? We realize that today users who have
+  a better middlware system can often lock other users from getting high
+  performance when they are running. If we have the concept of currency
+  which is eventually used in the same extent as node-hours, then users will
+  have to be able to think about how much storage and how much bandwidth
+  they can choose. One question that needs to be understood is if there are
+  times when the system sees that there are very few storage system
+  resources being used, then the lucky users can get the bandwidth cheaper
+  than at times of heavy usage. How can we enforce this? (Fairness)
+\item How do we manage all of the metdata not only from the principle
+  objects, but from the sub-objects? How well will this scale when we have
+  users who can potentially create billions of objects from their
+  simulation? (scalability and discovery)
 
@@
 \item Develop profiling system to determine how a data set is used during preparation runs prior to a capability run to determine how to optimize data placement for capability run analytics.
-\item
 
-\end{itemize}
+\end{enumerate}
+
+\input{narrative/data-description} % includes data-desc, apis, and qos-app.
+
+%\subsection{Application Interface}
+
+%\input{narrative/data-desc}
+\input{narrative/data-reduction} % includes data-reorg and the api specs move to the above section -
+\input{narrative/data-reorg}
+%\input{narrative/qos-app}
+\input{narrative/placement}
+%\input{narrative/policies}
+%\input{narrative/apis}
+
+%\subsection{Storage System}
+
+\input{narrative/migration}
+\input{narrative/resource-mgmt}
+\input{narrative/qos-storage}
+%\input{narrative/discovery}
+\input{narrative/eviction-metrics}
+\input{narrative/naming-service}
+%\input{narrative/time-estimation}
+\input{narrative/infrastructure}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "../proposal"
+%%% End:

### 2015-07-02 narrative/enabling_tech.tex
minor changes.
move some of the Cepth and ADIOS part into the enabling technologies, and try to re-organize to show the focus of this proposal is on theme 2.

@@
-\section*{Enabling Technologies}
-
+\section{Enabling Technologies}
+\label{sec:enabling-technologies}
 Significant existing work will serve as the foundation upon which this project
@@
 combination, we will be able to deliver a complete end-to-end system.
+
+In this project we will extend ADIOS~\cite{adios} with new APIs and novel techniques
+to coordinate with the storage back-end. The key idea is offering the ability to
+negotiate with the storage system to agree on the tradeoff between data quality
+and response time. We will test our ideas on the next generation OLCF,
+SNL/LANL, and NERSC systems.
+%
+We plan to leverage the many inovations we discovered in 
+the Lightweight File Systems~\cite{lwfs} (LWFS) project and the Ceph storage system project (ceph.com).
+In the LWFS project, Sandia sought to
+demonstrate stripping down a file system to the bare minimum features required
+for any storage system. Then, additional functionality and semantics, such as
+POSIX-style directories and consistency requirements, were added in using
+auxiliary services separate from the stack or by layering providing a new
+interface to the storage system. The project's first phase demonstrated that
+using a core object storage system with integrated security providing
+authentication and authorization services, customized storage interfaces could
+be developed with exactly the overhead required for different applications. The
+second phase of LWFS, known as Sirocco~\cite{sirocco}, is investigating how to
+manage storage devices for efficient and scalable storage performance.  Sirocco
+has been in process for three years and has developed a usable storage infrastructure.
+
+The Ceph project (ceph.com) started at UC Santa Cruz in 2004 and
+was initially funded by DOE/NNSA involving LLNL, LANL, and Sandia,
+with the goal to create an object-based parallel file system that
+addresses the well-known metadata service bottleneck of creating
+and accessing 1000s of files comprising a checkpoint~\cite{weil:osdi06}.
+However, after the Ceph kernel client was pulled into the Linux
+mainline in 2010, the focus of Ceph shifted to its object-based
+storage subsystem, also called Reliable Autonomous Distributed
+Object Store (RADOS), which now is the dominant storage system in
+OpenStack production-level deployments and was recently selected
+as the new backend for Flickr and Yahoo! Cloud Services. In this
+project we will leverage two of the key innovations of Ceph: (1)
+Controlled Replication Under Scalable Hashing (CRUSH)~\cite{weil:sc06}
+will help making resource identification and predictable performance
+scalable, and (2) extensible storage object classes~\cite{watkins:ucsctr15}
+will provide the main mechanisms of introducing new APIs,
+metadata vocabulary, and quality of service into the storage stack.

### 2015-07-02 narrative/intro.tex
move some of the Cepth and ADIOS part into the enabling technologies, and try to re-organize to show the focus of this proposal is on theme 2.

@@
 Our approach targets the expected characteristics of exascale storage
-hardware. The storage layer will be partitioned into multiple heterogenous
+hardware. The storage layer will be partitioned into multiple heterogeneous
 tiers with vastly different performance characteristics. The layers
@@
 %
+\section{Scalable Storage Software Infrastructure}
+Our focus in this proposal is in theme 2: Scalable Storage Software Infrastructure. We will focus on a state of the art
+SSIO solution to solve problems which researchers are beginning to face and will make science on exascale systems
+problematic without new insights and solutions.
+
 The overarching idea is that data from applications can be characterized by different
@@
 %indexing to better cope with the expected tsunami of data.
-
-In this project we will extend ADIOS~\cite{adios} with new APIs and novel techniques
-to coordinate with the storage back-end. The key idea is offering the ability to
-negotiate with the storage system to agree on the tradeoff between data quality
-and response time. We will test our ideas on the next generation OLCF,
-SNL/LANL, and NERSC systems.
-%
-We plan to leverage the many inovations we discovered in 
-the Lightweight File Systems~\cite{lwfs} (LWFS) project and the Ceph storage system project (ceph.com).
-In the LWFS project, Sandia sought to
-demonstrate stripping down a file system to the bare minimum features required
-for any storage system. Then, additional functionality and semantics, such as
-POSIX-style directories and consistency requirements, were added in using
-auxiliary services separate from the stack or by layering providing a new
-interface to the storage system. The project's first phase demonstrated that
-using a core object storage system with integrated security providing
-authentication and authorization services, customized storage interfaces could
-be developed with exactly the overhead required for different applications. The
-second phase of LWFS, known as Sirocco~\cite{sirocco}, is investigating how to
-manage storage devices for efficient and scalable storage performance.  Sirocco
-has been in process for three years and has developed a usable storage infrastructure.
-
-The Ceph project (ceph.com) started at UC Santa Cruz in 2004 and
-was initially funded by DOE/NNSA involving LLNL, LANL, and Sandia,
-with the goal to create an object-based parallel file system that
-addresses the well-known metadata service bottleneck of creating
-and accessing 1000s of files comprising a checkpoint~\cite{weil:osdi06}.
-However, after the Ceph kernel client was pulled into the Linux
-mainline in 2010, the focus of Ceph shifted to its object-based
-storage subsystem, also called Reliable Autonomous Distributed
-Object Store (RADOS), which now is the dominant storage system in
-OpenStack production-level deployments and was recently selected
-as the new backend for Flickr and Yahoo! Cloud Services. In this
-project we will leverage two of the key innovations of Ceph: (1)
-Controlled Replication Under Scalable Hashing (CRUSH)~\cite{weil:sc06}
-will help making resource identification and predictable performance
-scalable, and (2) extensible storage object classes~\cite{watkins:ucsctr15}
-will provide the main mechanisms of introducing new APIs,
-metadata vocabulary, and quality of service into the storage stack.
 

### 2015-07-02 narrative/intro.tex
table for our work from the CFP, and try to eventually map this into different sections in the proposal

@@
 
+Our proposal will be responsive to theme 2 in the call, and we summarize them in the following table.
+\begin{table}[htbp]
+\vspace{2ex}
+\begin{center}
+\begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{1.6cm} |}\hline
+\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant S2E2 work} & \multicolumn{1}{|c|}{\bf See} \\\hline\hline
+%\multicolumn{2}{|c|}{Collaboration} & 
+\multicolumn{2}{|p {2.4cm}|}{Approaches to improve the ability of SSIO software to support checkpoint/restart;} & 
+General piece we are doing for this... & 
+\S\ref{sec:?}, \S\ref{sec:??}, \\\hline
+1 & Data \hspace{1cm} abstractions &What will we do& \S\ref{sec:?}, \S\ref{sec:?}, \S\ref{sec:?} \\\hline
+2 & Mapping science data models onto hierarchical storage &What we are doing & \S\ref{sec:?}, \S\ref{sec:?} \\\hline
+3 & Mechanisms for data movement across the storage hierarchy  & what are we doing. & \S\ref{sec:?} \\\hline
+4 &  Extend the interaction of IO middleware with the resource management system & What we are doing & \S\ref{sec:?}, \S\ref{sec:?}  \\\hline
+5 & Design and implement of IO middleware architectures  &What we are doing & \S\ref{sec:?}, \S\ref{sec:?}, \S\ref{sec:?} \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
 %

### 2015-07-02 narrative/proposed-work.tex
table for our work from the CFP, and try to eventually map this into different sections in the proposal

@@
 \label{subsec:challenges}
+
 Our research into the storage system and middle ware layers must address
@@
 
+{\bf \color{red}we need to condense these challenges into a few basic principles - 4 at most - SAK}
+
+We propose to overcome these challenges by pioneering a new {\bf knowledge-centric approach} to a dynamic storage system
+and I/O layer which takes into account user prioritization and utility to optimize the SSIO layer. The approach is based on 
+four underlying principles:
+
+\underline{Principle 1: A knowledge-centric system} , let user knowledge  to define data policies. talk about this pieces...
+
+\underline{Principle 2: A SSIO system which integrates system knowledge together with application knowledge} to understand
+how to organize data in the deep storage hierarchy
+
+\underline{Principle 3: Predictable Performance in the SSIO layer} so that intelligent choices can be made by both the user and the system
+
 \input{narrative/data-description} % includes data-desc, apis, and qos-app.

### 2015-07-02 narrative/naming-service.tex
start to reshape the proposal
naming/discovery service.

@@
-\subsubsection{Naming and Discovery Service}
+\subsection{Naming and Discovery Service}
 
@@
 Our goal with this proposal is to offer a similar capability but with a finer
-granularity leveraging the inner structure of scientific data, 
-and manage discovering efficiently, where the data may
+granularity, leveraging the inner structure of scientific data,
+and manage data discovery efficiently, where the data may
 be split and placed on several storage layers, or may have migrated
-based on system pressures. Instead of a single file such as might be used to
-store an entire timestep output for a simulation, we will demonstrate offering
-the same capability at a subset of a single variable level. By shifting to a
+based on system pressures or user or system policies. Instead of a single file such as might be used to
+store an entire timestep output for a simulation, we will demonstrate extending
+this capability to a subset of a single variable level. By shifting to a
 finer-grained approach, we will enable more effective use of close/fast/small
-storage tiers and manage the impact when there is insufficient space forcing
-data to move. Traiditional HSM stores an entire file on a tier making room if
+storage tiers and prioritize objects with high importance.
+Traiditional HSM stores an entire file on a tier making room if
 insufficient space is available. With this shift to a partial variable
-granularity, a 1 PB output with 500 GB of ``high interest'' data can limit this
-costly tier ussage to just 500 GB greatly enhancing usability. This reduces
-pressure forcing data migration, but does not prevent it. We still need to be
-able to find the data should it move based on system pressure.
-
-The resulting naming and discovery service will attach the Sirocco
+granularity, a 1 PB output with 200 GB of ``high interest'' data can limit this
+costly tier ussage to just 200 GB greatly, improving usability and performance while still
+allowing science to be done. This effectively reduces
+pressure forcing data migration. Meanwhile, we need to be
+able to locate the data should it moves.
+Utilizing the description of data (section \ref{}) informed by users and packed by the middleware,
+the resulting naming and discovery service will attach the Sirocco
 contaier/object/fork/address tuple with a special attribute indicating that
-this record is of ``high interest'' strongly encouraging keeping the variable
-in close/fast/small storage as long as possible. In addition to maintaining
-this additional metadata, the application and/or middleware must provide this
-information. The storage system simply needs to act appropriately on this data
-based on the attribute.
+this record is of ``high interest'' strongly encouraging keeping the record
+in close/fast/small storage as long as possible.
 

### 2015-07-02 narrative/proposed-work.tex
moved around the sections to integrate them and give writing assignments, modifications

@@
 \label{subsec:challenges}
-
+{\bf {\color{red}Norbert will clean these pieces}}
 Our research into the storage system and middle ware layers must address
@@
 %\input{narrative/data-desc}
+
+{\bf {\color{red}Scott will integrate these pieces}}
 \input{narrative/data-reduction} % includes data-reorg and the api specs move to the above section -
 \input{narrative/data-reorg}
+
 %\input{narrative/qos-app}
+{\bf {\color{red}Hasan will integrate these pieces, Manish after Hasan}}
 \input{narrative/placement}
+\input{narrative/migration}
+\input{narrative/eviction-metrics}
+
 %\input{narrative/policies}
@@
 
-\input{narrative/migration}
+{\bf {\color{red}Gary will integrate these pieces}}
 \input{narrative/resource-mgmt}
@@
 %\input{narrative/discovery}
-\input{narrative/eviction-metrics}
+{\bf {\color{red}Jay  will integrate these pieces}}
 \input{narrative/naming-service}

### 2015-07-02 narrative/proposed-work.tex
Added some obvious research questions in the 4 areas.

--- 2015-07-02 narrative/proposed-work.tex
@@
 \label{subsec:challenges}
+\label{sec:proposed}
+
+\begin{enumerate}
+\item Re-factoring of the data
+  \begin{itemize}
+  \item What kind of reorganization would work when?
+  \item Would it take too much time and resources + memory and how can we
+    manage the cost of the reorganization
+  \item Will the users loose too much value from the data during refactoring
+    and how can we control/bound this loss of value
+  \end{itemize}
+\item Quality of Service
+  \begin{itemize}
+  \item Can we make any gurantees in a scalable decentralized system?
+  \item What quality of service metrics can be make gurantees on in a scalable
+    way?
+  \item What kind of error rates will we have in our estimation and what error
+    rates are acceptable?
+  \item How much will the qos scheduler, admission control, manager cost in
+    terms of resources and time
+  \item How do we make the scheduler scalable with the size of the system?
+  \end{itemize}
+\item Discovery and naming
+  \begin{itemize}
+  \item Which data block and what quality/type do we respond to a query
+    with?
+  \item How will we scale a fully distributed storage system's metadata
+    management
+  \item Can we accelerate the discovery process with the addition of an
+    application/user level metadata store that tracks where data was
+    originally written and how was it structured?
+  \item Can we provide time estimates for read/write calls in the presence of
+    an unbounded discovery system? If not how do we bound the discovery time?
+  \item Time cost of discovery vs time cost of actually reading the data
+    (latency vs bandwidth essentially)
+  \end{itemize}
+\item Migration
+  \begin{itemize}
+  \item Can we do scalable migration without making the discovery process
+    unbounded
+  \item What are the parameters and input for migration
+  \item Can we use the additional application level knowledge of data to purge
+    portions of the data from storage without making the data too much less
+    valuable?
+  \item With so many objects and no centralized directory how do we know what
+    is actually in the storage to be purged?
+  \item How do purge in a way that doesn't require a central authority,
+    doesn't interfere with other I/O and isn't bottlenecked by some sort of
+    global consistency
+  \end{itemize}
+\end{enumerate}
 
@@
 many of these challenges including
-\begin{enumerate}
-\item How do we describe the user intentions at the API layer and have this
-  communicated down to the middle-ware and storage layers?
-\item How do we allow users to define their user defined compression
-  techniques and have this data read back from the storage system layers?
-  Where do we execute the decompression techniques 
-\item How do we evaluate the tradeoffs at runtime to guide data placement,
-  including the movement of data across the network, across the different
-  storage tiers? (data movement and quality tradeoff)
-\item Can we use different forms of learning techniques as daemons on the
-  system to perform data migration from one storage tier to another? For
-  example, if we see that a user is looking at one time slice after another
-  for a certain object in their dataset which is in the slower tiers of the
-  hierarchy, do we automatically propane up the data which has not been
-  requested in the hope that this will be requested data? (prediction and
-  prefetching, user defined compression on movement)
-\item Can we understand the true need of campaign storage and understand the
-  possible impact of the campaign storage if we run a hadoop-like
-  file-system instead of a lustre/gpfs file system? (additional layers)
-\item Can we limit the amount of data duplication? Data space is going to be
-  very constrained on the exascale system so we must ensure that minimal
-  copies are made of data, and when data is duplicated, they are removed
-  through a garbage collection routine. (space management)
-\item Can we build a model to give us the time estimates during the reading
-  and writing phases, so that rules can be in place from the user
-  perspective to make adaptive decisions. For example, if the file system
-  says that reading will take 3 months, users can they place in rules to
-  then read in a subset of data, and they can understand which data can be
-  read quickly and which pieces will take more time. (estimation)
-\item Can we understand how to place code in the system which will allow
-  data-regeneration to take place. (regeneration)
-\item How does the storage system do a better job in managing request from
-  all of the users on a LCF than today? We realize that today users who have
-  a better middlware system can often lock other users from getting high
-  performance when they are running. If we have the concept of currency
-  which is eventually used in the same extent as node-hours, then users will
-  have to be able to think about how much storage and how much bandwidth
-  they can choose. One question that needs to be understood is if there are
-  times when the system sees that there are very few storage system
-  resources being used, then the lucky users can get the bandwidth cheaper
-  than at times of heavy usage. How can we enforce this? (Fairness)
-\item How do we manage all of the metdata not only from the principle
-  objects, but from the sub-objects? How well will this scale when we have
-  users who can potentially create billions of objects from their
-  simulation? (scalability and discovery)
+% \begin{enumerate}
+% \item How do we describe the user intentions at the API layer and have this
+%   communicated down to the middle-ware and storage layers?
+% \item How do we allow users to define their user defined compression
+%   techniques and have this data read back from the storage system layers?
+%   Where do we execute the decompression techniques 
+% \item How do we evaluate the tradeoffs at runtime to guide data placement,
+%   including the movement of data across the network, across the different
+%   storage tiers? (data movement and quality tradeoff)
+% \item Can we use different forms of learning techniques as daemons on the
+%   system to perform data migration from one storage tier to another? For
+%   example, if we see that a user is looking at one time slice after another
+%   for a certain object in their dataset which is in the slower tiers of the
+%   hierarchy, do we automatically propane up the data which has not been
+%   requested in the hope that this will be requested data? (prediction and
+%   prefetching, user defined compression on movement)
+% \item Can we understand the true need of campaign storage and understand the
+%   possible impact of the campaign storage if we run a hadoop-like
+%   file-system instead of a lustre/gpfs file system? (additional layers)
+% \item Can we limit the amount of data duplication? Data space is going to be
+%   very constrained on the exascale system so we must ensure that minimal
+%   copies are made of data, and when data is duplicated, they are removed
+%   through a garbage collection routine. (space management)
+% \item Can we build a model to give us the time estimates during the reading
+%   and writing phases, so that rules can be in place from the user
+%   perspective to make adaptive decisions. For example, if the file system
+%   says that reading will take 3 months, users can they place in rules to
+%   then read in a subset of data, and they can understand which data can be
+%   read quickly and which pieces will take more time. (estimation)
+% \item Can we understand how to place code in the system which will allow
+%   data-regeneration to take place. (regeneration)
+% \item How does the storage system do a better job in managing request from
+%   all of the users on a LCF than today? We realize that today users who have
+%   a better middlware system can often lock other users from getting high
+%   performance when they are running. If we have the concept of currency
+%   which is eventually used in the same extent as node-hours, then users will
+%   have to be able to think about how much storage and how much bandwidth
+%   they can choose. One question that needs to be understood is if there are
+%   times when the system sees that there are very few storage system
+%   resources being used, then the lucky users can get the bandwidth cheaper
+%   than at times of heavy usage. How can we enforce this? (Fairness)
+% \item How do we manage all of the metdata not only from the principle
+%   objects, but from the sub-objects? How well will this scale when we have
+%   users who can potentially create billions of objects from their
+%   simulation? (scalability and discovery)
+
+\paragraph{Deliverables and artifacts}
 
@@
 
-%\subsection{Application Interface}
+% \subsection{Application Interface}
 
-%\input{narrative/data-desc}
+% \input{narrative/data-desc}
 \input{narrative/data-reduction} % includes data-reorg and the api specs move to the above section -
 \input{narrative/data-reorg}
-%\input{narrative/qos-app}
+% \input{narrative/qos-app}
 \input{narrative/placement}
-%\input{narrative/policies}
-%\input{narrative/apis}
+% \input{narrative/policies}
+% \input{narrative/apis}
 
-%\subsection{Storage System}
+% \subsection{Storage System}
 
@@
 \input{narrative/qos-storage}
-%\input{narrative/discovery}
+% \input{narrative/discovery}
 \input{narrative/eviction-metrics}
 \input{narrative/naming-service}
-%\input{narrative/time-estimation}
+% \input{narrative/time-estimation}
 \input{narrative/infrastructure}

### 2015-07-02 narrative/data-description.tex
section refienment

@@
 the relationship between user data to be readily captured and embedded with the 
-raw data. As an example, retrieving a 3D field data generated from a simulation and then visualizing it
-requires the coordinates and connectivity variable 1) to be accessed together and 2) with low
-latency in order to achieve a good user experience. The description of data, 
-as enabled by managing data as objects, bridges 
+raw data. As an example, retrieving a 3D field data generated from a simulation 
+and then visualizing it requires the coordinates and connectivity variable 
+1) to be accessed together and 2) with low latency in order to achieve a good user experience. 
+The description of data, as enabled by managing data as objects, bridges 
 the semantics gap between applications, middleware, and storage systems,
@@
 optimize application and system performance.
+
+%The new APIs will enable the bridge by specifying selectable performance/quality/- cost tradeoffs 
+%from both the application and system perspectives based upon the user guided rules/policy and 
+%runtime system monitoring status. It allows the middleware to make best possible decisions from 
+%the feedback of storage system knowledge, such that it will embed user intentions and the available 
+%system storage. 
+%We want the system to give the users a certain amount of currency 
+%in terms of bandwidth, storage space on each level, and latency expectations. These notions will be fuzzy 
+%but they will allow the user to make ad-hoc decisions to figure out what needs to be saved.
 
@@
 to avoid the slow access to large-capacity lower-level storage, such as tapes.
-We propose a new technique, described by a utility function plugged in by a user, where objects
-can be cast into multiple buckets of data, and each bucket can be prioritized differently according
-to its utility value.
+We propose a new technique, described by a utility function plugged in by a user, 
+where objects can be cast into multiple buckets of data, and each bucket can be prioritized 
+differently according to its utility value.
 Once this is done, the middleware system will construct the description of data, re-organize
@@
 
-The user would like the ability to get a certain amount of Quality of Service 
+In the meanwhile, the user would like the ability to get a certain amount of Quality of Service 
 such that they can then make decisions when the expected bandwidth, 
@@
 the application with the storage system. Towards this end, we propose to
-address the quality of service requirements by proivding the application
+address the quality of service requirements by providing the application
 with mechanisms to specify the quality of I/O service, interrogate the
@@
 ADIOS \cite{lofstead2008flexible} we propose to explore the design of two
-mechanisms for this purpose. Firstly, we will explore the augmentation of
-I/O applicatition programing interfaces (I/O APIs) to allow applications to
+mechanisms for this purpose.
+
+Firstly, we will explore the augmentation of
+I/O application programing interfaces (I/O APIs) to allow applications to
 both specify timing and quality information and also query the storage
-system for timing estimates. Through this interface we expect the
+system for timing estimates. If user decide to write or read data to or from hierarchical storage systems, 
+they will rely on the API to send their intentions for inquiry and examine the system status, 
+including how much data they would like to write/read, desired bandwidth, data compression method, etc, 
+or the user can express their intension of writing data right now no matter what the traffic is now.
+Through this interface we expect the
 application and user to gain insights into how long a single output or input
@@
 
-We will also explore external data annotations, such as those
+Secondly, we will also explore external data annotations, such as those
 provided by the configuration file in ADIOS. Through the use of these
@@
 different storage layer.
+
+The capability of APIs provide dynamic runtime information to user and make the system 
+more transparent to them rather than a black box. The end user could analyze or visualize data 
+in a more comfortable position, where they are able to envision 
+how much longer they are going to wait for data.
 
@@
 
-\paragraph{Challenges}
+\paragraph{Challenges:}
 The design of new APIs posses both technical and adoption challenges. Here

### 2015-07-02 narrative/naming-service.tex
fixed typo in proposed-work.tex focused naming-service a bit

@@
 Given the well documented difficulties with POSIX-compliant naming and our
-approach that decomposes data into potentially large number of objects,
-we must take a new approach to naming.
-Data migration, as a result of system pressures or user/system policies,
-further adds to the complexity of managing naming and discovery. Our naming
-service must maintain sufficient metadata information, track objects
-as they move across storage layers, and facilitates data discovery.
+approach that decomposes what would be a single file into potentially large
+number of objects, we must take a new approach to naming.  Data migration, as a
+result of system pressures or user/system policies, further adds to the
+complexity of managing naming and discovery. Our naming service must maintain
+sufficient metadata information, track or be able to find objects as they move
+across storage layers, and facilitates data discovery.
 
-Hierarchical Storage Management (HSM) systems \cite{hsm} offer a strict caching approach
+Hierarchical Storage Management (HSM) systems~\cite{hsm} offer a strict caching approach
 for managing different storage capacities trading off performance for capacity.
@@
 Our goal with this proposal is to offer a similar capability but with a finer
-granularity, leveraging the inner structure of scientific data,
-and manage data discovery efficiently, where the data may
-be split and placed on several storage layers, or may have migrated
-based on system pressures or user or system policies. Instead of a single file such as might be used to
-store an entire timestep output for a simulation, we will demonstrate extending
-this capability to a subset of a single variable level. By shifting to a
-finer-grained approach, we will enable more effective use of close/fast/small
-storage tiers and prioritize objects with high importance.
-Traiditional HSM stores an entire file on a tier making room if
-insufficient space is available. With this shift to a partial variable
-granularity, a 1 PB output with 200 GB of ``high interest'' data can limit this
-costly tier ussage to just 200 GB greatly, improving usability and performance while still
-allowing science to be done. This effectively reduces
-pressure forcing data migration. Meanwhile, we need to be
-able to locate the data should it moves.
-Utilizing the description of data (section \ref{}) informed by users and packed by the middleware,
-the resulting naming and discovery service will attach the Sirocco
-contaier/object/fork/address tuple with a special attribute indicating that
-this record is of ``high interest'' strongly encouraging keeping the record
-in close/fast/small storage as long as possible.
+granularity leveraging the inner structure of scientific data. It must also
+handle data discovery efficiently for cases where data is spread across several
+storage layers or some of those pieces may have migrated based on system
+pressures or user or system policies. Instead of a single file, such as might
+be used to store an entire timestep output for a simulation, we will
+demonstrate a metadata service with naming at the science variable level, but
+with an associated list of all of the variable pieces and potentially different
+versions stored throughout the storage system. By shifting to a finer-grained
+approach, we will enable more effective use of close/fast/small storage tiers
+and prioritize objects with high importance.  Traiditional HSM stores an entire
+file on a tier making room if insufficient space is available.  With this shift
+to a partial variable granularity, a 1 PB output with 500 GB of ``high
+interest'' data can limit this costly tier usage to just 500 GB greatly
+improving storage availability in the close/fast/small tier and focus
+performance/capacity costs to get the most science from from the platform.  By
+only placing ``high interest'' data in the close/fast/small tier, we will
+effectively reduce pressure forcing data migration hurting both writing
+performance and the later data analytics required for generating scientific
+insights.  Sirocco identifies a data segment using a
+contaier/object/fork/address tuple that can have associated attributes. We will
+both add a special attribute indicating that this record is of ``high
+interest'' as well adapting the migration mechanisms to use this attribute, if
+present, to strongly encouraging keeping the record in close/fast/small storage
+as long as possible.
 
-The challenge for discovery is that potentially, data will migrate from where
-it is initially stored to a new location within the storage system. Sirocco
-offers an ability to search for data that has moved as well as forcing a
-particular resilience-based replica be the ``authoritative'' version. We will
-investigate if the current Sirocco functionality is capable of supporting the
-new operating modes we wish to offer. Initial expectations suggest having
-bounded time guarantees for finding data are critical for offering the quality
-of service guarantees we wish to offer. Some potential approaches can build on
-CRUSH~\cite{ceph} from Ceph to offer a map of where to search for data
-sequentially. Given our multiple storage tiers, we would extend this using some
-search limiting mechanism indicating to shift to the next tier to continue
-searching because the requested data never made it to this location. This new
-work will be an expansion of Sirocco's currently planned features.
+Meanwhile, we need to be able to locate or ``discover'' the data should it
+move.  The challenge for discovery is that potentially, data will migrate from
+where it is initially stored to a new location within the storage system.
+Sirocco offers an ability to search for data that has moved as well as forcing
+a particular resilience-based replica be the ``authoritative'' version
+migrating the data to a particular location. We will investigate if the current
+Sirocco functionality is capable of supporting the new operating modes we wish
+to offer. Initial expectations suggest having bounded time guarantees for
+finding data are critical for offering the quality of service guarantees we
+wish to offer. Some potential approaches can build on CRUSH~\cite{ceph} from
+Ceph to offer a map of where to search for data sequentially. Given our
+multiple storage tiers, we would extend this using some mechanism to shift to
+the next tier to continue searching because the requested data never made it to
+this location. This new work will be an expansion of Sirocco's currently
+planned features.
 
 The overriding theme for this proposal of providing a cooperative relationship
-between the user and the storage system extends into this area as well. In
-order for the user to negotiate with the storage system for data
-quality/retrieval time tradeoff, additional data is required. Providing this
-functionality requries the naming service be able to both return a list of data
-versions, but also some indication of data quality and estimated retrieval time
-for each. This estimate will be disucssed elsewhere in this proposal.
+between the user and the storage system extends into this area as well.
+Negotiating the data quality/retrieval time tradeoff reuqires additional data.
+Providing this functionality requries the naming service be able to both return
+a list of data versions and an indication of data quality. This will be
+combined with a retrieval time estimate for each verstion by the middleware to
+determine what data version should be retrieved.  This estimate will be
+disucssed elsewhere in this proposal.
 
@@
 compatibility. However, this view will not offer the same strict semantics
-POSIX requires. We must break these strict semantics to address scalability
-problems because of the serialized access to a single source for creating and
+POSIX defines. We must break these strict semantics to address scalability
+problems forced by the serialized access to a single source for creating and
 accessing files.  Several efforts~\cite{giga+,pvfs,others} have worked to
@@
 single directory or subtree. While these approaches help, they do not address
-this key scalability limitation. Instead, we ar willing to offer a short
-duration to eventually consistent POSIX view to offer the performance and
+this key scalability limitation of serialized access--even if the metadata
+service is spread across multiple nodes. Instead, we will offer a
+short duration to consistency POSIX view to offer the performance and
 consistency requirements an exascale application demands.
@@
 level completely can work well for scale out applications where data is created
-or consumed by a single process at a time rather than potentially O(1 million)
-processes all actively collectively for a single ``object''. To address this
-case, having some system integrated metadata services to associate names with
-these object is a preferable solution. While it will not have the same
-synchronous consistency semantics as POSIX, we will offer something as close as
-possible to address legacy application and command-line style data access and
-migration to other storage types using these POSIX semantics.
+or consumed by a single process at a time or just in a work queue rather than
+potentially O(1 million) processes all actively collectively for a single
+``object''. To address this case, having some system integrated metadata
+services to associate names with these object is a preferable solution. While
+it will not have the same synchronous consistency semantics as POSIX, we will
+offer something as close as possible to address legacy application and
+command-line style data access and migration to other storage types using these
+POSIX semantics.
 
@@
 to incorporate authorization capabilities. Sirocco currently integrates with a
-Kerberos service for authentication and authorization. Given a capability
-ticket, a user can access different objects as needed. This ticket structure
-offers protection services typically offered on POSIX directories and files,
-but can do it at the fork level instead. This allows a reduced quality data
-version to be available for the general users while the high-quality version
-would be limited to the data creator. This and other considerations for
-security must be incorporated into the entire naming and metadat service.
+Kerberos service for authentication. Given a capability ticket, a user can
+access different objects as needed. This ticket structure offers protection
+services typically offered on POSIX directories and files, but can do it at the
+fork level instead. This allows a reduced quality data version to be available
+for the general users while the high-quality version would be limited to the
+data creator. This and other considerations for security must be incorporated
+into the entire naming and metadata service.
 
@@
 \item Milestone 1: Demonstrate a metadata service capable of serving both POSIX
-clients and our clients, but without consideration for authentication, authorization, or data migration.
+clients and our clients, but without consideration for authentication,
+authorization, or data migration.
 
 \item Milestone 2: Demonstrate a time bounded search approach for finding data
-within the storage system to identify data the current location and retrieval
-performance characteristics for data since migration. This will update the
-cached information in the metadata service to short-circuit future requests for
-the data presuming that it does not migrate again.
+within the storage system to identify data and the current location.  This will
+update the cached information in the metadata service to short-circuit future
+requests for the data presuming that it does not migrate again soon.
 

### 2015-07-02 narrative/intro.tex
Added bio and minor restructuring of introductionZ

@@
 
-Avoiding impending bottlenecks in exascale scientific discovery will require
-new research into managing and storing the large amounts of data
-that will be produced by simulations and analyzed for months
-afterwards.
-%
-In this project we will demonstrate a cooperative approach for storing data
-where the user and the storage system work together to achieve the best possible data
-quality and performance given data features, storage tier characteristics,
-and current system state. We will investigate novel techniques to facilitate efficient
-mapping of data objects, even partitioning individual variables, from the user
-space onto multiple storage tiers and enable application-guided data
-reductions/transformations to address capacity and bandwidth bottlenecks while
-constraining the error to be within user provided bounds.
-% 
-We will address the associated I/O and storage challenges in the context of
-current and emerging storage landscapes and expedite insights into critical
-scientific processes, while demonstrating the validity of our approach in key DOE
-domains. We will research techniques for creating a Scalable
-Storage Software Infrastructure, integrating services from the middleware
-layer with Storage System features to support the
-novel strategy of distributing data both horizontally and vertically across the storage
-system. 
-The negotiation between these layers will be provided as a fundamental service, 
-allowing users to easily save ``the best'' amount of data, automatically placed
-throughout the entire storage stack.
-
-The metric we are most interested in optimizing is time to knowledge.
-Current approaches to addressing the I/O bottleneck fall into two broad
-categories: parallel file system approaches that optimize the throughput for
-an entire system, and I/O middleware approaches that optimize the
-performance of a single application. Both approaches have been successful to date, but
-are unlikely to overcome the remaining obstacles in reaching exascale. Instead
-of trying to optimize throughput, we will seek to reduce the time to
-knowledge. This is the most significant metric for scientific applications
-where the desired outcome is not storing data, but rather executing a
-knowledge extraction process on the data. Moreover, we aim to perform the
-optimization not for a single application, but rather for workloads within a
-multi-user multi-application system environment.
-
-Our approach targets the expected characteristics of exascale storage
-hardware. The storage layer will be partitioned into multiple heterogeneous
-tiers with vastly different performance characteristics. The layers
-will be further differentiated by the constraints on capacity
-and data lifetime within each layer. Tape archival storage will still maintain
-data long term, but access to this data will be orders of magnitude slower
-than the next layer and naively accessing data from archival storage will
-greatly impact productivity.
-%
 {\em The purpose of computing is insight, not numbers}~\cite{Hamming:book}.
@@
 instantly sympathize with the central objective of this proposal--minimizing
-the time to knowledge.  Our approach in this proposal is to enable researchers
-to place more knowledge into the SSIO system so that data can be recast from a bunch of bytes
-to information which the system can then place and optimize the time to gain knowledge from exascale simulations.
-We first motivate this by a problem our team has been working on in the past six months, and then recast this problem into
-a series of research questions and challenges which must be addressed in order to optimize the SSIO system for exascale science.
+the time to knowledge.
+
+Avoiding impending bottlenecks in exascale scientific discovery will require
+new research into managing and storing the large amounts of data that will be
+produced by simulations and analyzed for months afterwards.
+%
+In this project we will demonstrate a cooperative approach for storing data
+where the user and the storage system work together to achieve the best
+possible data quality and performance given data features, storage tier
+characteristics, and current system state. We will investigate novel techniques
+to facilitate efficient mapping of data objects, even partitioning individual
+variables, from the user space onto multiple storage tiers and enable
+application-guided data reductions/transformations to address capacity and
+bandwidth bottlenecks while constraining the error to be within user provided
+bounds.
+% 
+We will address the associated I/O and storage challenges in the context of
+current and emerging storage landscapes and expedite insights into critical
+scientific processes, while demonstrating the validity of our approach in key
+DOE domains. We will research techniques for creating a Scalable Storage
+Software Infrastructure, integrating services from the middleware layer with
+Storage System features to support the novel strategy of distributing data both
+horizontally and vertically across the storage system.  The negotiation between
+these layers will be provided as a fundamental service, allowing users to
+easily save ``the best'' amount of data, automatically placed throughout the
+entire storage stack.
+
+The metric we are most interested in optimizing is time to knowledge.  Current
+approaches to addressing the I/O bottleneck fall into two broad categories:
+parallel file system approaches that optimize the throughput for an entire
+system, and I/O middleware approaches that optimize the performance of a single
+application. Both approaches have been successful to date, but are unlikely to
+overcome the remaining obstacles in reaching exascale. Instead of trying to
+optimize throughput, we will seek to reduce the time to knowledge. This is the
+most significant metric for scientific applications where the desired outcome
+is not storing data, but rather executing a knowledge extraction process on the
+data. Moreover, we aim to perform the optimization not for a single
+application, but rather for workloads within a multi-user multi-application
+system environment.
+
+Our approach targets the expected characteristics of exascale storage hardware.
+The storage layer will be partitioned into multiple heterogeneous tiers with
+vastly different performance characteristics. The layers will be further
+differentiated by the constraints on capacity and data lifetime within each
+layer. Tape archival storage will still maintain data long term, but access to
+this data will be orders of magnitude slower than the next layer and naively
+accessing data from archival storage will greatly impact productivity.
+%
+Our approach in this proposal is to enable researchers to place more knowledge
+into the SSIO system so that data can be recast from a bunch of bytes to
+information which the system can then place and optimize the time to gain
+knowledge from exascale simulations.  We first motivate this by a problem our
+team has been working on in the past six months, and then recast this problem
+into a series of research questions and challenges which must be addressed in
+order to optimize the SSIO system for exascale science.
 
@@
 
-The overarching idea is that data from applications can be characterized by different
-levels of importance and different level of importance over time. 
-Codes such as the XGC1 simulation from C.
+Take for example the latest use case for the XGC1 simulation from C.
 S.
-Chang (PPPL), which is one of largest users of Leadership Class Facilities (over 300M hours at ANL, NERSC, and ORNL) has launched a series of simulations which need to write out 100 PB of data on the Titan system.
-Due to resource limitations, the simulation will only write about 10 PB, but
-when our team considered not only the time it would take to write out 10 PB in 10 days of run, but also where we would store this, which had to be tape since it would take
-over half of the file system, we decided that we couldn't write this amount of data. In other words we had to make decisions of which data to save from the original 100 PB.
-The first thing that we had to discover is how much space we could have on the parallel file system at any given time, and how much time it would take to archive 1 PB of data,
-which turned out to be 10 days, and we could only store 2 PB of data at once. This quickly led us to 1) decide which pieces of data to save from the large amount of data; in 
-other words, we had to prioritize which were the most important pieces. We had to further breakdown the large data into two categories, data which we think we will touch after
-the simulation is complete ( after the 10 days of wallclock time was finished), and what we would do with this data (the access pattern of the large data). We then had to 
-discover application specific data reduction techniques, and encode that into our solution, on top of ADIOS. We then had to implement a crude discovery system to understand
-which pieces of data were stored on the parallel file system and which pieces moved over to PPPL, and which pieces were on tape.  This solution worked well for one day
-but after the second day more problems occurred, the variability of the time it took to write the data made the I/O prohibitively costly, because even though we had the
-majority of the OLCF, the file system was shared and sometimes the writing time took less 100 seconds, and at other times it was over 1,000 seconds. We had no method
-to monitor this time and then write less data when the time was going to be longer.   After the simulation is complete we also need to visualize this reduced data, and 
-an additional problem appeared; just to get a first look at the visualization data seemed like it could take over a month because some of the timesteps were on tape and other
-timesteps were on the file system. We had no method to just look at the data quickly and understand which regions of the simulation to focus on.
+Chang (PPPL), who is one of the largest users of Leadership Class Facilities (over 300M hours at ALCF, NERSC, and OLCF).
+They launched a series of simulations, planning to produce 100~PB of data over 10 days of total runtime on the Titan system.
+An expert team was assembled from the user group, the middleware I/O team to help with the application I/O (using ADIOS), the local experts of the Lustre file-system and the tape archival system (HPSS).
+Due to physical resource limitations, the simulation was restricted to write about 10~PB initially, but then considering the time to write and read this much data on the tape archive system, and the financial costs to store it, this plan was cut again considerably.
+Note that such time and cost information was not known by the user; the HPSS expert had to inform everyone else about it.
+The user team had to make hard decisions of which data to save from the originally planned 100~PB.
+The file system experts told about the then-current file system limitations in size, performance and purge policy.
+The user then decided to write 0.5~PB per day (the individual daily simulation runs were and are going to be separated by many days or weeks), 5~PB in total, storing maximum of 2~PB of data on disks at a time and storing the rest on tape.
+
+This forced the team to decide which pieces of data to save from the large amount of data; in 
+other words, they had to prioritize which were the most important pieces and save only those.
+They had to further break down the still large data into two categories, data which they thought they would touch after one or two days of runs are completed and after the whole simulation is going to be completed (after ten days of runs), and what they would do with these different categories of data (the access pattern of the large data).
+
+They then had to develop new application specific data reduction techniques, and encode that into the application code itself on top of ADIOS, the middleware I/O layer.
+They then had to implement a crude discovery system to understand
+which pieces of data were stored on the parallel file system and which pieces were on tape, and also which pieces were moved over to PPPL for immediate analysis by the user.
+
+This solution worked well for the first day of run but after the second day more problems occurred: the variability of the time it took to write the data made the I/O prohibitively costly, because even though the project had reserved the majority of resources at the OLCF, the file system was shared and sometimes the writing time took less 100 seconds, and at other times it was over 1,000 seconds.
+There was no method for the application or the user to monitor and predict this resource usage and to dynamically change to write less data when the I/O time was going to be longer.
+The greatest problem, however, comes after the simulation is going to be completed: even to visualize the reduced data, just to get a first look at the visualization data, it could take over a month because some of the timesteps will be on tape and other
+timesteps will be on the file system. There is no method to just look at the data quickly at some level of quality quickly and understand which regions of the simulation to focus on in more detail.
+
+The overarching idea of this proposal is that data from applications can be characterized by different
+levels of importance and different levels of importance over time, which then can be used to guide the analysis and visualization later to access the most important data and deliver some level of knowledge without waiting for access to the whole original dataset.
+Information about the importance of the data at the application level, information about resource status and performance of the multi-level storage system, and flexible data reorganization and reduction operations at all levels can be utilized to provide a quality of service and allow users to gain insight in a timely manner.
 
@@
 which have more utility (more importance).
-Clearly there are many research challenges faced with
+There are many research challenges faced with
 this approach including: 1) How do we make this simple enough so that users can
@@
  %
-  In our simple scenario, a user wants to write 1 PB of data every hour
-for checkpoint/restart, and they wish to write 500TB of data every 15
-minutes to save analysis and visualization results which have already been reduced
-using in situ reduction techniques. At this stage, they will ask the system to
-write their data, and they would like to get an estimated time which they can
+  In our simple scenario, a user wants to write 1~PB of data every hour
+for checkpoint/restart, and they wish to write 500~TB of data every 15
+minutes to save data for analysis and visualization. At this stage, they will ask the system to
+write the data, and they would like to get an estimated time from which they can
 then figure out, through a series of rules, whether they want to write out all
 of the data, postpone the write until a later time, or write a reduced amount
-of data, along with some code
-container to allow them to regenerate the data with a specified level of
+of data using in situ reduction techniques, along with some code
+container that allows for regenerating the data with a specified level of
 accuracy.  Furthermore, they realize that the analysis data will be saved on
@@
 For each piece of data, the user will specify the lifetime so
-that the data with the highest utility is kept around for the longest
-period of time, but the data which has the least utility might be migrated to
-lower layers of the storage hierarchy more quickly.
+that the data with the highest utility is kept around in fast storage for the longest
+period of time, while the data with the least utility may be the first to be migrated to
+lower layers of the storage hierarchy.
 
@@
 writing or reading at the same time. By requiring a certain quality
-of service the system must potentially lock out users with lower currency
-they want to place in the reading or writing of their data. The storage
+of service the {\bf {\color {red} system must potentially lock out users with lower currency
+they want to place in the reading or writing of their data}}. The storage
 system must also automatically move data from one tier of the storage to
@@
 
-One of the clear research objectives of this project is to understand how the
-storage system can interact with the middleware (application-aware) piece so
+One of the main research objectives of this project is to understand how the
+storage system can interact with the middleware (which is application-aware)  so
 that the data with the highest utility is kept on the fastest layers of the
-storage system for as long as specified. Clearly we need to maintain 
-QoS such that fuzzy estimates of the time it will take to write/read data from
-the different layers informs any migration/eviction decisions, and clearly it
-is critical that the storage system can manage the data across all of the
-layers. We will use learning techniques to automatically migrate data
+storage system for as long as specified. We need to maintain 
+QoS and therefore, be able to provide fuzzy estimates of the time it will take to write/read data from
+the different layers to inform any migration/eviction decisions, an it
+is also critical that the storage system can manage (discover and retrieve) the data across all of the
+layers automatically, without user intervention. We will use learning techniques to automatically migrate data
 sub-objects across the hierarchy as well as to eventually evict data which is

### 2015-07-02 narrative/proposed-work.tex
I probably messed up narrative/proposed-work.tex trying to merge in the changes to the broken version I had downloaded previously. My git-foo is not good enough to fix this properly so someone else needs to make sure this is correct. I am pretty sure it is wrong now.

@@
 
+{\bf {\color{red}Norbert will clean these pieces}}
 Our research into the storage system and middle ware layers must address
@@
 
+\begin{enumerate}
 \item Add plug-in architecture to Sirocco to support selective data compression
@@
 
-% \input{narrative/data-desc}
+%\input{narrative/data-desc}
+
+{\bf {\color{red}Scott will integrate these pieces}}
 \input{narrative/data-reduction} % includes data-reorg and the api specs move to the above section -
 \input{narrative/data-reorg}
-% \input{narrative/qos-app}
+
+%\input{narrative/qos-app}
+{\bf {\color{red}Hasan will integrate these pieces, Manish after Hasan}}
 \input{narrative/placement}
-% \input{narrative/policies}
-% \input{narrative/apis}
+\input{narrative/migration}
+\input{narrative/eviction-metrics}
+
+%\input{narrative/policies}
+%\input{narrative/apis}
 
@@
 
-\input{narrative/migration}
+{\bf {\color{red}Gary will integrate these pieces}}
 \input{narrative/resource-mgmt}
 \input{narrative/qos-storage}
-% \input{narrative/discovery}
-\input{narrative/eviction-metrics}
+%\input{narrative/discovery}
+{\bf {\color{red}Jay  will integrate these pieces}}
 \input{narrative/naming-service}

### 2015-07-02 narrative/intro.tex
cleaned up table a bit. Will add things later.

@@
 %\multicolumn{2}{|c|}{Collaboration} & 
-\multicolumn{2}{|p {2.4cm}|}{Approaches to improve the ability of SSIO software to support checkpoint/restart;} & 
+\multicolumn{2}{|p {5.5cm}|}{Approaches to improve the ability of SSIO software to support checkpoint/restart;} & 
 General piece we are doing for this... & 
 \S\ref{sec:?}, \S\ref{sec:??}, \\\hline
-1 & Data \hspace{1cm} abstractions &What will we do& \S\ref{sec:?}, \S\ref{sec:?}, \S\ref{sec:?} \\\hline
-2 & Mapping science data models onto hierarchical storage &What we are doing & \S\ref{sec:?}, \S\ref{sec:?} \\\hline
+1 & Data abstractions & What will we do & \S\ref{sec:?}, \S\ref{sec:?}, \S\ref{sec:?} \\\hline
+2 & Mapping science data models onto hierarchical storage & What we are doing & \S\ref{sec:?}, \S\ref{sec:?} \\\hline
 3 & Mechanisms for data movement across the storage hierarchy & what are we doing. & \S\ref{sec:?} \\\hline
 4 &  Extend the interaction of IO middleware with the resource management system & What we are doing & \S\ref{sec:?}, \S\ref{sec:?}  \\\hline
-5 & Design and implement of IO middleware architectures  &What we are doing & \S\ref{sec:?}, \S\ref{sec:?}, \S\ref{sec:?} \\
+5 & Design and implement of IO middleware architectures  & What we are doing & \S\ref{sec:?}, \S\ref{sec:?}, \S\ref{sec:?} \\
 \hline

### 2015-07-03 narrative/intro.tex
some small changes -- a work in progress....

@@
 large scale scientific computation, is even more relevant today as we prepare
-for scientific simulation at the exascale. Hamming was a pioneer in the fields
-of coding theory, scientific computing and information theory who would
-instantly sympathize with the central objective of this proposal--minimizing
-the time to knowledge.
+for scientific simulation at exascale. 
+The central objective of this proposal, ``minimizing the time knowledge" from 
+scientific computing, aligns perfectly this sentiment. 
+%Hamming was a pioneer in the fields of coding theory, scientific computing and 
+%information theory who would instantly sympathize with the central objective of 
+%this proposal--minimizing the time to knowledge.
 
@@
 where the user and the storage system work together to achieve the best
-possible data quality and performance given data features, storage tier
+possible utility and performance given data features, storage tier
 characteristics, and current system state. We will investigate novel techniques
@@
 application-guided data reductions/transformations to address capacity and
-bandwidth bottlenecks while constraining the error to be within user provided
+bandwidth bottlenecks, while constraining errors to be within user provided
 bounds.
 %
+
 We will address the associated I/O and storage challenges in the context of
@@
 scientific processes, while demonstrating the validity of our approach in key
-DOE domains. We will research techniques for creating a Scalable Storage
-Software Infrastructure, integrating services from the middleware layer with
-Storage System features to support the novel strategy of distributing data both
-horizontally and vertically across the storage system.  The negotiation between
+DOE domains. We will research techniques for creating and using a Scalable 
+Storage Software Infrastructure, and integrating services from the middleware 
+layer with Storage System capabilities to support novel strategies for distributing 
+data both horizontally and vertically across the storage system.  Negotiation between
 these layers will be provided as a fundamental service, allowing users to
-easily save ``the best'' amount of data, automatically placed throughout the
+easily save ``the best'' amount of data, which is automatically placed throughout the
 entire storage stack.
@@
 application. Both approaches have been successful to date, but are unlikely to
-overcome the remaining obstacles in reaching exascale. Instead of trying to
-optimize throughput, we will seek to reduce the time to knowledge. This is the
-most significant metric for scientific applications where the desired outcome
-is not storing data, but rather executing a knowledge extraction process on the
-data. Moreover, we aim to perform the optimization not for a single
+overcome the remaining obstacles in reaching exascale. 
+Instead of trying to optimize throughput, we will seek to reduce the time to knowledge. 
+This is the most significant metric for scientific applications where the desired outcome
+is not storing data, but rather executing a knowledge creation and extraction processes 
+on the data. 
+Moreover, we aim to perform the optimization not for a single
 application, but rather for workloads within a multi-user multi-application
@@
 into the SSIO system so that data can be recast from a bunch of bytes to
-information which the system can then place and optimize the time to gain
+information, which the system can then place and optimize the time to gain
 knowledge from exascale simulations.  We first motivate this by a problem our
@@
 
-This solution worked well for the first day of run but after the second day more problems occurred: the variability of the time it took to write the data made the I/O prohibitively costly, because even though the project had reserved the majority of resources at the OLCF, the file system was shared and sometimes the writing time took less 100 seconds, and at other times it was over 1,000 seconds.
-There was no method for the application or the user to monitor and predict this resource usage and to dynamically change to write less data when the I/O time was going to be longer.
-The greatest problem, however, comes after the simulation is going to be completed: even to visualize the reduced data, just to get a first look at the visualization data, it could take over a month because some of the timesteps will be on tape and other
-timesteps will be on the file system. There is no method to just look at the data quickly at some level of quality quickly and understand which regions of the simulation to focus on in more detail.
+This solution worked well for the first day of run but after the second day more problems occurred: the variability in the time that it took to write the data made the I/O prohibitively expensive, because even though the project had reserved the majority of the resources at the OLCF, the file system was shared and sometimes the writing time took less 100 seconds, and at other times it was over 1,000 seconds.
+There was no method for the application or the user to monitor and predict this resource usage and to dynamically adapt to write less data when the I/O time was going to be longer.
+The greatest problem, however, comes after the simulation   completes: even to visualize the reduced data, just to get a first look at the visualization data, it could take over a month because some of the timesteps will be on tape and other timesteps will be on the file system.
+There is no method to just look at the data quickly at some level of quality quickly and understand which regions of the simulation to focus on in more detail.
 

### 2015-07-03 narrative/data-reduction.tex
Data refactoring...

@@
 
-It is useful to classify scientific data into two basic categories: structured
+It is critical that users give us more information about writing and reading from the SSIO layer
+in order to not only gain performance and predictability, but to also guide them to write and read
+the best possible data in a given period of time. In our example from C. S. Chang, along with most
+of the other scientist we work with, there is a prioritization of data, and this information must be
+sent over to our SSIO layer so that the best possible data is delivered to and from the SSIO layer, 
+and these decisions often involve refactoring (re-organizing and reducing the data) . In this section we
+describe the common methods to reduce data and to re-organize data. Since there are infinite possibilities 
+we need to understand what kind of data re-organization will work and when will it work?  We also need to 
+understand if it would take too much time and resources to perform the ``best methods'' and how can we manage the
+cost of the re-factoring of data compared to the eventual gains we will get?  Finally, another critical research question
+is if the scientist will loose too much value from the data during the refactoring process, and whether or not we can
+control this through error bounds?
+
+The first step of re-factoring the data is to understand how much this data can be reduced.  Therefore, 
+it is useful to classify scientific data into two basic categories: structured
 data (in the sense that it satisfies a known or relatively simple model) or
@@
 
-We propose a fresh approach to tackle the data overload problem head-on through
-using preprocessors in which information is not simply stored in the form of
-raw data, but rather in the form of appropriate parameterized algorithms. A
-priori information on the nature of the data is used to choose the basic
-algorithm while the actual values of the data are used to tune the parameters
-used in the reconstruction procedures. We shall refer to this paradigm as an
-Auditor.
+Once we realize that data can be reduced from many different techniques, we understand
+that with more application knowledge, we can reduce the total data size much more than without
+knowledge, so we need to be able to design the SSIO layers to be able to have generic
+data reduction techniques, and generic data re-organization techniques. But we need to system
+to allow give us more semantic knowledge of the data so that we can either: 1) Use an approach
+which can audit the calculation and save values of the objects when the auditor is off more than a given
+amount, and we can also save the auditor in a docker container, which can then be executed as part of 
+SSIO software layer.
 
-Scientific data is sometimes described as essentially incompressible owing to
-noise and other random effects present in the data, stemming from the numerical
-algorithm, software implementation, system specific behavior etc. The reasoning
-behind the description lies in the mathematical fact that it is impossible to
-achieve lossless compression on completely random data sets \cite{Gray:book}. Even
-recent state of the art lossless compression approaches \cite{GomezCappello} are only
-able to obtain 15\% improvement in lossless compression ratio while reducing
-the time spent in compression of floating point data by one half in some cases.
-Nevertheless, these limitations apply only if one ignores two key features:
-Firstly, randomness in scientific data generally occurs only below some
-threshold related to the accuracy of the measurement apparatus or the numerical
-algorithm. Secondly, lossy compression is perfectly acceptable provided the
-loss level falls below this threshold.
+Recognizing access patterns in an application is a key to potentially reducing
+future file read time. Scientific applications often read and analyze data alternately,
+thus by overlapping the two phases with prefetching can significantly
+reduce the overall execution time of the application. Accurate prefetching can
+be achieved with access pattern analysis. This requires an SSIO layer which can
+prefetch data in a asynchronous fashion to take the bytes on the file.
 
-Data compression broadly consists of three basic steps: (1) Pre-processing to
+There are many methods to re-organize the data, and to reduce the data and in this section we will list many
+different known methods.  We then list our different approaches to address the many research challenges using
+many well known methods, and a new method that we have been working on, known as an ``auditor'' which can be
+used ``in situ'' with the calculation during the data generation, and then used to dramatically reduce the impact on the SSIO
+layer for both reading and writing.
+
+\paragraph{State of the art:}
+{\bf Data Reduction Techniques} Data compression broadly consists of three basic steps: (1) Pre-processing to
 extract as much structure as feasible. This step is variously referred to as
@@
 standard compression approaches such as FPC \cite{BurtscherFPC} in which an XOR
-operation against the previous data is used as a pre-processor. The fact that
-even such a crude pre- processor is able to deliver reasonable results augurs
-well: Even a relatively crude Auditor can be expected to significantly
-out-perform the XOR and other similar approaches.
+operation against the previous data is used as a pre-processor.
+
+Our previous work~\cite{lakshminarasimhan2011compressing},~\cite{lakshminarasimhan2011compressing},~\cite{gong2012multi},
+~\cite{jenkins2012byte},~\cite{gong2013parlo},~\cite{boyuka2014transparent},~\cite{tang2014improving} has focused on many
+of these well known techniques to not only reduce the data but also re-organize the data. Previously we have examined:
+
+\begin{itemize}
+\item Precision based re-organization. This is where the most significant bytes of the data are
+all group together from each object. This data generally will have a higher utility then the data with
+the least significant bytes.  This process has the following three steps which need to be
+efficiently implemented when data gets placed to the storage system and from the storage 
+system. 
+%
+First, the data needs to be re-arranged which involves memory operations and needs
+to be done in-situ. This operation involves no communication.  Data needs to be copied, so
+the memory requirements are increased. At most this requires two copies an individual dataset
+in memory, and we will investigate techniques to allocate and deallocate this memory if the user
+will specify that the data will be overwritten after it is written to the storage system. This is often
+the case for many of the data quantities written form the simulation, but there are many cases
+where, for example, we want to write all of the particles from a Particle In Cell (PIC) simulation. 
+Since the particles will be used later in the calculation, we need to duplicate the storage. Our
+observation with working with the XGC1, GTC, Warp, and PiconGPU simulations (all leadership
+class simulations) is that we can temporary increase the storage of the particles, and then 
+release them since the temporary arrays used in the calculations are often freed when a PIC
+iteration is finished.. 
+%
+\item Frequency-based re-organization. 
+Another common technique to classify the importance of data is to first re-organize the data
+in frequency space and then specify that as the frequency increases the importance of data
+decreases.   Common technique which is used for streaming data, and data reduction techniques such as JPEG-2000~\cite{jpeg2000}
+is to  code streams which have regions of interest that offer several mechanisms to support spatial random access or region of interest access at varying degrees of granularity.
+It is possible to store different parts of the same data using different quality.
+%
+This allows us to place the lowest frequency pieces in the fastest storage, and
+the highest frequency either on the slowest storage tiers, or if the data sizes are prohibitively  
+costly, not even write out the highest frequency. This binning of data in frequency space can
+take as an input (storage bandwidth knowledge, storage size limitations, and user intentions
+of how long the majority of the data will be frequently accessed.
+
+In order to fully take advantage of frequency based re-organization we have devised a scheme where the data pre-conditioner first sorts the data
+in bins of 1024 words, and then we maintain the sort order, and then we used wavelets and spline fits to then re-organize the values.
+This re-organization allows us to understand that the resultant data is smoother now, and can be reduced in each one of the bins.
+\end{itemize}
+
+\paragraph{Proposed research approach:} 
+We will be looking at a multi-throng approach to data-refactoring. One approach allows the user to specify information about their data
+to indicate if the data is spatially varying, space-and time varying, just time varying. This allows us to store information in NVRAM on each
+node (when and where necessary) to later preform compression in both space and time . We will examine the different techniques for
+compression and re-organization along with examine techniques to use an error-estimator (gradients) to create a multi-resolution view of the
+data which can then be sampled, and placed in different quantiles of importance. These will then be mapped over into the different layers of 
+the storage hierarchy.  We will also investigate technique which can use ``plug-ins'' for detection of features and store higher resolution data
+around the feature and lower resolution values which are further away from the features. This re-organization must occur without increasinig the
+total amount of storage than what the users intended. We will also investigate the time it takes to re-organize the data and use the multi-layer
+approach to staging to understand where the data gets reduced and where it gets re-organized. Open questions are: 1) do we move the
+data over to other nodes to then reduce and re-organize, 2) do we first reduce the data and then later re-organize this data on different nodes?
+3) do we use different cores on a node to do data-refactoring. Similarly we will examine the cost of reading the data. It is clear that if we only
+read in the high-importance data it should improve the overall time to move the data from the SSIO layer to the application, but what happens
+when users want all of the data? There is an obvious penalty of having to take data in a hierarchical layout and then place this back to
+a single-layer view.  Additional research questions which must be addressed: Will some types of data (PIC data, finite element data, ...) be
+be conducive  for data-refactoring than others? How much semantic-domain knowledge will we need to get a dramatic improvement of the storage
+system over not doing anything special?
+
+We will measure these methods using the fusion XGC simulation and the corresponding analysis routines which they use for their largest
+simulations. We will also work with Jereon Tromp's SPECFM3D code, and the fusion GTC code from Z. Lin. In all cases
+we will measure their impact from their current simulations, and their projected simulations and analytics on exascale machines. Since 
+all three of these codes are part of the OLCF CAAR~\cite{CAAR} project for early access, there is strong belief that the simulations will
+be close to exascale simulations and we will be able to run these on prototypes of our SSIO layer to examine the full impact of data-refactoring.
+
+\paragraph{HIgh Impact Challenges:}
+Since the greatest form of data reduction on the SSIO layer would be to only write the code, and the input data,
+and then recalculate all of the results, we want to understand how to write a minimal set of information, and then place an auditor 
+code in a docker container so that data can be re-generated with an approximation of the calculation, but only valid over a limited amount 
+of time.This new approach to tackle the data overload by
+using preprocessors in which information is not simply stored in the form of
+raw data, but rather in the form of appropriate parameterized algorithms. A
+priori information on the nature of the data is used to choose the basic
+algorithm while the actual values of the data are used to tune the parameters
+used in the reconstruction procedures. We shall refer to this paradigm as an
+Auditor.
+
+We will examine the use of several auditors for the simulations which we will work with. One approach is to use a fluid-code to audit a kinetic
+code, which can be a good approximation over short periods of time. Once there is a large deviation (specified by the user) from the auditor
+to the original calculation, the data will be saved. This is a local operation, since the deviations will grow differently in space and in time.

### 2015-07-04 narrative/placement.tex
start to reshape the proposal
filled the lines for easier reading
merged all 3 texts into one .tex file

@@
-\subsubsection{Data Placement and Movement}
+\subsection{Data Placement and Movement}
 
-The application workflows targeted in this proposal generate very large amounts of data, and this data must be processed and analyzed before potential insights from the simulations can be realized.
-In-situ/in-transit data processing using in-memory data-staging approaches have been effectively used to address data-related challenges and enable simulations workflows on current high end systems -- the key idea is to use compute and storage resources on the high-end system to process the data close to where it is generated.
-However, given current architectural trends, indicate that emerging systems will have increasing numbers of cores per node and correspondingly decreasing amounts of DDRAM memory per core as well as decreasing memory bandwidth.
-Architectural trends also indicate deeper memory hierarchies and multilevel storage structures.
-As a results, to address the increases data volumes and data processing requirements requires that these levels of storage be effectively utilized and exposes as a uniform staging abstraction that can be used by the application workflows.
-A key requirements is application-driven runtime mechanisms for dynamically managing data placement across the layers of the distributed storage hierarchy, coordinating data movement and data sharing between the components of the application workflow so as to maximize its utility to the application and reduce access costs.
+The application workflows targeted in this proposal generate very large
+amounts of data, and this data must be processed and analyzed before
+potential insights from the simulations can be realized. In-situ/in-transit
+data processing using in-memory data-staging approaches have been
+effectively used to address data-related challenges and enable simulations
+workflows on current high end systems -- the key idea is to use compute and
+storage resources on the high-end system to process the data close to where
+it is generated. However, given current architectural trends, indicate that
+emerging systems will have increasing numbers of cores per node and
+correspondingly decreasing amounts of DDRAM memory per core as well as
+decreasing memory bandwidth. Architectural trends also indicate deeper
+memory hierarchies and multilevel storage structures. As a results, to
+address the increases data volumes and data processing requirements requires
+that these levels of storage be effectively utilized and exposes as a
+uniform staging abstraction that can be used by the application workflows. A
+key requirements is application-driven runtime mechanisms for dynamically
+managing data placement across the layers of the distributed storage
+hierarchy, coordinating data movement and data sharing between the
+components of the application workflow so as to maximize its utility to the
+application and reduce access costs.
 
-\paragraph{State of the art:} NAND flash-based devices are being increasingly used in HPC storage systems at different levels and for a variety of purposes.
-Several studies such as \cite{multitier}, \cite{sc10li} have investigated using compute node local SSDs as storage buffers, for example, to temporarily cache checkpoint data to support recovery from failures.
-Research efforts have also explored using deep memory devices for data analysis.
-Active Flash \cite{activeflash} proposes in-situ scientific data analysis by directly executing data analysis tasks on emerging storage devices.
-Minerva \cite{minerva} extends the conventional SSD architecture using a FPGA-based storage controller to offload data or I/O intensive application code to the SSD to accelerate data analysis.
-However, these solutions cannot deal with coupled application workflows, where the data exchange and access patterns can be both complex and dynamic.
-Furthermore, some of these solutions require hardware modifications or special access priorities to the HPC systems, which can limit their use to specific resources.
+\paragraph{State of the art:} NAND flash-based devices are being
+increasingly used in HPC storage systems at different levels and for a
+variety of purposes. Several studies such as \cite{multitier}, \cite{sc10li}
+have investigated using compute node local SSDs as storage buffers, for
+example, to temporarily cache checkpoint data to support recovery from
+failures. Research efforts have also explored using deep memory devices for
+data analysis. Active Flash \cite{activeflash} proposes in-situ scientific
+data analysis by directly executing data analysis tasks on emerging storage
+devices. Minerva \cite{minerva} extends the conventional SSD architecture
+using a FPGA-based storage controller to offload data or I/O intensive
+application code to the SSD to accelerate data analysis. However, these
+solutions cannot deal with coupled application workflows, where the data
+exchange and access patterns can be both complex and dynamic. Furthermore,
+some of these solutions require hardware modifications or special access
+priorities to the HPC systems, which can limit their use to specific
+resources.
 
-\paragraph{Proposed research approach:} In this project we will utilizes deep memory hierarchies and multilevel storage structures to support staging-based in-transit data management and processing for data-intensive scientific workflows at extreme scale.
-Specifically, we will address the coordination and data access/sharing requirements of application workflows, as well as the  online data management requirements such as for example, in-transit online data analysis,  check-pointing,  data reduction.
-We will also deal with the complexities associated with data placement of data and movement in the multiple layers and components of the storage hierarchy.
-The key components of our research include:
+\paragraph{Proposed research approach:} In this project we will utilizes
+deep memory hierarchies and multilevel storage structures to support
+staging-based in-transit data management and processing for data-intensive
+scientific workflows at extreme scale. Specifically, we will address the
+coordination and data access/sharing requirements of application workflows,
+as well as the online data management requirements such as for example,
+in-transit online data analysis, check-pointing, data reduction. We will
+also deal with the complexities associated with data placement of data and
+movement in the multiple layers and components of the storage hierarchy. The
+key components of our research include:
 
-{\em Application hints:}  Application hints can provide additional information to the runtime based on the users' knowledge of the application workflow as well as on past experiences, for example, possible data access patterns, nature of regions of interest, and domain specific information.
-This application level information will be added to the specification of the workflow and can provide advisory inputs to efficiently help the runtime make smarter decisions about the placement and/or movement of data objects and the scheduling and execution of tasks, as well as to effectively manage trade-offs between for example, power consumption and time-to-solution.
+{\em Application hints:} Application hints can provide additional
+information to the runtime based on the users' knowledge of the application
+workflow as well as on past experiences, for example, possible data access
+patterns, nature of regions of interest, and domain specific information.
+This application level information will be added to the specification of the
+workflow and can provide advisory inputs to efficiently help the runtime
+make smarter decisions about the placement and/or movement of data objects
+and the scheduling and execution of tasks, as well as to effectively manage
+trade-offs between for example, power consumption and time-to-solution.
 
-{\em Data utility:} We will formulate the concept of utility (from an application perspective) that is assigned data objects to quantify the benefits of placement and/or movement actions to application.
-These utilities may be based on the importance of the data objects, the frequency with which they are accessed, etc., and are used during runtime data management.
-For example, data objects will higher utility value may be placed closer (e.g., in faster memory) to the accessing application.
-Similarly, when it is time to evict data objects from a higher level storage layer, the decision about which object to evict may take the utility into account.
+{\em Data utility:} We will formulate the concept of utility (from an
+application perspective) that is assigned data objects to quantify the
+benefits of placement and/or movement actions to application. These
+utilities may be based on the importance of the data objects, the frequency
+with which they are accessed, etc., and are used during runtime data
+management. For example, data objects will higher utility value may be
+placed closer (e.g., in faster memory) to the accessing application.
+Similarly, when it is time to evict data objects from a higher level storage
+layer, the decision about which object to evict may take the utility into
+account.
 
-{\em Autonomic in-transit data management:} Efficiently placing data objects vertically across multiple memory hierarchy layers and storage levels as well as horizontally across different storage nodes requires balancing competing objectives.
-In this project we will explore autonomic data management strategies that can  evaluate utility/cost tradeoffs and appropriately place/move data objects at runtime.
-The autonomic approach will leverage both user hints and information gathered at  runtime (e.g., runtime data access history, network topology, etc.) to place data object.
+{\em Autonomic in-transit data management:} Efficiently placing data objects
+vertically across multiple memory hierarchy layers and storage levels as
+well as horizontally across different storage nodes requires balancing
+competing objectives. In this project we will explore autonomic data
+management strategies that can evaluate utility/cost tradeoffs and
+appropriately place/move data objects at runtime. The autonomic approach
+will leverage both user hints and information gathered at runtime (e.g.,
+runtime data access history, network topology, etc.) to place data object.
 
-{\em Runtime tracking and estimation:}
-Autonomic data placement and/or movement will leverage the information gathered at runtime together with predictive analysis requirements.
-This information may include runtime spatial-temporal data access patterns, physical network topology information, and etc., which will be tracked and gathered at runtime, and then be used to estimate and predict the behavior of coupled applications in the rest of simulation-time.
-A key research aspect here will be to anticipate accesses, for example using learning techniques or using application knowledge, and use this information to prefetch data objects.
-For example, we will identify and catalogue key phases and access patterns in the target workflows and then use it to characterize patterns at runtime and trigger appropriate place/movement actions.
+{\em Runtime tracking and estimation:} Autonomic data placement and/or
+movement will leverage the information gathered at runtime together with
+predictive analysis requirements. This information may include runtime
+spatial-temporal data access patterns, physical network topology
+information, and etc., which will be tracked and gathered at runtime, and
+then be used to estimate and predict the behavior of coupled applications in
+the rest of simulation-time. A key research aspect here will be to
+anticipate accesses, for example using learning techniques or using
+application knowledge, and use this information to prefetch data objects.
+For example, we will identify and catalogue key phases and access patterns
+in the target workflows and then use it to characterize patterns at runtime
+and trigger appropriate place/movement actions.
 
-\paragraph{Preliminary work and results:} Our recent work \cite{tongipdps15} explored a two-tiered staging method that spans both DRAM and solid state disks (SSD).
-It allows us to support both code coupling and data management for data intensive simulation workflows in a loosely coupled manner.
-In addition, our application-aware adaptive data placement has demonstrated the effectiveness of using user provided access information as ``hints'', and runtime data access pattern history to improve data access efficiency and overall end-to-end execution.
-Besides, additional data placement optimization \cite{qiansc15} that leverages physical network topology along with data access patterns has also demonstrated lower data access costs and good scalability.
-Our work have been deployed and proven with real applications, e.g., coupled combustion simulations (DNS-LES) that are part of the ExaCT Exascale Co-design Center and coupled Fusion simulations (XGC0-XGCa) that are part of the EPSI SciDAC on current high-end systems such Titan at ORNL.
+\subsection{Migration}
+
+One important research question in this proposal is whether and how a liberated
+Hierarchical Storage Manager designed for online use, such as Sirocco, might be
+employed in supporting the higher level concepts discussed previously.
+Conceptually, in this effort, data is accessed by the middleware IO libraries
+at various tiers in the lower-level storage hierarchy. While Sirocco does
+autonomously group the distributed set of media into like groups and manage
+movement of data between these groups the various motivations to do so include
+both the traditional ones found in classical HSM (Hierarchical Storage
+Management) as well as the novel.
+
+Sirocco groups like media into aggregated volumes on a service node, and like
+service nodes with like media into pseudo-tiers. It accomplishes both using
+attributes such as latency to first byte, bandwidth to and from the media, and
+some function representing the resilience capabilities of the attached media.
+In addition to these static attributes, when grouping servers into
+pseudo-tiers, Sirocco also may employ more dynamic attributes such as CPU load
+and the rate-of-change of the media use.
+
+Grouping into pseudo-tiers is not a formal, nor rigid, concept within Sirocco,
+at present. In classic HSM the tiering is often a formally expressed concept,
+embodied both in the architecture and the implementation. In Sirocco, though,
+it is not directly expressed. Instead, a client or server that attempts to make
+a decision relative to migration and staging uses the attributes discussed
+above to constrain the choice of candidates. For instance, a server acting as
+an off-node writeback cache for a compute node might only have access to
+volatile dynamic RAM for use as storage media and a client might deposit data
+that is marked for persistent media as a resilience constraint. At some point,
+then, such a server must copy or move the data to another Sirocco server with
+media matching, or exceeding, the resilience constraint. At that point in time,
+all servers managing appropriate persistent media are candidates but
+second-order attributes, such as latency and bandwidth of a candidate's managed
+media would tend to dominate the ranking. Thus, the concept of a ``tier''
+becomes an artifact of decision rather than a formal concept.
+
+Further complication arises when a Sirocco server is forced to make choices for
+capacity reasons. The Sirocco architecture dictates that a server function as a
+victim cache. As media managed by a server fills the server will begin to
+manage for capacity. In order to do so, currently, it need only verify or make
+a copy of the data on another server that meets the resiliency constraint of
+the held data prior to removing the local copy and in the current
+implementation the server makes no attempt either to remember it ever had the
+data or to notify others about what it is doing.  others that it has removed
+the local copy or remember that it has
+
+Taken together then, the above at least complicates the higher level research
+questions discussed in this proposal, potentially even renders some
+unachievable since a few involve determinism. We are forced, then, to consider
+either or both mitigating management policies within and between Sirocco
+servers as well as explicitly allowing clients to participate in capacity
+management decisions, for a time at least, on the relevant Sirocco servers.
+
+Our research intent involves short term and long term activities. In the short
+term we will modify the relevant clients to provide hit and miss related data
+and associated activities and explore an enhancement to our storage servers
+that would provide some, hopefully tightly, bounded estimate about time to
+first byte for a potential read operation. In the long term we hope that an
+augmented client-server API providing usage hints will allow us to modify local
+server policy algorithms to more effectively manage it's media and communicate
+to other servers holding copies of the relevant data how they might best manage
+theirs. In the event that this approach proves ineffective at meeting the
+overall approach described in this proposal we will explore modifying Sirocco
+servers to provide notifications about locations of the relevant data as they
+migrate and stage as well as supporting client-directed pinning of location,
+for a period at least. Client-directed pinning could introduce potentially
+serious negative impacts to Sirocco's overarching goals though, since the
+fundamental design of Sirocco assumes much about opaque and unrestricted data
+movement, so this is contemplated only as a last resort.
+
+\subsection{Eviction Metrics}
+
+% The title might have to change...
+
+\paragraph{Background:} There is a spectrum of approaches on how
+to leverage a memory and storage hierarchy: On one end of the
+spectrum is to copy or move data from one tier to another, the
+closer the tier to the hierarchy's top, the faster the access to
+the data it contains. On the other end of the spectrum is to use
+higher tiers exclusively for auxiliary data that to some extent
+represents the actual data on the lowest tier. Examples of auxiliary
+data are views, indices, lossy compressions, lower resolution data,
+and summary data (sometimes also known as metadata). There are a
+number of drivers that push approaches to one or the other extreme:
+(1) if a high tier has sufficient room for an application's working
+set, approaches that move the working set into that tier will perform
+well; Conversely, if a high tier very rarely captures working sets,
+it is better to fill it with auxiliary data that speeds up access
+to the working set captured by a lower tier; working sets only
+reduce capacity misses but the concept can be generalized to the
+likelihood of having the right data in place, i.e. proactively
+moving data to high tiers accessed in predictable patterns reduces
+compulsory misses as well; (2) if the application generates parallel
+read/write requests and strong consistency semantics are important,
+approaches that use auxiliary data to speed up access to persistent
+shared data might yield better performance than approaches that
+require coherence overhead to keep all copies on higher tiers
+consistent; (3) if the application's access patterns are mismatched
+with low tier access characteristics, a combination of approaches
+that use auxiliary data and move data between tiers in order to
+convert access patterns into a better match will perform well.
+
+% Any other drivers?
+
+\paragraph{Approach:} This spectrum enables interesting strategies
+to manage space pressure on high tiers: instead of just evicting
+pieces of data (blocks, pages, objects, files), one can free space
+by ``thinning'' data so that accesses within more limited views or
+lower resolutions can still occur without misses while requests
+outside of views or higher resolutions can leverage the information
+stored in auxiliary data. Conversely, data on a high tier can be
+``enriched'' (e.g. turned into a higher resolution) if there is
+value in doing so.
+
+\paragraph{Related Work:} Mention Stanford's Legion system.
+
+\paragraph{Challenge:} The decision on when to thin and when to
+enrich data on a particular level depends on \emph{access patterns}.
+Approaches that leveragee access patterns in application at national
+labs has shown to greatly improve performance and reduce the amount
+of overhead required for data management~\cite{he:hpdc13}. The
+challenge is for applications to communicate these access patterns
+to the I/O stack layers that manages tiers outside of the address
+space. In the proposed work we will develop an abstraction for
+\emph{locality} that allow applications or runtime profilers to
+describe locality properties of access patterns, and implement
+services that leverage locality properties to dynamically determine
+the value of thinning or enriching data at a particular tier.
+Consider the example of a road navigation system: the locality
+properties of each navigation client are based on physical constraints
+of speed, direction, likely resolution of map, and the fact that
+clients almost never go offroad.
+
+\paragraph{Preliminary work and results:} Our recent work \cite{tongipdps15}
+explored a two-tiered staging method that spans both DRAM and solid state
+disks (SSD). It allows us to support both code coupling and data management
+for data intensive simulation workflows in a loosely coupled manner. In
+addition, our application-aware adaptive data placement has demonstrated the
+effectiveness of using user provided access information as ``hints'', and
+runtime data access pattern history to improve data access efficiency and
+overall end-to-end execution. Besides, additional data placement
+optimization \cite{qiansc15} that leverages physical network topology along
+with data access patterns has also demonstrated lower data access costs and
+good scalability. Our work have been deployed and proven with real
+applications, e.g., coupled combustion simulations (DNS-LES) that are part
+of the ExaCT Exascale Co-design Center and coupled Fusion simulations
+(XGC0-XGCa) that are part of the EPSI SciDAC on current high-end systems
+such Titan at ORNL.

### 2015-07-04 narrative/placement.tex
started on the data placement section

@@
 \subsection{Data Placement and Movement}
+\newcommand{\Sir}{Sirocco}
 
 The application workflows targeted in this proposal generate very large
-amounts of data, and this data must be processed and analyzed before
-potential insights from the simulations can be realized. In-situ/in-transit
-data processing using in-memory data-staging approaches have been
-effectively used to address data-related challenges and enable simulations
-workflows on current high end systems -- the key idea is to use compute and
-storage resources on the high-end system to process the data close to where
-it is generated. However, given current architectural trends, indicate that
-emerging systems will have increasing numbers of cores per node and
-correspondingly decreasing amounts of DDRAM memory per core as well as
-decreasing memory bandwidth. Architectural trends also indicate deeper
-memory hierarchies and multilevel storage structures. As a results, to
-address the increases data volumes and data processing requirements requires
-that these levels of storage be effectively utilized and exposes as a
-uniform staging abstraction that can be used by the application workflows. A
-key requirements is application-driven runtime mechanisms for dynamically
-managing data placement across the layers of the distributed storage
-hierarchy, coordinating data movement and data sharing between the
-components of the application workflow so as to maximize its utility to the
-application and reduce access costs.
+amounts of data, which needs to be processed and analyzed before
+potential insights from the simulations can be realized. 
+In-situ and in-transit data staging approaches can alleviate some of the
+burden on storage, leveraging available compute and memory resources on high
+end systems to process the data close to where it is generated. 
+However, given current architectural trends, emerging systems will have a
+vastly great gulf between the computing capability (the capability to
+generate data) and the storage capability (the capability to store the
+generated data). Moreoever, the complexity and heterogeniety of the
+architecture will introduce new tiers in the storage system. Instead of
+being a flat, roughly homogenous space in which data can be stored, the new
+storage hierarchy will be multi-tiered and greatly heterogeneous.
+
+To take advantage of these architectural trends, we propose a triplet of
+data placement strategies. First, we will utilize the faster layer of the
+storage hierarhcies, such as DRAM, NVRAM and remote NVRAM, as the target for
+the initial write from an application. To mitigate the space concerns, we
+will use the techniques described in Section~\ref{sec:refactoring} to
+identify and partition the segments of data with the highest utility for the
+application and use case. The rest of the data will be stored on the
+parallel file system, as is done today. Second, we will utilize the
+migration of data supported by the existing implementation of \Sir to move the
+data to an appropriate tier. We will explore how migration can be guided by
+a combination of the utility function defined by the application and the
+resources avialable to the storage system. Knowing the specific structural
+and semantic information (including relationships between data segments)
+about the data will enable migration to partition the data further to
+optimize access to data for reader. 
+Finally, we will explore policies that combine the knowledge of data, its
+utility to users, the cost of maintaining data in long term storage and
+other metrics to decide how to purge irrelevant data from the storage
+system.
+
+Thus we propose to advance the state of the art in three connected areas:
+(1) Data placement,
+(2) Migration,
+and (3) Purging.
+
+\subsubsection{Data Placement}
+\label{sec:data-placement}
+
+Our past work on data
+staging\cite{tongipdps15,qiansc15}\cite{docan2012dataspaces}\cite{abbasi2010datastager}
+in HPC systems, coupled with the experience in developing high performance
+I/O transports for ADIOS \cite{liu2014hello,lofstead2008flexible} has
+provided us with insights on how to manage the complex I/O requirements for
+exascale applications, particularly in the face of the aforementioned
+multi-tier storage hierarchy. A key requirements is application-driven
+%
+runtime mechanisms for dynamically managing data placement across the layers
+of the distributed storage hierarchy,
+%
+coordinating data movement 
+%
+and data sharing between the components of the application workflow so as to
+maximize its utility to the application and reduce access costs. Our
+approach to data placement will leverage the increased set of knowledge
+about the data available to the middleware and storage system, potentially
+allowing for greater optimizations as compared to other storage solutions.
 
@@
 
-\paragraph{Proposed research approach:} In this project we will utilizes
+\paragraph{Proposed research approach:} In this project we will utilize
 deep memory hierarchies and multilevel storage structures to support
@@
 
-\subsection{Migration}
+\subsubsection{Migration}
+\label{sec:migration}
 
@@
 
-\subsection{Eviction Metrics}
+\subsubsection{Purging}
+\label{sec:purging}
 
@@
 such Titan at ORNL.
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "../proposal"
+%%% End:

### 2015-07-04 narrative/placement.tex
reorged the placement subsubsection

@@
 allowing for greater optimizations as compared to other storage solutions. 
+The key components of our research include
+
+\paragraph{Application Hints:}
+As described in Section~\ref{sec:refactor} one of the main components of our
+proposed storage system is the ability to refactor and reduce data as it is
+generated, and to reorgnized and regenerate the data as it is accessed. We
+carry this principle into the placement and movement of data by allowing
+applications to define hints and policies that guide \textbf{what} data is
+placed \textbf{where}. We will explore the use of application hints in two
+distinct areas. Firstly, we will study the challenges and tradeoffs of either
+augmenting the I/O interface with hints or allowing the additional of an
+external specification that defines the use case. Our experience with
+developing modern I/O interfaces has shown that both techniques have value,
+and we will investigate the set of hints that are embedded in the
+application code vs. those that are described within an non-compiled
+specification. Secondly, we will study how hints can guide the initial data
+placement as data is handed off from application to storage ( during a
+write) and from storage to application (during the read). In both cases we
+will study what minimal set of annotations and hints can allow the storage
+system to minimize data movement and optimize the resources consumed by I/O.
+
+\paragraph{Data Utility:} We will formulate the concept of utility (from an
+application perspective) that is assigned data objects to quantify the
+benefits of placement and/or movement actions to application. These
+utilities may be based on the importance of the data objects, the frequency
+with which they are accessed, etc., and are used during runtime data
+management. For example, data objects will higher utility value may be
+placed closer (e.g., in faster memory) to the accessing application.
+Similarly, when it is time to evict data objects from a higher level storage
+layer, the decision about which object to evict may take the utility into
+account.
+
+\paragraph{Autonomic in-transit data management:} 
+Efficiently placing data objects
+vertically across multiple memory hierarchy layers and storage levels as
+well as horizontally across different storage nodes requires balancing
+competing objectives. In this project we will explore autonomic data
+management strategies that can evaluate utility/cost tradeoffs and
+appropriately place/move data objects at runtime. The autonomic approach
+will leverage both user hints and information gathered at runtime (e.g.,
+runtime data access history, network topology, etc.) to place data object.
+
+\paragraph{ Runtime tracking and estimation:}
+Autonomic data placement and/or
+movement will leverage the information gathered at runtime together with
+predictive analysis requirements. This information may include runtime
+spatial-temporal data access patterns, physical network topology
+information, and etc., which will be tracked and gathered at runtime, and
+then be used to estimate and predict the behavior of coupled applications in
+the rest of simulation-time. A key research aspect here will be to
+anticipate accesses, for example using learning techniques or using
+application knowledge, and use this information to prefetch data objects.
+For example, we will identify and catalogue key phases and access patterns
+in the target workflows and then use it to characterize patterns at runtime
+and trigger appropriate place/movement actions.
 
@@
 resources.
-
-\paragraph{Proposed research approach:} In this project we will utilize
-deep memory hierarchies and multilevel storage structures to support
-staging-based in-transit data management and processing for data-intensive
-scientific workflows at extreme scale. Specifically, we will address the
-coordination and data access/sharing requirements of application workflows,
-as well as the online data management requirements such as for example,
-in-transit online data analysis, check-pointing, data reduction. We will
-also deal with the complexities associated with data placement of data and
-movement in the multiple layers and components of the storage hierarchy. The
-key components of our research include:
-
-{\em Application hints:} Application hints can provide additional
-information to the runtime based on the users' knowledge of the application
-workflow as well as on past experiences, for example, possible data access
-patterns, nature of regions of interest, and domain specific information.
-This application level information will be added to the specification of the
-workflow and can provide advisory inputs to efficiently help the runtime
-make smarter decisions about the placement and/or movement of data objects
-and the scheduling and execution of tasks, as well as to effectively manage
-trade-offs between for example, power consumption and time-to-solution.
-
-{\em Data utility:} We will formulate the concept of utility (from an
-application perspective) that is assigned data objects to quantify the
-benefits of placement and/or movement actions to application. These
-utilities may be based on the importance of the data objects, the frequency
-with which they are accessed, etc., and are used during runtime data
-management. For example, data objects will higher utility value may be
-placed closer (e.g., in faster memory) to the accessing application.
-Similarly, when it is time to evict data objects from a higher level storage
-layer, the decision about which object to evict may take the utility into
-account.
-
-{\em Autonomic in-transit data management:} Efficiently placing data objects
-vertically across multiple memory hierarchy layers and storage levels as
-well as horizontally across different storage nodes requires balancing
-competing objectives. In this project we will explore autonomic data
-management strategies that can evaluate utility/cost tradeoffs and
-appropriately place/move data objects at runtime. The autonomic approach
-will leverage both user hints and information gathered at runtime (e.g.,
-runtime data access history, network topology, etc.) to place data object.
-
-{\em Runtime tracking and estimation:} Autonomic data placement and/or
-movement will leverage the information gathered at runtime together with
-predictive analysis requirements. This information may include runtime
-spatial-temporal data access patterns, physical network topology
-information, and etc., which will be tracked and gathered at runtime, and
-then be used to estimate and predict the behavior of coupled applications in
-the rest of simulation-time. A key research aspect here will be to
-anticipate accesses, for example using learning techniques or using
-application knowledge, and use this information to prefetch data objects.
-For example, we will identify and catalogue key phases and access patterns
-in the target workflows and then use it to characterize patterns at runtime
-and trigger appropriate place/movement actions.
 

### 2015-07-04 narrative/placement.tex
some more reorganization of the text in placement

@@
 
-To take advantage of these architectural trends, we propose a triplet of
-data placement strategies. First, we will utilize the faster layer of the
+We proposed to advance the state of the art in the three major parts of the
+data lifecycle. First, we will explore the optimizations and techniques to
+scalably manage data when it is generated by the application. Second, we
+will study the management of resources and data at rest. And finally, we will
+investiage how data is consumed, either by being read and processed (for
+example within analytics) or by being purged from the system.
+
+Towards this end, we will utilize the faster layer of the
 storage hierarhcies, such as DRAM, NVRAM and remote NVRAM, as the target for
@@
 application and use case. The rest of the data will be stored on the
-parallel file system, as is done today. Second, we will utilize the
-migration of data supported by the existing implementation of \Sir to move the
-data to an appropriate tier. We will explore how migration can be guided by
-a combination of the utility function defined by the application and the
+parallel file system, as is done today. We will leverage the supported data
+mirgration techniques in \Sir to determine the most appropriate location to
+move the data once its in stable storage. This migration will be guided by a
+combination of the utility function defined by the application and the
 resources avialable to the storage system. Knowing the specific structural
@@
 optimize access to data for reader. 
-Finally, we will explore policies that combine the knowledge of data, its
+We will also explore policies that combine the knowledge of data, its
 utility to users, the cost of maintaining data in long term storage and
@@
 system.
-
-Thus we propose to advance the state of the art in three connected areas:
-(1) Data placement,
-(2) Migration,
-and (3) Purging.
 

### 2015-07-04 narrative/qos.tex
restructure qos and resource management.

--- 2015-07-02 narrative/qos-storage.tex
@@
+\subsection{Resource Management and QoS}
+
+\subsubsection{Resource Management}
+Sirocco currently offers automatic resource management capabilities. However,
+these features are strictly driven on resource capacity and data resilience
+requirements. For this project, we seek to expand the resource management
+capabilities to support additional metrics to guide resource management
+decisions. Our basic goal is to offer data-centric metrics that help guide the
+forced resource management decisions.
+
+As stated in the introduction, we seek to make a storage stack that works with
+the user. Towards that end, we must incorporate data attributes into this
+process. For example, particularly for a capability run, the storage system
+should give priority to data subset tagged with high importance to stay in fast
+storage tiers. Enabling this priority will be driven by other job
+characteristics such as job size or a manually set priority inserted by system
+administrators. It should also store the highest quality data possible given
+the performance requirement. Ideally, the storage system will store the raw
+data at full fidelity. When this tier approaches capacity, data should migrate
+not just based on age, but also based on this priority. This will help ensure
+the shortest time to insight by making sure the most important machine runs
+have priority to use system resources.
+
+Existing efforts on scheduing resources~\cite{io-cop,dorier} attempt to offer
+admission control to maximize storage performance. Unfortunately, these efforts
+are limited to participating applications and only applications from a single
+platform. For this approach to be effective, the storage system must offer an
+approach that handles clients from all connected platforms and does not
+require modification to use new storage access APIs. By using a priority
+tagging system, we will offer defaults suitable for all applications that can
+be informed by simple extensions to the job scheduler or more advanced
+management through new APIs.
+
+What else do we need to talk about?
+
 \subsubsection{QoS}
@@
 competiion for limited high performance resources. Unfettered, this competition
-will yield lower performance for all users than if applications attempt to reserve performance, thereby enabling the storage system to reason over its performance in terms of reservations and actual performance, accepting reservations that are within resource availability, rejecting those that would lead to overload.
+will yield lower performance for all users than if applications
+attempt to reserve performance, thereby enabling the storage system
+to reason over its performance in terms of reservations and actual
+performance, accepting reservations that are within resource
+availability, rejecting those that would lead to overload.
 
-It is important to note that we use performance reservations to reduce interference and to avoid overload conditions.
-If the system uses reservations intelligently, its performance is predictable and efficient because resources are not wasted on interference.
-However, in cases where an appication arrives that needs to run more urgently than any of the other already running applications, administrators can suspend less urgent applications which will save their state and relinquish their performance reservations.
-Upon restart, these applications make new performance reservations.
+It is important to note that we use performance reservations to
+reduce interference and to avoid overload conditions. If the system
+uses reservations intelligently, its performance is predictable and
+efficient because resources are not wasted on interference. However,
+in cases where an appication arrives that needs to run more urgently
+than any of the other already running applications, administrators
+can suspend less urgent applications which will save their state
+and relinquish their performance reservations. Upon restart, these
+applications make new performance reservations.
 
-There are two fundamental causes that can impede the I/O performance in an extreme-scale SSIO system.
-One is indirection, where layers of storage tiers are employed either for performance and/or for scalability reasons.
-The direct consequence is that there could be multiple traversing paths from application end to the rest place of data, and more often than not, the I/O paths are not under control under any single authority.
-The other cause is the shared use of resources.
-The best effort I/O request/response nature and lack of QoS mechanisms imply that there is little guarantee in terms of expected performance.
-Both indirection and shared use of resource contribute to a high probability of imbalanced use of resources thus the occurrence of congestion and degraded performance.
-Traditionally, there is a disconnection between storage infrastructure and rest of system software and applications: the infrastructure details are not exposed anywhere at all.
-One argument favoring this is to ensure a platform agnostic design.
+There are two fundamental causes that can impede the I/O performance
+in an extreme-scale SSIO system. One is indirection, where layers
+of storage tiers are employed either for performance and/or for
+scalability reasons.  The direct consequence is that there could
+be multiple traversing paths from application end to the rest place
+of data, and more often than not, the I/O paths are not under control
+under any single authority. The other cause is the shared use of
+resources. The best effort I/O request/response nature and lack of
+QoS mechanisms imply that there is little guarantee in terms of
+expected performance.  Both indirection and shared use of resource
+contribute to a high probability of imbalanced use of resources
+thus the occurrence of congestion and degraded performance.
+Traditionally, there is a disconnection between storage infrastructure
+and rest of system software and applications: the infrastructure
+details are not exposed anywhere at all. One argument favoring this
+is to ensure a platform agnostic design.
 
-As an example to demonstrate that this disconnection will hurt system performance: We launch 4096 processes with each process doing a single file I/O operation against half of the Spider II file system.
-The traces of those files are analyzed to examine the utilization distribution of different components.
-Figure (a), (b) and (c) shows the resource usage distribution for OSTs, OSSes, and LNETs, respectively.
-We observe that there exists a significant variation in usage across components of any given type (e.g., OST, OSS or LNET).
-For example, some OSTs are used more than 10 times while some others are never used (corresponding to zero frequency count).
-Similarly, OSSes and LNETs show significant imbalance in usage under the default placement strategy.
-Consequently, imbalanced resource utilization increases the contention at certain components more than others.
+As an example to demonstrate that this disconnection will hurt
+system performance: We launch 4096 processes with each process doing
+a single file I/O operation against half of the Spider II file
+system. The traces of those files are analyzed to examine the
+utilization distribution of different components. Figure (a), (b)
+and (c) shows the resource usage distribution for OSTs, OSSes, and
+LNETs, respectively. We observe that there exists a significant
+variation in usage across components of any given type (e.g., OST,
+OSS or LNET). For example, some OSTs are used more than 10 times
+while some others are never used (corresponding to zero frequency
+count). Similarly, OSSes and LNETs show significant imbalance in
+usage under the default placement strategy. Consequently, imbalanced
+resource utilization increases the contention at certain components
+more than others.
 
-Given these insights, we advocate the idea that the infrastructure knowledge should find a way to relay to the upper layer for better and more effective use of resources.
-We think this is more pertinent and critical given the recent development of multi-tier storage and storage component heterogeneity.
-There needs to be a way for application/middleware layer to gain more exposure of storage system for more intelligent processing logic.
-One prime example of such exposed knowledge can be request/response time.
-To most applications, this is a black box.
-Profiling it at the upper layer is neither efficient nor effective, as it doesn't reflect cross-layer characteristics.
-However, Most of storage layer does keep a detailed profiling of such information.
-We therefore envision and propose a histogram-based request/response profiling API that application and middleware layer can leverage and make more informative decisions.
+Given these insights, we advocate the idea that the infrastructure
+knowledge should find a way to relay to the upper layer for better
+and more effective use of resources.   We think this is more pertinent
+and critical given the recent development of multi-tier storage and
+storage component heterogeneity.  There needs to be a way for
+application/middleware layer to gain more exposure of storage system
+for more intelligent processing logic.  One prime example of such
+exposed knowledge can be request/response time. To most applications,
+this is a black box. Profiling it at the upper layer is neither
+efficient nor effective, as it doesn't reflect cross-layer
+characteristics. However, Most of storage layer does keep a detailed
+profiling of such information. We therefore envision and propose a
+histogram-based request/response profiling API that application and
+middleware layer can leverage and make more informative decisions.
 
@@
 ``Quality of Service'' (QoS) refers to the
-properties of the performance qualtiy of a particular service, in this case the storage system.
+properties of the performance qualtiy of a particular service, in
+this case the storage system.
 Performance quality can be expressed in either relative or absolute
@@
 terms are that they enable the implementation of strong
-guarantees: absolute terms enable agreements between application tasks and storage that we call \emph{reservations} and that do not change meaning depending
+guarantees: absolute terms enable agreements between application
+tasks and storage that we call \emph{reservations} and that do not
+change meaning depending
 on the storage system's workload. The key challenge of absolute
@@
 
-\subsubsection{QoS}
-
 \paragraph{Related Work:}

### 2015-07-05 narrative/qos.tex
get my own bib in.
started to work on qos.

@@
 
-What else do we need to talk about?
-
 \subsubsection{QoS}
@@
 \paragraph{Background:} One of the key differentiating features this project offers is driving storage
-decisions through quality of service (QoS) \emph{reservations} rather than continuing the current model of allowing
-all application to compete, without restriction, for part or all of the storage
+decisions through quality of service (QoS) \emph{reservations} rather than continuing the current model
+of allowing applications to compete, without restriction, for part or all of the storage
 resources. This traditional approach leads to intereference
-effects~\cite{lofstead:2010:adaptive} that can greatly impact IO performance
-predictability. Further, with the introduction of additional tiers in the
-storage stack, the performance variability will increase with greater
-competiion for limited high performance resources. Unfettered, this competition
-will yield lower performance for all users than if applications
+effects~\cite{lofstead:2010:adaptive,liu_hotstorage} that can greatly impact IO performance
+predictability. 
+%Further, with the introduction of additional tiers in the
+%storage stack, the performance variability will increase with greater
+%competiion for limited high performance resources. 
+%% Temporaily commented out the above sentence, because it may not be true for the next DOE systems. 
+Unfettered, this competition will yield lower performance for all users than if applications
 attempt to reserve performance, thereby enabling the storage system
@@
 availability, rejecting those that would lead to overload.
+
+{\bf {\color{red}Will connect to text on policies. Gary}}
 
@@
 than any of the other already running applications, administrators
-can suspend less urgent applications which will save their state
-and relinquish their performance reservations. Upon restart, these
-applications make new performance reservations.
+can rapidly respond to it by reducing the amount of resources taken 
+by less urgent applications, allowing the urgent application to finish,
+and reverting back to the previous committed resource level.
 

### 2015-07-05 narrative/qos.tex
more edits on qos.

@@
 
-\subsubsection{Resource Management}
-Sirocco currently offers automatic resource management capabilities. However,
-these features are strictly driven on resource capacity and data resilience
-requirements. For this project, we seek to expand the resource management
-capabilities to support additional metrics to guide resource management
-decisions. Our basic goal is to offer data-centric metrics that help guide the
-forced resource management decisions.
-
-As stated in the introduction, we seek to make a storage stack that works with
-the user. Towards that end, we must incorporate data attributes into this
-process. For example, particularly for a capability run, the storage system
-should give priority to data subset tagged with high importance to stay in fast
-storage tiers. Enabling this priority will be driven by other job
-characteristics such as job size or a manually set priority inserted by system
-administrators. It should also store the highest quality data possible given
-the performance requirement. Ideally, the storage system will store the raw
-data at full fidelity. When this tier approaches capacity, data should migrate
-not just based on age, but also based on this priority. This will help ensure
-the shortest time to insight by making sure the most important machine runs
-have priority to use system resources.
-
-Existing efforts on scheduing resources~\cite{io-cop,dorier} attempt to offer
-admission control to maximize storage performance. Unfortunately, these efforts
-are limited to participating applications and only applications from a single
-platform. For this approach to be effective, the storage system must offer an
-approach that handles clients from all connected platforms and does not
-require modification to use new storage access APIs. By using a priority
-tagging system, we will offer defaults suitable for all applications that can
-be informed by simple extensions to the job scheduler or more advanced
-management through new APIs.
-
-\subsubsection{QoS}
-
 \paragraph{Background:} One of the key differentiating features this project offers is driving storage
-decisions through quality of service (QoS) \emph{reservations} rather than continuing the current model
-of allowing applications to compete, without restriction, for part or all of the storage
-resources. This traditional approach leads to intereference
-effects~\cite{lofstead:2010:adaptive,liu_hotstorage} that can greatly impact IO performance
-predictability. 
+decisions through quality of service (QoS) \emph{reservations}, rather than continuing the current model
+of allowing applications to compete freely without restriction, for part or all of the storage
+resources. This traditional approach leads to intereference effects
+\cite{lofstead:2010:adaptive,liu_hotstorage} that can greatly impact IO performance predictability. 
 %Further, with the introduction of additional tiers in the
@@
 %% Temporaily commented out the above sentence, because it may not be true for the next DOE systems. 
-Unfettered, this competition will yield lower performance for all users than if applications
-attempt to reserve performance, thereby enabling the storage system
+Unfettered, this competition can yield a lower system and application performance 
+than the case where applications attempt to reserve storage resources, thereby enabling the storage system
 to reason over its performance in terms of reservations and actual
 performance, accepting reservations that are within resource
-availability, rejecting those that would lead to overload.
+availability and don't violate user or system policies.
 
-{\bf {\color{red}Will connect to text on policies. Gary}}
+{\bf {\color{red}Will connect to text on policies- Gary}}
 
@@
 in cases where an appication arrives that needs to run more urgently
-than any of the other already running applications, administrators
+than any of the other already running applications, the system
 can rapidly respond to it by reducing the amount of resources taken 
 by less urgent applications, allowing the urgent application to finish,
-and reverting back to the previous committed resource level.
+and reverting back to the previous committed resource level. This can
+be automatically enforced via user- (or adiminstrator-) specified \textit{policies}
+(section \ref{}).
 
@@
 
+\subsubsection{Resource Management}
+{\bf {\color{red} This section will be merged into the QoS - Gary}}
+Sirocco currently offers automatic resource management capabilities. However,
+these features are strictly driven on resource capacity and data resilience
+requirements. For this project, we seek to expand the resource management
+capabilities to support additional metrics to guide resource management
+decisions. Our basic goal is to offer data-centric metrics that help guide the
+forced resource management decisions.
+
+As stated in the introduction, we seek to make a storage stack that works with
+the user. Towards that end, we must incorporate data attributes into this
+process. For example, particularly for a capability run, the storage system
+should give priority to data subset tagged with high importance to stay in fast
+storage tiers. Enabling this priority will be driven by other job
+characteristics such as job size or a manually set priority inserted by system
+administrators. It should also store the highest quality data possible given
+the performance requirement. Ideally, the storage system will store the raw
+data at full fidelity. When this tier approaches capacity, data should migrate
+not just based on age, but also based on this priority. This will help ensure
+the shortest time to insight by making sure the most important machine runs
+have priority to use system resources.
+
+Existing efforts on scheduing resources~\cite{io-cop,dorier} attempt to offer
+admission control to maximize storage performance. Unfortunately, these efforts
+are limited to participating applications and only applications from a single
+platform. For this approach to be effective, the storage system must offer an
+approach that handles clients from all connected platforms and does not
+require modification to use new storage access APIs. By using a priority
+tagging system, we will offer defaults suitable for all applications that can
+be informed by simple extensions to the job scheduler or more advanced
+management through new APIs.
+
 \paragraph{Related Work:}

### 2015-07-05 narrative/qos.tex
more edits on QoS.

@@
 
-\paragraph{Background:} One of the key differentiating features this project offers is driving storage
-decisions through quality of service (QoS) \emph{reservations}, rather than continuing the current model
+\paragraph{Background:} 
+Quality of Service (QoS) refers to the
+properties of the performance quality of a particular service, in
+this case the storage system.
+Performance quality can be expressed in either relative or absolute
+terms. Examples of relative performance terms are ``fair'',
+``proportional'', or ``priority/class-based'' (``the higher the
+priority the better the service'' or ``1st class is better than
+2nd''). The key advantage of relative performance terms is that
+they are easy to implement. The key disadvantage is the difficulty
+of making any guarantees based on those implementations --
+even relative guarantees are mired by the well-known effect of
+priority inversion~\cite{lampson:cacm80}. Examples of absolute
+performance terms are ``rate-based'', ``soft real-time'', and
+``hard real-time''. The key advantage of absolute performance
+terms are that they enable the implementation of strong
+guarantees: absolute terms enable agreements between application
+tasks and storage that we call \emph{reservations} and that do not
+change meaning depending
+on the storage system's workload. The key challenge of absolute
+performance terms is that their implementation is non-trivial,
+especially in large-scale systems.
+
+One of the key differentiating features this project offers is driving storage
+decisions through QoS \emph{reservations}, rather than continuing the current model
 of allowing applications to compete freely without restriction, for part or all of the storage
@@
 availability and don't violate user or system policies.
-
-{\bf {\color{red}Will connect to text on policies- Gary}}
 
@@
 
-Given these insights, we advocate the idea that the infrastructure
-knowledge should find a way to relay to the upper layer for better
-and more effective use of resources. We think this is more pertinent
-and critical given the recent development of multi-tier storage and
-storage component heterogeneity.  There needs to be a way for
-application/middleware layer to gain more exposure of storage system
-for more intelligent processing logic.  One prime example of such
-exposed knowledge can be request/response time. To most applications,
-this is a black box. Profiling it at the upper layer is neither
-efficient nor effective, as it doesn't reflect cross-layer
-characteristics. However, Most of storage layer does keep a detailed
-profiling of such information. We therefore envision and propose a
-histogram-based request/response profiling API that application and
-middleware layer can leverage and make more informative decisions.
+% Don't know how to integrate this. Gary
+%Given these insights, we advocate the idea that the infrastructure
+%knowledge should find a way to relay to the upper layer for better
+%and more effective use of resources. We think this is more pertinent
+%and critical given the recent development of multi-tier storage and
+%storage component heterogeneity.  There needs to be a way for
+%application/middleware layer to gain more exposure of storage system
+%for more intelligent processing logic.  One prime example of such
+%exposed knowledge can be request/response time. To most applications,
+%this is a black box. Profiling it at the upper layer is neither
+%efficient nor effective, as it doesn't reflect cross-layer
+%characteristics. However, Most of storage layer does keep a detailed
+%profiling of such information. We therefore envision and propose a
+%histogram-based request/response profiling API that application and
+%middleware layer can leverage and make more informative decisions.
 
@@
 
-``Quality of Service'' (QoS) refers to the
-properties of the performance qualtiy of a particular service, in
-this case the storage system.
-Performance quality can be expressed in either relative or absolute
-terms. Examples of relative performance terms are ``fair'',
-``proportional'', or ``priority/class-based'' (``the higher the
-priority the better the service'' or ``1st class is better than
-2nd''). The key advantage of relative performance terms is that
-they are easy to implement. The key disadvantage is the difficulty
-of making any guarantees based on those implementations --
-even relative guarantees are mired by the well-known effect of
-priority inversion~\cite{lampson:cacm80}. Examples of absolute
-performance terms are ``rate-based'', ``soft real-time'', and
-``hard real-time''. The key advantage of absolute performance
-terms are that they enable the implementation of strong
-guarantees: absolute terms enable agreements between application
-tasks and storage that we call \emph{reservations} and that do not
-change meaning depending
-on the storage system's workload. The key challenge of absolute
-performance terms is that their implementation is non-trivial,
-especially in large-scale systems.
+% Lifted the following paragraph to the beginning. Gary
+%``Quality of Service'' (QoS) refers to the
+%properties of the performance qualtiy of a particular service, in
+%this case the storage system.
+%Performance quality can be expressed in either relative or absolute
+%terms. Examples of relative performance terms are ``fair'',
+%``proportional'', or ``priority/class-based'' (``the higher the
+%priority the better the service'' or ``1st class is better than
+%2nd''). The key advantage of relative performance terms is that
+%they are easy to implement. The key disadvantage is the difficulty
+%of making any guarantees based on those implementations --
+%even relative guarantees are mired by the well-known effect of
+%priority inversion~\cite{lampson:cacm80}. Examples of absolute
+%performance terms are ``rate-based'', ``soft real-time'', and
+%``hard real-time''. The key advantage of absolute performance
+%terms are that they enable the implementation of strong
+%guarantees: absolute terms enable agreements between application
+%tasks and storage that we call \emph{reservations} and that do not
+%change meaning depending
+%on the storage system's workload. The key challenge of absolute
+%performance terms is that their implementation is non-trivial,
+%especially in large-scale systems.
 

### 2015-07-05 narrative/qos.tex
more edits.

@@
 
-\subsubsection{Resource Management}
-{\bf {\color{red} This section will be merged into the QoS - Gary}}
-Sirocco currently offers automatic resource management capabilities. However,
-these features are strictly driven on resource capacity and data resilience
-requirements. For this project, we seek to expand the resource management
-capabilities to support additional metrics to guide resource management
-decisions. Our basic goal is to offer data-centric metrics that help guide the
+To achieve effective performance isolation, we seek to expand the resource management
+capabilities in Sirocco to support additional metrics to guide resource management
+decisions. Currently Sirocco offers automatic resource management capabilities,
+based on resource capacity and data resilience requirements. 
+ Our basic goal is to offer data-centric metrics that help guide the
 forced resource management decisions.
@@
 
-Existing efforts on scheduing resources~\cite{io-cop,dorier} attempt to offer
+\paragraph{Related Work:} Existing efforts on scheduing resources~\cite{io-cop,dorier} attempt to offer
 admission control to maximize storage performance. Unfortunately, these efforts
@@
 management through new APIs.
-
-\paragraph{Related Work:}
 

### 2015-07-05 narrative/qos.tex
modified:   qos.tex

@@
 
-\paragraph{Background:} 
+\paragraph{Resource Management Background:} There is an important connection between finding the resources that can help fulfill a particular request and the predictability of the performance of that request, both in terms of latency and resolution.
+This connection is particular important for reads: in very large scale systems, finding a resource to read can be a significant, if not dominant, component of overall read latency.
+If the search for resources is unbounded, it will be difficult to make any performance predictions.
+
+Locating resources is a typical distributed systems problem.
+Central solutions are limited in their scalability, especially if they have to track every update.
+Traditional file systems are good examples for this: most updates to files require updates to their centrally maintained block allocation tables.
+The opposite of file systems are ``pile'' systems which instead of keeping track of updates search for resources at request time.
+Pile systems allow very efficient writes because clients can write whereever there is free space and there is very bookkeeping overhead.
+Furthermore, pile systems have great flexibility for moving resources around, for example to ensure durability.
+However, reads require a potentially unbounded search, especially if it is not clear whether the resource actually exists.
+Methods to make searches more efficient without reverting to centralized bookkeeping usually involves some kind of partitioning of the search space and a method to limit the search to a small number of partitions.
+
+Frequently this involves a hash function that is relatively stable against changes to its range, such as failures or updates to the hardware of the system (see for example consistent hashing~\cite{karger:stoc97}).
+Consistent hashing is scalable since it can be well-known throughout a system of any size and only requires minimal movement of data in the presence of failures or system changes.
+Consistent hash functions can also have sets of ordered but infinite lists as their range (which works well with primary replication)~\cite{honicky:ipdps04}, implement declustering (limiting overlap between partitions), and support placement policy rules~\cite{weil:sc06}.
+
+In this proposal we will investigate the trade-off in terms of predictable peformance between ``filing'' data, making an upfront investment to ensure finding data quickly, and ``piling'' data, pushing the cost of finding to the moment when the data is actually needed.
+This trade-off has a parallel in data management where there are two alternative paths to achieve knowledge: the analogue to filing is ``extract, transform, and load'' (ETL) data into a warehouse, paying an upfront ingest cost to achieve fast querying; the analogue to piling is to not pay any upfront cost and query the raw data directly by using data processing platforms such as Hadoop.
+Recent work has found that the combination of these two techniques can be optimized to perform better than any one technique alone~\cite{lefevre:sigmod14a}.
+We believe that a similar opportunity exists for combining filing and piling approaches to achieve performance predictability without sacrificing too much flexibility.
+
+More specifically, the coarseness of a search space partitioning limits the search time within a partition, i.e.
+finely partitioned search spaces require little or no search within a partition but offer little placement flexibility.
+Partitioning can also be fine in one dimension and coarse in another to allow flexibility where it is most needed while still ensuring fast finding.
+For example, a hash function can finely partition a low storage tier but each partition can extend across multiple tiers so that a search still has to occur in order to find a resource on a particular tier.
+
+\paragraph{QoS Background:} 
 Quality of Service (QoS) refers to the

### 2015-07-05 narrative/data-description.tex
Change the data description to talk more about the choices that need to be made.

@@
 In this project, a key metadata attribute
-that enables QoS scheduling, data placement, and ???, is \textit{data utility value}, which
+that enables QoS scheduling, data placement, and lifetime, is \textit{data utility}, which
 captures the priority of objects - it is to our belief that
 for exascale science, users must prioritize a small set of objects to be saved on 
-higher-level capacity-limited storage layer
-to avoid the slow access to large-capacity lower-level storage, such as tapes.
+higher-level capacity-limited storage 
+to avoid the slow access to large-capacity lower-level storage, such as tape.
 We propose a new technique, described by a utility function plugged in by a user, 
@@
 In the meanwhile, the user would like the ability to get a certain amount of Quality of Service 
-such that they can then make decisions when the expected bandwidth, 
-for example is less than what they desire and will then place in a certain set 
-of rules which the system can make autonomic decisions to help decide what should be done.
+such that they can then make decisions when the expected bandwidth.
+%for example is less than what they desire and will then place in a certain set 
+%of rules which the system can make autonomic decisions to help decide what should be done.
 A key insight in this proposed project has been the increased interaction of
@@
 mechanisms for this purpose.
-
-Firstly, we will explore the augmentation of
-I/O application programing interfaces (I/O APIs) to allow applications to
-both specify timing and quality information and also query the storage
-system for timing estimates. If user decide to write or read data to or from hierarchical storage systems, 
-they will rely on the API to send their intentions for inquiry and examine the system status, 
-including how much data they would like to write/read, desired bandwidth, data compression method, etc, 
-or the user can express their intension of writing data right now no matter what the traffic is now.
-Through this interface we expect the
-application and user to gain insights into how long a single output or input
-call will take given the required quality information, and then react to
-these estimates by adjusting the quality or restricting the scope of the
-data required. Likewise we will explore how an application can provide
-timing information to the storage system to allow the storage system to make
-optimization decisions to best meet the requirements from the application.
-
-Secondly, we will also explore external data annotations, such as those
-provided by the configuration file in ADIOS. Through the use of these
-external augmentation the user can provide insights to the storage system on
-the relative value of the data, expected life time and performance
-characteristics, as well as relationships between different data sets. With
-this information the storage system can make optimizations specific to a use
-case. We expect these augmentations to be particularly important for
-eviction of data from a storage layer, and migration of data sets to a
-different storage layer.
-
-The capability of APIs provide dynamic runtime information to user and make the system 
-more transparent to them rather than a black box. The end user could analyze or visualize data 
-in a more comfortable position, where they are able to envision 
-how much longer they are going to wait for data.
 
@@
 data importance hasn't been explored, which we believe will be important for exascale storage solutions. 
+{\color{red}\bf there must be other work in this area, and we need to get more references}
 
@@
 by the middleware and utilized by the underlying storage system to management data placement.
+%
+%We will design an updated I/O API that will expand on the current POSIX I/O
+%semantics by adding new information to the I/O request.
 
-We will design an updated I/O API that will expand on the current POSIX I/O
-semantics by adding new information to the I/O request. We will explore the
+We will explore the
 set of application level hints that can be easily provided to the storage
@@
 
+\begin{wrapfigure}{r}{0.7\textwidth}
+%\vspace{0ex}
+        \begin{centering} 
+	\includegraphics[scale=0.77]{graphics/SSIO-bucket.pdf}
+        %\vspace{1ex}
+        \caption{Data is given as a set of objects which includes utility which contains the lifetime in a storage layer, and the priority of the layer.}
+        \label{fig:ssio-bucket}
+        \end{centering}
+%        \vspace{1ex}
+\end{wrapfigure}
+
+%
+We will explore these APIs through our ADIOS framework and work with many of our applications, 
+including XGC1, GTC, and SPECFM3D, and understand: 1) what hints the users can give us, 2) how
+the users can take away much of their application specific code to ``system level code'' through the use of these hints.
+Currently ADIOS has the ability to only perform write or read when an excute command is given; i.e. when users 
+ask to write and read, they are only done in  bulk, so data can be aggregated together for writes and reads. We will
+design an API that will integrate through the SSIO layer to give time estimations so that users can decide 
+which priorities of data will be written and read. The utility is given as a key,value pair where for each object: e.g. temperature
+we will have a user function executed that will bucket this into ``n''' bins, where the first bin is the most important data, and the
+last bin is the least important data. The key value pair will then be associated with each bin, which will give it a lifetime in the highest level
+of storage that the user thinks will be needed. For example: in Figure~\ref{fig:ssio-bucket} we have data that comes from the XGC1 code.
+The user can pick a data-refactoring scheme (see Sec~\ref{sec:data-refactor}) which will bucket the particle data into 4 levels of prioritization.
+The user can then pick a data utility which is 3 months for the highest priority items, and 12 months for the lowest priority data. 
+The user will then call an API when they specify all of the objects they want written (particles, and two fields and a mesh) in this example, and
+the system will return an estimate of the time it will take to write. The user can then use another API which will determine the lowest priority
+element (in this case the user will write all of the data (P1-P4). The data will then be stored on all of the layers of the file system, where the
+two highest priority buckets of the particle data will be stored in the Parallel File System, but the next  highest priority items will be stored
+lower in the storage stack. As the system needs to purge the data from the Parallel File System, it can understand that the P2 layer of
+data can be removed. 
+%
+
+We will design new APIs that the users can use as part of a data mover, or the system can automatically use, to raise or lower these sub-objects
+up and down the storage stack. The SSIO layer will then manage the system metadata so that the user can transparently access any of these
+sub-objects without knowledge of the layer. This requires that when reading the data, the APIs will then get a time estimate of reading, and then
+allow the users to determine if they will read less (only the highest priority sub-objects) if the time estimate for reading exceeds their expectations.
+%
+
+Users will also need to describe information about the data in order to choose the ``most optimal'' data-refactoring routine. Currently, we understand
+that users can tell us that the data represents: 1) phase-space (e.g. the particles in our example), 2) spatial-changes (example the mesh), 3) space-time
+(example the fields on the mesh). We need this information to get the best possible data refactoring methods. Although users can request 
+a user-defined data-refactoring method, system-level methods can use this information to make better decisions. We will investigate new APIS 
+and with new semantics to take this information and use this to best prioritize and re-organize information.
+
+Finally we will investigate additional information on top of the data model. This information will be used to talk about data associations which go
+beyond traditional data models. For example, there is a inherent relationship between the fields in our example and the mesh. A common data model
+will describe all of the variables in the mesh (the coordinates, the units, the types of elements if this were a finite-element mesh) . We will explore
+new mechanisms which will group data elements together so that we can understand that (M1, F1, and G1) need to be organized together. 
+In otherwords, in will not be helpful if data from M1 and M2 were organized together, along with only F1. When the users wants to visualize data from 
+the field, F1, they will need only M1, and M2 would be additional information. Likewise, if the user wants to see F1 and F2, then they will need the
+mesh from M1 and M2, and this requires that the time to access all of this data is approximately the same, which requires that all of this needs
+to be organized in the same layer of the storage system.  We will investigate what choices the SSIO system must make in order to keep these
+pieces together. Do we move everything to a lower layer when all of it doesn't fit, or can we move other user-data for this data set, evict for example
+P3 in order to keep F2, G2, and M2 together? We need to investigate the potential problems when this occurs, and how the users can specify
+additional information to ensure they get the best possible data from the SSIO layer.
+
+The capability of APIs provide dynamic runtime information to user and make the system 
+more transparent to them rather than a black box. The end user could analyze or visualize data 
+in a more comfortable position, where they are able to envision 
+how much longer they are going to wait for data.
+
+%Firstly, we will explore the augmentation of
+%I/O application programing interfaces (I/O APIs) to allow applications to
+%both specify timing and quality information and also query the storage
+%system for timing estimates. If user decide to write or read data to or from hierarchical storage systems, 
+%they will rely on the API to send their intentions for inquiry and examine the system status, 
+%including how much data they would like to write/read, desired bandwidth, data compression method, etc, 
+%or the user can express their intension of writing data right now no matter what the traffic is now.
+%Through this interface we expect the
+%application and user to gain insights into how long a single output or input
+%call will take given the required quality information, and then react to
+%these estimates by adjusting the quality or restricting the scope of the
+%data required. Likewise we will explore how an application can provide
+%timing information to the storage system to allow the storage system to make
+%optimization decisions to best meet the requirements from the application. 
+%
+%Secondly, we will also explore external data annotations, such as those
+%provided by the configuration file in ADIOS. Through the use of these
+%external augmentation the user can provide insights to the storage system on
+%the relative value of the data, expected life time and performance
+%characteristics, as well as relationships between different data sets. With
+%this information the storage system can make optimizations specific to a use
+%case. We expect these augmentations to be particularly important for
+%eviction of data from a storage layer, and migration of data sets to a
+%different storage layer.
+
 \paragraph{Challenges:}
@@
 enable the application to receive enough information to make adaptation
-decisions.
+decisions.  Since we will work closely with many of the leading LCF applications we need
+to make sure that our additional APIs and semantics in the SSIO layer will be accepted by them, and then
+hopefully by the rest of the community. This requires that we carefully examine all possible design choices to ensure
+that we don't keep changing the APIs and semantics, but we understand what we decisions can be made so that
+we don't burden the users with confusing APIS and semantics which can't even be used by the system.

### 2015-07-05 narrative/data-reduction.tex
figure
Change the data description to talk more about the choices that need to be made.

@@
 \subsection{Data Refactoring}
-
+\label{sec:data-refactor}
 It is critical that users give us more information about writing and reading from the SSIO layer
@@
 the storage hierarchy.  We will also investigate technique which can use ``plug-ins'' for detection of features and store higher resolution data
-around the feature and lower resolution values which are further away from the features. This re-organization must occur without increasinig the
+around the feature and lower resolution values which are further away from the features. This re-organization must occur without increasing the
 total amount of storage than what the users intended. We will also investigate the time it takes to re-organize the data and use the multi-layer

### 2015-07-05 abstract.tex
changes to the abstract

@@
 \enlargethispage{2\baselineskip}
-Exascale scientific discovery will be
-severely bottlenecked without sufficient new research into managing and
-storing the large amounts of data that will be produced during the
-simulation, and analyzed for months afterwards. In this project we will
+Scientific discovery at the exascale will not be possible without significant
+new research in the management, storage and retrieval over the long lifespan
+of the large amounts of data that will be produced.
+In this project we will
 demonstrate novel techniques to  facilitate efficient mapping of data
@@
 the user space onto multiple storage tiers, and enable application-guided
-data reductions/transformations to address capacity and bandwidth
+data reductions and transformations to address capacity and bandwidth
 bottlenecks. 
@@
 We will test our prototypes  on current and future DOE system with many of today's applications, including the
-s XGC1, GTC, QMCPack, and SpecFM3D simulations. 
+XGC1, GTC, QMCPack, and SpecFM3D simulations. 
 Our solutions will provide  new functionalities and APIs for

### 2015-07-05 narrative/placement.tex
Scott changes

@@
 
+%key points based on discussion with Carlos
+%1. Allocation tables and simple search won't be able to scale
+%2. Carlos suggests looking at a determinstic function to decide on
+%placement of data into the storage hierarchy. Instead of storing the
+%metadata that describes where data is stored (which server/tier) we can
+%simple associate a function for each application. 
+%3. Research into what would be an approrpriate function that allows even
+%splitting of data across many servers/tiers
+%4. Similar to our approach for defining utility over time
+%5. Migration can also use a function. We shouldn't distinguish migration
+%from data movement and placement too much
+%6. We need metrics that can decide how the data moves up and down in the
+%hierarchy (so we can enable pre-fetching)
+%7. Carlos suggested that instead of hints we call them service level
+%objectives (and that leads into quality of service)
 The application workflows targeted in this proposal generate very large
@@
 Towards this end, we will utilize the faster layer of the
-storage hierarhcies, such as DRAM, NVRAM and remote NVRAM, as the target for
+storage hierarchies, such as DRAM, NVRAM and remote NVRAM, as the target for
 the initial write from an application. To mitigate the space concerns, we
@@
 system.
+
+Our research will ultimately address the following questions:\\
+1) Can we do scalable migration without making the discovery process
+    unbounded;\\
+2) What are the parameters and input for migration;
+3)  Can we use the additional application level knowledge of data to purge
+    portions of the data from storage without making the data too much less
+    valuable?;\\
+3)  With so many objects and no centralized directory how do we know what
+    is actually in the storage to be purged?;\\
+4)  How do purge in a way that doesn't require a central authority,
+    doesn't interfere with other I/O and isn't bottlenecked by some sort of
+    global consistency?
 

### 2015-07-05 narrative/proposed-work.tex
Data refactoring...
restructure qos and resource management.
Scott changes

--- 2015-07-02 narrative/proposed-work.tex
@@
 
-\begin{enumerate}
-\item Re-factoring of the data
-  \begin{itemize}
-  \item What kind of reorganization would work when?
-  \item Would it take too much time and resources + memory and how can we
-    manage the cost of the reorganization
-  \item Will the users loose too much value from the data during refactoring
-    and how can we control/bound this loss of value
-  \end{itemize}
-\item Quality of Service
-  \begin{itemize}
-  \item Can we make any gurantees in a scalable decentralized system?
-  \item What quality of service metrics can be make gurantees on in a scalable
-    way?
-  \item What kind of error rates will we have in our estimation and what error
-    rates are acceptable?
-  \item How much will the qos scheduler, admission control, manager cost in
-    terms of resources and time
-  \item How do we make the scheduler scalable with the size of the system?
-  \end{itemize}
-\item Discovery and naming
-  \begin{itemize}
-  \item Which data block and what quality/type do we respond to a query
-    with?
-  \item How will we scale a fully distributed storage system's metadata
-    management
-  \item Can we accelerate the discovery process with the addition of an
-    application/user level metadata store that tracks where data was
-    originally written and how was it structured?
-  \item Can we provide time estimates for read/write calls in the presence of
-    an unbounded discovery system? If not how do we bound the discovery time?
-  \item Time cost of discovery vs time cost of actually reading the data
-    (latency vs bandwidth essentially)
-  \end{itemize}
-\item Migration
-  \begin{itemize}
-  \item Can we do scalable migration without making the discovery process
-    unbounded
-  \item What are the parameters and input for migration
-  \item Can we use the additional application level knowledge of data to purge
-    portions of the data from storage without making the data too much less
-    valuable?
-  \item With so many objects and no centralized directory how do we know what
-    is actually in the storage to be purged?
-  \item How do purge in a way that doesn't require a central authority,
-    doesn't interfere with other I/O and isn't bottlenecked by some sort of
-    global consistency
-  \end{itemize}
-\end{enumerate}
+%\begin{enumerate}
+%\item Re-factoring of the data
+%  \begin{itemize}
+%  \item What kind of reorganization would work when?
+%  \item Would it take too much time and resources + memory and how can we
+%    manage the cost of the reorganization
+%  \item Will the users loose too much value from the data during refactoring
+%    and how can we control/bound this loss of value
+%  \end{itemize}
+%\item Quality of Service
+%  \begin{itemize}
+%  \item Can we make any guarantees in a scalable decentralized system?
+%  \item What quality of service metrics can be make guarantees on in a scalable
+%    way?
+%  \item What kind of error rates will we have in our estimation and what error
+%    rates are acceptable?
+%  \item How much will the qos scheduler, admission control, manager cost in
+%    terms of resources and time
+%  \item How do we make the scheduler scalable with the size of the system?
+%  \end{itemize}
+%\item Discovery and naming
+%  \begin{itemize}
+%  \item Which data block and what quality/type do we respond to a query
+%    with?
+%  \item How will we scale a fully distributed storage system's metadata
+%    management
+%  \item Can we accelerate the discovery process with the addition of an
+%    application/user level metadata store that tracks where data was
+%    originally written and how was it structured?
+%  \item Can we provide time estimates for read/write calls in the presence of
+%    an unbounded discovery system? If not how do we bound the discovery time?
+%  \item Time cost of discovery vs time cost of actually reading the data
+%    (latency vs bandwidth essentially)
+%  \end{itemize}
+%\item Migration
+%  \begin{itemize}
+%  \item Can we do scalable migration without making the discovery process
+%    unbounded
+%  \item What are the parameters and input for migration
+%  \item Can we use the additional application level knowledge of data to purge
+%    portions of the data from storage without making the data too much less
+%    valuable?
+%  \item With so many objects and no centralized directory how do we know what
+%    is actually in the storage to be purged?
+%  \item How do purge in a way that doesn't require a central authority,
+%    doesn't interfere with other I/O and isn't bottlenecked by some sort of
+%    global consistency
+%  \end{itemize}
+%\end{enumerate}
 
-Our research into the storage system and middle ware layers must address
-many of these challenges including
+%Our research into the storage system and middle ware layers must address
+%many of these challenges including
 % \begin{enumerate}
@@
 
-\paragraph{Deliverables and artifacts}
-
-\item Add plug-in architecture to Sirocco to support selective data compression
-\item Add specific tier destinations command to Sirocco (assuming that the caching mechansism cannot achieve the desired effects).
-\item Develop profiling system to determine how a data set is used during preparation runs prior to a capability run to determine how to optimize data placement for capability run analytics.
-
-\end{enumerate}
-
-{\bf \color{red}we need to condense these challenges into a few basic principles - 4 at most - SAK}
+%\paragraph{Deliverables and artifacts}
+%
+%\begin{enumerate}
+%\item Add plug-in architecture to Sirocco to support selective data compression
+%\item Add specific tier destinations command to Sirocco (assuming that the caching mechansism cannot achieve the desired effects).
+%\item Develop profiling system to determine how a data set is used during preparation runs prior to a capability run to determine how to optimize data placement for capability run analytics.
+%
+%\end{enumerate}
+%
+%{\bf \color{red}we need to condense these challenges into a few basic principles - 4 at most - SAK}
 
@@
 and I/O layer which takes into account user prioritization and utility to optimize the SSIO layer. The approach is based on 
-four underlying principles:
+three underlying principles:
 
-\underline{Principle 1: A knowledge-centric system} , let user knowledge  to define data policies. talk about this pieces...
+\underline{Principle 1:} A knowledge-centric system , let user knowledge  to define data policies. Today Storage System and I/O
+layers are written in a stove-pipe fashion, and quite often don't allow optimizations to take place. We propose to re-design the
+layer in a highly integrated fashion where user will place their intentions into the system and actions will dynamically and statically
+take place to optimize for both the system and for individual requests.
 
-\underline{Principle 2: A SSIO system which integrates system knowledge together with application knowledge} to understand
-how to organize data in the deep storage hierarchy
+\underline{Principle 2}: Predictable Performance in the SSIO layer needs to be established so more science can be done on the
+exascale Systems. Without predictable performance, the I/O layer performance must be guessed by the users, which quite often
+mean that science is not performed well.
 
-\underline{Principle 3: Predictable Performance in the SSIO layer} so that intelligent choices can be made by both the user and the system
+\underline{Principle 3:} In a data-centric SSIO system, we need user goals to define the data policies. We introduce data-driven
+SSIO optimizations by enabling users to define data manipulations (re-organizations) at a high level and then we express
+policy goals by associating intentions (hints) with data;' e.g. ``this data at this level should be accessible in the next three months.''
+These high level specifications of desired results will be used in our system for data migration and eviction policies.
 
@@
 
-% \input{narrative/data-desc}
+%\input{narrative/data-desc}
+
 \input{narrative/data-reduction} % includes data-reorg and the api specs move to the above section -
-\input{narrative/data-reorg}
-% \input{narrative/qos-app}
+%\input{narrative/data-reorg}
+
+%\input{narrative/qos-app}
 \input{narrative/placement}
-% \input{narrative/policies}
-% \input{narrative/apis}
+\input{narrative/migration}
+\input{narrative/eviction-metrics}
+
+%\input{narrative/policies}
+%\input{narrative/apis}
 
@@
 
-\input{narrative/migration}
-\input{narrative/resource-mgmt}
-\input{narrative/qos-storage}
-% \input{narrative/discovery}
-\input{narrative/eviction-metrics}
+%\input{narrative/resource-mgmt}
+\input{narrative/qos}
+%\input{narrative/discovery}
+
 \input{narrative/naming-service}

### 2015-07-05 narrative/intro.tex
minor edits to the introduction

@@
 The central objective of this proposal, ``minimizing the time knowledge" from 
-scientific computing, aligns perfectly this sentiment. 
+scientific computing, aligns perfectly with this sentiment. 
 %Hamming was a pioneer in the fields of coding theory, scientific computing and 
@@
 Our approach in this proposal is to enable researchers to place more knowledge
-into the SSIO system so that data can be recast from a bunch of bytes to
+into the SSIO system so that data can be recast from a sequence of bytes to
 information, which the system can then place and optimize the time to gain
@@
 followed by a post-processing or analysis stage that might take place over
-several ensuing weeks or months. The overarching objective is to gain insight,
-or knowledge.  As already mentioned, the anticipated scale of computations
-means that this workflow is simply no longer viable.
+several ensuing weeks or months. Clearly, given the computation expected at the
+exascale, this type of workflow is simply not viable.
+
 %

### 2015-07-05 narrative/naming-service.tex

@@
 across storage layers, and facilitates data discovery.
+
+Our research in this area will address the following questions:\\
+   Which data block and what quality/type do we respond to a query
+    with?;\\
+   How will we scale a fully distributed storage system's metadata
+    management;\\
+   Can we accelerate the discovery process with the addition of an
+    application/user level metadata store that tracks where data was
+    originally written and how was it structured?;\\
+   Can we provide time estimates for read/write calls in the presence of
+    an unbounded discovery system? If not how do we bound the discovery time?
+   How do we amortize the time cost of discovery vs time cost of actually reading the data
+    (latency vs bandwidth essentially)?
 

### 2015-07-05 narrative/qos.tex

@@
 
-In this proposal we will investigate the trade-off in terms of predictable peformance between ``filing'' data, making an upfront investment to ensure finding data quickly, and ``piling'' data, pushing the cost of finding to the moment when the data is actually needed.
+In this proposal we will investigate the trade-off in terms of predictable performance between ``filing'' data, making an upfront investment to ensure finding data quickly, and ``piling'' data, pushing the cost of finding to the moment when the data is actually needed.
 This trade-off has a parallel in data management where there are two alternative paths to achieve knowledge: the analogue to filing is ``extract, transform, and load'' (ETL) data into a warehouse, paying an upfront ingest cost to achieve fast querying; the analogue to piling is to not pay any upfront cost and query the raw data directly by using data processing platforms such as Hadoop.
@@
 For example, a hash function can finely partition a low storage tier but each partition can extend across multiple tiers so that a search still has to occur in order to find a resource on a particular tier.
+
+Specifically, we will address the following research questions:\\
+  1) Can we make any guarantees in a scalable decentralized system?; \\
+  2) What quality of service metrics can be make guarantees on in a scalable
+    way?;\\
+  3) What kind of error rates will we have in our estimation and what error
+    rates are acceptable?;\\
+  4) How much will the qos scheduler, admission control, manager cost in
+    terms of resources and time?;\\
+  5) How do we make the scheduler scalable with the size of the system?
 

### 2015-07-05 narrative/intro.tex
cleanup the XGC1 use case

@@
 They launched a series of simulations, planning to produce 100~PB of data over 10 days of total runtime on the Titan system.
-An expert team was assembled from the user group, the middleware I/O team to help with the application I/O (using ADIOS), the local experts of the Lustre file-system and the tape archival system (HPSS).
-Due to physical resource limitations, the simulation was restricted to write about 10~PB initially, but then considering the time to write and read this much data on the tape archive system, and the financial costs to store it, this plan was cut again considerably.
-Note that such time and cost information was not known by the user; the HPSS expert had to inform everyone else about it.
-The user team had to make hard decisions of which data to save from the originally planned 100~PB.
-The file system experts told about the then-current file system limitations in size, performance and purge policy.
-The user then decided to write 0.5~PB per day (the individual daily simulation runs were and are going to be separated by many days or weeks), 5~PB in total, storing maximum of 2~PB of data on disks at a time and storing the rest on tape.
+An expert team was assembled from the user group, the middleware I/O team to help with the application I/O (ADIOS), the local experts of the Lustre file-system and the tape archival system (HPSS).
+Due to physical resource limitations, the simulation was restricted to write about 10~PB initially.
+However, once the time to write and read, together with the financial cost to archive this much data were more fully explored, this plan was again signficantly reduced.
+At each stage, the particular expert knowledge regarding the lustre file system and the archival tape storage were not known to the user team.
+As these particulars were realized, significant modifications in the plans were made.
+Eventually, the user was forced to make very hard decisions in reducing the desired 100~PB over a 10-day run to a mere 5~PB of data.
+A maximum of 2~PB would be on disk, and the rest would be archived.
 

### 2015-07-05 narrative/naming-service.tex
fixed the research questions

--- 2015-07-02 narrative/naming-service.tex
@@
 sufficient metadata information, track or be able to find objects as they move
-across storage layers, and facilitates data discovery.
+across storage layers, and facilitates data discovery. Since we need to also
+support access through POSIX APIs for existing application and system
+integration, we must address how to bridge the drastic differences between the
+decomposed objects of potentially varying fidelity with potentially multiple
+version with a POSIX API-compatible interface.
 
-Hierarchical Storage Management (HSM) systems~\cite{hsm} offer a strict caching approach
-for managing different storage capacities trading off performance for capacity.
-By maintaining a single namespace across all tiers, it is possible to list a
-single directory view with files stored at different tiers. While this approach
-to managing multi-tier storage works for whole files that fit on single logial
-devices, it is far from ideal for scientific simulations.
+Our research in this area will address the following questions:\\
+Since our metadata is distributed and partitioned, how do we respond to user
+  requests within a specified time bound with accuracy? How do we estimate the
+  completeness of the response?\\
+For POSIX API requests, how do we respond if only varying data fidelity levels
+  are available for the requested data object?\\
+What level of metadata should we maintain and keep in sync to enhance response
+  quality within time-bound requests?\\
+How much time should be allowed for searching for metadata given the user
+  specified limits for data retrieval since the user will still have to spend
+  the data retrieval time from different devices? How do we integrate the read
+  time overhead into the metadata requests?\\
+How do we provide appropriate security while maintaining scalability,
+  particularly with the distributed metadata stores?\\
+% Which data block and what quality/type do we respond to a query with?;\\
+% How will we scale a fully distributed storage system's metadata management;\\
+% Can we accelerate the discovery process with the addition of an
+%    application/user level metadata store that tracks where data was
+%    originally written and how was it structured?;\\
+% Can we provide time estimates for read/write calls in the presence of
+%    an unbounded discovery system? If not how do we bound the discovery time?
+% How do we amortize the time cost of discovery vs time cost of actually
+%    reading the data (latency vs bandwidth essentially)?
+
+Hierarchical Storage Management (HSM) systems~\cite{hsm} offer a strict caching
+approach for managing different storage capacities trading off performance for
+capacity.  By maintaining a single namespace across all tiers, it is possible
+to list a single directory view with files stored at different tiers. While
+this approach to managing multi-tier storage works for whole files that fit on
+single logial devices, it is far from ideal for scientific simulations.
 
@@
 between the user and the storage system extends into this area as well.
-Negotiating the data quality/retrieval time tradeoff reuqires additional data.
-Providing this functionality requries the naming service be able to both return
-a list of data versions and an indication of data quality. This will be
-combined with a retrieval time estimate for each verstion by the middleware to
-determine what data version should be retrieved.  This estimate will be
-disucssed elsewhere in this proposal.
+Negotiating the data quality/retrieval time tradeoff reuqires additional
+metadata.  Providing this functionality requries the naming service be able to
+both return a list of data versions and an indication of data quality. This
+will be combined with a retrieval time estimate for each verstion by the
+middleware to determine what data version should be retrieved.  This estimate
+will be disucssed elsewhere in this proposal.
 
@@
 reduce this contention by doing things like reducing the serialized scope to a
-single directory or subtree. While these approaches help, they do not address
-this key scalability limitation of serialized access--even if the metadata
-service is spread across multiple nodes. Instead, we will offer a
-short duration to consistency POSIX view to offer the performance and
-consistency requirements an exascale application demands.
+single directory or subtree or allowing a single process in a collective file
+operation talk with the metadata service and distribute a handle to other
+participants. While these approaches help, they do not address this key
+scalability limitation of serialized access--even if the metadata service is
+spread across multiple nodes. Instead, we will offer a short duration to
+consistency POSIX view to offer the performance and consistency requirements an
+exascale application demands.
 

### 2015-07-05 narrative/intro.tex
editing of intro

@@
 
-This forced the team to decide which pieces of data to save from the large amount of data; in 
-other words, they had to prioritize which were the most important pieces and save only those.
-They had to further break down the still large data into two categories, data which they thought they would touch after one or two days of runs are completed and after the whole simulation is going to be completed (after ten days of runs), and what they would do with these different categories of data (the access pattern of the large data).
+This forced the team to carefully scrutinize the large amount of data to be generated and save only the very most important pieces.
+Further, they had to break this greatly reduced data into two categories: first data they thought they would need to access while the simulation was running or shortly thereafter, and second data they would need after the simulation was finished.
+The former category would be stored on disk, and the later would be archived.
+Additionally, they needed to figure out how to efficiently retrive these different categories of data for post-processing, analysis and visualization.
 
-They then had to develop new application specific data reduction techniques, and encode that into the application code itself on top of ADIOS, the middleware I/O layer.
-They then had to implement a crude discovery system to understand
-which pieces of data were stored on the parallel file system and which pieces were on tape, and also which pieces were moved over to PPPL for immediate analysis by the user.
+Finally, the team had to develop new application specific data reduction techniques and encode them into the ADIOS middleware I/O layer.
+In order to do this, they had to implement a rudimentary discovery system to determine which data were located on disk, and which were on tape archive.
+Additionally, some pieces of data needed to be moved to PPPL for immediate analysis.
 
-This solution worked well for the first day of run but after the second day more problems occurred: the variability in the time that it took to write the data made the I/O prohibitively expensive, because even though the project had reserved the majority of the resources at the OLCF, the file system was shared and sometimes the writing time took less 100 seconds, and at other times it was over 1,000 seconds.
-There was no method for the application or the user to monitor and predict this resource usage and to dynamically adapt to write less data when the I/O time was going to be longer.
-The greatest problem, however, comes after the simulation   completes: even to visualize the reduced data, just to get a first look at the visualization data, it could take over a month because some of the timesteps will be on tape and other timesteps will be on the file system.
-There is no method to just look at the data quickly at some level of quality quickly and understand which regions of the simulation to focus on in more detail.
+This solution worked well for the first day of the run but after the second day significant problems occurred: I/O variability became prohibitively expensive.
+Although the team had reserved resources for the run, the file system was shared, and a 10X variability in I/O speeds was observed.
+What is needed for this problem is a system to monitor, predict and manage resource usage in a dynamically adaptable way.
+Such a system could write more data when the resource is idle, and less data when it is very busy.
+
+However, the most significant problem occurs after the simulation run completes.
+Basic visualizations from the reducted data, just to get a high-level view of the simulation, can take significant time.
+While some timesteps might be on disk, others are archived on tape (where transfer could take weeks).
+Further, visualization of detailed views of the simulation are even more complicated.
 

### 2015-07-05 narrative/placement.tex
reordering of placement

@@
-\subsection{Data Placement and Movement}
-\newcommand{\Sir}{Sirocco}
+\subsection{Managing Data in the Life Cycle}
+\label{sec:managing-data-life}
+
+\newcommand{\Sir}{Sirocco} %So I don't get the spelling wrong!
 
@@
 
-We proposed to advance the state of the art in the three major parts of the
-data lifecycle. First, we will explore the optimizations and techniques to
-scalably manage data when it is generated by the application. Second, we
-will study the management of resources and data at rest. And finally, we will
-investiage how data is consumed, either by being read and processed (for
-example within analytics) or by being purged from the system.
+We address three main concerns for the data lifecycle in this work. First,
+we will explore the initial placement of data during the output process.
+Second, we will study techniques that can enhance migration without making
+the eventual lookup of data unbounded. And finally, we will study how data
+at each tier is managed. %FIXME
 
@@
 
-\subsubsection{Data Placement}
-\label{sec:data-placement}
+\subsubsection{Initial Placement of Data}
+\label{sec:init-plac-data}
 
-Our past work on data
+\paragraph{Background:}
+When an application outputs data the storage and middleware layer needs to
+decide \textit{what} data is placed \textit{where} amongst the storage
+resource available to the system. This decision has far reaching impact
+throughout the life time of the data. Past work on data
 staging\cite{tongipdps15,qiansc15}\cite{docan2012dataspaces}\cite{abbasi2010datastager}
-in HPC systems, coupled with the experience in developing high performance
-I/O transports for ADIOS \cite{liu2014hello,lofstead2008flexible} has
-provided us with insights on how to manage the complex I/O requirements for
-exascale applications, particularly in the face of the aforementioned
-multi-tier storage hierarchy. A key requirements is application-driven
-%
-runtime mechanisms for dynamically managing data placement across the layers
-of the distributed storage hierarchy,
-%
-coordinating data movement 
-%
-and data sharing between the components of the application workflow so as to
-maximize its utility to the application and reduce access costs. Our
-approach to data placement will leverage the increased set of knowledge
-about the data available to the middleware and storage system, potentially
-allowing for greater optimizations as compared to other storage solutions. 
-The key components of our research include
+in HPC systems has shown that different output techniques targetting
+different layers of the storage hierarchy can have a significant impact on
+the overhead observed by the application for I/O operations. A key
+requirements is application-driven runtime mechanisms for dynamically
+managing data placement across the layers of the distributed storage
+hierarchy, coordinating data movement and data sharing between the
+components of the application workflow so as to maximize its utility to the
+application and reduce access costs. Our approach to data placement will
+leverage the increased set of knowledge about the data available to the
+middleware and storage system, potentially allowing for greater
+optimizations as compared to other storage solutions.
 
@@
 
-\paragraph{State of the art:} NAND flash-based devices are being
-increasingly used in HPC storage systems at different levels and for a
-variety of purposes. Several studies such as \cite{multitier}, \cite{sc10li}
-have investigated using compute node local SSDs as storage buffers, for
-example, to temporarily cache checkpoint data to support recovery from
-failures. Research efforts have also explored using deep memory devices for
-data analysis. Active Flash \cite{activeflash} proposes in-situ scientific
-data analysis by directly executing data analysis tasks on emerging storage
-devices. Minerva \cite{minerva} extends the conventional SSD architecture
-using a FPGA-based storage controller to offload data or I/O intensive
-application code to the SSD to accelerate data analysis. However, these
-solutions cannot deal with coupled application workflows, where the data
-exchange and access patterns can be both complex and dynamic. Furthermore,
-some of these solutions require hardware modifications or special access
-priorities to the HPC systems, which can limit their use to specific
-resources.
+% \paragraph{State of the art:} NAND flash-based devices are being
+% increasingly used in HPC storage systems at different levels and for a
+% variety of purposes. Several studies such as \cite{multitier}, \cite{sc10li}
+% have investigated using compute node local SSDs as storage buffers, for
+% example, to temporarily cache checkpoint data to support recovery from
+% failures. Research efforts have also explored using deep memory devices for
+% data analysis. Active Flash \cite{activeflash} proposes in-situ scientific
+% data analysis by directly executing data analysis tasks on emerging storage
+% devices. Minerva \cite{minerva} extends the conventional SSD architecture
+% using a FPGA-based storage controller to offload data or I/O intensive
+% application code to the SSD to accelerate data analysis. However, these
+% solutions cannot deal with coupled application workflows, where the data
+% exchange and access patterns can be both complex and dynamic. Furthermore,
+% some of these solutions require hardware modifications or special access
+% priorities to the HPC systems, which can limit their use to specific
+% resources.
+
+\subsubsection{Managing Data on Multiple Tiers}
+\label{sec:manag-data-mult}
+
+\paragraph{Background:} There is a spectrum of approaches on how
+to leverage a memory and storage hierarchy: On one end of the
+spectrum is to copy or move data from one tier to another, the
+closer the tier to the hierarchy's top, the faster the access to
+the data it contains. On the other end of the spectrum is to use
+higher tiers exclusively for auxiliary data that to some extent
+represents the actual data on the lowest tier. Examples of auxiliary
+data are views, indices, lossy compressions, lower resolution data,
+and summary data (sometimes also known as metadata). There are a
+number of drivers that push approaches to one or the other extreme:
+(1) if a high tier has sufficient room for an application's working
+set, approaches that move the working set into that tier will perform
+well; Conversely, if a high tier very rarely captures working sets,
+it is better to fill it with auxiliary data that speeds up access
+to the working set captured by a lower tier; working sets only
+reduce capacity misses but the concept can be generalized to the
+likelihood of having the right data in place, i.e. proactively
+moving data to high tiers accessed in predictable patterns reduces
+compulsory misses as well; (2) if the application generates parallel
+read/write requests and strong consistency semantics are important,
+approaches that use auxiliary data to speed up access to persistent
+shared data might yield better performance than approaches that
+require coherence overhead to keep all copies on higher tiers
+consistent; (3) if the application's access patterns are mismatched
+with low tier access characteristics, a combination of approaches
+that use auxiliary data and move data between tiers in order to
+convert access patterns into a better match will perform well.
+
+% Any other drivers?
+\paragraph{Approach:} This spectrum enables interesting strategies
+to manage space pressure on high tiers: instead of just evicting
+pieces of data (blocks, pages, objects, files), one can free space
+by ``thinning'' data so that accesses within more limited views or
+lower resolutions can still occur without misses while requests
+outside of views or higher resolutions can leverage the information
+stored in auxiliary data. Conversely, data on a high tier can be
+``enriched'' (e.g. turned into a higher resolution) if there is
+value in doing so.
+
+% \paragraph{Related Work:} Mention Stanford's Legion system.
+
+\paragraph{Challenge:} The decision on when to thin and when to
+enrich data on a particular level depends on \emph{access patterns}.
+Approaches that leveragee access patterns in application at national
+labs has shown to greatly improve performance and reduce the amount
+of overhead required for data management~\cite{he:hpdc13}. The
+challenge is for applications to communicate these access patterns
+to the I/O stack layers that manages tiers outside of the address
+space. In the proposed work we will develop an abstraction for
+\emph{locality} that allow applications or runtime profilers to
+describe locality properties of access patterns, and implement
+services that leverage locality properties to dynamically determine
+the value of thinning or enriching data at a particular tier.
+Consider the example of a road navigation system: the locality
+properties of each navigation client are based on physical constraints
+of speed, direction, likely resolution of map, and the fact that
+clients almost never go offroad.
 
@@
 
-One important research question in this proposal is whether and how a liberated
-Hierarchical Storage Manager designed for online use, such as Sirocco, might be
-employed in supporting the higher level concepts discussed previously.
-Conceptually, in this effort, data is accessed by the middleware IO libraries
-at various tiers in the lower-level storage hierarchy. While Sirocco does
-autonomously group the distributed set of media into like groups and manage
-movement of data between these groups the various motivations to do so include
-both the traditional ones found in classical HSM (Hierarchical Storage
-Management) as well as the novel.
+\paragraph{Background:} One important research question in this proposal is
+whether and how a liberated Hierarchical Storage Manager designed for online
+use, such as Sirocco, might be employed in supporting the higher level
+concepts discussed previously. Conceptually, in this effort, data is
+accessed by the middleware I/O libraries at various tiers in the lower-level
+storage hierarchy. While Sirocco does autonomously group the distributed set
+of media into like groups and manage movement of data between these groups
+the various motivations to do so include both the traditional ones found in
+classical HSM (Hierarchical Storage Management) as well as the novel.
 
@@
 
+\paragraph{Approach:}
 Our research intent involves short term and long term activities. In the short
@@
 movement, so this is contemplated only as a last resort.
-
-\subsubsection{Purging}
-\label{sec:purging}
-
-% The title might have to change...
-
-\paragraph{Background:} There is a spectrum of approaches on how
-to leverage a memory and storage hierarchy: On one end of the
-spectrum is to copy or move data from one tier to another, the
-closer the tier to the hierarchy's top, the faster the access to
-the data it contains. On the other end of the spectrum is to use
-higher tiers exclusively for auxiliary data that to some extent
-represents the actual data on the lowest tier. Examples of auxiliary
-data are views, indices, lossy compressions, lower resolution data,
-and summary data (sometimes also known as metadata). There are a
-number of drivers that push approaches to one or the other extreme:
-(1) if a high tier has sufficient room for an application's working
-set, approaches that move the working set into that tier will perform
-well; Conversely, if a high tier very rarely captures working sets,
-it is better to fill it with auxiliary data that speeds up access
-to the working set captured by a lower tier; working sets only
-reduce capacity misses but the concept can be generalized to the
-likelihood of having the right data in place, i.e. proactively
-moving data to high tiers accessed in predictable patterns reduces
-compulsory misses as well; (2) if the application generates parallel
-read/write requests and strong consistency semantics are important,
-approaches that use auxiliary data to speed up access to persistent
-shared data might yield better performance than approaches that
-require coherence overhead to keep all copies on higher tiers
-consistent; (3) if the application's access patterns are mismatched
-with low tier access characteristics, a combination of approaches
-that use auxiliary data and move data between tiers in order to
-convert access patterns into a better match will perform well.
-
-% Any other drivers?
-
-\paragraph{Approach:} This spectrum enables interesting strategies
-to manage space pressure on high tiers: instead of just evicting
-pieces of data (blocks, pages, objects, files), one can free space
-by ``thinning'' data so that accesses within more limited views or
-lower resolutions can still occur without misses while requests
-outside of views or higher resolutions can leverage the information
-stored in auxiliary data. Conversely, data on a high tier can be
-``enriched'' (e.g. turned into a higher resolution) if there is
-value in doing so.
-
-\paragraph{Related Work:} Mention Stanford's Legion system.
-
-\paragraph{Challenge:} The decision on when to thin and when to
-enrich data on a particular level depends on \emph{access patterns}.
-Approaches that leveragee access patterns in application at national
-labs has shown to greatly improve performance and reduce the amount
-of overhead required for data management~\cite{he:hpdc13}. The
-challenge is for applications to communicate these access patterns
-to the I/O stack layers that manages tiers outside of the address
-space. In the proposed work we will develop an abstraction for
-\emph{locality} that allow applications or runtime profilers to
-describe locality properties of access patterns, and implement
-services that leverage locality properties to dynamically determine
-the value of thinning or enriching data at a particular tier.
-Consider the example of a road navigation system: the locality
-properties of each navigation client are based on physical constraints
-of speed, direction, likely resolution of map, and the fact that
-clients almost never go offroad.
 

### 2015-07-05 narrative/enabling_tech.tex
fixed typo and clarifying enabling tech`

@@
 
-The Sirocco API is much more complex than a standard POSIX call prompting
-layering a more user friendly interface on top. The ADIOS~\cite{adios} API
-offers an API nearly as simple as POSIX, but with the ability to change the
-actual data transport mechanism without changing the source code. This affords
-writing to a single output API while shifting from writing to a POSIX file
-system, some middleware like DataSpaces~\cite{dataspaces}, or some other system
-like Sirocco all without changing the source code. ADIOS has proven effective
-for writing to POSIX file systems, data staging
-systems~\cite{predata,dataspaces,nessie}, workflow systems~\cite{flexpath}, and
-even nothing through the use of a NULL transport.
+The LWFS philosophy focuses on providing a universal core with additional
+semantics and services layered on top or run along side. For this project, the
+Sirocco API is much more complex than a standard POSIX call prompting layering
+a more user friendly interface on top. Placing a thin, optimized layer as part
+of the storage system will offer the kinds of functionality and semantics this
+project proposes. From the user perspective, something like the
+ADIOS~\cite{adios} API offers an API nearly as simple as POSIX, but with the
+ability to change the actual data transport mechanism without changing the
+source code. This affords writing to a single output API while shifting from
+writing to a POSIX file system, some middleware like
+DataSpaces~\cite{dataspaces}, or some other system like Sirocco all without
+changing the source code. ADIOS has proven effective for writing to POSIX file
+systems, data staging systems~\cite{predata,dataspaces,nessie}, workflow
+systems~\cite{flexpath}, and even nothing through the use of a NULL transport.
+Part of this project's goals seeks to understand where the dividing line
+between the storage layer shared by all users and middleware being used by a
+single or small number of users.
 

### 2015-07-05 narrative/intro.tex
editing of intro
fixed typo and clarifying enabling tech`

@@
 for scientific simulation at exascale. 
-The central objective of this proposal, ``minimizing the time knowledge" from 
+The central objective of this proposal, ``minimizing the time knowledge'' from 
 scientific computing, aligns perfectly with this sentiment. 
@@
 While some timesteps might be on disk, others are archived on tape (where transfer could take weeks).
-Further, visualization of detailed views of the simulation are even more complicated.
+Further, visualization and analysis of detailed views of the simulation are even more complicated.
 

### 2015-07-05 narrative/enabling_tech.tex
minor edits

@@
 
-For this project, we plan to use the Sirocco data storage mechanism to
-offer managing the actual data storage on all of the different tiers. We'll add
-a data management layer as part of the file system to handle data discovery and
-metadata tasks. As a middleware layer, we will investigate offering differing
-data compression, fidelity, and storage placement options. Through this 
-combination, we will be able to deliver a complete end-to-end system.
+For this project, we plan to use the Sirocco data storage mechanism to manage
+the actual data storage on all of the different tiers. Our initial thought on
+where the dividing line between the storage layer and the middleware is as
+follows. We will add a data management functionality as part of the storage
+system layer to handle data discovery and metadata tasks. At the middleware
+layer, we will investigate offering differing data compression, fidelity,
+storage placement, and retrieval selection options.  Through this combination,
+we will be able to deliver a complete end-to-end system. We will investigate
+which functionality works best at the different layers for usability and
+scalability.
 
-In this project we will extend ADIOS~\cite{adios} with new APIs and novel techniques
-to coordinate with the storage back-end. The key idea is offering the ability to
-negotiate with the storage system to agree on the tradeoff between data quality
-and response time. We will test our ideas on the next generation OLCF,
-SNL/LANL, and NERSC systems.
-%
-We plan to leverage the many inovations we discovered in 
-the Lightweight File Systems~\cite{lwfs} (LWFS) project and the Ceph storage system project (ceph.com).
-In the LWFS project, Sandia sought to
-demonstrate stripping down a file system to the bare minimum features required
-for any storage system. Then, additional functionality and semantics, such as
-POSIX-style directories and consistency requirements, were added in using
-auxiliary services separate from the stack or by layering providing a new
-interface to the storage system. The project's first phase demonstrated that
-using a core object storage system with integrated security providing
-authentication and authorization services, customized storage interfaces could
-be developed with exactly the overhead required for different applications. The
-second phase of LWFS, known as Sirocco~\cite{sirocco}, is investigating how to
-manage storage devices for efficient and scalable storage performance.  Sirocco
-has been in process for three years and has developed a usable storage infrastructure.
+In this project we will extend ADIOS~\cite{adios} with new APIs and novel
+techniques to coordinate with the storage back-end. The key idea is offering
+the ability to negotiate with the storage system to agree on the tradeoff
+between data quality and response time. We will test our ideas on the next
+generation OLCF, SNL/LANL, and NERSC systems.
 
-The Ceph project (ceph.com) started at UC Santa Cruz in 2004 and
-was initially funded by DOE/NNSA involving LLNL, LANL, and Sandia,
-with the goal to create an object-based parallel file system that
-addresses the well-known metadata service bottleneck of creating
-and accessing 1000s of files comprising a checkpoint~\cite{weil:osdi06}.
-However, after the Ceph kernel client was pulled into the Linux
-mainline in 2010, the focus of Ceph shifted to its object-based
-storage subsystem, also called Reliable Autonomous Distributed
-Object Store (RADOS), which now is the dominant storage system in
-OpenStack production-level deployments and was recently selected
-as the new backend for Flickr and Yahoo! Cloud Services. In this
-project we will leverage two of the key innovations of Ceph: (1)
-Controlled Replication Under Scalable Hashing (CRUSH)~\cite{weil:sc06}
-will help making resource identification and predictable performance
-scalable, and (2) extensible storage object classes~\cite{watkins:ucsctr15}
-will provide the main mechanisms of introducing new APIs,
-metadata vocabulary, and quality of service into the storage stack.
+We plan to leverage the many inovations we discovered in the Lightweight File
+Systems~\cite{lwfs} (LWFS) project and the Ceph~\cite{weil:osdi06} storage
+system project.  In the LWFS project, Sandia sought to demonstrate stripping
+down a file system to the bare minimum features required for any storage
+system. Then, additional functionality and semantics, such as POSIX-style
+directories and consistency requirements, were added in using auxiliary
+services separate from the stack or by layering providing a new interface to
+the storage system. The project's first phase demonstrated that using a core
+object storage system with integrated security providing authentication and
+authorization services, customized storage interfaces could be developed with
+exactly the overhead required for different applications. The second phase of
+LWFS, known as Sirocco~\cite{sirocco}, is investigating how to manage storage
+devices for efficient and scalable storage performance.  Sirocco has been in
+process for three years and has developed a usable storage infrastructure.
+
+The Ceph project (ceph.com) started at UC Santa Cruz in 2004 and was initially
+funded by DOE/NNSA involving LLNL, LANL, and Sandia, with the goal to create an
+object-based parallel file system that addresses the well-known metadata
+service bottleneck of creating and accessing 1000s of files comprising a
+checkpoint~\cite{weil:osdi06}.  However, after the Ceph kernel client was
+pulled into the Linux mainline in 2010, the focus of Ceph shifted to its
+object-based storage subsystem, also called Reliable Autonomous Distributed
+Object Store (RADOS), which now is the dominant storage system in OpenStack
+production-level deployments and was recently selected as the new backend for
+Flickr and Yahoo! Cloud Services. In this project we will leverage two of the
+key innovations of Ceph: (1) Controlled Replication Under Scalable Hashing
+(CRUSH)~\cite{weil:sc06} will help making resource identification and
+predictable performance scalable, and (2) extensible storage object
+classes~\cite{watkins:ucsctr15} will provide the main mechanisms of introducing
+new APIs, metadata vocabulary, and quality of service into the storage stack.

### 2015-07-05 narrative/proposed-work.tex
added labels
edits

@@
 
-\underline{Principle 1:} A knowledge-centric system , let user knowledge  to define data policies. Today Storage System and I/O
-layers are written in a stove-pipe fashion, and quite often don't allow optimizations to take place. We propose to re-design the
-layer in a highly integrated fashion where user will place their intentions into the system and actions will dynamically and statically
-take place to optimize for both the system and for individual requests.
+\underline{Principle 1:} A knowledge-centric system that allows user expertise to define data policies. Today Storage System and I/O
+layers are written in a stove-pipe fashion and usually do not allow for optimizations. We propose to re-design this
+layer in a highly integrated fashion that allows for user input.
+This re-designed layer will then statically, and dynamically optimize the use of resources for both the system and for individual users.
 
 \underline{Principle 2}: Predictable Performance in the SSIO layer needs to be established so more science can be done on the
-exascale Systems. Without predictable performance, the I/O layer performance must be guessed by the users, which quite often
-mean that science is not performed well.
+exascale Systems. Unpredicitable performance in the I/O layer is often met with unpredictable guesses from users which decreases the effectiveness of simulation runs and the associated science.
 
@@
 \input{narrative/placement}
-\input{narrative/migration}
-\input{narrative/eviction-metrics}
+% \input{narrative/migration}
+% \input{narrative/eviction-metrics}
 

### 2015-07-05 narrative/proposed-work.tex
edits

@@
 
-\underline{Principle 1:} A knowledge-centric system that allows user expertise to define data policies. Today Storage System and I/O
+\underline{Principle 1:} A knowledge-centric system that allows user expertise to define data policies. Today SSIO
 layers are written in a stove-pipe fashion and usually do not allow for optimizations. We propose to re-design this
@@
 
-\underline{Principle 2}: Predictable Performance in the SSIO layer needs to be established so more science can be done on the
-exascale Systems. Unpredicitable performance in the I/O layer is often met with unpredictable guesses from users which decreases the effectiveness of simulation runs and the associated science.
+\underline{Principle 2}: Predictable Performance in the SSIO layer needs to be established to increases the efficiency of science on exascale systems.
+Unpredicitable performance in the I/O layer is often met with unpredictable guesses from users which decreases the effectiveness of simulation runs and the associated science.
 
 \underline{Principle 3:} In a data-centric SSIO system, we need user goals to define the data policies. We introduce data-driven
-SSIO optimizations by enabling users to define data manipulations (re-organizations) at a high level and then we express
+SSIO optimizations by enabling users to define data manipulations (re-organizations) at a high level and then express
 policy goals by associating intentions (hints) with data;' e.g. ``this data at this level should be accessible in the next three months.''

### 2015-07-05 narrative/data-description.tex
Scott changes
Lots of edits for clarity.

--- 2015-07-02 narrative/data-description.tex
@@
-\subsection{Description of Data}
-In order for the middleware and storage layer to understand data,
-I/O intentions, and act accordingly, describing data in a way that can
-be understood by the system is needed. In this project, this is achieved 
-by managing data as \textit{objects}.
-An object is the smallest unit of data that consists of raw data and metadata,
-and a user space variable, such as \textit{temperature}, may consist of a collection
-of objects possibly with different accuracy or resolution.
-A key advantage of representing data as \textit{objects}, rather than 
-bytes, is that it allows data semantics, user intentions, QoS requirements, and
-the relationship between user data to be readily captured and embedded with the 
-raw data. As an example, retrieving a 3D field data generated from a simulation and then visualizing it
-requires the coordinates and connectivity variable 1) to be accessed together and 2) with low
-latency in order to achieve a good user experience. The description of data, 
-as enabled by managing data as objects, bridges 
-the semantics gap between applications, middleware, and storage systems,
-and allows the system to understand user-level data, execute QoS requirements and policies, and
-optimize application and system performance.
+\subsection{Data Description}
+In order for the middleware and storage layer to understand data, I/O
+intentions, and act accordingly, data must be described in a way that can be
+understood by the system. In this project, we use the smallest chunk of data
+stored in the storage system as a \textit{chunk}.  A chunk consists of raw
+data and metadata.  A user space variable, such as \textit{temperature}, may
+consist of a collection of chunks possibly with different accuracy or
+resolution with each chunk representing a different portion of the variable.
+A key advantage of having attributes at this lowest level is that data
+semantics, user intentions, QoS requirements, and the relationship between user
+data to be readily captured and embedded with the raw data so that these
+smallest pieces can each be managed independently or collectively, depending on
+the need. As an example, retrieving 3D field data generated from a simulation
+and then visualizing it requires the coordinates and connectivity variable be
+accessed both together and with low latency to achieve a good user experience.
+The data description, as enabled by managing data at this smallest level,
+bridges the semantics gap between applications, middleware, and storage
+systems and allows the system to understand user-level data, execute QoS
+requirements and policies, and optimize application and system performance.
 
-The description of data is essentially metadata and it may include conventional attributes, such as 
-data type, size, dimensionality, relationship to other data, and etc. In particular,
-relationship to other data objects can be captured by implementing \textit{soft links}
-and different types of relationships can be expressed by different types of soft links. 
-In this project, a key metadata attribute
-that enables QoS scheduling, data placement, and ???, is \textit{data utility value}, which
-captures the priority of objects - it is to our belief that
-for exascale science, users must prioritize a small set of objects to be saved on 
-higher-level capacity-limited storage layer
-to avoid the slow access to large-capacity lower-level storage, such as tapes.
-We propose a new technique, described by a utility function plugged in by a user, where objects
-can be cast into multiple buckets of data, and each bucket can be prioritized differently according
-to its utility value.
-Once this is done, the middleware system will construct the description of data, re-organize
-and place data to achieve the desired QoS goals and policies.
+%The new APIs will enable the bridge by specifying selectable
+%performance/quality/- cost tradeoffs from both the application and system
+%perspectives based upon the user guided rules/policy and runtime system
+%monitoring status. It allows the middleware to make best possible decisions
+%from the feedback of storage system knowledge, such that it will embed user
+%intentions and the available system storage.  We want the system to give the
+%users a certain amount of currency in terms of bandwidth, storage space on
+%each level, and latency expectations. These notions will be fuzzy but they
+%will allow the user to make ad-hoc decisions to figure out what needs to be
+%saved.
 
-The user would like the ability to get a certain amount of Quality of Service 
-such that they can then make decisions when the expected bandwidth, 
-for example is less than what they desire and will then place in a certain set 
-of rules which the system can make autonomic decisions to help decide what should be done.
+The data description may be stored with the individual chunks or within a
+seprate metadata service or both to aid QoS requirements. It would contain
+conventional attributes such as data type, size, dimensionality, and the
+relationship to other data. In particular, the relationship to other chunks can
+be captured by including \textit{typed links} each representing a different
+relationship type.  In this project, a key metric stored as a metadata
+attribute is \textit{data utility}.  It enables QoS scheduling, data placement
+decisions, data lifetime decisions, and captures the chunk priority. It is
+our belief that exascale science will require users to prioritize a small set
+of chunks to be saved on higher-level capacity-limited storage to avoid the
+slow access to large-capacity lower-level storage, such as tape.  We propose a
+new technique, described by a utility function provided by a user, where chunks
+can be cast into multiple data buckets and each bucket can be prioritized
+differently according to its utility value.  Once this is done, the middleware
+system will construct the data description and re-organize and place data to
+achieve the desired QoS goals and policies.
+
+In the meanwhile, the user would like a certain Quality of Service such that
+they can then make decisions about what data to access when.
+%for example is less than what they desire and will then place in a certain set
+%of rules which the system can make autonomic decisions to help decide what
+%should be done.
 A key insight in this proposed project has been the increased interaction of
 the application with the storage system. Towards this end, we propose to
-address the quality of service requirements by proivding the application
-with mechanisms to specify the quality of I/O service, interrogate the
-storage system, and react to the responses. Guided by our past work in
-ADIOS \cite{lofstead2008flexible} we propose to explore the design of two
-mechanisms for this purpose. Firstly, we will explore the augmentation of
-I/O applicatition programing interfaces (I/O APIs) to allow applications to
-both specify timing and quality information and also query the storage
-system for timing estimates. Through this interface we expect the
-application and user to gain insights into how long a single output or input
-call will take given the required quality information, and then react to
-these estimates by adjusting the quality or restricting the scope of the
-data required. Likewise we will explore how an application can provide
-timing information to the storage system to allow the storage system to make
-optimization decisions to best meet the requirements from the application.
+address the QoS requirements by providing the application with mechanisms to
+specify the quality of I/O service, interrogate the storage system, and react
+to the responses. Guided by our past work we propose to explore the design of
+mechanisms for this purpose.
 
-We will also explore external data annotations, such as those
-provided by the configuration file in ADIOS. Through the use of these
-external augmentation the user can provide insights to the storage system on
-the relative value of the data, expected life time and performance
-characteristics, as well as relationships between different data sets. With
-this information the storage system can make optimizations specific to a use
-case. We expect these augmentations to be particularly important for
-eviction of data from a storage layer, and migration of data sets to a
-different storage layer.
+Our research will address the following questions:
+1) How can users interact with the storage system efficiently?,
+2) How can users express their intentions into actionable items which can take place in the middleware and/or storage layer?
+3) How can users annotate data with a utility metric?
 
-\paragraph{State of the art:} ADIOS implements a binary-packed data format that allows
-data characteristics such as min, max and index to be wrapped around data chunks. A
-direct benefit is that each data chunk can be operated independently and I/O concurrently
-can be maximized. We will build upon this capability and further augment the format
-to express data utility. Recently DAMSEL \cite{damsel} provides a rich metadata representation
-and management layer that captures the relationship between data blocks for scientific applications.
-This allows application data to be mapped to storage system efficiently, without overburdening
-users to managing complex data model, such as AMR, from the user space. However, the aspect of 
-data importance hasn't been explored, which we believe will be important for exascale storage solutions.
+\paragraph{State of the art:}
+Our current data annotation techniques demonstrated in
+ADIOS~\cite{lofstead:2009:ipdps} implements a binary-packed data format that
+allows data characteristics such as min, max and index to be wrapped around
+data chunks. A direct benefit is that each data chunk can be operated upon
+independently and I/O concurrently can be maximized. We will build upon this
+capability and further augment the format to express data utility.  Recently
+DAMSEL~\cite{damsel} provides a rich metadata representation and management
+layer that captures the relationship between data blocks for scientific
+applications.  This allows application data to be mapped to storage system
+efficiently, without overburdening users to managing complex data model, such
+as AMR, from the user space. However, the aspect of data importance hasn't been
+explored, which we believe will be important for exascale storage solutions.
+{\color{red}\bf there must be other work in this area, and we need to get more
+references}
 
 \paragraph{Proposed research approach:} 
-We will design and develop new APIs through which users can describe the utility of data
-based on their expectations. This can be done by plugging user level functions that will
-partition user level data into objects and be executed by the middleware to calculate the 
-utility value for each object. The utility value along with other metadata will be encapsulated
-by the middleware and utilized by the underlying storage system to management data placement.
+We will design and develop new techniques for data management, description,
+and access that allows users to describe the data utility based on their
+expectations. This can be done by plugging user level functions that
+partition user level data into chunks and be executed by the middleware to
+calculate the utility value for each chunk. The utility value along with other
+metadata will be encapsulated by the middleware and utilized by the underlying
+storage system to management data placement.
+%
+%We will design an updated I/O API that will expand on the current POSIX I/O
+%semantics by adding new information to the I/O request.
 
-We will design an updated I/O API that will expand on the current POSIX I/O
-semantics by adding new information to the I/O request. We will explore the
-set of application level hints that can be easily provided to the storage
-system in order for the storage system to make the most appropriate
-optimization decisions. In particular, we expect the application to provide
-hints on the length of time before the output data is persistent and
-visible to users, and the expected lifetime of the data for data output
-calls. For read calls we expect the set of hints to combine latency and
-precision requirements. Additional hints will also be investigated.
+We will explore the set of application level hints that can be provided easily
+to the storage system. These hints will enable the storage system to make the
+most appropriate optimization decisions for this user in the context of the
+multi-user environment rather than trying to rely on a one-size-fits-all
+approach. In particular, we expect the application to provide hints on the
+length of time before the output data is persistent and visible to users and
+the expected data lifetime for output sets.  For read calls we expect the set
+of hints to combine latency and precision requirements.  Additional hints will
+also be investigated.
 
-We will also develop a new type of querying system that allows application
-to ask the storage system about completion timing information, the response
-being an estimate provided by the storage system. We will explore how these
-query functions can be integrated into common applications with minimal code
-disruptions, and the set of policies that applications can use to respond to
-this new information. We will also explore how applications can query the
-storage system to gain insights on available space in different tiers and
-how applications can tailor their output data to match the available space. 
+We will also develop a new type of querying system that allows an application
+to ask the storage system about completion timing information. The response
+would be an estimated elapsed time for the described operation. We will explore
+how these query functions can be integrated into common applications with
+minimal code disruptions and the set of policies that applications can use to
+respond to this new information. We will also explore how applications can
+query the storage system to gain insights on available space in different tiers
+and how applications can tailor their output data to match the available space.
 Finally we will study the use of an external data annotation system that can
-provide information to the storage system without requiring the application
-to be recompiled.
+provide information to the storage system without requiring the application to
+be recompiled.
 
-\paragraph{Challenges}
-The design of new APIs posses both technical and adoption challenges. Here
-we will only consider the technical challenges. The proposed set of APIs aim
-to expose broad system level environmental information to the application,
-and allow the application to adapt dynamically based on this information.
-One aspect of this challenge is the design of an interrogative API that
-enables a back-and-forth between the application and the storage system to
-enable the application to receive enough information to make adaptation
-decisions.
+%
+From an application level, we will explore these techniques through our ADIOS
+framework using many of the applications to which we have access, including
+XGC1, GTC, and SPECFM3D. We will understand: 1) what hints the users can give
+us, 2) how the users can take away much of their application specific code to
+``system level code'' through the use of these hints.  We will design an API
+that will integrate through the storage layer to give time estimations so that
+users can decide which data priorities will be written and read. The utility
+value for each chunk will place it into one of ``n'' bins where the first bin
+is the most important data and the last bin is the least important data. The
+bin will give it a lifetime in the highest level of storage that the user thinks
+will be needed. For example: in Figure~\ref{fig:ssio-bucket} we have data that
+comes from the XGC1 code.  The user can pick a data-refactoring scheme (see
+Sec~\ref{sec:data-refactor}) which will bucket the particle data into 4
+prioritization levels.  The user can then pick a data utility, such as 3 months
+for the highest priority items and 12 months for the lowest priority data.
+The user will then call an API when they specify all of the chunks they want
+written (particles, two fields, and a mesh in this example) and the system
+will return an estimate of the time it will take to write. The user can then
+use another API call which will determine the lowest priority element (in this
+case the user will write all of the data (P1-P4). The data will then be stored
+on all of the layers of the file system where the two highest priority buckets
+will be stored in the Parallel File System, but the next highest priority items
+will be stored lower in the storage stack. 
+%
+
+\begin{wrapfigure}{r}{0.7\textwidth}
+%\vspace{0ex}
+        \begin{centering} 
+	\includegraphics[scale=0.77]{graphics/SSIO-bucket.pdf}
+        %\vspace{1ex}
+        \caption{Data is given as a set of chunks which includes utility which contains the lifetime in a storage layer, and the priority of the layer.}
+        \label{fig:ssio-bucket}
+        \end{centering}
+%        \vspace{1ex}
+\end{wrapfigure}
+
+We will design new techniques that users can use as part of a data mover, or
+the system can automatically use, to raise or lower these sub-chunks up and
+down the storage stack. The SSIO layer will then manage the system metadata so
+that the user can transparently access any of these sub-chunks without
+knowledge of the layer. This requires that when reading the data, the APIs will
+then get a time estimate of reading, and then allow the users to determine if
+they will read less (only the highest priority sub-chunks) if the time estimate
+for reading exceeds their expectations. Existing Sirocco functionality may be
+sufficient for these purposes, but it must still be proven in practice.
+%
+
+Users will also need to describe information about the data in order to choose
+the ``most optimal'' data-refactoring routine. Currently, we understand that
+users can tell us that the data represents: 1) phase-space (e.g., the particles
+in our example), 2) spatial-changes (e.g., the mesh), 3) space-time (e.g.,
+the fields on the mesh). We need this information to get the best possible data
+refactoring methods. Although users can request a user-defined data-refactoring
+method, system-level methods can use this information to make better decisions.
+We will investigate new techniques and new semantics to take this information
+and use this to best prioritize and re-organize information.
+
+Finally we will investigate additional information on top of the data model.
+This information will be used to indicate data associations that exceed
+traditional data models. For example, the fields and mesh in our example have
+an inherent relationship. A common data model will describe all of the
+variables in the mesh (the coordinates, the units, and the types of elements if
+this were a finite-element mesh). We will explore new mechanisms that will
+group data together making it explicit that M1, F1, and G1 need to be organized
+together.  In otherwords, in will not be helpful if data from M1 and M2 were
+organized together along with only F1. When the users wants to visualize data
+from the field, F1, they will need only M1 and M2 would be additional
+information. Likewise, if the user wants to see F1 and F2, then they will need
+the mesh from M1 and M2. This requires that the time to access all of this
+data is approximately the same requiring similar access times for all of the
+data.  We will investigate what choices the SSIO system must make in order to
+keep these pieces ``together'', from a performance perspective. Can the storage
+system move everything to a lower layer when all of it doesn't fit or is there
+other, less tightly related data can be evicted to preserve the collective
+retrieval performance?  We need to investigate the potential problems when this
+occurs and how the users can specify additional information to ensure they get
+the best possible data from the SSIO layer.
+
+We seek to provide dynamic runtime information to the user and make the
+system more transparent. The end user could analyze or visualize data in a more
+predictable way since they have confidence how much longer they have to wait
+for the data.
+
+%Firstly, we will explore the augmentation of I/O application programing
+%interfaces (I/O APIs) to allow applications to both specify timing and quality
+%information and also query the storage system for timing estimates. If user
+%decide to write or read data to or from hierarchical storage systems, they
+%will rely on the API to send their intentions for inquiry and examine the
+%system status, including how much data they would like to write/read, desired
+%bandwidth, data compression method, etc, or the user can express their
+%intension of writing data right now no matter what the traffic is now.
+%Through this interface we expect the application and user to gain insights
+%into how long a single output or input call will take given the required
+%quality information, and then react to these estimates by adjusting the
+%quality or restricting the scope of the data required. Likewise we will
+%explore how an application can provide timing information to the storage
+%system to allow the storage system to make optimization decisions to best meet
+%the requirements from the application. 
+%
+%Secondly, we will also explore external data annotations, such as those
+%provided by the configuration file in ADIOS. Through the use of these external
+%augmentation the user can provide insights to the storage system on the
+%relative value of the data, expected life time and performance
+%characteristics, as well as relationships between different data sets. With
+%this information the storage system can make optimizations specific to a use
+%case. We expect these augmentations to be particularly important for eviction
+%of data from a storage layer, and migration of data sets to a different
+%storage layer.
+
+\paragraph{Challenges:}
+These new techniques offer both technical and adoption challenges. We will only
+consider the technical challenges. The proposed techniques and features aim to
+expose broad system level environmental information to the application and
+allow the application to adapt dynamically based on this information.  One
+aspect of this challenge is the design of an interrogative API that offers a
+negotiation between the application and the storage system to make adaptation
+decisions.  Since we will work closely with many of the leading LCF
+applications, we need to make sure that our additional APIs and semantics in
+the storage layer will be accepted by these applications and eventually by the
+rest of the community. This requires careful examination of all possible
+design choices to ensure stable APIs and semantics that do not burden the users
+with confusing APIS and semantics that cannot even be used by the system.

### 2015-07-05 narrative/proposed-work.tex
Lots of edits for clarity.

--- 2015-07-05 narrative/proposed-work.tex
@@
 
-We propose to overcome these challenges by pioneering a new {\bf knowledge-centric approach} to a dynamic storage system
-and I/O layer which takes into account user prioritization and utility to optimize the SSIO layer. The approach is based on 
-three underlying principles:
+We propose to overcome these challenges by pioneering a new {\bf
+knowledge-centric approach} to a dynamic storage system and I/O layer which
+takes into account user prioritization and utility to optimize the SSIO layer.
+The approach is based on three underlying principles:
 
-\underline{Principle 1:} A knowledge-centric system , let user knowledge  to define data policies. Today Storage System and I/O
-layers are written in a stove-pipe fashion, and quite often don't allow optimizations to take place. We propose to re-design the
-layer in a highly integrated fashion where user will place their intentions into the system and actions will dynamically and statically
-take place to optimize for both the system and for individual requests.
+\underline{Principle 1:} A knowledge-centric system , let user knowledge  to
+define data policies. Today Storage System and I/O layers are written in a
+stove-pipe fashion, and quite often do not allow optimizations to take place.
+We propose to re-design the layer in a highly integrated fashion where user
+will place their intentions into the system and actions will dynamically and
+statically take place to optimize for both the system and for individual
+requests.
 
-\underline{Principle 2}: Predictable Performance in the SSIO layer needs to be established so more science can be done on the
-exascale Systems. Without predictable performance, the I/O layer performance must be guessed by the users, which quite often
-mean that science is not performed well.
+\underline{Principle 2}: Predictable Performance in the SSIO layer needs to be
+established so more science can be done on the exascale systems. Without
+predictable performance, the I/O layer performance must be guessed by the
+users, which quite often mean that science is not performed well.
 
-\underline{Principle 3:} In a data-centric SSIO system, we need user goals to define the data policies. We introduce data-driven
-SSIO optimizations by enabling users to define data manipulations (re-organizations) at a high level and then we express
-policy goals by associating intentions (hints) with data;' e.g. ``this data at this level should be accessible in the next three months.''
-These high level specifications of desired results will be used in our system for data migration and eviction policies.
+\underline{Principle 3:} In a data-centric SSIO system, we need user goals to
+define the data policies. We introduce data-driven SSIO optimizations by
+enabling users to define data manipulations (re-organizations) at a high level
+and then we express policy goals by associating intentions (hints) with data;
+e.g. ``this data at this level should be accessible within 100 seconds in the
+next three months.'' These high level specifications of desired results will be
+used in our system for data migration and eviction policies.
 

### 2015-07-05 narrative/naming-service.tex
fixed some errors

@@
 
-Our research in this area will address the following questions:\\
-Since our metadata is distributed and partitioned, how do we respond to user
+Our research in this area will address the following questions:
+(1) Since our metadata is distributed and partitioned, how do we respond to user
   requests within a specified time bound with accuracy? How do we estimate the
-  completeness of the response?\\
-For POSIX API requests, how do we respond if only varying data fidelity levels
-  are available for the requested data object?\\
-What level of metadata should we maintain and keep in sync to enhance response
-  quality within time-bound requests?\\
-How much time should be allowed for searching for metadata given the user
+  completeness of the response?
+(2) For POSIX API requests, how do we respond if only varying data fidelity levels
+  are available for the requested data object?
+(3) What level of metadata should we maintain and keep in sync to enhance response
+  quality within time-bound requests?
+(4) How much time should be allowed for searching for metadata given the user
   specified limits for data retrieval since the user will still have to spend
   the data retrieval time from different devices? How do we integrate the read
-  time overhead into the metadata requests?\\
-How do we provide appropriate security while maintaining scalability,
-  particularly with the distributed metadata stores?\\
+  time overhead into the metadata requests?
+(5) How do we provide appropriate security while maintaining scalability,
+  particularly with the distributed metadata stores?
 % Which data block and what quality/type do we respond to a query with?;\\

### 2015-07-06 narrative/qos.tex
added labels
Make the QoS section a bit more concise.

@@
 \subsection{Resource Management and QoS}
+\label{sec:qos}
 
@@
 this case the storage system.
-Performance quality can be expressed in either relative or absolute
-terms. Examples of relative performance terms are ``fair'',
-``proportional'', or ``priority/class-based'' (``the higher the
-priority the better the service'' or ``1st class is better than
-2nd''). The key advantage of relative performance terms is that
-they are easy to implement. The key disadvantage is the difficulty
-of making any guarantees based on those implementations --
-even relative guarantees are mired by the well-known effect of
-priority inversion~\cite{lampson:cacm80}. Examples of absolute
-performance terms are ``rate-based'', ``soft real-time'', and
-``hard real-time''. The key advantage of absolute performance
-terms are that they enable the implementation of strong
-guarantees: absolute terms enable agreements between application
-tasks and storage that we call \emph{reservations} and that do not
-change meaning depending
-on the storage system's workload. The key challenge of absolute
-performance terms is that their implementation is non-trivial,
-especially in large-scale systems.
-
+%Performance quality can be expressed in either relative or absolute
+%terms. Examples of relative performance terms are ``fair'',
+%``proportional'', or ``priority/class-based'' (``the higher the
+%priority the better the service'' or ``1st class is better than
+%2nd''). The key advantage of relative performance terms is that
+%they are easy to implement. The key disadvantage is the difficulty
+%of making any guarantees based on those implementations --
+%even relative guarantees are mired by the well-known effect of
+%priority inversion~\cite{lampson:cacm80}. Examples of absolute
+%performance terms are ``rate-based'', ``soft real-time'', and
+%``hard real-time''. The key advantage of absolute performance
+%terms are that they enable the implementation of strong
+%guarantees: absolute terms enable agreements between application
+%tasks and storage that we call \emph{reservations} and that do not
+%change meaning depending
+%on the storage system's workload. The key challenge of absolute
+%performance terms is that their implementation is non-trivial,
+%especially in large-scale systems.
 One of the key differentiating features this project offers is driving storage
@@
 but least flexible strategy is to simply cap the latency of each
-task, irregardless of its completeness (see for
+task, regardless of its completeness (see for
 example~\cite{decandia:sosp07}). This strategy is applicable when

### 2015-07-06 narrative/qos.tex
minor changes.

@@
 to reason over its performance in terms of reservations and actual
-performance, accepting reservations that are within resource
-availability and don't violate user or system policies.
-
-It is important to note that we use performance reservations to
-reduce interference and to avoid overload conditions. If the system
-uses reservations intelligently, its performance is predictable and
+performance.
+If the system uses reservations intelligently, its performance is predictable and
 efficient because resources are not wasted on interference. However,

### 2015-07-06 narrative/qos.tex
one pass on qos.

@@
 If the system uses reservations intelligently, its performance is predictable and
-efficient because resources are not wasted on interference. However,
-in cases where an appication arrives that needs to run more urgently
+efficient because resources are not wasted on interference.
+In the cases where an appication arrives that needs to run more urgently
 than any of the other already running applications, the system
@@
 by less urgent applications, allowing the urgent application to finish,
-and reverting back to the previous committed resource level. This can
+and later reverting back to the previous committed resource level for the rest. This can
 be automatically enforced via user- (or adiminstrator-) specified \textit{policies}
@@
 this relationship only once.
-
 To achieve effective performance isolation, we seek to expand the resource management
-capabilities in Sirocco to support additional metrics to guide resource management
-decisions. Currently Sirocco offers automatic resource management capabilities,
-based on resource capacity and data resilience requirements. 
- Our basic goal is to offer data-centric metrics that help guide the
-forced resource management decisions.
+capabilities in Sirocco to support additional utilization 
+metrics to guide resource management
+decisions. Currently, Sirocco offers automatic resource management capabilities,
+based on resource capacity and data resilience requirements, and it needs to be extended
+to offer data-centric metrics that help guide the forced QoS decisions.
 
@@
 
-\paragraph{Related Work:} Existing efforts on scheduing resources~\cite{io-cop,dorier} attempt to offer
+\paragraph{Related Work:} On current HPC systems, absolute IO performance
+guarantees are not possible. Previous work \cite{lofstead:2010:adaptive,liu_hotstorage} 
+from this project team uses a client-assisted approach to mitigate IO variablity.
+Existing efforts on scheduing resources~\cite{io-cop,dorier} attempt to offer
 admission control to maximize storage performance. Unfortunately, these efforts

### 2015-07-06 narrative/data-description.tex

@@
 intentions, and act accordingly, data must be described in a way that can be
-understood by the system. In this project, we use the smallest chunk of data
-stored in the storage system as a \textit{chunk}.  A chunk consists of raw
+understood by the system. In this project, we use the label the smallest unit of data
+saved in the storage system as a \textit{chunk}.  A chunk consists of raw
 data and metadata.  A user space variable, such as \textit{temperature}, may
-consist of a collection of chunks possibly with different accuracy or
-resolution with each chunk representing a different portion of the variable.
+consist of a collection of chunks, possibly with different accuracy or
+resolution, with each chunk representing a different portion of the variable.
 A key advantage of having attributes at this lowest level is that data

### 2015-07-06 narrative/intro.tex
edits
changes to make this more readable.

@@
 for scientific simulation at exascale. 
-The central objective of this proposal, ``minimizing the time knowledge'' from 
+The central objective of this proposal, ``minimizing the time to knowledge'' from 
 scientific computing, aligns perfectly with this sentiment. 
@@
 
-Take for example the latest use case for the XGC1 simulation from C.
+Take for example the latest use case for the XGC1~\cite{chang2006integrated} simulation from C.
 S.
@@
 Due to physical resource limitations, the simulation was restricted to write about 10~PB initially.
-However, once the time to write and read, together with the financial cost to archive this much data were more fully explored, this plan was again signficantly reduced.
+However, once the time to write and read, together with the financial cost to archive this much data were more fully explored, this plan was again significantly reduced.
 At each stage, the particular expert knowledge regarding the lustre file system and the archival tape storage were not known to the user team.
@@
 The former category would be stored on disk, and the later would be archived.
-Additionally, they needed to figure out how to efficiently retrive these different categories of data for post-processing, analysis and visualization.
+Additionally, they needed to figure out how to efficiently retrieve these different categories of data for post-processing, analysis and visualization.
 
@@
 
+Our proposal will be responsive to theme 2 in the call, and we summarize them in the following table.
+\begin{table}[htbp]
+\vspace{2ex}
+\begin{center}
+\begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{1.6cm} |}\hline
+\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant S2E2 work} & \multicolumn{1}{|c|}{\bf See} \\\hline\hline
+%\multicolumn{2}{|c|}{Collaboration} & 
+\multicolumn{2}{|p {5.5cm}|}{Approaches to improve the ability of SSIO software to support checkpoint/restart;} & 
+General piece we are doing for this... & 
+\S\ref{sec:?}, \S\ref{sec:??}, \\\hline
+1 & Data abstractions & What will we do & \S\ref{sec:?}, \S\ref{sec:?}, \S\ref{sec:?} \\\hline
+2 & Mapping science data models onto hierarchical storage & What we are doing & \S\ref{sec:?}, \S\ref{sec:?} \\\hline
+3 & Mechanisms for data movement across the storage hierarchy & what are we doing. & \S\ref{sec:?} \\\hline
+4 &  Extend the interaction of IO middleware with the resource management system & What we are doing & \S\ref{sec:?}, \S\ref{sec:?}  \\\hline
+5 & Design and implement of IO middleware architectures  & What we are doing & \S\ref{sec:?}, \S\ref{sec:?}, \S\ref{sec:?} \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+
 \subsection{Writing Perspective}
@@
 writing or reading at the same time. By requiring a certain quality
-of service the {\bf {\color {red} system must potentially lock out users with lower currency
-they want to place in the reading or writing of their data}}. The storage
+of service the {\bf  system must potentially lock out users with lower currency
+they want to place in the reading or writing of their data}. The storage
 system must also automatically move data from one tier of the storage to
@@
 important pieces of the data.
-
-Our proposal will be responsive to theme 2 in the call, and we summarize them in the following table.
-\begin{table}[htbp]
-\vspace{2ex}
-\begin{center}
-\begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{1.6cm} |}\hline
-\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant S2E2 work} & \multicolumn{1}{|c|}{\bf See} \\\hline\hline
-%\multicolumn{2}{|c|}{Collaboration} & 
-\multicolumn{2}{|p {5.5cm}|}{Approaches to improve the ability of SSIO software to support checkpoint/restart;} & 
-General piece we are doing for this... & 
-\S\ref{sec:?}, \S\ref{sec:??}, \\\hline
-1 & Data abstractions & What will we do & \S\ref{sec:?}, \S\ref{sec:?}, \S\ref{sec:?} \\\hline
-2 & Mapping science data models onto hierarchical storage & What we are doing & \S\ref{sec:?}, \S\ref{sec:?} \\\hline
-3 & Mechanisms for data movement across the storage hierarchy & what are we doing. & \S\ref{sec:?} \\\hline
-4 &  Extend the interaction of IO middleware with the resource management system & What we are doing & \S\ref{sec:?}, \S\ref{sec:?}  \\\hline
-5 & Design and implement of IO middleware architectures  & What we are doing & \S\ref{sec:?}, \S\ref{sec:?}, \S\ref{sec:?} \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
 

### 2015-07-06 narrative/proposed-work.tex
changes to make this more readable.

@@
 
-\underline{Principle 1:} A knowledge-centric system , let user knowledge  to
+\underline{Principle 1:} {\bf A knowledge-centric system}, let user knowledge  to
 define data policies. Today Storage System and I/O layers are written in a
@@
 
-\underline{Principle 2}: Predictable Performance in the SSIO layer needs to be
+\underline{Principle 2}: {\bf Predictable Performance in the SSIO layer} needs to be
 established so more science can be done on the exascale systems. Without
@@
 
-\underline{Principle 3:} In a data-centric SSIO system, we need user goals to
-define the data policies. We introduce data-driven SSIO optimizations by
+\underline{Principle 3:} {\bf Users goals define data policies} In a data-centric SSIO system. 
+We introduce data-driven SSIO optimizations by
 enabling users to define data manipulations (re-organizations) at a high level

### 2015-07-06 narrative/intro.tex
Polishing intro

@@
 for scientific simulation at exascale. 
-The central objective of this proposal, ``minimizing the time to knowledge'' from 
-scientific computing, aligns perfectly with this sentiment. 
+The central objective of this proposal, ``minimizing the time to knowledge''  
+with respect to scientific computing, aligns perfectly with this sentiment. 
 %Hamming was a pioneer in the fields of coding theory, scientific computing and 
@@
 
-Avoiding impending bottlenecks in exascale scientific discovery will require
-new research into managing and storing the large amounts of data that will be
+As we move forward, avoiding impending bottlenecks in exascale scientific discovery will require
+new research into managing and storing the large volumes of data that are 
 produced by simulations and analyzed for months afterwards.
@@
 characteristics, and current system state. We will investigate novel techniques
-to facilitate efficient mapping of data objects, even partitioning individual
-variables, from the user space onto multiple storage tiers and enable
+to facilitate efficient mapping of data objects
+from the user space onto multiple storage tiers and enable
 application-guided data reductions/transformations to address capacity and

### 2015-07-06 narrative/enabling_tech.tex
working in some references

@@
 
-At the lowest level is the Lightweight File Systems (LWFS) project~\cite{lwfs}.
+At the lowest level is the Lightweight File Systems (LWFS) project~\cite{oldfield:lwfs}.
 As mentioned above, LWFS's goal is to provide the simplest core services for
@@
 project proposes. From the user perspective, something like the
-ADIOS~\cite{adios} API offers an API nearly as simple as POSIX, but with the
+ADIOS~\cite{lofstead:adaptible} API offers an API nearly as simple as POSIX, but with the
 ability to change the actual data transport mechanism without changing the
@@
 
-In this project we will extend ADIOS~\cite{adios} with new APIs and novel
+In this project we will extend ADIOS~\cite{lofstead:adaptible} with new APIs and novel
 techniques to coordinate with the storage back-end. The key idea is offering
@@
 We plan to leverage the many inovations we discovered in the Lightweight File
-Systems~\cite{lwfs} (LWFS) project and the Ceph~\cite{weil:osdi06} storage
+Systems~\cite{oldfield:lwfs} (LWFS) project and the Ceph~\cite{weil:osdi06} storage
 system project.  In the LWFS project, Sandia sought to demonstrate stripping

### 2015-07-06 narrative/naming-service.tex
working in some references

@@
 problems forced by the serialized access to a single source for creating and
-accessing files.  Several efforts~\cite{giga+,pvfs,others} have worked to
+accessing files.  Several efforts~\cite{patil:2007:giga+,carns:pvfs} have worked to
 reduce this contention by doing things like reducing the serialized scope to a
@@
 namespace on the side kept in sync using a transaction process like
-D2T~\cite{d2t} showing that this alternative approach can support traditional
+D2T~\cite{lofstead:2012:txn} showing that this alternative approach can support traditional
 POSIX API calls even though the underlying storage system uses a different

### 2015-07-06 narrative/qos.tex
working in some references

@@
 resources. This traditional approach leads to intereference effects
-\cite{lofstead:2010:adaptive,liu_hotstorage} that can greatly impact IO performance predictability. 
+\cite{lofstead:2010:io-variability,liu_hotstorage} that can greatly impact IO performance predictability. 
 %Further, with the introduction of additional tiers in the
@@
 \paragraph{Related Work:} On current HPC systems, absolute IO performance
-guarantees are not possible. Previous work \cite{lofstead:2010:adaptive,liu_hotstorage} 
+guarantees are not possible. Previous work \cite{lofstead:2010:io-variability,liu_hotstorage} 
 from this project team uses a client-assisted approach to mitigate IO variablity.

### 2015-07-06 narrative/data-description.tex
more refs added

@@
 Our current data annotation techniques demonstrated in
-ADIOS~\cite{lofstead:2009:ipdps} implements a binary-packed data format that
+ADIOS~\cite{lofstead:2009:adaptible} implements a binary-packed data format that
 allows data characteristics such as min, max and index to be wrapped around

### 2015-07-06 narrative/enabling_tech.tex
more refs added

@@
 project proposes. From the user perspective, something like the
-ADIOS~\cite{lofstead:adaptible} API offers an API nearly as simple as POSIX, but with the
+ADIOS~\cite{lofstead:2009:adaptible} API offers an API nearly as simple as POSIX, but with the
 ability to change the actual data transport mechanism without changing the
@@
 writing to a POSIX file system, some middleware like
-DataSpaces~\cite{dataspaces}, or some other system like Sirocco all without
+DataSpaces~\cite{docan:2010:dataspaces}, or some other system like Sirocco all without
 changing the source code. ADIOS has proven effective for writing to POSIX file
-systems, data staging systems~\cite{predata,dataspaces,nessie}, workflow
+systems, data staging systems~\cite{zheng:2010:predata,docan:2010:dataspaces,oldfield:2006:nessie}, workflow
 systems~\cite{flexpath}, and even nothing through the use of a NULL transport.
@@
 At a similar level as ADIOS, but can be used in conjunction, are tools like
-DataSpaces~\cite{dataspaces}. They offer a way to request data set portions
+DataSpaces~\cite{docan:2010:dataspaces}. They offer a way to request data set portions
 from a one component by the next componet in a data flow. In DataSpaces case,
@@
 
-In this project we will extend ADIOS~\cite{lofstead:adaptible} with new APIs and novel
+In this project we will extend ADIOS~\cite{lofstead:2009:adaptible} with new APIs and novel
 techniques to coordinate with the storage back-end. The key idea is offering

### 2015-07-06 narrative/enabling_tech.tex
more refs fixed

@@
 systems, data staging systems~\cite{zheng:2010:predata,docan:2010:dataspaces,oldfield:2006:nessie}, workflow
-systems~\cite{flexpath}, and even nothing through the use of a NULL transport.
+systems~\cite{Dayal:2014:flexpath}, and even nothing through the use of a NULL transport.
 Part of this project's goals seeks to understand where the dividing line

### 2015-07-06 narrative/naming-service.tex
fixed typos
more refs fixed

@@
 
-Hierarchical Storage Management (HSM) systems~\cite{hsm} offer a strict caching
+Hierarchical Storage Management (HSM) systems~\cite{blaze:1992:hsm} offer a strict caching
 approach for managing different storage capacities trading off performance for
@@
 finding data are critical for offering the quality of service guarantees we
-wish to offer. Some potential approaches can build on CRUSH~\cite{ceph} from
+wish to offer. Some potential approaches can build on CRUSH~\cite{weil:ceph} from
 Ceph to offer a map of where to search for data sequentially. Given our
@@
 While pure object stores, such as those popular in the big data
-domain~\cite{memcached,others} avoid this bottleneck by strictly offering an
+domain~\cite{Fitzpatrick:2004:memcached,others} avoid this bottleneck by strictly offering an
 object ID with the application required to manage how this ID maps to something

### 2015-07-06 narrative/intro.tex
Merged Manish' intro

@@
 
-As we move forward, avoiding impending bottlenecks in exascale scientific discovery will require
+Avoiding impending bottlenecks in exascale scientific discovery will require
 new research into managing and storing the large volumes of data that are 
@@
 from the user space onto multiple storage tiers and enable
-application-guided data reductions/transformations to address capacity and
+application-guided data reductions and transformations to address capacity and
 bandwidth bottlenecks, while constraining errors to be within user provided
@@
 % 
+Our proposed research program is guided by the following overarching objectives.
 
+{\em Address emerging storage architecture/technology landscape:}
 We will address the associated I/O and storage challenges in the context of
-current and emerging storage landscapes and expedite insights into critical
-scientific processes, while demonstrating the validity of our approach in key
-DOE domains. We will research techniques for creating and using a Scalable 
-Storage Software Infrastructure, and integrating services from the middleware 
-layer with Storage System capabilities to support novel strategies for distributing 
-data both horizontally and vertically across the storage system.  Negotiation between
-these layers will be provided as a fundamental service, allowing users to
+current and emerging storage landscapes,
+% and expedite insights into critical
+%scientific processes, while demonstrating the validity of our approach in key
+%DOE domains. 
+%We will research techniques for creating and using a Scalable 
+%Storage Software Infrastructure, and integrating services from the middleware 
+%layer with Storage System capabilities to support 
+and will develop novel strategies for distributing 
+data both horizontally and vertically across the layers of these storage systems.  
+Negotiation between these layers will be provided as a fundamental service, allowing users to
 easily save ``the best'' amount of data, which is automatically placed throughout the
@@
 
-The metric we are most interested in optimizing is time to knowledge.  Current
-approaches to addressing the I/O bottleneck fall into two broad categories:
-parallel file system approaches that optimize the throughput for an entire
-system, and I/O middleware approaches that optimize the performance of a single
-application. Both approaches have been successful to date, but are unlikely to
-overcome the remaining obstacles in reaching exascale. 
+{\em Develop a science driven, system aware approach:}
+%Our approach targets the expected characteristics of exascale storage hardware.
+%The storage layer will be partitioned into multiple heterogeneous tiers with
+%vastly different performance characteristics. The layers will be further
+%differentiated by the constraints on capacity and data lifetime within each
+%layer. Tape archival storage will still maintain data long term, but access to
+%this data will be orders of magnitude slower than the next layer and naively
+%accessing data from archival storage will greatly impact productivity.
+%
+Given that storage systems will be composed of multiple heterogeneous tiers with
+vastly different performance characteristics and constraints on capacity and 
+data lifetime they can support, our approach will enable researchers to 
+place more knowledge into the SSIO system so that data can be recast from 
+a sequence of bytes to information, which the system can then place and 
+optimize the time to gain knowledge from exascale simulations.
+
+%The traditional approach to transforming results of scientific computation into
+%knowledge consists of storing the numerical output of a simulation on disk,
+%followed by a post-processing or analysis stage that might take place over
+%several ensuing weeks or months. Clearly, given the computation expected at the
+%exascale, this type of workflow is simply not viable.
+
+{\em Emphasize impacts on science:}
+%The metric we are most interested in optimizing is time to knowledge.  
+%Current approaches to addressing the I/O bottleneck fall into two broad categories:
+%parallel file system approaches that optimize the throughput for an entire
+%system, and I/O middleware approaches that optimize the performance of a single
+%application. Both approaches have been successful to date, but are unlikely to
+%overcome the remaining obstacles in reaching exascale. 
 Instead of trying to optimize throughput, we will seek to reduce the time to knowledge.
 This is the most significant metric for scientific applications where the desired outcome
-is not storing data, but rather executing a knowledge creation and extraction processes 
+is not storing data, but rather executing knowledge creation and extraction processes 
 on the data. 
@@
 
-Our approach targets the expected characteristics of exascale storage hardware.
-The storage layer will be partitioned into multiple heterogeneous tiers with
-vastly different performance characteristics. The layers will be further
-differentiated by the constraints on capacity and data lifetime within each
-layer. Tape archival storage will still maintain data long term, but access to
-this data will be orders of magnitude slower than the next layer and naively
-accessing data from archival storage will greatly impact productivity.
-%
-Our approach in this proposal is to enable researchers to place more knowledge
-into the SSIO system so that data can be recast from a sequence of bytes to
-information, which the system can then place and optimize the time to gain
-knowledge from exascale simulations.  We first motivate this by a problem our
-team has been working on in the past six months, and then recast this problem
-into a series of research questions and challenges which must be addressed in
-order to optimize the SSIO system for exascale science.
+The overarching idea of this proposal is that data from applications can be characterized by different
+%levels of importance and different 
+levels of importance over time, which then can be used to guide the analysis and visualization later to access the most important data and deliver some level of knowledge without waiting for access to the whole original dataset.
+Information about the importance of the data at the application level, information about resource status and performance of the multi-level storage system, and flexible data reorganization and reduction operations at all levels can be utilized to provide a quality of service and allow users to gain insight in a timely manner.
 
-The traditional approach to transforming results of scientific computation into
-knowledge consists of storing the numerical output of a simulation on disk,
-followed by a post-processing or analysis stage that might take place over
-several ensuing weeks or months. Clearly, given the computation expected at the
-exascale, this type of workflow is simply not viable.
+This proposal combines knowledge and expertise 
+from a wide range of sources including ORNL applications, the ADIOS project,
+the SNL Lightweight File Systems project and storage system knowledge, the metadata
+management and storage systems knowledge of UCSC, and the middleware expertise
+of Rutgers. We will integrate this diverse expertise to investigate how to make
+a data, workflow, and application aware storage system.
+This project brings together a team with strong expertise in I/O middleware (ORNL, Rutgers), 
+file system (SNL, UCSC) and storage (UCSC), and connects and
+coordinates these key storage components in a seamless fashion.
 
 %
-\section{Scalable Storage Software Infrastructure}
-Our focus in this proposal is in theme 2: Scalable Storage Software Infrastructure. We will focus on a state of the art
-SSIO solution to solve problems which researchers are beginning to face and will make science on exascale systems
-problematic without new insights and solutions.
+%\section{Scalable Storage Software Infrastructure}
+%Our focus in this proposal is in Theme 2: {\em Scalable Storage Software Infrastructure} 
+%We will focus on a state of the art SSIO solution to solve problems which researchers are 
+%beginning to face and will make science on exascale systems
+%problematic without new insights and solutions.
+Our proposal will be responsive to Theme 2: {\em Scalable Storage Software Infrastructure} of the call, as summarize in Table~\ref{ab:call}.
+\begin{table}[htbp]
+\vspace{2ex}
+\begin{center}
+\begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{1.6cm} |}\hline
+\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant S2E2 work} & \multicolumn{1}{|c|}{\bf See} \\\hline\hline
+%\multicolumn{2}{|c|}{Collaboration} & 
+\multicolumn{2}{|p {5.5cm}|}{Approaches to improve the ability of SSIO software to support checkpoint/restart;} & 
+General piece we are doing for this... & 
+\S\ref{sec:?}, \S\ref{sec:??}, \\\hline
+1 & Data abstractions & What will we do & \S\ref{sec:?}, \S\ref{sec:?}, \S\ref{sec:?} \\\hline
+2 & Mapping science data models onto hierarchical storage & What we are doing & \S\ref{sec:?}, \S\ref{sec:?} \\\hline
+3 & Mechanisms for data movement across the storage hierarchy & what are we doing. & \S\ref{sec:?} \\\hline
+4 &  Extend the interaction of IO middleware with the resource management system & What we are doing & \S\ref{sec:?}, \S\ref{sec:?}  \\\hline
+5 & Design and implement of IO middleware architectures  & What we are doing & \S\ref{sec:?}, \S\ref{sec:?}, \S\ref{sec:?} \\
+\hline
+\end{tabular}
+\end{center}
+\label{tab:call}
+\end{table}
+
+%
+\section{An Illustrative Example}
+The challenges addressed by this proposal are illustrated by a problem our team has been working on in the past six months.
+We use this motivating example to formulate a series of research questions and challenges that must be addressed in order to support exascale science.
 
@@
 Eventually, the user was forced to make very hard decisions in reducing the desired 100~PB over a 10-day run to a mere 5~PB of data.
-A maximum of 2~PB would be on disk, and the rest would be archived.
+A maximum of 2~PB would be on disk at any time, and the rest would be archived.
 
 This forced the team to carefully scrutinize the large amount of data to be generated and save only the very most important pieces.
-Further, they had to break this greatly reduced data into two categories: first data they thought they would need to access while the simulation was running or shortly thereafter, and second data they would need after the simulation was finished.
+Further, they had to break this greatly reduced data into two categories: first data they thought they would need to access while the simulation was running or shortly thereafter, and second data they would need after the whole campaign was finished.
 The former category would be stored on disk, and the later would be archived.
@@
 However, the most significant problem occurs after the simulation run completes.
-Basic visualizations from the reducted data, just to get a high-level view of the simulation, can take significant time.
+Basic visualizations from the reduced data, just to get a high-level view of the simulation, can take significant time.
 While some timesteps might be on disk, others are archived on tape (where transfer could take weeks).
@@
 
-The overarching idea of this proposal is that data from applications can be characterized by different
-levels of importance and different levels of importance over time, which then can be used to guide the analysis and visualization later to access the most important data and deliver some level of knowledge without waiting for access to the whole original dataset.
-Information about the importance of the data at the application level, information about resource status and performance of the multi-level storage system, and flexible data reorganization and reduction operations at all levels can be utilized to provide a quality of service and allow users to gain insight in a timely manner.
-
-To address these challenges we are proposing to combine knowledge
-from a wide range of sources including ORNL applications, the ADIOS project,
-the SNL Lightweight File Systems project and storage system knowledge, the metadata
-management and storage systems knowledge of UCSC, and the middleware expertise
-of Rutgers. We will integrate this diverse expertise to investigate how to make
-a data, workflow, and application aware storage system.
+\subsection{Overview of the proposed approach (can be shortened)}
 
@@
 produced by a set of simulations.
-
-Our proposal will be responsive to theme 2 in the call, and we summarize them in the following table.
-\begin{table}[htbp]
-\vspace{2ex}
-\begin{center}
-\begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{1.6cm} |}\hline
-\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant S2E2 work} & \multicolumn{1}{|c|}{\bf See} \\\hline\hline
-%\multicolumn{2}{|c|}{Collaboration} & 
-\multicolumn{2}{|p {5.5cm}|}{Approaches to improve the ability of SSIO software to support checkpoint/restart;} & 
-General piece we are doing for this... & 
-\S\ref{sec:?}, \S\ref{sec:??}, \\\hline
-1 & Data abstractions & What will we do & \S\ref{sec:?}, \S\ref{sec:?}, \S\ref{sec:?} \\\hline
-2 & Mapping science data models onto hierarchical storage & What we are doing & \S\ref{sec:?}, \S\ref{sec:?} \\\hline
-3 & Mechanisms for data movement across the storage hierarchy & what are we doing. & \S\ref{sec:?} \\\hline
-4 &  Extend the interaction of IO middleware with the resource management system & What we are doing & \S\ref{sec:?}, \S\ref{sec:?}  \\\hline
-5 & Design and implement of IO middleware architectures  & What we are doing & \S\ref{sec:?}, \S\ref{sec:?}, \S\ref{sec:?} \\
-\hline
-\end{tabular}
-\end{center}
-\end{table}
 
@@
 tremendous amount of interference from other users on the system who may be
-writing or reading at the same time. By requiring a certain quality
-of service the {\bf  system must potentially lock out users with lower currency
-they want to place in the reading or writing of their data}. The storage
-system must also automatically move data from one tier of the storage to
+writing or reading at the same time and the time to perform an operation is unpredictable. 
+By requiring a certain quality of service the system could potentially lock out other users. 
+The storage system must also automatically move data from one tier of the storage to
 another, without affecting the quality of service that was offered at request

### 2015-07-06 narrative/data-description.tex
Polishing data description section

@@
 \subsection{Data Description}
-In order for the middleware and storage layer to understand data, I/O
-intentions, and act accordingly, data must be described in a way that can be
-understood by the system. In this project, we use the label the smallest unit of data
-saved in the storage system as a \textit{chunk}.  A chunk consists of raw
-data and metadata.  A user space variable, such as \textit{temperature}, may
+In order for the middleware and storage layer to make decisions
+that are appropriate given the user's intentions, data must be
+described in a way that communicates these intentions to
+the system. In this project, we label the smallest unit of data
+saved in the storage system as a \textit{chunk}.  A chunk consists of not only
+raw data, but also metadata that expresses additional
+knowledge about that data.
+A user space variable, such as \textit{temperature}, may
 consist of a collection of chunks, possibly with different accuracy or
-resolution, with each chunk representing a different portion of the variable.
-A key advantage of having attributes at this lowest level is that data
+resolution, and with each chunk representing
+a different portion of the variable.
+A key advantage of having attributes at this lowest level is that it allows data
 semantics, user intentions, QoS requirements, and the relationship between user
@@
 the need. As an example, retrieving 3D field data generated from a simulation
-and then visualizing it requires the coordinates and connectivity variable be
-accessed both together and with low latency to achieve a good user experience.
-The data description, as enabled by managing data at this smallest level,
+and then visualizing it requires the coordinates and connectivity variables be
+accessed simultaneously and with low latency to achieve a good user experience.
+Managing data at this low level
 bridges the semantics gap between applications, middleware, and storage

### 2015-07-06 narrative/data-description.tex
Polishing data description section

@@
 The data description may be stored with the individual chunks or within a
-seprate metadata service or both to aid QoS requirements. It would contain
+separate metadata service or both to aid QoS requirements. It would contain
 conventional attributes such as data type, size, dimensionality, and the
@@
 
-In the meanwhile, the user would like a certain Quality of Service such that
+A key insight in this proposed project has been the increased interaction of
+the application with the storage system.
+The user would like to have the ability to obtain information about, and even
+negotiate with the system to determin what
+level of Quality of Service is possible given
+a prospective set of I/O operations and current state of the system, so that
 they can then make decisions about what data to access when.
@@
 %should be done.
-A key insight in this proposed project has been the increased interaction of
-the application with the storage system. Towards this end, we propose to
+Towards this end, we propose to
 address the QoS requirements by providing the application with mechanisms to

### 2015-07-06 narrative/management.tex
getting changes for the evaluatoin

@@
-\section*{Management}
+\section{Management}
 \label{sec:management}
@@
 Dr. Klasky will oversee the outcome of this project,  Dr. Abbasi (ORNL) will lead the Application Interface area, and Dr. Lofstead
-will lead the Storage System. This project is focused on the integration of application aware-middleware which communications to the storage system,
-and each of these leaders have a strong track record of tight-collaboration, both working with Dr. Klasky for their Ph.D. at the Georgia
+will lead the Storage System. This project is focused on the tight integration of application aware-middleware and a storage system backend.
+Our team has a strong track record of tight-collaboration, with both Drs. Abbasi and Lofstead both working with Dr. Klasky for their Ph.D. at the Georgia
 Institute of Technology.  Our structure includes an area lead for each major research area plus
@@
 
-We will conduct six-monthly {\bf integration fests}  to validate our ability to combine models, apply
+We will conduct six-monthly {\bf integration tests}  to validate our ability to combine models, apply
 combined models to real problems, and cope with the challenges of real data. We apply this approach routinely
@@
 our ability to predict and explain performance of I/O workloads on the current machines., in different environments;
-our ability to reduce data and charcterize data from an object into multiple bins, 
+our ability to reduce data and characterize data from an object into multiple bins, 
 our ability to integrate our routines into the Sirocco file system and test the feasibily on current and next generation LCF systems, and
 how tools lead to workflow performance improvements, for both reading and writing data from our exemplar applications which we will test.
-
-\begin{table}[t]
-\caption{S2E2 Integration Tests. We list for each the technology area involved and new technologies expected.}
-\centering \footnotesize
-\vspace{2ex}
-\label{tab:ifests}
-%\begin{tabular}{| p{2cm} | p{2.5cm} | p{5cm} | p{5cm} | }
-\begin{tabular}{| l | l |l |l |} 
-\hline
-{\bf \#} &  {\bf Focus} & {\bf New technologies to be introduced} \\\hline
-1 & TBD			 & TBD\ \ \\
-2 &TBD & TBD\\
-3 & TBD         & TBD \\
-4 & TBD                                    & TBD  \\
-5 & TBD                                    & TBD \\
-6 & TBD                                    & TBD  \\\hline
-\end{tabular}
-\vspace{-8ex}
-\end{table}

### 2015-07-06 narrative/proposed-work.tex
getting changes for the evaluatoin

--- 2015-07-05 narrative/proposed-work.tex
@@
 takes into account user prioritization and utility to optimize the SSIO layer.
-The approach is based on three underlying principles:
+The approach is based on two underlying principles:
 
-\underline{Principle 1:} A knowledge-centric system , let user knowledge  to
+\underline{Principle 1:} {\bf A knowledge-centric system design}, let user knowledge  to
 define data policies. Today Storage System and I/O layers are written in a
@@
 
-\underline{Principle 2}: Predictable Performance in the SSIO layer needs to be
+\underline{Principle 2}: {\bf Predictable performance and quality of data in the SSIO layer} needs to be
 established so more science can be done on the exascale systems. Without
@@
 
-\underline{Principle 3:} In a data-centric SSIO system, we need user goals to
-define the data policies. We introduce data-driven SSIO optimizations by
-enabling users to define data manipulations (re-organizations) at a high level
-and then we express policy goals by associating intentions (hints) with data;
-e.g. ``this data at this level should be accessible within 100 seconds in the
-next three months.'' These high level specifications of desired results will be
-used in our system for data migration and eviction policies.
+%\underline{Principle 3:} {\bf Users goals define data policies} In a data-centric SSIO system. 
+%We introduce data-driven SSIO optimizations by
+%enabling users to define data manipulations (re-organizations) at a high level
+%and then we express policy goals by associating intentions (hints) with data;
+%e.g. ``this data at this level should be accessible within 100 seconds in the
+%next three months.'' These high level specifications of desired results will be
+%used in our system for data migration and eviction policies.
 

### 2015-07-06 narrative/data-reduction.tex
Pushing first cut of editing of data reduction.

@@
-\subsection{Data Refactoring}
-\label{sec:data-refactor}
-It is critical that users give us more information about writing and reading from the SSIO layer
-in order to not only gain performance and predictability, but to also guide them to write and read
-the best possible data in a given period of time. In our example from C. S. Chang, along with most
-of the other scientist we work with, there is a prioritization of data, and this information must be
-sent over to our SSIO layer so that the best possible data is delivered to and from the SSIO layer, 
-and these decisions often involve refactoring (re-organizing and reducing the data) . In this section we
-describe the common methods to reduce data and to re-organize data. Since there are infinite possibilities 
-we need to understand what kind of data re-organization will work and when will it work?  We also need to 
-understand if it would take too much time and resources to perform the ``best methods'' and how can we manage the
-cost of the re-factoring of data compared to the eventual gains we will get?  Finally, another critical research question
-is if the scientist will loose too much value from the data during the refactoring process, and whether or not we can
-control this through error bounds?
+\subsection{Data Refactoring} \label{sec:data-refactor}
 
-The first step of re-factoring the data is to understand how much this data can be reduced.  Therefore, 
-it is useful to classify scientific data into two basic categories: structured
-data (in the sense that it satisfies a known or relatively simple model) or
-unstructured data (in which information content follows no obvious or explicit
-model).  Although scientific data generally contains random components (due to
-finite precision, measurement and calibration effects etc.), scientific data is
-never purely random. Broadly speaking, the path from data to knowledge consists
-of extracting underlying models or patterns from the datasets, and interpreting
-the resulting models.
+The classical workflow whereby the entire dataset is written to storage for
+later analysis is no longer viable owing to the sheer amount of data being
+generated. In future, it will be vital to take advantage of a priori
+information when writing and reading from the SSIO layer in order to gain
+performance and predictability, and to prioritize I/O so that the data that
+is most useful for end-user can be written in the time available.
 
-Ideally, the scientist would perform the analysis stage {\em in situ} to
+In the case of C. S. Chang described earlier, in commmon with most other
+applications scientists with whom we collaborate, information is provided on
+which data should be sent over to our SSIO layer in order that the best
+possible information is available the subsequent analysis. The decision as how
+this should be accomplished generally involves refactoring (re-organizing and
+reducing the data).
+
+In this section we describe some common methods to reduce and re-organize data.
+There are many possibilities and the best choice will generally vary from case
+to case but, once settled, will not change from run to run. It will be
+important understand when the time and resources to required to identify and
+perform the ``best methods'' outweighs the eventual gains. Another critical
+research question concerns the quantification and control of the loss of
+information resulting from refactoring the data and using a reduced 
+dataset.
+
+A basic issue allied to refactoring is to understand how much information is
+actually present in a dataset and whether a refactoring based on a reduced
+order representation might prove effective. It is useful to classify scientific
+data into two basic categories: structured data (in the sense that it satisfies
+a known or relatively simple model) or unstructured data (in which information
+content follows no obvious or explicit model).  Although scientific data
+generally contains random components (due to finite precision, measurement and
+calibration effects etc.), useful scientific data is never purely random.
+Broadly speaking, the path from data to knowledge consists of extracting
+underlying models or patterns from the datasets, and interpreting the resulting
+models.
+
+Ideally, the scientist would perform the entire analysis stage {\em in situ} to
 extract the relevant information and, in the unlikely event of this being
@@
 nature, large scale simulations aim to discover new information that is often
-hidden as higher order effects amongst the data deluge. In particular, this
-means if data thinning or truncation is applied then one is in danger of
-effectively throwing out the baby with the bath water. One cannot store the
-entire data set in easily accessible source, due to sheer size, yet one cannot
-reduce it prior to archival without risking losing the information one is
-trying to extract.
+hidden in the form of higher order effects amongst the data deluge. In
+particular, this means if data thinning or truncation is applied willy nilly
+then one is in danger of effectively throwing out the baby with the bath water.
+One cannot store the entire data set in easily accessible source, due to sheer
+size, yet one cannot reduce it prior to archival without risking losing the
+information one is trying to extract.
 
@@
 
-Once we realize that data can be reduced from many different techniques, we understand
-that with more application knowledge, we can reduce the total data size much more than without
-knowledge, so we need to be able to design the SSIO layers to be able to have generic
-data reduction techniques, and generic data re-organization techniques. But we need to system
-to allow give us more semantic knowledge of the data so that we can either: 1) Use an approach
-which can audit the calculation and save values of the objects when the auditor is off more than a given
-amount, and we can also save the auditor in a docker container, which can then be executed as part of 
-SSIO software layer.
+Once we realize that data can be reduced from many different techniques, we
+understand that with more application knowledge, we can reduce the total data
+size much more than without knowledge, so we need to be able to design the SSIO
+layers to be able to have generic data reduction techniques, and generic data
+re-organization techniques. But we need to system to allow give us more
+semantic knowledge of the data so that we can either: 1) Use an approach which
+can audit the calculation and save values of the objects when the auditor is
+off more than a given amount, and we can also save the auditor in a docker
+container, which can then be executed as part of SSIO software layer.
 
 Recognizing access patterns in an application is a key to potentially reducing
-future file read time. Scientific applications often read and analyze data alternately,
-thus by overlapping the two phases with prefetching can significantly
-reduce the overall execution time of the application. Accurate prefetching can
-be achieved with access pattern analysis. This requires an SSIO layer which can
-prefetch data in a asynchronous fashion to take the bytes on the file.
+future file read time. Scientific applications often read and analyze data
+alternately, thus by overlapping the two phases with prefetching can
+significantly reduce the overall execution time of the application. Accurate
+prefetching can be achieved with access pattern analysis. This requires an SSIO
+layer which can prefetch data in a asynchronous fashion to take the bytes on
+the file.
 
-There are many methods to re-organize the data, and to reduce the data and in this section we will list many
-different known methods.  We then list our different approaches to address the many research challenges using
-many well known methods, and a new method that we have been working on, known as an ``auditor'' which can be
-used ``in situ'' with the calculation during the data generation, and then used to dramatically reduce the impact on the SSIO
-layer for both reading and writing.
+There are many methods to re-organize the data, and to reduce the data and in
+this section we will list many different known methods.  We then list our
+different approaches to address the many research challenges using many well
+known methods, and a new method that we have been working on, known as an
+``auditor'' which can be used ``in situ'' with the calculation during the data
+generation, and then used to dramatically reduce the impact on the SSIO layer
+for both reading and writing.
 
-\paragraph{State of the art:}
-{\bf Data Reduction Techniques} Data compression broadly consists of three basic steps: (1) Pre-processing to
-extract as much structure as feasible. This step is variously referred to as
-filtering, preconditioning, and pre- processing but in all cases, the objective
-is to represent the data in a form that becomes more amenable to compression.
-(2) Removal of redundancy such as omission of duplicates etc.  Finally, (3) the
+\paragraph{State of the art:} {\bf Data Reduction Techniques} Data compression
+broadly consists of three basic steps: (1) Pre-processing to extract as much
+structure as feasible. This step is variously referred to as filtering,
+preconditioning, and pre- processing but in all cases, the objective is to
+represent the data in a form that becomes more amenable to compression.  (2)
+Removal of redundancy such as omission of duplicates etc.  Finally, (3) the
 resulting data is compressed using entropy coding. The level of compression
@@
 
-Our previous work~\cite{lakshminarasimhan2011compressing},~\cite{lakshminarasimhan2011compressing},~\cite{gong2012multi},
-~\cite{jenkins2012byte},~\cite{gong2013parlo},~\cite{boyuka2014transparent},~\cite{tang2014improving} has focused on many
-of these well known techniques to not only reduce the data but also re-organize the data. Previously we have examined:
+Our previous
+work~\cite{lakshminarasimhan2011compressing,lakshminarasimhan2011compressing,%
+gong2012multi,jenkins2012byte,gong2013parlo,boyuka2014transparent,%
+tang2014improving}
+has focused on many of these well known techniques to not only reduce the data
+but also re-organize the data. Previously we have examined:
 
 \begin{itemize}
-\item Precision based re-organization. This is where the most significant bytes of the data are
-all group together from each object. This data generally will have a higher utility then the data with
-the least significant bytes.  This process has the following three steps which need to be
-efficiently implemented when data gets placed to the storage system and from the storage 
-system. 
+
+\item Precision based re-organization. This is where the most significant bytes
+	of the data are all group together from each object. This data
+	generally will have a higher utility then the data with the least
+	significant bytes.  This process has the following three steps which
+	need to be efficiently implemented when data gets placed to the storage
+	system and from the storage system.  First, the data needs to be
+	re-arranged which involves memory operations and needs to be done
+	in-situ. This operation involves no communication.  Data needs to be
+	copied, so the memory requirements are increased. At most this requires
+	two copies an individual dataset in memory, and we will investigate
+	techniques to allocate and deallocate this memory if the user will
+	specify that the data will be overwritten after it is written to the
+	storage system. This is often the case for many of the data quantities
+	written form the simulation, but there are many cases where, for
+	example, we want to write all of the particles from a Particle In Cell
+	(PIC) simulation.  Since the particles will be used later in the
+	calculation, we need to duplicate the storage. Our observation with
+	working with the XGC1, GTC, Warp, and PiconGPU simulations (all
+	leadership class simulations) is that we can temporary increase the
+	storage of the particles, and then release them since the temporary
+	arrays used in the calculations are often freed when a PIC iteration is
+	finished. 
 %
-First, the data needs to be re-arranged which involves memory operations and needs
-to be done in-situ. This operation involves no communication.  Data needs to be copied, so
-the memory requirements are increased. At most this requires two copies an individual dataset
-in memory, and we will investigate techniques to allocate and deallocate this memory if the user
-will specify that the data will be overwritten after it is written to the storage system. This is often
-the case for many of the data quantities written form the simulation, but there are many cases
-where, for example, we want to write all of the particles from a Particle In Cell (PIC) simulation. 
-Since the particles will be used later in the calculation, we need to duplicate the storage. Our
-observation with working with the XGC1, GTC, Warp, and PiconGPU simulations (all leadership
-class simulations) is that we can temporary increase the storage of the particles, and then 
-release them since the temporary arrays used in the calculations are often freed when a PIC
-iteration is finished.. 
-%
-\item Frequency-based re-organization. 
-Another common technique to classify the importance of data is to first re-organize the data
-in frequency space and then specify that as the frequency increases the importance of data
-decreases.   Common technique which is used for streaming data, and data reduction techniques such as JPEG-2000~\cite{jpeg2000}
-is to  code streams which have regions of interest that offer several mechanisms to support spatial random access or region of interest access at varying degrees of granularity.
-It is possible to store different parts of the same data using different quality.
-%
-This allows us to place the lowest frequency pieces in the fastest storage, and
-the highest frequency either on the slowest storage tiers, or if the data sizes are prohibitively  
-costly, not even write out the highest frequency. This binning of data in frequency space can
-take as an input (storage bandwidth knowledge, storage size limitations, and user intentions
-of how long the majority of the data will be frequently accessed.
+\item Frequency-based re-organization.  Another common technique to classify
+	the importance of data is to first re-organize the data in frequency
+	space and then specify that as the frequency increases the importance
+	of data decreases.   Common technique which is used for streaming data,
+	and data reduction techniques such as JPEG-2000~\cite{jpeg2000} is to
+	code streams which have regions of interest that offer several
+	mechanisms to support spatial random access or region of interest
+	access at varying degrees of granularity. It is possible to store
+	different parts of the same data using different quality.  This allows
+	us to place the lowest frequency pieces in the fastest storage, and the
+	highest frequency either on the slowest storage tiers, or if the data
+	sizes are prohibitively  costly, not even write out the highest
+	frequency. This binning of data in frequency space can take as an input
+	(storage bandwidth knowledge, storage size limitations, and user
+	intentions of how long the majority of the data will be frequently
+	accessed.  In order to fully take advantage of frequency based
+	re-organization we have devised a scheme where the data pre-conditioner
+	first sorts the data in bins of 1024 words, and then we maintain the
+	sort order, and then we used wavelets and spline fits to then
+	re-organize the values.  This re-organization allows us to understand
+	that the resultant data is smoother now, and can be reduced in each one
+	of the bins.  \end{itemize}
 
-In order to fully take advantage of frequency based re-organization we have devised a scheme where the data pre-conditioner first sorts the data
-in bins of 1024 words, and then we maintain the sort order, and then we used wavelets and spline fits to then re-organize the values.
-This re-organization allows us to understand that the resultant data is smoother now, and can be reduced in each one of the bins.
-\end{itemize}
+\paragraph{Proposed research approach:} We will be looking at a multi-throng
+approach to data-refactoring. One approach allows the user to specify
+information about their data to indicate if the data is spatially varying,
+space-and time varying, just time varying. This allows us to store information
+in NVRAM on each node (when and where necessary) to later preform compression
+in both space and time . We will examine the different techniques for
+compression and re-organization along with examine techniques to use an
+error-estimator (gradients) to create a multi-resolution view of the data which
+can then be sampled, and placed in different quantiles of importance. These
+will then be mapped over into the different layers of the storage hierarchy.
+We will also investigate technique which can use ``plug-ins'' for detection of
+features and store higher resolution data around the feature and lower
+resolution values which are further away from the features. This
+re-organization must occur without increasing the total amount of storage than
+what the users intended. We will also investigate the time it takes to
+re-organize the data and use the multi-layer approach to staging to understand
+where the data gets reduced and where it gets re-organized. Open questions are:
+1) do we move the data over to other nodes to then reduce and re-organize, 2)
+do we first reduce the data and then later re-organize this data on different
+nodes?  3) do we use different cores on a node to do data-refactoring.
+Similarly we will examine the cost of reading the data. It is clear that if we
+only read in the high-importance data it should improve the overall time to
+move the data from the SSIO layer to the application, but what happens when
+users want all of the data? There is an obvious penalty of having to take data
+in a hierarchical layout and then place this back to a single-layer view.
+Additional research questions which must be addressed: Will some types of data
+(PIC data, finite element data, ...) be be conducive  for data-refactoring than
+others? How much semantic-domain knowledge will we need to get a dramatic
+improvement of the storage system over not doing anything special?
 
-\paragraph{Proposed research approach:} 
-We will be looking at a multi-throng approach to data-refactoring. One approach allows the user to specify information about their data
-to indicate if the data is spatially varying, space-and time varying, just time varying. This allows us to store information in NVRAM on each
-node (when and where necessary) to later preform compression in both space and time . We will examine the different techniques for
-compression and re-organization along with examine techniques to use an error-estimator (gradients) to create a multi-resolution view of the
-data which can then be sampled, and placed in different quantiles of importance. These will then be mapped over into the different layers of 
-the storage hierarchy.  We will also investigate technique which can use ``plug-ins'' for detection of features and store higher resolution data
-around the feature and lower resolution values which are further away from the features. This re-organization must occur without increasing the
-total amount of storage than what the users intended. We will also investigate the time it takes to re-organize the data and use the multi-layer
-approach to staging to understand where the data gets reduced and where it gets re-organized. Open questions are: 1) do we move the
-data over to other nodes to then reduce and re-organize, 2) do we first reduce the data and then later re-organize this data on different nodes?
-3) do we use different cores on a node to do data-refactoring. Similarly we will examine the cost of reading the data. It is clear that if we only
-read in the high-importance data it should improve the overall time to move the data from the SSIO layer to the application, but what happens
-when users want all of the data? There is an obvious penalty of having to take data in a hierarchical layout and then place this back to
-a single-layer view.  Additional research questions which must be addressed: Will some types of data (PIC data, finite element data, ...) be
-be conducive  for data-refactoring than others? How much semantic-domain knowledge will we need to get a dramatic improvement of the storage
-system over not doing anything special?
+We will measure these methods using the fusion XGC simulation and the
+corresponding analysis routines which they use for their largest simulations.
+We will also work with Jeroen Tromp's SPECFM3D code, and the fusion GTC code
+from Z. Lin. In all cases we will measure their impact from their current
+simulations, and their projected simulations and analytics on exascale
+machines. Since all three of these codes are part of the OLCF CAAR~\cite{CAAR}
+project for early access, there is strong belief that the simulations will be
+close to exascale simulations and we will be able to run these on prototypes of
+our SSIO layer to examine the full impact of data-refactoring.
 
-We will measure these methods using the fusion XGC simulation and the corresponding analysis routines which they use for their largest
-simulations. We will also work with Jereon Tromp's SPECFM3D code, and the fusion GTC code from Z. Lin. In all cases
-we will measure their impact from their current simulations, and their projected simulations and analytics on exascale machines. Since 
-all three of these codes are part of the OLCF CAAR~\cite{CAAR} project for early access, there is strong belief that the simulations will
-be close to exascale simulations and we will be able to run these on prototypes of our SSIO layer to examine the full impact of data-refactoring.
+\paragraph{High Impact Challenges:} Since the greatest form of data reduction
+on the SSIO layer would be to only write the code, and the input data, and then
+recalculate all of the results, we want to understand how to write a minimal
+set of information, and then place an auditor code in a docker container so
+that data can be re-generated with an approximation of the calculation, but
+only valid over a limited amount of time.This new approach to tackle the data
+overload by using preprocessors in which information is not simply stored in
+the form of raw data, but rather in the form of appropriate parameterized
+algorithms. A priori information on the nature of the data is used to choose
+the basic algorithm while the actual values of the data are used to tune the
+parameters used in the reconstruction procedures. We shall refer to this
+paradigm as an Auditor.
 
-\paragraph{HIgh Impact Challenges:}
-Since the greatest form of data reduction on the SSIO layer would be to only write the code, and the input data,
-and then recalculate all of the results, we want to understand how to write a minimal set of information, and then place an auditor 
-code in a docker container so that data can be re-generated with an approximation of the calculation, but only valid over a limited amount 
-of time.This new approach to tackle the data overload by
-using preprocessors in which information is not simply stored in the form of
-raw data, but rather in the form of appropriate parameterized algorithms. A
-priori information on the nature of the data is used to choose the basic
-algorithm while the actual values of the data are used to tune the parameters
-used in the reconstruction procedures. We shall refer to this paradigm as an
-Auditor.
-
-We will examine the use of several auditors for the simulations which we will work with. One approach is to use a fluid-code to audit a kinetic
-code, which can be a good approximation over short periods of time. Once there is a large deviation (specified by the user) from the auditor
-to the original calculation, the data will be saved. This is a local operation, since the deviations will grow differently in space and in time.
+We will examine the use of several auditors for the simulations which we will
+work with. One approach is to use a fluid-code to audit a kinetic code, which
+can be a good approximation over short periods of time. Once there is a large
+deviation (specified by the user) from the auditor to the original calculation,
+the data will be saved. This is a local operation, since the deviations will
+grow differently in space and in time.

### 2015-07-06 narrative/data-reduction.tex
Another update

@@
 then one is in danger of effectively throwing out the baby with the bath water.
-One cannot store the entire data set in easily accessible source, due to sheer
-size, yet one cannot reduce it prior to archival without risking losing the
-information one is trying to extract.
+One cannot store the entire data set in an easily accessible source, due to
+sheer size, yet one cannot reduce it prior to archival without risking losing
+the information one is trying to extract.  Viewed in this way, the problem
+would appear intractable. However, the above discussion fails to recognize that
+much of the data is redundant in the information theoretic sense.  That is to
+say, the amount of information contained in the data is often significantly
+less than the amount of data.  The difficulty stems from the fact that one does
+not know about this redundancy in advance \emph{without the benefit of 
+priori knowledge}. This knowledge can often be provided by the user or can 
+be acquired through experience of dealing with differing runs of the same
+code.
 
-Viewed in this way, the problem would appear intractable. However, the above
-discussion fails to recognize that much of the data is redundant in the
-information theoretic sense.  That is to say, the amount of information
-contained in the data is often significantly less than the amount of data.  The
-difficulty stems from the fact that one does not know about this redundancy in
-advance.
+Application knowledge means one can sometimes achieve drastically superior
+reduction in the total data compared with what one might achieve otherwise.
+However, in the absence of such knowledge, we must equip the SSIO layers with
+generic data reduction techniques and re-organization techniques.
+Nevertheless, certain basic information on the semantics of the data is needed 
+and must be supported by the overall infrastructure.
 
-Once we realize that data can be reduced from many different techniques, we
-understand that with more application knowledge, we can reduce the total data
-size much more than without knowledge, so we need to be able to design the SSIO
-layers to be able to have generic data reduction techniques, and generic data
-re-organization techniques. But we need to system to allow give us more
-semantic knowledge of the data so that we can either: 1) Use an approach which
-can audit the calculation and save values of the objects when the auditor is
-off more than a given amount, and we can also save the auditor in a docker
-container, which can then be executed as part of SSIO software layer.
-
-Recognizing access patterns in an application is a key to potentially reducing
-future file read time. Scientific applications often read and analyze data
-alternately, thus by overlapping the two phases with prefetching can
-significantly reduce the overall execution time of the application. Accurate
-prefetching can be achieved with access pattern analysis. This requires an SSIO
-layer which can prefetch data in a asynchronous fashion to take the bytes on
-the file.
+Another important consideration is knowledge regarding typical access patterns
+for data in a given application. Scientific applications often read and analyze
+data in a structured fashion and failure to respect this pattern, or to exploit
+it through prefetching, can significantly increase the overall execution time
+of the application. Accurate prefetching can only be achieved with access
+pattern analysis. This requires an SSIO layer which can prefetch data in a
+asynchronous fashion to take the bytes on the file.
 

### 2015-07-06 narrative/data-description.tex

@@
 Our research will address the following questions:
-1) How can users interact with the storage system efficiently?,
+1) How can users interact with the storage system efficiently?
 2) How can users express their intentions into actionable items which can take place in the middleware and/or storage layer?
@@
 Our current data annotation techniques demonstrated in
-ADIOS~\cite{lofstead:2009:adaptible} implements a binary-packed data format that
+ADIOS~\cite{lofstead:2009:adaptible} implement a binary-packed data format that
 allows data characteristics such as min, max and index to be wrapped around
 data chunks. A direct benefit is that each data chunk can be operated upon
-independently and I/O concurrently can be maximized. We will build upon this
-capability and further augment the format to express data utility.  Recently
-DAMSEL~\cite{damsel} provides a rich metadata representation and management
+independently and I/O concurrency can be maximized. We will build upon this
+capability and further augment the format to express data utility.
+Recently
+DAMSEL~\cite{damsel} has provided a rich metadata representation and management
 layer that captures the relationship between data blocks for scientific
 applications.  This allows application data to be mapped to storage system
-efficiently, without overburdening users to managing complex data model, such
-as AMR, from the user space. However, the aspect of data importance hasn't been
-explored, which we believe will be important for exascale storage solutions.
+efficiently, without overburdening users with the management of complex data models, such
+as AMR, from the user space. Despite these efforts, the aspect of data importance has yet to be 
+explored, and we believe it will be vital for exascale storage solutions.
 {\color{red}\bf there must be other work in this area, and we need to get more
@@
 We will design and develop new techniques for data management, description,
-and access that allows users to describe the data utility based on their
-expectations. This can be done by plugging user level functions that
-partition user level data into chunks and be executed by the middleware to
-calculate the utility value for each chunk. The utility value along with other
-metadata will be encapsulated by the middleware and utilized by the underlying
-storage system to management data placement.
+and access that allow users to describe the data utility based on their
+expectations. This can be done by allowing users to plug in functions that
+partition user level data into chunks, which are then executed by the middleware to
+calculate the utility value for each chunk. This value, along with other
+metadata, will be encapsulated by the middleware and utilized by the underlying
+storage system to manage data placement.
 %
@@
 
-We will explore the set of application level hints that can be provided easily
+We will explore the space of application level hints that can be provided easily
 to the storage system. These hints will enable the storage system to make the
@@
 length of time before the output data is persistent and visible to users and
-the expected data lifetime for output sets.  For read calls we expect the set
-of hints to combine latency and precision requirements.  Additional hints will
+the expected data lifetime for output sets. For read calls we envision
+hints that provide latency and precision requirements. Additional hints will
 also be investigated.

### 2015-07-06 narrative/data-reduction.tex

@@
 to case but, once settled, will not change from run to run. It will be
-important understand when the time and resources to required to identify and
-perform the ``best methods'' outweighs the eventual gains. Another critical
+important to understand when the time and resources required to identify and
+perform the ``best methods'' outweighs any eventual gains. Another critical
 research question concerns the quantification and control of the loss of
@@
 Nevertheless, certain basic information on the semantics of the data is needed
-and must be supported by the overall infrastructure.
-
-Another important consideration is knowledge regarding typical access patterns
-for data in a given application. Scientific applications often read and analyze
-data in a structured fashion and failure to respect this pattern, or to exploit
-it through prefetching, can significantly increase the overall execution time
-of the application. Accurate prefetching can only be achieved with access
-pattern analysis. This requires an SSIO layer which can prefetch data in a
+and must be supported by the overall infrastructure.  Likewise, another
+important consideration is knowledge regarding typical access patterns for data
+in a given application. Scientific applications often read and analyze data in
+a structured fashion and failure to respect this pattern, or to exploit it
+through prefetching, can significantly increase the overall execution time of
+the application. Accurate prefetching can only be achieved with access pattern
+analysis. This requires an SSIO layer which is able to prefetch data in an
 asynchronous fashion to take the bytes on the file.
-
-There are many methods to re-organize the data, and to reduce the data and in
-this section we will list many different known methods.  We then list our
-different approaches to address the many research challenges using many well
-known methods, and a new method that we have been working on, known as an
-``auditor'' which can be used ``in situ'' with the calculation during the data
-generation, and then used to dramatically reduce the impact on the SSIO layer
-for both reading and writing.
 
@@
 structure as feasible. This step is variously referred to as filtering,
-preconditioning, and pre- processing but in all cases, the objective is to
+preconditioning, and preprocessing but in all cases, the objective is to
 represent the data in a form that becomes more amenable to compression.  (2)
@@
 transformation through to simply reordering the data.  What these, and other
-catchall methods, fail to exploit is a priori information on the nature and
-provenance of the data.
+catch-all methods, fail to exploit is a priori information on the nature and
+provenance of the data. Moreover, it is now widely recognized that any
+opportunities for {\em in situ} analysis should be exploited as far as
+possible. What is less well understood, is how one should incorporate a priori
+information and how one can identify possibilities for {\em in situ} analysis.
 
-As we have already observed, the incorporation of a priori information of this
-type is vital if one is to obtain significant data reduction. The use of an
-Auditor as a pre-processor exploits the wide (and increasing) cost gap between
-CPU operations versus storage-based operations. The preprocessor would ideally
-be employed in situ and is used in a similar fashion to pre-processors in
-standard compression approaches such as FPC \cite{BurtscherFPC} in which an XOR
-operation against the previous data is used as a pre-processor.
+We shall exploit a priori information in the preprocessing stage through the
+use of relatively computationally intensive preprocessors that attempt to
+extract structure in the data {\em in situ}. This approach takes advantage of
+the increasingly wide cost gap between CPU operations versus storage-based
+operations. The preprocessor would generally be employed in situ and is used in
+a similar fashion to pre-processors in standard compression approaches such as
+FPC \cite{BurtscherFPC} in which an XOR operation against the previous data is
+used as a pre-processor.
 

### 2015-07-06 narrative/evaluation.tex
evaluation section

\section{Evaluation}
\label{sec:evaluation}
In order to determine if our guiding principles are being met, and we can optimize the system from these guiding principles, 
we will evaluate our work using the Table~\ref{table:eval}.

\begin{tabular}{ | l | l | l | }

\hline
	{\bf Principle/Metric}                   & {\bf User perspective metric}                     & {\bf System perspective metric }\\ \hline
	{\bf Knowledge-centric system} & More knowledge = faster access            &  More knowledge = better resource\\
                                                               &    to higher quality data                               &    utilization across the entire hierarchy \\ \hline

        {\bf Predictable Performance/Accuracy} & More predictability =                     &  Accuracy of the time bounds across \\
                                                                             & better quality of data                      &  multiple layers of the storage hierarchy \\ 
                                                                             & wrote/read in time constrained\\
                                                                             & situations \\ \hline

\end{tabular}

In addition to benchmarking the various S2E2 methods and tools in isolation, we will work with our science partners at
Princeton University (J. Tromp), PPPL (C. S. Chang), and U. C. Irvine (Z. Lin), along with the OLCF  and SNL, to evaluate the 
leading edge applications along with the system performance. We will use the systems at both SNL and OLCF, which have 
multiple layers of storage. For example, the sith cluster at the OLCF (https://www.olcf.ornl.gov/computing-resources/sith/) 
has its own Lustre storage system, SSDs on each node, and is connected to the centerwide Lustre file system. We will perform 
test in both the system level (on sith) and in the user level (on titan), and evaluate our results using our application knowledge.

Part of our evaluation will include how accurate we can set the bounds on time estimates for applications. We will measure these
estimates when data is on just one level of the storage hierarchy, along with the time estimates when data chunks are split. We 
will evaluate how fast data can be discovered when it is split between the hierarchy, allowing us to better understand how many 
pieces data can be split into before it takes too long to access most of the data.  We will also evaluate our techniques to examine
how well we can migrate data and update our server to determine if data might get out-of-sync between data chunks. We will
need to clearly understand how well this system works on current systems and we will need to investigate this on the next generation
OLCF system (Summitt) which we will have early access, along with all of the applications we will be investigating.

Another area of investigation will be in how accurate data can be read in when there are time constraints due to users wanting
to read in many time slices in a given amount of time. For example, in J. Tromp's visualization, we may contain hundreds of time
slices, but when we visualize this data we would like very fast access to the ``first look' of the data. The question is whether the data
will be highly representative of the overall data. Can we make accurate visualization from the less accurate data?  We will work
with our application partners to access this, and we will determine the accuracy in time-constrained test vs. test which will not 
place any bounds on the time to retrieve the data.

We will also want to access how well our data migration and eviction policies work with user-knowledge. Will the overall system
be able to utilize more of the storage hierarchy by understanding these intentions and creating autonomic rules to self-optimize?
Furthermore we will want to see if users can give us more accurate policies which can then be used for better system resource
utilization. If we have more knowledge we would like to be able to illustrate that this will translate into higher levels of system
utilization.

We will also test to see the burden of users to place more knowledge into the system compared to the overall effect of more optimal
performance they will get. We realize that initially it will be a large burden for users to place initial knowledge into the system, but 
we will evaluate the results as users increase the knowledge placed into the system with the overall burden. Our hypothesis from 
working with many LCF users is that once they get over the initial burden of placing this knowledge the burden decreases as more 
knowledge is placed into the system, and thus, they will be more inclined to increase the knowledge into the system as the benefits
increase.

We will also need to evaluate how more knowledge into the system will lead into more performance they get from placing
the most important pieces at the higher layers of the hierarchy. We will evaluate this by increasing and decreasing this knowledge.
For example, if the user doesn't want their data broken into priority chunks, then the data will be at the mercy of what the system
does with no knowledge. If the user then gives us information that allows us to use ``generic'' rouitnes based on frequency, or 
accuracy, then we can use that knowledge to decompose the objects into chunks which can be placed. We can evaluate these
against users who will give us more knowledge which allows us to perform application level binning of data (for example for
C.S. Chang), or for level of features (for J. Tromp).

How will we quantify the quality of data?
We will look at errors after analytics.

How well does the system utilize all of its resources. We feel the more knowledge means the better the system can utilize all of the resources.

predictability ... are teh errors bounded,

### 2015-07-06 narrative/data-description.tex

@@
 expectations. This can be done by allowing users to plug in functions that
-partition user level data into chunks, which are then executed by the middleware to
-calculate the utility value for each chunk. This value, along with other
-metadata, will be encapsulated by the middleware and utilized by the underlying
-storage system to manage data placement.
+provide data-specific functionality to the middleware and storage system.
+These functions may be executed by the system to
+calculate the utility value for each chunk, handle custom compression
+and decompression, and perform other data-specific services, thus allowing
+the system to more effectively manage data placement and retrieval.
 %
@@
 to ask the storage system about completion timing information. The response
-would be an estimated elapsed time for the described operation. We will explore
+would be an estimated time for the described operation. We will explore
 how these query functions can be integrated into common applications with
@@
 query the storage system to gain insights on available space in different tiers
-and how applications can tailor their output data to match the available space.
+and how applications can tailor their output data to most effectively use the available space.
 Finally we will study the use of an external data annotation system that can
@@
 framework using many of the applications to which we have access, including
-XGC1, GTC, and SPECFM3D. We will understand: 1) what hints the users can give
-us, 2) how the users can take away much of their application specific code to
+XGC1, GTC, and SPECFM3D. We will investigate: 1) what hints the users can give
+the system, 2) how the users can replace much of their
+application specific code with
 ``system level code'' through the use of these hints.  We will design an API
-that will integrate through the storage layer to give time estimations so that
+that will be integrated with the storage layer to give time estimations so that
 users can decide which data priorities will be written and read. The utility
-value for each chunk will place it into one of ``n'' bins where the first bin
-is the most important data and the last bin is the least important data. The
-bin will give it a lifetime in the highest level of storage that the user thinks
-will be needed. For example: in Figure~\ref{fig:ssio-bucket} we have data that
+value for each chunk will be used to sort the chunks into prioritized bins.
+Chunks in a particular bin will be assigned an appropriate storage
+ tier and lifetime, which can be decided by the user through the API.
+For example: in Figure~\ref{fig:ssio-bucket} we have data that
 comes from the XGC1 code.  The user can pick a data-refactoring scheme (see
@@
 for the highest priority items and 12 months for the lowest priority data.
-The user will then call an API when they specify all of the chunks they want
+The user will then call a function, provided by the API, that allows them
+to specify all of the chunks they want
 written (particles, two fields, and a mesh in this example) and the system
 will return an estimate of the time it will take to write. The user can then
-use another API call which will determine the lowest priority element (in this
+use another call which will determine the lowest priority element (in this
 case the user will write all of the data (P1-P4). The data will then be stored
-on all of the layers of the file system where the two highest priority buckets
+across the layers of the file system where the two highest priority buckets
 will be stored in the Parallel File System, but the next highest priority items

### 2015-07-06 narrative/evaluation.tex
e
resized the table

@@
 	{\bf Principle/Metric}                   & {\bf User perspective metric}   & {\bf System perspective metric }\\ \hline
-	{\bf Knowledge-centric system} & More knowledge = faster access            &  More knowledge = better resource\\
-                                                               &    to higher quality data                               &    utilization across the entire hierarchy \\ \hline
+	{\bf Knowledge-centric}                  & More knowledge = faster access  &  More knowledge = better \\
+        {\bf system}                             & to higher quality data          &  resource utilization across \\
+                                                 &                                 & the entire hierarchy \\ \hline
 
-        {\bf Predictable Performance/Accuracy} & More predictability =                     &  Accuracy of the time bounds across \\
-                                                                             & better quality of data                      &  multiple layers of the storage hierarchy \\ 
-                                                                             & wrote/read in time constrained\\
-                                                                             & situations \\ \hline
+        {\bf Predictable Performance} & More predictability =                     &  Accuracy of the time bounds \\
+        {\bf Accuracy}                & better quality of data                    &  across multiple layers of the\\
+                                      &                                           &  storage hierarchy wrote/read \\ 
+                                      &                                           &  in time constrained\\
+                                      &                                           &  situations \\ \hline
 

### 2015-07-06 narrative/data-reduction.tex
small change

@@
 
-In the case of C. S. Chang described earlier, in commmon with most other
+In the case of C. S. Chang described earlier, in common with most other
 applications scientists with whom we collaborate, information is provided on
 which data should be sent over to our SSIO layer in order that the best
-possible information is available the subsequent analysis. The decision as how
-this should be accomplished generally involves refactoring (re-organizing and
+possible information is available for the subsequent analysis. The decision as how
+this should be accomplished generally involves data refactoring (re-organizing and
 reducing the data).
@@
 
-Ideally, the scientist would perform the entire analysis stage {\em in situ} to
+Ideally, the scientists would perform the entire analysis stage {\em in situ} to
 extract the relevant information and, in the unlikely event of this being
@@
 in NVRAM on each node (when and where necessary) to later preform compression
-in both space and time . We will examine the different techniques for
+in both space and time.
+
+We will examine the different techniques for
 compression and re-organization along with examine techniques to use an
@@
 will then be mapped over into the different layers of the storage hierarchy.
+
 We will also investigate technique which can use ``plug-ins'' for detection of
@@
 re-organization must occur without increasing the total amount of storage than
-what the users intended. We will also investigate the time it takes to
+what the users intended.
+
+We will also research the time it takes to
 re-organize the data and use the multi-layer approach to staging to understand

### 2015-07-06 coverpage.tex

@@
 
-\pagebreak
-
-\begin{tabular}{| l | l | l | r | r | r | r |} \hline
-Role & Name & Institution & Year 1 (\$K) & Year 2 (\$K) &   Year 3 (\$K) & Total (\$K) \\
-\hline
-Lead-PI & Scott Klasky & ORNL & 660 & 660 & 660 & 1,980 \\
-Lead PI & Jay Lofstead & Sandia  & 350 & 350 & 350 & 1,050 \\
-Co-PI & Carlos  Maltzahn & UCSC  & 160 & 160 & 160 & 480 \\
-Co-PI & Manish Parashar & Rutgers & 80 & 80 & 80 & 240 \\
-\hline
-{\bf Totals}  & & & {\bf 1,250} & {\bf 1,250} & {\bf 1,250} & {\bf 3,750}\\
-\hline
-\end{tabular}
-
 \newpage

### 2015-07-06 narrative/qos.tex
more refs fixed
some quick edits.

@@
 performance.
-If the system uses reservations intelligently, its performance is predictable and
-efficient because resources are not wasted on interference.
 In the cases where an appication arrives that needs to run more urgently
@@
 by less urgent applications, allowing the urgent application to finish,
-and later reverting back to the previous committed resource level for the rest. This can
-be automatically enforced via user- (or adiminstrator-) specified \textit{policies}
+and later reverting back to the previous committed resource level for the rest.
+This can be automatically enforced via user- (or adiminstrator-) specified \textit{policies}
 (section \ref{}).
@@
 thus the occurrence of congestion and degraded performance.
-Traditionally, there is a disconnection between storage infrastructure
-and rest of system software and applications: the infrastructure
-details are not exposed anywhere at all. One argument favoring this
-is to ensure a platform agnostic design.
 
@@
 from this project team uses a client-assisted approach to mitigate IO variablity.
-Existing efforts on scheduing resources~\cite{io-cop,dorier} attempt to offer
+Existing efforts on scheduing resources~\cite{thapaliya:2014:io-cop,dorier:2014:calciom} attempt to offer
 admission control to maximize storage performance. Unfortunately, these efforts

### 2015-07-06 narrative/qos.tex
qos.

@@
 
-\paragraph{Resource Management Background:} There is an important connection between finding the resources that can help fulfill a particular request and the predictability of the performance of that request, both in terms of latency and resolution.
-This connection is particular important for reads: in very large scale systems, finding a resource to read can be a significant, if not dominant, component of overall read latency.
-If the search for resources is unbounded, it will be difficult to make any performance predictions.
+\paragraph{Resource Management Background:} There is an important connection between finding the resources
+that can help fulfill a particular request and the performance predictability of that request, which demands
+careful resource management strategies. This connection is particularly important for reading - in very large 
+scale systems, finding a resource to read can be a significant, if not dominant, component of overall read 
+latency.
 
-Locating resources is a typical distributed systems problem.
-Central solutions are limited in their scalability, especially if they have to track every update.
-Traditional file systems are good examples for this: most updates to files require updates to their centrally maintained block allocation tables.
-The opposite of file systems are ``pile'' systems which instead of keeping track of updates search for resources at request time.
-Pile systems allow very efficient writes because clients can write whereever there is free space and there is very bookkeeping overhead.
-Furthermore, pile systems have great flexibility for moving resources around, for example to ensure durability.
-However, reads require a potentially unbounded search, especially if it is not clear whether the resource actually exists.
-Methods to make searches more efficient without reverting to centralized bookkeeping usually involves some kind of partitioning of the search space and a method to limit the search to a small number of partitions.
+Locating resources is a typical distributed systems problem. Centralized solutions are limited in their
+scalability, because typically there exists a single entity that tracks every update in the system. 
+Traditional file systems are good examples 
+for this: most updates to files require updates to their centrally maintained block allocation tables. 
+The opposite of file systems are ``pile'' systems, which, instead of keeping track of updates, search for 
+resources at request time. Pile systems allow very efficient writes because clients can write whereever
+there is free space and there is very low bookkeeping overhead. Furthermore, pile systems have great
+flexibility for moving resources around, for example to ensure durability. However, reads require a 
+potentially unbounded search, especially if it is not clear whether the resource actually exists. 
+Methods to make searches more efficient without reverting to centralized bookkeeping usually involves 
+some kind of partitioning of the search space and a method to limit the search to a small number of partitions.
 

### 2015-07-06 narrative/qos.tex
done for now and will come back to it.

@@
 \paragraph{Resource Management Background:} There is an important connection between finding the resources
-that can help fulfill a particular request and the performance predictability of that request, which demands
-careful resource management strategies. This connection is particularly important for reading - in very large 
-scale systems, finding a resource to read can be a significant, if not dominant, component of overall read 
-latency.
-
+that can help fulfill a particular request and the performance predictability of that request, and this demands
+careful resource management strategies that deal with large number of data chunks spread across storage hierarchy. 
+%This connection is particularly important for data retrieval 
+%- in large scale storage systems, finding a resource to read can be a significant, if not dominant, 
+%component of overall read latency. 
 Locating resources is a typical distributed systems problem. Centralized solutions are limited in their
-scalability, because typically there exists a single entity that tracks every update in the system. 
-Traditional file systems are good examples 
-for this: most updates to files require updates to their centrally maintained block allocation tables. 
+scalability, because there exists a single entity that tracks updates in the system. 
+Traditional file systems are good examples for this: most updates to files require updates 
+to their centrally maintained block allocation tables. 
 The opposite of file systems are ``pile'' systems, which, instead of keeping track of updates, search for 
 resources at request time. Pile systems allow very efficient writes because clients can write whereever
-there is free space and there is very low bookkeeping overhead. Furthermore, pile systems have great
-flexibility for moving resources around, for example to ensure durability. However, reads require a 
-potentially unbounded search, especially if it is not clear whether the resource actually exists. 
+there is free space and there is very low bookkeeping overhead. However, unbounded search is possible, 
+if it is unclear whether the resource actually exists. 
 Methods to make searches more efficient without reverting to centralized bookkeeping usually involves 

### 2015-07-06 narrative/evaluation.tex
filling the lines so the text is somewhat readable

@@
 \label{sec:evaluation}
+
 In order to determine if our guiding principles are being met, and we can optimize the system from these guiding principles, 
@@
 
-In addition to benchmarking the various S2E2 methods and tools in isolation, we will work with our science partners at
-Princeton University (J. Tromp), PPPL (C. S. Chang), and U. C. Irvine (Z. Lin), along with the OLCF  and SNL, to evaluate the 
-leading edge applications along with the system performance. We will use the systems at both SNL and OLCF, which have 
-multiple layers of storage. For example, the sith cluster at the OLCF (https://www.olcf.ornl.gov/computing-resources/sith/) 
-has its own Lustre storage system, SSDs on each node, and is connected to the centerwide Lustre file system. We will perform 
-test in both the system level (on sith) and in the user level (on titan), and evaluate our results using our application knowledge.
+In addition to benchmarking the various S2E2 methods and tools in isolation,
+we will work with our science partners at Princeton University (J. Tromp),
+PPPL (C. S. Chang), and U. C. Irvine (Z. Lin), along with the OLCF and SNL,
+to evaluate the leading edge applications along with the system performance.
+We will use the systems at both SNL and OLCF, which have multiple layers of
+storage. For example, the sith cluster at the OLCF
+(https://www.olcf.ornl.gov/computing-resources/sith/) has its own Lustre
+storage system, SSDs on each node, and is connected to the centerwide Lustre
+file system. We will perform test in both the system level (on sith) and in
+the user level (on titan), and evaluate our results using our application
+knowledge.
 
-Part of our evaluation will include how accurate we can set the bounds on time estimates for applications. We will measure these
-estimates when data is on just one level of the storage hierarchy, along with the time estimates when data chunks are split. We 
-will evaluate how fast data can be discovered when it is split between the hierarchy, allowing us to better understand how many 
-pieces data can be split into before it takes too long to access most of the data.  We will also evaluate our techniques to examine
-how well we can migrate data and update our server to determine if data might get out-of-sync between data chunks. We will
-need to clearly understand how well this system works on current systems and we will need to investigate this on the next generation
-OLCF system (Summitt) which we will have early access, along with all of the applications we will be investigating.
+Part of our evaluation will include how accurate we can set the bounds on
+time estimates for applications. We will measure these estimates when data
+is on just one level of the storage hierarchy, along with the time estimates
+when data chunks are split. We will evaluate how fast data can be discovered
+when it is split between the hierarchy, allowing us to better understand how
+many pieces data can be split into before it takes too long to access most
+of the data. We will also evaluate our techniques to examine how well we can
+migrate data and update our server to determine if data might get
+out-of-sync between data chunks. We will need to clearly understand how well
+this system works on current systems and we will need to investigate this on
+the next generation OLCF system (Summitt) which we will have early access,
+along with all of the applications we will be investigating.
 
-Another area of investigation will be in how accurate data can be read in when there are time constraints due to users wanting
-to read in many time slices in a given amount of time. For example, in J. Tromp's visualization, we may contain hundreds of time
-slices, but when we visualize this data we would like very fast access to the ``first look' of the data. The question is whether the data
-will be highly representative of the overall data. Can we make accurate visualization from the less accurate data?  We will work
-with our application partners to access this, and we will determine the accuracy in time-constrained test vs. test which will not 
-place any bounds on the time to retrieve the data.
+Another area of investigation will be in how accurate data can be read in
+when there are time constraints due to users wanting to read in many time
+slices in a given amount of time. For example, in J. Tromp's visualization,
+we may contain hundreds of time slices, but when we visualize this data we
+would like very fast access to the ``first look' of the data. The question
+is whether the data will be highly representative of the overall data. Can
+we make accurate visualization from the less accurate data? We will work
+with our application partners to access this, and we will determine the
+accuracy in time-constrained test vs. test which will not place any bounds
+on the time to retrieve the data.
 
-We will also want to access how well our data migration and eviction policies work with user-knowledge. Will the overall system
-be able to utilize more of the storage hierarchy by understanding these intentions and creating autonomic rules to self-optimize?
-Furthermore we will want to see if users can give us more accurate policies which can then be used for better system resource
-utilization. If we have more knowledge we would like to be able to illustrate that this will translate into higher levels of system
-utilization.
+We will also want to access how well our data migration and eviction
+policies work with user-knowledge. Will the overall system be able to
+utilize more of the storage hierarchy by understanding these intentions and
+creating autonomic rules to self-optimize? Furthermore we will want to see
+if users can give us more accurate policies which can then be used for
+better system resource utilization. If we have more knowledge we would like
+to be able to illustrate that this will translate into higher levels of
+system utilization.
 
-We will also test to see the burden of users to place more knowledge into the system compared to the overall effect of more optimal
-performance they will get. We realize that initially it will be a large burden for users to place initial knowledge into the system, but 
-we will evaluate the results as users increase the knowledge placed into the system with the overall burden. Our hypothesis from 
-working with many LCF users is that once they get over the initial burden of placing this knowledge the burden decreases as more 
-knowledge is placed into the system, and thus, they will be more inclined to increase the knowledge into the system as the benefits
-increase.
+We will also test to see the burden of users to place more knowledge into
+the system compared to the overall effect of more optimal performance they
+will get. We realize that initially it will be a large burden for users to
+place initial knowledge into the system, but we will evaluate the results as
+users increase the knowledge placed into the system with the overall burden.
+Our hypothesis from working with many LCF users is that once they get over
+the initial burden of placing this knowledge the burden decreases as more
+knowledge is placed into the system, and thus, they will be more inclined to
+increase the knowledge into the system as the benefits increase.
 
-We will also need to evaluate how more knowledge into the system will lead into more performance they get from placing
-the most important pieces at the higher layers of the hierarchy. We will evaluate this by increasing and decreasing this knowledge.
-For example, if the user doesn't want their data broken into priority chunks, then the data will be at the mercy of what the system
-does with no knowledge. If the user then gives us information that allows us to use ``generic'' rouitnes based on frequency, or 
-accuracy, then we can use that knowledge to decompose the objects into chunks which can be placed. We can evaluate these
-against users who will give us more knowledge which allows us to perform application level binning of data (for example for
-C.S. Chang), or for level of features (for J. Tromp).
+We will also need to evaluate how more knowledge into the system will lead
+into more performance they get from placing the most important pieces at the
+higher layers of the hierarchy. We will evaluate this by increasing and
+decreasing this knowledge. For example, if the user doesn't want their data
+broken into priority chunks, then the data will be at the mercy of what the
+system does with no knowledge. If the user then gives us information that
+allows us to use ``generic'' rouitnes based on frequency, or accuracy, then
+we can use that knowledge to decompose the objects into chunks which can be
+placed. We can evaluate these against users who will give us more knowledge
+which allows us to perform application level binning of data (for example
+for C.S. Chang), or for level of features (for J. Tromp).
 
-How will we quantify the quality of data?
-We will look at errors after analytics.
+How will we quantify the quality of data? We will look at errors after
+analytics.
 
-How well does the system utilize all of its resources. We feel the more knowledge means the better the system can utilize all of the resources.
+How well does the system utilize all of its resources. We feel the more
+knowledge means the better the system can utilize all of the resources.
 
 predictability ... are teh errors bounded,
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "../proposal"
+%%% End:

### 2015-07-06 narrative/naming-service.tex
restructured naming and discovery to fit the standard pattern.

@@
 
+\paragraph{Background:}
 Given the well documented difficulties with POSIX-compliant naming and our
@@
 
-Our research in this area will address the following questions:
-(1) Since our metadata is distributed and partitioned, how do we respond to user
-  requests within a specified time bound with accuracy? How do we estimate the
-  completeness of the response?
-(2) For POSIX API requests, how do we respond if only varying data fidelity levels
-  are available for the requested data object?
-(3) What level of metadata should we maintain and keep in sync to enhance response
-  quality within time-bound requests?
-(4) How much time should be allowed for searching for metadata given the user
-  specified limits for data retrieval since the user will still have to spend
-  the data retrieval time from different devices? How do we integrate the read
-  time overhead into the metadata requests?
-(5) How do we provide appropriate security while maintaining scalability,
-  particularly with the distributed metadata stores?
-% Which data block and what quality/type do we respond to a query with?;\\
-% How will we scale a fully distributed storage system's metadata management;\\
-% Can we accelerate the discovery process with the addition of an
-%    application/user level metadata store that tracks where data was
-%    originally written and how was it structured?;\\
-% Can we provide time estimates for read/write calls in the presence of
-%    an unbounded discovery system? If not how do we bound the discovery time?
-% How do we amortize the time cost of discovery vs time cost of actually
-%    reading the data (latency vs bandwidth essentially)?
+Hierarchical Storage Management (HSM) systems~\cite{blaze:1992:hsm} offer a
+strict caching approach for managing different storage capacities trading off
+performance for capacity.  By maintaining a single namespace across all tiers,
+it is possible to list a single directory view with files stored at different
+tiers. While this approach to managing multi-tier storage works for whole files
+that fit on single logial devices, it is far from ideal for scientific
+simulations.
 
-Hierarchical Storage Management (HSM) systems~\cite{blaze:1992:hsm} offer a strict caching
-approach for managing different storage capacities trading off performance for
-capacity.  By maintaining a single namespace across all tiers, it is possible
-to list a single directory view with files stored at different tiers. While
-this approach to managing multi-tier storage works for whole files that fit on
-single logial devices, it is far from ideal for scientific simulations.
+The overriding theme for this proposal of providing a cooperative relationship
+between the user and the storage system extends into this area as well.
+Negotiating the data quality/retrieval time tradeoff reuqires additional
+metadata.  Providing this functionality requries the naming service be able to
+both return a list of data versions and an indication of data quality. This
+will be combined with a retrieval time estimate for each verstion by the
+middleware to determine what data version should be retrieved.  This estimate
+will be disucssed elsewhere in this proposal.
 
+In addition to the basic naming and data tracking operations, we will also need
+to incorporate authorization capabilities. Sirocco currently integrates with a
+Kerberos service for authentication. Given a capability ticket, a user can
+access different objects as needed. This ticket structure offers protection
+services typically offered on POSIX directories and files, but can do it at the
+fork level instead. This allows a reduced quality data version to be available
+for the general users while the high-quality version would be limited to the
+data creator. This and other considerations for security must be incorporated
+into the entire naming and metadata service.
+
+While strongly encouraging data to stay in a particular type of storage tier,
+ultimately it must be possible to flush this data to make room for other uses.
+Sirocco offers a ``flush'' operation that forcibly migrates data to the
+long-term resilience requirements freeing space in the scare resources. In
+many, if not most cases, this will move data towards or actually on to a device
+like tape intended for long-term storage.
+
+\paragraph{Approach:}
 Our goal with this proposal is to offer a similar capability but with a finer
@@
 finding data are critical for offering the quality of service guarantees we
-wish to offer. Some potential approaches can build on CRUSH~\cite{weil:ceph} from
-Ceph to offer a map of where to search for data sequentially. Given our
+wish to offer. Some potential approaches can build on CRUSH~\cite{weil:ceph}
+from Ceph to offer a map of where to search for data sequentially. Given our
 multiple storage tiers, we would extend this using some mechanism to shift to
@@
 
-The overriding theme for this proposal of providing a cooperative relationship
-between the user and the storage system extends into this area as well.
-Negotiating the data quality/retrieval time tradeoff reuqires additional
-metadata.  Providing this functionality requries the naming service be able to
-both return a list of data versions and an indication of data quality. This
-will be combined with a retrieval time estimate for each verstion by the
-middleware to determine what data version should be retrieved.  This estimate
-will be disucssed elsewhere in this proposal.
-
-Structurally, the traditional POSIX naming service offeriing a hierarchical
-space consisting of directories and files may be maintained for backwards
-compatibility. However, this view will not offer the same strict semantics
-POSIX defines. We must break these strict semantics to address scalability
-problems forced by the serialized access to a single source for creating and
-accessing files.  Several efforts~\cite{patil:2007:giga+,carns:pvfs} have worked to
-reduce this contention by doing things like reducing the serialized scope to a
-single directory or subtree or allowing a single process in a collective file
-operation talk with the metadata service and distribute a handle to other
-participants. While these approaches help, they do not address this key
-scalability limitation of serialized access--even if the metadata service is
-spread across multiple nodes. Instead, we will offer a short duration to
-consistency POSIX view to offer the performance and consistency requirements an
-exascale application demands.
-
-While pure object stores, such as those popular in the big data
-domain~\cite{Fitzpatrick:2004:memcached,others} avoid this bottleneck by strictly offering an
-object ID with the application required to manage how this ID maps to something
-of interest. This approach of removing the metadata service from the system
-level completely can work well for scale out applications where data is created
-or consumed by a single process at a time or just in a work queue rather than
-potentially O(1 million) processes all actively collectively for a single
-``object''. To address this case, having some system integrated metadata
-services to associate names with these object is a preferable solution. While
-it will not have the same synchronous consistency semantics as POSIX, we will
-offer something as close as possible to address legacy application and
-command-line style data access and migration to other storage types using these
-POSIX semantics.
-
-We plan to address this scalability problem by continuing with the LWFS and
-Sirocco model. LWFS and Sirocco have take an approach similar to the pure
-object stores, but with a focus on the HPC setting. They have abandonded a
-fully POSIX compliant metadata service as the default model in favor of a
+We plan to address scalability problems by continuing with the LWFS and Sirocco
+model. LWFS and Sirocco have take an approach similar to the pure object
+stores, but with a focus on the HPC setting. They have abandonded a fully POSIX
+compliant metadata service as the default model in favor of a
 container/object/fork/addrress tuple for identifying data similar to those used
@@
 namespace on the side kept in sync using a transaction process like
-D2T~\cite{lofstead:2012:txn} showing that this alternative approach can support traditional
-POSIX API calls even though the underlying storage system uses a different
-model. Because we are not requiring the entire storage space be addressible
-from a single tree root, we can offer multiple independent roots using the
-Sirocco object storage. We will investigate how to make the naming and
-discovery service scalable under the circumstances outlined above.
+D2T~\cite{lofstead:2012:txn} showing that this alternative approach can support
+traditional POSIX API calls even though the underlying storage system uses a
+different model. Because we are not requiring the entire storage space be
+addressible from a single tree root, we can offer multiple independent roots
+using the Sirocco object storage. We will investigate how to make the naming
+and discovery service scalable under the circumstances outlined above.
 
-In addition to the basic naming and data tracking operations, we will also need
-to incorporate authorization capabilities. Sirocco currently integrates with a
-Kerberos service for authentication. Given a capability ticket, a user can
-access different objects as needed. This ticket structure offers protection
-services typically offered on POSIX directories and files, but can do it at the
-fork level instead. This allows a reduced quality data version to be available
-for the general users while the high-quality version would be limited to the
-data creator. This and other considerations for security must be incorporated
-into the entire naming and metadata service.
+\paragraph{Related Work:}
+Structurally, the traditional POSIX naming service offeriing a hierarchical
+space consisting of directories and files may be maintained for backwards
+compatibility. However, this view will not offer the same strict semantics
+POSIX defines. We must break these strict semantics to address scalability
+problems forced by the serialized access to a single source for creating and
+accessing files.  Several efforts~\cite{patil:2007:giga+,carns:pvfs} have
+worked to reduce this contention by doing things like reducing the serialized
+scope to a single directory or subtree or allowing a single process in a
+collective file operation talk with the metadata service and distribute a
+handle to other participants. While these approaches help, they do not address
+this key scalability limitation of serialized access--even if the metadata
+service is spread across multiple nodes. Instead, we will offer a short
+duration to consistency POSIX view to offer the performance and consistency
+requirements an exascale application demands.
 
-While strongly encouraging data to stay in a particular type of storage tier,
-ultimately it must be possible to flush this data to make room for other uses.
-Sirocco offers a ``flush'' operation that forcibly migrates data to the
-long-term resilience requirements freeing space in the scare resources. In
-many, if not most cases, this will move data towards or actually on to a device
-like tape intended for long-term storage.
+While pure object stores, such as those popular in the big data
+domain~\cite{Fitzpatrick:2004:memcached,others} avoid this bottleneck by
+strictly offering an object ID with the application required to manage how this
+ID maps to something of interest. This approach of removing the metadata
+service from the system level completely can work well for scale out
+applications where data is created or consumed by a single process at a time or
+just in a work queue rather than potentially O(1 million) processes all
+actively collectively for a single ``object''. To address this case, having
+some system integrated metadata services to associate names with these object
+is a preferable solution. While it will not have the same synchronous
+consistency semantics as POSIX, we will offer something as close as possible to
+address legacy application and command-line style data access and migration to
+other storage types using these POSIX semantics.
 
+\paragraph{Challenge:}
 The main challenge of incorporating the additional, rich metadata will be
@@
 functionality.
+
+Our research in this area will address the following questions:
+(1) Since our metadata is distributed and partitioned, how do we respond to user
+  requests within a specified time bound with accuracy? How do we estimate the
+  completeness of the response?
+(2) For POSIX API requests, how do we respond if only varying data fidelity
+  levels are available for the requested data object?
+(3) What level of metadata should we maintain and keep in sync to enhance
+  response quality within time-bound requests?
+(4) How much time should be allowed for searching for metadata given the user
+  specified limits for data retrieval since the user will still have to spend
+  the data retrieval time from different devices? How do we integrate the read
+  time overhead into the metadata requests?
+(5) How do we provide appropriate security while maintaining scalability,
+  particularly with the distributed metadata stores?
 

### 2015-07-06 narrative/evaluation.tex
Tweaked evaluation section

@@
 
-In order to determine if our guiding principles are being met, and we can optimize the system from these guiding principles, 
-we will evaluate our work using the Table~\ref{table:eval}.
+In order to determine if our guiding principles are being met, and we can
+optimize the system from these guiding principles, we will evaluate our work
+using the Table~\ref{table:eval}.
 
+To determine the effectiveness of the components of our proposed work we will
+partition the problem along the lines of our two guiding principles. 
+We will evaluate how the knowledge centric system design we employ can
+provide a more predictable level of performance and data quality.
+
+%I dont think this table is right. Knowledge is our independent variable
+%and predictability is our dependent variable
 \begin{tabular}{ | l | l | l | }
@@
 (https://www.olcf.ornl.gov/computing-resources/sith/) has its own Lustre
-storage system, SSDs on each node, and is connected to the centerwide Lustre
-file system. We will perform test in both the system level (on sith) and in
+storage system, SSDs on each node, and is connected to the center-wide Lustre
+file system. We will perform testing in both the system level (on sith) and in
 the user level (on titan), and evaluate our results using our application
@@
 
-Part of our evaluation will include how accurate we can set the bounds on
+Our evaluation will include how accurately we can set the bounds on
 time estimates for applications. We will measure these estimates when data
-is on just one level of the storage hierarchy, along with the time estimates
-when data chunks are split. We will evaluate how fast data can be discovered
-when it is split between the hierarchy, allowing us to better understand how
-many pieces data can be split into before it takes too long to access most
-of the data. We will also evaluate our techniques to examine how well we can
+is on just one level of the storage hierarchy, as well as
+accesses that include data chunks on multiple levels.
+We will evaluate how fast data can be discovered
+when it is split between levels, allowing us to better understand how
+finely data can be split before it takes too long to access.
+We will also evaluate our techniques to examine how well we can
 migrate data and update our server to determine if data might get
@@
 this system works on current systems and we will need to investigate this on
-the next generation OLCF system (Summitt) which we will have early access,
-along with all of the applications we will be investigating.
+the next generation OLCF system such as Summit, to which we will have early access.
 
-Another area of investigation will be in how accurate data can be read in
+Another area of investigation will be in how accurately data can be read 
 when there are time constraints due to users wanting to read in many time
 slices in a given amount of time. For example, in J. Tromp's visualization,
-we may contain hundreds of time slices, but when we visualize this data we
-would like very fast access to the ``first look' of the data. The question
-is whether the data will be highly representative of the overall data. Can
+we may be interested in looking at hundreds of time slices, but when we visualize this data we
+would like very fast access to the ``first look'' of the data. The question
+is whether the high priority data will be sufficiently representative of the overall data. Can
 we make accurate visualization from the less accurate data? We will work
-with our application partners to access this, and we will determine the
+with our application partners to assess this, and we will determine the
 accuracy in time-constrained test vs. test which will not place any bounds
@@
 
-We will also want to access how well our data migration and eviction
+We will also want to assess how well our data migration and eviction
 policies work with user-knowledge. Will the overall system be able to
@@
 better system resource utilization. If we have more knowledge we would like
-to be able to illustrate that this will translate into higher levels of
+to be able to show that this will translate into higher levels of
 system utilization.
 
-We will also test to see the burden of users to place more knowledge into
-the system compared to the overall effect of more optimal performance they
-will get. We realize that initially it will be a large burden for users to
+One of the risks we face in our design is the unwillingness of users to add
+more knowledge to the system. To better understand the burden on the user we
+will attempt to quantify the additional burden on the users in relation to
+the additional knowledge required for optimization. 
+ We realize that initially it will be a large burden for users to
 place initial knowledge into the system, but we will evaluate the results as
@@
 
-We will also need to evaluate how more knowledge into the system will lead
-into more performance they get from placing the most important pieces at the
+We will also need to evaluate to what extent more knowledge about
+the data will lead to increased performance achieved by placing the most important pieces at the
 higher layers of the hierarchy. We will evaluate this by increasing and
-decreasing this knowledge. For example, if the user doesn't want their data
+decreasing this knowledge, and comparing the effects. For example, if the user doesn't want their data
 broken into priority chunks, then the data will be at the mercy of what the
 system does with no knowledge. If the user then gives us information that
-allows us to use ``generic'' rouitnes based on frequency, or accuracy, then
+allows us to use ``generic'' routines based on frequency, or accuracy, then
 we can use that knowledge to decompose the objects into chunks which can be
-placed. We can evaluate these against users who will give us more knowledge
-which allows us to perform application level binning of data (for example
+placed. We can evaluate these questions by working with specific users who
+are in a position give us more knowledge,
+allowing us to perform application level binning of data (for example
 for C.S. Chang), or for level of features (for J. Tromp).
 
-How will we quantify the quality of data? We will look at errors after
-analytics.
+% How will we quantify the quality of data? We will look at errors after
+% analytics.
 
-How well does the system utilize all of its resources. We feel the more
-knowledge means the better the system can utilize all of the resources.
+% How well does the system utilize all of its resources. We feel the more
+% knowledge means the better the system can utilize all of the resources.
 
-predictability ... are teh errors bounded,
+%predictability ... are the errors bounded,
 

### 2015-07-06 narrative/data-description.tex
Adjusting size of figure 1.1
Tweaked figure caption
Adjusted size of figure 1.1

@@
         \begin{centering} 
-	\includegraphics[scale=0.77]{graphics/SSIO-bucket.pdf}
+	\includegraphics[scale=0.75]{graphics/SSIO-bucket.pdf}
         %\vspace{1ex}
-        \caption{Data is given as a set of chunks which includes utility which contains the lifetime in a storage layer, and the priority of the layer.}
+        \caption{The system manages the data across the various stages of the data lifecycle}
         \label{fig:ssio-bucket}
@@
 
-We will design new techniques that users can use as part of a data mover, or
-the system can automatically use, to raise or lower these sub-chunks up and
+We will design new techniques, available to both users and the storage system,
+to raise or lower these sub-chunks up and
 down the storage stack. The SSIO layer will then manage the system metadata so
 that the user can transparently access any of these sub-chunks without
-knowledge of the layer. This requires that when reading the data, the APIs will
-then get a time estimate of reading, and then allow the users to determine if
-they will read less (only the highest priority sub-chunks) if the time estimate
+the need to know where the data currently resides.
+This requires that when reading the data, the middleware will
+provide a time estimate of reading, and then allow the users to decide if
+they will read less (e.g., only the highest priority sub-chunks)
+if the time estimate
 for reading exceeds their expectations. Existing Sirocco functionality may be
@@
 Users will also need to describe information about the data in order to choose
-the ``most optimal'' data-refactoring routine. Currently, we understand that
+the optimal data-refactoring routine. Currently, we understand that
 users can tell us that the data represents: 1) phase-space (e.g., the particles
 in our example), 2) spatial-changes (e.g., the mesh), 3) space-time (e.g.,
-the fields on the mesh). We need this information to get the best possible data
-refactoring methods. Although users can request a user-defined data-refactoring
-method, system-level methods can use this information to make better decisions.
-We will investigate new techniques and new semantics to take this information
-and use this to best prioritize and re-organize information.
+the fields on the mesh). We need this information to select
+the best possible data
+refactoring methods. Although users can insert a user-defined data-refactoring
+method, system-level methods can also use this information to make better decisions.
+We will investigate new techniques and new semantics capable of using
+this information to best prioritize and re-organize data.
 
-Finally we will investigate additional information on top of the data model.
+Finally we will investigate the use of additional information
+on top of the data model.
 This information will be used to indicate data associations that exceed
@@
 group data together making it explicit that M1, F1, and G1 need to be organized
-together.  In otherwords, in will not be helpful if data from M1 and M2 were
+together.  In other words, it will not be helpful if data from M1 and M2 were
 organized together along with only F1. When the users wants to visualize data
@@
 We seek to provide dynamic runtime information to the user and make the
-system more transparent. The end user could analyze or visualize data in a more
-predictable way since they have confidence how much longer they have to wait
-for the data.
+system more transparent. This would allow the end user to analyze or
+visualize data in a more
+predictable way since they would have more confidence of how much time
+the data retrieval process would take.
 
@@
 the storage layer will be accepted by these applications and eventually by the
-rest of the community. This requires careful examination of all possible
-design choices to ensure stable APIs and semantics that do not burden the users
-with confusing APIS and semantics that cannot even be used by the system.
+rest of the community. This requires careful examination of the full range
+of possible
+design choices to ensure stable APIs and useful semantics.

### 2015-07-06 narrative/evaluation.tex

--- 2015-07-06 narrative/evaluation.tex
@@
 
-Part of our evaluation will include how accurate we can set the bounds on
+Our evaluation will include how accurately we can set the bounds on
 time estimates for applications. We will measure these estimates when data
@@
 
-We will also test to see the burden of users to place more knowledge into
-the system compared to the overall effect of more optimal performance they
-will get. We realize that initially it will be a large burden for users to
+One of the risks we face in our design is the unwillingness of users to add
+more knowledge to the system. To better understand the burden on the user we
+will attempt to quantify the additional burden on the users in relation to
+the additional knowledge required for optimization. 
+ We realize that initially it will be a large burden for users to
 place initial knowledge into the system, but we will evaluate the results as
@@
 
-How will we quantify the quality of data? We will look at errors after
-analytics.
+% How will we quantify the quality of data? We will look at errors after
+% analytics.
 
-How well does the system utilize all of its resources. We feel the more
-knowledge means the better the system can utilize all of the resources.
+% How well does the system utilize all of its resources. We feel the more
+% knowledge means the better the system can utilize all of the resources.
 
-predictability ... are teh errors bounded,
+%predictability ... are the errors bounded,
 

### 2015-07-06 narrative/evaluation.tex
significant readability improvements and hopefully additional clarity

--- 2015-07-06 narrative/evaluation.tex
@@
 
-In order to determine if our guiding principles are being met, and we can
-optimize the system from these guiding principles, we will evaluate our work
-using the Table~\ref{table:eval}.
-
-To determine the effectiveness of the components of our proposed work we will
-partition the problem along the lines of our two guiding principles. 
-We will evaluate how the knowledge centric system design we employ can
-provide a more predictable level of performance and data quality.
+Our guiding principles are the basis for the evaluation. These principles are
+described in Table~\ref{table:eval}.  We will partition the problem along these
+lines and evaluate how the knowledge centric system design we employ can
+provide more predictable performance and data quality.
 
@@
 %and predictability is our dependent variable
-\begin{tabular}{ | l | l | l | }
-
+\begin{table}[ht]
+\caption{Guiding Principles}
+\label{table:eval}
+\centering
+\begin{tabular}{ | p{1.4in} | p{2in} | p{2in} | }
   \hline
   {\bf Principle/Metric}                   & {\bf User perspective metric}   & {\bf System perspective metric }\\ \hline
-  {\bf Knowledge-centric}                  & More knowledge = faster access  &  More knowledge = better \\
-  {\bf system}                             & to higher quality data          &  resource utilization across \\
-                                           &                                 & the entire hierarchy \\ \hline
-
-  {\bf Predictable Performance} & More predictability =                     &  Accuracy of the time bounds \\
-  {\bf Accuracy}                & better quality of data                    &  across multiple layers of the\\
-                                           &                                           &  storage hierarchy wrote/read \\ 
-                                           &                                           &  in time constrained\\
-                                           &                                           &  situations \\ \hline
-
+  {\bf Knowledge-centric system}           & More knowledge = faster access to higher quality data &  More knowledge = better resource utilization across the entire hierarchy \\ \hline
+  {\bf Predictable Performance Accuracy}   & More predictability = better quality of data &  Accuracy of the time bounds across multiple layers of the storage hierarchy wrote/read in time constrained situations\\ \hline
 \end{tabular}
+\end{table}
 
@@
 PPPL (C. S. Chang), and U. C. Irvine (Z. Lin), along with the OLCF and SNL,
-to evaluate the leading edge applications along with the system performance.
-We will use the systems at both SNL and OLCF, which have multiple layers of
-storage. For example, the sith cluster at the OLCF
+to evaluate the leading edge applications working with the storage system to
+determine performance predictability and data quality.  Our evaluation will
+will use the multiple storage layers available on systems at both SNL and the
+OLCF.  For example, the sith cluster at the OLCF
 (https://www.olcf.ornl.gov/computing-resources/sith/) has its own Lustre
 storage system, SSDs on each node, and is connected to the center-wide Lustre
-file system. We will perform testing in both the system level (on sith) and in
+file system. We will perform test in both the system level (on sith) and in
 the user level (on titan), and evaluate our results using our application
@@
 
-Our evaluation will include how accurately we can set the bounds on
-time estimates for applications. We will measure these estimates when data
-is on just one level of the storage hierarchy, as well as
-accesses that include data chunks on multiple levels.
-We will evaluate how fast data can be discovered
-when it is split between levels, allowing us to better understand how
-finely data can be split before it takes too long to access.
-We will also evaluate our techniques to examine how well we can
-migrate data and update our server to determine if data might get
-out-of-sync between data chunks. We will need to clearly understand how well
-this system works on current systems and we will need to investigate this on
-the next generation OLCF system such as Summit, to which we will have early access.
+The first part of our evaluation will cover the following characteristics: (1)
+the accuracy of time bounds estimates for applications when data is on just one
+level of the storage hierarchy as well as when the data chunks are split; (2)
+data discovery speed when chunks are spread across the hierarchy informing how
+many pieces data can be split into before it takes too long to access most of
+the data; and (3) how well we can migrate data and update our server to
+determine if data might get out-of-sync between data chunks.  We will need to
+clearly understand how well this system works on current systems and we will
+need to investigate this on the next generation OLCF system (Summitt) and the
+SNL system (Trinity) which we will have early access, along with all of the
+applications we will be investigating.
 
-Another area of investigation will be in how accurately data can be read 
-when there are time constraints due to users wanting to read in many time
-slices in a given amount of time. For example, in J. Tromp's visualization,
-we may be interested in looking at hundreds of time slices, but when we visualize this data we
-would like very fast access to the ``first look'' of the data. The question
-is whether the high priority data will be sufficiently representative of the overall data. Can
-we make accurate visualization from the less accurate data? We will work
-with our application partners to assess this, and we will determine the
-accuracy in time-constrained test vs. test which will not place any bounds
-on the time to retrieve the data.
+The second part of our evaluation will cover data accuracy for data read when
+there are time constraints due to users wanting to read in many time slices in
+a given amount of time. For example, J. Tromp's visualization may contain
+hundreds of time slices, but when we visualize this data, we would like very
+fast access to the data ``first look''. We will determine if the data will be
+sufficiently representative of the overall data.  We will work with our
+application partners to assess the visualization accuracy using different data
+qualities including comparisons between a time-constrained test vs. a test
+without any time constraints for data retrieval.
 
-We will also want to assess how well our data migration and eviction
-policies work with user-knowledge. Will the overall system be able to
-utilize more of the storage hierarchy by understanding these intentions and
-creating autonomic rules to self-optimize? Furthermore we will want to see
-if users can give us more accurate policies which can then be used for
-better system resource utilization. If we have more knowledge we would like
-to be able to show that this will translate into higher levels of
-system utilization.
+The third part of the evaluation will assess how well our data migration and
+eviction policies work with user-knowledge. We will determine if the overall
+system is able to use more of the storage hierarchy by understanding these
+intentions and creating autonomic rules to self-optimize. Furthermore, we will
+determine if users can give more accurate policies that lead to better system
+resource utilization. We specifically will evaluate if more knowledge
+translates into higher levels of system utilization.
 
-One of the risks we face in our design is the unwillingness of users to add
-more knowledge to the system. To better understand the burden on the user we
-will attempt to quantify the additional burden on the users in relation to
-the additional knowledge required for optimization. 
- We realize that initially it will be a large burden for users to
-place initial knowledge into the system, but we will evaluate the results as
-users increase the knowledge placed into the system with the overall burden.
-Our hypothesis from working with many LCF users is that once they get over
-the initial burden of placing this knowledge the burden decreases as more
-knowledge is placed into the system, and thus, they will be more inclined to
-increase the knowledge into the system as the benefits increase.
+The fourth part of the evaluation focuses on the additional user burden. One of
+the risks in the proposed design is the unwillingness of users to add more
+knowledge to the system. To better understand the user burden, we will attempt
+to quantify the additional requirements to improve the automatic optimization.
+We expect a substantial minimal burden to inform the automatic optimization.
+Our evaluation will focus on the improvements gained through additional user
+provided data and attempt to relate the pain/benefit tradeoff. The initial
+hypothesis from working with many LCF users suggests that once the initial
+burden to facilitate the automatic optimization, the additional burden required
+to increase benefits is small enough to encourage users to participate more
+heavily for greater benefit.
 
-We will also need to evaluate to what extent more knowledge about
-the data will lead to increased performance achieved by placing the most important pieces at the
-higher layers of the hierarchy. We will evaluate this by increasing and
-decreasing this knowledge, and comparing the effects. For example, if the user doesn't want their data
-broken into priority chunks, then the data will be at the mercy of what the
-system does with no knowledge. If the user then gives us information that
-allows us to use ``generic'' routines based on frequency, or accuracy, then
-we can use that knowledge to decompose the objects into chunks which can be
-placed. We can evaluate these questions by working with specific users who
-are in a position give us more knowledge,
-allowing us to perform application level binning of data (for example
-for C.S. Chang), or for level of features (for J. Tromp).
+The fifth part of the evalation will evaluate how well additional data
+knowledge in the system will lead to more performance by placing the most
+important pieces at the higher layers of the hierarchy. We will test different
+knowledge levels and evaluate the system impact.  For example, if the user
+doesn't want the data broken into priority chunks, then the data will be at
+the mercy of what the system does when there is no knowledge. If the user then 
+provides sufficient information to use ``generic'' rouitnes based on frequency
+or accuracy, then that knowledge will be used to decompose the objects into
+specially placed chunks. We will evaluate these no and minimal knowledge cases
+against those that provide additional knowledge allowing application level data
+binning (for C.S. Chang) or for level of features (for J. Tromp).
 

### 2015-07-07 other-support/lofstead-cp.tex
fixed C and P

@@
 
-%\begin{tabbing}
-%Project Title \hspace{5.5em}\=: \= Code Coupling for Scalability and Robust Data Analytics\\
-%Source of Support           \>: \> Sandia National Laboratories LDRD \\
-%Award Amount                \>: \> \$600K/year to SNL\\
-%Period Covered              \>: \> 8/2014 -- 7/2017\\
-%Investigator Commitment     \>: \> 0.5 FTE
-%\end{tabbing}
+\begin{tabbing}
+Project Title \hspace{5.5em}\=: \= SCADS: Scalable Deep Storage Hierarchy for Exascale Computing\\
+Source of Support           \>: \> DOE Storage Systems and Input/Output for Extreme Scale Science\\
+Award Amount                \>: \> \$176/year to SNL\\
+Period Covered              \>: \> 10/2015 -- 9/2018\\
+Investigator Commitment     \>: \> 0.25 FTE
+\end{tabbing}
 

### 2015-07-07 narrative/qos.tex
split into two subsections for QoS.

@@
 
-\paragraph{Resource Management Background:} There is an important connection between finding the resources
-that can help fulfill a particular request and the performance predictability of that request, and this demands
-careful resource management strategies that deal with large number of data chunks spread across storage hierarchy. 
+\subsubsection{Resource Management}
+\paragraph{Background:} There is an important connection between finding the resources
+that can fulfill a particular request and the performance predictability of that request, and this demands
+careful resource management strategies that can deal with placing and retrieving a large number of data 
+chunks spread across the entire storage hierarchy. 
 %This connection is particularly important for data retrieval 
@@
 Locating resources is a typical distributed systems problem. Centralized solutions are limited in their
-scalability, because there exists a single entity that tracks updates in the system. 
-Traditional file systems are good examples for this: most updates to files require updates 
+scalability, because they require a single entity that tracks updates in the system. 
+Traditional file systems are good examples for this - most updates to files require updates 
 to their centrally maintained block allocation tables. 
@@
 resources at request time. Pile systems allow very efficient writes because clients can write whereever
-there is free space and there is very low bookkeeping overhead. However, unbounded search is possible, 
-if it is unclear whether the resource actually exists. 
+there is free space and there is very low bookkeeping overhead. However, unbounded search in pile systems
+is possible, e.g., when it is unclear whether the resource actually exists. 
 Methods to make searches more efficient without reverting to centralized bookkeeping usually involves 
-some kind of partitioning of the search space and a method to limit the search to a small number of partitions.
+partitioning of the search space and a method to limit the search to a small number of partitions.
 
-Frequently this involves a hash function that is relatively stable against changes to its range, such as failures or updates to the hardware of the system (see for example consistent hashing~\cite{karger:stoc97}).
-Consistent hashing is scalable since it can be well-known throughout a system of any size and only requires minimal movement of data in the presence of failures or system changes.
-Consistent hash functions can also have sets of ordered but infinite lists as their range (which works well with primary replication)~\cite{honicky:ipdps04}, implement declustering (limiting overlap between partitions), and support placement policy rules~\cite{weil:sc06}.
+Frequently this involves a hash function that is relatively stable against changes to its range, 
+such as failures or updates to the hardware of the system. Consistent hashing~\cite{karger:stoc97}
+is scalable since it can be well-known throughout a system of any size and only requires minimal 
+movement of data in the presence of failures or system changes. 
+%Consistent hash functions can also
+%have sets of ordered but infinite lists as their range (which works well with primary replication)
+%~\cite{honicky:ipdps04}, implement declustering (limiting overlap between partitions), and support
+%placement policy rules~\cite{weil:sc06}.
 
-In this proposal we will investigate the trade-off in terms of predictable performance between ``filing'' data, making an upfront investment to ensure finding data quickly, and ``piling'' data, pushing the cost of finding to the moment when the data is actually needed.
-This trade-off has a parallel in data management where there are two alternative paths to achieve knowledge: the analogue to filing is ``extract, transform, and load'' (ETL) data into a warehouse, paying an upfront ingest cost to achieve fast querying; the analogue to piling is to not pay any upfront cost and query the raw data directly by using data processing platforms such as Hadoop.
-Recent work has found that the combination of these two techniques can be optimized to perform better than any one technique alone~\cite{lefevre:sigmod14a}.
-We believe that a similar opportunity exists for combining filing and piling approaches to achieve performance predictability without sacrificing too much flexibility.
+\paragraph{Approach:} 
+In this proposal we will investigate the trade-off in terms of predictable performance between 
+``filing'' data, making an upfront investment to ensure finding data quickly, and ``piling'' 
+data, pushing the cost of finding to the moment when the data is actually needed. This trade-off 
+has a parallel in data management where there are two alternative paths to achieve knowledge: 
+the analogue to filing is ``extract, transform, and load'' (ETL) data into a warehouse, paying 
+an upfront ingest cost to achieve fast querying; the analogue to piling is to not pay any upfront 
+cost and query the raw data directly by using data processing platforms such as Hadoop. Recent work 
+has found that the combination of these two techniques can be optimized to perform better than 
+any one technique alone~\cite{lefevre:sigmod14a}. We believe that a similar opportunity exists 
+for combining filing and piling approaches to achieve performance predictability without sacrificing 
+too much flexibility.
 
-More specifically, the coarseness of a search space partitioning limits the search time within a partition, i.e.
-finely partitioned search spaces require little or no search within a partition but offer little placement flexibility.
-Partitioning can also be fine in one dimension and coarse in another to allow flexibility where it is most needed while still ensuring fast finding.
-For example, a hash function can finely partition a low storage tier but each partition can extend across multiple tiers so that a search still has to occur in order to find a resource on a particular tier.
+More specifically, the coarseness of a search space partitioning limits the search time within 
+a partition, i.e. finely partitioned search spaces require little or no search within a partition 
+but offer little placement flexibility. Partitioning can also be fine in one dimension and coarse 
+in another to allow flexibility where it is most needed while still ensuring fast finding. For 
+example, a hash function can finely partition a low storage tier but each partition can extend 
+across multiple tiers so that a search still has to occur in order to find a resource on a particular 
+tier.
 
@@
 
-\paragraph{QoS Background:} 
-Quality of Service (QoS) refers to the
+\subsubsection{QoS}
+\paragraph{Background:} Quality of Service (QoS) refers to the
 properties of the performance quality of a particular service, in

### 2015-07-07 narrative/intro.tex
new evaluation...
edits of intro

@@
 Avoiding impending bottlenecks in exascale scientific discovery will require
-new research into managing and storing the large volumes of data that are 
+new research into managing, storing and retreiving the large volumes of data that are 
 produced by simulations and analyzed for months afterwards.
@@
 In this project we will demonstrate a cooperative approach for storing data
-where the user and the storage system work together to achieve the best
-possible utility and performance given data features, storage tier
-characteristics, and current system state. We will investigate novel techniques
-to facilitate efficient mapping of data objects
-from the user space onto multiple storage tiers and enable
-application-guided data reductions and transformations to address capacity and
-bandwidth bottlenecks, while constraining errors to be within user provided
-bounds.
+where the user and the storage system work together to optimize performance with
+respect to user requirements, current system state, and characteristics of the tiered-storage system.
+Our investigation will include the novel techniques for the efficient mapping of data objects
+from the user space across a hierarchical tiered storage. These techniques will
+include support for application-guided data reductions with user-defined error bounds
+and transformations to optimize performance, capacity and minimize bandwidth bottlenecks.
+
 % 
-Our proposed research program is guided by the following overarching objectives.
-
-{\em Address emerging storage architecture/technology landscape:}
+Our proposed research program is guided by the following overarching objectives:
+\begin{itemize}
+\item
+{\em Address emerging storage architecture and technology landscape:}
 We will address the associated I/O and storage challenges in the context of
@@
 entire storage stack.
-
+\item
 {\em Develop a science driven, system aware approach:}
@@
 vastly different performance characteristics and constraints on capacity and 
-data lifetime they can support, our approach will enable researchers to 
+supported data lifetime, our approach will enable researchers to 
 place more knowledge into the SSIO system so that data can be recast from 
-a sequence of bytes to information, which the system can then place and 
-optimize the time to gain knowledge from exascale simulations.
+a sequence of bytes to information rich data. This information rich data can the
+be used by the system to placement within the tiered storage that will optimize the time
+to gain knowledge and insight from exascale simulations.
 
@@
 
+\item
 {\em Emphasize impacts on science:}
@@
 Instead of trying to optimize throughput, we will seek to reduce the time to knowledge.
-This is the most significant metric for scientific applications where the desired outcome
-is not storing data, but rather executing knowledge creation and extraction processes 
-on the data. 
-Moreover, we aim to perform the optimization not for a single
-application, but rather for workloads within a multi-user multi-application
-system environment.
+As extracting knowledge is the only metric that matters for scientific applications, we will
+focus not only upon data storage, but upon efficient execution of knowledge create and extraction
+workflows.
+Moreover, we aim to perform the optimization not only for a single
+application, but rather for multiple workloads running on a multi-user, multi-application
+system.
+\end{itemize}
 
-The overarching idea of this proposal is that data from applications can be characterized by different
-%levels of importance and different 
-levels of importance over time, which then can be used to guide the analysis and visualization later to access the most important data and deliver some level of knowledge without waiting for access to the whole original dataset.
+This proposal can be summarized as research into using the varying importance of simulation data over time to
+ensure that the right portions of data, in the right format and resolution are available to scientists when they are needed.
+Application-specific information about the importance of data will be used in conjuction with system- and resource-specific information
+to optimize the scientists experience interacting with, and extracing knowledge and insight from simulation data.
+
+%The overarching idea of this proposal is that data from applications can be characterized by different
+%%levels of importance and different 
+%levels of importance over time, which then can be used to guide the analysis and visualization later to access the most important data and deliver some level of knowledge without waiting for access to the whole original dataset.
 Information about the importance of the data at the application level, information about resource status and performance of the multi-level storage system, and flexible data reorganization and reduction operations at all levels can be utilized to provide a quality of service and allow users to gain insight in a timely manner.
@@
 This proposal combines knowledge and expertise 
-from a wide range of sources including ORNL applications, the ADIOS project,
+from a wide range of sources including applications running at ORNL, the ADIOS project,
 the SNL Lightweight File Systems project and storage system knowledge, the metadata
 management and storage systems knowledge of UCSC, and the middleware expertise
-of Rutgers. We will integrate this diverse expertise to investigate how to make
+of Rutgers. We will integrate this diverse expertise to investigate how to create
 a data, workflow, and application aware storage system.
@@
 %problematic without new insights and solutions.
-Our proposal will be responsive to Theme 2: {\em Scalable Storage Software Infrastructure} of the call, as summarize in Table~\ref{ab:call}.
+Our proposal will be responsive to Theme 2: {\em Scalable Storage Software Infrastructure} of the call, as summarize in Table~\ref{tab:call}.
+
 \begin{table}[htbp]
@@
 \end{center}
+%\caption {WRITE A CAPTION FOR THIS TABLE}
+\caption {}
 \label{tab:call}
@@
 
-\subsection{Writing Perspective}
+\subsection{Writing Perspective (is there a better name for this section????)}
 \label{subsec:sim-perspective}
 The simulation scientist ideally should be able to
-obtain an estimate of how much time will be required to write their data,
-and how much storage space they will be able to get during the lifetime of
-their simulations. Furthermore, they would like the ability to get an assurance
-of Quality of Service such that they can react appropriately when the
-expected bandwidth, for example, is less than what they desire. Ideally, the user
-would specify a set of rules with which the system can make autonomic
-decisions to determine what should be done.
+obtain an estimate of how much time will be required to write the data,
+and how much storage space is available during the simulations lifetime.
+Further, they would like Quality of Service assurances so that reactions
+to varying bandwidth can be handled appropriately. Ideally, a set of
+user-specified rules would allow the system to make autonomic decisions dynamically.
 %
 
-One of the key aspects of our approach is to allow the users to define the
-utility of their data; in other-words, the system should provide a
-mechanism such that, for each data object, a user can specify pieces of the object
-which have more utility (more importance).
-There are many research challenges faced with
-this approach including: 1) How do we make this simple enough so that users can
-do this, 2) Can we define a model to allow the system to make
-autonomic decisions to save the data which has the highest utility, and 3) Can
-we allow the back-end storage system to interact with this model so that
-it can optimize not just a single application but across all of the
-applications.
+Central to our approach is giving users the ability to define the importance, or utility of their data.
+This includes the ability to assign varying levels of importance to individual data objects in their entirety, or specific
+subsets and pieces.
+Research challenges include 1) an expressive and easy to use interface, 2) a model for the system supporting
+autonomic decision for saving data using the user-defined importance and utility of data, and 3) the interface between
+the system and the back-end storage across both single, and multiple running applications.
+
  %
   In our simple scenario, a user wants to write 1~PB of data every hour
-for checkpoint/restart, and they wish to write 500~TB of data every 15
-minutes to save data for analysis and visualization. At this stage, they will ask the system to
-write the data, and they would like to get an estimated time from which they can
-then figure out, through a series of rules, whether they want to write out all
-of the data, postpone the write until a later time, or write a reduced amount
-of data using in situ reduction techniques, along with some code
-container that allows for regenerating the data with a specified level of
-accuracy.  Furthermore, they realize that the analysis data will be saved on
-either the parallel file system, long-term tape storage, or on the campaign storage,
-which is a medium-term storage
-area with lower bandwidth than the parallel file system.
-For each piece of data, the user will specify the lifetime so
-that the data with the highest utility is kept around in fast storage for the longest
-period of time, while the data with the least utility may be the first to be migrated to
-lower layers of the storage hierarchy.
+for checkpoint/restart, and write 500~TB of data every 15
+minutes for analysis and visualization.
+The user would then specify a set of rules that governed the actual behaviour while the simulation is running.
+If the system resources are available and the estimated write times are in order, the write can proceed.
+Or, the write can be postponed to a later time-step, or data processed via in situ techniques
+or a reduced set of data along with a regenerating code container at a given accuracy can be saved.
+Data importance or utility will be used to place the various data on the proper resources
+either the parallel file system, long-term tape storage, or on the campaign storage (medium-term storage with lower bandwidth).
+The user specified lifetime for each data ensures that the data with the highest utility is preserved in fast storage
+for the longest period of time, and the least utility are migrated to lower layers in the storage hierarchy.
 

### 2015-07-07 narrative/intro.tex
edits of intro

@@
 tremendous amount of interference from other users on the system who may be
-writing or reading at the same time and the time to perform an operation is unpredictable. 
+writing or reading at the same time, leading to unpredictable performance.
 By requiring a certain quality of service the system could potentially lock out other users. 

### 2015-07-07 narrative/intro.tex
Merge branch 'master' of https://bitbucket.org/habbasi/ssio-proposal
Conflicts:
	narrative/intro orig.tex

@@
 Avoiding impending bottlenecks in exascale scientific discovery will require
-new research into managing, storing and retreiving the large volumes of data that are 
+new research into managing, storing and retrieving the large volumes of data that are 
 produced by simulations and analyzed for months afterwards.

### 2015-07-07 narrative/data-description.tex
changes to the introduction, budget and add labels.

@@
 \subsection{Data Description}
+\label{sec:data-description}
 In order for the middleware and storage layer to make decisions

### 2015-07-07 narrative/intro.tex
changes to the introduction, budget and add labels.

@@
 %
-In this project we will demonstrate a cooperative approach for storing data
+In this project we will demonstrate a cooperative approach for storing and retrieving  data
 where the user and the storage system work together to optimize performance with
 respect to user requirements, current system state, and characteristics of the tiered-storage system.
-Our investigation will include the novel techniques for the efficient mapping of data objects
-from the user space across a hierarchical tiered storage. These techniques will
+Our investigation will include  novel techniques for the efficient mapping of data objects
+from the user space across to a hierarchical tiered storage layer. These techniques will
 include support for application-guided data reductions with user-defined error bounds
@@
 Negotiation between these layers will be provided as a fundamental service, allowing users to
-easily save ``the best'' amount of data, which is automatically placed throughout the
+easily save ``the best''  data, which is automatically placed throughout the
 entire storage stack.
@@
 place more knowledge into the SSIO system so that data can be recast from 
-a sequence of bytes to information rich data. This information rich data can the
-be used by the system to placement within the tiered storage that will optimize the time
+a sequence of bytes to information rich data. This information rich data can then
+be used by the system to guide the  placement within the tiered storage and will optimize the time
 to gain knowledge and insight from exascale simulations.
@@
 As extracting knowledge is the only metric that matters for scientific applications, we will
-focus not only upon data storage, but upon efficient execution of knowledge create and extraction
+focus not only upon data storage, but upon efficient  retrieval of knowledge from their 
 workflows.
@@
 This proposal can be summarized as research into using the varying importance of simulation data over time to
-ensure that the right portions of data, in the right format and resolution are available to scientists when they are needed.
-Application-specific information about the importance of data will be used in conjuction with system- and resource-specific information
-to optimize the scientists experience interacting with, and extracing knowledge and insight from simulation data.
+ensure that the right portions of data, in the right quality  are available to scientists when they are needed.
+Application-specific information about the importance of data will be used in conjunction with system and resource-specific information
+to optimize  scientists experience in interacting with, and extracting knowledge and insight from their  data.
 
@@
 This proposal combines knowledge and expertise 
-from a wide range of sources including applications running at ORNL, the ADIOS project,
-the SNL Lightweight File Systems project and storage system knowledge, the metadata
+from a wide range of sources including many leadership class applications using the Adaptable Input/Output System (ADIOS) framework,
+the Sandia National Laboratory (SNL) Lightweight File Systems (LWFS) project and storage system knowledge, the metadata
 management and storage systems knowledge of UCSC, and the middleware expertise
 of Rutgers. We will integrate this diverse expertise to investigate how to create
-a data, workflow, and application aware storage system.
-This project brings together a team with strong expertise in I/O middleware (ORNL, Rutgers),
-file system (SNL, UCSC) and storage (UCSC), and connects and
-coordinates these key storage components in a seamless fashion.
+a data, workflow, and application aware storage and I/O system.
+%This project brings together a team with strong expertise in I/O middleware (ORNL, Rutgers),
+%file system (SNL, UCSC) and storage (UCSC), and connects and
+%coordinates these key storage components in a seamless fashion.
 
@@
 \begin{table}[htbp]
+\small
 \vspace{2ex}
 \begin{center}
-\begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{1.6cm} |}\hline
+\begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{2.2cm} |}\hline
 \multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant S2E2 work} & \multicolumn{1}{|c|}{\bf See} \\\hline\hline
-%\multicolumn{2}{|c|}{Collaboration} & 
-\multicolumn{2}{|p {5.5cm}|}{Approaches to improve the ability of SSIO software to support checkpoint/restart;} & 
-General piece we are doing for this... & 
-\S\ref{sec:?}, \S\ref{sec:??}, \\\hline
-1 & Data abstractions & What will we do & \S\ref{sec:?}, \S\ref{sec:?}, \S\ref{sec:?} \\\hline
-2 & Mapping science data models onto hierarchical storage & What we are doing & \S\ref{sec:?}, \S\ref{sec:?} \\\hline
-3 & Mechanisms for data movement across the storage hierarchy & what are we doing. & \S\ref{sec:?} \\\hline
-4 &  Extend the interaction of IO middleware with the resource management system & What we are doing & \S\ref{sec:?}, \S\ref{sec:?}  \\\hline
-5 & Design and implement of IO middleware architectures  & What we are doing & \S\ref{sec:?}, \S\ref{sec:?}, \S\ref{sec:?} \\
+1 & Data abstractions & Create new APIs  and annotations for users to guide the system for placement  & \S\ref{sec:data-description} \\\hline
+2 & Mapping science data models onto hierarchical storage &  Create the ability to re-organize, reduce and prioritize data chunks from an object& \S\ref{sec:data-refactor}, \S\ref{sec:data-description},  \S\ref{sec:managing-data-life}\\\hline
+3 & Mechanisms for data movement across the storage hierarchy & what are we doing. & \S \ref{sec:managing-data-life},  \S\ref{sec:migration} \\\hline
+4 &  Extend the interaction of IO middleware with the resource management system & We will manage data from the application  to the middleware to the storage system and allow them to have an autonomic interaction & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
+5 & Design and implement of IO middleware architectures  & We will create APIs to guide the placement of chunks of data across the SSIO and autonomically move data when necessary& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
+6 & Approaches to improve the ability of SSIO software to support checkpoint/restart & We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then purged when it is no longer needed & \S \ref{sec:???}  \\
 \hline
@@
 %\caption {WRITE A CAPTION FOR THIS TABLE}
+\vskip -0.5cm
 \caption {}
@@
 %
-\section{An Illustrative Example}
+\paragraph{An Illustrative Example}
 The challenges addressed by this proposal are illustrated by a problem our team has been working on in the past six months.
@@
 
-Take for example the latest use case for the XGC1~\cite{chang2006integrated} simulation from C.
+The latest use case is from  the XGC1~\cite{chang2006integrated} simulation from C.
 S.
-Chang (PPPL), who is one of the largest users of Leadership Class Facilities (over 300M hours at ALCF, NERSC, and OLCF).
+Chang (PPPL), who is one of the largest users of Leadership Class Facilities (over 300M hours at DOE facilities).
 They launched a series of simulations, planning to produce 100~PB of data over 10 days of total runtime on the Titan system.
-An expert team was assembled from the user group, the middleware I/O team to help with the application I/O (ADIOS), the local experts of the Lustre file-system and the tape archival system (HPSS).
+An expert team was assembled from the user group, the  I/O, and storage  team to help with ensuring that the maximum amount of information would be
+saved.
 Due to physical resource limitations, the simulation was restricted to write about 10~PB initially.
-However, once the time to write and read, together with the financial cost to archive this much data were more fully explored, this plan was again significantly reduced.
-At each stage, the particular expert knowledge regarding the lustre file system and the archival tape storage were not known to the user team.
+However, once the time to write and read, together with the financial cost to archive this much data were fully explored, this plan was significantly reduced.
+At each stage, the particular expert knowledge regarding the file system and the archival storage were not known to the user team.
 As these particulars were realized, significant modifications in the plans were made.
-Eventually, the user was forced to make very hard decisions in reducing the desired 100~PB over a 10-day run to a mere 5~PB of data.
-A maximum of 2~PB would be on disk at any time, and the rest would be archived.
+Eventually, they were forced to make very hard decisions in reducing the desired 100~PB over a 10-day run to only 5~PB of data.
 
@@
 In order to do this, they had to implement a rudimentary discovery system to determine which data were located on disk, and which were on tape archive.
-Additionally, some pieces of data needed to be moved to PPPL for immediate analysis.
 
@@
 Such a system could write more data when the resource is idle, and less data when it is very busy.
-
 However, the most significant problem occurs after the simulation run completes.
-Basic visualizations from the reduced data, just to get a high-level view of the simulation, can take significant time.
+Basic visualizations, just to get a high-level view of the simulation, can take significant time.
 While some timesteps might be on disk, others are archived on tape (where transfer could take weeks).
@@
 
-\subsection{Overview of the proposed approach (can be shortened)}
+\subsection{Overview of the proposed approach}
 
@@
 
-We will examine the research challenges from the perspective of
+We  examine the research challenges from the perspective of
 %
-(1) the user of the system who is attempting to run an exascale simulation,
+(1) the scientist who runs the exascale simulation,
 %
@@
 
-\subsection{Writing Perspective (is there a better name for this section????)}
+\paragraph{Simulation Perspective}
 \label{subsec:sim-perspective}
@@
  %
-  In our simple scenario, a user wants to write 1~PB of data every hour
-for checkpoint/restart, and write 500~TB of data every 15
-minutes for analysis and visualization.
-The user would then specify a set of rules that governed the actual behaviour while the simulation is running.
+  In our  scenario we would like the scientist would then specify a set of rules that governed the actual behavior while the simulation is running.
 If the system resources are available and the estimated write times are in order, the write can proceed.
-Or, the write can be postponed to a later time-step, or data processed via in situ techniques
-or a reduced set of data along with a regenerating code container at a given accuracy can be saved.
-Data importance or utility will be used to place the various data on the proper resources
-either the parallel file system, long-term tape storage, or on the campaign storage (medium-term storage with lower bandwidth).
+Or, the write can be postponed to a later time-step, or data reduced via in situ techniques can be saved.
+Data importance or utility will be used to place the various data on the proper SSIO layer.
 The user specified lifetime for each data ensures that the data with the highest utility is preserved in fast storage
 for the longest period of time, and the least utility are migrated to lower layers in the storage hierarchy.
+%
+%Since the data will be a collection of objects placed on the storage system,
+%some of the objects will be broken up into multiple pieces by ``plug-ins''
+%to the system which will allow data to be characterized first in terms of the most
+%important pieces, and then by subsequent levels of detail. We can think of this as
+%something similar to an Adaptive Mesh Refinement scheme which keeps track of
+%the places where there are features such as steep gradients, which contain large errors on the
+%coarser views, as opposed to other areas with smaller errors. These distinctions allow us to make decisions
+%about which sub-objects to save (all of the data, which will then go to
+%different layers) or just some of the sub-objects? The research questions
+%are numerous: How does the user specify their intentions? How does the user
+%prioritize the different sub-objects of the data? Where does the data go
+%into the storage hierarchy? Does the data get replicated in the lower tiers
+%of the hierarchy? When the user specifies the time they want the data saved,
+%how does this get into the storage system?
 
-Since the data will be a collection of objects placed on the storage system,
-some of the objects will be broken up into multiple pieces by ``plug-ins''
-to the system which will allow data to be characterized first in terms of the most
-important pieces, and then by subsequent levels of detail. We can think of this as
-something similar to an Adaptive Mesh Refinement scheme which keeps track of
-the places where there are features such as steep gradients, which contain large errors on the
-coarser views, as opposed to other areas with smaller errors. These distinctions allow us to make decisions
-about which sub-objects to save (all of the data, which will then go to
-different layers) or just some of the sub-objects? The research questions
-are numerous: How does the user specify their intentions? How does the user
-prioritize the different sub-objects of the data? Where does the data go
-into the storage hierarchy? Does the data get replicated in the lower tiers
-of the hierarchy? When the user specifies the time they want the data saved,
-how does this get into the storage system?
-
-\subsection{Storage System Perspective}
+\paragraph{Storage System Perspective}
 \label{subsec:storage-perspective}
@@
 
-One of the main research objectives of this project is to understand how the
-storage system can interact with the middleware (which is application-aware)  so
+One of our  research objectives is to understand how the
+storage system can interact with the middleware   so
 that the data with the highest utility is kept on the fastest layers of the
@@
 is also critical that the storage system can manage (discover and retrieve) the data across all of the
-layers automatically, without user intervention. We will use learning techniques to automatically migrate data
-sub-objects across the hierarchy as well as to eventually evict data which is
-large and has very low utility.
+layers automatically, without user intervention. 
+%We will use learning techniques to automatically migrate data
+%sub-objects across the hierarchy as well as to eventually evict data which is
+%large and has very low utility.
 
-\subsection{Reading  Perspective}
+\paragraph{Reading  Perspective}
 \label{subsec:reading-perspective}

### 2015-07-07 narrative/enabling_tech.tex
clean up the enabling technologies section a bit and shortened this.

--- 2015-07-06 narrative/enabling_tech.tex
@@
 will be performed.
-
-At the lowest level is the Lightweight File Systems (LWFS) project~\cite{oldfield:lwfs}.
-As mentioned above, LWFS's goal is to provide the simplest core services for
-a storage system and investigate adding functionality by using auxiliary
-services or layering an API on top that offers more functionality. For example,
-by eliminating consistency checks from the core system, applications already
-managing consistency can eliminate the redundant check enhancing performance.
-
-The current LWFS project phase is investigating new ways for managing storage
-devices. This project, Sirocco, offers a peer-to-peer style data storage device
-mesh wtih data migrating based on use, required resilience, and available
+%1
+At the lowest level is the Lightweight File Systems ({\bf LWFS}) project~\cite{oldfield:lwfs}.
+which had the  goal  to provide the simplest core services for
+a storage system and having  added functionality  using auxiliary
+services. 
+%For example,
+%by eliminating consistency checks from the core system, applications already
+%managing consistency can eliminate the redundant check enhancing performance.
+%
+%The current LWFS project phase is investigating new ways for managing storage
+%devices. This project, 
+{\bf Sirocco} cam out of the LWFS project and offers a peer-to-peer style data storage device
+mesh with data migrating based on use, required resilience, and available
 resources. For example, when writing data initially into Sirocco, required
@@
 
-The LWFS philosophy focuses on providing a universal core with additional
-semantics and services layered on top or run along side. For this project, the
-Sirocco API is much more complex than a standard POSIX call prompting layering
-a more user friendly interface on top. Placing a thin, optimized layer as part
-of the storage system will offer the kinds of functionality and semantics this
-project proposes. From the user perspective, something like the
-ADIOS~\cite{lofstead:adaptible} API offers an API nearly as simple as POSIX, but with the
+%The LWFS philosophy focuses on providing a universal core with additional
+%semantics and services layered on top or run along side. For this project, the
+%Sirocco API is much more complex than a standard POSIX call prompting layering
+%a more user friendly interface on top. Placing a thin, optimized layer as part
+%of the storage system will offer the kinds of functionality and semantics this
+%project proposes.
+
+  From the user perspective, something like the
+{\bf ADIOS}~\cite{lofstead:2009:adaptible} API offers an API nearly as simple as POSIX, but with the
 ability to change the actual data transport mechanism without changing the
 source code. This affords writing to a single output API while shifting from
-writing to a POSIX file system, some middleware like
-DataSpaces~\cite{dataspaces}, or some other system like Sirocco all without
+writing to a file system to operating on the data using in situ techniques, without
 changing the source code. ADIOS has proven effective for writing to POSIX file
-systems, data staging systems~\cite{predata,dataspaces,nessie}, workflow
-systems~\cite{flexpath}, and even nothing through the use of a NULL transport.
+systems, data staging systems~\cite{zheng:2010:predata,docan:2010:dataspaces,oldfield:2006:nessie}, workflow
+systems~\cite{Dayal:2014:flexpath}, and even through the WAN network 
 Part of this project's goals seeks to understand where the dividing line
 between the storage layer shared by all users and middleware being used by a
-single or small number of users.
-
-At a similar level as ADIOS, but can be used in conjunction, are tools like
-DataSpaces~\cite{dataspaces}. They offer a way to request data set portions
-from a one component by the next componet in a data flow. In DataSpaces case,
-the data transfer does not have to correspond to the data from a single
-process.  Instead, DataSpaces will determine where the data is and issue
-requests to pull the data from the source to the sink in an on demand process.
+single or small number of users. We will utilize DataSpaces~\cite{docan:2010:dataspaces} to 
+ further integrate the I/O (ADIOS) layer with the storage layer (Sirocco).
+%
+% 
+%They offer a way to request data set portions
+%from a one component by the next componet in a data flow. In DataSpaces case,
+%the data transfer does not have to correspond to the data from a single
+%process.  Instead, DataSpaces will determine where the data is and issue
+%requests to pull the data from the source to the sink in an on demand process.
 
 For this project, we plan to use the Sirocco data storage mechanism to manage
-the actual data storage on all of the different tiers. Our initial thought on
-where the dividing line between the storage layer and the middleware is as
-follows. We will add a data management functionality as part of the storage
+the actual data storage on all of the different tiers. We will add data management functionality as part of the storage
 system layer to handle data discovery and metadata tasks. At the middleware
@@
 which functionality works best at the different layers for usability and
-scalability.
+scalability. Furthermore we will  extend ADIOS~\cite{lofstead:2009:adaptible} with new APIs and novel
+techniques to coordinate with the storage back-end. The key idea is offering
+the ability to negotiate with the storage system to agree on the trade-off
+between data quality and response time.
 
-In this project we will extend ADIOS~\cite{lofstead:adaptible} with new APIs and novel
-techniques to coordinate with the storage back-end. The key idea is offering
-the ability to negotiate with the storage system to agree on the tradeoff
-between data quality and response time. We will test our ideas on the next
-generation OLCF, SNL/LANL, and NERSC systems.
-
-We plan to leverage the many inovations we discovered in the Lightweight File
-Systems~\cite{oldfield:lwfs} (LWFS) project and the Ceph~\cite{weil:osdi06} storage
-system project.  In the LWFS project, Sandia sought to demonstrate stripping
-down a file system to the bare minimum features required for any storage
-system. Then, additional functionality and semantics, such as POSIX-style
-directories and consistency requirements, were added in using auxiliary
-services separate from the stack or by layering providing a new interface to
-the storage system. The project's first phase demonstrated that using a core
-object storage system with integrated security providing authentication and
-authorization services, customized storage interfaces could be developed with
-exactly the overhead required for different applications. The second phase of
-LWFS, known as Sirocco~\cite{sirocco}, is investigating how to manage storage
-devices for efficient and scalable storage performance.  Sirocco has been in
-process for three years and has developed a usable storage infrastructure.
-
+We will also take advantage of the many innovations we discovered the Ceph~\cite{weil:osdi06} storage
+system project. 
+%
+%  In the LWFS project, Sandia sought to demonstrate stripping
+%down a file system to the bare minimum features required for any storage
+%system. Then, additional functionality and semantics, such as POSIX-style
+%directories and consistency requirements, were added in using auxiliary
+%services separate from the stack or by layering providing a new interface to
+%the storage system. The project's first phase demonstrated that using a core
+%object storage system with integrated security providing authentication and
+%authorization services, customized storage interfaces could be developed with
+%exactly the overhead required for different applications. The second phase of
+%LWFS, known as Sirocco~\cite{sirocco}, is investigating how to manage storage
+%devices for efficient and scalable storage performance.  Sirocco has been in
+%process for three years and has developed a usable storage infrastructure.
+%
 The Ceph project (ceph.com) started at UC Santa Cruz in 2004 and was initially

### 2015-07-07 narrative/placement.tex
some changes

@@
 generate data) and the storage capability (the capability to store the
-generated data). Moreoever, the complexity and heterogeniety of the
+generated data). Moreover, the complexity and heterogeneity of the
 architecture will introduce new tiers in the storage system. Instead of
-being a flat, roughly homogenous space in which data can be stored, the new
+being a flat, roughly homogeneous space in which data can be stored, the new
 storage hierarchy will be multi-tiered and greatly heterogeneous. 
-
-We address three main concerns for the data lifecycle in this work. First,
-we will explore the initial placement of data during the output process.
-Second, we will study techniques that can enhance migration without making
-the eventual lookup of data unbounded. And finally, we will study how data
-at each tier is managed. %FIXME
+%
+We address three main concerns for the data life-cycle in this work. First,
+we will {\it explore the initial placement} of data during the output process.
+Second, we will study techniques that can {\it enhance migration} without making
+the eventual lookup of data unbounded. And finally, we will study {\it manage each tier} in a scalable fashion. %FIXME
 
 Towards this end, we will utilize the faster layer of the
-storage hierarchies, such as DRAM, NVRAM and remote NVRAM, as the target for
+storage hierarchies, such as DRAM and NVRAM as the target for
 the initial write from an application. To mitigate the space concerns, we
-will use the techniques described in Section~\ref{sec:refactoring} to
+will use the techniques described in \S~\ref{sec:data-refactor} to
 identify and partition the segments of data with the highest utility for the
 application and use case. The rest of the data will be stored on the
-parallel file system, as is done today. We will leverage the supported data
-mirgration techniques in \Sir to determine the most appropriate location to
+lower layers of the file system, such as the parallel file system, as is done today. 
+We will leverage the supported data
+migration techniques in \Sir~to determine the most appropriate location to
 move the data once its in stable storage. This migration will be guided by a
 combination of the utility function defined by the application and the
-resources avialable to the storage system. Knowing the specific structural
+resources available to the storage system. Knowing the specific structural
 and semantic information (including relationships between data segments)
 about the data will enable migration to partition the data further to
-optimize access to data for reader. 
+optimize access to data for reader. For example, in Fig.~\ref{fig:ssio-bucket},
+P3 is initially placed in Campaign storage, since it's size and utility allow our system
+to place data there, but as the utility of data allows this chunk to be migrated to the 
+Long Term, and slower, storage layer to allow other data chunks from other users to take advantage
+of this space. Similar, P2 migrates from the Parallel FS to Campaign storage to clear room for other
+users. We will utilize this migration and eviction strategy for Checkpoint/Restart (C/R) data as well, 
+where data can be initially placed on NVRAM and then purged after the next (or next two) C/R files
+are written to NVRAM.
+
 We will also explore policies that combine the knowledge of data, its

### 2015-07-07 narrative/intro.tex

@@
 5 & Design and implement of IO middleware architectures  & We will create APIs to guide the placement of chunks of data across the SSIO and autonomically move data when necessary& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
-6 & Approaches to improve the ability of SSIO software to support checkpoint/restart & We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then purged when it is no longer needed & \S \ref{sec:???}  \\
+6 & Approaches to improve the ability of SSIO software to support checkpoint/restart & We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then purged when it is no longer needed & \S \ref{fig:ssio-bucket}  \\
 \hline

### 2015-07-07 narrative/enabling_tech.tex
edits for 1.2

@@
 %1
-At the lowest level is the Lightweight File Systems ({\bf LWFS}) project~\cite{oldfield:lwfs}.
-which had the  goal  to provide the simplest core services for
-a storage system and having  added functionality  using auxiliary
-services. 
+At the lowest level is the Lightweight File Systems ({\bf LWFS}) project~\cite{oldfield:lwfs},
+which provides the simplest core services for a storage system with additional functionality
+for using auxiliary services.
 %For example,

### 2015-07-07 narrative/enabling_tech.tex
edits for 1.2

@@
 %devices. This project, 
-{\bf Sirocco} cam out of the LWFS project and offers a peer-to-peer style data storage device
+{\bf Sirocco} came out of the LWFS project and offers a peer-to-peer style data storage device
 mesh with data migrating based on use, required resilience, and available
@@
 resilience to move the data to before deleting the local copy. Further, when
-data is requested, a client can ask any Sirocco server and Sirocco will find
-the data returning it to the client. If desired, the canonical data location
-can be moved on client request to offer better performance for the client. In
-additional features, a stored object can have multiple data forks that can
+data are requested, a client can ask any Sirocco server and Sirocco will find
+the data and return it to the client. If desired, the canonical data location
+can be moved on client request to offer better performance. Additionally,
+a stored object can have multiple data forks that can
 store different versions or aspects of the same data.  For this project, we

### 2015-07-07 narrative/enabling_tech.tex
edits for 1.2

@@
 
-  From the user perspective, something like the
-{\bf ADIOS}~\cite{lofstead:2009:adaptible} API offers an API nearly as simple as POSIX, but with the
+From the user perspective,
+{\bf ADIOS}~\cite{lofstead:2009:adaptible} offers an API nearly as simple as POSIX, but with the
 ability to change the actual data transport mechanism without changing the
-source code. This affords writing to a single output API while shifting from
-writing to a file system to operating on the data using in situ techniques, without
-changing the source code. ADIOS has proven effective for writing to POSIX file
+source code. This flexibility allows users to shift from writing data to a file system
+to operating on data with in situ techniques without changing the source code.
+ADIOS has proven effective for writing to POSIX file
 systems, data staging systems~\cite{zheng:2010:predata,docan:2010:dataspaces,oldfield:2006:nessie}, workflow
 systems~\cite{Dayal:2014:flexpath}, and even through the WAN network 
-Part of this project's goals seeks to understand where the dividing line
+Part of this project's goals are to understand where the dividing line
 between the storage layer shared by all users and middleware being used by a

### 2015-07-07 narrative/data-reduction.tex

@@
 generated. In future, it will be vital to take advantage of a priori
-information when writing and reading from the SSIO layer in order to gain
+information when a) writing and reading from the SSIO layer in order to gain
 performance and predictability, and to prioritize I/O so that the data that
-is most useful for end-user can be written in the time available.
+is most useful for end-user can be written in the time available, and b) in
+performing {\em in situ} analysis before storing the information.
 
@@
 which data should be sent over to our SSIO layer in order that the best
-possible information is available for the subsequent analysis. The decision as how
-this should be accomplished generally involves data refactoring (re-organizing and
-reducing the data).
+possible information is available for the subsequent analysis. The decision as
+how this should be accomplished generally involves data refactoring
+(re-organizing and reducing the data).
 
-In this section we describe some common methods to reduce and re-organize data.
-There are many possibilities and the best choice will generally vary from case
-to case but, once settled, will not change from run to run. It will be
-important to understand when the time and resources required to identify and
-perform the ``best methods'' outweighs any eventual gains. Another critical
-research question concerns the quantification and control of the loss of
-information resulting from refactoring the data and using a reduced 
-dataset.
+There are many possibilities for refactoring and the best choice will generally
+vary from case to case but, once settled, will not change from run to run. It
+will be important to understand when the time and resources required to
+identify and perform the ``best methods'' outweighs any eventual gains. Another
+critical research question concerns the quantification and control of the loss
+of information resulting from refactoring the data and using a reduced dataset.
 
@@
 
-Ideally, the scientists would perform the entire analysis stage {\em in situ} to
-extract the relevant information and, in the unlikely event of this being
-possible, would effectively circumvent the large data issue completely.  The
-problem of course is that this is unlikely to be possible since, by their
-nature, large scale simulations aim to discover new information that is often
-hidden in the form of higher order effects amongst the data deluge. In
-particular, this means if data thinning or truncation is applied willy nilly
-then one is in danger of effectively throwing out the baby with the bath water.
-One cannot store the entire data set in an easily accessible source, due to
-sheer size, yet one cannot reduce it prior to archival without risking losing
-the information one is trying to extract.  Viewed in this way, the problem
-would appear intractable. However, the above discussion fails to recognize that
-much of the data is redundant in the information theoretic sense.  That is to
-say, the amount of information contained in the data is often significantly
-less than the amount of data.  The difficulty stems from the fact that one does
-not know about this redundancy in advance \emph{without the benefit of 
-priori knowledge}. This knowledge can often be provided by the user or can 
-be acquired through experience of dealing with differing runs of the same
-code.
+Ideally, the scientists would perform the entire analysis stage {\em in situ}
+to extract the relevant information and in so doing would effectively
+circumvent the large data issue completely.  The catch, of course, is that this
+is unlikely to be possible since, by their nature, large scale simulations aim
+to discover new information that is often hidden in the form of higher order
+effects amongst the data deluge. In particular, this means if data thinning or
+truncation is applied willy nilly then one is in danger of effectively throwing
+out the information one is seeking. It would seem that one cannot store the
+entire data set in an easily accessible source, due to sheer size, yet one
+cannot reduce it prior to archival without risking losing the information one
+is trying to extract.  Viewed in this way, the problem would appear
+intractable. However, the above discussion fails to recognize that much of the
+data is redundant in an information theoretic sense.  That is to say, the
+amount of information contained in the dataset is often significantly less than
+the amount of data.  The difficulty stems from the fact that one does not know
+about this redundancy in advance \emph{without the benefit of priori
+knowledge}. This knowledge can often be provided by the user or can be acquired
+through experience of dealing with differing runs of the same code.
 
-Application knowledge means one can sometimes achieve drastically superior
+Deep application knowledge means one can sometimes achieve drastically superior
 reduction in the total data compared with what one might achieve otherwise.
-However, in the absence of such knowledge, we must equip the SSIO layers with
-generic data reduction techniques and re-organization techniques.
-Nevertheless, certain basic information on the semantics of the data is needed
-and must be supported by the overall infrastructure.  Likewise, another
-important consideration is knowledge regarding typical access patterns for data
-in a given application. Scientific applications often read and analyze data in
-a structured fashion and failure to respect this pattern, or to exploit it
-through prefetching, can significantly increase the overall execution time of
-the application. Accurate prefetching can only be achieved with access pattern
-analysis. This requires an SSIO layer which is able to prefetch data in an
-asynchronous fashion to take the bytes on the file.
+However, even in the absence of such high level knowledge, we must equip the
+SSIO layers with generic data reduction techniques and re-organization
+techniques. For instance, certain basic information on the semantics of the
+data is needed and must be supported by the overall infrastructure. Likewise,
+information regarding typical access patterns for data in a given application
+should be placed in the SSIO layers. Scientific applications often read and
+analyze data in a structured fashion and failure to respect this pattern, or to
+exploit it through prefetching, can significantly increase the overall
+execution time of the application. Accurate prefetching can only be achieved
+with access pattern analysis. This requires an SSIO layer which is capable of
+prefetching data in an asynchronous fashion to take the bytes on the file.
 
@@
 catch-all methods, fail to exploit is a priori information on the nature and
-provenance of the data. Moreover, it is now widely recognized that any
-opportunities for {\em in situ} analysis should be exploited as far as
-possible. What is less well understood, is how one should incorporate a priori
-information and how one can identify possibilities for {\em in situ} analysis.
-
-We shall exploit a priori information in the preprocessing stage through the
-use of relatively computationally intensive preprocessors that attempt to
-extract structure in the data {\em in situ}. This approach takes advantage of
-the increasingly wide cost gap between CPU operations versus storage-based
-operations. The preprocessor would generally be employed in situ and is used in
-a similar fashion to pre-processors in standard compression approaches such as
-FPC \cite{BurtscherFPC} in which an XOR operation against the previous data is
-used as a pre-processor.
+provenance of the data.
 
@@
 tang2014improving}
-has focused on many of these well known techniques to not only reduce the data
+has focused on several alternative techniques to not only reduce the data
 but also re-organize the data. Previously we have examined:
@@
 
-\paragraph{Proposed research approach:} We will be looking at a multi-throng
+\paragraph{Proposed research approach:} It is now widely recognized that any
+opportunities for {\em in situ} analysis should be exploited as far as
+possible. What is less well understood, is how one should incorporate a priori
+information and how one can identify possibilities for {\em in situ} analysis.
+A preprocessor would be employed in situ and used in a similar fashion to
+pre-processors in standard compression approaches such as FPC
+\cite{BurtscherFPC} in which an XOR operation against the previous data is used
+as a pre-processor.  We shall exploit a priori information in the preprocessing
+stage through the use of relatively computationally intensive preprocessors
+(compared with FPC) that seek to extract structure in the data {\em in situ}.
+This approach takes advantage of the increasingly wide cost gap between CPU
+operations versus storage-based operations.
+
+Ideally, the scientists would perform the entire analysis stage {\em in situ}
+to extract the relevant information and in so doing would effectively
+circumvent the large data issue completely.  The problem of course is that this
+is unlikely to be possible since, by their nature, large scale simulations aim
+to discover new information that is often hidden in the form of higher order
+effects amongst the data deluge. In particular, this means if data thinning or
+truncation is applied willy nilly then one is in danger of effectively throwing
+out the information one is seeking. It would seem that one cannot store the
+entire data set in an easily accessible source, due to sheer size, yet one
+cannot reduce it prior to archival without risking losing the information one
+is trying to extract.  Viewed in this way, the problem would appear
+intractable. However, the above discussion fails to recognize that much of the
+data is redundant in an information theoretic sense.  That is to say, the
+amount of information contained in the dataset is often significantly less than
+the amount of data.  The difficulty stems from the fact that one does not know
+about this redundancy in advance \emph{without the benefit of priori
+knowledge}. This knowledge can sometimes be provided by the user, can be
+acquired through experience of dealing with differing runs of the same code, or
+acquired during a run by adaptively selecting and tuning the algorithm based on
+metrics during runtime.
+
+We will assemble a suite of general techniques for refactoring of data that
+will be embedded in the system and can be selected either by the user or by the
+application itself. The techniques will range from crude catch all methods such
+as those employed by general compression routines such as FPC, through to
+multi-resolution analysis using wavelets, including the methods on which we
+have worked (cited above).  These will then be mapped over into the different
+layers of the storage hierarchy.  We will also cater for user supplied
+``plug-ins'' for detection of features and store higher resolution data around
+the feature and lower resolution values which are further away from the
+features. The plug-ins could be used to override the general embedded
+techniques and would provide a mechanism for the transfer of a priori knowledge
+from the user to the hierarchy without imposing a burden on either.
+
+We will also research the time it takes to re-organize the data and use the
+multi-layer approach to staging to understand where the data gets reduced and
+where it gets re-organized. Open questions are: 1) do we move the data over to
+other nodes to then reduce and re-organize, 2) do we first reduce the data and
+then later re-organize this data on different nodes?  3) do we use different
+cores on a node to do data-refactoring.  Similarly we will examine the cost of
+reading the data. It is clear that if we only read in the high-importance data
+it should improve the overall time to move the data from the SSIO layer to the
+application, but what happens when users want all of the data? There is an
+obvious penalty of having to take data in a hierarchical layout and then place
+this back to a single-layer view.  Additional research questions which must be
+addressed: Will some types of data (PIC data, finite element data, ...) be be
+conducive  for data-refactoring than others? How much semantic-domain knowledge
+will we need to get a dramatic improvement of the storage system over not doing
+anything special?
+
+We will measure these methods using the fusion XGC simulation and the
+corresponding analysis routines which they use for their largest simulations.
+We will also work with Jeroen Tromp's SPECFM3D~\cite{SPECFEM3D} code, and the
+fusion GTC~\cite{klasky2003grid} code.  In all cases we will measure their
+impact from their current simulations, and their projected simulations and
+analytics on exascale machines. Since all three of these codes are part of the
+OLCF CAAR~\cite{CAAR} project for early access, there is strong belief that the
+simulations will be close to exascale simulations and we will be able to run
+these on prototypes of our SSIO layer to examine the full impact of
+data-refactoring.
+
+\paragraph{High Impact Challenges:} Maximal data reduction on the SSIO layer
+would result from storing only the code and the input parameters, and simply
+recalculating all of the results. This extreme is not feasible in practice
+owingto the size of the computation involved, but sheds light on the path to
+achieving a step change in data reduction. We shall investigate the possibility
+of writing a minimal set of information, and placing an ``auditor'' code in a
+docker container so that data can be re-generated with a {\em local}
+approximation of the full simulation. This is an entirely new approach to
+tackle the data overload by using preprocessors in which information is not
+simply stored in the form of raw data, but rather in the form of appropriate
+parameterized algorithms. A priori information on the nature of the data is
+used to choose the basic algorithm while the actual values of the data are used
+to tune the parameters used in the reconstruction procedures. We shall refer to
+this paradigm as an Auditor.  We will examine the use of several auditors for
+the simulations which we will work with. One approach is to use a fluid-code to
+audit a kinetic code, which can be a good approximation over short periods of
+time. Once there is a large deviation (specified by the user) from the auditor
+to the original calculation, the data will be saved. This is a local operation,
+since the deviations will grow differently in space and in time.
+
+%%%%%%%%%%%%%
+We will be looking at a multi-throng
 approach to data-refactoring. One approach allows the user to specify
@@
 in both space and time.
-
-We will examine the different techniques for
-compression and re-organization along with examine techniques to use an
-error-estimator (gradients) to create a multi-resolution view of the data which
-can then be sampled, and placed in different quantiles of importance. These
-will then be mapped over into the different layers of the storage hierarchy.
-
-We will also investigate technique which can use ``plug-ins'' for detection of
-features and store higher resolution data around the feature and lower
-resolution values which are further away from the features. This
-re-organization must occur without increasing the total amount of storage than
-what the users intended.
-
-We will also research the time it takes to
-re-organize the data and use the multi-layer approach to staging to understand
-where the data gets reduced and where it gets re-organized. Open questions are:
-1) do we move the data over to other nodes to then reduce and re-organize, 2)
-do we first reduce the data and then later re-organize this data on different
-nodes?  3) do we use different cores on a node to do data-refactoring.
-Similarly we will examine the cost of reading the data. It is clear that if we
-only read in the high-importance data it should improve the overall time to
-move the data from the SSIO layer to the application, but what happens when
-users want all of the data? There is an obvious penalty of having to take data
-in a hierarchical layout and then place this back to a single-layer view.
-Additional research questions which must be addressed: Will some types of data
-(PIC data, finite element data, ...) be be conducive  for data-refactoring than
-others? How much semantic-domain knowledge will we need to get a dramatic
-improvement of the storage system over not doing anything special?
-
-We will measure these methods using the fusion XGC simulation and the
-corresponding analysis routines which they use for their largest simulations.
-We will also work with Jeroen Tromp's SPECFM3D code, and the fusion GTC code
-from Z. Lin. In all cases we will measure their impact from their current
-simulations, and their projected simulations and analytics on exascale
-machines. Since all three of these codes are part of the OLCF CAAR~\cite{CAAR}
-project for early access, there is strong belief that the simulations will be
-close to exascale simulations and we will be able to run these on prototypes of
-our SSIO layer to examine the full impact of data-refactoring.
-
-\paragraph{High Impact Challenges:} Since the greatest form of data reduction
-on the SSIO layer would be to only write the code, and the input data, and then
-recalculate all of the results, we want to understand how to write a minimal
-set of information, and then place an auditor code in a docker container so
-that data can be re-generated with an approximation of the calculation, but
-only valid over a limited amount of time.This new approach to tackle the data
-overload by using preprocessors in which information is not simply stored in
-the form of raw data, but rather in the form of appropriate parameterized
-algorithms. A priori information on the nature of the data is used to choose
-the basic algorithm while the actual values of the data are used to tune the
-parameters used in the reconstruction procedures. We shall refer to this
-paradigm as an Auditor.
-
-We will examine the use of several auditors for the simulations which we will
-work with. One approach is to use a fluid-code to audit a kinetic code, which
-can be a good approximation over short periods of time. Once there is a large
-deviation (specified by the user) from the auditor to the original calculation,
-the data will be saved. This is a local operation, since the deviations will
-grow differently in space and in time.

### 2015-07-07 narrative/data-description.tex
clarified example and figure in data description

@@
 The user would like to have the ability to obtain information about, and even
-negotiate with the system to determin what
+negotiate with the system to determine what
 level of Quality of Service is possible given
@@
 Sec~\ref{sec:data-refactor}) which will bucket the particle data into 4
-prioritization levels.  The user can then pick a data utility, such as 3 months
+prioritization levels in the memory. The user can then pick a data utility, such as 3 months
 for the highest priority items and 12 months for the lowest priority data.
@@
 written (particles, two fields, and a mesh in this example) and the system
-will return an estimate of the time it will take to write. The user can then
-use another call which will determine the lowest priority element (in this
-case the user will write all of the data (P1-P4). The data will then be stored
+will return an estimate of the time it will take to write. If that it unacceptably large, the user can 
+use another call to specify the lowest priority element that they want to actually write, 
+skipping the rest too save time (in this
+example the user will write all of the data). The data will then be stored
 across the layers of the file system where the two highest priority buckets
 will be stored in the Parallel File System, but the next highest priority items
-will be stored lower in the storage stack. 
+will be stored lower in the storage stack. Depending on the capabilities of the actual storage system, 
+SSIO will decide if the data has to go through the parallel file system 
+towards a lower stack or can be staged to that stack directly. 
 %
 
-\begin{wrapfigure}{r}{0.7\textwidth}
-%\vspace{0ex}
+\begin{wrapfigure}{r}{0.6\textwidth}
+        \vspace{-2ex}
         \begin{centering} 
-	\includegraphics[scale=0.75]{graphics/SSIO-bucket.pdf}
+	\includegraphics[scale=0.8]{graphics/SSIO-bucket.pdf}
         %\vspace{1ex}
@@
         \end{centering}
-%        \vspace{1ex}
+       \vspace{-1ex}
 \end{wrapfigure}
@@
 the optimal data-refactoring routine. Currently, we understand that
-users can tell us that the data represents: 1) phase-space (e.g., the particles
-in our example), 2) spatial-changes (e.g., the mesh), 3) space-time (e.g.,
+users can tell us that the data represents: 1)~phase-space (e.g., the particles
+in our example), 2)~spatial-changes (e.g., the mesh), 3)~space-time (e.g.,
 the fields on the mesh). We need this information to select

### 2015-07-07 narrative/data-reduction.tex
Slight change to refactoring

@@
 
-Ideally, the scientists would perform the entire analysis stage {\em in situ}
-to extract the relevant information and in so doing would effectively
-circumvent the large data issue completely.  The problem of course is that this
-is unlikely to be possible since, by their nature, large scale simulations aim
-to discover new information that is often hidden in the form of higher order
-effects amongst the data deluge. In particular, this means if data thinning or
-truncation is applied willy nilly then one is in danger of effectively throwing
-out the information one is seeking. It would seem that one cannot store the
-entire data set in an easily accessible source, due to sheer size, yet one
-cannot reduce it prior to archival without risking losing the information one
-is trying to extract.  Viewed in this way, the problem would appear
-intractable. However, the above discussion fails to recognize that much of the
-data is redundant in an information theoretic sense.  That is to say, the
-amount of information contained in the dataset is often significantly less than
-the amount of data.  The difficulty stems from the fact that one does not know
-about this redundancy in advance \emph{without the benefit of priori
-knowledge}. This knowledge can sometimes be provided by the user, can be
-acquired through experience of dealing with differing runs of the same code, or
-acquired during a run by adaptively selecting and tuning the algorithm based on
-metrics during runtime.
-
 We will assemble a suite of general techniques for refactoring of data that

### 2015-07-07 narrative/enabling_tech.tex
edits for 1.2

@@
 For this project, we plan to use the Sirocco data storage mechanism to manage
-the actual data storage on all of the different tiers. We will add data management functionality as part of the storage
+the actual data storage across the different tiers. We will add data management functionality as part of the storage
 system layer to handle data discovery and metadata tasks. At the middleware
@@
 
-We will also take advantage of the many innovations we discovered the Ceph~\cite{weil:osdi06} storage
+We will also take advantage of the many innovations discovered in the Ceph~\cite{weil:osdi06} storage
 system project. 
@@
 object-based parallel file system that addresses the well-known metadata
-service bottleneck of creating and accessing 1000s of files comprising a
+service bottleneck of creating and accessing 1000s of files from a typical 
 checkpoint~\cite{weil:osdi06}.  However, after the Ceph kernel client was
@@
 key innovations of Ceph: (1) Controlled Replication Under Scalable Hashing
-(CRUSH)~\cite{weil:sc06} will help making resource identification and
+(CRUSH)~\cite{weil:sc06} will help make resource identification and
 predictable performance scalable, and (2) extensible storage object

### 2015-07-07 narrative/proposed-work.tex
edits for 1.3

@@
 
-\underline{Principle 1:} {\bf A knowledge-centric system design}, let user knowledge  to
+\underline{Principle 1:} {\bf A knowledge-centric system design} that allows user knowledge  to
 define data policies. Today Storage System and I/O layers are written in a
 stove-pipe fashion, and quite often do not allow optimizations to take place.
-We propose to re-design the layer in a highly integrated fashion where user
-will place their intentions into the system and actions will dynamically and
-statically take place to optimize for both the system and for individual
+We propose to re-design the layer in a highly integrated fashion where users
+will place their intentions into the system and actions will statically and dynamically
+take place to optimize for both the system and for individual
 requests.

### 2015-07-07 narrative/data-description.tex
edits for 1.3

--- 2015-07-07 narrative/data-description.tex
@@
 consist of a collection of chunks, possibly with different accuracy or
-resolution, and with each chunk representing
+resolution, and each chunk representing
 a different portion of the variable.
 A key advantage of having attributes at this lowest level is that it allows data
-semantics, user intentions, QoS requirements, and the relationship between user
+semantics, user intentions, QoS requirements, and the relationships between user
 data to be readily captured and embedded with the raw data so that these
 smallest pieces can each be managed independently or collectively, depending on
-the need. As an example, retrieving 3D field data generated from a simulation
-and then visualizing it requires the coordinates and connectivity variables be
+the need. As an example, retrieving and visualizing 3D field data generated from a simulation
+requires the mesh (consiting of coordinates and connectivity variables) be
 accessed simultaneously and with low latency to achieve a good user experience.

### 2015-07-07 narrative/data-description.tex
edits for 1.3

@@
 relationship to other data. In particular, the relationship to other chunks can
-be captured by including \textit{typed links} each representing a different
+be captured by including \textit{typed links}, each representing a different
 relationship type.  In this project, a key metric stored as a metadata
@@
 negotiate with the system to determine what
-level of Quality of Service is possible given
+level of QofS is possible given
 a prospective set of I/O operations and current state of the system, so that

### 2015-07-07 narrative/data-description.tex
edits for 1.3

@@
 capability and further augment the format to express data utility.
-Recently
+Recently,
 DAMSEL~\cite{damsel} has provided a rich metadata representation and management
@@
 and access that allow users to describe the data utility based on their
-expectations. This can be done by allowing users to plug in functions that
+expectations. This can be done by allowing users to apply plugins
 provide data-specific functionality to the middleware and storage system.
-These functions may be executed by the system to
+These plugins may be executed by the system to
 calculate the utility value for each chunk, handle custom compression
@@
 
-We will explore the space of application level hints that can be provided easily
+We will explore the space of application level hints that can be easily provided
 to the storage system. These hints will enable the storage system to make the

### 2015-07-07 narrative/data-reduction.tex

@@
 \item Precision based re-organization. This is where the most significant bytes
-	of the data are all group together from each object. This data
+	of the data are all grouped together from each object. This data
 	generally will have a higher utility then the data with the least
@@
 	storage system. This is often the case for many of the data quantities
-	written form the simulation, but there are many cases where, for
+	written from the simulation, but there are many cases where, for
 	example, we want to write all of the particles from a Particle In Cell
@@
 this back to a single-layer view.  Additional research questions which must be
-addressed: Will some types of data (PIC data, finite element data, ...) be be
+addressed: Will some types of data (PIC data, finite element data, ...) be
 conducive  for data-refactoring than others? How much semantic-domain knowledge
@@
 recalculating all of the results. This extreme is not feasible in practice
-owingto the size of the computation involved, but sheds light on the path to
+owing to the size of the computation involved, but sheds light on the path to
 achieving a step change in data reduction. We shall investigate the possibility

### 2015-07-07 appendix-data-management-plan.tex
DM plan

\section*{Appendix 6: Data Management plan}

\vspace{1ex}
\noindent
The relevant data for the proposed work will be from the software we develop from our team, the documentation for our software, the performance data we use in our publications,
and the requirement documents we get from any of the users who we are working with.

Our team is  will be sharing these data artifacts with the broader community. ORNL will setup a dedicated website for this project which will have all relevant information, along with
a link to all of the new software that will be developed using an open source license.  The source code will be maintained using a versioning system and repository, Github, that is widely used and that we expect will also be well known for the next 10 years.

The documentation will be made available online on the project web site. The performance data will be distributed through publications and technical reports. Publications and technical reports will be made available on the project web site as well. In addition, journal, conference, and workshop papers are also typically archived by publishers (e.g. ACM, IEEE, or Springer). Similarly, Oak Ridge and Sandia National Laboratories maintain archives of all the technical reports. The project web site will maintain pointers to all publications and reports, along with all data that was used in the publications.

### 2015-07-07 narrative/qos.tex
minor changes.

@@
 The opposite of file systems are ``pile'' systems, which, instead of keeping track of updates, search for 
-resources at request time. Pile systems allow very efficient writes because clients can write whereever
+resources at request time. Pile systems allow very efficient writes because clients can write wherever
 there is free space and there is very low bookkeeping overhead. However, unbounded search in pile systems
@@
 of allowing applications to compete freely without restriction, for part or all of the storage
-resources. This traditional approach leads to intereference effects
+resources. This traditional approach leads to interference effects
 \cite{lofstead:2010:io-variability,liu_hotstorage} that can greatly impact IO performance predictability. 
@@
 performance.
-In the cases where an appication arrives that needs to run more urgently
+In the cases where an application arrives that needs to run more urgently
 than any of the other already running applications, the system
@@
 and later reverting back to the previous committed resource level for the rest.
-This can be automatically enforced via user- (or adiminstrator-) specified \textit{policies}
+This can be automatically enforced via user- (or administrator-) specified \textit{policies}
 (section \ref{}).
@@
 
-\paragraph{Challenges:} The deep and heterogenous memory and storage
+\paragraph{Challenges:} The deep and heterogeneous memory and storage
 hierarchy we are assuming for this proposal complicates the
@@
 \item The complexity of a utilization model involving the entire
-storage hieararchy with 100,000s of devices is potentially daunting.
+storage hierarchy with 100,000s of devices is potentially daunting.
 However, the utilization model can be simplified by modeling classes

### 2015-07-07 narrative/placement.tex
minor changes for a reference

@@
 \paragraph{Application Hints:}
-As described in Section~\ref{sec:refactor} one of the main components of our
+As described in \S\ref{sec:data-refactor} one of the main components of our
 proposed storage system is the ability to refactor and reduce data as it is

### 2015-07-07 narrative/placement.tex
Still keeping lock before call with Manish

--- 2015-07-05 narrative/placement.tex
@@
 %objectives (and that leads into quality of service)
+
 The application workflows targeted in this proposal generate very large
@@
 potential insights from the simulations can be realized. 
-In-situ and in-transit data staging approaches can alleviate some of the
-burden on storage, leveraging available compute and memory resources on high
-end systems to process the data close to where it is generated. 
-However, given current architectural trends, emerging systems will have a
-vastly great gulf between the computing capability (the capability to
-generate data) and the storage capability (the capability to store the
-generated data). Moreoever, the complexity and heterogeniety of the
-architecture will introduce new tiers in the storage system. Instead of
-being a flat, roughly homogenous space in which data can be stored, the new
-storage hierarchy will be multi-tiered and greatly heterogeneous.
+The overall data life cycle consists of three phases. Data starts
+when it is generated by a simulation or through a data acquisition system,
+in the case of experimental and observational data (EOD). 
+In our target heterogenous multi-tier storage platform, the system must
+decide where to place the output data at this time, and how to partition and
+reorganize the data to optimize this initial placement. The key optimization
+metrics during this phase is the percieved I/O overhead for the application.
 
-We address three main concerns for the data lifecycle in this work. First,
-we will explore the initial placement of data during the output process.
-Second, we will study techniques that can enhance migration without making
-the eventual lookup of data unbounded. And finally, we will study how data
-at each tier is managed. %FIXME
+In the second phase of the data life cycle, the system takes over the
+management of the data and performs the necessary steps to meet service
+level objectives (SLOs). These steps include migration, reorganization, and
+reductions for the data. During this phase the system optimizes for resource
+utilization, interference with running applications, and availability and
+resiliency objectives. Data can be moved into archival storage (tape or cold
+storage) during this phase.
 
-Towards this end, we will utilize the faster layer of the
-storage hierarchies, such as DRAM, NVRAM and remote NVRAM, as the target for
-the initial write from an application. To mitigate the space concerns, we
-will use the techniques described in Section~\ref{sec:refactoring} to
-identify and partition the segments of data with the highest utility for the
-application and use case. The rest of the data will be stored on the
-parallel file system, as is done today. We will leverage the supported data
-mirgration techniques in \Sir to determine the most appropriate location to
-move the data once its in stable storage. This migration will be guided by a
-combination of the utility function defined by the application and the
-resources avialable to the storage system. Knowing the specific structural
-and semantic information (including relationships between data segments)
-about the data will enable migration to partition the data further to
-optimize access to data for reader. 
-We will also explore policies that combine the knowledge of data, its
-utility to users, the cost of maintaining data in long term storage and
-other metrics to decide how to purge irrelevant data from the storage
-system.
+In the third phase the data is consumed for knowledge discovery (for example
+read into analytics to produce new scientific insights). It must be noted
+that the knowledge discovery process does not remove the data from the
+system - the data remains available until it is purged. During this phase
+the storage system optimizes for read performance and data availability.
 
-Our research will ultimately address the following questions:\\
-1) Can we do scalable migration without making the discovery process
-    unbounded;\\
-2) What are the parameters and input for migration;
-3)  Can we use the additional application level knowledge of data to purge
-    portions of the data from storage without making the data too much less
-    valuable?;\\
-3)  With so many objects and no centralized directory how do we know what
-    is actually in the storage to be purged?;\\
-4)  How do purge in a way that doesn't require a central authority,
-    doesn't interfere with other I/O and isn't bottlenecked by some sort of
-    global consistency?
+% In the final phase the data is purged from the system altogether. The
+% decision to purge is made by the system based on operational metrics from
+% system administrators. This phase optimizes the space available in the
+% system to prevent out of space concerns for applications.
+
+Our proposed research explores the management of this entire life cycle
+given our thesis that more knowledge, and more specific knowledge about the
+data, can allow the management of data to be optimized with reduced cost.
+Thus, we will explore the following high level research questions: (1) What methods can
+be utilized in the initial placement of data that do not impose expensive
+costs on discovery, consumption and migration? (2) How can the placement of
+data be optimized for the multi-tiered hierarchy expected in exascale
+systems? (3) How can we migrate data throughout the layers of storage
+without making the discovery process unbounded? (4) How does the additional
+knowledge aid in preparing the data for reading in a timely manner? (5) How
+do we make the decision on purging using this additional knowledge?
+
+ % Next we provide details on the challenges and proposed research through each
+ % phase of the data life cycle.
+
+% We address three main concerns for the data life-cycle in this work. First,
+% we will {\it explore the initial placement} of data during the output
+% process. Second, we will study techniques that can {\it enhance migration}
+% without making the eventual lookup of data unbounded. And finally, we will
+% study who to {\it manage each tier} in a scalable fashion. %FIXME
+
+% \subsubsection{Initial Placement of Data}
+% \label{sec:init-plac-data}
+
+% The generation of data from simulations or EOD sources will require the
+% system to make initial placement decisions. These decisions will leverage
+% the description and refactoring of data mentioned earlier, and also the
+% tiered design of the storage system. The volume and velocity of data at
+% exascale can be alleviated partially by utilizing in-situ and in-transit
+% data staging to process the data before the storage layer. Eventually,
+% however, the data, processed and transformed, has to be sent to storage.
+
+% % In-situ and in-transit data staging approaches can alleviate some of the
+% % burden on storage, leveraging available compute and memory resources on high
+% % end systems to process the data close to where it is generated. 
+% % However, given current architectural trends, emerging systems will have a
+% % vastly great gulf between the computing capability (the capability to
+% % generate data) and the storage capability (the capability to store the
+% % generated data). Moreover, the complexity and heterogeneity of the
+% % architecture will introduce new tiers in the storage system. Instead of
+% % being a flat, roughly homogeneous space in which data can be stored, the new
+% % storage hierarchy will be multi-tiered and greatly heterogeneous. 
+% %
+
+% For the initial placement of data, we will utilize the faster layers of the storage
+% hierarchy, such as DRAM and NVRAM as the target for the initial write from
+% an application as much as possible. These targets can be on-node (such as in
+% the design of the Summit system at ORNL) or on separate nodes (as in the
+% Trinity system at LANL). Regardless, the space limitations might require
+% some of the data to be directly transferred to the parallel file system.
+% Partitioning of data to accomplish this has been described in
+% ~\S~\ref{sec:data-refactor}.
+
+% To mitigate the space concerns, we will use the techniques
+% described in \S~\ref{sec:data-refactor} to identify and partition the
+% segments of data with the highest utility for the application and use case.
+% The rest of the data will be stored on the lower layers of the file system,
+% such as the parallel file system, as is done today. We will leverage the
+% supported data migration techniques in \Sir~to determine the most
+% appropriate location to move the data once its in stable storage. This
+% migration will be guided by a combination of the utility function defined by
+% the application and the resources available to the storage system. Knowing
+% the specific structural and semantic information (including relationships
+% between data segments) about the data will enable migration to partition the
+% data further to optimize access to data for reader. For example, in
+% Fig.~\ref{fig:ssio-bucket}, P3 is initially placed in Campaign storage,
+% since it's size and utility allow our system to place data there, but as the
+% utility of data allows this chunk to be migrated to the Long Term, and
+% slower, storage layer to allow other data chunks from other users to take
+% advantage of this space. Similar, P2 migrates from the Parallel FS to
+% Campaign storage to clear room for other users. We will utilize this
+% migration and eviction strategy for Checkpoint/Restart (C/R) data as well,
+% where data can be initially placed on NVRAM and then purged after the next
+% (or next two) C/R files are written to NVRAM.
+
+% We will also explore policies that combine the knowledge of data, its
+% utility to users, the cost of maintaining data in long term storage and
+% other metrics to decide how to purge irrelevant data from the storage
+% system.
+
+% Our research will ultimately address the following questions:\\
+% 1) Can we do scalable migration without making the discovery process
+%     unbounded;\\
+% 2) What are the parameters and input for migration;
+% 3)  Can we use the additional application level knowledge of data to purge
+%     portions of the data from storage without making the data too much less
+%     valuable?;\\
+% 3)  With so many objects and no centralized directory how do we know what
+%     is actually in the storage to be purged?;\\
+% 4)  How do purge in a way that doesn't require a central authority,
+%     doesn't interfere with other I/O and isn't bottlenecked by some sort of
+%     global consistency?
 
@@
 different layers of the storage hierarchy can have a significant impact on
-the overhead observed by the application for I/O operations. A key
-requirements is application-driven runtime mechanisms for dynamically
+the overhead observed by the application for I/O operations. The key
+requirements here are application-driven runtime mechanisms for dynamically, and
+deterministically, 
 managing data placement across the layers of the distributed storage
@@
 well as horizontally across different storage nodes requires balancing
-competing objectives. In this project we will explore autonomic data
+competing objectives. We will explore autonomic data
 management strategies that can evaluate utility/cost tradeoffs and
@@
 % Any other drivers?
-\paragraph{Approach:} This spectrum enables interesting strategies
+\paragraph{Thinning of Data:} 
+This spectrum enables interesting strategies
 to manage space pressure on high tiers: instead of just evicting
@@
 
-% \paragraph{Related Work:} Mention Stanford's Legion system.
-
-\paragraph{Challenge:} The decision on when to thin and when to
+\paragraph{Locality Properties:} The decision on when to thin and when to
 enrich data on a particular level depends on \emph{access patterns}.
-Approaches that leveragee access patterns in application at national
-labs has shown to greatly improve performance and reduce the amount
+Approaches that leverage access patterns in applications at leadership facilities
+ have shown to greatly improve performance and reduce the amount
 of overhead required for data management~\cite{he:hpdc13}. The
@@
 
-\paragraph{Approach:}
+\paragraph{Bounded estimates for access:}
 Our research intent involves short term and long term activities. In the short
@@
 that would provide some, hopefully tightly, bounded estimate about time to
-first byte for a potential read operation. In the long term we hope that an
+first byte for a potential read operation.
+
+\paragraph{Migration Policies:}
+In the long term we hope that an
 augmented client-server API providing usage hints will allow us to modify local
@@
 
-\paragraph{Preliminary work and results:} Our recent work \cite{tongipdps15}
+\subsubsection{Preliminary work and results} 
+\label{sec:prework}
+
+Our recent work \cite{tongipdps15}
 explored a two-tiered staging method that spans both DRAM and solid state

### 2015-07-07 narrative/data-reduction.tex

@@
 audit a kinetic code, which can be a good approximation over short periods of
-time. Once there is a large deviation (specified by the user) from the auditor
-to the original calculation, the data will be saved. This is a local operation,
-since the deviations will grow differently in space and in time.
-
-%%%%%%%%%%%%%
-We will be looking at a multi-throng
-approach to data-refactoring. One approach allows the user to specify
-information about their data to indicate if the data is spatially varying,
-space-and time varying, just time varying. This allows us to store information
-in NVRAM on each node (when and where necessary) to later preform compression
-in both space and time.
+time.

### 2015-07-07 narrative/deliverables.tex
getting changes for the evaluatoin

@@
-\section*{Deliverables}
+\section{Deliverables}
+The following are the per year project deliverables:
 
+\textbf{Year 1}
 \begin{itemize}
-\item plug-in architecture for Sirocco
-\item storage APIs to manipulate storing and retrieving data to specific tiers
-\item metadata management for data stored in multiple tiers using different compression
-\item metadata management for deeper data knowledge
-\item middleware hooks to manage new storage APIs
-\item example lossy and lossless compression plug-ins to demonstrate the effectiveness of this approach
+\item Develop new techniques for data description that allow users to 
+describe the data utility based on their expectations. (ORNL)
+
+\item Assemble and evaluate general techniques for refactoring of data
+that will be embedded in the system and can be selected by users or applications. (Brown/ORNL)
+\item Explore the use of application hints and utility of data to guide the initial placement of data. 
+(Rutgers/ORNL)
+\item Investigate the trade-off between ``filing'' and ``piling'' data. (UCSC/Sandia)
+\item Demonstrate a metadata service capable of serving both POSIX clients and our clients,
+but without consideration for authentication, authorization, or data migration. (Sandia/UCSC)
 \end{itemize}
+
+\textbf{Year 2}
+\begin{itemize}
+\item Develop a new type of querying system that allows an application to ask the storage system
+about completion timing information. (ORNL)
+\item Research autonomic data management strategies that
+can evaluate utility/cost tradeoff, and appropriately place/move data objects at runtime.
+\item Demonstrate a time bounded search approach for finding data within the storage system
+to identify data and the current location. This will update the cached information in the metadata
+service to short-circuit future requests for the data presuming that it does not migrate again soon. 
+\item Extend Sirocco to enforce data-centric metrics that help guide the QoS decisions.
+(Sandia/UCSC)
+\end{itemize}
+
+\textbf{Year 3}
+\begin{itemize}
+\item Work with DOE applicationexsiting ADIOS-enabled applications, framework using many of the applications to which we have access, including XGC1, GTC, and SPECFM3D
+%\item storage APIs to manipulate storing and retrieving data to specific tiers
+%\item metadata management for data stored in multiple tiers using different compression
+%\item metadata management for deeper data knowledge
+%\item middleware hooks to manage new storage APIs
+%\item example lossy and lossless compression plug-ins to demonstrate the effectiveness of this approach
+\end{itemize}

### 2015-07-07 narrative/enabling_tech.tex

@@
 %
-The Ceph project (ceph.com) started at UC Santa Cruz in 2004 and was initially
+The {\bf Ceph} project (ceph.com) started at UC Santa Cruz in 2004 and was initially
 funded by DOE/NNSA involving LLNL, LANL, and Sandia, with the goal to create an

### 2015-07-07 narrative/data-description.tex
editing section 1.3

@@
 We will also develop a new type of querying system that allows an application
-to ask the storage system about completion timing information. The response
-would be an estimated time for the described operation. We will explore
+to get estimated timing information for operations on the storage system.
+We will explore
 how these query functions can be integrated into common applications with
@@
 respond to this new information. We will also explore how applications can
-query the storage system to gain insights on available space in different tiers
-and how applications can tailor their output data to most effectively use the available space.
+query the storage system on space availability across different tiers and how
+applications their tailor their output data to make the most effective use of
+available space.
 Finally we will study the use of an external data annotation system that can
-provide information to the storage system without requiring the application to
-be recompiled.
+provide information to the storage system without requiring recompilation of the application.
 
@@
 From an application level, we will explore these techniques through our ADIOS
-framework using many of the applications to which we have access, including
+framework using many of our existing application partners, including
 XGC1, GTC, and SPECFM3D. We will investigate: 1) what hints the users can give
@@
 Chunks in a particular bin will be assigned an appropriate storage
- tier and lifetime, which can be decided by the user through the API.
-For example: in Figure~\ref{fig:ssio-bucket} we have data that
-comes from the XGC1 code.  The user can pick a data-refactoring scheme (see
-Sec~\ref{sec:data-refactor}) which will bucket the particle data into 4
-prioritization levels in the memory. The user can then pick a data utility, such as 3 months
+tier and lifetime, which can be specified by the user through the API.
+As an example, Figure~\ref{fig:ssio-bucket} illustrates a use case for the XGC1 code.
+The user can pick a data-refactoring scheme (see
+Sec~\ref{sec:data-refactor}) which will classify the particle data into 4 different
+prioritization levels. The user can then pick a data utility, such as 3 months
 for the highest priority items and 12 months for the lowest priority data.
-The user will then call a function, provided by the API, that allows them
+The user will then call an API function that allows them
 to specify all of the chunks they want
 written (particles, two fields, and a mesh in this example) and the system
-will return an estimate of the time it will take to write. If that it unacceptably large, the user can 
-use another call to specify the lowest priority element that they want to actually write, 
-skipping the rest too save time (in this
-example the user will write all of the data). The data will then be stored
+will return a time estimate for this write operation.
+If this time estimate is too large, an API call can be made to 
+specify the lowest priority element to be written, skipping the rest in order to save time.
+The data will then be stored
 across the layers of the file system where the two highest priority buckets

### 2015-07-07 narrative/qos.tex
another reference

@@
 across multiple tiers so that a search still has to occur in order to find a resource on a particular 
-tier.
-
-Specifically, we will address the following research questions:\\
-  1) Can we make any guarantees in a scalable decentralized system?; \\
-  2) What quality of service metrics can be make guarantees on in a scalable
-    way?;\\
-  3) What kind of error rates will we have in our estimation and what error
-    rates are acceptable?;\\
-  4) How much will the qos scheduler, admission control, manager cost in
-    terms of resources and time?;\\
-  5) How do we make the scheduler scalable with the size of the system?
+tier.  Specifically, we will address the following research questions:
+\begin{itemize}
+  \item Can we make any guarantees in a scalable decentralized system?; 
+ \item  What quality of service metrics can be make guarantees on in a scalable
+    way?;
+  \item What kind of error rates will we have in our estimation and what error
+    rates are acceptable?;
+  \item How much will the qos scheduler, admission control, manager cost in
+    terms of resources and time?;
+  \item  How do we make the scheduler scalable with the size of the system?
+\end{itemize}
 
@@
 and later reverting back to the previous committed resource level for the rest.
-This can be automatically enforced via user- (or administrator-) specified \textit{policies}
-(section \ref{}).
+This can be automatically enforced via user- (or administrator-) specified \textit{policies}\S\ref{sec:data-description}.
 

### 2015-07-07 narrative/data-description.tex
editing section 1.3

@@
 the need to know where the data currently resides.
-This requires that when reading the data, the middleware will
-provide a time estimate of reading, and then allow the users to decide if
+At read time, the middleware will be required to 
+provide a time estimate, and then allow the users to decide if
 they will read less (e.g., only the highest priority sub-chunks)
 if the time estimate
-for reading exceeds their expectations. Existing Sirocco functionality may be
+for reading does not meet requirements. Existing Sirocco functionality may be
 sufficient for these purposes, but it must still be proven in practice.

### 2015-07-07 narrative/enabling_tech.tex

@@
 systems, data staging systems~\cite{zheng:2010:predata,docan:2010:dataspaces,oldfield:2006:nessie}, workflow
-systems~\cite{Dayal:2014:flexpath}, and even through the WAN network 
+systems~\cite{Dayal:2014:flexpath}, and even through the WAN network. 
 Part of this project's goals are to understand where the dividing line

### 2015-07-07 narrative/naming-service.tex
changes to the introduction, budget and add labels.

@@
 \subsection{Naming and Discovery Service}
-
+\label{sec:naming-discovery}
 \paragraph{Background:}
@@
 
-\item Milestone 1: Demonstrate a metadata service capable of serving both POSIX
+\item {\bf Milestone 1}: Demonstrate a metadata service capable of serving both POSIX
 clients and our clients, but without consideration for authentication,
@@
 
-\item Milestone 2: Demonstrate a time bounded search approach for finding data
+\item {\bf Milestone 2}: Demonstrate a time bounded search approach for finding data
 within the storage system to identify data and the current location.  This will
@@
 
-\item Milestone 3: Demonstrate incorporating A\&A and show scalability under
+\item {\bf Milestone 3:} Demonstrate incorporating A\&A and show scalability under
 both application load as well as storage system pressures showing that we can

### 2015-07-07 narrative/data-reduction.tex
minor changes to help Tammy work with the proposal
editing 1.3.2

@@
 generated. In future, it will be vital to take advantage of a priori
-information when a) writing and reading from the SSIO layer in order to gain
+information when 1) writing and reading from the SSIO layer in order to gain
 performance and predictability, and to prioritize I/O so that the data that
-is most useful for end-user can be written in the time available, and b) in
-performing {\em in situ} analysis before storing the information.
+is most useful for end-user can be written in the time available, and 2) in
+performing {\em in situ} operations and analysis before storing the information.
 
@@
 operations versus storage-based operations. 
-
 We will assemble a suite of general techniques for refactoring of data that

### 2015-07-07 narrative/data-reduction.tex
editing 1.3.2

@@
 
-In the case of C. S. Chang described earlier, in common with most other
+In the case of C. S. Chang described earlier, and common with most other
 applications scientists with whom we collaborate, information is provided on
-which data should be sent over to our SSIO layer in order that the best
-possible information is available for the subsequent analysis. The decision as
-how this should be accomplished generally involves data refactoring
-(re-organizing and reducing the data).
+which data should be sent over to the SSIO layer so that the best, and most
+important information is available for subsequent analysis.
+This is generally accomplished using refactoring, which includes re-organization and reductions of the data.
 

### 2015-07-07 narrative/placement.tex
restructured the entire life cycle text. Purging is still incomplete. Someone need to sanity check the migration portions also. We also need milestones for each of the subareas here

@@
 potential insights from the simulations can be realized. 
-The overall data life cycle consists of three phases. Data starts
+The overall data life cycle consists of four phases. Data starts
 when it is generated by a simulation or through a data acquisition system,
@@
 
-% In the final phase the data is purged from the system altogether. The
-% decision to purge is made by the system based on operational metrics from
-% system administrators. This phase optimizes the space available in the
-% system to prevent out of space concerns for applications.
+In the final phase the data is purged from the system altogether. The
+decision to purge is made by the system based on operational metrics from
+system administrators. This phase optimizes the space available in the
+system to prevent out of space concerns for applications.
 
@@
 
- % Next we provide details on the challenges and proposed research through each
+%Next we provide details on the challenges and proposed research through each
 % phase of the data life cycle.
@@
 layer, the decision about which object to evict may take the utility into
-account.
-
-\paragraph{Autonomic in-transit data management:} 
-Efficiently placing data objects
-vertically across multiple memory hierarchy layers and storage levels as
-well as horizontally across different storage nodes requires balancing
-competing objectives. We will explore autonomic data
-management strategies that can evaluate utility/cost tradeoffs and
-appropriately place/move data objects at runtime. The autonomic approach
-will leverage both user hints and information gathered at runtime (e.g.,
-runtime data access history, network topology, etc.) to place data object.
+account. We will also explore the balancing of the data across the memory
+hierarchy autonomically. The autonomic approach will leverage both user
+hints and information gathered at runtime (e.g., runtime data access
+history, network topology, etc.) to place data object.
 
@@
 % resources.
+\subsubsection{Migration}
+\label{sec:migration}
 
-\subsubsection{Managing Data on Multiple Tiers}
+\paragraph{Background:} 
+Once the system has completed the initial data placement with minimal
+overhead to the application, it is up to the storage system to manage the
+migration of data to other storage servers and across the hierarchy. An
+important reserach aspect here is how a liberated Hierarchical Storage
+Manager (HSM) designed for online use, such as Sirocco, might be employed in
+providing support for the migration of data. Sirocco autonomously groups the
+distributed set of media and service nodes into like groups and manages the
+movement of data between these groups. The grouping is managed using
+attributes such as latency to first byte, bandwidth to and from the media, and
+some function representing the resilience capabilities of the attached media.
+In addition to these static attributes, when grouping servers into
+pseudo-tiers, Sirocco also may employ more dynamic attributes such as CPU load
+and the rate-of-change of the media use.
+
+Grouping into pseudo-tiers is not a formal, nor rigid, concept within
+Sirocco, at present. In classic HSM the tiering is often a formally
+expressed concept, embodied both in the architecture and the implementation.
+In Sirocco, though, it is not directly expressed. Instead, a client or
+server that attempts to make a decision relative to migration and staging
+uses these attributes to constrain the choice of candidates. For instance, a
+server acting as an off-node writeback cache for a compute node might only
+have access to volatile dynamic RAM for use as storage media and a client
+might deposit data that is marked for persistent media as a resilience
+constraint. At some point, then, such a server must copy or move the data to
+another Sirocco server with media matching, or exceeding, the resilience
+constraint. At that point in time, all servers managing appropriate
+persistent media are candidates but second-order attributes, such as latency
+and bandwidth of a candidate's managed media would tend to dominate the
+ranking. Thus, the concept of a ``tier'' becomes an artifact of decision
+rather than a formal concept.
+
+Give these design decisions, our research in this project will study how
+migration can serve to manage storage resident data within the constraints
+defined by the application and the user.
+
+\paragraph{Migration Policies:}
+The policies that govern the migration of data across groups and across
+tiers have similar challenges as initial data placement. The migration
+policy must encapsulate, at a high level, the constraints defined by the
+application, but must also manage resources on each server for capacity and
+load reasons. We will explore policy formulations that allow \Sir to move
+data within a tier to continue to meet service objectives. Another component
+of migration is the temporal function that determines the utility of data
+through its life time on storage. If they utility decays over time, for
+example, the initial constraints of data (such as resilience requirements)
+will be modified and allow more leeway in how the system handles migration.
+However, as data is widely distributed the scalability of this approach must
+be investigated. 
+We will explore how the utility function can be used to modify migration
+parameters in a scalable manner, without requiring a centralized authority
+to make decisions.
+
+\paragraph{Bounded estimates for lookup:}
+One challenge for storage systems with continous decentralized migration is
+the timeliness of lookup when applications issue read calls. This is due to
+how no central directory exists for where data is at anytime. There are
+solutions to this problem that utilize lookup tables which are updated
+whenever a server initiates a migration process. However, these solutions
+are not scalable. In order to have predictable performance the lookup
+process must complete in a bounded time. Another solution here is to use a
+distributed hashing function to determine exactly where data is placed,
+reducing lookup to an $O(1)$ operation. We explore the discovery of data in
+\S~\ref{sec:naming-discovery}. To support this discovery process, we will
+investiagte how the migration of data can be performed in a way that enables
+bounded estimates about time to first byte for a potential read operation.
+
+% We will explore the use of a  hashing function
+% that will leverage determinism to initiate the lookup process. If at most
+% $k$ migration steps can happen 
+% Our research intent involves short term and long term activities. In the short
+% term we will modify the relevant clients to provide hit and miss related data
+% and associated activities and explore an enhancement to our storage servers
+% that would provide some, hopefully tightly, bounded estimate about time to
+% first byte for a potential read operation.
+
+% Further complication arises when a Sirocco server is forced to make choices for
+% capacity reasons. The Sirocco architecture dictates that a server function as a
+% victim cache. As media managed by a server fills the server will begin to
+% manage for capacity. In order to do so, currently, it need only verify or make
+% a copy of the data on another server that meets the resiliency constraint of
+% the held data prior to removing the local copy and in the current
+% implementation the server makes no attempt either to remember it ever had the
+% data or to notify others about what it is doing.  others that it has removed
+% the local copy or remember that it has
+
+% Taken together then, the above at least complicates the higher level research
+% questions discussed in this proposal, potentially even renders some
+% unachievable since a few involve determinism. We are forced, then, to consider
+% either or both mitigating management policies within and between Sirocco
+% servers as well as explicitly allowing clients to participate in capacity
+% management decisions, for a time at least, on the relevant Sirocco servers.
+
+% \paragraph{Migration Policies:}
+% In the long term we hope that an
+% augmented client-server API providing usage hints will allow us to modify local
+% server policy algorithms to more effectively manage it's media and communicate
+% to other servers holding copies of the relevant data how they might best manage
+% theirs. In the event that this approach proves ineffective at meeting the
+% overall approach described in this proposal we will explore modifying Sirocco
+% servers to provide notifications about locations of the relevant data as they
+% migrate and stage as well as supporting client-directed pinning of location,
+% for a period at least. Client-directed pinning could introduce potentially
+% serious negative impacts to Sirocco's overarching goals though, since the
+% fundamental design of Sirocco assumes much about opaque and unrestricted data
+% movement, so this is contemplated only as a last resort.
+
+\subsubsection{Making data avilable on multiple tiers}
 \label{sec:manag-data-mult}
 
-\paragraph{Background:} There is a spectrum of approaches on how
-to leverage a memory and storage hierarchy: On one end of the
-spectrum is to copy or move data from one tier to another, the
-closer the tier to the hierarchy's top, the faster the access to
-the data it contains. On the other end of the spectrum is to use
-higher tiers exclusively for auxiliary data that to some extent
-represents the actual data on the lowest tier. Examples of auxiliary
-data are views, indices, lossy compressions, lower resolution data,
-and summary data (sometimes also known as metadata). There are a
+\paragraph{Background:} Providing data to users for knowledge discovery is
+one of the primary purposes of our storage system design. There is
+There is a
+spectrum of approaches on how
+this can be accomplished: On one end the storage system can aim to maintain
+full resolution data in the highest tiers of the storage hierarchy. This
+approach is, obviously, limited due to the disparity between size of data
+and available space in the higher tiers. Another approach is to utilize the
+techniques described in \ref{sec:data-refactor} to create auxiliary views of
+the data that, with user defined error, represent the data, and place these
+smaller data sets on the higher tiers for faster access.
+
+There are a
 number of drivers that push approaches to one or the other extreme:
@@
 ``enriched'' (e.g. turned into a higher resolution) if there is
-value in doing so.
+value in doing so. We will explore how the storage system can perform these
+operations to meet the service level objectives for each use case, while
+optimizing the overall utility from the system perpective.
 
@@
 
-\subsubsection{Migration}
-\label{sec:migration}
+\subsubsection{Purging data from the system}
+\label{sec:purging-data}
+(only outline since I have to go)
+1. Purging is important from the perspective of preventing uncontrolled
+growth of data on a system
+2. Purging is hard because no one wants to lose data and we need to explore
+the policies that govern it
+3. If we have a good utility function and a good way to refactor the data we
+could purge the ``less important'' pieces and maintain a high level of
+knowledge without requiring very high amount of storage
+3. Purging needs to happen in a decentralized way but global utility has to
+be maximized - thats a pretty hard problem
 
-\paragraph{Background:} One important research question in this proposal is
-whether and how a liberated Hierarchical Storage Manager designed for online
-use, such as Sirocco, might be employed in supporting the higher level
-concepts discussed previously. Conceptually, in this effort, data is
-accessed by the middleware I/O libraries at various tiers in the lower-level
-storage hierarchy. While Sirocco does autonomously group the distributed set
-of media into like groups and manage movement of data between these groups
-the various motivations to do so include both the traditional ones found in
-classical HSM (Hierarchical Storage Management) as well as the novel.
+% \subsubsection{Preliminary work and results} 
+% \label{sec:prework}
 
-Sirocco groups like media into aggregated volumes on a service node, and like
-service nodes with like media into pseudo-tiers. It accomplishes both using
-attributes such as latency to first byte, bandwidth to and from the media, and
-some function representing the resilience capabilities of the attached media.
-In addition to these static attributes, when grouping servers into
-pseudo-tiers, Sirocco also may employ more dynamic attributes such as CPU load
-and the rate-of-change of the media use.
-
-Grouping into pseudo-tiers is not a formal, nor rigid, concept within Sirocco,
-at present. In classic HSM the tiering is often a formally expressed concept,
-embodied both in the architecture and the implementation. In Sirocco, though,
-it is not directly expressed. Instead, a client or server that attempts to make
-a decision relative to migration and staging uses the attributes discussed
-above to constrain the choice of candidates. For instance, a server acting as
-an off-node writeback cache for a compute node might only have access to
-volatile dynamic RAM for use as storage media and a client might deposit data
-that is marked for persistent media as a resilience constraint. At some point,
-then, such a server must copy or move the data to another Sirocco server with
-media matching, or exceeding, the resilience constraint. At that point in time,
-all servers managing appropriate persistent media are candidates but
-second-order attributes, such as latency and bandwidth of a candidate's managed
-media would tend to dominate the ranking. Thus, the concept of a ``tier''
-becomes an artifact of decision rather than a formal concept.
-
-Further complication arises when a Sirocco server is forced to make choices for
-capacity reasons. The Sirocco architecture dictates that a server function as a
-victim cache. As media managed by a server fills the server will begin to
-manage for capacity. In order to do so, currently, it need only verify or make
-a copy of the data on another server that meets the resiliency constraint of
-the held data prior to removing the local copy and in the current
-implementation the server makes no attempt either to remember it ever had the
-data or to notify others about what it is doing.  others that it has removed
-the local copy or remember that it has
-
-Taken together then, the above at least complicates the higher level research
-questions discussed in this proposal, potentially even renders some
-unachievable since a few involve determinism. We are forced, then, to consider
-either or both mitigating management policies within and between Sirocco
-servers as well as explicitly allowing clients to participate in capacity
-management decisions, for a time at least, on the relevant Sirocco servers.
-
-\paragraph{Bounded estimates for access:}
-Our research intent involves short term and long term activities. In the short
-term we will modify the relevant clients to provide hit and miss related data
-and associated activities and explore an enhancement to our storage servers
-that would provide some, hopefully tightly, bounded estimate about time to
-first byte for a potential read operation.
-
-\paragraph{Migration Policies:}
-In the long term we hope that an
-augmented client-server API providing usage hints will allow us to modify local
-server policy algorithms to more effectively manage it's media and communicate
-to other servers holding copies of the relevant data how they might best manage
-theirs. In the event that this approach proves ineffective at meeting the
-overall approach described in this proposal we will explore modifying Sirocco
-servers to provide notifications about locations of the relevant data as they
-migrate and stage as well as supporting client-directed pinning of location,
-for a period at least. Client-directed pinning could introduce potentially
-serious negative impacts to Sirocco's overarching goals though, since the
-fundamental design of Sirocco assumes much about opaque and unrestricted data
-movement, so this is contemplated only as a last resort.
-
-\subsubsection{Preliminary work and results} 
-\label{sec:prework}
-
-Our recent work \cite{tongipdps15}
-explored a two-tiered staging method that spans both DRAM and solid state
-disks (SSD). It allows us to support both code coupling and data management
-for data intensive simulation workflows in a loosely coupled manner. In
-addition, our application-aware adaptive data placement has demonstrated the
-effectiveness of using user provided access information as ``hints'', and
-runtime data access pattern history to improve data access efficiency and
-overall end-to-end execution. Besides, additional data placement
-optimization \cite{qiansc15} that leverages physical network topology along
-with data access patterns has also demonstrated lower data access costs and
-good scalability. Our work have been deployed and proven with real
-applications, e.g., coupled combustion simulations (DNS-LES) that are part
-of the ExaCT Exascale Co-design Center and coupled Fusion simulations
-(XGC0-XGCa) that are part of the EPSI SciDAC on current high-end systems
-such Titan at ORNL.
+% Our recent work \cite{tongipdps15}
+% explored a two-tiered staging method that spans both DRAM and solid state
+% disks (SSD). It allows us to support both code coupling and data management
+% for data intensive simulation workflows in a loosely coupled manner. In
+% addition, our application-aware adaptive data placement has demonstrated the
+% effectiveness of using user provided access information as ``hints'', and
+% runtime data access pattern history to improve data access efficiency and
+% overall end-to-end execution. Besides, additional data placement
+% optimization \cite{qiansc15} that leverages physical network topology along
+% with data access patterns has also demonstrated lower data access costs and
+% good scalability. Our work have been deployed and proven with real
+% applications, e.g., coupled combustion simulations (DNS-LES) that are part
+% of the ExaCT Exascale Co-design Center and coupled Fusion simulations
+% (XGC0-XGCa) that are part of the EPSI SciDAC on current high-end systems
+% such Titan at ORNL.
 

### 2015-07-07 narrative/data-description.tex

@@
 users can tell us that the data represents: 1)~phase-space (e.g., the particles
-in our example), 2)~spatial-changes (e.g., the mesh), 3)~space-time (e.g.,
+in our example), 2)~spatial-changes (e.g., the mesh), and 3)~space-time (e.g.,
 the fields on the mesh). We need this information to select
@@
 organized together along with only F1. When the users wants to visualize data
-from the field, F1, they will need only M1 and M2 would be additional
+from the field, F1, they will need only M1, and M2 would be additional
 information. Likewise, if the user wants to see F1 and F2, then they will need
 the mesh from M1 and M2. This requires that the time to access all of this
-data is approximately the same requiring similar access times for all of the
-data.  We will investigate what choices the SSIO system must make in order to
+data is approximately the same.
+We will investigate what choices the SSIO system must make in order to
 keep these pieces ``together'', from a performance perspective. Can the storage
 system move everything to a lower layer when all of it doesn't fit or is there
-other, less tightly related data can be evicted to preserve the collective
+other, less tightly related data that can be evicted to preserve the collective
 retrieval performance?  We need to investigate the potential problems when this

### 2015-07-07 narrative/deliverables.tex

@@
 (Rutgers/ORNL)
-\item Investigate the trade-off between ``filing'' and ``piling'' data. (UCSC/Sandia)
-\item Demonstrate a metadata service capable of serving both POSIX clients and our clients,
-but without consideration for authentication, authorization, or data migration. (Sandia/UCSC)
+\item Investigate the trade-off between ``filing'' and ``piling'' data. 
+Demonstrate a time bounded search approach for finding data within the storage system
+to identify data and the current location. (UCSC/Sandia)
+\item Demonstrate a metadata service capable of serving both POSIX clients and our clients. (Sandia/UCSC)
 \end{itemize}
@@
 about completion timing information. (ORNL)
+\item Evaluate the effectiveness and overhead to re-organize data and use the multi-layer approach 
+to staging. (Brown/ORNL)
 \item Research autonomic data management strategies that
 can evaluate utility/cost tradeoff, and appropriately place/move data objects at runtime. 
-\item Demonstrate a time bounded search approach for finding data within the storage system
-to identify data and the current location. This will update the cached information in the metadata
-service to short-circuit future requests for the data presuming that it does not migrate again soon. 
-\item Extend Sirocco to enforce data-centric metrics that help guide the QoS decisions.
+Research runtime tracking and estimation. (Rutgers/ORNL)
+\item Extend Sirocco to enforce data-centric metrics that help guide the QoS decisions. (Sandia)
 (Sandia/UCSC)
@@
 \begin{itemize}
-\item Work with DOE applicationexsiting ADIOS-enabled applications, framework using many of the applications to which we have access, including XGC1, GTC, and SPECFM3D
+\item Work with DOE applications, particularly XGC1, GTC, and SPECFM3D, to assess 
+the new I/O interface including both APIs and hints. (ORNL)
+\item Research techniques that can enhance migration without making the eventual 
+lookup of data unbounded, and manage each tier in a scalable fashion. (ORNL)
+\item Demonstrate incorporating A\&A and show scalability under both application load as
+well as storage system pressures showing that we can tolerate both loads and maintain quality of
+service. (Sandia)
 %\item storage APIs to manipulate storing and retrieving data to specific tiers

### 2015-07-07 other-support/curry.tex

\documentclass[10pt]{article}
\usepackage[english]{babel}
\usepackage{cite}
\usepackage{url}
\usepackage{times}
\usepackage{graphicx}
\usepackage[sectionbib]{bibunits}
\pagestyle{empty}

% Set page geometry
\usepackage{geometry}
\geometry{verbose,letterpaper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

\begin{document}

\begin{center}
{\Large Current and Pending Support for Matthew L. Curry}
\end{center}

\section*{Current Support}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \= Scalable Data Management\\
Principal Investigator      \>: \> Gerald F. Lofstead\\
Source of Support           \>: \> NNSA/Advanced Simulation and Computing\\
Period Covered              \>: \> 10/2014-9/2015\\
Commitment                  \>: \> 0.05 FTE\\
Description                 \>: \> Research and development activities to improve NNSA application performance.\\
Overlap                     \>: \> None\\
\end{tabbing}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \= Tajique\\
Principal Investigator      \>: \> Anthony Vestal\\
Source of Support           \>: \> OGA\\
Period Covered              \>: \> 10/2014-9/2015\\
Commitment                  \>: \> 0.1 FTE\\
Description                 \>: \> Characterize distributed file system behavior\\
Overlap                     \>: \> None\\
\end{tabbing}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \= Sirocco\\
Principal Investigator      \>: \> Lee Ward\\
Source of Support           \>: \> NNSA/Advanced Simulation and Computing\\
Period Covered              \>: \> 10/2014-9/2015\\
Commitment                  \>: \> 0.85 FTE\\
Description                 \>: \> Development activities for Sirocco, a novel storage systemfor Exascale computers.\\
Overlap                     \>: \> None\\
\end{tabbing}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Pending Proposals (in addition to this)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Pending Proposals (in addition to this)}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \= Managing Node-Based NVMs for HPC Clusters\\
Principal Investigator      \>: \> Ethan Miller, Darrell Long, and Lee Ward\\
Source of Support           \>: \> ASCR\\
Award Amount                \>: \> Approximately \$1,500,000\\
Period Covered              \>: \> 10/2015-9/2018\\
Commitment                  \>: \> 0.1 FTE\\
Description                 \>: \> Develop local and global mechanisms for managing non-volatile memories integrated \\
 \> \>into nodes of supercomputers.\\
Overlap                     \>: \> None\\
\end{tabbing}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \= Modeling Impacts of Resilience Architectures for Extreme-Scale Storage Systems\\
Principal Investigator      \>: \> Remzi Arpaci-Dusseau, Matthew L. Curry, and Nathan DeBardeleben\\
Source of Support           \>: \> ASCR\\
Award Amount                \>: \> Approximately \$1,500,000\\
Period Covered              \>: \> 10/2015-9/2018\\
Commitment                  \>: \> 0.25 FTE\\
Description                 \>: \> Project to model multi-level resilience mechanisms for large-scale storage systems\\
Overlap                     \>: \> None\\
\end{tabbing}

\end{document}

### 2015-07-08 narrative/placement.tex
work in progress....

@@
 potential insights from the simulations can be realized. 
-The overall data life cycle consists of four phases. Data starts
-when it is generated by a simulation or through a data acquisition system,
-in the case of experimental and observational data (EOD). 
-In our target heterogenous multi-tier storage platform, the system must
-decide where to place the output data at this time, and how to partition and
-reorganize the data to optimize this initial placement. The key optimization
-metrics during this phase is the percieved I/O overhead for the application.
+The overall data life cycle can be viewed as consisting of four key phases. In the first phase data 
+is generated by a simulation or through a data acquisition system,
+in the case of experimental and observational data (EOD). This data has 
+to be appropriately initially placed (and possibly partitioned) within the heterogenous 
+multi-tier storage to optimize application I/O overheads.
+%In our proposed heterogenous multi-tier storage platform, the system must
+%decide where to place the output data at this time, and how to partition and
+%reorganize the data to optimize this initial placement. The key optimization
+%metrics during this phase is the perceived I/O overhead for the application.
 
-In the second phase of the data life cycle, the system takes over the
-management of the data and performs the necessary steps to meet service
-level objectives (SLOs). These steps include migration, reorganization, and
-reductions for the data. During this phase the system optimizes for resource
+In the second phase of the data life cycle, the data is managed by the to 
+meet application requirements and constraints. This includes migration, 
+reorganization and reduction of the data, etc., to optimize resource
 utilization, interference with running applications, and availability and
-resiliency objectives. Data can be moved into archival storage (tape or cold
-storage) during this phase.
+resiliency objectives. 
+%Data can be moved into archival storage (tape or cold storage) during this phase. 
+%
+%In the second phase of the data life cycle the system takes over the
+%management of the data and performs the necessary steps to meet service
+%level objectives (SLOs). These steps include migration, reorganization, and
+%reductions for the data. During this phase the system optimizes for resource
+%utilization, interference with running applications, and availability and
+%resiliency objectives. Data can be moved into archival storage (tape or cold
+%storage) during this phase.
 
@@
 
-In the final phase the data is purged from the system altogether. The
-decision to purge is made by the system based on operational metrics from
-system administrators. This phase optimizes the space available in the
-system to prevent out of space concerns for applications.
+In the final phase the data is purged from the system. The
+decision to purge may be made by the system based on operational metrics from
+system administrators or by the user. This phase optimizes the space available in the
+system. % to prevent out of space concerns for applications.
 
-Our proposed research explores the management of this entire life cycle
-given our thesis that more knowledge, and more specific knowledge about the
-data, can allow the management of data to be optimized with reduced cost.
-Thus, we will explore the following high level research questions: (1) What methods can
-be utilized in the initial placement of data that do not impose expensive
-costs on discovery, consumption and migration? (2) How can the placement of
-data be optimized for the multi-tiered hierarchy expected in exascale
-systems? (3) How can we migrate data throughout the layers of storage
-without making the discovery process unbounded? (4) How does the additional
-knowledge aid in preparing the data for reading in a timely manner? (5) How
-do we make the decision on purging using this additional knowledge?
+Our proposed research will address the management of data across the entire life cycle 
+in an application driven and system aware manner. 
+%given our thesis that more knowledge, and more specific knowledge about the
+%data, can allow the management of data to be optimized with reduced cost.
+Some of the high level research questions that we will explore are as follows: 
+(1) How can we initially place data so that it can be discovered and consumed efficiently?
+%What methods can be utilized in the initial placement of data that do not impose expensive
+%costs on discovery, consumption and migration? 
+(2) How can the placement and migration of data across a multi-tiered storage hierarchy be 
+optimized at runtime, both from the application and system perspective?
+%(3) How can we migrate data throughout the layers of storage
+%without making the discovery process unbounded? 
+(4) How can knowledge about the application used to better prepare the data for consumption? 
+(5) When and how do we make the decision to purge data?  %using this additional knowledge?
 

### 2015-07-08 narrative/data-reduction.tex
editing 1.3.2

@@
 
-The classical workflow whereby the entire dataset is written to storage for
-later analysis is no longer viable owing to the sheer amount of data being
-generated. In future, it will be vital to take advantage of a priori
+The classical workflow where the entire dataset is written to storage for
+later analysis is no longer viable simply because the amount of generated data is too large.
+In future, it will be vital to take advantage of a priori
 information when 1) writing and reading from the SSIO layer in order to gain
 performance and predictability, and to prioritize I/O so that the data that
-is most useful for end-user can be written in the time available, and 2) in
+is most useful for end-user can be written in the time available, and 2)
 performing {\em in situ} operations and analysis before storing the information.
@@
 There are many possibilities for refactoring and the best choice will generally
-vary from case to case but, once settled, will not change from run to run. It
+vary from case to case but, once settled, will typically not change from run to run. It
 will be important to understand when the time and resources required to

### 2015-07-08 narrative/data-reduction.tex
editing 1.3.2

@@
 structure as feasible. This step is variously referred to as filtering,
-preconditioning, and preprocessing but in all cases, the objective is to
+preconditioning, and preprocessing. But in all cases, the objective is to
 represent the data in a form that becomes more amenable to compression.  (2)
@@
 removal of redundancy is generally minor. The main scope for achieving
-significant compression (a compression rate of two or better say) therefore
+significant compression (e.g. a compression rate of two) therefore
 lies in the pre-processing step.

### 2015-07-08 narrative/qos.tex
editing....

@@
 \paragraph{Background:} There is an important connection between finding the resources
-that can fulfill a particular request and the performance predictability of that request, and this demands
-careful resource management strategies that can deal with placing and retrieving a large number of data 
-chunks spread across the entire storage hierarchy. 
+that can fulfill a particular request and the performance predictability of that request. As such, this demands
+careful resource management strategies that can deal with placement and retrieval of 
+a large number of data chunks spread across the entire storage hierarchy. 
 %This connection is particularly important for data retrieval 
@@
 tier.  Specifically, we will address the following research questions:
-\begin{itemize}
+\begin{tightItemize}
   \item Can we make any guarantees in a scalable decentralized system?; 
@@
   \item  How do we make the scheduler scalable with the size of the system?
-\end{itemize}
+\end{tightItemize}
 
@@
 
-\begin{itemize}
+\begin{tightItemize}
 
@@
 
-\end{itemize}
+\end{tightItemize}
 

### 2015-07-08 narrative/placement.tex
editing 1.3.2
Work in progress

@@
 
-In the second phase of the data life cycle, the data is managed by the to 
+In the second phase of the data life cycle, the data is managed by the system to 
 meet application requirements and constraints. This includes migration, 
@@
 
-Our proposed research will address the management of data across the entire life cycle 
-in an application driven and system aware manner. 
+A key requirement is application-driven runtime mechanisms for dynamically managing 
+data placement across the layers of the distributed storage hierarchy throughout the 
+data lifecycle, coordinating data movement and data sharing between the components 
+of the application workflow, with the overarching goal of maximizing the relative utility 
+to the application as well as the system while reducing access costs. As notes before,
+the complexities of heterogeneous multi-level storage structures requires adaptive 
+placement policies are required to be implemented to optimally utilize storage resources 
+vertically (across deep memory hierarchies) and horizontally (across nodes within a 
+memory level) while accommodating dynamic application requirements and  transient 
+system states.
+
+Our proposed research will address data management challenges across the entire 
+data life cycle in an application driven and system aware manner. 
 %given our thesis that more knowledge, and more specific knowledge about the
@@
 
-\subsubsection{Initial Placement of Data}
+\subsubsection{Data Placement}
 \label{sec:init-plac-data}
@@
 \paragraph{Background:}
-When an application outputs data the storage and middleware layer needs to
-decide \textit{what} data is placed \textit{where} amongst the storage
-resource available to the system. This decision has far reaching impact
-throughout the life time of the data. Past work on data
-staging\cite{tongipdps15,qiansc15}\cite{docan2012dataspaces}\cite{abbasi2010datastager}
-in HPC systems has shown that different output techniques targetting
-different layers of the storage hierarchy can have a significant impact on
-the overhead observed by the application for I/O operations. The key
-requirements here are application-driven runtime mechanisms for dynamically, and
-deterministically, 
-managing data placement across the layers of the distributed storage
-hierarchy, coordinating data movement and data sharing between the
-components of the application workflow so as to maximize its utility to the
-application and reduce access costs. Our approach to data placement will
-leverage the increased set of knowledge about the data available to the
-middleware and storage system, potentially allowing for greater
-optimizations as compared to other storage solutions.
+When an application outputs data the storage and middleware layers needs to
+decide \textit{what} data is placed \textit{where} in the multi-level storage system. 
+This placement decision can have a significant impact on data management 
+throughout the life time of the data. For example, our past work on data 
+staging~\cite{tongipdps15,qiansc15,docan2012dataspaces,abbasi2010datastager} 
+on HPC systems with multi-level memory structures has shown that different 
+output techniques targeting different layers of the storage hierarchy can have a 
+significant impact on the overhead observed by the application for I/O operations.
 
-\paragraph{Application Hints:}
+%The key requirements here are application-driven runtime mechanisms for dynamically, 
+%and deterministically, managing data placement across the layers of the distributed storage 
+%hierarchy, coordinating data movement and data sharing between the components of the 
+%application workflow so as to maximize its utility to the application and reduce access costs.
+
+Our approach to data placement will leverage the increased set of knowledge about the data available to the middleware and storage system, potentially allowing for greater optimizations as compared to other storage solutions.
+
+\paragraph{Soliciting application {\em hints}:}
 As described in \S\ref{sec:data-refactor} one of the main components of our
@@
 
-\paragraph{Data Utility:} We will formulate the concept of utility (from an
+\paragraph{Defining data utility:} We will formulate the concept of utility (from an
 application perspective) that is assigned data objects to quantify the

### 2015-07-08 narrative/placement.tex
Work in progress

@@
 system. % to prevent out of space concerns for applications.
-
-A key requirement is application-driven runtime mechanisms for dynamically managing 
-data placement across the layers of the distributed storage hierarchy throughout the 
-data lifecycle, coordinating data movement and data sharing between the components 
-of the application workflow, with the overarching goal of maximizing the relative utility 
-to the application as well as the system while reducing access costs. As notes before,
-the complexities of heterogeneous multi-level storage structures requires adaptive 
-placement policies are required to be implemented to optimally utilize storage resources 
-vertically (across deep memory hierarchies) and horizontally (across nodes within a 
-memory level) while accommodating dynamic application requirements and  transient 
-system states.
 
@@
 
-Our approach to data placement will leverage the increased set of knowledge about the data available to the middleware and storage system, potentially allowing for greater optimizations as compared to other storage solutions.
+A key requirement is application-driven runtime mechanisms for dynamically managing 
+data placement across the layers of the distributed storage hierarchy throughout the 
+data lifecycle, coordinating data movement and data sharing between the components 
+of the application workflow, with the overarching goal of maximizing the relative utility 
+to the application as well as the system while reducing access costs. 
+%
+As noted before, the complexities of heterogeneous multi-level storage structures requires 
+adaptive placement policies are required to be implemented to optimally utilize storage 
+resources vertically (across deep memory hierarchies) and horizontally (across nodes 
+within a memory level) while accommodating dynamic application requirements and transient 
+system states. 
+%
+Our approach in the proposal is to increase knowledge about the data and its use within the 
+application, and leverage this knowledge to drive data placement and overall management.
+
+%Our approach to data placement will leverage the increased set of knowledge about the data 
+%available to the middleware and storage system, potentially allowing for greater optimizations 
+%as compared to other storage solutions.
 
@@
 proposed storage system is the ability to refactor and reduce data as it is
-generated, and to reorgnized and regenerate the data as it is accessed. We
+generated, and to reorganized and regenerate the data as it is accessed. We
 carry this principle into the placement and movement of data by allowing

### 2015-07-08 narrative/placement.tex
Work in progress...

@@
 system to minimize data movement and optimize the resources consumed by I/O.
+
+Application hints can provide additional information to the runtime based on the users' knowledge of the application workflow as well as on past experiences, for example, possible data access patterns, nature of regions of interest, and domain specific information.
+This application level information will be added to the specification of the workflow and can provide advisory inputs to efficiently help the runtime make smarter decisions about the placement and/or movement of data objects and the scheduling and execution of tasks, as well as to effectively manage trade-offs between for example, power consumption and time-to-solution.
 

### 2015-07-08 abstract.tex
small modifications

@@
 Our goal in this project is
-to address the associated I/O and storage challenges in the context of
+to address the associated Input/Output (I/O) and storage challenges in the context of
 current and emerging storage landscapes, and expedite insights into mission
-critical scientific processes; which is associated with {\bf theme two} of the FOA. To that end, we will build on the
-capabilities offered by ADIOS and DataSpaces at ORNL that provide I/O abstractions
+critical scientific processes; which is associated with {\bf theme two} of the Funding Opportunity Announcement. To that end, we will build on the
+capabilities offered by Adaptable I/O System (ADIOS) and DataSpaces at ORNL that provide I/O abstractions
 and services, the Sirocco peer-to-peer file system at
@@
 the available system resources.  
-We will test our prototypes  on current and future DOE system with many of today's applications, including the
+We will test our prototypes  on current and future Department of Energy (DOE) system with many of today's applications, including the
 XGC1, GTC, QMCPack, and SpecFM3D simulations. 

### 2015-07-08 narrative/qos.tex
temporarily removed "research questions" from QoS as they don't match the text.

@@
 across multiple tiers so that a search still has to occur in order to find a resource on a particular 
-tier.  Specifically, we will address the following research questions:
-\begin{tightItemize}
-  \item Can we make any guarantees in a scalable decentralized system?; 
- \item  What quality of service metrics can be make guarantees on in a scalable
-    way?;
-  \item What kind of error rates will we have in our estimation and what error
-    rates are acceptable?;
-  \item How much will the qos scheduler, admission control, manager cost in
-    terms of resources and time?;
-  \item  How do we make the scheduler scalable with the size of the system?
-\end{tightItemize}
+tier.
+
+%Specifically, we will address the following research questions:
+%\begin{tightItemize}
+%  \item Can we make any guarantees in a scalable decentralized system?; 
+% \item  What quality of service metrics can be make guarantees on in a scalable
+%    way?;
+%  \item What kind of error rates will we have in our estimation and what error
+%    rates are acceptable?;
+%  \item How much will the qos scheduler, admission control, manager cost in
+%    terms of resources and time?;
+%  \item  How do we make the scheduler scalable with the size of the system?
+%\end{tightItemize}
 

### 2015-07-08 bios/ward.tex
Lee's bio

\documentclass[11pt]{nsfcv}

%\newcommand{\elm}{E.~L. Miller}
%\newcommand{\exm}{E.~Miller}
%\newcommand{\person}[5]{\item #2, #1: #3 #4 (#5);\xspace}

\begin{document}
\begin{center}
  {\bf\Large Lee Ward} \\
  {\bf\Large Principal Member of Technical Staff} \\
  Sandia National Laboratories \\
  Phone: +1 (505) 844-9545 \\
  Email: \verb+lee@sandia.gov+
\end{center}

%\section*{Education}
%\begin{yearlist}
%\item[1995] Ph.\,D. (computer science), University of California at Berkeley
%\item[1990] M.\,S. (computer science), University of California at Berkeley
%\item[1987] Sc.\,B. (computer science), Brown University
%\end{yearlist}

\section*{Professional Experience}
\begin{yearlist}

\item[2000--\hfill]
  Principal Member of Technical Staff, Sandia National Laboratories. Conduct research
  on high performance file systems and storage.

\end{yearlist}

\section*{Qualifications Summary}

\begin{itemtight}
\item Designed and wrote the eNFS filel system service for Sandia's CPlant clusters
\item Designed and wrote the libsysio virtual file system interface for Cray-XT class
  machines running the Catamount operating system
\item Received two patents for error correcting codes applicable to 3D stacked-silicon memories
\end{itemtight}

\section*{Selected Recent Publications}
\begin{publist}

\item
  Liu, Guoxin, Haiying Shen, and Lee Ward.
  ``An efficient and trustworthy P2P and social network integrated file sharing system.'' Computers, IEEE Transactions on 64, no. 1 (2015): 54-70.

\item
  Karakoyunlu, Cengiz, Dries Kimpe, Philip Carns, Kevin Harms, Robert Ross, and Lee Ward.
  ``Toward a unified object storage foundation for scalable storage systems.''
  In Cluster Computing (CLUSTER), 2013 IEEE International Conference on, pp. 1-8. IEEE, 2013.

\item
  Carns, Philip, Kevin Harms, Dries Kimpe, Justin M. Wozniak, Robert Ross, Lee Ward, Matthew Curry et al.
  ``A case for optimistic coordination in hpc storage systems.'' In High Performance Computing, Networking, Storage and Analysis (SCC), 2012 SC Companion:, pp. 48-53. IEEE, 2012.

\item
  Wildani, Avani, Ethan L. Miller, and Lee Ward.
  ``Efficiently identifying working sets in block i/o streams."
  In Proceedings of the 4th Annual International Conference on Systems and Storage, p. 5. ACM, 2011.

\item
  Curry, Matthew L., Anthony Skjellum, H. Lee Ward, and Ron Brightwell.
  ``Gibraltar: A Reed‐Solomon coding library for storage applications on programmable graphics processors.'' Concurrency and Computation: Practice and Experience 23, no. 18 (2011): 2477-2495.

\end{publist}

\newpage
\section*{Conflicts of Interest}

\subsection*{Collaborators and Co-authors (in past 48 months)}
\begin{itemtight}
\item Arpaci-Dusseau, Remzi  (University of Wisconsin - Madison): collaborator
\item Brooks, Richard (Clemson University): collaborator
\item Carns, Phil (Argonne National Laboratory): co-author; collaborator
\item Harms, Kevin (Argonne National Laboratory): co-author
\item Kelly, Suzanne (Sandia National Laboratories): co-author
\item Kimpe, Dries (Argonne National Laboratory): co-author
\item Reddy, Narasimha (Texas A\&M University): co-author; collaborator
\item Resnick, David (Sandia National Laboratory): co-author
\item Ross, Robert (Argonne National Laboratory): co-author; collaborator
\item Shen, Haiying (Clemson University): co-author; collaborator
\item Skjellum, Anthony (Auburn University): co-author; collaborator
\item Sun, Zhiwei (Data Direct Networks): co-author
\item Long, Darrell (University of California, Santa Cruz); co-author, collaborator
\item Miller, Ethan (University of California, Santa Cruz); co-author, collaborator
\end{itemtight}

\end{document}

### 2015-07-08 narrative/data-reduction.tex
spelling, etc. changes

@@
 Our previous
-work~\cite{lakshminarasimhan2011compressing,lakshminarasimhan2011compressing,%
+work~\cite{lakshminarasimhan2011compressing,%
 gong2012multi,jenkins2012byte,gong2013parlo,boyuka2014transparent,%
@@
 this back to a single-layer view.  Additional research questions which must be
-addressed: Will some types of data (PIC data, finite element data, ...) be
+addressed: Will some types of data (e.g. PIC data) be more
 conducive  for data-refactoring than others? How much semantic-domain knowledge
@@
 corresponding analysis routines which they use for their largest simulations.
-We will also work with Jeroen Tromp's SPECFM3D~\cite{SPECFEM3D} code, and the
+We will also work with the SPECFM3D~\cite{SPECFEM3D} code, and the
 fusion GTC~\cite{klasky2003grid} code.  In all cases we will measure their

### 2015-07-08 participant-table.tex
add my name to the personnel.

@@
 \setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Scott Klasky} (\emph{klasky@ornl.gov})}}}
-\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Gerald Lofstead} (\emph{glofstead@sandia.gov})}}}
+\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Gerald Lofstead} (\emph{gflofst@sandia.gov})}}}
 \setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Carlos Maltzahn} (\emph{carlosm@soe.ucsc.edu})}}}
@@
 \\
-\textbf{\ornl} & \makebox[\longestpientry][l]{\textbf{Scott Klasky} (\emph{klasky@ornl.gov})}\hspace*{1em}Hasan Abbasi, Kimmy Mu, Manish Parashar\\
-\textbf{\snl} & \makebox[\longestpientry][l]{\textbf{Gerald Lofstead} (\emph{glofstead@sandia.gov})}\hspace*{1em}Matthew Curry, Lee Ward\\
+\textbf{\ornl} & \makebox[\longestpientry][l]{\textbf{Scott Klasky} (\emph{klasky@ornl.gov})}\hspace*{1em}Hasan Abbasi, Gary Liu, Kimmy Mu, Manish Parashar\\
+\textbf{\snl} & \makebox[\longestpientry][l]{\textbf{Gerald Lofstead} (\emph{gflofst@sandia.gov})}\hspace*{1em}Matthew Curry, Lee Ward\\
 \textbf{\ucsc} & \makebox[\longestpientry][l]{\textbf{Carlos Maltzahn} (\emph{carlosm@soe.ucsc.edu})}\hspace*{1em}\

### 2015-07-08 abstract.tex

@@
 
-\vspace{-\belowdisplayskip}
+%\vspace{-\belowdisplayskip}
 \vspace{-\abovedisplayskip}
@@
 new research in the management, storage and retrieval over the long lifespan
-of the large amounts of data that will be produced.
+of the extreme amounts of data that will be produced.
 In this project we will

### 2015-07-08 coverpage.tex

@@
 \hspace*{10mm} \= \makebox[65mm][l]{\textbf{Applicant Institution:}} \= Oak Ridge National Laboratory \\
-\>  \>PO BOX 2008 MS6057 Oak Ridge TN 37831-6057\\
+\>  \>1 Bethel Valley Road\\
+\> \> Oak Ridge Tennessee 37831-6057\\
 \end{tabbing}
@@
 \vrule height0.4pt depth 0pt width 2.5in \\
-\> \> Scott KIasky,  Distinguished Scientist \\
+\> \> Scott Klasky,  Distinguished Scientist \\
 \> \> Oak Ridge National Laboratory \\
 \> \> Computer Science and Mathematics Division \\
-\> \> Oak Ridge, TN 37831-6057 \\
+\> \> Oak Ridge, Tennessee 37831-6057 \\
 \> \> (T) 865-241-9980 (Fax) 865-241-4811 \\
-\> \> klasky@ornl.gov
+\> \> Email: klasky@ornl.gov
 \end{tabbing}
@@
 \vrule height0.4pt depth 0pt width 2.5in \\
-\> \> Tammy Reed \\
+\> \> Barney Maccabe \\
 \> \> Oak Ridge National Laboratory \\
-\> \>Computer Science and Mathematics Division \\
+\> \> Computer Science and Mathematics Division \\
 \> \> Oak Ridge, TN 37831-6057 \\
-\> \> (T) 865-574-1936 (Fax) 865-241-4811\\
-\> \> tsr@ornl.gov
+\> \> (T) 865-241-6504 \\
+\> \> Email: maccabe@ornl.gov
 \end{tabbing}
@@
 \begin{tabbing}
-\hspace*{10mm} \= \makebox[80mm][l]{\textbf{Program Announcement Number:}} \= LAB 15-1338 \\
+\hspace*{10mm} \= \makebox[65mm][l]{\textbf{Program Announcement Number:}} \= LAB 15-1338\\
+\> \textbf{PAMS Tracking No.:}			    \> 0000219877\\
 \> \textbf{DOE/Office of Science Program Office:} \> Office of Advanced Scientific Computing Research\\
+\>									    \> Storage Systems and Input/Output for Extreme Scale Science\\
 \> \textbf{DOE/Office of Science Program Office}  \> \\
-\> \makebox[70mm][r]{\textbf{Technical Contact:}} \> Lucy Nowell\\
+\> \makebox[64mm][r]{\textbf{Technical Contact:}} \>  Lucy Nowell\\
+\\
+\> \textbf{Research Area:} \>Scalable Storage Software Infrastructure
+
 \end{tabbing}
 
-\medskip
-
 \newpage

### 2015-07-08 narrative/data-description.tex

@@
 negotiate with the system to determine what
-level of QofS is possible given
+level of QoS is possible given
 a prospective set of I/O operations and current state of the system, so that
@@
 Our research will address the following questions:
-1) How can users interact with the storage system efficiently?
-2) How can users express their intentions into actionable items which can take place in the middleware and/or storage layer?
+1) How can users interact with the storage system efficiently?,
+2) How can users express their intentions into actionable items which can take place in the middleware and/or storage layer?, and
 3) How can users annotate data with a utility metric?

### 2015-07-08 narrative/data-reduction.tex

@@
 
-Ideally, the scientists would perform the entire analysis stage {\em in situ}
+Ideally,  scientists would like to perform the entire analysis stage {\em in situ}
 to extract the relevant information and in so doing would effectively
@@
 effects amongst the data deluge. In particular, this means if data thinning or
-truncation is applied willy nilly then one is in danger of effectively throwing
+truncation is applied haphazardly then one is in danger of effectively throwing
 out the information one is seeking. It would seem that one cannot store the

### 2015-07-08 narrative/intro.tex

@@
 %problematic without new insights and solutions.
-Our proposal will be responsive to Theme 2: {\em Scalable Storage Software Infrastructure} of the call, as summarize in Table~\ref{tab:call}.
+Our proposal will be responsive to Theme 2: {\em Scalable Storage Software Infrastructure} of the call, as summarized in Table~\ref{tab:call}.
 
@@
 This forced the team to carefully scrutinize the large amount of data to be generated and save only the very most important pieces.
-Further, they had to break this greatly reduced data into two categories: first data they thought they would need to access while the simulation was running or shortly thereafter, and second data they would need after the whole campaign was finished.
+Further, they had to break this greatly reduced data into two categories: first data they thought they would need to access while the simulation was running or shortly thereafter, and second, data they would need after the whole campaign was finished.
 The former category would be stored on disk, and the later would be archived.
@@
 
-We  examine the research challenges from the perspective of
+We  examine the research challenges from the perspective of:
 %
@@
 and how much storage space is available during the simulations lifetime.
-Further, they would like Quality of Service assurances so that reactions
+Further, they would like Quality of Service (QoS) assurances so that reactions
 to varying bandwidth can be handled appropriately. Ideally, a set of
@@
 subsets and pieces.
-Research challenges include 1) an expressive and easy to use interface, 2) a model for the system supporting
+Research challenges include: 1) an expressive and easy to use interface, 2) a model for the system supporting
 autonomic decision for saving data using the user-defined importance and utility of data, and 3) the interface between
@@
 QoS and therefore, be able to provide fuzzy estimates of the time it will take to write/read data from
-the different layers to inform any migration/eviction decisions, an it
+the different layers to inform any migration/eviction decisions, and it
 is also critical that the storage system can manage (discover and retrieve) the data across all of the

### 2015-07-08 narrative/naming-service.tex
tight itemize, tight enumerate

@@
 tiers. While this approach to managing multi-tier storage works for whole files
-that fit on single logial devices, it is far from ideal for scientific
+that fit on single logical devices, it is far from ideal for scientific
 simulations.
@@
 between the user and the storage system extends into this area as well.
-Negotiating the data quality/retrieval time tradeoff reuqires additional
-metadata.  Providing this functionality requries the naming service be able to
+Negotiating the data quality/retrieval time tradeoff requires additional
+metadata.  Providing this functionality requires the naming service be able to
 both return a list of data versions and an indication of data quality. This
-will be combined with a retrieval time estimate for each verstion by the
+will be combined with a retrieval time estimate for each version by the
 middleware to determine what data version should be retrieved.  This estimate
-will be disucssed elsewhere in this proposal.
+will be discussed elsewhere in this proposal.
 
@@
 approach, we will enable more effective use of close/fast/small storage tiers
-and prioritize objects with high importance.  Traiditional HSM stores an entire
+and prioritize objects with high importance.  Traditional HSM stores an entire
 file on a tier making room if insufficient space is available.  With this shift
@@
 
-\begin{enumerate}
+\begin{tightEnumerate}
 
@@
 
-\end{enumerate}
+\end{tightEnumerate}

### 2015-07-08 narrative/qos.tex

@@
 
-\paragraph{Related Work:} On current HPC systems, absolute IO performance
+\paragraph{Related Work:} On current production HPC systems, absolute IO performance
 guarantees are not possible. Previous work \cite{lofstead:2010:io-variability,liu_hotstorage} 
-from this project team uses a client-assisted approach to mitigate IO variablity.
+from this project team uses a client-assisted approach to mitigate IO variablity at scale (100,000-core).
+Recently server side QoS scheduling has also been studied for HPC applications \cite{Dai:2014}.
+However, the key hurdle of scalability has not been solved.
 Existing efforts on scheduing resources~\cite{thapaliya:2014:io-cop,dorier:2014:calciom} attempt to offer
@@
 be informed by simple extensions to the job scheduler or more advanced
-management through new APIs.
+management through new APIs. On the other hand, storage QoS has been
+studied for enterprise applications and systems \cite {Gulati:2007,Gulati:2010,Gulati:2012}
+in virtualized environments. It is not straightforward to adapt these solutions
+to HPC storage systems because of the scalability and efficiency concerns, despite
+they achieved very good performance isolations.
 

### 2015-07-08 other-support/lofstead-cp.tex

--- 2015-06-22 other-support/lofstead-cp.tex
@@
 Period Covered              \>: \> 10/2014 -- 9/2014\\
-Investigator Commitment     \>: \> 0.2 FTE
+Investigator Commitment     \>: \> 0.2 FTE\\
+Research Performed          \>: \> Integrated Application Workflow infrastructure support\\
+Overlaps                    \>: \> Focused on IAWs in general. This proposal looks at NVM impacts.\\
 \end{tabbing}
@@
 Period Covered              \>: \> 10/2014 -- 9/2018\\
-Investigator Commitment     \>: \> 0.25 FTE
+Investigator Commitment     \>: \> 0.25 FTE\\
+Research Performed          \>: \> data warehouse for task-based programming system.\\
+Overlaps                    \>: \> None.\\
 \end{tabbing}
@@
 Period Covered              \>: \> 8/2013 -- 7/2016\\
-Investigator Commitment     \>: \> 0.35 FTE
+Investigator Commitment     \>: \> 0.35 FTE\\
+Research Performed          \>: \> Application composition operating system support.\\
+Overlaps                    \>: \> Example workflows that will be explored in this proposal.\\
 \end{tabbing}
@@
 Period Covered              \>: \> 8/2014 -- 7/2017\\
-Investigator Commitment     \>: \> 0.2 FTE
+Investigator Commitment     \>: \> 0.2 FTE\\
+Research Performed          \>: \> Develop generic workflow components for HPC\\
+Overlaps                    \>: \> Example workflows that will be explored in this proposal.\\
 \end{tabbing}
@@
 
-%\begin{tabbing}
-%Project Title \hspace{5.5em}\=: \= Code Coupling for Scalability and Robust Data Analytics\\
-%Source of Support           \>: \> Sandia National Laboratories LDRD \\
-%Award Amount                \>: \> \$600K/year to SNL\\
-%Period Covered              \>: \> 8/2014 -- 7/2017\\
-%Investigator Commitment     \>: \> 0.5 FTE
-%\end{tabbing}
+\begin{tabbing}
+Project Title \hspace{5.5em}\=: \= SCADS: Scalable Deep Storage Hierarchy for Exascale Computing\\
+Source of Support           \>: \> DOE Storage Systems and Input/Output for Extreme Scale Science\\
+Award Amount                \>: \> \$176/year to SNL\\
+Period Covered              \>: \> 10/2015 -- 9/2018\\
+Investigator Commitment     \>: \> 0.25 FTE\\
+Research Performed          \>: \> effective NVM placement and capacity for best end-to-end science throughput\\
+Overlaps                    \>: \> None.\\
+\end{tabbing}
 

### 2015-07-08 bios/curry-src.tex
Renamed curry.tex, to prevent broken build due to smart include.

--- 2015-06-26 bios/curry.tex
@@
 \usepackage{hyperref}
+\usepackage{enumitem}
+\usepackage{times}
+\setlist{nolistsep} % Makes itemize tighter
 % Max 3 pages
@@
 \subsection{Education and Training}
+%{\bf Undergraduate, graduate and postdoctoral training, provide institution, major/area, degree and year.}
 \begin{itemize}
-\item Ph.D., Computer and Information Sciences\\ University of Alabama at Birmingham, December 2010
-\item M.S., Computer and Information Sciences\\ University of Alabama at Birmingham, May 2008
-\item B.S., Computer and Information Sciences\\ University of Alabama at Birmingham, December 2004
+\item Ph.D., Computer and Information Sciences, University of Alabama at Birmingham, December 2010.
+\item M.S., Computer and Information Sciences, University of Alabama at Birmingham, May 2008.
+\item B.S., Computer and Information Sciences, University of Alabama at Birmingham, December 2004.
 \end{itemize}
@@
 \subsection{Research and Professional Experience}
+%{\bf Beginning with the current position list, in re- verse chronological order, professional/academic positions with a brief description of responsibilities.}
 \begin{itemize}
 \item {\bf Senior Member of Technical Staff, Scalable System Software,
-  Sandia National Laboratories}, 2010-Present.\\ Performing research
+  Sandia National Laboratories}, 2010-Present. Performing research
   and development activities associated with high-performance storage
@@
   Sciences, University of Alabama at Birmingham}, January 2009 - May
-  2009.\\ Supported research software development activities for NSF
+  2009. Supported research software development activities for NSF
   Award CNS-0821497, ``MRI: Development of a GPU-Enabled Integrated
@@
 \item {\bf Student Intern, Sandia National Laboratories}, Summers
-  2005-2008 and May 2009-December 2010.\\ Performed research and
+  2005-2008 and May 2009-December 2010. Performed research and
   development activities associated with high-performance storage and
@@
   spring and fall semesters beginning January 2005 and ending May
-  2009.\\ Performed research, taught courses (Operating Systems,
+  2009. Performed research, taught courses (Operating Systems,
   Introduction to C++), gave guest lectures (Parallel Computing,
@@
 \subsection{Publications}
+%{\bf Provide a list of up to 10 publications most closely related to the proposed project.
+For each publication, identify the names of all authors (in the same sequence in which they appear in the publication), the article title, book or journal title, volume number, page numbers, year of publication, and website address if available electroni- cally.
+Patents, copyrights and software systems developed may be provided in addition to or substituted for publications.
+.
+An abbreviated style such as the Physical Re- view Letters (PRL) convention for citations (list only the first author) may be used for publications with more than 10 authors.}
 
@@
 \subsection{Synergistic Activities}
+%{\bf List no more than 5 professional and scholarly activities related to the effort proposed.}
 \begin{itemize}
@@
 \subsection{Identification of Potential Conflicts of Interest or Bias in Selection of Reviewers}
+% This part can extend past the three pages
+%\subsubsection{Collaborators and Co-Editors}
+% {\bf List in alphabetical order all persons, including their first and last names and current organizational affiliation, who are, or who have been, collaborators or co-authors with you on a research project, book or book article, report, abstract, or paper during the 48 months preceding the submission of this application.
+For publications or collaborations with more than 10 authors or participants, only list those individuals in the core group with whom the Principal Investigator interacted on a regular basis while the research was being done.
+Also, list any individuals who are currently, or have been, co-editors with you on a special issue of a journal, compendium, or conference proceedings during the 24 months preceding the submission of this application.
+For each COI, include the source of the conflict (e.g., co-author, collaborator, co-editor, graduate student, advisor).
+If there are no collaborators or co-editors to report, state None.
+The COI list may require additional pages beyond the three-page limit.}
+\iffalse
 \begin{itemize}
@@
 \end{itemize}
+\fi
+Purushotham V. Bangalore, University of Alabama at Birmingham (collaborator).
+Ron Brightwell, Sandia National Laboratories (co-author, collaborator).
+Phil Carns, Argonne National Laboratory (co-author, collaborator).
+Geoff Danielson, Sandia National Laboratories (collaborator).
+Jill Gemmill, Clemson University (co-author).
+Kevin Harms, Argonne National Laboratory (co-author).
+Jay Harris, Clemson University (co-author).
+Dries Kimpe, Argonne National Laboratory (co-author, collaborator).
+Ruth Klundt, Sandia National Laboratories (co-author).
+Thomas Kroeger, Sandia National Laboratories (collaborator).
+Darrell Long, University of California-Santa Cruz (collaborator).
+Carlos Maltzahn, University of California-Santa Cruz (co-author, collaborator).
+Ethan Miller, University of California-Santa Cruz (collaborator).
+Ron Oldfield, Sandia National Laboratories (collaborator).
+David R. Resnick, Sandia National Laboratories (co-author).
+Robert Ricci, University of Utah (co-author).
+Robert Ross, Argonne National Laboratory (co-author, collaborator).
+Haiying Shen, Clemson University (collaborator).
+Anthony Skjellum, Auburn University (co-author, collaborator, Ph.D. advisor).
+Weibin Sun, Google (co-author).
+Zhiwei Sun, Data Direct Networks (co-author, collaborator).
+Lee Ward, Sandia National Laboratories (co-author, collaborator).
+
+%\subsubsection{Graduate and Postdoctoral Advisors and Advisees}
+%{\bf List the names and current or- ganizational affiliations of your graduate advisor(s) and principal postdoctoral spon- sor(s).
+Also, list the names and current organizational affiliations of your graduate students and postdoctoral associates.}
 

### 2015-07-08 abstract.tex
changes by Tammy

@@
 and services, the Sirocco peer-to-peer file system at
-Sandia and the object storage and annotation expertise of UC Santa Cruz, to
+Sandia, and the object storage and annotation expertise of UC Santa Cruz, to
 explore application and multi-tier storage aware data management
@@
 XGC1, GTC, QMCPack, and SpecFM3D simulations. 
-Our solutions will provide  new functionalities and APIs for
+Our solutions will provide  new functionalities and Application Program Interfaces (API) for:
 1) specifying, at the application level, data annotations that enable the

### 2015-07-08 coverpage.tex
changes by Tammy

@@
 \begin{tabbing}
-\hspace*{10mm} \= \makebox[65mm][l]{\textbf{Lead Principal Investigator:}} \=
-\vrule height0.4pt depth 0pt width 2.5in \\
-\> \> Scott Klasky,  Distinguished Scientist \\
+\hspace*{10mm} \= \makebox[65mm][l]{\textbf{Lead Principal Investigator:}} \= Scott Klasky,  Distinguished Scientist \\
 \> \> Oak Ridge National Laboratory \\
@@
 \end{tabbing}
+%\vrule height0.4pt depth 0pt width 2.5in \\
 
@@
 \begin{tabbing}
-\hspace*{10mm} \= \makebox[65mm][l]{\textbf{Administrative Point of Contact:}} \=
-\vrule height0.4pt depth 0pt width 2.5in \\
-\> \> Barney Maccabe \\
+\hspace*{10mm} \= \makebox[65mm][l]{\textbf{Administrative Point of Contact:}} \= Barney Maccabe \\
 \> \> Oak Ridge National Laboratory \\

### 2015-07-08 narrative/data-description.tex
changes by Tammy

@@
 We will design and develop new techniques for data management, description,
-and access that allow users to describe the data utility based on their
+and access that allows users to describe the data utility based on their
 expectations. This can be done by allowing users to apply plugins
@@
 available space.
-Finally we will study the use of an external data annotation system that can
+Finally, we will study the use of an external data annotation system that can
 provide information to the storage system without requiring recompilation of the application.
@@
 XGC1, GTC, and SPECFM3D. We will investigate: 1) what hints the users can give
-the system, 2) how the users can replace much of their
+the system, and 2) how the users can replace much of their
 application specific code with
@@
 The user can pick a data-refactoring scheme (see
-Sec~\ref{sec:data-refactor}) which will classify the particle data into 4 different
-prioritization levels. The user can then pick a data utility, such as 3 months
+Sec~\ref{sec:data-refactor}) which will classify the particle data into four different
+prioritization levels. The user can then pick a data utility, such as three months
 for the highest priority items and 12 months for the lowest priority data.
@@
 We will investigate what choices the SSIO system must make in order to
-keep these pieces ``together'', from a performance perspective. Can the storage
-system move everything to a lower layer when all of it doesn't fit or is there
+keep these pieces "together,'' from a performance perspective. Can the storage
+system move everything to a lower layer when all of it doesn't fit, or is there
 other, less tightly related data that can be evicted to preserve the collective

### 2015-07-08 narrative/data-reduction.tex
changes by Tammy

@@
 later analysis is no longer viable simply because the amount of generated data is too large.
-In future, it will be vital to take advantage of a priori
+In the future, it will be vital to take advantage of a priori
 information when 1) writing and reading from the SSIO layer in order to gain
@@
 
-\paragraph{State of the art:} {\bf Data Reduction Techniques} Data compression
+\paragraph{State of the art:} {\bf Data Reduction Techniques.} Data compression
 broadly consists of three basic steps: (1) Pre-processing to extract as much
@@
 represent the data in a form that becomes more amenable to compression.  (2)
-Removal of redundancy such as omission of duplicates etc.  Finally, (3) the
+Removal of redundancy such as omission of duplicates, etc.  Finally, (3) the
 resulting data is compressed using entropy coding. The level of compression
@@
 transformation through to simply reordering the data.  What these, and other
-catch-all methods, fail to exploit is a priori information on the nature and
+catch-all methods fail to exploit is a priori information on the nature and
 provenance of the data.
@@
 	re-organization we have devised a scheme where the data pre-conditioner
-	first sorts the data in bins of 1024 words, and then we maintain the
+	first sorts the data in bins of 1024 words, then we maintain the
 	sort order, and then we used wavelets and spline fits to then
@@
 other nodes to then reduce and re-organize, 2) do we first reduce the data and
-then later re-organize this data on different nodes?  3) do we use different
+then later re-organize this data on different nodes?, or 3) do we use different
 cores on a node to do data-refactoring.  Similarly we will examine the cost of

### 2015-07-08 narrative/intro.tex
changes by Tammy

@@
 supported data lifetime, our approach will enable researchers to 
-place more knowledge into the SSIO system so that data can be recast from 
+place more knowledge into the Storage Systems and Input/Output (SSIO) system so that data can be recast from 
 a sequence of bytes to information rich data. This information rich data can then
@@
 3 & Mechanisms for data movement across the storage hierarchy & what are we doing. & \S \ref{sec:managing-data-life},  \S\ref{sec:migration} \\\hline
-4 &  Extend the interaction of IO middleware with the resource management system & We will manage data from the application  to the middleware to the storage system and allow them to have an autonomic interaction & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
-5 & Design and implement of IO middleware architectures  & We will create APIs to guide the placement of chunks of data across the SSIO and autonomically move data when necessary& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
+4 &  Extend the interaction of I/O middleware with the resource management system & We will manage data from the application  to the middleware to the storage system and allow them to have an autonomic interaction & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
+5 & Design and implement of I/O middleware architectures  & We will create APIs to guide the placement of chunks of data across the SSIO and autonomically move data when necessary& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
 6 & Approaches to improve the ability of SSIO software to support checkpoint/restart & We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then purged when it is no longer needed & \S \ref{fig:ssio-bucket}  \\

### 2015-07-08 narrative/placement.tex
changes by Tammy

@@
 Some of the high level research questions that we will explore are as follows: 
-(1) How can we initially place data so that it can be discovered and consumed efficiently?
+(1) How can we initially place data so that it can be discovered and consumed efficiently?,
 %What methods can be utilized in the initial placement of data that do not impose expensive
@@
 (2) How can the placement and migration of data across a multi-tiered storage hierarchy be 
-optimized at runtime, both from the application and system perspective?
+optimized at runtime, both from the application and system perspective?,
 %(3) How can we migrate data throughout the layers of storage
 %without making the discovery process unbounded? 
-(4) How can knowledge about the application used to better prepare the data for consumption? 
+(4) How can knowledge about the application used to better prepare the data for consumption?, and 
 (5) When and how do we make the decision to purge data?  %using this additional knowledge?

### 2015-07-08 narrative/data-description.tex
utility function description.

@@
 explored, and we believe it will be vital for exascale storage solutions.
-{\color{red}\bf there must be other work in this area, and we need to get more
-references}
 
@@
 towards a lower stack or can be staged to that stack directly. 
+% Description of the utility function research
+
+An important research question which me must address is the semantics of
+describing the utility function, and the possibility that the users will
+accept this, and if the system can then use this information, and make 
+decisions based on this utility function. We realize that the system will
+try to optimize across all of the different user request, and the utility
+function must have defaults that users can inquire into, and change these
+when they see that these defaults can have a possibile negative impact on
+their knowledge discovery.
+
+The assumptions which we will make in choosing a utility function is that:
+1) There will only be a small number of priority levels that users will use
+to re-prioritze their data; 2) The number of tiers in the storage system will
+also be small (e.g. NVRAM, PFS, Campaign Storage, HPSS), 3) The default
+lifetime of objects in each tier will be specified by the resource by
+a tuple (size, lifetime).  The user can then utilize this information to
+place information to the system such as: store the highest priority data
+as high up on the hiearchy for 6 months, and then migrate this data to tape.
+The user can also define the utility for CR files as that they have high
+priority but their lifetime is for a short period of time; i.e. when the
+next CR file appears on the system or the next two CR files. Research in
+describing this utility function will be in conjunction with our users
+and we will evaluate how policies can be placed in the system and how well
+this utility function can be observed by the user.
+
 %
-
 \begin{wrapfigure}{r}{0.6\textwidth}

### 2015-07-08 participant-table.tex
utility function description.

@@
 \\
-\textbf{\ornl} & \makebox[\longestpientry][l]{\textbf{Scott Klasky} (\emph{klasky@ornl.gov})}\hspace*{1em}Hasan Abbasi, Gary Liu, Kimmy Mu, Manish Parashar\\
+\textbf{\ornl} & \makebox[\longestpientry][l]{\textbf{Scott Klasky} (\emph{klasky@ornl.gov})}\hspace*{1em}Hasan Abbasi, Gary Liu, Kimmy Mu\\
 \textbf{\snl} & \makebox[\longestpientry][l]{\textbf{Gerald Lofstead} (\emph{gflofst@sandia.gov})}\hspace*{1em}Matthew Curry, Lee Ward\\
-\textbf{\ucsc} & \makebox[\longestpientry][l]{\textbf{Carlos Maltzahn} (\emph{carlosm@soe.ucsc.edu})}\hspace*{1em}\
+\textbf{\rutgers} & \makebox[\longestpientry][l]{\textbf{Manish Parashar} (\emph{parashar@rutgers.edu})}\hspace*{1em}\\
+\textbf{\ucsc} & \makebox[\longestpientry][l]{\textbf{Carlos Maltzahn} (\emph{carlosm@soe.ucsc.edu})}\hspace*{1em}
 \end{tabularx}

### 2015-07-08 narrative/proposed-work.tex

@@
 predictable performance, the I/O layer performance must be guessed by the
-users, which quite often mean that science is not performed well.
+users, which quite often means that science is not performed well.
+
+Our work is to take much of the ``magic'' and ``tricks''
+away from optimizations necessary for users to understand to gain optimizations
+on current file systems, and provide a systematic approach to describing enough
+of their intentions and knolwedge from the {\it user} space along with much
+of the hidden knowledge of the underlying storage layers. For example,
+   ADIOS contains most of the ``best practices'' for I/O optimizations for
+   each file system, and as users place more knowledge into ADIOS, ADIOS
+   can make optimizations. This becomes much more difficult as users want to
+   describe more about the storage hiearchy and more about the priorities of
+   different chunks of data. Our research is aimed to allow users to simply
+   describe this and get the descriptions of system knowledge so that the
+   best optimizations can be made.
 

### 2015-07-08 narrative/proposed-work.tex
Reworked new paragraph in proposed research section

--- 2015-07-07 narrative/proposed-work.tex
@@
 predictable performance, the I/O layer performance must be guessed by the
-users, which quite often mean that science is not performed well.
+users, which quite often means that science is not performed well.
+
+We aim to alleviate the need for the ``magic'' and ``tricks''
+that are currently required to optimize application I/O performance
+on today's file systems. To accomplish this, we will provide a systematic
+approach for describing intentions and other knowledge from {\it user}
+space, as well as hidden knowledge of the underlying storage layers.
+%For example, ADIOS contains most of the ``best practices'' for I/O optimizations for
+%   each file system, and as users place more knowledge into ADIOS, ADIOS
+%   can make optimizations.
+By capturing user intentions and acting upon them in the middleware, we
+free the user from polluting application code with system specific
+optimizations which may provide better performance in the short term,
+but are detrimental to the long-term maintainability of applications, and
+are more sensitive to changes to system configuration. We have successfully
+employed this strategy of separation of concerns in ADIOS, and believe that
+it will become increasingly important as users want to describe more about
+the storage hierarchy and the priorities of different chunks of data. 
+Our research is aimed to allow users to simply describe these things and
+utilize the descriptions so that the best optimizations can be made.
 

### 2015-07-08 narrative/data-reduction.tex
syntax changes

@@
 data into two basic categories: structured data (in the sense that it satisfies
-a known or relatively simple model) or unstructured data (in which information
+a known or relatively simple model), or unstructured data (in which information
 content follows no obvious or explicit model).  Although scientific data
@@
 	in-situ. This operation involves no communication.  Data needs to be
-	copied, so the memory requirements are increased. At most this requires
+	copied, so the memory requirements are increased. {\color{red}At most this requires
 	two copies an individual dataset in memory, and we will investigate
@@
 	specify that the data will be overwritten after it is written to the
-	storage system. This is often the case for many of the data quantities
+	storage system CHECK sentence-tsr}. This is often the case for many of the data quantities
 	written from the simulation, but there are many cases where, for
@@
 	frequency. This binning of data in frequency space can take as an input
-	(storage bandwidth knowledge, storage size limitations, and user
+	(storage  {\color{red}ending )? CHECK-tsr} bandwidth knowledge, storage size limitations, and user
 	intentions of how long the majority of the data will be frequently
@@
 will be embedded in the system and can be selected either by the user or by the
-application itself. The techniques will range from crude catch all methods such
+application itself. The techniques will range from crude catch-all methods such
 as those employed by general compression routines such as FPC, through to

### 2015-07-08 narrative/evaluation.tex
tight itemize, tight enumerate
spelling, etc. changes
syntax changes

@@
 \begin{table}[ht]
-\caption{Guiding Principles}
+
 \label{table:eval}
 \centering
-\begin{tabular}{ | p{1.4in} | p{2in} | p{2in} | }
+\begin{tabular}{ | p{1.4in} | p{2in} | p{2.5in} | }
   \hline
@@
 \end{tabular}
+
+\caption{Guiding Principles}
 \end{table}
@@
 to evaluate the leading edge applications working with the storage system to
-determine performance predictability and data quality.  Our evaluation will
+determine performance predictability and data quality.  Our evaluation 
 will use the multiple storage layers available on systems at both SNL and the
@@
 storage system, SSDs on each node, and is connected to the center-wide Lustre
-file system. We will perform test in both the system level (on sith) and in
+file system. We will perform tests in both the system level (on sith) and in
 the user level (on titan), and evaluate our results using our application
@@
 need to investigate this on the next generation OLCF system (Summitt) and the
-SNL system (Trinity) which we will have early access, along with all of the
+SNL system (Trinity) which we will have early access along with all of the
 applications we will be investigating.
@@
 hundreds of time slices, but when we visualize this data, we would like very
-fast access to the data ``first look''. We will determine if the data will be
+fast access to the data ``first look.'' We will determine if the data will be
 sufficiently representative of the overall data.  We will work with our

### 2015-07-08 narrative/intro.tex
syntax changes

@@
 2 & Mapping science data models onto hierarchical storage &  Create the ability to re-organize, reduce and prioritize data chunks from an object& \S\ref{sec:data-refactor}, \S\ref{sec:data-description},  \S\ref{sec:managing-data-life}\\\hline
-3 & Mechanisms for data movement across the storage hierarchy & what are we doing. & \S \ref{sec:managing-data-life},  \S\ref{sec:migration} \\\hline
+3 & Mechanisms for data movement across the storage hierarchy & {\color{red}what are we doing. CHECK-tsr} & \S \ref{sec:managing-data-life},  \S\ref{sec:migration} \\\hline
 4 &  Extend the interaction of I/O middleware with the resource management system & We will manage data from the application  to the middleware to the storage system and allow them to have an autonomic interaction & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
@@
 This forced the team to carefully scrutinize the large amount of data to be generated and save only the very most important pieces.
-Further, they had to break this greatly reduced data into two categories: first data they thought they would need to access while the simulation was running or shortly thereafter, and second, data they would need after the whole campaign was finished.
+Further, they had to break this greatly reduced data into two categories: first, data they thought they would need to access while the simulation was running or shortly thereafter, and second, data they would need after the whole campaign was finished.
 The former category would be stored on disk, and the later would be archived.
@@
 writing or reading at the same time, leading to unpredictable performance.
-By requiring a certain quality of service the system could potentially lock out other users. 
+By requiring a certain QoS of service the system could potentially lock out other users. 
 The storage system must also automatically move data from one tier of the storage to
-another, without affecting the quality of service that was offered at request
+another, without affecting the QoS that was offered at request
 time. The storage system must be able to organize
@@
 of this class of user is that they want a reduction of latency and they
-require a certain amount of quality of service as well. When someone is
+require a certain amount of QoS as well. When someone is
 doing interactive analysis and expects that the data they read in is about

### 2015-07-08 narrative/naming-service.tex
syntax changes

@@
 decomposed objects of potentially varying fidelity with potentially multiple
-version with a POSIX API-compatible interface.
+version{\color{red}s?} with a POSIX API-compatible interface.
 
@@
 Sirocco offers a ``flush'' operation that forcibly migrates data to the
-long-term resilience requirements freeing space in the scare resources. In
+long-term resilience requirements, freeing space in the {\color{red}scare CHECK-tsr} resources. In
 many, if not most cases, this will move data towards or actually on to a device
-like tape intended for long-term storage.
+like tape, intended for long-term storage.
 
@@
 to a partial variable granularity, a 1 PB output with 500 GB of ``high
-interest'' data can limit this costly tier usage to just 500 GB greatly
+interest'' data can limit this costly tier usage to just 500 GB, greatly
 improving storage availability in the close/fast/small tier and focus
-performance/capacity costs to get the most science from from the platform.  By
+performance/capacity costs to get the most science from the platform.  By
 only placing ``high interest'' data in the close/fast/small tier, we will
@@
 insights.  Sirocco identifies a data segment using a
-contaier/object/fork/address tuple that can have associated attributes. We will
-both add a special attribute indicating that this record is of ``high
+container/object/fork/address tuple that can have associated attributes. {\color{red}We will
+add both a special attribute indicating that this record is of ``high
 interest'' as well adapting the migration mechanisms to use this attribute, if
 present, to strongly encouraging keeping the record in close/fast/small storage
-as long as possible.
+as long as possible. CHECK sentence-tsr}
 
@@
 We plan to address scalability problems by continuing with the LWFS and Sirocco
-model. LWFS and Sirocco have take an approach similar to the pure object
+model. LWFS and Sirocco have taken an approach similar to the pure object
 stores, but with a focus on the HPC setting. They have abandonded a fully POSIX
 compliant metadata service as the default model in favor of a
-container/object/fork/addrress tuple for identifying data similar to those used
+container/object/fork/address tuple for identifying data similar to those used
 for pure object stores. By having a service that addresses the object
@@
 \paragraph{Related Work:}
-Structurally, the traditional POSIX naming service offeriing a hierarchical
+Structurally, the traditional POSIX naming service offering a hierarchical
 space consisting of directories and files may be maintained for backwards
@@
 worked to reduce this contention by doing things like reducing the serialized
-scope to a single directory or subtree or allowing a single process in a
+scope to a single directory or subtree, or allowing a single process in a
 collective file operation talk with the metadata service and distribute a
@@
   requests within a specified time bound with accuracy? How do we estimate the
-  completeness of the response?
+  completeness of the response?, 
 (2) For POSIX API requests, how do we respond if only varying data fidelity
-  levels are available for the requested data object?
+  levels are available for the requested data object?, 
 (3) What level of metadata should we maintain and keep in sync to enhance
-  response quality within time-bound requests?
+  response quality within time-bound requests?, 
 (4) How much time should be allowed for searching for metadata given the user
@@
   the data retrieval time from different devices? How do we integrate the read
-  time overhead into the metadata requests?
+  time overhead into the metadata requests?, and
 (5) How do we provide appropriate security while maintaining scalability,
@@
 
-\item {\bf Milestone 3:} Demonstrate incorporating A\&A and show scalability under
-both application load as well as storage system pressures showing that we can
+\item {\bf Milestone 3:} Demonstrate incorporating {\color{red}A\&A (what does this stand for CHECK-tsr)} and show scalability under
+both application load as well as storage system pressures, showing that we can
 tolerate both loads and maintain quality of service.

### 2015-07-08 narrative/placement.tex
syntax changes

@@
 (2) How can the placement and migration of data across a multi-tiered storage hierarchy be 
-optimized at runtime, both from the application and system perspective?,
+optimized at runtime, both from the application and system perspective?, {\color{red} where is 3?} 
 %(3) How can we migrate data throughout the layers of storage
@@
 proposed storage system is the ability to refactor and reduce data as it is
-generated, and to reorganized and regenerate the data as it is accessed. We
+generated, and to reorganize and regenerate the data as it is accessed. We
 carry this principle into the placement and movement of data by allowing
@@
 distinct areas. Firstly, we will study the challenges and tradeoffs of either
-augmenting the I/O interface with hints or allowing the additional of an
+augmenting the I/O interface with hints, or allowing the addition of an
 external specification that defines the use case. Our experience with
@@
 specification. Secondly, we will study how hints can guide the initial data
-placement as data is handed off from application to storage ( during a
+placement as data is handed off from application to storage (during a
 write) and from storage to application (during the read). In both cases we
@@
 with which they are accessed, etc., and are used during runtime data
-management. For example, data objects will higher utility value may be
+management. For example, data objects with higher utility value may be
 placed closer (e.g., in faster memory) to the accessing application.
@@
 spatial-temporal data access patterns, physical network topology
-information, and etc., which will be tracked and gathered at runtime, and
+information, {\color {red}and etc., CHECK-tsr} which will be tracked and gathered at runtime, and
 then be used to estimate and predict the behavior of coupled applications in
@@
 \paragraph{Background:} Providing data to users for knowledge discovery is
-one of the primary purposes of our storage system design. There is
-There is a
+one of the primary purposes of our storage system design. There is a
 spectrum of approaches on how
@@
 \label{sec:purging-data}
-(only outline since I have to go)
+{\color{red}(only outline since I have to go)}
 1. Purging is important from the perspective of preventing uncontrolled

### 2015-07-08 narrative/qos.tex
syntax changes

@@
 to their centrally maintained block allocation tables. 
-The opposite of file systems are ``pile'' systems, which, instead of keeping track of updates, search for 
+The opposite of file systems are ``pile'' systems, which, instead of keeping track of updates, searchs for 
 resources at request time. Pile systems allow very efficient writes because clients can write wherever
@@
 ``filing'' data, making an upfront investment to ensure finding data quickly, and ``piling'' 
-data, pushing the cost of finding to the moment when the data is actually needed. This trade-off 
+data, pushing the cost of finding to the moment when the data is actually needed. {\color{red}This trade-off 
 has a parallel in data management where there are two alternative paths to achieve knowledge: 
@@
 an upfront ingest cost to achieve fast querying; the analogue to piling is to not pay any upfront 
-cost and query the raw data directly by using data processing platforms such as Hadoop. Recent work 
+cost and query the raw data directly by using data processing platforms such as Hadoop sentence CHECK-tsr}. Recent work 
 has found that the combination of these two techniques can be optimized to perform better than 
@@
 a partition, i.e. finely partitioned search spaces require little or no search within a partition 
-but offer little placement flexibility. Partitioning can also be fine in one dimension and coarse 
+but offer little placement flexibility. Partitioning can also be fine in one dimension, and coarse 
 in another to allow flexibility where it is most needed while still ensuring fast finding. For 
@@
 \subsubsection{QoS}
-\paragraph{Background:} Quality of Service (QoS) refers to the
+\paragraph{Background:} QoS refers to the
 properties of the performance quality of a particular service, in
@@
 resources. This traditional approach leads to interference effects
-\cite{lofstead:2010:io-variability,liu_hotstorage} that can greatly impact IO performance predictability. 
+\cite{lofstead:2010:io-variability,liu_hotstorage} that can greatly impact I/O performance predictability. 
 %Further, with the introduction of additional tiers in the
@@
   \includegraphics[width=\columnwidth]{graphics/infrastructure.pdf}\vspace{-1.2in}
-  \caption{Resource usage distribution for OST(a), OSS(b) and LNETs(c). }
+  \caption{Resource usage distribution for OST(a), OSS(b) and LNETs(c) }
 \end{figure}
@@
 
-\paragraph{Related Work:} On current production HPC systems, absolute IO performance
+\paragraph{Related Work:} On current production HPC systems, absolute I/O performance
 guarantees are not possible. Previous work \cite{lofstead:2010:io-variability,liu_hotstorage} 
-from this project team uses a client-assisted approach to mitigate IO variablity at scale (100,000-core).
+from this project team uses a client-assisted approach to mitigate I/O variablity at scale (100,000-core).
 Recently server side QoS scheduling has also been studied for HPC applications \cite{Dai:2014}.

### 2015-07-09 narrative/data-reduction.tex
some changes to fix syntax problems

@@
 	in-situ. This operation involves no communication.  Data needs to be
-	copied, so the memory requirements are increased. {\color{red}At most this requires
-	two copies an individual dataset in memory, and we will investigate
+	copied, so the memory requirements are increased. At most this requires
+	two copies of an individual dataset in memory, and we will investigate
 	techniques to allocate and deallocate this memory if the user will
 	specify that the data will be overwritten after it is written to the
-	storage system CHECK sentence-tsr}. This is often the case for many of the data quantities
+	storage system. This is often the case for many of the data quantities
 	written from the simulation, but there are many cases where, for
@@
 	different parts of the same data using different quality.  This allows
-	us to place the lowest frequency pieces in the fastest storage, and the
-	highest frequency either on the slowest storage tiers, or if the data
-	sizes are prohibitively  costly, not even write out the highest
-	frequency. This binning of data in frequency space can take as an input
-	(storage  {\color{red}ending )? CHECK-tsr} bandwidth knowledge, storage size limitations, and user
-	intentions of how long the majority of the data will be frequently
-	accessed.  In order to fully take advantage of frequency based
+	us to place the lowest frequency chunks in the fastest storage, and the
+	highest frequency chunks  on either  the slowest storage tiers, or if the data
+	sizes are prohibitively  costly, not even write out  these pieces. 
+%	This binning of data in frequency space can take as an input
+%	(storage  {\color{red}ending )? CHECK-tsr} bandwidth knowledge, storage size limitations, and user
+%	intentions of how long the majority of the data will be frequently
+%	accessed. 
+	 In order to fully take advantage of frequency based
 	re-organization we have devised a scheme where the data pre-conditioner

### 2015-07-09 narrative/data-reduction.tex
minor changes for readability.

@@
 	specify that the data will be overwritten after it is written to the
-	storage system. This is often the case for many of the data quantities
+	storage system. This is often the case for many of the  quantities
 	written from the simulation, but there are many cases where, for
@@
 	calculation, we need to duplicate the storage. Our observation with
-	working with the XGC1, GTC, Warp, and PiconGPU simulations (all
-	leadership class simulations) is that we can temporary increase the
+	working with the XGC1, GTC, and PiconGPU simulations 
+	 is that we can temporary increase the
 	storage of the particles, and then release them since the temporary
@@
 	space and then specify that as the frequency increases the importance
-	of data decreases.   Common technique which is used for streaming data,
+	of data decreases.   Common techniques  used for streaming data,
 	and data reduction techniques such as JPEG-2000~\cite{jpeg2000} is to

### 2015-07-09 narrative/intro.tex
fixed the table in intro.

@@
 \begin{center}
+\caption {Relevance to SSIO}
+\label{tab:call}
 \begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{2.2cm} |}\hline
@@
 2 & Mapping science data models onto hierarchical storage &  Create the ability to re-organize, reduce and prioritize data chunks from an object& \S\ref{sec:data-refactor}, \S\ref{sec:data-description},  \S\ref{sec:managing-data-life}\\\hline
-3 & Mechanisms for data movement across the storage hierarchy & {\color{red}what are we doing. CHECK-tsr} & \S \ref{sec:managing-data-life},  \S\ref{sec:migration} \\\hline
+3 & Mechanisms for data movement across the storage hierarchy & Manage data migration based on policies& \S \ref{sec:managing-data-life},  \S\ref{sec:migration} \\\hline
 4 &  Extend the interaction of I/O middleware with the resource management system & We will manage data from the application  to the middleware to the storage system and allow them to have an autonomic interaction & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
@@
 \end{center}
-%\caption {WRITE A CAPTION FOR THIS TABLE}
 \vskip -0.5cm
-\caption {}
-\label{tab:call}
 \end{table}

### 2015-07-09 narrative/data-description.tex

@@
 Our current data annotation techniques demonstrated in
-ADIOS~\cite{lofstead:2009:adaptible} implement a binary-packed data format that
+ADIOS~\cite{lofstead:2009:adaptible} implement a binary-packed (BP) data format that
 allows data characteristics such as min, max and index to be wrapped around
@@
 efficiently, without overburdening users with the management of complex data models, such
-as AMR, from the user space. Despite these efforts, the aspect of data importance has yet to be 
+as adaptive mesh refinement (AMR), from the user space. 
+FFS~\cite{ffs} implements a type system that provides highly efficient binary
+data communication and XML-like data description, facilitating data sharing
+between collaborators. Additionally, NetCDF~\cite{netcdf} and HDF5~\cite{hdf5}
+uses a fixed data format to provide a hierarchical data descriptions for scientific data.
+Despite these efforts, the aspect of data importance has yet to be 
 explored, and we believe it will be vital for exascale storage solutions.

### 2015-07-09 narrative/naming-service.tex

@@
 decomposed objects of potentially varying fidelity with potentially multiple
-version{\color{red}s?} with a POSIX API-compatible interface.
+versions with a POSIX API-compatible interface.
 
@@
 Sirocco offers a ``flush'' operation that forcibly migrates data to the
-long-term resilience requirements, freeing space in the {\color{red}scare CHECK-tsr} resources. In
+long-term resilience requirements, freeing space in the scarce resources. In
 many, if not most cases, this will move data towards or actually on to a device
@@
 insights.  Sirocco identifies a data segment using a
-container/object/fork/address tuple that can have associated attributes. {\color{red}We will
+container/object/fork/address tuple that can have associated attributes. We will
 add both a special attribute indicating that this record is of ``high
-interest'' as well adapting the migration mechanisms to use this attribute, if
-present, to strongly encouraging keeping the record in close/fast/small storage
-as long as possible. CHECK sentence-tsr}
+interest'' as well adapting a migration mechanisms to use this attribute, if
+present, to strongly encourage keeping the record in  the highest storage tier as possible.
 
@@
 model. LWFS and Sirocco have taken an approach similar to the pure object
-stores, but with a focus on the HPC setting. They have abandonded a fully POSIX
+stores, but with a focus on the HPC setting. They have abandoned a fully POSIX
 compliant metadata service as the default model in favor of a
@@
 While pure object stores, such as those popular in the big data
-domain~\cite{Fitzpatrick:2004:memcached,others} avoid this bottleneck by
+domain~\cite{Fitzpatrick:2004:memcached}, others avoid this bottleneck by
 strictly offering an object ID with the application required to manage how this

### 2015-07-09 narrative/placement.tex

@@
 predictive analysis requirements. This information may include runtime
-spatial-temporal data access patterns, physical network topology
-information, {\color {red}and etc., CHECK-tsr} which will be tracked and gathered at runtime, and
+spatial-temporal data access patterns, and/or physical network topology
+information, which will be tracked and gathered at runtime, and
 then be used to estimate and predict the behavior of coupled applications in
@@
 application knowledge, and use this information to prefetch data objects.
-For example, we will identify and catalogue key phases and access patterns
+For example, we will identify and catalog key phases and access patterns
 in the target workflows and then use it to characterize patterns at runtime

### 2015-07-09 narrative/qos.tex

@@
 ``filing'' data, making an upfront investment to ensure finding data quickly, and ``piling'' 
-data, pushing the cost of finding to the moment when the data is actually needed. {\color{red}This trade-off 
+data, pushing the cost of finding to the moment when the data is actually needed. This trade-off 
 has a parallel in data management where there are two alternative paths to achieve knowledge: 
-the analogue to filing is ``extract, transform, and load'' (ETL) data into a warehouse, paying 
-an upfront ingest cost to achieve fast querying; the analogue to piling is to not pay any upfront 
-cost and query the raw data directly by using data processing platforms such as Hadoop sentence CHECK-tsr}. Recent work 
+the analogue to filing is  to ``extract, transform, and load'' (ETL) data into a warehouse, paying 
+an upfront ingest cost to achieve fast querying; and the analogue to piling is to not pay any upfront 
+cost and later querying   the raw data directly by using a data processing platform such as Hadoop.  Recent work 
 has found that the combination of these two techniques can be optimized to perform better than 
@@
 
-\subsubsection{QoS}
+\subsubsection{Quality of Service (QoS)}
 \paragraph{Background:} QoS refers to the
@@
 \end{figure}
+\vskip -1.5in
 

### 2015-07-09 other-support/ward-cp.tex

\documentclass[10pt]{article}
\usepackage[english]{babel}
\usepackage{cite}
\usepackage{url}
\usepackage{times}
\usepackage{graphicx}
\usepackage[sectionbib]{bibunits}
\pagestyle{empty}

% Set page geometry
\usepackage{geometry}
\geometry{verbose,letterpaper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}

\begin{document}

\begin{center}
{\Large Current and Pending Support for H. Lee Ward}
\end{center}

Mr. Ward is a principle member of the technical staff in the Center for
Computing Research group at Sandia National Laboratories (SNL). The following
table identifies current and pending support. This funding provides for effort,
travel, and resources. The effort levels will be adjusted if pending proposals
are funded, and as current projects finish.

\section*{Current Support}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \= Sirocco\\
Principal Investigator      \>: \> Lee Ward\\
Source of Support           \>: \> DOE NNSA/Advanced Simulation and Computing\\
Period Covered              \>: \> 10/2014-9/2015\\
Commitment                  \>: \> 0.85 FTE\\
Description                 \>: \> Development activities for Sirocco, a novel storage systemfor Exascale computers.\\
Overlap                     \>: \> None\\
\end{tabbing}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Pending Proposals (in addition to this)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Pending Proposals (in addition to this)}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \= Exascale Storage Authorization and Data-Location Research\\
Principal Investigator      \>: \> Anthony Skjellum\\
Source of Support           \>: \> DOE Office of Science\\
Period Covered              \>: \> 10/2015-9/2018\\
Commitment                  \>: \> 0.058 FTE\\
Description                 \>: \> Design, consulting, mentoring in support of research using Sirocco as a vehicle.\\
Overlap                     \>: \> None\\
\end{tabbing}

\begin{tabbing}
Project Title \hspace{5.5em}\=: \= Managing Node-Based Non-Volatile Memories for HPC Clusters\\
Principal Investigator      \>: \> Ethan Miller\\
Source of Support           \>: \> DOE Office of Science\\
Period Covered              \>: \> 10/2015-9/2018\\
Commitment                  \>: \> 0.25 FTE\\
Description                 \>: \> Design and development activities for Sirocco in support of research using Sirocco as a vehicle.\\
Overlap                     \>: \> None\\
\end{tabbing}

\end{document}

### 2015-07-09 narrative/data-description.tex
spelling, etc.

@@
 the need. As an example, retrieving and visualizing 3D field data generated from a simulation
-requires the mesh (consiting of coordinates and connectivity variables) be
+requires the mesh (consisting of coordinates and connectivity variables) be
 accessed simultaneously and with low latency to achieve a good user experience.
@@
 %The new APIs will enable the bridge by specifying selectable
-%performance/quality/- cost tradeoffs from both the application and system
+%performance/quality/- cost trade-offs from both the application and system
 %perspectives based upon the user guided rules/policy and runtime system
@@
 function must have defaults that users can inquire into, and change these
-when they see that these defaults can have a possibile negative impact on
+when they see that these defaults can have a possible negative impact on
 their knowledge discovery.
+
+\begin{wrapfigure}{r}{0.6\textwidth}
+        \vspace{-6ex}
+        \begin{centering} 
+	\includegraphics[scale=0.8]{graphics/SSIO-bucket.pdf}
+        %\vspace{1ex}
+        \caption{The system manages the data across the various stages of the data lifecycle}
+        \label{fig:ssio-bucket}
+        \end{centering}
+      % \vspace{-1ex}
+\end{wrapfigure}
 
@@
 1) There will only be a small number of priority levels that users will use
-to re-prioritze their data; 2) The number of tiers in the storage system will
+to re-prioritize their data; 2) The number of tiers in the storage system will
 also be small (e.g. NVRAM, PFS, Campaign Storage, HPSS), 3) The default
@@
 place information to the system such as: store the highest priority data
-as high up on the hiearchy for 6 months, and then migrate this data to tape.
+as high up on the hierarchy for 6 months, and then migrate this data to tape.
 The user can also define the utility for CR files as that they have high
@@
 %
-\begin{wrapfigure}{r}{0.6\textwidth}
-        \vspace{-2ex}
-        \begin{centering} 
-	\includegraphics[scale=0.8]{graphics/SSIO-bucket.pdf}
-        %\vspace{1ex}
-        \caption{The system manages the data across the various stages of the data lifecycle}
-        \label{fig:ssio-bucket}
-        \end{centering}
-       \vspace{-1ex}
-\end{wrapfigure}
 

### 2015-07-09 narrative/placement.tex
spelling, etc.

@@
 in the case of experimental and observational data (EOD). This data has 
-to be appropriately initially placed (and possibly partitioned) within the heterogenous 
+to be appropriately initially placed (and possibly partitioned) within the heterogeneous 
 multi-tier storage to optimize application I/O overheads.
@@
 placed \textbf{where}. We will explore the use of application hints in two
-distinct areas. Firstly, we will study the challenges and tradeoffs of either
+distinct areas. Firstly, we will study the challenges and trade-offs of either
 augmenting the I/O interface with hints, or allowing the addition of an
@@
 migration of data to other storage servers and across the hierarchy. An
-important reserach aspect here is how a liberated Hierarchical Storage
-Manager (HSM) designed for online use, such as Sirocco, might be employed in
+important research aspect here is how a liberated Hierarchical Storage
+Manager (HSM) designed for on-line use, such as Sirocco, might be employed in
 providing support for the migration of data. Sirocco autonomously groups the
@@
 \paragraph{Bounded estimates for lookup:}
-One challenge for storage systems with continous decentralized migration is
+One challenge for storage systems with continuous decentralized migration is
 the timeliness of lookup when applications issue read calls. This is due to
@@
 \S~\ref{sec:naming-discovery}. To support this discovery process, we will
-investiagte how the migration of data can be performed in a way that enables
+investigate how the migration of data can be performed in a way that enables
 bounded estimates about time to first byte for a potential read operation.
@@
 
-\subsubsection{Making data avilable on multiple tiers}
+\subsubsection*{Making data available on multiple tiers}
 \label{sec:manag-data-mult}
@@
 and available space in the higher tiers. Another approach is to utilize the
-techniques described in \ref{sec:data-refactor} to create auxiliary views of
+techniques described in \S\ref{sec:data-refactor} to create auxiliary views of
 the data that, with user defined error, represent the data, and place these
@@
 reduce capacity misses but the concept can be generalized to the
-likelihood of having the right data in place, i.e. proactively
+likelihood of having the right data in place, i.e. pro-actively
 moving data to high tiers accessed in predictable patterns reduces
@@
 operations to meet the service level objectives for each use case, while
-optimizing the overall utility from the system perpective.
+optimizing the overall utility from the system perspective.
 
@@
 of speed, direction, likely resolution of map, and the fact that
-clients almost never go offroad.
+clients almost never go off-road.
 

### 2015-07-09 narrative/data-description.tex
short description on the auditor

@@
 
-%
 From an application level, we will explore these techniques through our ADIOS
 framework using many of our existing application partners, including
-XGC1, GTC, and SPECFM3D. We will investigate: 1) what hints the users can give
+XGC1~\cite{chang2006integrated}, GTC~\cite{klasky2003grid} , and SPECFM3D~\cite{SPECFEM3D}. We will investigate: 1) what hints the users can give
 the system, and 2) how the users can replace much of their

### 2015-07-09 narrative/data-reduction.tex
short description on the auditor

@@
 simply stored in the form of raw data, but rather in the form of appropriate
-parameterized algorithms. A priori information on the nature of the data is
-used to choose the basic algorithm while the actual values of the data are used
-to tune the parameters used in the reconstruction procedures. We shall refer to
-this paradigm as an Auditor.  We will examine the use of several auditors for
-the simulations which we will work with. One approach is to use a fluid-code to
-audit a kinetic code, which can be a good approximation over short periods of
-time.
+parameterized algorithms. An ``auditor'' is a code which mimics the exascale simulation, 
+but typically can run using less degrees of freedom. This means that the code is a good
+local (in space and time) approximation of the exascale simulation and will be much less
+expensive to execute (so that it can be run in situ), and follows the behavior of the simulation
+for short periods of time. This allows us to ``audit'' the simulation and when the differences 
+become large in a local region of space, we write out the results of the data in this local region.
+The hope is that this will be done only locally and that with a good ``auditor''  this will not happen
+too frequently. This then allows us to use the ``auditor'' to reproduce results during the analysis 
+process using an inexpensive computation and put less pressure on the SSIO layer.
+%
+% 
+%   A priori information on the nature of the data is
+%used to choose the basic algorithm while the actual values of the data are used
+%to tune the parameters used in the reconstruction procedures. We shall refer to
+%this paradigm as an Auditor.  We will examine the use of several auditors for
+%the simulations which we will work with. One approach is to use a fluid-code to
+%audit a kinetic code, which can be a good approximation over short periods of
+%time.

### 2015-07-09 narrative/evaluation.tex
minor changes for readability.
spelling, etc.

@@
 
-\label{table:eval}
 \centering
@@
 \caption{Guiding Principles}
+\label{table:eval}
 \end{table}
@@
 clearly understand how well this system works on current systems and we will
-need to investigate this on the next generation OLCF system (Summitt) and the
+need to investigate this on the next generation OLCF system (Summit) and the
 SNL system (Trinity) which we will have early access along with all of the
@@
 Our evaluation will focus on the improvements gained through additional user
-provided data and attempt to relate the pain/benefit tradeoff. The initial
+provided data and attempt to relate the pain/benefit trade-off. The initial
 hypothesis from working with many LCF users suggests that once the initial

### 2015-07-09 narrative/placement.tex
Edits throughout the section. Add some text for purging .

@@
 
-\subsubsection{Data Placement}
+\subsubsection{Data Placement and Movement}
 \label{sec:init-plac-data}
@@
 \paragraph{Background:}
-When an application outputs data the storage and middleware layers needs to
+When an application outputs or accesses data the storage and middleware layers needs to
 decide \textit{what} data is placed \textit{where} in the multi-level storage system. 
@@
 As described in \S\ref{sec:data-refactor} one of the main components of our
-proposed storage system is the ability to refactor and reduce data as it is
-generated, and to reorganize and regenerate the data as it is accessed. We
-carry this principle into the placement and movement of data by allowing
+proposed storage system is the ability to reorganize, refactor and reduce data 
+as it is generated, and to reorganize and possibly regenerate the data as it is accessed. 
+We carry this principle into the placement and movement of data by allowing
 applications to define hints and policies that guide \textbf{what} data is
 placed \textbf{where}. We will explore the use of application hints in two
-distinct areas. Firstly, we will study the challenges and trade-offs of either
+distinct areas. First, we will study the challenges and trade-offs of either
 augmenting the I/O interface with hints, or allowing the addition of an
 external specification that defines the use case. Our experience with
-developing modern I/O interfaces has shown that both techniques have value,
-and we will investigate the set of hints that are embedded in the
+developing modern I/O interfaces has shown that both techniques 
+have value~\cite{tongipdps15,qiansc15}, and we will investigate the set of hints that are embedded in the
 application code vs. those that are described within an non-compiled
-specification. Secondly, we will study how hints can guide the initial data
+specification. Second, we will study how hints can guide data
 placement as data is handed off from application to storage (during a
@@
 
-Application hints can provide additional information to the runtime based on the users' knowledge of the application workflow as well as on past experiences, for example, possible data access patterns, nature of regions of interest, and domain specific information.
-This application level information will be added to the specification of the workflow and can provide advisory inputs to efficiently help the runtime make smarter decisions about the placement and/or movement of data objects and the scheduling and execution of tasks, as well as to effectively manage trade-offs between for example, power consumption and time-to-solution.
+%Application hints can provide additional information to the runtime based on the users' knowledge of the application workflow as well as on past experiences, for example, possible data access patterns, nature of regions of interest, and domain specific information.
+%This application level information will be added to the specification of the workflow and can provide advisory inputs to efficiently help the runtime make smarter decisions about the placement and/or movement of data objects and the scheduling and execution of tasks, as well as to effectively manage trade-offs between for example, power consumption and time-to-solution.
 
@@
 utilities may be based on the importance of the data objects, the frequency
-with which they are accessed, etc., and are used during runtime data
-management. For example, data objects with higher utility value may be
+with which they are accessed, etc., and can be used during runtime data
+management to evaluate and compare different placement/movement options. 
+For example, data objects with higher utility value may be
 placed closer (e.g., in faster memory) to the accessing application.
@@
 layer, the decision about which object to evict may take the utility into
-account. We will also explore the balancing of the data across the memory
-hierarchy autonomically. The autonomic approach will leverage both user
-hints and information gathered at runtime (e.g., runtime data access
-history, network topology, etc.) to place data object.
+account. Our initial work in defining such a utility is presented in~\cite{tongipdps15}.
+ We will also explore autonomic approaches for balancing of the placement and 
+ movement of data across the memory hierarchy autonomically, which will be driven 
+ by the data utility and will leverage both user hints and information gathered at runtime (e.g., runtime data access
+history, network topology, etc.).
+Finally, we will explore a complementary notion of utility from a system-wide perspective 
+that will capture and quantify, for example, the impact of placement decisions on 
+the overall utilization of storage across multiple applications, and will enable system level runtime optimization.
 
 \paragraph{ Runtime tracking and estimation:}
-Autonomic data placement and/or
-movement will leverage the information gathered at runtime together with
-predictive analysis requirements. This information may include runtime
-spatial-temporal data access patterns, and/or physical network topology
-information, which will be tracked and gathered at runtime, and
-then be used to estimate and predict the behavior of coupled applications in
-the rest of simulation-time. A key research aspect here will be to
-anticipate accesses, for example using learning techniques or using
+Autonomic data placement and/or movement will leverage the information gathered 
+at runtime together with predictive analysis requirements. This information may include 
+runtime spatial-temporal data access patterns, and/or physical network topology
+information, which will be tracked at runtime, and then be used to estimate and predict 
+the data access behaviors of application workflow at runtime. 
+A key research aspect here will be to anticipate accesses, for example using learning techniques or using
 application knowledge, and use this information to prefetch data objects.
@@
 in the target workflows and then use it to characterize patterns at runtime
-and trigger appropriate place/movement actions.
+and trigger appropriate placement and/or movement actions. Our initial 
+work explore these ideas is presented in~\cite{qiansc15}.
 
@@
 % resources.
+
 \subsubsection{Migration}
@@
 \paragraph{Background:} 
-Once the system has completed the initial data placement with minimal
-overhead to the application, it is up to the storage system to manage the
-migration of data to other storage servers and across the hierarchy. An
-important research aspect here is how a liberated Hierarchical Storage
+A complementary aspects of the data lifecycle management is the management and 
+%Once the system has completed the initial data placement with minimal
+%overhead to the application, it is up to the storage system to manage the
+migration of data across storage servers and across the hierarchy. 
+An important research aspect here is how a liberated Hierarchical Storage
 Manager (HSM) designed for on-line use, such as Sirocco, might be employed in
@@
 The policies that govern the migration of data across groups and across
-tiers have similar challenges as initial data placement. The migration
+tiers have similar challenges as data placement. The migration
 policy must encapsulate, at a high level, the constraints defined by the
@@
 
-\subsubsection{Purging data from the system}
+\subsubsection{Purging}
 \label{sec:purging-data}
-{\color{red}(only outline since I have to go)}
-1. Purging is important from the perspective of preventing uncontrolled
-growth of data on a system
-2. Purging is hard because no one wants to lose data and we need to explore
-the policies that govern it
-3. If we have a good utility function and a good way to refactor the data we
-could purge the ``less important'' pieces and maintain a high level of
-knowledge without requiring very high amount of storage
-3. Purging needs to happen in a decentralized way but global utility has to
-be maximized - thats a pretty hard problem
+An important phase of the data lifecycle is the purging of data from the system, 
+which presents significant challenges, both from the systems and policy perspectives. 
+Deciding what data to purge and when to purge it can be non-trivial, and it is 
+essential that this is done with the application in the loop. One approach we will 
+explore in this efforts will be extending the notion of system and application utility. 
+The former can be used to determine when data object need to purged while the 
+latter can help determine the relative importance of data objects to the 
+application and identify which data objects can be purged. Developing policies 
+for purging is a more challenging problem and will required engaging with the 
+user community. However, a key research problem that will be addressed in
+ this effort is developing the abstractions that can be used to consistently and 
+ unambiguously express these policies as well as the mechanisms that can be 
+ used to correctly enforce these policies.
+
+%{\color{red}(only outline since I have to go)}
+%1. Purging is important from the perspective of preventing uncontrolled
+%growth of data on a system
+%2. Purging is hard because no one wants to lose data and we need to explore
+%the policies that govern it
+%3. If we have a good utility function and a good way to refactor the data we
+%could purge the ``less important'' pieces and maintain a high level of
+%knowledge without requiring very high amount of storage
+%3. Purging needs to happen in a decentralized way but global utility has to
+%be maximized - thats a pretty hard problem
 

### 2015-07-09 narrative/data-description.tex
reference fixes

--- 2015-07-09 narrative/data-description.tex
@@
 Our current data annotation techniques demonstrated in
-ADIOS~\cite{lofstead:2009:adaptible} implement a binary-packed (BP) data format that
+ADIOS~\cite{lofstead:2009:adaptible} implement a binary-packed data format that
 allows data characteristics such as min, max and index to be wrapped around
@@
 efficiently, without overburdening users with the management of complex data models, such
-as adaptive mesh refinement (AMR), from the user space. 
-FFS~\cite{ffs} implements a type system that provides highly efficient binary
-data communication and XML-like data description, facilitating data sharing
-between collaborators. Additionally, NetCDF~\cite{netcdf} and HDF5~\cite{hdf5}
-uses a fixed data format to provide a hierarchical data descriptions for scientific data.
-Despite these efforts, the aspect of data importance has yet to be 
+as AMR, from the user space. Despite these efforts, the aspect of data importance has yet to be 
 explored, and we believe it will be vital for exascale storage solutions.
@@
 design choices to ensure stable APIs and useful semantics.
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "../proposal"
+%%% End:

### 2015-07-09 coverpage.tex

@@
 \begin{tabbing}
-\hspace*{10mm} \= \makebox[65mm][l]{\textbf{Applicant Institution:}} \= Oak Ridge National Laboratory \\
+\hspace*{10mm} \= \makebox[65mm][l]{Applicant/Institution:} \= Oak Ridge National Laboratory \\
 \>  \>1 Bethel Valley Road\\
@@
 
-\medskip
+%\medskip
 
-%% Lead PI name, telephone number, email:
+%% Postal Address
 \begin{tabbing}
-\hspace*{10mm} \= \makebox[65mm][l]{\textbf{Lead Principal Investigator:}} \= Scott Klasky,  Distinguished Scientist \\
-\> \> Oak Ridge National Laboratory \\
-\> \> Computer Science and Mathematics Division \\
+\hspace*{10mm} \= \makebox[65mm][l]{Postal Address:} \= P. O. Box 2008 \\
+%\> \> Oak Ridge National Laboratory \\
+%\> \> Computer Science and Mathematics Division \\
 \> \> Oak Ridge, Tennessee 37831-6057 \\
-\> \> (T) 865-241-9980 (Fax) 865-241-4811 \\
-\> \> Email: klasky@ornl.gov
+%\> \> (T) 865-241-9980 (Fax) 865-241-4811 \\
+%\> \> Email: klasky@ornl.gov
 \end{tabbing}
@@
 
-\medskip
+%\medskip
 
@@
 \begin{tabbing}
-\hspace*{10mm} \= \makebox[65mm][l]{\textbf{Administrative Point of Contact:}} \= Barney Maccabe \\
-\> \> Oak Ridge National Laboratory \\
-\> \> Computer Science and Mathematics Division \\
-\> \> Oak Ridge, TN 37831-6057 \\
-\> \> (T) 865-241-6504 \\
+\hspace*{10mm} \= \makebox[65mm][l]{Administrative POC:} \= Barney Maccabe \\
+%\> \> Oak Ridge National Laboratory \\
+%\> \> Computer Science and Mathematics Division \\
+%\> \> Oak Ridge, TN 37831-6057 \\
+\> \> Phone: 865-241-6504 \\
 \> \> Email: maccabe@ornl.gov
@@
 
-\bigskip
+%\medskip
+
+%% Lead PI name, telephone number, email:
+\begin{tabbing}
+\hspace*{10mm} \= \makebox[65mm][l]{Lead PI:} \= Scott Klasky, Oak Ridge National Laboratory \\
+%\> \> Oak Ridge National Laboratory \\
+%\> \> Computer Science and Mathematics Division \\
+%\> \> Oak Ridge, Tennessee 37831-6057 \\
+%\> \> (T) 865-241-9980 (Fax) 865-241-4811 \\
+%\> \> Email: klasky@ornl.gov
+\end{tabbing}
+%\vrule height0.4pt depth 0pt width 2.5in \\
+
+%\medskip
 
 \begin{tabbing}
-\hspace*{10mm} \= \makebox[65mm][l]{\textbf{Program Announcement Number:}} \= LAB 15-1338\\
-\> \textbf{PAMS Tracking No.:}			    \> 0000219877\\
-\> \textbf{DOE/Office of Science Program Office:} \> Office of Advanced Scientific Computing Research\\
-\>									    \> Storage Systems and Input/Output for Extreme Scale Science\\
-\> \textbf{DOE/Office of Science Program Office}  \> \\
-\> \makebox[64mm][r]{\textbf{Technical Contact:}} \>  Lucy Nowell\\
+\hspace*{10mm} \= \makebox[65mm][l]{DOE National Lab Announcement No.:} \= \textbf{LAB 15-1338}\\
 \\
-\> \textbf{Research Area:} \>Scalable Storage Software Infrastructure
+\> {PAMS Letter of Intent Tracking Number:}			    \> \textbf{0000219877}\\
+\\
+\> {DOE/Office of Science Program Office:} \> \textbf{Office of Advanced Scientific Computing}\\
+\>									    \> \textbf{Research Storage Systems and Input/Output}\\
+\>									    \> \textbf{(SSIO) for Extreme Scale Science}\\
+\\
+\> {DOE/Office of Science}  \> \\
+\> \makebox[64mm][r]{Program Office Technical Contact:} \>  \textbf{Lucy Nowell}\\
+\\
+\> {Research Area:} \>\textbf{Scalable Storage Software Infrastructure}
 

### 2015-07-09 narrative/apis.tex

@@
 
-The new APIs will provide new functionality of specifying selectable performance/quality/- cost tradeoffs from both the application and system perspectives based upon the user guided rules/policy and runtime system monitoring status.
+The new APIs will provide new functionality of specifying selectable performance/quality/- cost trade-offs from both the application and system perspectives based upon the user guided rules/policy and runtime system monitoring status.
 It allows the middleware to make best possible decisions from the feedback of storage system knowledge, such that it will embed user intentions and the available system storage.
@@
 
+We will extend ADIOS with new APIs to provide such functionalities.
 If user decide to write or read data to or from hierarchical storage systems, they will rely on the API to send their intentions for inquiry and examine the system status, including how much data they would like to write/read, desired bandwidth, data compression method, etc, or the user can express their intension of writing data right now no matter what the traffic is now.

### 2015-07-09 narrative/enabling_tech.tex

@@
 classes~\cite{watkins:ucsctr15} will provide the main mechanisms of introducing
-new APIs, metadata vocabulary, and quality of service into the storage stack.
+new APIs, metadata vocabulary, and QoS into the storage stack.

### 2015-07-09 narrative/intro.tex

@@
 \begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{2.2cm} |}\hline
-\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant S2E2 work} & \multicolumn{1}{|c|}{\bf See} \\\hline\hline
+\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant S2E2 work} & \multicolumn{1}{|c|}{\bf Proposed research} \\\hline\hline
 1 & Data abstractions & Create new APIs  and annotations for users to guide the system for placement  & \S\ref{sec:data-description} \\\hline

### 2015-07-09 narrative/management.tex

@@
 will lead the Storage System. This project is focused on the tight integration of application aware-middleware and a storage system backend.
-Our team has a strong track record of tight-collaboration, with both Drs. Abbasi and Lofstead both working with Dr. Klasky for their Ph.D. at the Georgia
+Our team has a strong track record of tight-collaboration, with both Drs. Abbasi and Lofstead working with Dr. Klasky for their Ph.D. at the Georgia
 Institute of Technology.  Our structure includes an area lead for each major research area plus
@@
 form the management team, which will meet weekly to review progress and revise plans as needed.
-We will also have a {\bf Science Advisory Board} consisting of one application Scientist, Dr. C. S. Chang (PPPL), one storage system expert from the 
-OLCF, Dr. Vazhkudai (ORNL) who we will advice us yearly on our project goals and outcomes and make adjustments accordingly. We will also have our
-lead Math Lead, Dr. Ainsworth, to work with the applied mathematics community to ensure that the hierarchical decomposition of data will work in theory
+We will also have a {\bf Science Advisory Board} consisting of one application scientist, Dr. C. S. Chang (PPPL), and one storage system expert from the 
+OLCF, Dr. Vazhkudai (ORNL), who will advise us yearly on our project goals and outcomes and make adjustments accordingly. We will also have our
+Math Lead, Dr. Ainsworth, to work with the applied mathematics community to ensure that the hierarchical decomposition of data will work in theory
 for current and next generation exascale simulations.
@@
 scheduling, and running critical post hoc reviews of integration tests will be coordinated by the Integration Lead, Dr. Parashar. 
-These meetings will bring all hands together at least once a year.
+These meetings will bring all hands together at least once per year.
 
 We will monitor and report {\bf success metrics} including: 
-research publications;
-our ability to predict and explain performance of I/O workloads on the current machines., in different environments;
+research publications, 
+our ability to predict and explain performance of I/O workloads on the current machines in different environments, 
 our ability to reduce data and characterize data from an object into multiple bins, 
-our ability to integrate our routines into the Sirocco file system and test the feasibily on current and next generation LCF systems, and
+our ability to integrate our routines into the Sirocco file system and test the feasibility on current and next generation LCF systems, and
 how tools lead to workflow performance improvements, for both reading and writing data from our exemplar applications which we will test.

### 2015-07-09 narrative/intro.tex
changed the acronym.

--- 2015-07-09 narrative/intro.tex
@@
 \begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{2.2cm} |}\hline
-\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant S2E2 work} & \multicolumn{1}{|c|}{\bf See} \\\hline\hline
+\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant SIRIUS work} & \multicolumn{1}{|c|}{\bf Proposed research} \\\hline\hline
 1 & Data abstractions & Create new APIs  and annotations for users to guide the system for placement  & \S\ref{sec:data-description} \\\hline

### 2015-07-09 narrative/data-description.tex
minor changes

--- 2015-07-09 narrative/data-description.tex
@@
 The user can pick a data-refactoring scheme (see
-Sec~\ref{sec:data-refactor}) which will classify the particle data into four different
+\S~\ref{sec:data-refactor}) which will classify the particle data into four different
 prioritization levels. The user can then pick a data utility, such as three months
@@
 to re-prioritize their data; 2) The number of tiers in the storage system will
-also be small (e.g. NVRAM, PFS, Campaign Storage, HPSS), 3) The default
+be small (e.g. NVRAM, PFS, Campaign Storage, HPSS), 3) The default
 lifetime of objects in each tier will be specified by the resource by
@@
 next CR file appears on the system or the next two CR files. Research in
-describing this utility function will be in conjunction with our users
+describing this utility function will be in conjunction with users
 and we will evaluate how policies can be placed in the system and how well

### 2015-07-09 narrative/intro.tex
minor changes

--- 2015-07-09 narrative/intro.tex
@@
 3 & Mechanisms for data movement across the storage hierarchy & Manage data migration based on policies& \S \ref{sec:managing-data-life},  \S\ref{sec:migration} \\\hline
-4 &  Extend the interaction of I/O middleware with the resource management system & We will manage data from the application  to the middleware to the storage system and allow them to have an autonomic interaction & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
-5 & Design and implement of I/O middleware architectures  & We will create APIs to guide the placement of chunks of data across the SSIO and autonomically move data when necessary& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
-6 & Approaches to improve the ability of SSIO software to support checkpoint/restart & We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then purged when it is no longer needed & \S \ref{fig:ssio-bucket}  \\
+4 &  Extend the interaction of middleware with the storage system & Manage data from the application  to middleware to the storage system using an autonomic interaction & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
+5 & Design and implement of I/O middleware architectures  & Create APIs to guide the placement of chunks of data across the SSIO and autonomically move data& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
+6 & Approaches to improve the ability of SSIO software to support checkpoint/restart & Define a utility functions for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then purged when it is no longer needed & \S \ref{fig:ssio-bucket}  \\
 \hline
@@
 As these particulars were realized, significant modifications in the plans were made.
-Eventually, they were forced to make very hard decisions in reducing the desired 100~PB over a 10-day run to only 5~PB of data.
+Eventually, they were forced to make very hard decisions in reducing the desired 100~PB to only 5~PB of data.
 

### 2015-07-09 narrative/migration.tex
start to reshape the proposal
scott changes

@@
-\subsubsection{Migration}
+\subsection{Migration}
 
@@
 In addition to these static attributes, when grouping servers into
-pseudo-tiers, Sirocco also may employ more dynamic attributes such as CPU load
+pseudo-tiers, Sirocco may also employ more dynamic attributes such as CPU load
 and the rate-of-change of the media use.
@@
 at present. In classic HSM the tiering is often a formally expressed concept,
-embodied both in the architecture and the implementation. In Sirocco, though,
-it is not directly expressed. Instead, a client or server that attempts to make
+embodied both in the architecture and the implementation. In Sirocco
+it is not directly expressed; instead, a client or server that attempts to make
 a decision relative to migration and staging uses the attributes discussed

### 2015-07-09 narrative/enabling_tech.tex
added Sirocco ref

@@
 %devices. This project, 
-{\bf Sirocco} came out of the LWFS project and offers a peer-to-peer style data storage device
+{\bf Sirocco}~\cite{sirocco} came out of the LWFS project and offers a peer-to-peer style data storage device
 mesh with data migrating based on use, required resilience, and available

### 2015-07-09 narrative/qos.tex

@@
 \subsubsection{Resource Management}
-\paragraph{Background:} There is an important connection between finding the resources
+There is an important connection between finding the resources
 that can fulfill a particular request and the performance predictability of that request. As such, this demands
@@
 scalability, because they require a single entity that tracks updates in the system. 
-Traditional file systems are good examples for this - most updates to files require updates 
+Traditional file systems are good examples for this ; most updates to files require updates 
 to their centrally maintained block allocation tables. 
@@
 \subsubsection{Quality of Service (QoS)}
-\paragraph{Background:} QoS refers to the
+QoS refers to the
 properties of the performance quality of a particular service, in
@@
   \centering
-  \includegraphics[width=\columnwidth]{graphics/infrastructure.pdf}\vspace{-1.2in}
+  \includegraphics[width=\columnwidth]{graphics/infrastructure.pdf}
   \caption{Resource usage distribution for OST(a), OSS(b) and LNETs(c) }
 \end{figure}
-\vskip -1.5in
+\vskip-0.2in
 

### 2015-07-09 narrative/placement.tex

@@
 
-Our proposed research will address data management challenges across the entire 
+Our research will address data management challenges across the entire 
 data life cycle in an application driven and system aware manner. 
@@
 (2) How can the placement and migration of data across a multi-tiered storage hierarchy be 
-optimized at runtime, both from the application and system perspective?, {\color{red} where is 3?} 
+optimized at runtime, both from the application and system perspective?,
 %(3) How can we migrate data throughout the layers of storage
 %without making the discovery process unbounded? 
-(4) How can knowledge about the application used to better prepare the data for consumption?, and 
-(5) When and how do we make the decision to purge data?  %using this additional knowledge?
+(3) How can knowledge about the application used to better prepare the data for consumption?, and 
+(4) When and how do we make the decision to purge data?  %using this additional knowledge?
 
@@
 \label{sec:init-plac-data}
-
-\paragraph{Background:}
 When an application outputs or accesses data the storage and middleware layers needs to
@@
 %
-Our approach in the proposal is to increase knowledge about the data and its use within the 
+Our approach is to increase knowledge about the data and its use within the 
 application, and leverage this knowledge to drive data placement and overall management.
@@
 and trigger appropriate placement and/or movement actions. Our initial 
-work explore these ideas is presented in~\cite{qiansc15}.
+work explore these ideas is presented in~\cite{qiansc15,choi2012mining}.
 
@@
 \label{sec:migration}
-
-\paragraph{Background:} 
 A complementary aspects of the data lifecycle management is the management and 
@@
 expressed concept, embodied both in the architecture and the implementation.
-In Sirocco, though, it is not directly expressed. Instead, a client or
+In Sirocco, though, it is not directly expressed; instead, a client or
 server that attempts to make a decision relative to migration and staging
@@
 application, but must also manage resources on each server for capacity and
-load reasons. We will explore policy formulations that allow \Sir to move
-data within a tier to continue to meet service objectives. Another component
+load reasons. We will explore policy formulations that allow \Sir~to move
+data within a tier and continue to meet service objectives. Another component
 of migration is the temporal function that determines the utility of data
-through its life time on storage. If they utility decays over time, for
+through its life time on storage. If the utility decays over time, for
 example, the initial constraints of data (such as resilience requirements)
-will be modified and allow more leeway in how the system handles migration.
-However, as data is widely distributed the scalability of this approach must
-be investigated. 
+will be modified and allow more leeway in how the system will handle migration.
+However, as data is widely distributed {\it the scalability of this approach must
+be investigated}. 
 We will explore how the utility function can be used to modify migration
@@
 are not scalable. In order to have predictable performance the lookup
-process must complete in a bounded time. Another solution here is to use a
+process must complete in a bounded time. Our potential solution here is to use a
 distributed hashing function to determine exactly where data is placed,
@@
 investigate how the migration of data can be performed in a way that enables
-bounded estimates about time to first byte for a potential read operation.
+bounded estimates  for a potential read operation.
 
@@
 \label{sec:manag-data-mult}
-
-\paragraph{Background:} Providing data to users for knowledge discovery is
+ Providing data to users for knowledge discovery is
 one of the primary purposes of our storage system design. There is a
@@
 this can be accomplished: On one end the storage system can aim to maintain
-full resolution data in the highest tiers of the storage hierarchy. This
+fully accurate data in the highest tiers of the storage hierarchy. This
 approach is, obviously, limited due to the disparity between size of data
@@
 the data that, with user defined error, represent the data, and place these
-smaller data sets on the higher tiers for faster access.
+high priority  data sets on the higher tiers for faster access.
 
@@
 the value of thinning or enriching data at a particular tier.
-Consider the example of a road navigation system: the locality
-properties of each navigation client are based on physical constraints
-of speed, direction, likely resolution of map, and the fact that
-clients almost never go off-road.
+%Consider the example of a road navigation system: the locality
+%properties of each navigation client are based on physical constraints
+%of speed, direction, likely resolution of map, and the fact that
+%clients almost never go off-road.
 
-\subsubsection{Purging}
+\subsubsection*{Purging}
 \label{sec:purging-data}
@@
 application and identify which data objects can be purged. Developing policies 
-for purging is a more challenging problem and will required engaging with the 
+for purging is a more challenging problem and will require engaging  the 
 user community. However, a key research problem that will be addressed in

### 2015-07-09 narrative/proposed-work.tex
more changes for readability.

@@
 
+One of the underlying principles of our approach is that we recognize that there are a small set
+of important I/O patterns to optimize for~\cite{lofstead2011six,polte2009and,tian2011edo,tian2012system},
+and our representative applications have worked with us on all of these I/O patterns. Throughout our project
+we will continuously focus on creating research artifacts which can optimize these patterns on current generation through
+exascale systems.  One important pattern is Checkpoint/Restarts (CR) and we will focus on optimizing these through our
+research, and also using many of the important research techniques for using NVRAM and neighbor's NVRAM for storing 
+the CR data. The migration and purging of this will be optimized by our SSIO layer, and we will have the user data intentions
+signify the utility of CR data so that this data can be high-priority but stored for short periods of time.
+
 %\underline{Principle 3:} {\bf Users goals define data policies} In a data-centric SSIO system. 

### 2015-07-09 narrative/qos.tex
more changes for readability.

@@
 
-\subsubsection{Resource Management}
+\subsubsection{(Resource Management}
 There is an important connection between finding the resources

### 2015-07-09 narrative/placement.tex
reduce the repitition a bit in the utility function definition.

@@
 %This application level information will be added to the specification of the workflow and can provide advisory inputs to efficiently help the runtime make smarter decisions about the placement and/or movement of data objects and the scheduling and execution of tasks, as well as to effectively manage trade-offs between for example, power consumption and time-to-solution.
-
-\paragraph{Defining data utility:} We will formulate the concept of utility (from an
-application perspective) that is assigned data objects to quantify the
-benefits of placement and/or movement actions to application. These
-utilities may be based on the importance of the data objects, the frequency
-with which they are accessed, etc., and can be used during runtime data
-management to evaluate and compare different placement/movement options. 
-For example, data objects with higher utility value may be
-placed closer (e.g., in faster memory) to the accessing application.
-Similarly, when it is time to evict data objects from a higher level storage
-layer, the decision about which object to evict may take the utility into
-account. Our initial work in defining such a utility is presented in~\cite{tongipdps15}.
- We will also explore autonomic approaches for balancing of the placement and 
- movement of data across the memory hierarchy autonomically, which will be driven 
- by the data utility and will leverage both user hints and information gathered at runtime (e.g., runtime data access
-history, network topology, etc.).
-Finally, we will explore a complementary notion of utility from a system-wide perspective 
-that will capture and quantify, for example, the impact of placement decisions on 
-the overall utilization of storage across multiple applications, and will enable system level runtime optimization.
 

### 2015-07-09 narrative/qos.tex
tied together and smoothed wording a bit.

--- 2015-07-09 narrative/qos.tex
@@
 \subsubsection{Resource Management}
-There is an important connection between finding the resources
-that can fulfill a particular request and the performance predictability of that request. As such, this demands
-careful resource management strategies that can deal with placement and retrieval of 
-a large number of data chunks spread across the entire storage hierarchy. 
-%This connection is particularly important for data retrieval 
-%- in large scale storage systems, finding a resource to read can be a significant, if not dominant, 
-%component of overall read latency. 
-Locating resources is a typical distributed systems problem. Centralized solutions are limited in their
-scalability, because they require a single entity that tracks updates in the system. 
-Traditional file systems are good examples for this ; most updates to files require updates 
-to their centrally maintained block allocation tables. 
-The opposite of file systems are ``pile'' systems, which, instead of keeping track of updates, searchs for 
-resources at request time. Pile systems allow very efficient writes because clients can write wherever
-there is free space and there is very low bookkeeping overhead. However, unbounded search in pile systems
-is possible, e.g., when it is unclear whether the resource actually exists. 
-Methods to make searches more efficient without reverting to centralized bookkeeping usually involves 
-partitioning of the search space and a method to limit the search to a small number of partitions.
+Careful resource management strategies that can deal with placement and
+retrieval of a large number of data chunks spread across the entire storage
+hierarchy are key for predictable performance. 
+%This connection is particularly important for data retrieval - in large scale
+%storage systems, finding a resource to read can be a significant, if not
+%dominant, component of overall read latency. 
+Locating resources is typical a distributed systems problem. Centralized
+solutions suffer scalability challenges due to centralized, serialized system
+updates.  POSIX-compliant file systems are good examples of this
+strucutre--most updates to files require updates to their centrally maintained
+block allocation tables.
 
-Frequently this involves a hash function that is relatively stable against changes to its range, 
-such as failures or updates to the hardware of the system. Consistent hashing~\cite{karger:stoc97}
-is scalable since it can be well-known throughout a system of any size and only requires minimal 
-movement of data in the presence of failures or system changes. 
-%Consistent hash functions can also
-%have sets of ordered but infinite lists as their range (which works well with primary replication)
-%~\cite{honicky:ipdps04}, implement declustering (limiting overlap between partitions), and support
-%placement policy rules~\cite{weil:sc06}.
+An opposite example of this central management strategy are ``pile'' systems,
+which, instead of tracking updates, they search for resources upon request.
+Pile systems feature efficient writes because clients can write wherever there
+is free space with low bookkeeping overhead.  However, unbounded search is
+possible, e.g., when it is unclear whether the resource actually exists.
+Methods to make searches more efficient without reverting to centralized
+bookkeeping usually involves search space partitioning and a method to limit
+the search to a small number of partitions.  Frequently this involves a
+hash function that is relatively stable against changes to its range, such as
+failures or updates to the hardware of the system. Consistent
+hashing~\cite{karger:stoc97} is scalable since it can be well-known throughout
+a system of any size and only requires minimal movement of data in the presence
+of failures or system changes. 
+%Consistent hash functions can also have sets of ordered but infinite lists as
+%their range (which works well with primary replication)
+%~\cite{honicky:ipdps04}, implement declustering (limiting overlap between
+%partitions), and support placement policy rules~\cite{weil:sc06}.
+
+Sirocco offers more of a pile struture rather than a file structure. One of the
+challenges we face is offering performance predictability more similar to
+POSIX-compliant file systems while we take advantage of the distributed,
+flexible storage management offered by Sirocco.
 
 \paragraph{Approach:} 
-In this proposal we will investigate the trade-off in terms of predictable performance between 
-``filing'' data, making an upfront investment to ensure finding data quickly, and ``piling'' 
-data, pushing the cost of finding to the moment when the data is actually needed. This trade-off 
-has a parallel in data management where there are two alternative paths to achieve knowledge: 
-the analogue to filing is  to ``extract, transform, and load'' (ETL) data into a warehouse, paying 
-an upfront ingest cost to achieve fast querying; and the analogue to piling is to not pay any upfront 
-cost and later querying   the raw data directly by using a data processing platform such as Hadoop.  Recent work 
-has found that the combination of these two techniques can be optimized to perform better than 
-any one technique alone~\cite{lefevre:sigmod14a}. We believe that a similar opportunity exists 
-for combining filing and piling approaches to achieve performance predictability without sacrificing 
-too much flexibility.
+In this proposal, we will investigate the trade-off in terms of predictable
+performance between ``filing'' and ``piling'' data.  This trade-off has a
+parallel in data management where there are two alternative paths to achieve
+knowledge. The analogue to filing is  to ``extract, transform, and load'' (ETL)
+data into a warehouse paying an upfront ingest cost to achieve fast querying.
+The analogue to piling is to not pay any upfront cost and later querying the
+raw data directly by using a data processing platform such as Hadoop.  Recent
+work has found that the combination of these two techniques can be optimized to
+perform better than any one technique alone~\cite{lefevre:sigmod14a}. We
+believe that a similar opportunity exists for combining filing and piling
+approaches to achieve performance predictability without sacrificing too much
+flexibility. Our additional challenge is incorporating both the storage tiers
+and potential data quality metrics.
 
-More specifically, the coarseness of a search space partitioning limits the search time within 
-a partition, i.e. finely partitioned search spaces require little or no search within a partition 
-but offer little placement flexibility. Partitioning can also be fine in one dimension, and coarse 
-in another to allow flexibility where it is most needed while still ensuring fast finding. For 
-example, a hash function can finely partition a low storage tier but each partition can extend 
-across multiple tiers so that a search still has to occur in order to find a resource on a particular 
-tier.
+More specifically, the coarseness of a search space partitioning limits the
+search time within a partition, i.e., finely partitioned search spaces require
+little or no search within a partition but offer little placement flexibility.
+Partitioning can also be fine in one dimension and coarse in another to allow
+flexibility where it is most needed while still ensuring fast finding. For
+example, a hash function can finely partition a low storage tier but each
+partition can extend across multiple tiers so that a search must occur in order
+to find a resource on a particular tier.
 
@@
 \subsubsection{Quality of Service (QoS)}
-QoS refers to the
-properties of the performance quality of a particular service, in
-this case the storage system.
+Our ultimate goal, rather than just performance predictability, is to offer
+predictable Quality of Service (QoS) for the storage system.
 %Performance quality can be expressed in either relative or absolute
@@
 One of the key differentiating features this project offers is driving storage
-decisions through QoS \emph{reservations}, rather than continuing the current model
-of allowing applications to compete freely without restriction, for part or all of the storage
-resources. This traditional approach leads to interference effects
-\cite{lofstead:2010:io-variability,liu_hotstorage} that can greatly impact I/O performance predictability. 
+decisions through QoS~\emph{reservations} rather than continuing the current
+model of allowing applications to compete freely without restriction for part
+or all of the storage resources. The current approach leads to interference
+effects~\cite{lofstead:2010:io-variability,liu_hotstorage} that can greatly
+impact I/O performance predictability. 
 %Further, with the introduction of additional tiers in the
@@
 %competiion for limited high performance resources. 
-%% Temporaily commented out the above sentence, because it may not be true for the next DOE systems. 
-Unfettered, this competition can yield a lower system and application performance 
-than the case where applications attempt to reserve storage resources, thereby enabling the storage system
-to reason over its performance in terms of reservations and actual
+%% Temporaily commented out the above sentence, because it may not be true for
+%the next DOE systems. 
+Unfettered, this competition can yield lower system and application performance
+than if applications attempt to reserve storage resources thereby enabling the
+storage system to reason over its performance in terms of reservations and
+actual performance.  In the cases where an application arrives that needs to
+run more urgently than any of the other already running applications, the
+system can rapidly respond to it by reducing the amount of resources taken by
+less urgent applications, allowing the urgent application to finish, and later
+reverting back to the previous committed resource level for the rest.  This can
+be automatically enforced via user- or administrator- specified
+\textit{policies}\S\ref{sec:data-description}.
+
+There are two fundamental causes that can impede the I/O performance in an
+extreme-scale SSIO system. One is indirection, where layers of storage tiers
+are employed either for performance and/or for scalability reasons.  The direct
+consequence is that there could be multiple traversing paths from application
+end to the data resting place, and more often than not, the I/O paths are not
+under control under of any single authority. The other cause is shared
+resources use. The best effort I/O request/response nature and lack of QoS
+mechanisms imply that there is little expected performance guarantee.  Both
+indirection and shared resources use contribute to a high probability of
+imbalanced resources use and thus the occurrence of congestion and degraded
 performance.
-In the cases where an application arrives that needs to run more urgently
-than any of the other already running applications, the system
-can rapidly respond to it by reducing the amount of resources taken 
-by less urgent applications, allowing the urgent application to finish,
-and later reverting back to the previous committed resource level for the rest.
-This can be automatically enforced via user- (or administrator-) specified \textit{policies}\S\ref{sec:data-description}.
 
-There are two fundamental causes that can impede the I/O performance
-in an extreme-scale SSIO system. One is indirection, where layers
-of storage tiers are employed either for performance and/or for
-scalability reasons.  The direct consequence is that there could
-be multiple traversing paths from application end to the rest place
-of data, and more often than not, the I/O paths are not under control
-under any single authority. The other cause is the shared use of
-resources. The best effort I/O request/response nature and lack of
-QoS mechanisms imply that there is little guarantee in terms of
-expected performance. Both indirection and shared use of resource
-contribute to a high probability of imbalanced use of resources
-thus the occurrence of congestion and degraded performance.
-
-As an example to demonstrate that this disconnection will hurt
-system performance: We launch 4096 processes with each process doing
-a single file I/O operation against half of the Spider II file
-system. The traces of those files are analyzed to examine the
-utilization distribution of different components. Figure (a), (b)
-and (c) shows the resource usage distribution for OSTs, OSSes, and
-LNETs, respectively. We observe that there exists a significant
-variation in usage across components of any given type (e.g., OST,
-OSS or LNET). For example, some OSTs are used more than 10 times
-while some others are never used (corresponding to zero frequency
-count). Similarly, OSSes and LNETs show significant imbalance in
-usage under the default placement strategy. Consequently, imbalanced
-resource utilization increases the contention at certain components
-more than others.
+Consider this example demonstrating this disconnection will hurt system
+performance. We launch 4096 processes with each process doing a single file I/O
+operation against half of the Spider II file system at ORNL. The file traces
+are analyzed to examine the utilization distribution of different components.
+Figures (a), (b) and (c) show the resource usage distribution for OSTs, OSSes,
+and LNETs, respectively. We observe a significant usage variation
+across components of any given type (e.g., OST, OSS or LNET). For example, some
+OSTs are used more than 10 times while some others are never used
+(corresponding to zero frequency count). Similarly, OSSes and LNETs show
+significant usage imbalance under the default placement strategy.
+Consequently, imbalanced resource utilization increases the contention at
+certain components more than others.
 
@@
 
-% This is about finding the relationshiop of workload-dependent to workload-independent metrics. -- Carlos
+% This is about finding the relationshiop of workload-dependent to
+% workload-independent metrics. -- Carlos
 %are driven from multiple different approaches. First, supporting
@@
 
-\paragraph{Approach:} The goals of this proposal require applications
-to demand guarantees in terms of absolute performance qualities.
-The first key enabler for any service to make absolute guarantees
-is \emph{performance isolation}, i.e. the ability to shield the
-performance impact of one task from another. The most straight-forward
-but least flexible strategy is to simply cap the latency of each
-task, regardless of its completeness (see for
-example~\cite{decandia:sosp07}). This strategy is applicable when
-incomplete tasks still have value and missing results are either
-not important or can be easily retrieved in subsequent tasks. A
-more generally applicable strategy is to control the admission of
-tasks to the system: \emph{admission control} maintains a
+\paragraph{Approach:}
+The goals of this proposal require applications to demand guarantees in terms
+of absolute performance qualities.  The first key enabler for any service to
+make absolute guarantees is \emph{performance isolation}, i.e., the ability to
+shield the performance impact of one task from another. The most
+straight-forward but least flexible strategy is to simply cap the latency of
+each task, regardless of its completeness (see for
+example~\cite{decandia:sosp07}). This strategy is applicable when incomplete
+tasks still have value and missing results are either not important or can be
+easily retrieved in subsequent tasks. A more generally applicable strategy is
+to control task admission to the system. \emph{Admission control} maintains a
 \emph{utilization model} that can predict the utilization of system
-resources given a new task, estimates the current utilization of
-the system, and decides whether the task can be admitted without
-overloading the system. Thus, admission control avoids system
-overload conditions that lead to hard-to-model chaotic behaviors.
-Consequently, utilization models can frequently be approximated by
-automatically calibrated linear models (see for
+resources given a new task, estimates the current utilization of the system,
+and decides whether the task can be admitted without overloading the system.
+Thus, admission control avoids system overload conditions that lead to
+hard-to-model chaotic behaviors.  Consequently, utilization models can
+frequently be approximated by automatically calibrated linear models (see for
 example~\cite{skourtis:hpdc12}).
 
-Another essential component of performance isolation is the
-\emph{charging model} which determines which task(s) should be
-charged how much for a request. Performance isolation is implemented
-by the charging model by ensuring that performance of a task only
-depends on its reservation and its workload behavior (but not on
-other workloads in the system). Each task has a budget based on its
-reservation and spends it based on the cost of each request determined
-by the charging model. A charging model determines the cost based
-on the utilization model and the interactions a request can have
-with other requests (e.g. a random request to spinning media
-interrupting a stream of sequential requests).
+Another essential component of performance isolation is the \emph{charging
+model} that determines which task(s) should be charged how much for a request.
+Performance isolation is implemented by the charging model by ensuring that
+task performance only depends on its reservation and its workload behavior (but
+not on other workloads in the system). Each task has a budget based on its
+reservation and spends it based on the cost of each request determined by the
+charging model. A charging model determines the cost based on the utilization
+model and the interactions a request can have with other requests (e.g. a
+random request to spinning media interrupting a stream of sequential requests).
 
-Performance isolation (including its utilization and charging models)
-critically depends on \emph{workload-independent metrics}. Even
-though throughput and latency are frequently the most meaningful
-metrics for an application, they are not workload-independent and
-therefore are inadequate for measuring the utilization of a device
-across all workloads, unless one accepts worst-case estimates that
-can lead to gross underutilization of resources (e.g. the throughput
-of random and sequential workloads in spinning media differ by
-orders of magnitudes). A common workload-independent metric is
-\emph{time utilization}, i.e. the amount of time a resource is
-utilized in a given time interval. Time utilization has the advantage
-of having an always defined maximum and therefore makes resources
-fully reservable. Even though a workload-independent metric might
-at first not appear meaningful for an application, given a particular
-workload an application can discover the relationship between its
-workload-dependent metric and the workload-independent metric.
-Because of performance isolation, the application has to discover
-this relationship only once.
-To achieve effective performance isolation, we seek to expand the resource management
-capabilities in Sirocco to support additional utilization 
-metrics to guide resource management
-decisions. Currently, Sirocco offers automatic resource management capabilities,
-based on resource capacity and data resilience requirements, and it needs to be extended
-to offer data-centric metrics that help guide the forced QoS decisions.
+Performance isolation, including its utilization and charging models,
+critically depends on \emph{workload-independent metrics}. Even though
+throughput and latency are frequently the most meaningful metrics for an
+application, they are not workload-independent and therefore are inadequate for
+measuring device utilization across all workloads unless one accepts
+worst-case estimates that can lead to gross resource underutilization. For
+example, the throughput of random and sequential workloads in spinning media
+differ by orders of magnitudes. A common workload-independent metric is
+\emph{time utilization}, i.e., the amount of time a resource is utilized in a
+given time interval. Time utilization has the advantage of having an always
+defined maximum and therefore makes resources fully reservable. Even though a
+workload-independent metric might at first not appear meaningful for an
+application, given a particular workload an application can discover the
+relationship between its workload-dependent metric and the workload-independent
+metric.  Because of performance isolation, the application has to discover this
+relationship only once.  To achieve effective performance isolation, we seek to
+expand the resource management capabilities in Sirocco to support additional
+utilization metrics to guide resource management decisions. Currently, Sirocco
+offers automatic resource management capabilities, based on resource capacity
+and data resilience requirements, and it needs to be extended to offer
+data-centric metrics that help guide the forced QoS decisions.
 
@@
 
-\paragraph{Related Work:} On current production HPC systems, absolute I/O performance
-guarantees are not possible. Previous work \cite{lofstead:2010:io-variability,liu_hotstorage} 
-from this project team uses a client-assisted approach to mitigate I/O variablity at scale (100,000-core).
-Recently server side QoS scheduling has also been studied for HPC applications \cite{Dai:2014}.
-However, the key hurdle of scalability has not been solved.
-Existing efforts on scheduing resources~\cite{thapaliya:2014:io-cop,dorier:2014:calciom} attempt to offer
+\paragraph{Related Work:}
+On current production HPC systems, absolute I/O performance guarantees are not
+possible. Previous work~\cite{lofstead:2010:io-variability,liu_hotstorage} from
+this project team uses a client-assisted approach to mitigate I/O variablity at
+scale (100,000-core).  Recently, server side QoS scheduling has also been
+studied for HPC applications~\cite{Dai:2014}.  However, the key hurdle of
+scalability has not been solved.  Existing efforts on scheduing
+resources~\cite{thapaliya:2014:io-cop,dorier:2014:calciom} attempt to offer
 admission control to maximize storage performance. Unfortunately, these efforts
@@
 platform. For this approach to be effective, the storage system must offer an
-approach that handles clients from all connected platforms and does not
-require modification to use new storage access APIs. By using a priority
-tagging system, we will offer defaults suitable for all applications that can
-be informed by simple extensions to the job scheduler or more advanced
-management through new APIs. On the other hand, storage QoS has been
-studied for enterprise applications and systems \cite {Gulati:2007,Gulati:2010,Gulati:2012}
+approach that handles clients from all connected platforms and does not require
+modification to use new storage access APIs. By using a priority tagging
+system, we will offer defaults suitable for all applications that can be
+informed by simple extensions to the job scheduler or more advanced management
+through new APIs. On the other hand, storage QoS has been studied for
+enterprise applications and systems~\cite{Gulati:2007,Gulati:2010,Gulati:2012}
 in virtualized environments. It is not straightforward to adapt these solutions
-to HPC storage systems because of the scalability and efficiency concerns, despite
-they achieved very good performance isolations.
+to HPC storage systems because of the scalability and efficiency concerns,
+despite they achieved very good performance isolations.
 
-\paragraph{Challenges:} The deep and heterogeneous memory and storage
-hierarchy we are assuming for this proposal complicates the
-relationship between workload-dependent and workload-independent
-metrics: the performance of a task can significantly differ depending
-on what levels of the hierarchy are involved. Furthermore, an
-important goal of the proposed project is to enable applications
-to reason about the trade-off between resolution and latency, adding
-yet another dimension to how tasks are mapped to resources.
+\paragraph{Challenges:}
+The deep and heterogeneous memory and storage hierarchy we are assuming for
+this proposal complicates the relationship between workload-dependent and
+workload-independent metrics: the performance of a task can significantly
+differ depending on what levels of the hierarchy are involved. Furthermore, an
+important goal of the proposed project is to enable applications to reason
+about the trade-off between resolution and latency, adding yet another
+dimension to how tasks are mapped to resources.
 

### 2015-07-09 narrative/data-description.tex
more changes for readability.

@@
 \begin{wrapfigure}{r}{0.6\textwidth}
-        \vspace{-6ex}
         \begin{centering} 
@@
 
-The assumptions which we will make in choosing a utility function is that:
-1) There will only be a small number of priority levels that users will use
-to re-prioritize their data; 2) The number of tiers in the storage system will
-be small (e.g. NVRAM, PFS, Campaign Storage, HPSS), 3) The default
-lifetime of objects in each tier will be specified by the resource by
-a tuple (size, lifetime).  The user can then utilize this information to
-place information to the system such as: store the highest priority data
-as high up on the hierarchy for 6 months, and then migrate this data to tape.
-The user can also define the utility for CR files as that they have high
-priority but their lifetime is for a short period of time; i.e. when the
-next CR file appears on the system or the next two CR files. Research in
-describing this utility function will be in conjunction with users
-and we will evaluate how policies can be placed in the system and how well
-this utility function can be observed by the user.
+%The assumptions which we will make in choosing a utility function is that:
+%1) There will only be a small number of priority levels that users will use
+%to re-prioritize their data; 2) The number of tiers in the storage system will
+%be small (e.g. NVRAM, PFS, Campaign Storage, HPSS), 3) The default
+%lifetime of objects in each tier will be specified by the resource by
+%a tuple (size, lifetime).  The user can then utilize this information to
+%place information to the system such as: store the highest priority data
+%as high up on the hierarchy for 6 months, and then migrate this data to tape.
+%The user can also define the utility for CR files as that they have high
+%priority but their lifetime is for a short period of time; i.e. when the
+%next CR file appears on the system or the next two CR files. Research in
+%describing this utility function will be in conjunction with users
+%and we will evaluate how policies can be placed in the system and how well
+%this utility function can be observed by the user.
+
+ We will formulate the concept of utility (from an
+application perspective) that is assigned data objects to quantify the
+benefits of placement and/or movement actions to application. These
+utilities may be based on the importance of the data objects, the frequency
+with which they are accessed, etc., and can be used during runtime data
+management to evaluate and compare different placement/movement options. 
+For example, data objects with higher utility value may be
+placed closer (e.g., in faster memory) to the accessing application.
+Similarly, when it is time to evict data objects from a higher level storage
+layer, the decision about which object to evict may take the utility into
+account. Our initial work in defining such a utility is presented in~\cite{tongipdps15}.
+ We will also explore autonomic approaches for balancing of the placement and 
+ movement of data across the memory hierarchy autonomically, which will be driven 
+ by the data utility and will leverage both user hints and information gathered at runtime (e.g., runtime data access
+history, network topology, etc.).
+Finally, we will explore a complementary notion of utility from a system-wide perspective 
+that will capture and quantify, for example, the impact of placement decisions on 
+the overall utilization of storage across multiple applications, and will enable system level runtime optimization.
 

### 2015-07-09 narrative/intro.tex

--- 2015-07-09 narrative/intro.tex
@@
 3 & Mechanisms for data movement across the storage hierarchy & Manage data migration based on policies& \S \ref{sec:managing-data-life},  \S\ref{sec:migration} \\\hline
-4 &  Extend the interaction of I/O middleware with the resource management system & We will manage data from the application  to the middleware to the storage system and allow them to have an autonomic interaction & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
-5 & Design and implement of I/O middleware architectures  & We will create APIs to guide the placement of chunks of data across the SSIO and autonomically move data when necessary& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
-6 & Approaches to improve the ability of SSIO software to support checkpoint/restart & We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then purged when it is no longer needed & \S \ref{fig:ssio-bucket}  \\
+4 &  Extend the interaction of middleware with the storage system & Manage data from the application  to middleware to the storage system using an autonomic interaction & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
+5 & Design and implement of I/O middleware architectures  & Create APIs to guide the placement of chunks of data across the SSIO and autonomically move data& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
+
+6 & Approaches to improve the ability of SSIO software to support checkpoint/restart & We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then purged when it is no longer needed & \S \ref{fig:ssio-bucket}, \S\ref{sec:qos}, \S\ref{sec:managing-data-life}  \\
 \hline

### 2015-07-09 participant-table.tex
participant table

--- 2015-07-08 participant-table.tex
@@
-% Generated by bin/config-participant-table on Thu Jun 18 15:07:03 EDT 2015
+% Generated by bin/config-participant-table on Thu, Jul 09, 2015  4:57:56 PM
 \newlength{\longestpientry}
@@
 \\
-\textbf{\ornl} & \makebox[\longestpientry][l]{\textbf{Scott Klasky} (\emph{klasky@ornl.gov})}\hspace*{1em}Hasan Abbasi, Gary Liu, Kimmy Mu, Manish Parashar\\
+\textbf{\ornl} & \makebox[\longestpientry][l]{\textbf{Scott Klasky} (\emph{klasky@ornl.gov})}\hspace*{1em}Hasan Abbasi, Kimmy Mu, Manish Parashar, Feiyi Wang\\
 \textbf{\snl} & \makebox[\longestpientry][l]{\textbf{Gerald Lofstead} (\emph{gflofst@sandia.gov})}\hspace*{1em}Matthew Curry, Lee Ward\\
-\textbf{\ucsc} & \makebox[\longestpientry][l]{\textbf{Carlos Maltzahn} (\emph{carlosm@soe.ucsc.edu})}\hspace*{1em}\
+\textbf{\ucsc} & \makebox[\longestpientry][l]{\textbf{Carlos Maltzahn} (\emph{carlosm@soe.ucsc.edu})}\hspace*{1em}\\
 \end{tabularx}

### 2015-07-09 participant-table.tex
figure with R instead of F, and I need Rutgers as a separate institution

@@
 \\
-\textbf{\ornl} & \makebox[\longestpientry][l]{\textbf{Scott Klasky} (\emph{klasky@ornl.gov})}\hspace*{1em}Hasan Abbasi, Kimmy Mu, Manish Parashar, Feiyi Wang\\
+\textbf{\ornl} & \makebox[\longestpientry][l]{\textbf{Scott Klasky} (\emph{klasky@ornl.gov})}\hspace*{1em}Hasan Abbasi, Kimmy Mu, Feiyi Wang\\
 \textbf{\snl} & \makebox[\longestpientry][l]{\textbf{Gerald Lofstead} (\emph{gflofst@sandia.gov})}\hspace*{1em}Matthew Curry, Lee Ward\\
-\textbf{\ucsc} & \makebox[\longestpientry][l]{\textbf{Carlos Maltzahn} (\emph{carlosm@soe.ucsc.edu})}\hspace*{1em}\\
+\textbf{\ucsc} & \makebox[\longestpientry][l]{\textbf{Carlos Maltzahn}
+ (\emph{carlosm@soe.ucsc.edu})}\hspace*{1em}\\
+ \textbf{Rutgers} & \makebox[\longestpientry][l]{\textbf{Manish Parashar}
+ (\emph{parashar.rutgers.edu})}\hspace*{1em}\\
 \end{tabularx}

### 2015-07-10 narrative/intro.tex
minor changes again.

@@
 Our proposed research program is guided by the following overarching objectives:
-\begin{itemize}
+\begin{tightItemize}
 \item
@@
 system.
-\end{itemize}
+\end{tightItemize}
 
@@
 
-6 & Approaches to improve the ability of SSIO software to support checkpoint/restart & We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then purged when it is no longer needed & \S \ref{fig:ssio-bucket}, \S\ref{sec:qos}, \S\ref{sec:managing-data-life}  \\
+6 & Approaches to improve the ability of SSIO software to support checkpoint/restart & We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then purged when it is no longer needed & \S \ref{sec:data-description}, \S\ref{sec:qos}, \S\ref{sec:managing-data-life}  \\
 \hline
@@
 to varying bandwidth can be handled appropriately. Ideally, a set of
-user-specified rules would allow the system to make autonomic decisions dynamically.
+user-specified rules would allow the system to make autonomic decisions dynamically
+~\S~\ref{sec:qos}.
 %
 
-Central to our approach is giving users the ability to define the importance, or utility of their data.
+Central to our approach is giving users the ability to define the importance, or utility of their data    
+ \S\ref{sec:data-description}.
 This includes the ability to assign varying levels of importance to individual data objects in their entirety, or specific
@@
 Research challenges include: 1) an expressive and easy to use interface, 2) a model for the system supporting
-autonomic decision for saving data using the user-defined importance and utility of data, and 3) the interface between
-the system and the back-end storage across both single, and multiple running applications.
+autonomic decision for saving data using the user-defined importance and utility of data~\S~\ref{sec:managing-data-life})
+, and 3) the interface between
+the system and the back-end storage across both single, and multiple running applications~\S~\ref{sec:qos}.
 

### 2015-07-10 narrative/intro.tex
added some text about performance bounds. can someone with storage experience clean this up?

@@
 layers automatically, without user intervention. 
+Central to our approach is that the storage system needs to define bounds of time estimates
+(min time, max time), and as our system evolves, we want to allow plug-ins which can help
+reduce the range of the bounds, and furthermore, with optimizations we would like to minimize 
+the times for both reading and writing. Thus, we want predictable performance, and methods
+which can obtain better performance bounds in time.
+
 %We will use learning techniques to automatically migrate data

### 2015-07-10 narrative/data-description.tex
figure with R instead of F, and I need Rutgers as a separate institution
trying to distinguish now between value and priority.

@@
 attribute is \textit{data utility}.  It enables QoS scheduling, data placement
-decisions, data lifetime decisions, and captures the chunk priority. It is
+decisions, data lifetime decisions, and captures the value of the chunk. It is
 our belief that exascale science will require users to prioritize a small set
@@
 \S~\ref{sec:data-refactor}) which will classify the particle data into four different
-prioritization levels. The user can then pick a data utility, such as three months
-for the highest priority items and 12 months for the lowest priority data.
+value levels. The user can then pick a data utility, such as three months
+for the highest value items and 12 months for the lowest value data.
 The user will then call an API function that allows them
@@
 If this time estimate is too large, an API call can be made to 
-specify the lowest priority element to be written, skipping the rest in order to save time.
+specify the lowest value element to be written, skipping the rest in order to save time.
 The data will then be stored
-across the layers of the file system where the two highest priority buckets
-will be stored in the Parallel File System, but the next highest priority items
+across the layers of the file system where the two highest value buckets
+will be stored in the Parallel File System, but the next highest value items
 will be stored lower in the storage stack. Depending on the capabilities of the actual storage system, 
@@
 their knowledge discovery.
-
-\begin{wrapfigure}{r}{0.6\textwidth}
-        \begin{centering} 
-	\includegraphics[scale=0.8]{graphics/SSIO-bucket.pdf}
-        %\vspace{1ex}
-        \caption{The system manages the data across the various stages of the data lifecycle}
-        \label{fig:ssio-bucket}
-        \end{centering}
-      % \vspace{-1ex}
-\end{wrapfigure}
 
@@
 the overall utilization of storage across multiple applications, and will enable system level runtime optimization. 
+\begin{wrapfigure}{R}{0.6\textwidth}
+        \begin{centering} 
+	\includegraphics[scale=0.8]{graphics/SSIO-bucket.pdf}
+        %\vspace{1ex}
+        \caption{The system manages the data across the various stages of the data lifecycle}
+        \label{fig:ssio-bucket}
+        \end{centering}
+      % \vspace{-1ex}
+\end{wrapfigure}
 
@@
 provide a time estimate, and then allow the users to decide if
-they will read less (e.g., only the highest priority sub-chunks)
+they will read less (e.g., only the highest value sub-chunks)
 if the time estimate

### 2015-07-10 narrative/evaluation.tex
reference fixes
changed the acronym.
put table title on top.
trying to distinguish now between value and priority.

@@
 \centering
+\caption{Guiding Principles}
+\label{table:eval}
 \begin{tabular}{ | p{1.4in} | p{2in} | p{2.5in} | }
@@
 \end{tabular}
-
-\caption{Guiding Principles}
-\label{table:eval}
 \end{table}
 
-In addition to benchmarking the various S2E2 methods and tools in isolation,
+In addition to benchmarking the various SIRIUS methods and tools in isolation,
 we will work with our science partners at Princeton University (J. Tromp),
@@
 
-The fifth part of the evalation will evaluate how well additional data
+The fifth part of the evaluation will evaluate how well additional data
 knowledge in the system will lead to more performance by placing the most
@@
 knowledge levels and evaluate the system impact.  For example, if the user
-doesn't want the data broken into priority chunks, then the data will be at
+doesn't want the data broken into {\it value} chunks, then the data will be at
 the mercy of what the system does when there is no knowledge. If the user then 
-provides sufficient information to use ``generic'' rouitnes based on frequency
+provides sufficient information to use ``generic'' routines based on frequency
 or accuracy, then that knowledge will be used to decompose the objects into

### 2015-07-10 narrative/placement.tex
trying to distinguish now between value and priority.

@@
 the data that, with user defined error, represent the data, and place these
-high priority  data sets on the higher tiers for faster access.
+{\it high value}  data chunks on the higher tiers for faster access.
 

### 2015-07-10 narrative/proposed-work.tex
trying to distinguish now between value and priority.

@@
 the CR data. The migration and purging of this will be optimized by our SSIO layer, and we will have the user data intentions
-signify the utility of CR data so that this data can be high-priority but stored for short periods of time.
+signify the utility of CR data so that this data can be high-value but stored for short periods of time. The value of a chunk of data
+from an object signifies it's importance to the users relative to the other chunks of their data.
 

### 2015-07-10 narrative/qos.tex
trying to distinguish now between value and priority.

@@
 process. For example, particularly for a capability run, the storage system
-should give priority to data subset tagged with high importance to stay in fast
-storage tiers. Enabling this priority will be driven by other job
-characteristics such as job size or a manually set priority inserted by system
-administrators. It should also store the highest quality data possible given
+should give (on a user by user basis), higher priority to data subset tagged with high importance to stay in fast
+storage tiers.
+% Enabling this priority will be driven by other job
+%characteristics such as job size or a manually set priority inserted by system
+administrators.
+ It should also store the highest value data possible given
 the performance requirement. Ideally, the storage system will store the raw
 data at full fidelity. When this tier approaches capacity, data should migrate
-not just based on age, but also based on this priority. This will help ensure
-the shortest time to insight by making sure the most important machine runs
-have priority to use system resources.
+not just based on age, but also based on this value. One area of research will be in
+trying to see if we can retain some guarantees of QoS while giving some high-profile users
+greater SSIO a higher {\bf priority} than other users on the system. We realize that this makes
+it very difficult to main overall system QoS, but we will investigate techniques when one
+user is running on the majority of the exascale resource and we can allow them to lock out other SSIO
+request during small periods of time during the lifetime of their exascale simulation.
 
@@
 approach that handles clients from all connected platforms and does not require
-modification to use new storage access APIs. By using a priority tagging
+modification to use new storage access APIs. By using a {\it value} tagging
 system, we will offer defaults suitable for all applications that can be

### 2015-07-10 narrative/deliverables.tex
spelling, etc.
minor changes
more formatting and fixing.

@@
 The following are the per year project deliverables:
+\small
 
 \textbf{Year 1}
-\begin{itemize}
-\item Develop new techniques for data description that allow users to 
-describe the data utility based on their expectations. (ORNL)
+\begin{tightItemize}
+\item Develop  techniques for data description  to 
+describe the data utility based on user expectations. (ORNL)
 
-\item Assemble and evaluate general techniques for refactoring of data
-that will be embedded in the system and can be selected by users or applications. (Brown/ORNL)
+\item Evaluate general  techniques for refactoring of data
+that will be embedded in the system . (ORNL)
 \item Explore the use of application hints and utility of data to guide the initial placement of data. 
-(Rutgers/ORNL)
+(Rutgers)
 \item Investigate the trade-off between ``filing'' and ``piling'' data. 
 Demonstrate a time bounded search approach for finding data within the storage system
-to identify data and the current location. (UCSC/Sandia)
-\item Demonstrate a metadata service capable of serving both POSIX clients and our clients. (Sandia/UCSC)
-\end{itemize}
+to identify data and the current location. (UCSC)
+\item Demonstrate a metadata service capable of serving both POSIX clients and our clients. (Sandia)
+\end{tightItemize}
 
 \textbf{Year 2}
-\begin{itemize}
-\item Develop a new type of querying system that allows an application to ask the storage system
+\begin{tightItemize}
+\item Develop a querying system to allow apps to inquire
 about completion timing information. (ORNL)
 \item Evaluate the effectiveness and overhead to re-organize data and use the multi-layer approach 
-to staging. (Brown/ORNL)
+to staging. (ORNL)
 \item Research autonomic data management strategies that
-can evaluate utility/cost tradeoff, and appropriately place/move data objects at runtime. 
-Research runtime tracking and estimation. (Rutgers/ORNL)
+can evaluate utility/cost trade-off, and appropriately place/move data objects at runtime. 
+Research runtime tracking and estimation. (Rutgers)
 \item Extend Sirocco to enforce data-centric metrics that help guide the QoS decisions. (Sandia)
-(Sandia/UCSC)
-\end{itemize}
-
+\item Prototype time bounded search approach for finding data within the storage system (UCSC)
+\end{tightItemize}
 \textbf{Year 3}
-\begin{itemize}
-\item Work with DOE applications, particularly XGC1, GTC, and SPECFM3D, to assess 
-the new I/O interface including both APIs and hints. (ORNL)
-\item Research techniques that can enhance migration without making the eventual 
-lookup of data unbounded, and manage each tier in a scalable fashion. (ORNL)
-\item Demonstrate incorporating A\&A and show scalability under both application load as
-well as storage system pressures showing that we can tolerate both loads and maintain quality of
-service. (Sandia)
+\begin{tightItemize}
+\item Work with DOE applications to assess 
+the new I/O interface including both APIs and hints. (ORNL, Rutgers)
+\item Prototype new QoS methods for storing \& retrieving data in given time bounds. (ORNL, UCSC)
+\item Demonstrate admission control and show scalability for  application loads and 
+ storage system pressures to maintain quality of
+service. (Sandia, UCSC)
 %\item storage APIs to manipulate storing and retrieving data to specific tiers
@@
 %\item example lossy and lossless compression plug-ins to demonstrate the effectiveness of this approach
-\end{itemize}
+\end{tightItemize}
+\normalsize

### 2015-07-10 narrative/enabling_tech.tex
more formatting and fixing.

--- 2015-07-09 narrative/enabling_tech.tex
@@
 %devices. This project, 
-{\bf Sirocco} came out of the LWFS project and offers a peer-to-peer style data storage device
+In this project we will leverage {\bf Sirocco}~\cite{sirocco}, which  offers a peer-to-peer style data storage device
 mesh with data migrating based on use, required resilience, and available
-resources. For example, when writing data initially into Sirocco, required
+resources. When writing data initially into Sirocco, required
 resilience characteristics are provided prompting Sirocco to ensure the data is
@@
 When a particular device is full, Sirocco determines if it should migrate the
-data to a new location or if it is a copy beyond the resilience requirement. If
-this copy must still be protected, Sirocco will find a place of sufficient 
-resilience to move the data to before deleting the local copy. Further, when
-data are requested, a client can ask any Sirocco server and Sirocco will find
-the data and return it to the client. If desired, the canonical data location
-can be moved on client request to offer better performance. Additionally,
-a stored object can have multiple data forks that can
+data to a new location or if it is a copy beyond the resilience requirement. Sirocco also allows
+a stored object to have multiple data forks that can
 store different versions or aspects of the same data.  For this project, we
@@
 
-We will also take advantage of the many innovations discovered in the Ceph~\cite{weil:osdi06} storage
-system project. 
+We will also take advantage of the many innovations discovered in the {\bf Ceph}~\cite{weil:osdi06} storage
+system. 
 %
@@
 %
-The {\bf Ceph} project (ceph.com) started at UC Santa Cruz in 2004 and was initially
-funded by DOE/NNSA involving LLNL, LANL, and Sandia, with the goal to create an
+%The {\bf Ceph} project
+% (ceph.com) started at UC Santa Cruz in 2004 and was initially
+%funded by DOE/NNSA involving LLNL, LANL, and Sandia, with the\
+ The project goal was to create an
 object-based parallel file system that addresses the well-known metadata

### 2015-07-10 narrative/evaluation.tex
more formatting and fixing.
more wordsmyth and trying to make more sense and uniformity.

@@
 Our guiding principles are the basis for the evaluation. These principles are
-described in Table~\ref{table:eval}.  We will partition the problem along these
+described in Table~\ref{table:eval}, and we will partition the problem along these
 lines and evaluate how the knowledge centric system design we employ can
 provide more predictable performance and data quality. 
+In addition to benchmarking the various SIRIUS methods and tools in isolation,
+we will work with our science partners at Princeton University (J. Tromp, C. S. Chang),
+and U. C. Irvine (Z. Lin), along with the OLCF and SNL,
+to evaluate their applications working with the storage system to
+determine performance predictability and data quality.  Our evaluation 
+will use the multiple storage layers available on systems at both SNL and the
+OLCF, including the sith cluster discussed in the Equipment appendix. 
+ We will perform tests in both the system level (on sith) and in
+the user level (on titan), and evaluate our results using our application
+knowledge.
 
@@
 \begin{table}[ht]
-
+\small
 \centering
@@
 \end{tabular}
+\vspace{-2ex}
+\normalsize
 \end{table}
-
-In addition to benchmarking the various SIRIUS methods and tools in isolation,
-we will work with our science partners at Princeton University (J. Tromp),
-PPPL (C. S. Chang), and U. C. Irvine (Z. Lin), along with the OLCF and SNL,
-to evaluate the leading edge applications working with the storage system to
-determine performance predictability and data quality.  Our evaluation 
-will use the multiple storage layers available on systems at both SNL and the
-OLCF.  For example, the sith cluster at the OLCF
-(https://www.olcf.ornl.gov/computing-resources/sith/) has its own Lustre
-storage system, SSDs on each node, and is connected to the center-wide Lustre
-file system. We will perform tests in both the system level (on sith) and in
-the user level (on titan), and evaluate our results using our application
-knowledge.
 

### 2015-07-10 narrative/management.tex
more wordsmyth and trying to make more sense and uniformity.

@@
 Careful management will be required to ensure that the problems and delays that are inevitable in any research project
-do not unduly disrupt overall progress. To this end, we define a detailed workplan,
+do not  disrupt overall progress.Here we define a detailed workplan,
 establish a  management structure with defined oversight responsibilities, and establish a schedule of regular meetings and
@@
 will lead the Storage System. This project is focused on the tight integration of application aware-middleware and a storage system backend.
-Our team has a strong track record of tight-collaboration, with both Drs. Abbasi and Lofstead working with Dr. Klasky for their Ph.D. at the Georgia
-Institute of Technology.  Our structure includes an area lead for each major research area plus
+%Our team has a strong track record of tight-collaboration, with both Drs. Abbasi and Lofstead working with Dr. Klasky for their Ph.D. at the Georgia
+%Institute of Technology.  
+Our structure includes an area lead for each major research area plus
 an integration lead responsible for coordinating periodic integration tests. The area leads plus the project PIs
 form the management team, which will meet weekly to review progress and revise plans as needed.
-We will also have a {\bf Science Advisory Board} consisting of one application scientist, Dr. C. S. Chang (PPPL), and one storage system expert from the 
-OLCF, Dr. Vazhkudai (ORNL), who will advise us yearly on our project goals and outcomes and make adjustments accordingly. We will also have our
-Math Lead, Dr. Ainsworth, to work with the applied mathematics community to ensure that the hierarchical decomposition of data will work in theory
-for current and next generation exascale simulations.
+We will also have a  yearly meeting with  one application scientist we will partner with,  and one storage system expert from the 
+OLCF, who will advise us on our project goals and outcomes and we will make adjustments accordingly. We will also have our
+Math Lead, Dr. Ainsworth, to work with the applied mathematics community to ensure that our data-refactoring will work 
+for current and next generation exascale math algorithms.
 

### 2015-07-10 narrative/naming-service.tex
spelling, etc.
more wordsmyth and trying to make more sense and uniformity.

@@
 between the user and the storage system extends into this area as well.
-Negotiating the data quality/retrieval time tradeoff requires additional
+Negotiating the data quality/retrieval time trade-off requires additional
 metadata.  Providing this functionality requires the naming service be able to
@@
 different model. Because we are not requiring the entire storage space be
-addressible from a single tree root, we can offer multiple independent roots
+addressable from a single tree root, we can offer multiple independent roots
 using the Sirocco object storage. We will investigate how to make the naming
@@
 
-\begin{tightEnumerate}
-
-\item {\bf Milestone 1}: Demonstrate a metadata service capable of serving both POSIX
+In this project we will first 
+%\begin{tightEnumerate}
+%
+%\item {\bf Milestone 1}: 
+demonstrate a metadata service capable of serving both POSIX
 clients and our clients, but without consideration for authentication,
 authorization, or data migration.
-
-\item {\bf Milestone 2}: Demonstrate a time bounded search approach for finding data
+Secondly we will 
+%\item {\bf Milestone 2}: D
+demonstrate a time bounded search approach for finding data
 within the storage system to identify data and the current location.  This will
@@
 requests for the data presuming that it does not migrate again soon.
-
-\item {\bf Milestone 3:} Demonstrate incorporating {\color{red}A\&A (what does this stand for CHECK-tsr)} and show scalability under
+Finally, we will 
+%\item {\bf Milestone 3:} D
+demonstrate admission control and show scalability under
 both application load as well as storage system pressures, showing that we can
@@
 
-\end{tightEnumerate}
+%\end{tightEnumerate}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "../proposal"
+%%% End:

### 2015-07-10 narrative/intro.tex

@@
 %
-(1) the scientist who runs the exascale simulation,
+(1) scientists who run exascale simulations,
 %
@@
 %
-(3) the user of the system who is attempting to understand the data
+(3) users of the system who attempt to understand the data
 produced by a set of simulations.
@@
 \label{subsec:sim-perspective}
-The simulation scientist ideally should be able to
+A simulation scientist ideally should be able to
 obtain an estimate of how much time will be required to write the data,
@@
 user-specified rules would allow the system to make autonomic decisions dynamically
-~\S~\ref{sec:qos}.
+~\S~\ref{sec:managing-data-life}~\S~\ref{sec:qos}.
 %
@@
 writing or reading at the same time, leading to unpredictable performance.
-By requiring a certain QoS of service the system could potentially lock out other users. 
+By requiring a certain QoS, the system could potentially lock out and throttle other users. 
 The storage system must also automatically move data from one tier of the storage to

### 2015-07-10 narrative/data-reduction.tex
spelling
some more refs...

@@
 
-Our previous
+Previous
 work~\cite{lakshminarasimhan2011compressing,%
 gong2012multi,jenkins2012byte,gong2013parlo,boyuka2014transparent,%
-tang2014improving}
+tang2014improving, sato2014user}
 has focused on several alternative techniques to not only reduce the data
@@
 	re-arranged which involves memory operations and needs to be done
-	in-situ. This operation involves no communication.  Data needs to be
+	in situ. This operation involves no communication.  Data needs to be
 	copied, so the memory requirements are increased. At most this requires

### 2015-07-10 narrative/qos.tex
more wordsmyth and trying to make more sense and uniformity.
some more refs...

@@
 \end{figure}
-\vskip-0.2in
+ \vspace{-2ex}
 
@@
 should give (on a user by user basis), higher priority to data subset tagged with high importance to stay in fast
-storage tiers.
-% Enabling this priority will be driven by other job
-%characteristics such as job size or a manually set priority inserted by system
+storage tier
 administrators.

### 2015-07-10 narrative/intro.tex
fixed reading perspective.

@@
 %indexing to better cope with the expected tsunami of data.
-
+Overall this project deals with I/O and storage problems
+for both checkpoint/restart (C/R) and post-processing.
 We examine the research challenges from the perspective of:
@@
 \label{subsec:reading-perspective}
-From the perspective of the users who want to read in the data from their
-simulations, and then operate on this data and possibly write data, we see
-that in general these users will be running on much less compute resources
-than the users who are running the simulation. One of the important aspects
-of this class of user is that they want a reduction of latency and they
+From the perspective of reading, this project will allow both
+simulation restart and post-processing. For the former case,
+we will allow both traditional file-based approach leveraging fast storage tier
+such as NVRAM (non-volatile RAM), and memory-based approach via retrieving
+data directly from NVRAM. For the latter case,
+an important aspect is that they want a reduction of latency and they
 require a certain amount of QoS as well. When someone is

### 2015-07-10 narrative/intro.tex
reading perspective.
work in progress.

--- 2015-07-10 narrative/intro.tex
@@
 Our investigation will include  novel techniques for the efficient mapping of data objects
-from the user space across to a hierarchical tiered storage layer. These techniques will
+from the user space to a hierarchical tiered storage layer. These techniques will
 include support for application-guided data reductions with user-defined error bounds
@@
 Negotiation between these layers will be provided as a fundamental service, allowing users to
-easily save ``the best''  data, which is automatically placed throughout the
-entire storage stack.
+easily save ``the best''  data, which is automatically and appropriately placed within the
+ storage stack.
 \item
@@
 supported data lifetime, our approach will enable researchers to 
-place more knowledge into the Storage Systems and Input/Output (SSIO) system so that data can be recast from 
-a sequence of bytes to information rich data. This information rich data can then
-be used by the system to guide the  placement within the tiered storage and will optimize the time
-to gain knowledge and insight from exascale simulations.
+place more knowledge into the Storage Systems and Input/Output (SSIO) system 
+so that data can be recast from a sequence of bytes to information rich data. This 
+information rich data can then be used by the system to guide the  placement within 
+the tiered storage and will optimize the time to gain knowledge and insight from 
+exascale simulations.
 
@@
 As extracting knowledge is the only metric that matters for scientific applications, we will
-focus not only upon data storage, but upon efficient  retrieval of knowledge from their 
+focus not only on data storage, but also on efficient retrieval of knowledge from their 
 workflows.
@@
 
+%%% Manish: I am not sure I fully understand the first part of the paragraph below.
 This proposal can be summarized as research into using the varying importance of simulation data over time to
-ensure that the right portions of data, in the right quality  are available to scientists when they are needed.
+ensure that the right portions of data, with the right quality  are available to scientists when they are needed.
 Application-specific information about the importance of data will be used in conjunction with system and resource-specific information
-to optimize  scientists experience in interacting with, and extracting knowledge and insight from their  data.
+to optimize the scientists' experience in interacting with, and extracting knowledge and insight from their data.
 
@@
 the Sandia National Laboratory (SNL) Lightweight File Systems (LWFS) project and storage system knowledge, the metadata
-management and storage systems knowledge of UCSC, and the middleware expertise
-of Rutgers. We will integrate this diverse expertise to investigate how to create
+management and storage systems knowledge at UCSC, and the middleware expertise
+at Rutgers. We will integrate this diverse expertise to investigate how to create
 a data, workflow, and application aware storage and I/O system.
@@
 \multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant S2E2 work} & \multicolumn{1}{|c|}{\bf Proposed research} \\\hline\hline
-1 & Data abstractions & Create new APIs  and annotations for users to guide the system for placement  & \S\ref{sec:data-description} \\\hline
+1 & Data abstractions & Create new APIs  and annotations to guide data placement across the storage system & \S\ref{sec:data-description} \\\hline
 2 & Mapping science data models onto hierarchical storage &  Create the ability to re-organize, reduce and prioritize data chunks from an object& \S\ref{sec:data-refactor}, \S\ref{sec:data-description},  \S\ref{sec:managing-data-life}\\\hline

### 2015-07-10 narrative/intro.tex

@@
 \label{subsec:reading-perspective}
-From the perspective of the users who want to read in the data from their
-simulations, and then operate on this data and possibly write data, we see
-that in general these users will be running on much less compute resources
-than the users who are running the simulation. One of the important aspects
-of this class of user is that they want a reduction of latency and they
+From the perspective of reading, this project will look at both
+simulation restart and post-processing. For the former case,
+we will allow both traditional file-based approach leveraging fast storage tier
+such as NVRAM (non-volatile RAM), and memory-based approach via retrieving
+data directly from NVRAM. For the latter case,
+an important aspect is that they want a reduction of latency and they
 require a certain amount of QoS as well. When someone is

### 2015-07-10 narrative/qos.tex
removed priorty portion.
added a sentence to charging model

@@
 model and the interactions a request can have with other requests (e.g. a
-random request to spinning media interrupting a stream of sequential requests).
+random request to spinning media interrupting a stream of sequential
+requests). Thus users with a higher allocation can still meet their higher
+requirements for I/O, without violating the gurantees given to other users.
 
@@
 
-As stated in the introduction, we seek to make a storage stack that works with
-the user. Towards that end, we must incorporate data attributes into this
-process. For example, particularly for a capability run, the storage system
-should give (on a user by user basis), higher priority to data subset tagged with high importance to stay in fast
-storage tier
-administrators.
- It should also store the highest value data possible given
-the performance requirement. Ideally, the storage system will store the raw
-data at full fidelity. When this tier approaches capacity, data should migrate
-not just based on age, but also based on this value. One area of research will be in
-trying to see if we can retain some guarantees of QoS while giving some high-profile users
-greater SSIO a higher {\bf priority} than other users on the system. We realize that this makes
-it very difficult to main overall system QoS, but we will investigate techniques when one
-user is running on the majority of the exascale resource and we can allow them to lock out other SSIO
-request during small periods of time during the lifetime of their exascale simulation.
+% As stated in the introduction, we seek to make a storage stack that works with
+% the user. Towards that end, we must incorporate data attributes into this
+% process. For example, particularly for a capability run, the storage system
+% should give (on a user by user basis), higher priority to data subset tagged with high importance to stay in fast
+% storage tier
+% administrators.
+%  It should also store the highest value data possible given
+% the performance requirement. Ideally, the storage system will store the raw
+% data at full fidelity. 
+% When this tier approaches capacity, data should migrate
+% not just based on age, but also based on this value. One area of research will be in
+% trying to see if we can retain some guarantees of QoS while providing
+% policies that can govern admissio
+% greater SSIO a higher {\bf priority} than other users on the system. 
+% We realize that this makes
+% it very difficult to main overall system QoS, but we will investigate techniques when one
+% user is running on the majority of the exascale resource and we can allow them to lock out other SSIO
+% request during small periods of time during the lifetime of their exascale simulation.
 

### 2015-07-10 participant-table.tex
updated personnel.
added mark, gary to personnel list and change manish's affiliation

--- 2015-07-09 participant-table.tex
@@
-% Generated by bin/config-participant-table on Thu, Jul 09, 2015  4:57:56 PM
+% Generated by bin/config-participant-table on Fri, Jul 10, 2015  1:33:09 PM
 \newlength{\longestpientry}
 \setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Scott Klasky} (\emph{klasky@ornl.gov})}}}
+\setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Manish Parashar} (\emph{parashar@rutgers.edu})}}}
 \setlength{\longestpientry}{\maxof{\longestpientry}{\widthof{\textbf{Gerald Lofstead} (\emph{gflofst@sandia.gov})}}}
@@
 \\
-\textbf{\ornl} & \makebox[\longestpientry][l]{\textbf{Scott Klasky} (\emph{klasky@ornl.gov})}\hspace*{1em}Hasan Abbasi, Kimmy Mu, Manish Parashar, Feiyi Wang\\
+\textbf{\ornl} & \makebox[\longestpientry][l]{\textbf{Scott Klasky} (\emph{klasky@ornl.gov})}\hspace*{1em}Hasan Abbasi, Mark Ainsworth, Qing Liu, Kimmy Mu, Feiyi Wang\\
+\textbf{\rutgers} & \makebox[\longestpientry][l]{\textbf{Manish Parashar} (\emph{parashar@rutgers.edu})}\hspace*{1em}\\
 \textbf{\snl} & \makebox[\longestpientry][l]{\textbf{Gerald Lofstead} (\emph{gflofst@sandia.gov})}\hspace*{1em}Matthew Curry, Lee Ward\\

### 2015-07-10 narrative/intro.tex
Work in progress...

@@
 1 & Data abstractions & Create new APIs  and annotations to guide data placement across the storage system & \S\ref{sec:data-description} \\\hline
-2 & Mapping science data models onto hierarchical storage &  Create the ability to re-organize, reduce and prioritize data chunks from an object& \S\ref{sec:data-refactor}, \S\ref{sec:data-description},  \S\ref{sec:managing-data-life}\\\hline
-3 & Mechanisms for data movement across the storage hierarchy & Manage data migration based on policies& \S \ref{sec:managing-data-life},  \S\ref{sec:migration} \\\hline
-4 &  Extend the interaction of middleware with the storage system & Manage data from the application  to middleware to the storage system using an autonomic interaction & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
+%2 & Mapping science data models onto hierarchical storage &  Create the ability to map, re-organize, reduce and prioritize data chunks from an object& \S\ref{sec:data-refactor}, \S\ref{sec:data-description},  \S\ref{sec:managing-data-life}\\\hline
+2 & Mapping science data models onto hierarchical storage &  Enable science-driven mapping, re-organization, reduction and prioritization of data chunks derive from application objects & \S\ref{sec:data-refactor}, \S\ref{sec:data-description},  \S\ref{sec:managing-data-life}\\\hline
+3 & Mechanisms for data movement across the storage hierarchy & Enable science-driven data placement and migration based on policies& \S \ref{sec:managing-data-life},  \S\ref{sec:migration} \\\hline
+4 &  Extend the interaction of middleware with the storage system & Manage data placement and migration between the application, middleware and storage system using an autonomic cross-layer approach & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
 5 & Design and implement of I/O middleware architectures  & Create APIs to guide the placement of chunks of data across the SSIO and autonomically move data& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
-
 6 & Approaches to improve the ability of SSIO software to support checkpoint/restart & We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then purged when it is no longer needed & \S \ref{sec:data-description}, \S\ref{sec:qos}, \S\ref{sec:managing-data-life}  \\
@@
 \paragraph{An Illustrative Example}
-The challenges addressed by this proposal are illustrated by a problem our team has been working on in the past six months.
-We use this motivating example to formulate a series of research questions and challenges that must be addressed in order to support exascale science.
+We use an illustrative Fusion simulation use case that our team as been working on to highlight the challenges addressed in this proposal, and to 
+%The challenges addressed by this proposal are illustrated by a problem our team has been working on in the past six months. We use this motivating example to 
+formulate a series of SSIO research questions that must be addressed in order to support exascale science.
 
-The latest use case is from  the XGC1~\cite{chang2006integrated} simulation from C.
+The use case describes a series of XGC1~\cite{chang2006integrated} simulations that C.
 S.
-Chang (PPPL), who is one of the largest users of Leadership Class Facilities (over 300M hours at DOE facilities).
-They launched a series of simulations, planning to produce 100~PB of data over 10 days of total runtime on the Titan system.
-An expert team was assembled from the user group, the  I/O, and storage  team to help with ensuring that the maximum amount of information would be
-saved.
-Due to physical resource limitations, the simulation was restricted to write about 10~PB initially.
-However, once the time to write and read, together with the financial cost to archive this much data were fully explored, this plan was significantly reduced.
-At each stage, the particular expert knowledge regarding the file system and the archival storage were not known to the user team.
+Chang (PPPL), one of the largest users of Leadership Class Facilities (over 300M hours at DOE facilities), planned to launch.
+These simulations were expected to produce 100~PB of data over 10 days of total runtime on the Titan system, and assembled a team of experts including the user group and I/O and storage personnel  to help ensure that the maximum amount of information would be saved.
+Due to physical resource limitations, the simulation was immediately restricted to write about 10~PB.
+However, when the time to write and read, together with the financial cost to archive this volume of data were fully explored, the goal was even further reduced and they would only be able to write 5~PB over the 10-day run.
+%At each stage, the particular expert knowledge regarding the file system and the archival storage were not known to the user team.
 As these particulars were realized, significant modifications in the plans were made.
@@
 This forced the team to carefully scrutinize the large amount of data to be generated and save only the very most important pieces.
-Further, they had to break this greatly reduced data into two categories: first, data they thought they would need to access while the simulation was running or shortly thereafter, and second, data they would need after the whole campaign was finished.
-The former category would be stored on disk, and the later would be archived.
+Further, they had to divide this  reduced data into two categories: first, data they thought they would need to access while the simulation was running or shortly thereafter, and second, data they would need after the whole campaign was finished.
+The former category of data would be stored on disk, and the latter would be archived.
 Additionally, they needed to figure out how to efficiently retrieve these different categories of data for post-processing, analysis and visualization.
@@
 In order to do this, they had to implement a rudimentary discovery system to determine which data were located on disk, and which were on tape archive.
+This solution worked well for initially but after the second day significant problems occurred: I/O variability became prohibitively expensive.
+Although the team had reserved resources for the run, the file system was shared, and a 10X variability in I/O speeds was observed.
 
-This solution worked well for the first day of the run but after the second day significant problems occurred: I/O variability became prohibitively expensive.
-Although the team had reserved resources for the run, the file system was shared, and a 10X variability in I/O speeds was observed.
-What is needed for this problem is a system to monitor, predict and manage resource usage in a dynamically adaptable way.
-Such a system could write more data when the resource is idle, and less data when it is very busy.
+What is needed here is a system to monitor, predict and manage resource usage in a dynamically adaptable way.
+Such a system can write more data when the resource was idle, and less data when it is busy.
 However, the most significant problem occurs after the simulation run completes.
-Basic visualizations, just to get a high-level view of the simulation, can take significant time.
-While some timesteps might be on disk, others are archived on tape (where transfer could take weeks).
-Further, visualization and analysis of detailed views of the simulation are even more complicated.
+Basic visualizations, to get a high-level view of the simulation, can take significant time due to data access latencies.
+While some timesteps might be on disk, others are archived on tape (where transfers could take weeks).
+Detailed visualization and analysis of the simulation data are even more complicated.
 
@@
 %indexing to better cope with the expected tsunami of data.
-Overall this project deals with I/O and storage problems
-for both checkpoint/restart (C/R) and post-processing.
-We examine the research challenges from the perspective of:
+%
+%Overall this project deals with I/O and storage problems
+%for both checkpoint/restart (C/R) and post-processing.
+This proposal aims to address core I/O and storage challenges faced by 
+extreme scale simulation workflows for typical I/O patterns including 
+checkpoint/restart (C/R), in-situ analysis, and post-processing.
+%
+Furthermore, we examine these research challenges from the perspective of:
 %
@@
 %
-(2) the Storage System and I/O layer which needs to negotiate between all of
+(2) the Storage System and I/O layer, which needs to negotiate between all of
 the users, and finally
@@
 \label{subsec:sim-perspective}
-A simulation scientist ideally should be able to
-obtain an estimate of how much time will be required to write the data,
+A simulation scientist should ideally be able to
+obtain an estimate of how much time will be required to write data,
 and how much storage space is available during the simulations lifetime.
-Further, they would like Quality of Service (QoS) assurances so that reactions
-to varying bandwidth can be handled appropriately. Ideally, a set of
-user-specified rules would allow the system to make autonomic decisions dynamically
+Further, they would like Quality of Service (QoS) assurances, and that
+issues such as varying bandwidth can be handled appropriately and 
+automatically. For example, a set of user-specified rules would allow 
+the system to make autonomic decisions at runtime
 ~\S~\ref{sec:managing-data-life}~\S~\ref{sec:qos}.
@@
  \S\ref{sec:data-description}.
-This includes the ability to assign varying levels of importance to individual data objects in their entirety, or specific
+This includes the ability to assign varying levels of importance to individual data objects in their entirety, or to specific
 subsets and pieces.
-Research challenges include: 1) an expressive and easy to use interface, 2) a model for the system supporting
-autonomic decision for saving data using the user-defined importance and utility of data~\S~\ref{sec:managing-data-life})
-, and 3) the interface between
-the system and the back-end storage across both single, and multiple running applications~\S~\ref{sec:qos}.
+Key research challenges include: 1) developing an expressive and easy to use programinterface, 2) formulating a system model that can support 
+autonomic decision, for example, for saving data based on user-defined importance and utility of data~\S~\ref{sec:managing-data-life})
+, and 3) the interface between the system and the back-end storage across both single, and multiple running applications~\S~\ref{sec:qos}.
 

### 2015-07-10 narrative/intro.tex
minor changes for CR

--- 2015-07-10 narrative/intro.tex
@@
 %coordinates these key storage components in a seamless fashion.
-
 %
@@
 1 & Data abstractions & Create new APIs  and annotations to guide data placement across the storage system & \S\ref{sec:data-description} \\\hline
-2 & Mapping science data models onto hierarchical storage &  Create the ability to re-organize, reduce and prioritize data chunks from an object& \S\ref{sec:data-refactor}, \S\ref{sec:data-description},  \S\ref{sec:managing-data-life}\\\hline
+2 & Mapping science data models onto hierarchical storage &  Create the ability to re-organize, reduce and integrate utility with data chunks from an object& \S\ref{sec:data-refactor}, \S\ref{sec:data-description},  \S\ref{sec:managing-data-life}\\\hline
 3 & Mechanisms for data movement across the storage hierarchy & Manage data migration based on policies& \S \ref{sec:managing-data-life},  \S\ref{sec:migration} \\\hline
@@
 
-6 & Approaches to improve the ability of SSIO software to support checkpoint/restart & We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then purged when it is no longer needed & \S \ref{sec:data-description}, \S\ref{sec:qos}, \S\ref{sec:managing-data-life}  \\
+6 & Approaches to improve the ability of SSIO software to support checkpoint/restart & We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then later purged & \S \ref{sec:data-description}, \S\ref{sec:qos}, \S\ref{sec:managing-data-life}  \\
 \hline

### 2015-07-10 narrative/enabling_tech.tex
adding references.

@@
 From the user perspective,
-{\bf ADIOS}~\cite{lofstead:2009:adaptible} offers an API nearly as simple as POSIX, but with the
+{\bf ADIOS}~\cite{liu_helloadios,lofstead:2009:adaptible} offers an API nearly as simple as POSIX, but with the
 ability to change the actual data transport mechanism without changing the

### 2015-07-10 narrative/intro.tex

--- 2015-07-10 narrative/intro.tex
@@
 5 & Design and implement of I/O middleware architectures  & Create APIs to guide the placement of chunks of data across the SSIO and autonomically move data& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
-6 & Approaches to improve the ability of SSIO software to support checkpoint/restart & We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then purged when it is no longer needed & \S \ref{sec:data-description}, \S\ref{sec:qos}, \S\ref{sec:managing-data-life}  \\
+6 & Approaches to improve the ability of SSIO software to support checkpoint/restart & We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then later purged & \S \ref{sec:data-description}, \S\ref{sec:qos}, \S\ref{sec:managing-data-life}
+
 \hline

### 2015-07-10 narrative/intro.tex
stupid error

@@
 1 & Data abstractions & Create new APIs  and annotations to guide data placement across the storage system & \S\ref{sec:data-description} \\\hline
-%2 & Mapping science data models onto hierarchical storage &  Create the ability to map, re-organize, reduce and prioritize data chunks from an object& \S\ref{sec:data-refactor}, \S\ref{sec:data-description},  \S\ref{sec:managing-data-life}\\\hline
 2 & Mapping science data models onto hierarchical storage &  Enable science-driven mapping, re-organization, reduction and prioritization of data chunks derive from application objects & \S\ref{sec:data-refactor}, \S\ref{sec:data-description},  \S\ref{sec:managing-data-life}\\\hline
@@
 5 & Design and implement of I/O middleware architectures  & Create APIs to guide the placement of chunks of data across the SSIO and autonomically move data& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
-6 & Approaches to improve the ability of SSIO software to support checkpoint/restart & We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then later purged & \S \ref{sec:data-description}, \S\ref{sec:qos}, \S\ref{sec:managing-data-life}
 
-\hline
 \end{tabular}

### 2015-07-10 narrative/intro.tex

@@
 5 & Design and implement of I/O middleware architectures  & Create APIs to guide the placement of chunks of data across the SSIO and autonomically move data& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
+6 &Approaches to improve the ability of SSIO software to support checkpoint/restart &  We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then later purged & \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:data-description} \\ \hline
 

### 2015-07-10 narrative/intro.tex
Work in progress...

--- 2015-07-10 narrative/intro.tex
@@
 %coordinates these key storage components in a seamless fashion.
+
 %
@@
 5 & Design and implement of I/O middleware architectures  & Create APIs to guide the placement of chunks of data across the SSIO and autonomically move data& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
-6 & Approaches to improve the ability of SSIO software to support checkpoint/restart & We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then later purged & \S \ref{sec:data-description}, \S\ref{sec:qos}, \S\ref{sec:managing-data-life}
-
+6 & Approaches to improve the ability of SSIO software to support checkpoint/restart & We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then purged when it is no longer needed & \S \ref{sec:data-description}, \S\ref{sec:qos}, \S\ref{sec:managing-data-life}  \\
 \hline
@@
 autonomic decision, for example, for saving data based on user-defined importance and utility of data~\S~\ref{sec:managing-data-life})
-, and 3) the interface between the system and the back-end storage across both single, and multiple running applications~\S~\ref{sec:qos}.
+, and 3) defining the interface between the system and the back-end storage for both single and multiple running applications~\S~\ref{sec:qos}.
 
  %
-  In our  scenario we would like the scientist would then specify a set of rules that governed the actual behavior while the simulation is running.
-If the system resources are available and the estimated write times are in order, the write can proceed.
-Or, the write can be postponed to a later time-step, or data reduced via in situ techniques can be saved.
-Data importance or utility will be used to place the various data on the proper SSIO layer.
-The user specified lifetime for each data ensures that the data with the highest utility is preserved in fast storage
-for the longest period of time, and the least utility are migrated to lower layers in the storage hierarchy.
+In the scenario above, the scientist would then specify a set of rules that would guide I/O and data management 
+at runtime. For example, if system resources are available and the estimated write times acceptable, the write 
+can proceed, but if this is not the case, the write may be postponed to a later time-step, or the data reduced using
+an in-situ techniques before it is saved. Data ``importance or utility'' can be used to place various data at  
+proper SSIO layers -- data with the highest utility (e.g., based on user defined importance or the frequency of access) 
+is maintained in faster storage for long periods of time, while data with lower utility can be migrated to lower layers 
+in the storage hierarchy.
+%The user specified lifetime for each data ensures that the data with the highest utility is preserved in fast storage
+%for the longest period of time, and the least utility are migrated to lower layers in the storage hierarchy.
 %

### 2015-07-10 narrative/intro.tex
intro minor changes

--- 2015-07-10 narrative/intro.tex
@@
 
-Central to our approach is giving users the ability to define the importance, or utility of their data    
+Central to our approach is giving users the ability to define the utility of their data    
  \S\ref{sec:data-description}.
@@
 subsets and pieces.
-Key research challenges include: 1) developing an expressive and easy to use programinterface, 2) formulating a system model that can support 
-autonomic decision, for example, for saving data based on user-defined importance and utility of data~\S~\ref{sec:managing-data-life})
-, and 3) the interface between the system and the back-end storage across both single, and multiple running applications~\S~\ref{sec:qos}.
+Key research challenges include: 1) developing an expressive and easy to use program interface, 2) formulating a system model that can support 
+autonomic decision, for example, for saving data based on user-defined importance and utility of data~\S~\ref{sec:managing-data-life}),
+ and 3) the interface between the system and the back-end storage across both single, and multiple running applications~\S~\ref{sec:qos}.
 
@@
 If the system resources are available and the estimated write times are in order, the write can proceed.
-Or, the write can be postponed to a later time-step, or data reduced via in situ techniques can be saved.
-Data importance or utility will be used to place the various data on the proper SSIO layer.
+%Or, the write can be postponed to a later time-step, or data reduced via in situ techniques can be saved.
+Data  utility will be used to place the various data on the proper SSIO layer.
 The user specified lifetime for each data ensures that the data with the highest utility is preserved in fast storage
-for the longest period of time, and the least utility are migrated to lower layers in the storage hierarchy.
+for the longest period of time, and the least utility are migrated to lower storage  layers.
 %
@@
 writing or reading at the same time, leading to unpredictable performance.
-By requiring a certain QoS, the system could potentially lock out and throttle other users. 
+By requiring a certain QoS~\S\ref{sec:qos}, the system could potentially lock out and throttle other users. 
 The storage system must also automatically move data from one tier of the storage to
@@
 the times for both reading and writing. Thus, we want predictable performance, and methods
-which can obtain better performance bounds in time.
+which can obtain better performance bounds in time. Furthermore, since our system will place chunks
+in all of the tiers  we need to quickly discover where the data is in the hierarchy~\S\ref{sec:naming-discovery}
 
@@
 From the perspective of reading, this project will look at both
-simulation restart and post-processing. For the former case,
+CR  and post-processing I/O patterns. For the former case,
 we will allow both traditional file-based approach leveraging fast storage tier
@@
 small sample of the data without much knowledge if they are missing
-important pieces of the data.
+important pieces of the data~\S\ref{sec:migration}.
 

### 2015-07-10 narrative/intro.tex

@@
 subsets and pieces.
-Key research challenges include: 1) developing an expressive and easy to use program interface, 2) formulating a system model that can support 
+Key research challenges include: 1) developing an expressive and easy to use a program interface, 2) formulating a system model that can support 
 autonomic decision, for example, for saving data based on user-defined importance and utility of data~\S~\ref{sec:managing-data-life}),
- and 3) the interface between the system and the back-end storage across both single, and multiple running applications~\S~\ref{sec:qos}.
+, and 3) defining the interface between the system and the back-end storage for both single and multiple running applications~\S~\ref{sec:qos}.
 
- %
-  In our  scenario we would like the scientist would then specify a set of rules that governed the actual behavior while the simulation is running.
-If the system resources are available and the estimated write times are in order, the write can proceed.
-%Or, the write can be postponed to a later time-step, or data reduced via in situ techniques can be saved.
-Data  utility will be used to place the various data on the proper SSIO layer.
-The user specified lifetime for each data ensures that the data with the highest utility is preserved in fast storage
-for the longest period of time, and the least utility are migrated to lower storage  layers.
+In the scenario above, the scientist would then specify a set of rules that would guide I/O and data management 
+at runtime. For example, if system resources are available and the estimated write times acceptable, the write 
+can proceed, but if this is not the case, the write may be postponed to a later time-step, or the data reduced using
+an in-situ techniques before it is saved. Data ``importance or utility'' can be used to place various data at  
+proper SSIO layers -- data with the highest utility (e.g., based on user defined importance or the frequency of access) 
+is maintained in faster storage for long periods of time, while data with lower utility can be migrated to lower layers 
+in the storage hierarchy.
+
 %

### 2015-07-10 narrative/proposed-work.tex
minor changes for value...
rephrase some text

--- 2015-07-08 narrative/proposed-work.tex
@@
 
+One of the underlying principles of our approach is that we recognize that there are a small set
+of important I/O patterns to optimize for~\cite{lofstead2011six,polte2009and,tian2011edo,tian2012system},
+and our representative applications contain all of these I/O patterns. Throughout our project
+we will continuously focus on creating research artifacts which can optimize these patterns on current  and
+exascale systems.  One important pattern is Checkpoint/Restarts (CR) and we will focus on optimizing this pattern,
+ and also using many of the important research techniques for using NVRAM and neighbor's NVRAM for storing and retrieving
+the CR data. The migration and purging of this will be optimized by our SSIO layer using the utility function of the CR data so that this can  be stored for short periods of time, and then purged.
+
 %\underline{Principle 3:} {\bf Users goals define data policies} In a data-centric SSIO system. 

### 2015-07-10 narrative/proposed-work.tex
guiding principle 2.

--- 2015-07-10 narrative/proposed-work.tex
@@
 \underline{Principle 1:} {\bf A knowledge-centric system design} that allows user knowledge  to
-define data policies. Today Storage System and I/O layers are written in a
+define data policies. Today SSIO layers are written in a
 stove-pipe fashion, and quite often do not allow optimizations to take place.
-We propose to re-design the layer in a highly integrated fashion where users
+We propose to re-design the layers in a highly integrated fashion where users
 will place their intentions into the system and actions will statically and dynamically
@@
 
-\underline{Principle 2}: {\bf Predictable performance and quality of data in the SSIO layer} needs to be
-established so more science can be done on the exascale systems. Without
-predictable performance, the I/O layer performance must be guessed by the
-users, which quite often means that science is not performed well.
+\underline{Principle 2}: {\bf Predictable performance and quality of data in the SSIO layers} needs to be
+established so science can be done on the exascale systems in a more efficient manner. Without
+predictable performance, not only can the runs be slowed down because of contentions \cite{liu_hotstorage}, 
+but also it affects key science decisions, e.g., how much data reduction should be performed.
+It is clear that if a run is expected to be slow, an aggresive data reduction needs to be performed
+so that science can be done in a timely fashion, which has been reflected in the XGC run (\S~\ref{sec:introduction})
+the PI recently participated.
 
@@
 signify the utility of CR data so that this data can be high-value but stored for short periods of time. The value of a chunk of data
-from an object signifies it's importance to the users relative to the other chunks of their data.
+from an object signifies its importance to the users relative to the other chunks of their data.
 

### 2015-07-10 narrative/proposed-work.tex
fixed a few thing in the proposed work.

@@
 so that science can be done in a timely fashion, which has been reflected in the XGC run (\S~\ref{sec:introduction})
-the PI recently participated.
+that the PI recently participated.
 
@@
 approach for describing intentions and other knowledge from {\it user}
-space, as well as hidden knowledge of the underlying storage layers.
+space, as well as allow performance estimations and guarantees 
+from the underlyingree storage.
 %For example, ADIOS contains most of the ``best practices'' for I/O optimizations for
@@
 employed this strategy of separation of concerns in ADIOS, and believe that
-it will become increasingly important as users want to describe more about
-the storage hierarchy and the priorities of different chunks of data. 
-Our research is aimed to allow users to simply describe these things and
-utilize the descriptions so that the best optimizations can be made.
+it will become increasingly important because saving all data may not be possible, and
+users want the ability to describe and prioritize different chunks of data. 
+Our research is aimed to allow users to provide simple descriptions of their data,
+which further allows the middleware and storage to optimize application I/O performance.
 

### 2015-07-10 narrative/proposed-work.tex
more.

@@
 
-One of the underlying principles of our approach is that we recognize that there are a small set
-of important I/O patterns to optimize for~\cite{lofstead2011six,polte2009and,tian2011edo,tian2012system},
-and our representative applications contain all of these I/O patterns. Throughout our project
-we will continuously focus on creating research artifacts which can optimize these patterns on current  and
-exascale systems.  One important pattern is Checkpoint/Restarts (CR) and we will focus on optimizing this pattern,
+Additionally we recognize that HPC applications exhibit a set
+of common I/O patterns \cite{lofstead2011six,polte2009and,tian2011edo,tian2012system},
+Throughout our project we will continuously focus on creating research artifacts 
+which can optimize these I/O patterns on current and
+exascale systems. One important pattern is Checkpoint/Restart (C/R) and we will focus on optimizing this pattern,
  and also using many of the important research techniques for using NVRAM and neighbor's NVRAM for storing and retrieving
-the CR data. The migration and purging of this will be optimized by our SSIO layer using the utility function of the CR data so that this can  be stored for short periods of time, and then purged.
+the C/R data. The migration and purging of this will be optimized by our SSIO layer using the utility function of the C/R data so that this can be stored for short periods of time, and then purged.
 

### 2015-07-10 cover-collab.tex
supplement for collaboration

\vfill
\begin{center}
\textbf{\huge Cover Page Supplement for Collaborations}
\end{center}

\vfill

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Model content for table, really included below
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \begin{center}
%% \textbf{\large Collaborating Institutions}
%% \medskip
%% \begin{tabular}{lll}
%% \hline\hline
%%   & \emph{Institution} & \emph{Principal Investigator} \\
%% \hline
%% \textbf{Lead PI} & \textbf{Oak Ridge National Laboratory} 
%%   & \textbf{Sudharshan Vazhkudai}\\
%% Co-PI & Lawrence Berkeley National Laboratory & Ramakrishnan Lavanya \\
%% \hline\hline
%% \end{tabular}
%% \end{center}

\input{cover-collab-table}

\vfill
\begin{center}
\textbf{\large Leadership Structure}
\medskip
\end{center}

This is a collaborative proposal between 
Oak Ridge National Laboratory, and
Sandia National Laboratories, and
Rutgers University, and
University of California, Santa-Cruz.

%% Provide a statement explaining the leadership structure of the
%% collaboration.

\vfill
\begin{center}
\textbf{\large Facilities, Equipment, and Resources}
\medskip
\end{center}

This proposal brings together the expertiese and facilities from two
national laboratories, Oak Ridge National Laboratory and Sandia National
Laboratories, and two universities, Rutgers University and University of
California, Santa-Cruz. 
 In addition, by
sharing experiese and personnel costs, it represents
a collaboration with other large programs within the Office of
Science, including on data management and exascale monitoring, as well as
ongoing development of the Sirocco file-system. 
The proposed work will leverage existing facilities at all three
sites, including leadership machines at ORNL and SNL, and also utilize
storage oriented cluster systems at all locations. The PIs will also combine
their experience with production middleware and storage system software
development and their collaboration with leading application scientists to
better guide the assumptions underlying this effort.

%% Include a description of each collaborating institution’s facilities,
%% equipment, and resources that will be made available to the
%% collaborative group.

\vfill
\begin{center}
\textbf{\large Summary Budget Information}
\medskip
\begin{tabular}{| l | l | l | r | r | r | r |} \hline
Role & Name & Institution & Year 1 (\$K) & Year 2 (\$K) &   Year 3 (\$K) & Total (\$K) \\
\hline
Lead-PI & Scott Klasky & ORNL & {\$}570 &  {\$}570 &  {\$}570 &   {\$}1,710\\
Co-PI & Jay Lofstead & Sandia  &  {\$}350 &  {\$}350 &  {\$}350 &  {\$}1,050 \\
Co-PI & Carlos  Maltzahn & UCSC  &  {\$}160 &  {\$}160 &  {\$}160 &  {\$}480 \\
Co-PI & Manish Parashar & Rutgers &  {\$}170 &  {\$}170 &  {\$}170 &  {\$}510  \\
\hline
{\bf Totals}  & & & {\bf  {\$}1,250} & {\bf  {\$}1,250} & {\bf  {\$}1,250} & {\bf  {\$}3,750}\\
\hline
\end{tabular}
\end{center}

\noindent
{Period of Performance: October 1, 2015 - September 31, 2018}\\
\noindent
\vfill

### 2015-07-10 abstract.tex

@@
 4) leveraging techniques such as application-aware data compression, 
-and I/O prioritization to enforce these
+and  I/O utility to enforce these
 policies.

### 2015-07-10 narrative/intro.tex

@@
 (min time, max time), and as our system evolves, we want to allow plug-ins which can help
-reduce the range of the bounds, and furthermore, with optimizations we would like to minimize 
-the times for both reading and writing. Thus, we want predictable performance, and methods
-which can obtain better performance bounds in time. Furthermore, since our system will place chunks
-in all of the tiers  we need to quickly discover where the data is in the hierarchy~\S\ref{sec:naming-discovery}
+reduce the range of the bounds.
+Overall, our goal is to enable predictable performance acorss the storage system, and we will develop mechanisms to enable this
+~\S\ref{sec:naming-discovery}.
 

### 2015-07-10 narrative/intro.tex
changed from manish, as well as other minor changes.

@@
 \label{subsec:reading-perspective}
-From the perspective of reading, this project will look at both
-CR  and post-processing I/O patterns. For the former case,
-we will allow both traditional file-based approach leveraging fast storage tier
-such as NVRAM (non-volatile RAM), and memory-based approach via retrieving
-data directly from NVRAM. For the latter case,
-an important aspect is that they want a reduction of latency and they
-require a certain amount of QoS as well. When someone is
-doing interactive analysis and expects that the data they read in is about
+From the perspective of reading, this project will look at key I/O patterns such as checkpoint-restart (CR), in-situ analytics and post-processing. 
+For example, in case of CR, we will support both, traditional file-based approaches leveraging fast storage tiers
+such as NVRAM (non-volatile RAM), as well as memory-based approach based on retrieving
+data directly from NVRAM. In case of in-situ and post-processing,  an important aspect is the latency -- here reducing latency and providing QoS 
+ becomes important. 
+ When someone is doing interactive analysis and expects that the data they read in is about
 10s, but it turns into 1,000s, the user typically tends to either wait

### 2015-07-10 narrative/proposed-work.tex
changed from manish, as well as other minor changes.

@@
 Additionally we recognize that HPC applications exhibit a set
-of common I/O patterns \cite{lofstead2011six,polte2009and,tian2011edo,tian2012system},
+of common I/O patterns \cite{lofstead2011six,polte2009and,tian2011edo,tian2012system}.
 Throughout our project we will continuously focus on creating research artifacts 
 which can optimize these I/O patterns on current and
-exascale systems. One important pattern is Checkpoint/Restart (C/R) and we will focus on optimizing this pattern,
- and also using many of the important research techniques for using NVRAM and neighbor's NVRAM for storing and retrieving
-the C/R data. The migration and purging of this will be optimized by our SSIO layer using the utility function of the C/R data so that this can be stored for short periods of time, and then purged.
+exascale systems. One key focus is Checkpoint/Restart (C/R), leveraging NVRAM to accelerate
+checkpoint and data retrieval. 
+The migration and purging of this will be optimized by SIRUS using the utility function of 
+the C/R data so that this can be stored for short periods of time, and then purged.
 

### 2015-07-11 cover-collab.tex
made addtions to the collaborative cover page

@@
-\vfill
+\newpage
+
 \begin{center}
@@
 \end{center}
-
-\vfill
 
@@
 
-\input{cover-collab-table}
-
-\vfill
+%\input{cover-collab-table}
 \begin{center}
 \textbf{\large Leadership Structure}
-\medskip
 \end{center}
@@
 University of California, Santa-Cruz. 
+PI Klasky will be responsible for project oversight, reporting to DOE, and
+final decision making in event of major disagreements.
+We define an area lead for each major research area plus
+an integration lead responsible for coordinating periodic integration tests. The area leads plus the project PIs
+form the management team, which will meet weekly to review progress and revise plans as needed.
 
@@
 
-\vfill
 \begin{center}
 \textbf{\large Facilities, Equipment, and Resources}
-\medskip
 \end{center}
-
-This proposal brings together the expertiese and facilities from two
+This proposal brings together the expertise and facilities from two
 national laboratories, Oak Ridge National Laboratory and Sandia National
@@
  In addition, by
-sharing experiese and personnel costs, it represents
+sharing expertise and personnel costs, it represents
 a collaboration with other large programs within the Office of
@@
 %% collaborative group.
+\vspace{2ex}
 
-\vfill
+\noindent
+{\bf Description of facilities, equipment, and resources}: We summarize these here; detailed information is provided in the supplementary documents.
+\begin{itemize}
+\item
+{\bf Oak Ridge National Laboratory}. \underline{Computers}:
+Oak Ridge Leadership Facility (discretionary time for SNS experiments), including
+Titan, Gaea (a Cray XE6 system), the ORNL Institutional Cluster (OIC),
+and Frost. A cloud resource, CADES, can be made available to
+researchers in this project, as well as the sith cluster which has a Lustre file
+sysem, 2 SSD on each node, and access to the cneter wide Spider file system.
+ \underline{Networks}: 100 Gbps connection to ESnet.
+\vspace{-1ex}
+\item
+{\bf Rutgers University}. \underline{Computers}: Excalibur BGP, Clusters, CometCloud, and CAPER testbeds. \underline{Networks}: 10 Gbps to Internet2.
+\vspace{-1ex}
+\item
+{\bf Sandia National Laboratory}.
+\underline{Computers}: 264 TF Red Sky system and many testbed systems
+\underline{Networks}: 100 Gbps to ESnet
+\item
+{\bf UCSC}.
+\underline{Computers}:  48 node system and access to the ISSDM storage resource
+\underline{Networks}: 1 Gbps to ESnet
+
+\end{itemize}
+
 \begin{center}

### 2015-07-11 narrative/qos.tex
minor edits.

@@
 Our ultimate goal, rather than just performance predictability, is to offer
-predictable Quality of Service (QoS) for the storage system.
+predictable QoS for the storage system.
 %Performance quality can be expressed in either relative or absolute

### 2015-07-11 narrative/enabling_tech.tex
try to rewrite the data description

@@
 systems, data staging systems~\cite{zheng:2010:predata,docan:2010:dataspaces,oldfield:2006:nessie}, workflow
-systems~\cite{Dayal:2014:flexpath}, and even through the WAN network. 
+systems~\cite{Dayal:2014:flexpath}, and even through the WAN network~\cite{zhang2014active}. Further,
+I/O rerouting and throttling techniques \cite{liu_hotstorage,liu_helloadios} have been developed 
+within ADIOS to reduce contentions on best-effort storage systems and this is our initial
+effort in tackling storage load imbalance and improving storage QoS on current DOE systems.
 Part of this project's goals are to understand where the dividing line

### 2015-07-11 narrative/qos.tex
minor edits.
minor thing on QoS.

@@
 block allocation tables.
-
 An opposite example of this central management strategy are ``pile'' systems,
@@
 flexibility. Our additional challenge is incorporating both the storage tiers
-and potential data quality metrics.
+and data utility.
 
@@
 are analyzed to examine the utilization distribution of different components.
-Figures (a), (b) and (c) show the resource usage distribution for OSTs, OSSes,
-and LNETs, respectively. We observe a significant usage variation
+Figures (a), (b) and (c) show the resource usage distribution for OSTs (object storage
+target), OSSes (object storage server),
+and LNETs (Lustre Networking), respectively. We observe a significant usage variation
 across components of any given type (e.g., OST, OSS or LNET). For example, some

### 2015-07-11 narrative/qos.tex
finished qos.

@@
 be automatically enforced via user- or administrator- specified
-\textit{policies}\S\ref{sec:data-description}.
+policies.
 
@@
 important goal of the proposed project is to enable applications to reason
-about the trade-off between resolution and latency, adding yet another
+about the trade-off between accuracy and performance, adding yet another
 dimension to how tasks are mapped to resources.
@@
 
-\item To offer latency/latency trade-offs, the system must be able
+\item To offer performance and accuracy trade-offs, the system must be able
 to quickly generate a number of data production plans, involving
-different parts of the hierarchy and different data resolutions.
-It will then use the utilization model to estimate the latency for
-each production plan and the resolutions they can provide. Here
+different parts of the hierarchy and different data accuracy.
+It will then use the utilization model to estimate the performance for
+each production plan and the accuracy they can provide. Here
 again, pseudo-random selection could reduce the number of resources

### 2015-07-11 narrative/placement.tex
minor edits on life cycle.

@@
 Deciding what data to purge and when to purge it can be non-trivial, and it is 
-essential that this is done with the application in the loop. One approach we will 
-explore in this efforts will be extending the notion of system and application utility. 
-The former can be used to determine when data object need to purged while the 
-latter can help determine the relative importance of data objects to the 
-application and identify which data objects can be purged. Developing policies 
+essential that this is done with the application in the loop. 
+%One approach we will 
+%explore in this efforts will be extending the notion of system and application utility. 
+%The former can be used to determine when data chunks need to purged while the 
+%latter can help determine the relative importance of data chunks to the 
+%application and identify which data chunks can be purged. 
+Developing policies 
 for purging is a more challenging problem and will require engaging the 

### 2015-07-11 narrative/qos.tex
modified:   bib/maltzahn.bib 	modified:   narrative/qos.tex

@@
 
+Resource management and QoS have to be carefully coordinated.
+We use the term ``resource'' both for hardware, including computational, communication, and storage devices, and for data items such as data structures, blocks, objects, and files.
+To satisfy a request, the system has to locate resources that can satisfy the request: both devices and data items (or allocate the space for data items in case of writes).
+
 \subsubsection{Resource Management}
 Careful resource management strategies that can deal with placement and
-retrieval of a large number of data chunks spread across the entire storage
+retrieval of a large number of data items spread across the entire storage
 hierarchy are key for predictable performance. 
@@
 %dominant, component of overall read latency. 
-Locating resources is typical a distributed systems problem. Centralized
+Locating resources in a distributed system is a well-known problem: centralized
 solutions suffer scalability challenges due to centralized, serialized system
-updates.  POSIX-compliant file systems are good examples of this
-strucutre--most updates to files require updates to their centrally maintained
-block allocation tables.
-An opposite example of this central management strategy are ``pile'' systems,
-which, instead of tracking updates, they search for resources upon request.
-Pile systems feature efficient writes because clients can write wherever there
-is free space with low bookkeeping overhead.  However, unbounded search is
-possible, e.g., when it is unclear whether the resource actually exists.
-Methods to make searches more efficient without reverting to centralized
+updates. Many POSIX-compliant distributed file systems are good examples as updates to files require updates to their centrally maintained
+metadata. %block allocation tables. 
+% Carlos says: Parallel, POSIX-compliant file systems like Panasas,
+% Lustre, and Ceph have never used central block allocation tables
+% but only in local file systems they stripe over.
+Distributed metadata management solutions on the other hand pose
+challenges in keeping metadata management consistent and load-balanced
+across different workloads~\cite{sevilla:sc15}. Methods to make
+searches more efficient without reverting to centralized
 bookkeeping usually involves search space partitioning and a method to limit
@@
 
-Sirocco offers more of a pile struture rather than a file structure. One of the
-challenges we face is offering performance predictability more similar to
-POSIX-compliant file systems while we take advantage of the distributed,
-flexible storage management offered by Sirocco.
+\paragraph{Approach:} In this proposal we introduce the concept of ``pile systems'' as a contrast to file systems. The name
+is derived from personal information management user studies that divide users into ``pilers'' and ``filers''~\cite{malone:tois83}.
+More recent user studies showed that users generally use both piling and filing strategies, but to different degrees~\cite{whittaker:tochi01} (the lead author of this paper is now at UC Santa Cruz).
+Pile systems are systems that are capable of adding data with little ingest or metadata overhead or any placement restrictions and instead pay the cost of finding data at the time at which the data is retrieved.
+Pile systems are in contrast to file systems which maintain metadata and coordinate placement to minimize the cost of finding data.
+Pile systems enable efficient writes and flexible placement because clients can write wherever there
+is free space with little bookkeeping overhead.
+However, pile systems can have unbounded search times, especially when it is unclear whether the resource actually exists.
+A good example of a pile system design is Sirocco which offers exceptionally flexible storage management.
 
-\paragraph{Approach:} 
-In this proposal, we will investigate the trade-off in terms of predictable
-performance between ``filing'' and ``piling'' data.  This trade-off has a
+% Sirocco offers more of a pile struture rather than a file structure. One of the
+% challenges we face is offering performance predictability more similar to
+% POSIX-compliant file systems while we take advantage of the distributed,
+% flexible storage management offered by Sirocco.
+
+In the proposed project, we will investigate the trade-off in terms of predictable
+performance between ``filing'' and ``piling'' data, and we will
+prototype combinations of filing and piling strategies in Sirocco.
+The performance trade-off has a
 parallel in data management where there are two alternative paths to achieve
 knowledge. The analogue to filing is  to ``extract, transform, and load'' (ETL)
-data into a warehouse paying an upfront ingest cost to achieve fast querying.
-The analogue to piling is to not pay any upfront cost and later querying the
+data into a warehouse paying an upfront ingest cost to achieve fast 
+querying~\cite{vassiliadis:ijdwm09}.
+The analogue to piling is to not pay any upfront cost and instead query the
 raw data directly by using a data processing platform such as Hadoop.  Recent
 work has found that the combination of these two techniques can be optimized to
-perform better than any one technique alone~\cite{lefevre:sigmod14a}. We
+perform better than any one technique alone~\cite{lefevre:sigmod14a} (this is work supported in part by the DOE DAMASC project). We
 believe that a similar opportunity exists for combining filing and piling

### 2015-07-11 narrative/naming-service.tex
first bunch of changes in data description.

@@
 integration, we must address how to bridge the drastic differences between the
-decomposed objects of potentially varying fidelity with potentially multiple
-versions with a POSIX API-compatible interface.
+decomposed objects of potentially varying fidelity
+with a POSIX API-compatible interface.
 
@@
 metadata.  Providing this functionality requires the naming service be able to
-both return a list of data versions and an indication of data quality. This
-will be combined with a retrieval time estimate for each version by the
-middleware to determine what data version should be retrieved.  This estimate
+both return a list of data chunks and an indication of data quality. This
+will be combined with a retrieval time estimate by the
+middleware/application to determine what data chunks should be retrieved.  This estimate
 will be discussed elsewhere in this proposal.
@@
 services typically offered on POSIX directories and files, but can do it at the
-fork level instead. This allows a reduced quality data version to be available
-for the general users while the high-quality version would be limited to the
+fork level instead. This allows a reduced quality data chunk to be available
+for the general users while the high-quality data would be limited to the
 data creator. This and other considerations for security must be incorporated

### 2015-07-12 narrative/naming-service.tex
minor edits.

@@
 number of objects, we must take a new approach to naming.  Data migration, as a
-result of system pressures or user/system policies, further adds to the
+result of capacity pressures or user/system policies, further adds to the
 complexity of managing naming and discovery. Our naming service must maintain
 sufficient metadata information, track or be able to find objects as they move
-across storage layers, and facilitates data discovery. Since we need to also
-support access through POSIX APIs for existing application and system
+across storage tiers, and facilitates data discovery. Since we need to also
+support access through POSIX APIs for existing applications and system
 integration, we must address how to bridge the drastic differences between the
@@
 will be combined with a retrieval time estimate by the
-middleware/application to determine what data chunks should be retrieved.  This estimate
+middleware/application to determine what version should be retrieved.  This estimate
 will be discussed elsewhere in this proposal.
@@
 services typically offered on POSIX directories and files, but can do it at the
-fork level instead. This allows a reduced quality data chunk to be available
-for the general users while the high-quality data would be limited to the
+fork level instead. This allows a reduced quality verion to be available
+for the general users while the high-quality version would be limited to the
 data creator. This and other considerations for security must be incorporated
@@
 Our goal with this proposal is to offer a similar capability but with a finer
-granularity leveraging the inner structure of scientific data. It must also
+granularity leveraging the inner structure and utility of scientific data. It must
 handle data discovery efficiently for cases where data is spread across several
-storage layers or some of those pieces may have migrated based on system
-pressures or user or system policies. Instead of a single file, such as might
+storage tiers or some of those pieces may have migrated based on capacity
+pressures or user/system policies. Instead of a single file, such as might
 be used to store an entire timestep output for a simulation, we will
-demonstrate a metadata service with naming at the science variable level, but
+demonstrate a metadata service with naming at the application variable (e.g., particle) level, but
 with an associated list of all of the variable pieces and potentially different
 versions stored throughout the storage system. By shifting to a finer-grained
-approach, we will enable more effective use of close/fast/small storage tiers
+approach, we will enable more effective use of fast yet small storage tiers
 and prioritize objects with high importance.  Traditional HSM stores an entire
@@
 interest'' data can limit this costly tier usage to just 500 GB, greatly
-improving storage availability in the close/fast/small tier and focus
+improving storage availability in the fast tiers and focus
 performance/capacity costs to get the most science from the platform.  By
-only placing ``high interest'' data in the close/fast/small tier, we will
-effectively reduce pressure forcing data migration hurting both writing
+only placing ``high interest'' data in the highest tier, we will
+effectively reduce the pressure forcing data migration, which hurts both writing
 performance and the later data analytics required for generating scientific
@@
 container/object/fork/address tuple that can have associated attributes. We will
-add both a special attribute indicating that this record is of ``high
-interest'' as well adapting a migration mechanisms to use this attribute, if
+add a special attribute indicating that this record is of ``high
+interest'' as well adapting a migration mechanisms to take advantage of this attribute, if
 present, to strongly encourage keeping the record in the highest storage tier as possible.
@@
 to offer. Initial expectations suggest having bounded time guarantees for
-finding data are critical for offering the quality of service guarantees we
-wish to offer. Some potential approaches can build on CRUSH~\cite{weil:ceph}
+finding data are critical for offering the QoS guarantees.
+Some potential approaches can build on CRUSH~\cite{weil:ceph}
 from Ceph to offer a map of where to search for data sequentially. Given our
@@
 for pure object stores. By having a service that addresses the object
-collections that comprise a single thing, such as a variable or timestep
-output, just enough metadata is maintained to make the storage system usable
+collections that comprise a single entity, such as a variable or timestep
+output, sufficient metadata is maintained to make the storage system usable
 without additional heavy lifting by clients.  LWFS demonstrated a POSIX-style
@@
 accessing files.  Several efforts~\cite{patil:2007:giga+,carns:pvfs} have
-worked to reduce this contention by doing things like reducing the serialized
+worked to reduce this contention by reducing the serialized
 scope to a single directory or subtree, or allowing a single process in a
@@
 handle to other participants. While these approaches help, they do not address
-this key scalability limitation of serialized access--even if the metadata
+this key scalability limitation of serialized access - even if the metadata
 service is spread across multiple nodes. Instead, we will offer a short

### 2015-07-12 narrative/naming-service.tex
finishing up naming service.

@@
 
-While pure object stores, such as those popular in the big data
-domain~\cite{Fitzpatrick:2004:memcached}, others avoid this bottleneck by
+In pure object stores, such as those popular in the big data
+domain~\cite{Fitzpatrick:2004:memcached}, one avoids this bottleneck by
 strictly offering an object ID with the application required to manage how this
 ID maps to something of interest. This approach of removing the metadata
-service from the system level completely can work well for scale out
+service from the system level completely can work well for scaling out
 applications where data is created or consumed by a single process at a time or
-just in a work queue rather than potentially O(1 million) processes all
-actively collectively for a single ``object''. To address this case, having
+just in a work queue, rather than potentially O(1 million) processes all
+actively requesting for a single ``object''. To address this case, having
 some system integrated metadata services to associate names with these object
-is a preferable solution. While it will not have the same synchronous
-consistency semantics as POSIX, we will offer something as close as possible to
+is a preferable solution. We will offer solutions to
 address legacy application and command-line style data access and migration to
-other storage types using these POSIX semantics.
+other storage types using these POSIX semantics, without providing the same synchronous
+consistency semantics as POSIX.
 
@@
 joined by the additional challenge of coordinating with the other storage and
-application-layer services to offer the best access times possible for data
+application layer services to offer the best access times possible for data
 stored in the system. The developed metadata services that drive data
@@
 
-Our research in this area will address the following questions:
-(1) Since our metadata is distributed and partitioned, how do we respond to user
+Our research in this area will address the following specific questions:
+(1) Since the metadata will be distributed and partitioned, how do we respond to user
   requests within a specified time bound with accuracy? How do we estimate the

### 2015-07-12 narrative/data-reduction.tex
first few paragraphs of data refactoring.

@@
 The classical workflow where the entire dataset is written to storage for
-later analysis is no longer viable simply because the amount of generated data is too large.
-In the future, it will be vital to take advantage of a priori
+later analysis will no longer be viable in exascale, simply because the amount of generated data will 
+be too large due to capacity and performance reasons.
+In the future, it will be vital to take advantage of \textit{a priori}
 information when 1) writing and reading from the SSIO layer in order to gain
-performance and predictability, and to prioritize I/O so that the data that
-is most useful for end-user can be written in the time available, and 2)
+higher performance and predictability, and to prioritize the data that
+is most useful for end users so that I/O can be finished in the time available, and 2)
 performing {\em in situ} operations and analysis before storing the information.
+In the fusion use case described earlier which shares commonalities with many other
+DOE applications, \textit{a priori} information can indeed be provided by application
+scientists regarding which data should be sent over to the SSIO layer so that
+minimally, the most science relevant data can be available for subsequent analysis.
+This allows the science to be done even when the storage is in busy state - we believe
+in a shared multi-user environment, it occurs often \cite{liu_hotstorage}.
+In this project, this capability of prioritizing data is accomplished by data refactoring, 
+which includes re-organization and reductions of the data.
+There are many methodologies to refactor data and the best choice will generally
+be application dependent. However, our observation is that, once the choice is settled for an application, 
+it will not typically change from run to run.
 
-In the case of C. S. Chang described earlier, and common with most other
-applications scientists with whom we collaborate, information is provided on
-which data should be sent over to the SSIO layer so that the best, and most
-important information is available for subsequent analysis.
-This is generally accomplished using refactoring, which includes re-organization and reductions of the data.
-
-There are many possibilities for refactoring and the best choice will generally
-vary from case to case but, once settled, will typically not change from run to run. It
-will be important to understand when the time and resources required to
-identify and perform the ``best methods'' outweighs any eventual gains. Another
+To refactor data effectively and efficiently, it is important to understand
+the critical point when the time and resources required to
+identify and perform the ``best methods'' outweighs the intended gains. Another
 critical research question concerns the quantification and control of the loss
 of information resulting from refactoring the data and using a reduced dataset. 
-
 A basic issue allied to refactoring is to understand how much information is

### 2015-07-12 LOI/inc2.tex
Final committed pre-proposal with organized conflict lists.
substantial changes to intro and particularly the application use case

@@
 
-Our objective here is to reduce the time to knowledge, not just for a single
+Our objective here is to reduce the time to insight, not just for a single
 application, but for the entire workload in a multi-user environment, where
@@
  %% Manish: Commented this since it seems to be replicated below.  
-%  Our objective here is to reduce the time to knowledge, not just for a single
+%  Our objective here is to reduce the time to insight, not just for a single
 %application or workflow, but for the entire workload in the system. This is
@@
 as well as what's happening on the entire system. The impact of such a solution 
-will be a significant reduction in the time to knowledge, and the overall increase in 
+will be a significant reduction in the time to insight, and the overall increase in 
 the effectiveness and utility of exascale simulations.
@@
 
-  \includepdf[pages={1}]{COI/COI-all.pdf}
-  \includepdf[pages={2}]{COI/COI-all.pdf}
+  \includepdf[pages=-]{COI/COI-all.pdf}
 \end{document}

### 2015-07-12 abstract.tex
substantial changes to intro and particularly the application use case

@@
 new research in the management, storage and retrieval over the long lifespan
-of the extreme amounts of data that will be produced.
+of the extreme amounts of data that will be produced. Our thesis in this
+project is that adding application level knowledge
+about data to guide the actions of the storage system provides a substantial
+benefits to the organization, storage, and access to extreme scale data,
+resulting in improved productivity for computational science. 
 In this project we will
@@
 bottlenecks. 
-Our goal in this project is
+Our goal here is
 to address the associated Input/Output (I/O) and storage challenges in the context of
@@
 critical scientific processes; which is associated with {\bf theme two} of the Funding Opportunity Announcement. To that end, we will build on the
-capabilities offered by Adaptable I/O System (ADIOS) and DataSpaces at ORNL that provide I/O abstractions
+capabilities offered by Adaptable I/O System (ADIOS) and DataSpaces at ORNL
+and Rutgers University that provide I/O abstractions
 and services, the Sirocco peer-to-peer file system at
@@
 
-Our objective here is to reduce the time to knowledge, not just for a single
+We aim to reduce the time to insight, not just for a single
 application, but for the entire workload in a multi-user environment, where

### 2015-07-12 narrative/intro.tex
substantial changes to intro and particularly the application use case

@@
 for scientific simulation at exascale. 
-The central objective of this proposal, ``minimizing the time to knowledge''  
+The central objective of this proposal, ``minimizing the time to insight''  
 with respect to scientific computing, aligns perfectly with this sentiment. 
@@
 information rich data can then be used by the system to guide the  placement within 
-the tiered storage and will optimize the time to gain knowledge and insight from 
+the tiered storage and will optimize the time to gain insight from 
 exascale simulations.
@@
 %overcome the remaining obstacles in reaching exascale. 
-Instead of trying to optimize throughput, we will seek to reduce the time to knowledge.
+Instead of trying to optimize throughput, we will seek to reduce the time to insight.
 As extracting knowledge is the only metric that matters for scientific applications, we will
@@
 
-%%% Manish: I am not sure I fully understand the first part of the paragraph below.
-This proposal can be summarized as research into using the varying importance of simulation data over time to
-ensure that the right portions of data, with the right quality  are available to scientists when they are needed.
-Application-specific information about the importance of data will be used in conjunction with system and resource-specific information
-to optimize the scientists' experience in interacting with, and extracting knowledge and insight from their data.
-
+%%% Manish: I am not sure I fully understand the first part of the paragraph
+%%% below.
+The proposed work explores how application-specific information about the
+importance of data will be used in conjunction with system and
+resource-specific information to optimize the scientists' experience in
+interacting with, and extracting knowledge and insight from their data.
+%
 %The overarching idea of this proposal is that data from applications can be characterized by different
 %%levels of importance and different 
-%levels of importance over time, which then can be used to guide the analysis and visualization later to access the most important data and deliver some level of knowledge without waiting for access to the whole original dataset.
-Information about the importance of the data at the application level, information about resource status and performance of the multi-level storage system, and flexible data reorganization and reduction operations at all levels can be utilized to provide a quality of service and allow users to gain insight in a timely manner.
-
-This proposal combines knowledge and expertise 
+%levels of importance over time, which then can be used to guide the
+% analysis and visualization later to access the most important data and
+% deliver some level of knowledge without waiting for access to the whole
+% original dataset. Information about the importance of the data at the
+% application level, information about resource status and performance of
+% the multi-level storage system, and flexible data reorganization and
+% reduction operations at all levels can be utilized to provide a quality of
+% service and allow users to gain insight in a timely manner.
+The project combines knowledge and expertise 
 from a wide range of sources including many leadership class applications using the Adaptable Input/Output System (ADIOS) framework,
@@
 \paragraph{An Illustrative Example}
-We use an illustrative Fusion simulation use case that our team as been working on to highlight the challenges addressed in this proposal, and to 
+We use an illustrative fusion simulation use case that our team as been working on to highlight the challenges addressed in this proposal, and to 
 %The challenges addressed by this proposal are illustrated by a problem our team has been working on in the past six months. We use this motivating example to 
@@
 
-The use case describes a series of XGC1~\cite{chang2006integrated} simulations that C.
-S.
-Chang (PPPL), one of the largest users of Leadership Class Facilities (over 300M hours at DOE facilities), planned to launch.
-These simulations were expected to produce 100~PB of data over 10 days of total runtime on the Titan system, and assembled a team of experts including the user group and I/O and storage personnel  to help ensure that the maximum amount of information would be saved.
-Due to physical resource limitations, the simulation was immediately restricted to write about 10~PB.
-However, when the time to write and read, together with the financial cost to archive this volume of data were fully explored, the goal was even further reduced and they would only be able to write 5~PB over the 10-day run.
-%At each stage, the particular expert knowledge regarding the file system and the archival storage were not known to the user team.
-As these particulars were realized, significant modifications in the plans were made.
-Eventually, they were forced to make very hard decisions in reducing the desired 100~PB over a 10-day run to only 5~PB of data.
+The motivating use case is a series of simulations of the
+ITER~\cite{dietz1996iter} fusion experiments using the
+XGC1~\cite{chang2006integrated} application. XGC1 is one of the largest
+applications used at the Department of Energy's (DOE's) Leadership Class
+Facilities, with an allocation of over 300M hours. 
+% % simulations
+% % simulations planne
+% %that C. S. Chang (PPPL), 
+% one of the largest users of Leadership Class Facilities (LCFs) (over 300M hours at DOE facilities), planned to launch.
+These simulations were expected to produce 100~PB of data over 10 days of
+total runtime on the Titan system at the Oak Ridge Leadership Computing
+Facility (OLCF), and required a team of experts including
+the user group and I/O and storage personnel to help ensure that the maximum
+amount of information would be saved. Due to physical resource limitations,
+the simulation was immediately restricted to write about 10~PB. However,
+when the time to write and read, together with the financial cost to archive
+this volume of data was fully explored, the data size goal was further
+reduced to only 5~PB over the 10 day run.
 
-This forced the team to carefully scrutinize the large amount of data to be generated and save only the very most important pieces.
-Further, they had to divide this  reduced data into two categories: first, data they thought they would need to access while the simulation was running or shortly thereafter, and second, data they would need after the whole campaign was finished.
-The former category of data would be stored on disk, and the latter would be archived.
-Additionally, they needed to figure out how to efficiently retrieve these different categories of data for post-processing, analysis and visualization.
+This forced a careful scrutiny of the large amount of data to be generated
+and restriction of the output to only the most important pieces. Further,
+this reduced data set had to be divided into two categories: first, the data
+that would be accessed while the simulation was running or shortly
+thereafter, and second, the data that would be needed after the whole
+campaign was finished. The former category of data would be stored on disk,
+and the latter would be archived. Additionally, the entire team needed to
+figure out how to efficiently retrieve these different categories of data
+for post-processing, analysis and visualization.
 
-Finally, the team had to develop new application specific data reduction techniques and encode them into the ADIOS middleware I/O layer.
-In order to do this, they had to implement a rudimentary discovery system to determine which data were located on disk, and which were on tape archive.
-This solution worked well for initially but after the second day significant problems occurred: I/O variability became prohibitively expensive.
-Although the team had reserved resources for the run, the file system was shared, and a 10X variability in I/O speeds was observed.
+To accomplish this the application and storage teams developed new
+application specific data reduction techniques and added them to the ADIOS
+I/O middleware layer. A rudimentary discovery system was also created to
+track what data was on disk and what was on the tape archive. Initially this
+solution was sufficient, but eventually a new problem was introduced:
+Although the team had reserved resources for the run, the entire file system
+was shared \cite{shipman2012next} resulting in very high I/O variability
+\cite{lofstead:2010:io-variability}. Addressing this variability, caused by
+contention on shared resources, is an ongoing concern.
 
-What is needed here is a system to monitor, predict and manage resource usage in a dynamically adaptable way.
-Such a system can write more data when the resource was idle, and less data when it is busy.
-However, the most significant problem occurs after the simulation run completes.
-Basic visualizations, to get a high-level view of the simulation, can take significant time due to data access latencies.
-While some timesteps might be on disk, others are archived on tape (where transfers could take weeks).
-Detailed visualization and analysis of the simulation data are even more complicated.
+The problems faced by this user workflow motivate our research here. The use
+case demonstrated the three tiers of solutions that are required -
+application level knowledge of data and how it will be used, middleware
+management of data and resources, and storage system level scheduling of
+resources. It also provided us with new insights: First, the users and developers
+of the application data can, without a great loss of encoded knowledge,
+reduce the size of data that needs to be stored. Second, data storage must
+be managed with input from the users to correctly decide the target of
+storage operations. Third, shared resources must be managed in a way that
+provides sufficient performance to all applications without encumbering the
+applications with high variability in performance.
+
+These insights drive our proposed research approach.
+
+% Finally, the team had to develop new application specific data reduction
+% techniques and encode them into the ADIOS middleware I/O layer. In order to
+% do this, they  implemented a rudimentary discovery system to determine
+% which data was located on disk, and which was on the tape archive. This
+% solution worked well  initially but after the second day significant
+% problems occurred: I/O variability became prohibitive. Although
+% the team had reserved resources for the run, the file system was shared, and
+% a 10X variability in I/O speeds was observed.
+
+% What is needed here is a system to monitor, predict and manage resource
+% usage in a dynamically adaptable way. Such a system can write more data when
+% the resource was idle, and less data when it is busy. However, the most
+% significant problem occurs after the simulation run completes. Basic
+% visualizations, to get a high-level view of the simulation, can take
+% significant time due to data access latencies. While some timesteps might be
+% on disk, others are archived on tape (where transfers could take weeks).
+% Detailed visualization and analysis of the simulation data are even more
+% complicated.
 

### 2015-07-12 narrative/data-reduction.tex

@@
 execution time of the application. Accurate prefetching can only be achieved
-with access pattern analysis. This requires an SSIO layer which is capable of
+with access pattern analysis (\S\ref{sec:managing-data-life}). This requires an SSIO layer which is capable of
 prefetching data in an asynchronous fashion to take the bytes on the file.
@@
 	of the data are all grouped together from each object. This data
-	generally will have a higher utility then the data with the least
-	significant bytes.  This process has the following three steps which
-	need to be efficiently implemented when data gets placed to the storage
-	system and from the storage system.  First, the data needs to be
-	re-arranged which involves memory operations and needs to be done
-	in situ. This operation involves no communication.  Data needs to be
-	copied, so the memory requirements are increased. At most this requires
-	two copies of an individual dataset in memory, and we will investigate
-	techniques to allocate and deallocate this memory if the user will
-	specify that the data will be overwritten after it is written to the
-	storage system. This is often the case for many of the  quantities
-	written from the simulation, but there are many cases where, for
-	example, we want to write all of the particles from a Particle In Cell
-	(PIC) simulation.  Since the particles will be used later in the
-	calculation, we need to duplicate the storage. Our observation with
-	working with the XGC1, GTC, and PiconGPU simulations 
-	 is that we can temporary increase the
-	storage of the particles, and then release them since the temporary
-	arrays used in the calculations are often freed when a PIC iteration is
-	finished. 
+	generally will have a higher utility than the data with the least
+	significant bytes. 
+        %This process has the following three steps which
+	%need to be efficiently implemented when data gets placed to or retrieved from the storage.
+	The data needs to be re-arranged which involves potentially intensive 
+        memory operations and needs to be done as much in situ as possible without
+        involving the slow storage systems, and inter-processor communications.
+	%At most this requires
+	%two copies of an individual dataset in memory, and we will investigate
+	%techniques to allocate and deallocate this memory if the user will
+	%specify that the data will be overwritten after it is written to the
+	%storage system. This is often the case for many of the  quantities
+	%written from the simulation, but there are many cases where, for
+	%example, we want to write all of the particles from a Particle In Cell
+	%(PIC) simulation.  Since the particles will be used later in the
+	%calculation, we need to duplicate the storage. Our observation with
+	%working with the XGC1, GTC, and PiconGPU simulations 
+	% is that we can temporary increase the
+	%storage of the particles, and then release them since the temporary
+	%arrays used in the calculations are often freed when a PIC iteration is
+	%finished. 
 %
-\item Frequency-based re-organization.  Another common technique to classify
-	the importance of data is to first re-organize the data in frequency
-	space and then specify that as the frequency increases the importance
-	of data decreases.   Common techniques  used for streaming data,
-	and data reduction techniques such as JPEG-2000~\cite{jpeg2000} is to
-	code streams which have regions of interest that offer several
-	mechanisms to support spatial random access or region of interest
-	access at varying degrees of granularity. It is possible to store
-	different parts of the same data using different quality.  This allows
-	us to place the lowest frequency chunks in the fastest storage, and the
+\item Frequency-based re-organization.  Another common approach to classifying
+	the importance of data is to re-organize data according to it frequency.
+	This is commonly used in streaming data, and data reduction techniques
+        such as those used JPEG-2000~\cite{jpeg2000}, which allow mechanisms
+	to support spatial random access or region of interest
+	access at varying degrees of granularity. It demonstrates the possibility of storing
+	the same data using different quality. In relation to this project, this
+        capability allows us to place the lowest frequency chunks in the fastest storage, and the
 	highest frequency chunks on either the slowest storage tiers, or if the data
@@
 	 In order to fully take advantage of frequency based
-	re-organization we have devised a scheme where the data pre-conditioner
-	first sorts the data in bins of 1024 words, then we maintain the
-	sort order, and then we used wavelets and spline fits to then
-	re-organize the values.  This re-organization allows us to understand
-	that the resultant data is smoother now, and can be reduced in each one
-	of the bins.  \end{itemize}
+	re-organization, the team has devised a scheme where the data pre-conditioner
+	first sorts the data in bins with fixed length, and then wavelets and spline fits are used to
+	re-organize the values.  This re-organization effectively produces
+	a smoother data, allowing for much better reduction later.
+\end{itemize}
 

### 2015-07-12 narrative/naming-service.tex

@@
 \paragraph{Background:}
-Given the well documented difficulties with POSIX-compliant naming and our
-approach that decomposes what would be a single file into potentially large
-number of objects, we must take a new approach to naming.  Data migration, as a
-result of capacity pressures or user/system policies, further adds to the
-complexity of managing naming and discovery. Our naming service must maintain
-sufficient metadata information, track or be able to find objects as they move
-across storage tiers, and facilitates data discovery. Since we need to also
-support access through POSIX APIs for existing applications and system
-integration, we must address how to bridge the drastic differences between the
-decomposed objects of potentially varying fidelity
-with a POSIX API-compatible interface.
-
-Hierarchical Storage Management (HSM) systems~\cite{blaze:1992:hsm} offer a
-strict caching approach for managing different storage capacities trading off
-performance for capacity.  By maintaining a single namespace across all tiers,
-it is possible to list a single directory view with files stored at different
-tiers. While this approach to managing multi-tier storage works for whole files
-that fit on single logical devices, it is far from ideal for scientific
-simulations.
+POSIX-compliant naming suffers from well documented scalability limitations due
+to required operation serialization. Our approach would exascerbate this
+problem because we decompose what would be a single file into potentially large
+numbers of objects each broken into potentially multiple chunks to represent
+data utility at a finer grained level to facilitate efficient storage usage.
+Data migration, as a result of capacity pressures or user/system policies,
+further adds to the naming management and discovery complexity.  Temporarily to
+address performance requirements, we may make additional chunk replicas. In
+some cases, such as with tape, having a highly compressed replica available in
+a faster access medium near the tape system will offer efficient data probing
+before forcing data retrieval from tape. In general, the performance replicas
+will only exist as long as necessary to address performance requirements while
+the tape overview version would serve as a metadata-like interface to the tape
+archive. Since we must also support access through POSIX APIs for existing
+applications and system integration, we must address how to bridge the drastic
+differences between the objects and chunks of potentially varying fidelity with
+a POSIX API-compatible interface.  Our naming service must maintain sufficient
+metadata information to track or be able to find objects and chunks as they
+move across storage tiers facilitating data discovery and be able to offer a
+POSIX-compatible view for existing infrastructure integration.
 
@@
 between the user and the storage system extends into this area as well.
-Negotiating the data quality/retrieval time trade-off requires additional
-metadata.  Providing this functionality requires the naming service be able to
-both return a list of data chunks and an indication of data quality. This
-will be combined with a retrieval time estimate by the
-middleware/application to determine what version should be retrieved.  This estimate
-will be discussed elsewhere in this proposal.
+Users should be able to understand the available data quality and approximate
+retrieval times to enable making an appropriate trade-off for the current task.
+Supporting this interaction requires additional metadata.  The naming service
+must be able to both return a list of data chunks with their current
+location and an indication of data quality. This will be combined with a
+retrieval time estimate by the middleware/application to determine what version
+should be retrieved.  This retrieval time estimate will be discussed elsewhere
+in this proposal.
+
+While, Hierarchical Storage Management (HSM) systems~\cite{blaze:1992:hsm}
+seems similar to what we propose, that view only applies at the surface level.
+HSM offers a strict caching approach for managing different storage capacities
+trading off performance for capacity.  By maintaining a single namespace across
+all tiers, it is possible to list a single directory view with files stored at
+different tiers. HSM uses a whole file granularity. While this approach to
+managing multi-tier storage works for whole files that fit on single logical
+devices, it is far from ideal for scientific simulations. We will further
+optimize the HSM approach to make the most efficient use of scarce storage
+resources by only pulling the important/requested portions of a data set close
+to the user. We propose to address this HSM limitation and support the required
+complex naming and discovery metadata.
 
@@
 data creator. This and other considerations for security must be incorporated
-into the entire naming and metadata service.
+into the entire naming and metadata service, particularly for
+POSIX-compatibility.
 
-While strongly encouraging data to stay in a particular type of storage tier,
+While strongly encouraging data to stay in a particular storage tier,
 ultimately it must be possible to flush this data to make room for other uses.
@@
 \paragraph{Approach:}
-Our goal with this proposal is to offer a similar capability but with a finer
-granularity leveraging the inner structure and utility of scientific data. It must
-handle data discovery efficiently for cases where data is spread across several
-storage tiers or some of those pieces may have migrated based on capacity
-pressures or user/system policies. Instead of a single file, such as might
-be used to store an entire timestep output for a simulation, we will
-demonstrate a metadata service with naming at the application variable (e.g., particle) level, but
-with an associated list of all of the variable pieces and potentially different
-versions stored throughout the storage system. By shifting to a finer-grained
-approach, we will enable more effective use of fast yet small storage tiers
-and prioritize objects with high importance.  Traditional HSM stores an entire
-file on a tier making room if insufficient space is available.  With this shift
-to a partial variable granularity, a 1 PB output with 500 GB of ``high
-interest'' data can limit this costly tier usage to just 500 GB, greatly
-improving storage availability in the fast tiers and focus
-performance/capacity costs to get the most science from the platform.  By
-only placing ``high interest'' data in the highest tier, we will
-effectively reduce the pressure forcing data migration, which hurts both writing
-performance and the later data analytics required for generating scientific
-insights.  Sirocco identifies a data segment using a
-container/object/fork/address tuple that can have associated attributes. We will
-add a special attribute indicating that this record is of ``high
-interest'' as well adapting a migration mechanisms to take advantage of this attribute, if
-present, to strongly encourage keeping the record in the highest storage tier as possible.
+Our goal with this proposal is to offer an HSM-style capability but with a
+finer granularity leveraging the inner structure and utility of scientific
+data. It must handle data discovery efficiently for cases where data is spread
+across several storage tiers or some of those pieces may have migrated based on
+capacity pressures or user/system policies. Instead of a single file, such as
+might be used to store an entire timestep output for a simulation, we will
+demonstrate a metadata service with naming at the application variable (e.g.,
+particle) level, but with an associated list of all of the variable pieces and
+potentially different versions stored throughout the storage system (i.e., the
+chunk list). By shifting to a finer-grained approach, we will enable more
+effective use of fast yet small storage tiers and prioritize objects with high
+importance.  With this shift to a partial variable granularity, a 1 PB output
+with 500 GB of ``high interest'' data can limit this costly tier usage to just
+500 GB, greatly improving storage availability in the fast tiers and focus
+performance/capacity costs to get the most science from the platform.  By only
+placing ``high interest'' data in the highest tier, we will effectively reduce
+the pressure forcing data migration which hurts both writing performance and
+the later data analytics required for generating scientific insights.  Sirocco
+identifies a data segment using a container/object/fork/address tuple that can
+have associated attributes. We will add a special attribute indicating that
+this record is of ``high interest'' as well adapting a migration mechanisms to
+take advantage of this attribute, if present, to strongly encourage keeping the
+record in the highest storage tier as possible.
 
-Meanwhile, we need to be able to locate or ``discover'' the data should it
-move.  The challenge for discovery is that potentially, data will migrate from
-where it is initially stored to a new location within the storage system.
-Sirocco offers an ability to search for data that has moved as well as forcing
-a particular resilience-based replica be the ``authoritative'' version
-migrating the data to a particular location. We will investigate if the current
-Sirocco functionality is capable of supporting the new operating modes we wish
-to offer. Initial expectations suggest having bounded time guarantees for
-finding data are critical for offering the QoS guarantees.
-Some potential approaches can build on CRUSH~\cite{weil:ceph}
-from Ceph to offer a map of where to search for data sequentially. Given our
+We also must be able to locate or ``discover'' the data should it move.  The
+challenge for discovery is that potentially, data will migrate from where it is
+initially stored to a new location within the storage system.  Sirocco offers
+an ability to search for data that has moved as well as forcing a particular
+resilience or quality-based replica be the ``authoritative'' version migrating
+the data to a particular location. We will investigate if the current Sirocco
+functionality is capable of supporting the new operating modes we wish to
+offer. Initial expectations suggest having bounded time guarantees for finding
+data are critical for offering the QoS guarantees or even strong best effort
+support.  Some potential approaches can build on CRUSH~\cite{weil:ceph} from
+Ceph to offer a map of where to search for data sequentially. Given our
 multiple storage tiers, we would extend this using some mechanism to shift to
@@
 using the Sirocco object storage. We will investigate how to make the naming
-and discovery service scalable under the circumstances outlined above.
+and discovery service scalable under the circumstances outlined above,
+particularly to offer a coherent view given the potentially varying data
+quality across different chunks.
 
@@
 worked to reduce this contention by reducing the serialized
-scope to a single directory or subtree, or allowing a single process in a
+scope to a single directory or subtree or by allowing a single process in a
 collective file operation talk with the metadata service and distribute a
@@
 just in a work queue, rather than potentially O(1 million) processes all
-actively requesting for a single ``object''. To address this case, having
-some system integrated metadata services to associate names with these object
-is a preferable solution. We will offer solutions to
-address legacy application and command-line style data access and migration to
-other storage types using these POSIX semantics, without providing the same synchronous
-consistency semantics as POSIX.
+actively requesting for a single ``object''. To address this case, having some
+system integrated metadata services to associate names with these object is a
+preferable solution. We will offer solutions to address legacy application and
+command-line style data access and migration to other storage types using these
+POSIX semantics, without providing the same synchronous consistency semantics
+as POSIX.
 
@@
 application layer services to offer the best access times possible for data
-stored in the system. The developed metadata services that drive data
-compression and subsetting operations must have easy, consistent, and ideally
-non-blocking or locking access to this metadata service. We must investigate
-how to build such a metadata and naming service that also incorporates and
-maintains the additional metadata required to support our advanced
-functionality.
+stored in the system at available data quality levels. The developed metadata
+services that drive data compression and subsetting operations must have easy,
+consistent, and ideally non-blocking or locking access to this metadata
+service. We must investigate how to build such a metadata and naming service
+that also incorporates and maintains the additional metadata required to
+support our advanced functionality.
 
 Our research in this area will address the following specific questions:
-(1) Since the metadata will be distributed and partitioned, how do we respond to user
-  requests within a specified time bound with accuracy? How do we estimate the
-  completeness of the response?, 
+(1) Since the metadata will be distributed and partitioned, how do we respond
+to user requests within a specified time bound with accuracy? How do we
+estimate the completeness of the response?, 
 (2) For POSIX API requests, how do we respond if only varying data fidelity
@@
 (4) How much time should be allowed for searching for metadata given the user
-  specified limits for data retrieval since the user will still have to spend
-  the data retrieval time from different devices? How do we integrate the read
-  time overhead into the metadata requests?, and
+specified limits for data retrieval since the user will still have to spend the
+data retrieval time from different devices? How do we integrate the read time
+overhead into the metadata requests?, and
 (5) How do we provide appropriate security while maintaining scalability,

### 2015-07-12 narrative/data-reduction.tex
more on data refactoring.

@@
 
-We will also research the time it takes to re-organize the data and use the
-multi-layer approach to staging to understand where the data gets reduced and
-where it gets re-organized. Open questions are: 1) do we move the data over to
-other nodes to then reduce and re-organize, 2) do we first reduce the data and
-then later re-organize this data on different nodes?, or 3) do we use different
-cores on a node to do data-refactoring.  Similarly we will examine the cost of
-reading the data. It is clear that if we only read in the high-importance data
-it should improve the overall time to move the data from the SSIO layer to the
-application, but what happens when users want all of the data? There is an
-obvious penalty of having to take data in a hierarchical layout and then place
-this back to a single-layer view.  Additional research questions which must be
-addressed: Will some types of data (e.g. PIC data) be more
-conducive  for data-refactoring than others? How much semantic-domain knowledge
-will we need to get a dramatic improvement of the storage system over not doing
-anything special?
+To implement these plugin capabilities, we will build upon our existing data staging 
+capability and further allow data to be
+efficiently staged, re-organized and mapped to a multi-tiered storage. Various
+processing scenarios such as in-situ, in-transit or a combination of the two
+will be researched and developed to meet our needs.
+%The open questions that will be addressed include: 
+%1) do we move the data over to
+%other nodes to then reduce and re-organize, 2) do we first reduce the data and
+%then later re-organize this data on different nodes?, or 3) do we use different
+%cores on a node to do data-refactoring.  
+Similarly we will examine the cost of reading the data. In the case where only high importance being retrieved,
+overall time to move the data from the SSIO layer to the
+application is expected to fast. A key challenge is in the assessment of retrieving
+both high and low important data that reside across storage tiers and possibly migrate to
+new locations as a result of capacity pressure. There is a
+clear penalty of having to discovery data (\S\ref{sec:naming-discovery}), take data in a hierarchical layout and then re-compose
+this back to a single-layer view.  
+%Additional research questions which must be
+%addressed: Will some types of data (e.g. PIC data) be more
+%conducive  for data-refactoring than others? How much semantic-domain knowledge
+%will we need to get a dramatic improvement of the storage system over not doing
+%anything special?
 
-We will measure these methods using the fusion XGC simulation and the
+We will measure these methods using the fusion XGC, SPECFM3D~\cite{SPECFEM3D}, and the
+fusion GTC~\cite{klasky2003grid}, and the
 corresponding analysis routines which they use for their largest simulations.
-We will also work with the SPECFM3D~\cite{SPECFEM3D} code, and the
-fusion GTC~\cite{klasky2003grid} code.  In all cases we will measure their
-impact from their current simulations, and their projected simulations and
-analytics on exascale machines. Since all three of these codes are part of the
-OLCF CAAR~\cite{CAAR} project for early access, there is strong belief that the
+Since all three of these codes are part of the
+OLCF CAAR~\cite{CAAR} project for early access, there is strong belief that these
 simulations will be close to exascale simulations and we will be able to run
-these on prototypes of our SSIO layer to examine the full impact of
-data-refactoring.
+these application on SIRIUS to examine the full impact of data-refactoring.
 
-\paragraph{High Impact Challenges:} Maximal data reduction on the SSIO layer
+\paragraph{Challenges:} Maximal data reduction on the SSIO layer
 would result from storing only the code and the input parameters, and simply

### 2015-07-12 narrative/intro.tex
rewrote the perspective portions

@@
 
-\begin{table}[htbp]
-\small
-\vspace{2ex}
-\begin{center}
-\caption {Relevance to SSIO}
-\label{tab:call}
-\begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{2.2cm} |}\hline
-\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant S2E2 work} & \multicolumn{1}{|c|}{\bf Proposed research} \\\hline\hline
-1 & Data abstractions & Create new APIs  and annotations to guide data placement across the storage system & \S\ref{sec:data-description} \\\hline
-2 & Mapping science data models onto hierarchical storage &  Enable science-driven mapping, re-organization, reduction and prioritization of data chunks derive from application objects & \S\ref{sec:data-refactor}, \S\ref{sec:data-description},  \S\ref{sec:managing-data-life}\\\hline
-3 & Mechanisms for data movement across the storage hierarchy & Enable science-driven data placement and migration based on policies& \S \ref{sec:managing-data-life},  \S\ref{sec:migration} \\\hline
-4 &  Extend the interaction of middleware with the storage system & Manage data placement and migration between the application, middleware and storage system using an autonomic cross-layer approach & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
-5 & Design and implement of I/O middleware architectures  & Create APIs to guide the placement of chunks of data across the SSIO and autonomically move data& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
-6 &Approaches to improve the ability of SSIO software to support checkpoint/restart &  We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then later purged & \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:data-description} \\ \hline
-
-\end{tabular}
-\end{center}
-\vskip -0.5cm
-\end{table}
-
 %
@@
 
-To accomplish this the application and storage teams developed new
+To accomplish this, the application and storage teams developed new
 application specific data reduction techniques and added them to the ADIOS
@@
 \subsection{Overview of the proposed approach}
+
+\begin{table}[htbp]
+\small
+\vspace{2ex}
+\begin{center}
+\caption {Relevance to SSIO}
+\label{tab:call}
+\begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{2.2cm} |}\hline
+\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant S2E2 work} & \multicolumn{1}{|c|}{\bf Proposed research} \\\hline\hline
+1 & Data abstractions & Create new APIs  and annotations to guide data placement across the storage system & \S\ref{sec:data-description} \\\hline
+2 & Mapping science data models onto hierarchical storage &  Enable science-driven mapping, re-organization, reduction and prioritization of data chunks derive from application objects & \S\ref{sec:data-refactor}, \S\ref{sec:data-description},  \S\ref{sec:managing-data-life}\\\hline
+3 & Mechanisms for data movement across the storage hierarchy & Enable science-driven data placement and migration based on policies& \S \ref{sec:managing-data-life},  \S\ref{sec:migration} \\\hline
+4 &  Extend the interaction of middleware with the storage system & Manage data placement and migration between the application, middleware and storage system using an autonomic cross-layer approach & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
+5 & Design and implement of I/O middleware architectures  & Create APIs to guide the placement of chunks of data across the SSIO and autonomically move data& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
+6 &Approaches to improve the ability of SSIO software to support checkpoint/restart &  We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then later purged & \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:data-description} \\ \hline
+\end{tabular}
+\end{center}
+\vskip -0.5cm
+\end{table}
 
@@
 %
+Our approach is based on giving users the ability to define the utility of
+their data \S\ref{sec:data-description}. This includes the ability to assign
+varying levels of importance to individual data objects in their entirety,
+or to specific subsets and pieces. Key research challenges include: 1)
+developing an expressive and easy to use a program interface, 2) formulating
+a system model that can support autonomic decision, for example, for saving
+data based on user-defined importance and utility of
+data~\S~\ref{sec:managing-data-life}), and 3) defining the interface between
+the system and the back-end storage for both single and multiple running
+applications~\S~\ref{sec:qos}.
 
-Central to our approach is giving users the ability to define the utility of their data    
- \S\ref{sec:data-description}.
-This includes the ability to assign varying levels of importance to individual data objects in their entirety, or to specific
-subsets and pieces.
-Key research challenges include: 1) developing an expressive and easy to use a program interface, 2) formulating a system model that can support 
-autonomic decision, for example, for saving data based on user-defined importance and utility of data~\S~\ref{sec:managing-data-life}),
-, and 3) defining the interface between the system and the back-end storage for both single and multiple running applications~\S~\ref{sec:qos}.
-
-In the scenario above, the scientist would then specify a set of rules that would guide I/O and data management 
-at runtime. For example, if system resources are available and the estimated write times acceptable, the write 
-can proceed, but if this is not the case, the write may be postponed to a later time-step, or the data reduced using
-an in-situ techniques before it is saved. Data ``importance or utility'' can be used to place various data at  
-proper SSIO layers -- data with the highest utility (e.g., based on user defined importance or the frequency of access) 
-is maintained in faster storage for long periods of time, while data with lower utility can be migrated to lower layers 
-in the storage hierarchy.
+The scientist will be able to specify a set of rules that
+would guide I/O and data management at runtime. For example, if system
+resources are available and the estimated write times acceptable, the write
+can proceed, but if this is not the case, the write may be postponed to a
+later time-step, or the data reduced using an in-situ techniques before it
+is saved. Data ``importance or utility'' can be used to place various data
+at proper SSIO layers -- data with the highest utility (e.g., based on user
+defined importance or the frequency of access) is maintained in faster
+storage for long periods of time, while data with lower utility can be
+migrated to lower layers in the storage hierarchy.
 
@@
 \label{subsec:storage-perspective}
-From the perspective of the storage system, it is necessary to manage the data from
-a given user request amongst all of the requests from all users, and try to optimize the
-entire system accordingly. Today the problem is that when a user writes there is a
-tremendous amount of interference from other users on the system who may be
-writing or reading at the same time, leading to unpredictable performance.
-By requiring a certain QoS~\S\ref{sec:qos}, the system could potentially lock out and throttle other users. 
-The storage system must also automatically move data from one tier of the storage to
-another, without affecting the QoS that was offered at request
-time. The storage system must be able to organize
-data amongst all of the tiers of storage in a known, consistent way.
+From the perspective of the storage system, it is necessary to manage a
+single I/O request from a user in the context of all requests from all
+users, and try to optimize the entire system accordingly. This entails
+providing quality of service (~\S\ref{sec:qos} to the user to minimize
+resource contention, distribute the data across all of the tiers of storage
+in a known consistent manner (\S~\ref{sec:init-plac-data}), and automatically move data from one part of
+the storage hierarchy to another (~\ref{sec:migration}).
 
@@
 layers automatically, without user intervention. 
-Central to our approach is that the storage system needs to define bounds of time estimates
+Our approach here is that the storage system needs to define bounds of time estimates
 (min time, max time), and as our system evolves, we want to allow plug-ins which can help
@@
 \label{subsec:reading-perspective}
-From the perspective of reading, this project will look at key I/O patterns such as checkpoint-restart (CR), in-situ analytics and post-processing. 
-For example, in case of CR, we will support both, traditional file-based approaches leveraging fast storage tiers
-such as NVRAM (non-volatile RAM), as well as memory-based approach based on retrieving
-data directly from NVRAM. In case of in-situ and post-processing,  an important aspect is the latency -- here reducing latency and providing QoS 
- becomes important. 
- When someone is doing interactive analysis and expects that the data they read in is about
-10s, but it turns into 1,000s, the user typically tends to either wait
-another day for doing their analysis or gets frustrated and reads much less
-data than they really need to. Furthermore the user tends to read in just a
-small sample of the data without much knowledge if they are missing
-important pieces of the data~\S\ref{sec:migration}.
+
+From the reading perspective, the data must be accessible in a fast and
+predictable manner for all read access patterns, with different access
+patterns having their own requirements for latency, throughput and
+consistency. Here the user must be able to describe the read patterns and
+query for performance estimations (~\S~\ref{sec:data-description}), be able
+to quickly access the most important data (\S~\ref{sec:data-refactor}) and recieve
+performance gurantees (~\S\ref{sec:qos}) on the time required to complete
+the read request. The challenge here is to address the variety of read
+patterns that drive knowledge discover.
+
+For example, consider the case of Checkpoint/Restart (C/R) which needs a lot
+of data at high throughput to complete the restart in case of application
+failure, with latency being less important due to the batch nature of the
+request, and the data must be of sufficient quality to support an
+application restart without significant errors. 
+This is in contrast to interactive analysis, where the most important metric
+is the latency to start the analysis process. The quality here is more
+variable and based greatly on the expectations of the user, but performance
+predictability is important due to the interactive nature of the process.
+
+Our research objective here is to explore the interface that reading code
+can utilize to discover the important properites of data, the expectations
+about the time to access the data, and make decisions on the quality of data
+required to complete the task.
 

### 2015-07-12 narrative/naming-service.tex
more edits for clarity

@@
 this key scalability limitation of serialized access - even if the metadata
-service is spread across multiple nodes. Instead, we will offer a short
-duration to consistency POSIX view to offer the performance and consistency
-requirements an exascale application demands.
+service is spread across multiple nodes. The problem is the centralized listing
+of where file stripes are stored or other, similar information. Instead, we
+will offer a short duration to consistency POSIX view to offer the performance
+and consistency requirements an exascale application demands.
 
-In pure object stores, such as those popular in the big data
-domain~\cite{Fitzpatrick:2004:memcached}, one avoids this bottleneck by
-strictly offering an object ID with the application required to manage how this
-ID maps to something of interest. This approach of removing the metadata
-service from the system level completely can work well for scaling out
-applications where data is created or consumed by a single process at a time or
-just in a work queue, rather than potentially O(1 million) processes all
-actively requesting for a single ``object''. To address this case, having some
-system integrated metadata services to associate names with these object is a
-preferable solution. We will offer solutions to address legacy application and
-command-line style data access and migration to other storage types using these
-POSIX semantics, without providing the same synchronous consistency semantics
-as POSIX.
+Pure object stores, such as those popular in the big data
+domain~\cite{Fitzpatrick:2004:memcached}, avoid this bottleneck by only
+offering an object ID and offloading the ID to name mapping to the application.
+Removing the metadata service completely from the system level can work well
+for scale-out applications where data is created or consumed by a single
+process at a time or just in a work queue rather than potentially O(1 million)
+processes all actively requesting a single ``object''. To address this case,
+having some system integrated metadata services to associate names with object
+and chunk IDs is a preferable solution. We will offer solutions to address
+legacy application and command-line style data access and migration to other
+storage types using these POSIX semantics, without providing the same
+synchronous consistency semantics as POSIX.
 
@@
 The main challenge of incorporating the additional, rich metadata will be
-joined by the additional challenge of coordinating with the other storage and
+joined by coordinating with the other storage and
 application layer services to offer the best access times possible for data

### 2015-07-12 narrative/data-reduction.tex

@@
 of writing a minimal set of information, and placing an ``auditor'' code in a
-docker container so that data can be re-generated with a {\em local}
+container-like environment so that data can be re-generated with a {\em local}
 approximation of the full simulation. This is an entirely new approach to

### 2015-07-12 narrative/intro.tex
refer to our system as SIRUS.

@@
 at Rutgers. We will integrate this diverse expertise to investigate how to create
-a data, workflow, and application aware storage and I/O system.
+a data, workflow, and application aware storage and I/O system, which we refer to as
+{\bf SIRUS}.
 %This project brings together a team with strong expertise in I/O middleware (ORNL, Rutgers),

### 2015-07-12 narrative/data-reduction.tex
simple changes to reflect SIRUS

@@
 DOE applications, \textit{a priori} information can indeed be provided by application
-scientists regarding which data should be sent over to the SSIO layer so that
+scientists regarding which data should be sent over to the SSIO layer, which we refer
+to in our proposal as {\bf SIRUS}, so that
 minimally, the most science relevant data can be available for subsequent analysis.
 This allows the science to be done even when the storage is in busy state - we believe
-in a shared multi-user environment, it occurs often \cite{liu_hotstorage}.
+in a shared multi-user environment, it occurs often \cite{liu_hotstorage}, and gives users
+high performance variability.
 In this project, this capability of prioritizing data is accomplished by data refactoring, 
@@
 reduction in the total data compared with what one might achieve otherwise.
-However, even in the absence of such high level knowledge, we must equip the
-SSIO layers with generic data reduction techniques and re-organization
+However, even in the absence of such high level knowledge, we must equip SIRUS with generic data reduction techniques and re-organization
 techniques. For instance, certain basic information on the semantics of the
@@
 information regarding typical access patterns for data in a given application
-should be placed in the SSIO layers. Scientific applications often read and
+should be placed in SIRUS. Scientific applications often read and
 analyze data in a structured fashion and failure to respect this pattern, or to
@@
 execution time of the application. Accurate prefetching can only be achieved
-with access pattern analysis (\S\ref{sec:managing-data-life}). This requires an SSIO layer which is capable of
-prefetching data in an asynchronous fashion to take the bytes on the file.
+with access pattern analysis (\S\ref{sec:managing-data-life}).
+
+%% I don't understand this sententce This requires SIRUS
+% to be capable of
+%prefetching data in an asynchronous fashion to take the bytes on the file.
 
@@
 Similarly we will examine the cost of reading the data. In the case where only high importance being retrieved,
-overall time to move the data from the SSIO layer to the
+overall time to move the data from SIRUS to the
 application is expected to fast. A key challenge is in the assessment of retrieving
@@
 
-\paragraph{Challenges:} Maximal data reduction on the SSIO layer
+\paragraph{Challenges:} Maximal data reduction in SIRUS
 would result from storing only the code and the input parameters, and simply
@@
 too frequently. This then allows us to use the ``auditor'' to reproduce results during the analysis 
-process using an inexpensive computation and put less pressure on the SSIO layer.
+process using an inexpensive computation and put less pressure on SIRUS
 %

### 2015-07-12 narrative/evaluation.tex
eva.

@@
   {\bf Knowledge-centric system}           & More knowledge = faster access to higher quality data &  More knowledge = better resource utilization across the entire hierarchy \\ \hline
-  {\bf Predictable Performance Accuracy}   & More predictability = better quality of data &  Accuracy of the time bounds across multiple layers of the storage hierarchy wrote/read in time constrained situations\\ \hline
+  {\bf Predictable Performance and Accuracy}   & More predictability = better quality of data &  Accuracy of the time bounds across multiple layers of the storage hierarchy wrote/read in time constrained situations\\ \hline
 \end{tabular}

### 2015-07-13 narrative/data-reduction.tex
small wording changes

@@
 
-To implement these plugin capabilities, we will build upon our existing data staging 
+To implement these plug-in capabilities, we will build upon our existing data staging 
 capability and further allow data to be

### 2015-07-13 narrative/intro.tex
small wording changes

@@
 \begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{2.2cm} |}\hline
-\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant S2E2 work} & \multicolumn{1}{|c|}{\bf Proposed research} \\\hline\hline
+\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant SIRIUS work} & \multicolumn{1}{|c|}{\bf Proposed research} \\\hline\hline
 1 & Data abstractions & Create new APIs  and annotations to guide data placement across the storage system & \S\ref{sec:data-description} \\\hline

### 2015-07-13 narrative/proposed-work.tex
small wording changes

@@
 
-We propose to overcome these challenges by pioneering a new {\bf
+We propose to overcome the challenges described earlier by pioneering a new {\bf
 knowledge-centric approach} to a dynamic storage system and I/O layer which
-takes into account user prioritization and utility to optimize the SSIO layer.
-The approach is based on two underlying principles:
+optimize the SSIO layer on the following two principles:
+% The approach is based on two underlying principles:
 
@@
 established so science can be done on the exascale systems in a more efficient manner. Without
-predictable performance, not only can the runs be slowed down because of contentions \cite{liu_hotstorage}, 
+predictable performance, not only can the runs be slowed down because of
+shared resource contention \cite{liu_hotstorage,lofstead2010managing}, 
 but also it affects key science decisions, e.g., how much data reduction should be performed.
-It is clear that if a run is expected to be slow, an aggresive data reduction needs to be performed
-so that science can be done in a timely fashion, which has been reflected in the XGC run (\S~\ref{sec:introduction})
-that the PI recently participated.
+% It is clear that if a run is expected to be slow, an aggresive data reduction needs to be performed
+% so that science can be done in a timely fashion, which has been reflected in the XGC run (\S~\ref{sec:introduction})
+% that the PI recently participated.
 
@@
 on today's file systems. To accomplish this, we will provide a systematic
-approach for describing intentions and other knowledge from {\it user}
-space, as well as allow performance estimations and guarantees 
-from the underlyingree storage.
+approach for describing intentions and other knowledge from the {\it user}, 
+as well as allow performance estimations and guarantees 
+from the underlying storage.
 %For example, ADIOS contains most of the ``best practices'' for I/O optimizations for
@@
 but are detrimental to the long-term maintainability of applications, and
-are more sensitive to changes to system configuration. We have successfully
+are more sensitive to changes in the system configuration. We have successfully
 employed this strategy of separation of concerns in ADIOS, and believe that
@@
 which can optimize these I/O patterns on current and
-exascale systems. One key focus is Checkpoint/Restart (C/R), leveraging NVRAM to accelerate
+exascale systems. One area of focus is Checkpoint/Restart (C/R), leveraging NVRAM to accelerate
 checkpoint and data retrieval. 

### 2015-07-13 narrative/data-reduction.tex
Based on title its SIRIUS not SIRUS.

@@
 scientists regarding which data should be sent over to the SSIO layer, which we refer
-to in our proposal as {\bf SIRUS}, so that
+to in our proposal as {\bf SIRIUS}, so that
 minimally, the most science relevant data can be available for subsequent analysis.
@@
 reduction in the total data compared with what one might achieve otherwise.
-However, even in the absence of such high level knowledge, we must equip SIRUS with generic data reduction techniques and re-organization
+However, even in the absence of such high level knowledge, we must equip SIRIUS with generic data reduction techniques and re-organization
 techniques. For instance, certain basic information on the semantics of the
@@
 information regarding typical access patterns for data in a given application
-should be placed in SIRUS. Scientific applications often read and
+should be placed in SIRIUS. Scientific applications often read and
 analyze data in a structured fashion and failure to respect this pattern, or to
@@
 Similarly we will examine the cost of reading the data. In the case where only high importance being retrieved,
-overall time to move the data from SIRUS to the
+overall time to move the data from SIRIUS to the
 application is expected to fast. A key challenge is in the assessment of retrieving
@@
 
-\paragraph{Challenges:} Maximal data reduction in SIRUS
+\paragraph{Challenges:} Maximal data reduction in SIRIUS
 would result from storing only the code and the input parameters, and simply
@@
 too frequently. This then allows us to use the ``auditor'' to reproduce results during the analysis 
-process using an inexpensive computation and put less pressure on SIRUS
+process using an inexpensive computation and put less pressure on SIRIUS
 %

### 2015-07-13 narrative/intro.tex
Based on title its SIRIUS not SIRUS.

@@
 a data, workflow, and application aware storage and I/O system, which we refer to as
-{\bf SIRUS}.
+{\bf SIRIUS}.
 %This project brings together a team with strong expertise in I/O middleware (ORNL, Rutgers),

### 2015-07-13 cover-collab.tex
substantial changes to intro and particularly the application use case
added additional Sandia computing resources

@@
 {\bf Description of facilities, equipment, and resources}: We summarize these here; detailed information is provided in the supplementary documents.
-\begin{itemize}
+\begin{itemize}[itemsep=1pt,topsep=2pt]
 \item
@@
 {\bf Sandia National Laboratory}.
-\underline{Computers}: 264 TF Red Sky system and many testbed systems
-\underline{Networks}: 100 Gbps to ESnet
+\underline{Computers}: 1.1 PF Cielo, {\textasciitilde}40 PF Trinity, 264 TF Red Sky system and many testbed systems.
+\underline{Networks}: 100 Gbps to ESnet.
 \item
-{\bf UCSC}.
-\underline{Computers}:  48 node system and access to the ISSDM storage resource
-\underline{Networks}: 1 Gbps to ESnet
-
+{\bf University of California Santa-Cruz}.
+\underline{Computers}:  48 node system and access to the ISSDM storage resource.
+\underline{Networks}: 1 Gbps to ESnet.
 \end{itemize}
@@
 \vfill
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: "proposal"
+%%% End:

### 2015-07-13 narrative/data-reduction.tex
minor edits for formatting consistency.

@@
 In the future, it will be vital to take advantage of \textit{a priori}
-information when 1) writing and reading from the SSIO layer in order to gain
+information when (1) writing and reading from the SSIO layer in order to gain
 higher performance and predictability, and to prioritize the data that
-is most useful for end users so that I/O can be finished in the time available, and 2)
+is most useful for end users so that I/O can be finished in the time available, and (2)
 performing {\em in situ} operations and analysis before storing the information.
@@
 %The open questions that will be addressed include: 
-%1) do we move the data over to
-%other nodes to then reduce and re-organize, 2) do we first reduce the data and
-%then later re-organize this data on different nodes?, or 3) do we use different
+%(1) do we move the data over to
+%other nodes to then reduce and re-organize, (2) do we first reduce the data and
+%then later re-organize this data on different nodes?, or (3) do we use different
 %cores on a node to do data-refactoring.  

### 2015-07-13 narrative/data-reorg.tex
minor edits for formatting consistency.

@@
 First, the data needs to be re-arranged which involves memory operations and needs
-to be done in-situ. This operation involves no communication.  Data needs to be copied, so
+to be done in situ. This operation involves no communication.  Data needs to be copied, so
 the memory requirements are increased. At most this requires two copies an individual dataset
@@
 and then re-arranging the data in real-space once we re-organize this. The research questions
-we must address are: 1) what is the cost of re-organzing the data, 2) how do we
+we must address are: (1) what is the cost of re-organzing the data, (2) how do we
 

### 2015-07-13 narrative/intro.tex
minor edits for formatting consistency.

@@
 2 & Mapping science data models onto hierarchical storage &  Enable science-driven mapping, re-organization, reduction and prioritization of data chunks derive from application objects & \S\ref{sec:data-refactor}, \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}\\\hline
-3 & Mechanisms for data movement across the storage hierarchy & Enable science-driven data placement and migration based on policies& \S \ref{sec:managing-data-life},  \S\ref{sec:migration} \\\hline
+3 & Mechanisms for data movement across the storage hierarchy & Enable science-driven data placement and migration based on policies& \S\ref{sec:managing-data-life}, \S\ref{sec:migration} \\\hline
 4 &  Extend the interaction of middleware with the storage system & Manage data placement and migration between the application, middleware and storage system using an autonomic cross-layer approach & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
@@
 the system to make autonomic decisions at runtime
-~\S~\ref{sec:managing-data-life}~\S~\ref{sec:qos}.
+\S\ref{sec:managing-data-life} \S\ref{sec:qos}.
 %
@@
 varying levels of importance to individual data objects in their entirety,
-or to specific subsets and pieces. Key research challenges include: 1)
-developing an expressive and easy to use a program interface, 2) formulating
+or to specific subsets and pieces. Key research challenges include: (1)
+developing an expressive and easy to use a program interface, (2) formulating
 a system model that can support autonomic decision, for example, for saving
 data based on user-defined importance and utility of
-data~\S~\ref{sec:managing-data-life}), and 3) defining the interface between
+data \S\ref{sec:managing-data-life}), and (3) defining the interface between
 the system and the back-end storage for both single and multiple running
-applications~\S~\ref{sec:qos}.
+applications \S\ref{sec:qos}.
 
@@
 users, and try to optimize the entire system accordingly. This entails
-providing quality of service (~\S\ref{sec:qos} to the user to minimize
+providing quality of service (\S\ref{sec:qos} to the user to minimize
 resource contention, distribute the data across all of the tiers of storage
-in a known consistent manner (\S~\ref{sec:init-plac-data}), and automatically move data from one part of
-the storage hierarchy to another (~\ref{sec:migration}).
+in a known consistent manner (\S\ref{sec:init-plac-data}), and automatically move data from one part of
+the storage hierarchy to another (\S\ref{sec:migration}).
 
@@
 Overall, our goal is to enable predictable performance acorss the storage system, and we will develop mechanisms to enable this
-~\S\ref{sec:naming-discovery}.
+\S\ref{sec:naming-discovery}.
 
@@
 consistency. Here the user must be able to describe the read patterns and
-query for performance estimations (~\S~\ref{sec:data-description}), be able
-to quickly access the most important data (\S~\ref{sec:data-refactor}) and recieve
-performance gurantees (~\S\ref{sec:qos}) on the time required to complete
+query for performance estimations (\S\ref{sec:data-description}), be able
+to quickly access the most important data (\S\ref{sec:data-refactor}) and recieve
+performance gurantees (\S\ref{sec:qos}) on the time required to complete
 the read request. The challenge here is to address the variety of read

### 2015-07-13 narrative/enabling_tech.tex

--- 2015-07-10 narrative/enabling_tech.tex
@@
 ADIOS has proven effective for writing to POSIX file
-systems, data staging systems~\cite{zheng:2010:predata,docan:2010:dataspaces,oldfield:2006:nessie}, workflow
-systems~\cite{Dayal:2014:flexpath}, and even through the WAN network. 
+systems~\cite{lofstead:2009:adaptible}, data staging systems~\cite{zheng:2010:predata,docan:2010:dataspaces,oldfield:2006:nessie}, workflow
+systems~\cite{Dayal:2014:flexpath}, and even through the WAN network~\cite{zhang2014active}. Further,
+I/O rerouting and throttling techniques \cite{liu_hotstorage,liu_helloadios} have been developed 
+within ADIOS to reduce contentions on best-effort storage systems. This is our initial
+effort in tackling storage load imbalance and improving storage QoS on current DOE systems.
 Part of this project's goals are to understand where the dividing line

### 2015-07-13 narrative/intro.tex
format consistency

@@
 the system to make autonomic decisions at runtime
-\S\ref{sec:managing-data-life} \S\ref{sec:qos}.
+(\S\ref{sec:managing-data-life}, \S\ref{sec:qos}).
 %
 Our approach is based on giving users the ability to define the utility of
-their data \S\ref{sec:data-description}. This includes the ability to assign
+their data (\S\ref{sec:data-description}). This includes the ability to assign
 varying levels of importance to individual data objects in their entirety,
@@
 data based on user-defined importance and utility of
-data \S\ref{sec:managing-data-life}), and (3) defining the interface between
+data (\S\ref{sec:managing-data-life}), and (3) defining the interface between
 the system and the back-end storage for both single and multiple running
-applications \S\ref{sec:qos}.
+applications (\S\ref{sec:qos}).
 
@@
 users, and try to optimize the entire system accordingly. This entails
-providing quality of service (\S\ref{sec:qos} to the user to minimize
-resource contention, distribute the data across all of the tiers of storage
+providing quality of service to the user to minimize
+resource contention (\S\ref{sec:qos}), distribute the data across all of the tiers of storage
 in a known consistent manner (\S\ref{sec:init-plac-data}), and automatically move data from one part of
@@
 Overall, our goal is to enable predictable performance acorss the storage system, and we will develop mechanisms to enable this
-\S\ref{sec:naming-discovery}.
+(\S\ref{sec:naming-discovery}).
 

### 2015-07-13 narrative/data-reduction.tex

@@
 The classical workflow where the entire dataset is written to storage for
-later analysis will no longer be viable in exascale, simply because the amount of generated data will 
+later analysis will no longer be viable in exascale simply because the amount of generated data will 
 be too large due to capacity and performance reasons.

### 2015-07-13 narrative/data-reduction.tex
Lots of edits for clarity

@@
 In the future, it will be vital to take advantage of \textit{a priori}
-information when (1) writing and reading from the SSIO layer in order to gain
-higher performance and predictability, and to prioritize the data that
-is most useful for end users so that I/O can be finished in the time available, and (2)
+information when (1) writing and reading from the storage system to gain
+higher performance and predictability and prioritize the most useful data 
+for end users so that I/O can be finished in the time available, and (2)
 performing {\em in situ} operations and analysis before storing the information.
@@
 DOE applications, \textit{a priori} information can indeed be provided by application
-scientists regarding which data should be sent over to the SSIO layer, which we refer
+scientists regarding which data should be sent to the storage system, which we refer
 to in our proposal as {\bf SIRIUS}, so that
 minimally, the most science relevant data can be available for subsequent analysis.
-This allows the science to be done even when the storage is in busy state - we believe
-in a shared multi-user environment, it occurs often \cite{liu_hotstorage}, and gives users
+This allows the science to be done even when the storage is busy (we believe
+in a shared multi-user environment). It occurs often~\cite{liu_hotstorage} and causes
 high performance variability.
 In this project, this capability of prioritizing data is accomplished by data refactoring, 
-which includes re-organization and reductions of the data.
+which includes data re-organization and reductions.
 There are many methodologies to refactor data and the best choice will generally
@@
 content follows no obvious or explicit model).  Although scientific data
-generally contains random components (due to finite precision, measurement and
-calibration effects etc.), useful scientific data is never purely random.
+generally contains random components (due to finite precision and measurement and
+calibration effects), useful scientific data is never purely random.
 Broadly speaking, the path from data to knowledge consists of extracting
-underlying models or patterns from the datasets, and interpreting the resulting
+underlying models or patterns from the datasets and interpreting the resulting
 models.
 
-Ideally, scientists would like to perform the entire analysis stage {\em in situ}
-to extract the relevant information and in so doing would effectively
-circumvent the large data issue completely.  The catch, of course, is that this
+Ideally, scientists would like to perform the entire analysis stage {\em in
+situ} to extract the relevant information effectively circumventing the large
+data issue completely.  The catch, of course, is that this
 is unlikely to be possible since, by their nature, large scale simulations aim
-to discover new information that is often hidden in the form of higher order
+to discover new information often hidden in the form of higher order
 effects amongst the data deluge. In particular, this means if data thinning or
-truncation is applied haphazardly then one is in danger of effectively throwing
-out the information one is seeking. It would seem that one cannot store the
-entire data set in an easily accessible source, due to sheer size, yet one
-cannot reduce it prior to archival without risking losing the information one
-is trying to extract.  Viewed in this way, the problem would appear
+truncation is applied haphazardly, the sought data may be eliminated.
+The
+entire data set cannot be stored in an easily accessible source due to sheer size. Yet, the data cannot 
+be reduced prior to archiving without risking losing the desired information.
+Viewed in this way, the problem would appear
 intractable. However, the above discussion fails to recognize that much of the
@@
 the amount of data.  The difficulty stems from the fact that one does not know
-about this redundancy in advance \emph{without the benefit of priori
+about this redundancy in advance \emph{without the benefit of a priori
 knowledge}. This knowledge can often be provided by the user or can be acquired
@@
 
-Deep application knowledge means one can sometimes achieve drastically superior
-reduction in the total data compared with what one might achieve otherwise.
-However, even in the absence of such high level knowledge, we must equip SIRIUS with generic data reduction techniques and re-organization
-techniques. For instance, certain basic information on the semantics of the
-data is needed and must be supported by the overall infrastructure. Likewise,
-information regarding typical access patterns for data in a given application
+Deep application knowledge means one can sometimes achieve dramatically superior
+data reduction compared with what one might achieve otherwise.
+However, even in the absence of such high level knowledge, we must equip SIRIUS with generic data reduction and re-organization
+techniques. For instance, certain basic data semantics information 
+is needed and must be supported by the overall infrastructure. Likewise,
+information regarding typical data access patterns in a given application
 should be placed in SIRIUS. Scientific applications often read and
-analyze data in a structured fashion and failure to respect this pattern, or to
+analyze data in a structured fashion. Failure to respect this pattern or 
 exploit it through prefetching, can significantly increase the overall
-execution time of the application. Accurate prefetching can only be achieved
+application execution time. Accurate prefetching can only be achieved
 with access pattern analysis (\S\ref{sec:managing-data-life}).
@@
 represent the data in a form that becomes more amenable to compression.  (2)
-Removal of redundancy such as omission of duplicates, etc.  Finally, (3) the
+Redundancy removal such as duplicates omission.  Finally, (3) the
 resulting data is compressed using entropy coding. The level of compression
 that can be achieved in the final step is limited by the Shannon information
-content of the data \cite{Gray:book}, and many standard tools are available for
+content of the data~\cite{Gray:book} and many standard tools are available for
 efficient and perfect entropy coding. The potential for compression through
 removal of redundancy is generally minor. The main scope for achieving
-significant compression (e.g. a compression rate of two) therefore
+significant compression (e.g., a compression rate of two) therefore
 lies in the pre-processing step.
@@
 Research in the construction of pre-processing techniques is still in its
-infancy, especially in context of exascale simulations. Techniques that have
-been advocated range from simple difference filtering, bit-sorting, prefix
-transformation through to simply reordering the data.  What these, and other
-catch-all methods fail to exploit is a priori information on the nature and
-provenance of the data.
+infancy, especially in context of exascale simulations. Advocated techniques 
+range from simple difference filtering, bit-sorting, and prefix
+transformation to simply reordering the data.  What these, and other
+catch-all methods fail to exploit is a priori information on the data nature and
+provenance.
 
@@
 
-\item Precision based re-organization. This is where the most significant bytes
+\item {\em Precision based re-organization}. This is where the most significant bytes
 	of the data are all grouped together from each object. This data
@@
         memory operations and needs to be done as much in situ as possible without
-        involving the slow storage systems, and inter-processor communications.
+        involving the slow storage systems and inter-processor communications.
 	%At most this requires
@@
 %
-\item Frequency-based re-organization.  Another common approach to classifying
+\item {\em Frequency-based re-organization}.  Another common approach to classifying
 	the importance of data is to re-organize data according to it frequency.
-	This is commonly used in streaming data, and data reduction techniques
-        such as those used JPEG-2000~\cite{jpeg2000}, which allow mechanisms
-	to support spatial random access or region of interest
+	This is commonly used in streaming data and data reduction techniques
+        such as those used JPEG-2000~\cite{jpeg2000}. These mechanisms
+	support spatial random access or region of interest
 	access at varying degrees of granularity. It demonstrates the possibility of storing
 	the same data using different quality. In relation to this project, this
-        capability allows us to place the lowest frequency chunks in the fastest storage, and the
-	highest frequency chunks on either the slowest storage tiers, or if the data
+        capability allows us to place the lowest frequency chunks in the fastest storage and the
+	highest frequency chunks on either the slowest storage tiers, or if writing out the data
 	sizes are prohibitively costly, not even write out these pieces. 
@@
 	re-organize the values.  This re-organization effectively produces
-	a smoother data, allowing for much better reduction later.
+	a smoother data allowing for much better reduction later.
 \end{itemize}
@@
 capability and further allow data to be
-efficiently staged, re-organized and mapped to a multi-tiered storage. Various
-processing scenarios such as in-situ, in-transit or a combination of the two
-will be researched and developed to meet our needs.
+efficiently staged, re-organized, and mapped onto multi-tiered storage. Various
+processing scenarios such as in situ, in transit, or a combination of the two
+will be investigated to meet our needs.
 %The open questions that will be addressed include: 
@@
 %cores on a node to do data-refactoring.  
-Similarly we will examine the cost of reading the data. In the case where only high importance being retrieved,
+Similarly, we will examine the data reading cost. In the case where only high importance data is retrieved,
 overall time to move the data from SIRIUS to the
-application is expected to fast. A key challenge is in the assessment of retrieving
-both high and low important data that reside across storage tiers and possibly migrate to
+application is expected to be fast. A key challenge is in assessing retrieving
+both high and low importance data that reside across storage tiers and possibly migrate to
 new locations as a result of capacity pressure. There is a
-clear penalty of having to discovery data (\S\ref{sec:naming-discovery}), take data in a hierarchical layout and then re-compose
+clear penalty of having to discover data (\S\ref{sec:naming-discovery}) or take data in a hierarchical layout and then re-compose
 this back to a single-layer view.  
@@
 \paragraph{Challenges:} Maximal data reduction in SIRIUS
-would result from storing only the code and the input parameters, and simply
+would result from storing only the code and the input parameters and simply
 recalculating all of the results. This extreme is not feasible in practice
@@
 achieving a step change in data reduction. We shall investigate the possibility
-of writing a minimal set of information, and placing an ``auditor'' code in a
+of writing a minimal set of information and placing an ``auditor'' code in a
 container-like environment so that data can be re-generated with a {\em local}

### 2015-07-13 narrative/placement.tex
edits for clarity

@@
 potential insights from the simulations can be realized. 
-The overall data life cycle can be viewed as consisting of four key phases. In the first phase data 
+The overall data life cycle can be viewed as consisting of four key phases. In the first phase, data 
 is generated by a simulation or through a data acquisition system,
@@
 meet application requirements and constraints. This includes migration, 
-reorganization and reduction of the data, etc., to optimize resource
+reorganization and reduction of the data to optimize resource
 utilization, interference with running applications, and availability and
@@
 
-In the third phase the data is consumed for knowledge discovery (for example
-read into analytics to produce new scientific insights). It must be noted
+In the third phase, the data is consumed for knowledge discovery. For example
+read into analytics to produce new scientific insights. It must be noted
 that the knowledge discovery process does not remove the data from the
@@
 
-In the final phase the data is purged from the system. The
+In the final phase, the data is purged from the system. The
 decision to purge may be made by the system based on operational metrics from
@@
 % Our research will ultimately address the following questions:\\
-% 1) Can we do scalable migration without making the discovery process
+% (1) Can we do scalable migration without making the discovery process
 %     unbounded;\\
-% 2) What are the parameters and input for migration;
-% 3)  Can we use the additional application level knowledge of data to purge
+% (2) What are the parameters and input for migration;
+% (3)  Can we use the additional application level knowledge of data to purge
 %     portions of the data from storage without making the data too much less
 %     valuable?;\\
-% 3)  With so many objects and no centralized directory how do we know what
+% (4)  With so many objects and no centralized directory how do we know what
 %     is actually in the storage to be purged?;\\
-% 4)  How do purge in a way that doesn't require a central authority,
+% (5)  How do purge in a way that doesn't require a central authority,
 %     doesn't interfere with other I/O and isn't bottlenecked by some sort of
@@
 \label{sec:init-plac-data}
-When an application outputs or accesses data the storage and middleware layers needs to
+When an application outputs or accesses data, the storage and middleware layers needs to
 decide \textit{what} data is placed \textit{where} in the multi-level storage system. 
@@
 Our approach is to increase knowledge about the data and its use within the 
-application, and leverage this knowledge to drive data placement and overall management.
+application and leverage this knowledge to drive data placement and overall management.
 
@@
 \paragraph{Soliciting application {\em hints}:}
-As described in \S\ref{sec:data-refactor} one of the main components of our
-proposed storage system is the ability to reorganize, refactor and reduce data 
-as it is generated, and to reorganize and possibly regenerate the data as it is accessed. 
+As described in \S\ref{sec:data-refactor}, one of the main components of our
+proposed storage system is the ability to reorganize, refactor, and reduce data 
+as it is generated and to reorganize and possibly regenerate the data as it is accessed. 
 We carry this principle into the placement and movement of data by allowing
@@
 distinct areas. First, we will study the challenges and trade-offs of either
-augmenting the I/O interface with hints, or allowing the addition of an
+augmenting the I/O interface with hints or allowing the addition of an
 external specification that defines the use case. Our experience with
 developing modern I/O interfaces has shown that both techniques 
-have value~\cite{tongipdps15,qiansc15}, and we will investigate the set of hints that are embedded in the
+have value~\cite{tongipdps15,qiansc15} and we will investigate the set of hints that are embedded in the
 application code vs. those that are described within an non-compiled
@@
 at runtime together with predictive analysis requirements. This information may include 
-runtime spatial-temporal data access patterns, and/or physical network topology
-information, which will be tracked at runtime, and then be used to estimate and predict 
+runtime spatial-temporal data access patterns and/or physical network topology
+information, which will be tracked at runtime and then be used to estimate and predict 
 the data access behaviors of application workflow at runtime. 
-A key research aspect here will be to anticipate accesses, for example using learning techniques or using
-application knowledge, and use this information to prefetch data objects.
+A key research aspect here will be to anticipate accesses, for example, using learning techniques or using
+application knowledge and use this information to prefetch data objects.
 For example, we will identify and catalog key phases and access patterns

### 2015-07-13 abstract.tex
move abstract up.
minor edits.

@@
 \phantomsection
-\addstarredchapter{Abstract}
+%\addstarredchapter{Abstract}
 %%\adjustmtc
@@
 project is that adding application level knowledge
-about data to guide the actions of the storage system provides a substantial
+about data to guide the actions of the storage system provides substantial
 benefits to the organization, storage, and access to extreme scale data,

### 2015-07-13 cover-collab.tex
minor edits.

@@
 This is a collaborative proposal between 
-Oak Ridge National Laboratory, and
+Oak Ridge National Laboratory,
 Sandia National Laboratories, and
@@
 Science, including on data management and exascale monitoring, as well as
-ongoing development of the Sirocco file-system. 
+ongoing development of the Sirocco file system. 
 The proposed work will leverage existing facilities at all three

### 2015-07-13 narrative/management.tex
spell check

@@
 Careful management will be required to ensure that the problems and delays that are inevitable in any research project
-do not  disrupt overall progress.Here we define a detailed workplan,
+do not  disrupt overall progress.Here we define a detailed work plan,
 establish a  management structure with defined oversight responsibilities, and establish a schedule of regular meetings and
-reviews to ensure progress.  We will use google docs to share documents, emails lists for informal communications and a 
+reviews to ensure progress.  We will use Google Docs (http://docs.google.com) to share documents, emails lists for informal communications and a 
 weekly phone call in the main research areas to discuss project tasks and milestones.

### 2015-07-13 narrative/enabling_tech.tex

@@
 systems~\cite{Dayal:2014:flexpath}, and even through the WAN network~\cite{zhang2014active}. Further,
-I/O rerouting and throttling techniques \cite{liu_hotstorage,liu_helloadios} have been developed 
+I/O re-routing and throttling techniques \cite{liu_hotstorage,liu_helloadios} have been developed 
 within ADIOS to reduce contentions on best-effort storage systems. This is our initial
@@
 %They offer a way to request data set portions
-%from a one component by the next componet in a data flow. In DataSpaces case,
+%from a one component by the next component in a data flow. In DataSpaces case,
 %the data transfer does not have to correspond to the data from a single

### 2015-07-13 narrative/evaluation.tex

@@
 will use the multiple storage layers available on systems at both SNL and the
-OLCF, including the sith cluster discussed in the Equipment appendix. 
- We will perform tests in both the system level (on sith) and in
+OLCF, including the Sith cluster discussed in the Equipment appendix. 
+ We will perform tests in both the system level (on Sith) and in
 the user level (on titan), and evaluate our results using our application

### 2015-08-12 abstract.tex
Modified the abstract following Lucy's directive to shorten to half a page

--- 2015-07-10 abstract.tex
@@
 \enlargethispage{2\baselineskip}
+
 Scientific discovery at the exascale will not be possible without significant
-new research in the management, storage and retrieval over the long lifespan
-of the extreme amounts of data that will be produced.
-In this project we will
-demonstrate novel techniques to  facilitate efficient mapping of data
-objects, even partitioning individual variables, from
-the user space onto multiple storage tiers, and enable application-guided
-data reductions and transformations to address capacity and bandwidth
-bottlenecks. 
-Our goal in this project is
-to address the associated Input/Output (I/O) and storage challenges in the context of
-current and emerging storage landscapes, and expedite insights into mission
-critical scientific processes; which is associated with {\bf theme two} of the Funding Opportunity Announcement. To that end, we will build on the
-capabilities offered by Adaptable I/O System (ADIOS) and DataSpaces at ORNL that provide I/O abstractions
-and services, the Sirocco peer-to-peer file system at
-Sandia, and the object storage and annotation expertise of UC Santa Cruz, to
-explore application and multi-tier storage aware data management
-solutions.    This project
-brings together a  team with strong expertise in I/O middleware (ORNL, Rutgers), file system (SNL, UCSC) and storage (UCSC), and connects and
-coordinates these key storage components in a seamless fashion.
+new research into the management, storage and retrieval over the full lifespan
+of the extreme amounts of data that will be produced. Our thesis is that by
+adding application level knowledge about data to guide the actions of the
+storage system we obtain substantial benefits in the organization, storage, and
+access to extreme scale data, resulting in improved productivity for
+computational science. We will demonstrate novel techniques to facilitate
+efficient mapping of data objects from the user space onto multiple storage
+tiers, and enable application-guided data reductions and transformations to
+address capacity and bandwidth bottlenecks.  Our goal is to address the
+associated Input/Output (I/O) and storage challenges in the context of current
+and emerging storage landscapes, and expedite insights into mission critical
+scientific processes.  We aim to reduce the time to insight, not just for a
+single application, but for the entire workload in a multi-user environment,
+where storage is shared across users. We shall achieve this by allowing
+selectable data quality, through trading in accuracy and error in order to meet
+the time or resource constraint. We will explore beyond checkpoint/restart I/O,
+and will address the challenges posed by key data access patterns in the
+knowledge gathering process. Ultimately, we will take the knowledge from the
+storage system to provide vital feedback to the middleware so that the best
+possible decisions can be autonomically made between the user intentions and
+the available system resources.  We will test our prototypes on current and
+future Department of Energy (DOE) system with many of today's cutting edge
+applications to ensure our techniques can be used within our framework on
+current and next generation systems.
 
-Our objective here is to reduce the time to knowledge, not just for a single
-application, but for the entire workload in a multi-user environment, where
-the storage is shared among users.
-We achieve this goal by allowing selectable data quality, by trading its accuracy and error
-in order to meet the time or resource constraint. 
-We will explore beyond 
-checkpoint/restart I/O, and will address the challenges posed by 
-key
-data access patterns in the knowledge gathering process.
-Ultimately, we will take the knowledge from the storage system to provide vital feedback to the middleware 
-so that the best possible decisions can be autonomically made between the user intentions and
-the available system resources.  
-We will test our prototypes  on current and future Department of Energy (DOE) system with many of today's applications, including the
-XGC1, GTC, QMCPack, and SpecFM3D simulations. 
-Our solutions will provide  new functionalities and Application Program Interfaces (API) for:
-1) specifying, at the application level, data annotations that enable the
-quantification of the relative importance and utility of data objects and
-enable partitioning  across the storage hierarchy;
-2) specifying selectable performance/quality/cost tradeoffs from both the
-application and system perspectives;
-3) evaluating these tradeoffs at runtime during data placement and movement,
-and executing the resultant policies in an autonomic system using models,
-heuristics and continuous learning; and
-4) leveraging techniques such as application-aware data compression, 
-and  I/O utility to enforce these
-policies.
 \noindent

### 2015-08-13 abstract.tex
new figure

@@
 
-Scientific discovery at the exascale will not be possible without significant
-new research into the management, storage and retrieval over the full lifespan
-of the extreme amounts of data that will be produced. Our thesis is that by
-adding application level knowledge about data to guide the actions of the
-storage system we obtain substantial benefits in the organization, storage, and
-access to extreme scale data, resulting in improved productivity for
-computational science. We will demonstrate novel techniques to facilitate
-efficient mapping of data objects from the user space onto multiple storage
-tiers, and enable application-guided data reductions and transformations to
-address capacity and bandwidth bottlenecks.  Our goal is to address the
-associated Input/Output (I/O) and storage challenges in the context of current
-and emerging storage landscapes, and expedite insights into mission critical
-scientific processes.  We aim to reduce the time to insight, not just for a
-single application, but for the entire workload in a multi-user environment,
-where storage is shared across users. We shall achieve this by allowing
-selectable data quality, through trading in accuracy and error in order to meet
-the time or resource constraint. We will explore beyond checkpoint/restart I/O,
-and will address the challenges posed by key data access patterns in the
-knowledge gathering process. Ultimately, we will take the knowledge from the
-storage system to provide vital feedback to the middleware so that the best
-possible decisions can be autonomically made between the user intentions and
-the available system resources.  We will test our prototypes on current and
-future Department of Energy (DOE) system with many of today's cutting edge
-applications to ensure our techniques can be used within our framework on
-current and next generation systems.
+% Scientific discovery at the exascale will not be possible without significant
+% new research into the management, storage and retrieval over the full lifespan
+% of the extreme amounts of data that will be produced. Our thesis is that by
+% adding application level knowledge about data to guide the actions of the
+% storage system we obtain substantial benefits in the organization, storage, and
+% access to extreme scale data, resulting in improved productivity for
+% computational science. We will demonstrate novel techniques to facilitate
+% efficient mapping of data objects from the user space onto multiple storage
+% tiers, and enable application-guided data reductions and transformations to
+% address capacity and bandwidth bottlenecks.  Our goal is to address the
+% associated Input/Output (I/O) and storage challenges in the context of current
+% and emerging storage landscapes, and expedite insights into mission critical
+% scientific processes.  We aim to reduce the time to insight, not just for a
+% single application, but for the entire workload in a multi-user environment,
+% where storage is shared across users. We shall achieve this by allowing
+% selectable data quality, through trading in accuracy and error in order to meet
+% the time or resource constraint. We will explore beyond checkpoint/restart I/O,
+% and will address the challenges posed by key data access patterns in the
+% knowledge gathering process. Ultimately, we will take the knowledge from the
+% storage system to provide vital feedback to the middleware so that the best
+% possible decisions can be autonomically made between the user intentions and
+% the available system resources.  We will test our prototypes on current and
+% future Department of Energy (DOE) system with many of today's cutting edge
+% applications to ensure our techniques can be used within our framework on
+% current and next generation systems.
+
+This project explores the use of application level knowledge to reduce the
+time to insight for a workload (not just a single application) in a
+multi-user environment where storage and other resources are shared among
+users. Our approach is based on the notion of selectable data quality to
+enable a tradeoff between result accuracy and near-real-time data movement
+requirements for oversubscribed computational and storage resources. For the
+past several decades, extreme scale computational science has been conducted
+on dedicated computational and storage resources. Shared resources were
+overprovisioned resulting in little or no impact from resource contention.
+This changed recently. Execution times for scientific applications vary
+dramatically depending on external contention for shared resources. This
+execution time variance is particularly challenging in light of new usage
+models that have near-real-time requirements, e.g., providing information
+needed to control an experimental instrument like the Spallation Neutron
+Source.
+
+More specifically, we will explore the resource contention issues
+with spreading a single data set horizontally across a single tier in the
+storage hierarchy, such as spinning media, as well as vertically across
+different kinds of storage media, including non-volatile memory. To manage
+the varying storage media capacities and bandwidths, we provide mechanisms
+to express the required data quality and the near real-time requirements of
+the application. Our thesis is that by adding application level knowledge
+about data to guide storage system behaviors, we obtain substantial benefits
+in the organization, storage, and access to extreme scale data resulting in
+improved productivity for computational science. We will demonstrate novel
+techniques to facilitate efficient and effective data placement onto
+multiple storage tiers and enable application-guided data reductions and
+transformations to address capacity and bandwidth bottlenecks. Our goal is
+to address the associated data management challenges in the context of
+current and emerging storage landscapes and expedite insights into mission
+critical scientific processes
 

### 2015-08-13 abstract.tex
abstract word file and update

@@
 
-This project explores the use of application level knowledge to reduce the
-time to insight for a workload (not just a single application) in a
-multi-user environment where storage and other resources are shared among
-users. Our approach is based on the notion of selectable data quality to
-enable a tradeoff between result accuracy and near-real-time data movement
-requirements for oversubscribed computational and storage resources. For the
-past several decades, extreme scale computational science has been conducted
-on dedicated computational and storage resources. Shared resources were
-overprovisioned resulting in little or no impact from resource contention.
-This changed recently. Execution times for scientific applications vary
-dramatically depending on external contention for shared resources. This
-execution time variance is particularly challenging in light of new usage
-models that have near-real-time requirements, e.g., providing information
-needed to control an experimental instrument like the Spallation Neutron
-Source.
+This project explores the use of application level knowledge to optimize the
+times to insight across a workload of multiple applications in a multi-user
+environment with shared storage and network resources. The basis for this
+work is the notion of selectable data quality to explore the tradeoffs
+between accuracy of results, resource requirements, and time to insight on
+systems with shared, oversubscribed computational and storage resources. 
+In the last several years extreme scale computational science was conducted
+using dedicated computational resources, with shared resources, such as
+network and storage, being overprovisioned. This resulted in little to no
+impact from resource contention. The increased scale of recent, and incoming
+future systems, has placed a greater emphasis on mitigating contention.
 
-More specifically, we will explore the resource contention issues
-with spreading a single data set horizontally across a single tier in the
-storage hierarchy, such as spinning media, as well as vertically across
-different kinds of storage media, including non-volatile memory. To manage
-the varying storage media capacities and bandwidths, we provide mechanisms
-to express the required data quality and the near real-time requirements of
-the application. Our thesis is that by adding application level knowledge
-about data to guide storage system behaviors, we obtain substantial benefits
+Our thesis is that by adding application level knowledge
+about data to guide storage system behaviors, we will obtain substantial benefits
 in the organization, storage, and access to extreme scale data resulting in
@@
 current and emerging storage landscapes and expedite insights into mission
-critical scientific processes
+critical scientific processes.
 

### 2015-08-13 abstract.tex
updates

@@
 systems with shared, oversubscribed computational and storage resources.
-In the last several years extreme scale computational science was conducted
-using dedicated computational resources, with shared resources, such as
-network and storage, being overprovisioned. This resulted in little to no
-impact from resource contention. The increased scale of recent, and incoming
-future systems, has placed a greater emphasis on mitigating contention.
 
-Our thesis is that by adding application level knowledge
-about data to guide storage system behaviors, we will obtain substantial benefits
-in the organization, storage, and access to extreme scale data resulting in
+Our thesis is that by adding application level knowledge about data to guide
+storage system behaviors, we will obtain substantial benefits in the
+organization, storage, and access to extreme scale data resulting in
 improved productivity for computational science. We will demonstrate novel
-techniques to facilitate efficient and effective data placement onto
-multiple storage tiers and enable application-guided data reductions and
-transformations to address capacity and bandwidth bottlenecks. Our goal is
-to address the associated data management challenges in the context of
-current and emerging storage landscapes and expedite insights into mission
-critical scientific processes.
+techniques to facilitate efficient data placement onto multiple storage
+tiers, and enable application-guided data reductions to address capacity,
+bandwidth, and latency. Our goal is to address the associated data management
+challenges in the context of current and emerging storage landscapes and
+expedite insights into mission critical scientific processes.
 

### 2015-10-21 proposal/acronym.tex
moved proposal into a subdirectory in preparation for adding the project artifacts to this repo

%%%%% Acronym definitions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% \SVN $Id$

% Avoid printing the chapter heading if we don't want the acronyms printed
\ifthenelse{\boolean{printacro}}{
   \chapter{Glossary of Acronyms}
   \typeout{DEBUGGING: In acronym.tex chapter heading}
}{\newpage}

\typeout{DEBUGGING: In acronym.tex}

%%% Definitions of acronyms: \acro{id}[short name]{full name}
%%% In text, reference with \ac{id}
\begin{acronym}

% Autogenerated institutional acronyms with all-lower and all-upper versions
\input{acronym-inst}

% Autogenerated acronyms for the institutional PIs
\input{acronym-pis}

%%% Definitions of acronyms: \acro{id}[short name]{full name}

\acro{AACE}[AACE]{Architecture-Aware Compilation Environment}
\acro{ACTS}[ACTS]{Advanced Computational Software}
\acro{ALCF}[ALCF]{Argonne Leadership Computing Facility}
\acro{AM}[AM]{applied math}
\acro{APGAS}[APGAS]{asynchronous partitioned global address space}
\acro{API}[API]{application program interface}
\acro{ASCI}[ASCI]{Accelerated Strategic Computing Initiative}
\acro{ASCR}[ASCR]{Office of Advanced Scientific Computing Research}
\acro{ASC}[ASC]{Advanced Simulation and Computing}
\acro{AST}[AST]{abstract syntax tree}
\acro{BES}[BES]{Office of Basic Energy Sciences}
\acro{BOFs}[BOFs]{``Birds of a Feather'' sessions}
\acro{CASC}[CASC]{Center for Applied Scientific Computing}
\acro{CASL}[CASL]{Consortium for Advanced Simulation of Light Water Reactors}
\acro{CBHPC}[CBHPC]{Component-Based High-Performance Computing}
\acro{CCA}[CCA]{Common Component Architecture}
\acro{CCTTSS}[CCTTSS]{Center for Component Technology for Terascale Simulation Software}
\acro{CESAR}[CESAR]{Center for Exascale Simulation of Advanced Reactors}
\acro{CESM}[CESM]{Community Earth System Model}
\acro{CGM}[CGM]{Common Geometry Module}
\acro{CSE}[CSE]{computational science and engineering}
\acro{CSP}[CSP]{communicating sequential processes}
\acro{CS}[CS]{computer science}
\acro{DAMSEL}[DAMSEL]{Data Model Storage Library for Exascale Science}
\acro{DARPA}[DARPA]{Defense Advanced Research Projects Agency}
\acro{DFT}[DFT]{density functional theory}
\acro{DOE}[DOE]{Dept.~of Energy}
\acro{DSL}[DSL]{domain-specific language} % Cite with \acp to get plural
\acro{DVI}[DVI]{differential variational inequality}
\acro{ECDC}[ECDC]{Exascale Co-Design Consortium}
\acro{eDSL}[eDSL]{embedded domain-specific language}
\acro{EMSL}[EMSL]{Environmental Molecular Sciences Laboratory}
\acro{ESMF}[ESMF]{Earth System Modeling Framework}
\acro{ExMatEx}[ExMatEx]{Exascale Co-Design Center for Materials in Extreme Environments}
\acro{ExaCT}[ExaCT]{Exascale Simulation of Combustion in Turbulence}
\acro{FACETS}[FACETS]{Framework Application for Core-Edge Transport Simulations}
\acro{FASTMath}[FASTMath]{Frameworks, Algorithms, and Scalable Technologies for Mathematics Institute}
\acro{FES}[FES]{Office of Fusion Energy Sciences}
\acro{FLOPS}[FLOPS]{Floating Point Operations Per Second}
\acro{FNAL}[Fermilab]{Fermi National Accelerator Laboratory}
\acro{FOA}[FOA]{Funding Opportunity Announcement}
\acro{FSP}[FSP]{Fusion Simulation Program}
\acro{GPU}[GPU]{graphical processing unit}
\acro{GUI}[GUI]{graphical user interface}
\acro{HPC}[HPC]{high-performance computing}
\acro{IDE}[IDE]{interactive development environment}
\acro{INCITE}[INCITE]{Innovative and Novel Computational Impact on Theory and Experiment}
\acro{IPM}[IPM]{Integrated Performance Monitoring}
\acro{IPS}[IPS]{Integrated Plasma Simulator}
\acro{IR}[IR]{intermediate representation}
\acro{ISA}[ISA]{instruction set architecture}
\acro{ITAPS}[ITAPS]{Interoperable Tools for Advance Petascale Simulation}
\acro{IT}[IT]{Information Technology}
\acro{LCF}[LCF]{Leadership Computing Facility}
\acro{MCMD}[MCMD]{Multiple-Component Multiple-Data}
\acro{MC}[MC]{Monte Carlo}
\acro{MDD}[MDD]{model-driven design}
\acro{MD}[MD]{molecular dynamics}
\acro{MPI}[MPI]{Message Passing Interface}
\acro{MSCF}[MCSF]{Molecular Science Computing Facility}
\acro{MS}[M\&S]{modeling and simulation}
\acro{NDA}[NDA]{non-disclosure agreement}
\acro{NERSC}[NERSC]{National Energy Research Scientific Computing Center}
\acro{NNSA}[NNSA]{National Nuclear Security Administration}
\acro{NSF}[NSF]{National Science Foundation}
\acro{NUMA}[NUMA]{non-uniform memory access}
\acro{OASCR}[ASCR]{Office of Advanced Scientific Computing Research}
\acro{OLCF}[OLCF]{Oak Ridge Leadership Computing Facility}
\acro{OO}[OO]{object-oriented}
\acro{OS}[OS]{operating system}
\acro{PCIe}[PCIe]{Peripheral Component Interconnect Express}
\acro{PGAS}[PGAS]{partitioned global address space}
\acro{QA}[QA]{quality assurance}
\acro{RD}[R\&D]{research and development}
\acro{RDMA}[RDMA]{remote direct memory access}
\acro{SC}[SC]{Office of Science}
\acro{SDMAV}[SDMAV]{the anticipated SciDAC Institute in Scientific Data Management, Analysis, and Visualization}
\acro{SE}[SE]{software engineering}
\acro{SIMD}[SIMD]{single-instruction multiple-data}
\acro{SPMD}[SPMD]{single-program multiple-data}
\acro{SQA}[SQA]{software quality assurance}
\acro{SST}[SST]{Structured Simulation Toolkit}
\acro{SUPER}[SUPER]{Institute for Sustained Performance, Energy, and Resilience}\acro{QUEST}[QUEST]{Quantification of Uncertainty in Extreme Scale Computations}\acro{FASTMath}[FASTMath]{Frameworks, Algorithms, and Scalable Technologies for Mathematics}
\acro{SciDAC}[SciDAC]{Scientific Discovery through Advanced Computing}
\acro{TASCS}[TASCS]{Center for Technology for Advanced Scientific Component Software}
\acro{TCE}[TCE]{Tensor Contraction Engine}
\acro{TDD}[TDD]{test-driven development}
\acro{UHPC}[UHPC]{Ubiquitious High-Performance Computing}
\acro{UI}[UI]{user interface}
\acro{UML}[UML]{Unified Modeling Language}
\acro{UM}[UM]{University of Maryland}
\acro{UQ}[UQ]{uncertainty quantification}
\acro{VCS}[VCS]{version control system}
\acro{VV}[V\&V]{verification and validation}

\acro{LULESH}[LULESH]{Livermore Unstructured Lagrangian Explicit Shock Hydrodynamics}
\acro{SSCA}[SSCA]{Scalable Synthetic Compact Application}
\acro{DoD}[DoD]{Dept.~of Defense}
\acro{HPCMP}[HPCMP]{High Performance Computing Modernization Program}

\acro{DVFS}[DVFS]{dynamic voltage and frequency scaling}
\acro{TMR}[TMR]{triple modular redundancy}
\acro{ABFT}[ABFT]{algorithm-based fault tolerance}
\end{acronym}

### 2015-10-21 proposal/narrative/data-desc.tex
moved proposal into a subdirectory in preparation for adding the project artifacts to this repo

--- 2015-07-01 narrative/data-desc.tex
@@
 raw data. As an example, retrieving a 3D field data generated from a simulation and then visualizing it
-requires the coordinates and connectivity variable 1) to be accessed together and 2) with low
+requires the coordinates and connectivity variable (1) to be accessed together and (2) with low
 latency in order to achieve a good user experience. The description of data, 

### 2015-10-21 proposal/narrative/data-reduction.tex
moved proposal into a subdirectory in preparation for adding the project artifacts to this repo

--- 2015-07-13 narrative/data-reduction.tex
@@
 order representation might prove effective. It is useful to classify scientific
-data into two basic categories: structured data (in the sense that it satisfies
-a known or relatively simple model), or unstructured data (in which information
+data into two basic categories: regular data (in the sense that it satisfies
+a known or relatively simple model), or irregular data (in which information
 content follows no obvious or explicit model).  Although scientific data

### 2015-10-21 proposal/narrative/eviction-metrics.tex
moved proposal into a subdirectory in preparation for adding the project artifacts to this repo

--- 2015-07-01 narrative/eviction-metrics.tex
@@
-\subsubsection{Eviction Metrics}
+\subsection{Eviction Metrics}
 
@@
 lower resolutions can still occur without misses while requests
-outside of views or higher resolutions can leverage the information
+accessing outside of views or higher resolutions can leverage the information
 stored in auxiliary data. Conversely, data on a high tier can be

### 2015-10-21 proposal/narrative/intro.tex
moved proposal into a subdirectory in preparation for adding the project artifacts to this repo

--- 2015-07-13 narrative/intro.tex
@@
 % % simulations
-% % simulations planne
+% % simulations plane
 % %that C. S. Chang (PPPL), 
@@
 was shared \cite{shipman2012next} resulting in very high I/O variability
-\cite{lofstead:2010:io-variability}. Addressing this variability, caused by
+\cite{lofstead:2010:io-variability,liu_hotstorage}. Addressing this variability, caused by
 contention on shared resources, is an ongoing concern.
@@
 applications with high variability in performance. 
-
 These insights drive our proposed research approach.
@@
 5 & Design and implement of I/O middleware architectures  & Create APIs to guide the placement of chunks of data across the SSIO and autonomically move data& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
-6 &Approaches to improve the ability of SSIO software to support checkpoint/restart &  We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then later purged & \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:data-description} \\ \hline
+6 &Approaches to improve the ability of SSIO software to support checkpoint/restart &  We will have utility functions defined for the lifetime of data, and allow C/R data to be automatically placed in the fastest storage and then later purged & \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:data-description} \\ \hline
 \end{tabular}
@@
 reduce the range of the bounds.
-Overall, our goal is to enable predictable performance acorss the storage system, and we will develop mechanisms to enable this
+Overall, our goal is to enable predictable performance across the storage system, and we will develop mechanisms to enable this
 (\S\ref{sec:naming-discovery}).
@@
 query for performance estimations (\S\ref{sec:data-description}), be able
-to quickly access the most important data (\S\ref{sec:data-refactor}) and recieve
-performance gurantees (\S\ref{sec:qos}) on the time required to complete
+to quickly access the most important data (\S\ref{sec:data-refactor}) and receive
+performance guarantees (\S\ref{sec:qos}) on the time required to complete
 the read request. The challenge here is to address the variety of read
@@
 Our research objective here is to explore the interface that reading code
-can utilize to discover the important properites of data, the expectations
+can utilize to discover the important properties of data, the expectations
 about the time to access the data, and make decisions on the quality of data
@@
 %layer which will talk to the applications, with the Storage system. The
-%negotation between these layers will be a fundamental service which we will
+%negotiation between these layers will be a fundamental service which we will
 %create in this project to ensure that users will be able to save `'the
@@
 %Our approach leverages the expected characteristics of exascale storage
-%hardware. The storage layer will be partitioned into multiple heterogenous
+%hardware. The storage layer will be partitioned into multiple heterogeneous
 %tiers with vastly different performance characteristics. This difference

### 2015-10-21 proposal/narrative/migration.tex
moved proposal into a subdirectory in preparation for adding the project artifacts to this repo

--- 2015-07-09 narrative/migration.tex
@@
 at various tiers in the lower-level storage hierarchy. While Sirocco does
-autonomously group the distributed set of media into like groups and manage
+autonomously group the distributed set of media into groups and manage
 movement of data between these groups the various motivations to do so include

### 2015-10-21 proposal/narrative/proposed-work.tex
moved proposal into a subdirectory in preparation for adding the project artifacts to this repo

--- 2015-07-13 narrative/proposed-work.tex
@@
 checkpoint and data retrieval. 
-The migration and purging of this will be optimized by SIRUS using the utility function of 
+The migration and purging of this will be optimized by SIRIUS using the utility function of 
 the C/R data so that this can be stored for short periods of time, and then purged.

### 2015-10-21 proposal/participant-table.tex
moved proposal into a subdirectory in preparation for adding the project artifacts to this repo

--- 2015-07-10 participant-table.tex
@@
 \\
-\textbf{\ornl} & \makebox[\longestpientry][l]{\textbf{Scott Klasky} (\emph{klasky@ornl.gov})}\hspace*{1em}Hasan Abbasi, Mark Ainsworth, Qing Liu, Kimmy Mu, Feiyi Wang\\
+\textbf{\ornl} & \makebox[\longestpientry][l]{\textbf{Scott Klasky} (\emph{klasky@ornl.gov})}\hspace*{1em}Hasan Abbasi, Mark Ainsworth,\\
+\textbf{} & \makebox[\longestpientry][l]{\textbf{ }\hspace*{1in}}       Qing Liu, Kimmy Mu, Feiyi Wang\\
 \textbf{\rutgers} & \makebox[\longestpientry][l]{\textbf{Manish Parashar} (\emph{parashar@rutgers.edu})}\hspace*{1em}\\

### 2015-12-23 papers/2016-HotStorage/paper.tex
added template for Hot Storage paper

\documentclass{sig-alt-gov2}
\pdfpagewidth=8.5in
\pdfpageheight=11in
%\usepackage{subfig}
\usepackage{subfigure}
\usepackage[pdftex]{}
\usepackage{graphicx}
%\usepackage{todonotes}
\usepackage{listings}
\lstset {% general command to set parameter(s)
         language=C,
     basicstyle=\footnotesize,               % print whole listing small
%     keywordstyle=\color{black}\bfseries, % underlined bold black keywords
%     identifierstyle =\color{black},  % nothing happens
%     commentstyle=\color{black}\emph, % white comments
     %stringstyle=\ttfamily,          % typewriter type for strings
%     stringstyle=\color{black},       % typewriter type for strings
         tabsize=4,
         showtabs=false,
     showstringspaces=false}
%can't figure this one out for particles bandwidth
%\usepackage[caption,label]{subfig}

%\newcommand{\TITLE}{D\textsuperscript{\huge{2}}T: Doubly Distributed Transactions for High Performance and Distributed Computing}

\newcommand{\DDT}{D\textsuperscript{2}T~}
\newcommand{\DDTns}{D\textsuperscript{2}T}

\hyphenation{sub-trans-ac-tion}

\addtolength{\parskip}{-0.08in}

% in order for balance columns to work, it has to be before begin document...
%\balancecolumns

\begin{document}

%\conferenceinfo{PDSW'15,} {November 17, 2015, Austin, Texas}
%\CopyrightYear{2015}
%\crdata{978-1-4503-1103-8/11/11}
%\clubpenalty=10000
%\widowpenalty = 10000

\title{A Revolutionary Approach to I/O and Storage for Exascale Scientific Computing}

\numberofauthors{1}
\author{
%\alignauthor {Jay Lofstead\footnotemark[1], Ivo Jimenez\footnotemark[2],Carlos Maltzahn\footnotemark[2], Quincey Koziol\footnotemark[3], John Bent\footnotemark[4], Eric Barton\footnotemark[5]}\\
\alignauthor {Jay Lofstead\footnotemark[1], Carlos Maltzahn\footnotemark[2], Scott Klasky\footnotemark[3], Hasan Abbasi\footnotemark[3], Qing Liu\footnotemark[3]\\
Matthew Curry\footnotemark[1], Mark Ainsworth\footnotemark[5], Manish Parashar\footnotemark[4]}\\
\footnotemark[1]~~Sandia National Laboratories gflofst@sandia.gov\\
\footnotemark[2]~~University of California, Santa Cruz carlosm@soe.ucsc.edu\\
\footnotemark[3]~~Oak Ridge National Laboratory\\
\footnotemark[4]~~Rutgers University\\
\footnotemark[5]~~Brown University\\
}
\maketitle

\begin{abstract}

The longest existing parallel I/O APIs, HDF5, PnetCDF, and NetCDF, are focused
on treating the entire output from a paralell application as a single unit.
These libraries work to gather the data from all processes and arrange it such
that it is organized much like as if it were written from a single process.
This approach, in particular the two-phase collective I/O with data sieving
portion, has proven difficult to scale for 3-D domain decompositions at scale.
ADIOS shifted the focus from treating the entire application output as a single
entity to treating the output from each writer as a self-contained entity. This
eliminated the need for the problematic two-phase collective I/O and has
demonstrated good scalability well into the petascale era. Another shift is
required to address the needs of exascale systems. New system I/O bandwidth is
not keeping pace with the compute acceleration leading to accute bottlenecks.
While ``burst buffer'' technology will help address the performance gap, it
does not address the problem completely. This paper describes a new approach
to not just an I/O library, but also the underlying storage infrastructure to
address the needs of exascacle applications cognizant of projected exascale
platform characteristics and in the context of current industry trends.

\end{abstract}

%\category{D.4}{Software}{Operating Systems}
%\category{D.4.7}{Operating Systems}{Organization and Design}[hierarchical design]

%\terms{Design, Performance}

\section{Introduction}

%\begin{figure}[htbp]
%\vspace{-0.10in}
%\centering
%\includegraphics[width=\columnwidth]{images/arch-mapping}
%\vspace{-0.20in}
%\caption{Architecture and Component Mapping}
%\label{fig:arch-mapping}
%\vspace{-0.15in}
%\end{figure}

The rest of the paper is organized as follows. A brief overview of related work
is presented first in Section~\ref{sec:related}. Section~\ref{sec:end-user}
discusses the programmatic interface end users will see when interacting with
the storage array.
Section~\ref{sec:storage}
discusses the challenges and opportunities for deploying scalable storage
infrastructure.
Section~\ref{sec:qos} 
discusses the quality of service features proposed for this project.
Next is an exploration of the metadata challenges required to support such a
system in Section~\ref{sec:metadata}.
An initial prototype demonstration of the functionality is presented in
Section~\ref{sec:evaluation}.
The paper is concluded in
Section~\ref{sec:conclusion} with a summary of the broad issues covered in the
paper.

\section{Related Work}
\label{sec:related}

Ceph~\cite{weil:ceph} is a distributed object store and file system. It offers
both a POSIX and object interface including features typically found in parallel
file systems.
Ceph's unique striping approach uses pseudo-random numbers with a
known seed eliminating the need for the metadata service to track where each
stripe in a parallel file is placed.
PVFS~\cite{carns:pvfs} offers optimizations to reduce metadata server load,
such as a single process opening a file and sharing the handle.
It has been
commercialized in recent years as OrangeFS.
Lustre~\cite{braam:lustre-arch} has become the de facto standard on most major
clusters offering scalable performance and fine-grained end-user and
programmatic control over how data is placed in the storage system.
GPFS~\cite{schmuck:gpfs} offers a hands-off approach for providing good
performance for scaling parallel IO tasks and is used extensively by its owner,
IBM.
Panasas~\cite{panasas:architecture} seeks to offer a dynamically adaptive
striping system that detects the need for additional stripes for performance
and adjusts the file layout as necessary.

Other file systems, like GoogleFS~\cite{ghemawat:googlefs} and
HDFS~\cite{Shvachko:2010:hdfs}, address distributed rather than parallel
computing and cannot be compared directly.
The primary difference between
distributed and parallel file systems is the ability of the file system to
store and retrieve data simultaneously from multiple clients, in parallel, and
treat the resulting collection of pieces as a single object.  Distributed file
systems rely on a single client creating a file, but distributing the set of
files across a wide array of storage devices.
The other, popular distributed
file system of note is NFS~\cite{powlowski:1994:nfs3} that has been used for
decades for enterprise file systems.
These other file systems are mainly of
interest in the context of the ACG features of FFSIO and will be discussed more
in Section~\ref{sec:acg}.

\section{End-User API Layer}
\label{sec:end-user}

\section{Storage Challenges and Opportunities}
\label{sec:storage}

\section{Quality of Service Features}
\label{sec:qos}

\section{Metadata Challenges}
\label{sec:metadata}

\section{Demonstration}
\label{sec:evaluation}

This stack has an early prototype implementation intended to test concepts
rather than performance and scalability. It has focused on examining the
interaction of the different APIs for each layer to flesh out any detailed
requirements or concerns that may have been missed in the conceptualization of
this IO stack. To demonstrate the viability of the IO stack described in this
paper, we show some very early performance results from the untuned prototype.

\section{Conclusions}
\label{sec:conclusion}

\section{Acknowledgements}

\includegraphics[scale=0.07]{logos/doe_logo}
\includegraphics[scale=0.30]{logos/snl_logo}
\includegraphics[scale=0.35]{logos/nnsa_logo}
Sandia National Laboratories is a multi-program laboratory managed and operated
by Sandia Corporation, a wholly owned subsidiary of Lockheed Martin
Corporation, for the U.S. Department of Energy's National Nuclear Security
Administration under contract DE-AC04-94AL85000.

\bibliographystyle{abbrv}
\bibliography{paper}

\vfill\eject

\end{document}

### 2016-01-18 papers/2016-HotStorage/paper.tex

@@
 
-\title{A Revolutionary Approach to I/O and Storage for Exascale Scientific Computing}
+\title{SIRIUS: A Revolutionary Approach to I/O and Storage for Exascale Scientific Computing}
 
@@
 Matthew Curry\footnotemark[1], Mark Ainsworth\footnotemark[5], Manish Parashar\footnotemark[4]}\\
-\footnotemark[1]~~Sandia National Laboratories gflofst@sandia.gov\\
-\footnotemark[2]~~University of California, Santa Cruz carlosm@soe.ucsc.edu\\
-\footnotemark[3]~~Oak Ridge National Laboratory\\
-\footnotemark[4]~~Rutgers University\\
+\footnotemark[1]~~Sandia National Laboratories
+\footnotemark[2]~~University of California\\
+\footnotemark[3]~~Oak Ridge National Laboratory
+\footnotemark[4]~~Rutgers University
 \footnotemark[5]~~Brown University\\
@@
 \section{Introduction}
+
+The first generation of standardized APIs for scientific computing IO focused
+on serial output from a single process to a distributed or local file system.
+These systems still exist in the form of serial HDF5~\cite{hdf5} and the
+end-of-life NetCDF 3.x~\cite{netcdf}. The key feature of these APIs was the
+logical data model that eliminated any parallel data decomoposition artifacts
+from the actual data storage. This offered consistent performance and regular
+data access patterns.
+
+The second generation is really an evolution of the first with HDF-5 adding
+parallel capabilities and NetCDF bifurcating into NetCDF4~\cite{netcdf4} and
+PnetCDF~\cite{pnetcdf} with PnetCDF almost completely maintaining the data
+model and NetCDF4 shifting to use HDF5 underneath. The other improvement is
+incorporating knowledge of an underlying parallel file system to support large
+files written in unision from many processes.
+
+The third generation is a reaction to the second based on recognizing that
+maintaining the logical data model in physical form became inordinately
+expensive and offered a performance penalty in most cases when running at
+scale. Libraries like ADIOS~\cite{adios} shifted from writing complete
+variables from distributed processes as a single entity to treating all data
+from a single source to be treated as a unit. This eliminated the data
+reorganization at the cost of more complicated reading patterns. In spite of
+this extra complication, the performance is generally
+better~\cite{six-degrees}.
+
+SIRIUS represents the next generation. New platforms are introducting
+additional storage layers to accelerate IO operations and to compensate for
+plateauing storage array performance. At the same time, data sizes continue
+to grow. The fourth generation IO API will have to both address the IO API
+and middleware layer as well as the deeper, more complex storage hierarchy.
+
+Systems like ADIOS showed the performance advantage for decomposed data
+layouts, but still rely on a uniform storage layer. SIRIUS represents a shift
+to consider the local portion of each globally distributed variable as a
+management entity with data placement decisions across the entire storage
+hierarhy rather than on a single tier.  Decisions on what data to place where
+is driven by the data contents itself according to end-user specified
+priorities and system characteristics. For example, if temperature and pressure
+are nearly always retrieved together, organizing the data to keep just these
+two arrays together without any interleaving from other variables or scalars
+will accelerate later reading. If these variables could also be ``processed''
+such that the important data features are maintained in high resolution while
+less interesting regions are stored in highly compressed--even using lossy
+compression--forms, the amount of data written can be greatly increased. While
+accelerating writing is important. This approach also supports accelerating
+reading.
+
+This paper presents an overview of the efforts to build such a fourth
+generation storage and IO system. A discussion of the end-user API, storage
+challenges, quality of service features, and metadata challenges are all
+discussed briefly.
 
@@
 decades for enterprise file systems.
-These other file systems are mainly of
-interest in the context of the ACG features of FFSIO and will be discussed more
-in Section~\ref{sec:acg}.
 
@@
 
+The initial thoughts about an end-use API is to try to work within established
+APIs to ease user migration. However, some of the key features, such as the
+data ``refactor'' functionality and the quality of service requirements for
+both writing and reading are new. To manage these new requirements, an extended
+API is proposed. At a high level, the common API broad categories are
+discussed.
+
+\noindent\textbf{Open and Close} need to add additional information to include
+intent information. The intent hints to the system that....
+
+\noindent\textbf{Refactoring} may be a uniform operation across an
+entire variable. For example, splitting a double into three parts--two byes,
+four bytes, and two bytes, yields a two-byte value that is the majority of the
+data precision. The next four bytes would be most of the rest of the data
+precision. The last two bytes are the remainder of the precision. For quick
+overviews, just reading the two-byte portion is sufficient to get a rough idea
+of data contents. If particular data looks interesting, more and more data can
+be read to improve the precision. This can easily be applied uniformly by each
+process without coordinating. A more dynamic approach would be to apply a
+wavelet compression~\cite{wavelet} approach that generates different compressed
+data sizes at a fixed error rate depending on the data entropy. This is still a
+uniform operator, but the resulting data can be of radically different sizes.
+
+A more complex, but interesting approach is to offer an {\em Auditor} that
+performs the same computation, but with some simplification that greatly
+reduces computation intensity with only a minor loss of precision over the
+short term. Longer term, this auditor would need to be reset to bring it back
+into alignment with the full precision simulation. The much smaller reduced
+precision data could be used instead of the full precision in many
+circumstances.
+
+A middle ground approach between these two refactoring approaches is to use a
+non-uniform operator across a variable. Instead of, for example, splitting on
+byte boundaries to make multiple values, a function that identifies data
+features through a local analytics function could identify ``interesting'' data
+that should be maintained in full precision while other areas can be more
+highly compressed. Such an operator would decompose the data from a process
+further than a single process or node depending on the data features. This
+approach may balance precision better than wavelet compression, but at a higher
+analytics cost.
+
+We plan to investigate all three approaches as different applications will have
+different techniques that work best. Any performant storage system should
+offer the flexibility to incorporate application specific data refactoring
+techniques and sufficient metadata support for an arbitrary client to access
+the data successfully.
+
+\noindent\textbf{Quality of Service} deadlines are important for driving how
+much time is spent performing IO. At the same time, data quality must be
+maintained at a sufficient level for the resulting science to be possible.
+The system must offer a way for an end-user to negotiate with the system for
+trading off between time spent in IO and the amount of data written or read.
+Part of that tradeoff will be increased or reduced data quality, depending on
+the system.
+
 \section{Storage Challenges and Opportunities}
 \label{sec:storage}
+
+The supercomputing platforms being deployed today are all incorporating a
+Non-volatile memory (NVM) layer as a fast cache between compute and the storage
+array. The assumption is that nearly all IO will be performed against this
+layer to hide the much greater latency to the storage array. The challenge is
+the NVM capacity is expensive limiting the amount deployed. The unspoken
+limitation is that this fast layer still must drain to the relatively slow
+storage array. This performance mismatch will limit capacity and performance
+just due to the drain and data stage-in operations.
+
+A more effective way to use these resources is to only store the most
+important data in the fast tier and store lower precision representations of
+the rest of the data in the slower tiers. This lowers pressures on the NVM
+layer increasing performance overall by greatly reducing the drain/stage-in
+operations and reducing the data volumes stored on a per application basis.
 
@@
 
+SIRIUS offers a novel feature related to managing IO time as a first class
+characteristic. Rather than being at the mercy of the storage array
+performance, an end-user can describe the local data to the storage system and
+receive an estimate on how long that data will take to write. If that is
+acceptable within a margin of error, the API can write. If it is unacceptable,
+the end-user can make decisions about skipping the output or maybe spending a
+little time with a strong data refactoring to reduce the data size.
+
+On reading, a user may specify that they are willing to wait for an approximate
+time for an operation. This will negotiate with the storage system to determine
+the data quality returned. If the requested data quality cannot be retrieved
+within a given time, the storage array may be able to apply an additional
+data reduction operator to decrease the amount of time spent in IO.
+Alternatively, if multiple different refactored data versions are stored,
+alternative versions can be requested instead.
+
 \section{Metadata Challenges}
 \label{sec:metadata}
+
+The metadata system has a daunting challenge. Traditional metadata systems for
+both distributed and parallel file systems only have to deal with a single name
+space and storage hierarchy. While hierarichal storage management~\cite{hsm}
+systems offer a way to show a single name space, the reality is that the tier
+closest to the user is a cache for the other tiers with data migrating as
+necessary to support user interaction. SIRIUS is working in a fundamentally
+different way. Each tier is treated as a first class citizen with direct
+client access possible. Where data is actually retrieved from is where it is
+actually stored rather than forcing data migration and caching. Systems like
+Sirocco~\cite{sirocco} offer a way to support this model directly and are a
+fundamental part of our system.
+
+The additional challege SIRIUS metadata services face is incorporating
+user-specified attributes including data refactoring functions. Since an
+arbitrary user must be able to access any data written for which they have
+permissions. Having data locked behind data compression that has changed over
+time may result in effectively encrypting data and losing the decryption key.
+
+We also seek to offer the ability to also store multiple versions of the same
+data using different refactoring techniques. This will offer additional options
+for higher precision OR higher performance reading.
 
@@
 \label{sec:conclusion}
+
+Overall, the SIRIUS storage system is ushering in a new generation of both IO
+APIs and storage system managed together. It incorporates additional metadata
+to help accelerate data access and ultimately time to insight. Our initial
+efforts show promise that such a system is both possible and can offer
+attractive performance to better use the supercomputing platform.
 

### 2016-02-05 papers/2016-Supercomputing/paper.tex
initial work on the Supercomputing paper

--- 2015-12-23 papers/2016-HotStorage/paper.tex
@@
 
-\newcommand{\DDT}{D\textsuperscript{2}T~}
-\newcommand{\DDTns}{D\textsuperscript{2}T}
-
 \hyphenation{sub-trans-ac-tion}
@@
 
-\title{A Revolutionary Approach to I/O and Storage for Exascale Scientific Computing}
+\title{Restructuring Data for Exascale Scientific Productivity}
 
@@
 \author{
-%\alignauthor {Jay Lofstead\footnotemark[1], Ivo Jimenez\footnotemark[2],Carlos Maltzahn\footnotemark[2], Quincey Koziol\footnotemark[3], John Bent\footnotemark[4], Eric Barton\footnotemark[5]}\\
-\alignauthor {Jay Lofstead\footnotemark[1], Carlos Maltzahn\footnotemark[2], Scott Klasky\footnotemark[3], Hasan Abbasi\footnotemark[3], Qing Liu\footnotemark[3]\\
-Matthew Curry\footnotemark[1], Mark Ainsworth\footnotemark[5], Manish Parashar\footnotemark[4]}\\
-\footnotemark[1]~~Sandia National Laboratories gflofst@sandia.gov\\
-\footnotemark[2]~~University of California, Santa Cruz carlosm@soe.ucsc.edu\\
-\footnotemark[3]~~Oak Ridge National Laboratory\\
-\footnotemark[4]~~Rutgers University\\
-\footnotemark[5]~~Brown University\\
+\alignauthor {
+Jay Lofstead\footnotemark[1]~~,
+Hasan Abbasi\footnotemark[3]~~,
+Mark Ainsworth\footnotemark[6]~\footnotemark[3]~~,
+Jong Choi\footnotemark[3]~~,
+Matthew Curry\footnotemark[1]~~,\\
+Scott Klasky\footnotemark[3]~\footnotemark[4]~\footnotemark[7]~~,
+Tashin Kurc\footnotemark[9]~~,
+Qing Liu\footnotemark[3]~~,
+Carlos Maltzahn\footnotemark[2]~~,
+Manish Parashar\footnotemark[5]~~,\\
+Norbert Podhorzski\footnotemark[3]~~,
+Feiyi Wang\footnotemark[3]~~,
+Matthew Wolf\footnotemark[7]~~,
+C. S. Chang\footnotemark[8]~~,
+Michael Churchill\footnotemark[8]~~,
+Stephane Ethier\footnotemark[8]~~
+}\\
+\footnotemark[1]~~Sandia National Laboratories,
+\footnotemark[2]~~University of California Santa Cruz,\\
+\footnotemark[3]~~Oak Ridge National Laboratory,
+\footnotemark[4]~~U. Tenn. Knoxville,
+\footnotemark[5]~~Rutgers University,\\
+\footnotemark[6]~~Brown University,
+\footnotemark[7]~~Georgia Tech,
+\footnotemark[8]~~PPPL,
+\footnotemark[9]~~Stony Brook
 }
@@
 
-The longest existing parallel I/O APIs, HDF5, PnetCDF, and NetCDF, are focused
-on treating the entire output from a paralell application as a single unit.
-These libraries work to gather the data from all processes and arrange it such
-that it is organized much like as if it were written from a single process.
-This approach, in particular the two-phase collective I/O with data sieving
-portion, has proven difficult to scale for 3-D domain decompositions at scale.
-ADIOS shifted the focus from treating the entire application output as a single
-entity to treating the output from each writer as a self-contained entity. This
-eliminated the need for the problematic two-phase collective I/O and has
-demonstrated good scalability well into the petascale era. Another shift is
-required to address the needs of exascale systems. New system I/O bandwidth is
-not keeping pace with the compute acceleration leading to accute bottlenecks.
-While ``burst buffer'' technology will help address the performance gap, it
-does not address the problem completely. This paper describes a new approach
-to not just an I/O library, but also the underlying storage infrastructure to
-address the needs of exascacle applications cognizant of projected exascale
-platform characteristics and in the context of current industry trends.
+As the exascale computing age emerges, data related issues are becoming
+critical factors that determine how and where we do computing. Popular
+approaches used by traditional I/O solution and storage libraries become
+increasingly bottlenecked due to their assumptions about data movement,
+re-organization, and storage. While, new technologies, such as ``burst
+buffers'', can help address some of the short-term performance issues, it is
+essential that we reexamine the underlying storage and I/O infrastructure to
+effectively support requirements and challenges at exascale and beyond.
+
+In this paper we present
+- different data needs different techniques
+- write overhead for the different refactoring/auditing approaches
+- derived quantities in viz after precision based refactoring and differences
+  from full fidelity data sets.
+- linear auditing approaches for data refactoring and the effects on data
+  sizes, io performance, and viz quality. Do every point, but split at 24 bits
+and 40 bits.
 
@@
 the storage array.
-Section~\ref{sec:storage}
-discusses the challenges and opportunities for deploying scalable storage
-infrastructure.
-Section~\ref{sec:qos} 
-discusses the quality of service features proposed for this project.
-Next is an exploration of the metadata challenges required to support such a
-system in Section~\ref{sec:metadata}.
-An initial prototype demonstration of the functionality is presented in
+Section~\ref{sec:iof}
+briefly discusses the motivation and proposal for the IO forwarding layer.
+Section~\ref{sec:iod} describes the IO Dispatcher layer and the broad
+functionality it offers.
+Section~\ref{sec:daos} discusses how the DAOS
+layer functions.
+The VOSD layer is discussed in Section~\ref{sec:vosd}.
+Next is an exploration of cross-cutting features like transactions and
+metadata management in Section~\ref{sec:broader}.
+A demonstration of the functionality is presented in
 Section~\ref{sec:evaluation}.
@@
 \label{sec:related}
+
+Many projects over the last couple of decades have sought to address some
+challenging aspect of parallel file system design. The recent rise of ``Big
+Data'' applications with different characteristic IO patterns have somewhat
+complicated the picture. Extreme scale machines will be expected to handle both
+the traditional simulation-related workloads as well as applications more
+squarely in the Big Data arena. This will require some adjustments to the
+underlying system for good performance for both scenarios.
+
+The major previous work is really limited to full file systems rather than the
+mountain of file system refinements made over the years. A selection of these
+other file systems and some features that make it relatively unique are
+described below.
 
@@
 
-\section{End-User API Layer}
-\label{sec:end-user}
-
-\section{Storage Challenges and Opportunities}
-\label{sec:storage}
-
-\section{Quality of Service Features}
-\label{sec:qos}
-
-\section{Metadata Challenges}
-\label{sec:metadata}
-
 \section{Demonstration}
 \label{sec:evaluation}
-
-This stack has an early prototype implementation intended to test concepts
-rather than performance and scalability. It has focused on examining the
-interaction of the different APIs for each layer to flesh out any detailed
-requirements or concerns that may have been missed in the conceptualization of
-this IO stack. To demonstrate the viability of the IO stack described in this
-paper, we show some very early performance results from the untuned prototype.
 
@@
 
+The Fast Forward Storage and IO Stack project has designed a good first pass at
+addressing the requirements for an extreme scale data storage mechanism.  By
+preferring a high level user API like HDF5 rather than using the POSIX
+interface, more advanced functionality can be incorporated with less end-user
+impact. The introduction of the IOD layer with buffering will absorb the
+difference between the compute node IO demands and the available bandwidth in
+the storage array. With DAOS supporting translating the container and object
+model to the underlying storage options, different storage technologies can be
+deployed over time.
+
+With the overall stack design a prototype implementation complete, refinements,
+such as fault detection and recovery, can be designed and tested.  These and
+other activities for phase 2 will ultimately generate what is likely to be the
+next generation storage stack for extreme scale platforms.
+
 \section{Acknowledgements}
 
-\includegraphics[scale=0.07]{logos/doe_logo}
-\includegraphics[scale=0.30]{logos/snl_logo}
-\includegraphics[scale=0.35]{logos/nnsa_logo}
 Sandia National Laboratories is a multi-program laboratory managed and operated

### 2016-02-11 papers/2016-Supercomputing/paper.tex
fixed Tahsin's name spelling.

@@
 Scott Klasky\footnotemark[3]~\footnotemark[4]~\footnotemark[7]~~,
-Tashin Kurc\footnotemark[9]~~,
+Tahsin Kurc\footnotemark[9]~~,
 Qing Liu\footnotemark[3]~~,
@@
 \section{Introduction}
+
+Be sure to address the following:\\
+1. Why should Chang or Chen care from the apps side.\\
+2.why should ken or hank care from the viz side\\
+3.why would Garth care from the storage side\\
 

### 2016-03-02 papers/2016-Supercomputing/paper.tex
added notes from the PLAN.JPEG picture from our in person paper plans

@@
 1. Why should Chang or Chen care from the apps side.\\
-2.why should ken or hank care from the viz side\\
-3.why would Garth care from the storage side\\
+2. why should ken or hank care from the viz side\\
+3. why would Garth care from the storage side\\
 
@@
 %\end{figure}
+
+Contributions:\\
+\begin{enumerate}
+\item XGC1 Byte-split data (3/5)
+\begin{enumerate}
+\item decomposition/compression
+\item auditor
+\item data challenges preventing arbitrary use
+\end{enumerate}
+\item XGC1/S3D Data Interpolation (timestep to timestep)
+\end{enumerate}
+
+Engineering:\\
+\begin{enumerate}
+\item ADIOS write transports
+\item VisIt auditor reader
+\item interfaces
+\end{enumerate}
+
+validation/experiments (XGC1 only for now, 1 & 2 buckets, all other factors):\\
+\begin{enumerate}
+\item Performance writing (A orig, B 2 buckets)
+\item Performance reading (A orig, B 2 buckets, C bucket 1, D, bucket 2)
+\item Visualization diff (3-d)
+\item numeric evaluation diff (to get past criticism of 3-d hiding errors)
+\item derived quantity
+\item interpolation validation
+\end{enumerate}
+
+target: Lustre, DataSpaces\\
+nodes: 10, 100, 1000 (Sith 35 compute + 5 storage)\\
+data/node: 10M, 1G\\
+vars: 10, 100\\
+per var: 1, 2 buckets (4 if we have time)\\
+timesteps: 2, 4, 8, 16\\
+time between timesteps: 10s, 100s\\
+
+XGC1:\\
+1d array format\\
+nparticles (N x 8)\\
+fields (nv) (3-5 vars)\\
+
+Pixie3D/S3D:\\
+3d array: (1k x 1k x 1k)\\
+
+Use ITER mesh input for XGC1
+
+Viz diff
+
+\{reduced, full, diff\} images, numerical eval of diff\\
+
+Responsibilities:
+\begin{itemize}
+\item Jay - write draft/drive paper
+\item everyone - experiment runs
+\item Ivan - DataSpaces help
+\item Tahsin/Jong - Use real XGC1 data in test harness
+\item Tahsin/Jong - Big run for byte split example
+\item Tahsin/Jong - small run for interpolation example
+\item Hadan - Real S3D run
+\item Hasan/Norbert - ADIOS writer
+\item James/Dave - ADIOS reader/viz
+\item Hasan/Ben - ADIOS interpolation pieces
+\end{itemize}
 

### 2016-03-02 papers/2016-Supercomputing/paper.tex
added notes on pland and progress

@@
 
+COMPLETED WORK:\\
+1.
+XGC1 ADIOS write to VisIt read, 3/5, 4/4, 2/6 split (4/4 converts the 4 double bytes to a proper float).
+Output is written to disk and read is directly.
+DataSpaces to do still.
+all scalars are written to both files.\\
+2. Can we do this to Integers too?\\
+3. XGC1 interpolation uses every 4th element only and did 3/5 split\\
+4. Hasan to talk with Jeremy about how to make the interpolation work properly.
+5. Doing off the particles for the first pass. Checking to see if they can use the fields instead.\\
+6. Also measure the diff with the viz and the numerical eval for the 3/5 split\\
+7. Make it neutral in terms of tech used to make double blind possible. Add in
+the rest of the details once accepted\\
+
 The rest of the paper is organized as follows. A brief overview of related work

### 2016-03-05 papers/2016-HotStorage/paper.tex
shifted format to USENIX style and fixed all refs

@@
-\documentclass{sig-alt-gov2}
-\pdfpagewidth=8.5in
-\pdfpageheight=11in
-%\usepackage{subfig}
-\usepackage{subfigure}
-\usepackage[pdftex]{}
-\usepackage{graphicx}
-%\usepackage{todonotes}
-\usepackage{listings}
-\lstset {% general command to set parameter(s)
-         language=C,
-     basicstyle=\footnotesize,               % print whole listing small
-%     keywordstyle=\color{black}\bfseries, % underlined bold black keywords
-%     identifierstyle =\color{black},  % nothing happens
-%     commentstyle=\color{black}\emph, % white comments
-     %stringstyle=\ttfamily,          % typewriter type for strings
-%     stringstyle=\color{black},       % typewriter type for strings
-         tabsize=4,
-         showtabs=false,
-     showstringspaces=false}
-%can't figure this one out for particles bandwidth
-%\usepackage[caption,label]{subfig}
-
-%\newcommand{\TITLE}{D\textsuperscript{\huge{2}}T: Doubly Distributed Transactions for High Performance and Distributed Computing}
-
-\newcommand{\DDT}{D\textsuperscript{2}T~}
-\newcommand{\DDTns}{D\textsuperscript{2}T}
-
-\hyphenation{sub-trans-ac-tion}
-
-\addtolength{\parskip}{-0.08in}
-
-% in order for balance columns to work, it has to be before begin document...
-%\balancecolumns
-
+% This version uses the latex2e styles, not the very ancient 2.09 stuff.
+\documentclass[letterpaper,twocolumn,10pt]{article}
+\usepackage{usenix,epsfig,endnotes,url}
 \begin{document}
 
-%\conferenceinfo{PDSW'15,} {November 17, 2015, Austin, Texas}
-%\CopyrightYear{2015}
-%\crdata{978-1-4503-1103-8/11/11}
-%\clubpenalty=10000
-%\widowpenalty = 10000
+%don't want date printed
+\date{}
 
-\title{SIRIUS: A Revolutionary Approach to I/O and Storage for Exascale Scientific Computing}
+%make title bold and 14 pt font (Latex default is non-bold, 16 pt)
+\title{\Large \bf SIRIUS: A Revolutionary Approach to I/O and Storage for Exascale Scientific Computing}
 
-\numberofauthors{1}
+%for single author (just remove % characters)
 \author{
-%\alignauthor {Jay Lofstead\footnotemark[1], Ivo Jimenez\footnotemark[2],Carlos Maltzahn\footnotemark[2], Quincey Koziol\footnotemark[3], John Bent\footnotemark[4], Eric Barton\footnotemark[5]}\\
-\alignauthor {Jay Lofstead\footnotemark[1], Carlos Maltzahn\footnotemark[2], Scott Klasky\footnotemark[3], Hasan Abbasi\footnotemark[3], Qing Liu\footnotemark[3]\\
-Matthew Curry\footnotemark[1], Mark Ainsworth\footnotemark[5], Manish Parashar\footnotemark[4]}\\
-\footnotemark[1]~~Sandia National Laboratories
-\footnotemark[2]~~University of California\\
-\footnotemark[3]~~Oak Ridge National Laboratory
-\footnotemark[4]~~Rutgers University
-\footnotemark[5]~~Brown University\\
-}
+{\rm Your N.\ Here}\\
+Your Institution
+\and
+{\rm Second Name}\\
+Second Institution
+% copy the following lines to add more authors
+% \and
+% {\rm Name}\\
+%Name Institution
+} % end author
+
+%\author{
+%\alignauthor {Jay Lofstead\footnotemark[1], Carlos Maltzahn\footnotemark[2], Scott Klasky\footnotemark[3], Hasan Abbasi\footnotemark[3], Qing Liu\footnotemark[3]\\
+%Matthew Curry\footnotemark[1], Mark Ainsworth\footnotemark[5], Manish Parashar\footnotemark[4]}\\
+%\footnotemark[1]~~Sandia National Laboratories
+%\footnotemark[2]~~University of California\\
+%\footnotemark[3]~~Oak Ridge National Laboratory
+%\footnotemark[4]~~Rutgers University
+%\footnotemark[5]~~Brown University\\
+%}
+
 \maketitle
 
-\begin{abstract}
+% Use the following at camera-ready time to suppress page numbers.
+% Comment it out when you first submit the paper for review.
+\thispagestyle{empty}
 
+\subsection*{Abstract}
 The longest existing parallel I/O APIs, HDF5, PnetCDF, and NetCDF, are focused
@@
 
-\end{abstract}
-
-%\category{D.4}{Software}{Operating Systems}
-%\category{D.4.7}{Operating Systems}{Organization and Design}[hierarchical design]
-
-%\terms{Design, Performance}
-
 \section{Introduction}
@@
 on serial output from a single process to a distributed or local file system.
-These systems still exist in the form of serial HDF5~\cite{hdf5} and the
-end-of-life NetCDF 3.x~\cite{netcdf}. The key feature of these APIs was the
+These systems still exist in the form of serial HDF5~\cite{chilan:2006:hdf5-performance} and the
+end-of-life NetCDF 3.x~\cite{netcdf3}. The key feature of these APIs was the
 logical data model that eliminated any parallel data decomoposition artifacts
@@
 The second generation is really an evolution of the first with HDF-5 adding
-parallel capabilities and NetCDF bifurcating into NetCDF4~\cite{netcdf4} and
-PnetCDF~\cite{pnetcdf} with PnetCDF almost completely maintaining the data
+parallel capabilities and NetCDF bifurcating into NetCDF4~\cite{rew:2011:netcdf-guide} and
+PnetCDF~\cite{torrellas:2003:pnetcdf} with PnetCDF almost completely maintaining the data
 model and NetCDF4 shifting to use HDF5 underneath. The other improvement is
@@
 expensive and offered a performance penalty in most cases when running at
-scale. Libraries like ADIOS~\cite{adios} shifted from writing complete
+scale. Libraries like ADIOS~\cite{lofstead:2009:adaptable} shifted from writing complete
 variables from distributed processes as a single entity to treating all data
@@
 this extra complication, the performance is generally
-better~\cite{six-degrees}.
+better~\cite{lofstead:2011:six-degrees}.
 
@@
 process without coordinating. A more dynamic approach would be to apply a
-wavelet compression~\cite{wavelet} approach that generates different compressed
+wavelet compression~\cite{klappenecker:1995:wavelet} approach that generates different compressed
 data sizes at a fixed error rate depending on the data entropy. This is still a
@@
 both distributed and parallel file systems only have to deal with a single name
-space and storage hierarchy. While hierarichal storage management~\cite{hsm}
+space and storage hierarchy. While hierarichal storage management~\cite{blaze:1992:hsm}
 systems offer a way to show a single name space, the reality is that the tier
@@
 actually stored rather than forcing data migration and caching. Systems like
-Sirocco~\cite{sirocco} offer a way to support this model directly and are a
+Sirocco~\cite{curry:2015:sirocco} offer a way to support this model directly and are a
 fundamental part of our system.
@@
 
-\section{Acknowledgements}
+%\subsection{How to Build Your Paper}
+%
+%You have to run {\tt latex} once to prepare your references for
+%munging.  Then run {\tt bibtex} to build your bibliography metadata.
+%Then run {\tt latex} twice to ensure all references have been resolved.
+%If your source file is called {\tt usenixTemplate.tex} and your {\tt
+%  bibtex} file is called {\tt usenixTemplate.bib}, here's what you do:
+%{\tt \small
+%\begin{verbatim}
+%latex usenixTemplate
+%bibtex usenixTemplate
+%latex usenixTemplate
+%latex usenixTemplate
+%\end{verbatim}
+%}
+%
 
-\includegraphics[scale=0.07]{logos/doe_logo}
-\includegraphics[scale=0.30]{logos/snl_logo}
-\includegraphics[scale=0.35]{logos/nnsa_logo}
+\section{Acknowledgments}
+
+%\includegraphics[scale=0.07]{logos/doe_logo}
+%\includegraphics[scale=0.30]{logos/snl_logo}
+%\includegraphics[scale=0.35]{logos/nnsa_logo}
 Sandia National Laboratories is a multi-program laboratory managed and operated
@@
 
-\bibliographystyle{abbrv}
-\bibliography{paper}
+{\footnotesize \bibliographystyle{acm}
+\bibliography{paper}}
 
-\vfill\eject
+%\theendnotes
 

### 2016-03-07 papers/2016-HotStorage/paper.tex
added table for reading options

@@
 
-The first generation of standardized APIs for scientific computing IO focused
-on serial output from a single process to a distributed or local file system.
-These systems still exist in the form of serial HDF5~\cite{chilan:2006:hdf5-performance} and the
-end-of-life NetCDF 3.x~\cite{netcdf3}. The key feature of these APIs was the
-logical data model that eliminated any parallel data decomoposition artifacts
-from the actual data storage. This offered consistent performance and regular
-data access patterns.
+High performance computing simulations have driven tightly coupled platform
+scaling for decades. The output from these simulations are analyzed using
+regular tools and techniques optimized for this environment.  The first
+generation of standardized APIs for scientific computing IO focused on serial
+output from a single process to a distributed or local file system.  These
+systems still exist in the form of serial
+HDF5~\cite{chilan:2006:hdf5-performance} and the end-of-life NetCDF
+3.x~\cite{netcdf3}. The key feature of these APIs was the logical data model
+that eliminated any parallel data decomoposition artifacts from the actual data
+storage. This offered consistent performance and regular data access patterns.
 
 The second generation is really an evolution of the first with HDF-5 adding
-parallel capabilities and NetCDF bifurcating into NetCDF4~\cite{rew:2011:netcdf-guide} and
-PnetCDF~\cite{torrellas:2003:pnetcdf} with PnetCDF almost completely maintaining the data
-model and NetCDF4 shifting to use HDF5 underneath. The other improvement is
-incorporating knowledge of an underlying parallel file system to support large
-files written in unision from many processes.
+parallel capabilities and NetCDF bifurcating into
+NetCDF4~\cite{rew:2011:netcdf-guide} and PnetCDF~\cite{torrellas:2003:pnetcdf}
+with PnetCDF almost completely maintaining the data model and NetCDF4 shifting
+to use HDF5 underneath. The other improvement is incorporating knowledge of an
+underlying parallel file system to support large files written in unision from
+many processes.
 
@@
 maintaining the logical data model in physical form became inordinately
-expensive and offered a performance penalty in most cases when running at
-scale. Libraries like ADIOS~\cite{lofstead:2009:adaptable} shifted from writing complete
-variables from distributed processes as a single entity to treating all data
-from a single source to be treated as a unit. This eliminated the data
-reorganization at the cost of more complicated reading patterns. In spite of
-this extra complication, the performance is generally
+expensive~\cite{lofstead:2011:six-degrees} and offered a performance penalty in
+most cases when running at scale. Libraries like
+ADIOS~\cite{lofstead:2009:adaptable} shifted from writing complete variables
+from distributed processes as a single entity to treating all data from a
+single source to be treated as a unit. This eliminated the data reorganization
+at the cost of more complicated reading patterns. In spite of this extra
+complication, the performance is generally
 better~\cite{lofstead:2011:six-degrees}.
@@
 and middleware layer as well as the deeper, more complex storage hierarchy.
-
-Systems like ADIOS showed the performance advantage for decomposed data
-layouts, but still rely on a uniform storage layer. SIRIUS represents a shift
-to consider the local portion of each globally distributed variable as a
+SIRIUS considers the local portion of each globally distributed variable as a
 management entity with data placement decisions across the entire storage
@@
 
-\noindent\textbf{Open and Close} need to add additional information to include
-intent information. The intent hints to the system that....
-
 \noindent\textbf{Refactoring} may be a uniform operation across an
@@
 be read to improve the precision. This can easily be applied uniformly by each
-process without coordinating. A more dynamic approach would be to apply a
-wavelet compression~\cite{klappenecker:1995:wavelet} approach that generates different compressed
-data sizes at a fixed error rate depending on the data entropy. This is still a
-uniform operator, but the resulting data can be of radically different sizes.
+process without coordinating. From a storage hierarchy perspective, the primary
+value portion could be written to high performance, low capacity storage while
+the rest is written to slower storage. That would reduce the data quantity
+written to NVM, for example, by 75\% for a 2/6 split enabling far more output
+to be written to this fast tier with the rest spooled asynchronously.
 
 A more complex, but interesting approach is to offer an {\em Auditor} that
-performs the same computation, but with some simplification that greatly
-reduces computation intensity with only a minor loss of precision over the
-short term. Longer term, this auditor would need to be reset to bring it back
-into alignment with the full precision simulation. The much smaller reduced
+performs a similar, but lower precision computation that greatly reduces
+computation intensity with only a minor loss of precision over short spatial
+and temporal domains. As the simulation progresses, the auditor drift would
+need to be reset to bring it back into alignment. The much smaller, reduced
 precision data could be used instead of the full precision in many
-circumstances.
+circumstances.  A more dynamic lossy approach would be to apply a wavelet
+compression~\cite{klappenecker:1995:wavelet} approach that generates different
+compressed data sizes at a fixed error rate depending on the data entropy.
+These are still uniform operators, but the resulting data can be of radically
+different sizes. For this approach, some or all of the lower precision data
+can be written to the fast tier while full precision can be spooled
+asynchrnously to the slower tier(s).
 
@@
 approach may balance precision better than wavelet compression, but at a higher
-analytics cost.
+analytics cost. We would put data areas with interesting features on the fast
+tier while the rest could be spooled to the slow tier asynchronously.
 
 We plan to investigate all three approaches as different applications will have
-different techniques that work best. Any performant storage system should
-offer the flexibility to incorporate application specific data refactoring
-techniques and sufficient metadata support for an arbitrary client to access
-the data successfully.
+different techniques that work best. Our initial tests have demonstrated the
+potential for the auditor approach in an isolated environment. We have also
+built a prototype ADIOS transport method for the byte-split apporach. Any
+performant storage system should offer the flexibility to incorporate
+application specific data refactoring techniques and sufficient metadata
+support for an arbitrary client to access the data successfully.
 
@@
 Part of that tradeoff will be increased or reduced data quality, depending on
-the system.
+the system. For example, when data is written, a few different forms may be
+output given sufficient space and time. One may be every fourth element
+across a regular mesh reduces the data size to $1/64^th$ the full data set.
+Further dividing that using byte-splitting can reduce it by another 75\%. For
+quick overviews to search data sets, this reduced form can be sufficient
+driving which data sets should be investigated in more detail. The key driving
+factor is time. The applicaiton scientist can read 0.4\% of the full data
+volume with a similar time reduction. SIRIUS will offer a negotiation process
+to read any data.
+
+First, a user will request a particular variable and timesetp. SIRIUS will
+return a table similar to Table~\ref{tab:results}. For a given variable ``A'',
+the end-user can then select the data quality and approximate reading time for
+the operation. In this case, there are 5 options for obtaining the whole data
+set (1/2/3/4, 5/6, 7/8, 9, 10). The first four represent a spatial and byte
+split, 5/6 represent just strided writing, 7/8 represent just a byte-split, 9
+represents a wavelet compressed version, and 10 is the full precision data.
+The user would specify to read a particular data instance that would proceed
+normally.  Quality of Service soft guarantees specified as part of the request
+will result in SIRIUS allocating soft bandwidth reservations to meet the
+request.
+
+\begin{table}[tbp]
+\centering
+\caption{Reading Options for Variable ``A''}
+\label{tab:results}
+\begin{tabular}{|l|l|l|l|l|l|}
+\hline
+\# & Size & Time & Time Err & Refactoring & Data Err\\
+\hline
+1 & $(\frac{1}{4})^3(\frac{3}{8})$A & 10s & $\pm$ 3s & stride, byte-split & 99\% \\
+\hline
+2 & $(\frac{3}{4})^3(\frac{3}{8})$A & 90s & $\pm$ 30s & stride, byte-split & 58\% \\
+\hline
+3 & $(\frac{1}{4})^3(\frac{5}{8})$A & 16s & $\pm$ 5s & stride, byte-split & 0.01\% \\
+\hline
+4 & $(\frac{3}{4})^3(\frac{5}{8})$A & 120s & $\pm$ 50s & stride, byte-split & 0.01\% \\
+\hline
+\hline
+5 & $(\frac{1}{4})^3$A & 1200s & $\pm$ 30s & stride & 98\% \\
+\hline
+6 & $(\frac{3}{4})^3$A & 2400s & $\pm$ 90s & stride & 58\% \\
+\hline
+\hline
+7 & $(\frac{3}{8})$A & 1350s & $\pm$ 120s & byte-split & 5\% \\
+\hline
+8 & $(\frac{5}{8})$A & 2250s & $\pm$ 120s & byte-split & 0.01\% \\
+\hline
+\hline
+9 & A & 36s & $\pm$ 6s & wavelet & 1\% \\
+\hline
+\hline
+10 & A & 3600s & $\pm$ 600s & none & 0\% \\
+\hline
+\end{tabular}
+\end{table}
 

### 2016-03-07 papers/2016-HotStorage/paper.tex
fixed some formatting

@@
 output given sufficient space and time. One may be every fourth element
-across a regular mesh reduces the data size to $1/64^th$ the full data set.
+across a regular mesh reduces the data size to $1/64^{th}$ the full data set.
 Further dividing that using byte-splitting can reduce it by another 75\%. For
@@
 will result in SIRIUS allocating soft bandwidth reservations to meet the
-request.
+request. While some of the error rates seem to render the data unusable,
+consider a 3000x3000x3000 simulation domain. That would yield a 27 gigapixel
+image should all of the data be used. Visualization will already radically
+reduce the data sizes to make usable visualizations that work on a standard
+display. A 99\% reduction would yield a 270 megapixel image--still quite large.
 

### 2016-03-08 papers/2016-Supercomputing/paper.tex
starting intro

@@
 
+  Big data and the associated changes in the computing platform are key
+  components for the advances from exascale systems for scientific discovery
+  with computing. Data will increasingly drive the next generation of scientific
+  discoveries, but limitations in the rate of progress for data storage and
+  movement technologies is likely to hamper this revolution. Storage performance
+  and capacity bottlenecks have the potential to greatly limit how much data can
+  be output from exascale simulations, how often this data can be output, and
+  how much knowledge can be derived from the data.
+
+  New techniques in storage and I/O hardware is one direction towards addressing
+  these issues. In this paper we present an alternate method for addressing the
+  data deluge, by utilizing domain specific knowledge about the data to manage
+  the organization, content and precision of the output data. The tradeoff
+  between system level quality of service and domain specific quality of
+  information can be used to circumvent the storage bottlenecks without
+  exploding the cost or energy usage of storage systems.
+
+  In this paper we present a lossless data partitioning technique that
+  intelligently utilizes hierarchical storage in next-generation leadership
+  computing systems. We show that lower accuracy partitions can be used to
+  service the need for most common use cases, and additional accuracy can be
+  added for more demanding scenarios. The performance benefits are .....
+
+  \begin{comment}
     As the exascale computing age emerges, data related issues are becoming
@@
     and 40 bits.
+  \end{comment}
 
@@
 
+\begin{comment}
   Be sure to address the following:\\
@@
   3. why would Garth care from the storage side\\
+\end{comment}
+
+  Big data and the associated changes in the computing platform are key
+  components for the advances from exascale systems for scientific discovery
+  with computing. Data will increasingly drive the next generation of scientific
+  discoveries, but limitations in the rate of progress for data storage and
+  movement technologies is likely to hamper this revolution. Storage performance
+  and capacity bottlenecks have the potential to greatly limit how much data can
+  be output from exascale simulations, how often this data can be output, and
+  how much knowledge can be derived from the data.
+
+  New techniques in storage and I/O hardware is one direction towards addressing
+  these issues. In this paper we present an alternate method for addressing the
+  data deluge, by utilizing domain specific knowledge about the data to manage
+  the organization, content and precision of the output data. The tradeoff
+  between system level quality of service and domain specific quality of
+  information can be used to circumvent the storage bottlenecks without
+  exploding the cost or energy usage of storage systems.
+
+  We have undertaken this effort as part of a larger rethinking of the role of
+  storage in high performance computing applications, the SIRIUS project. In
+  this paper we demonstrate the initial efforts towards creating precision
+  aware I/O techniques for different types of large scale data sets. Our
+  experiments show that selecting for output accuracy based on the intended
+  usage for data can yeild substantial performance benefits. Moreover, we show
+  that lossless data partitioning techniques can be utilize in lieu of lossy
+  compression methods to utilize the hierarchical storage structure of
+  next-generation leadership computing platforms. By partitioning data into
+  buckets of varying precision, many usecases can be met with smaller, less
+  accurate buckets. For other use cases where only the highest precision will
+  suffice, the additional cost can be paid as and when needed.
 

### 2016-03-10 papers/2016-HotStorage/paper.tex
added Usenix style file and some unknown edits.

@@
 The rest of the paper is organized as follows. A brief overview of related work
-is presented first in Section~\ref{sec:related}. Section~\ref{sec:end-user}
-discusses the programmatic interface end users will see when interacting with
-the storage array.
+is presented first in Section~\ref{sec:related}. Section~\ref{sec:features}
+discusses the differentiating platform features SIRIUS will provide.
 Section~\ref{sec:storage}
@@
 
-\section{End-User API Layer}
-\label{sec:end-user}
+\section{Platform Features}
+\label{sec:features}
 
-The initial thoughts about an end-use API is to try to work within established
-APIs to ease user migration. However, some of the key features, such as the
-data ``refactor'' functionality and the quality of service requirements for
-both writing and reading are new. To manage these new requirements, an extended
-API is proposed. At a high level, the common API broad categories are
-discussed.
+The platform has two key features that drive the rest of the design. First is
+the data refactoring/compression approach. Second are quality of service
+estimates.
 
@@
 operations and reducing the data volumes stored on a per application basis.
+As part of the data refactoring, SIRIUS will use the entire storage hierarchy
+to store different portions of each variable as illustrated in
+Table~\ref{tab:results}. For example, for the 1/2/3/4 setup, 4 different tiers
+may be used to store this single variable.
 

### 2016-03-11 papers/2016-HotStorage/paper.tex
added results and fixed error percentages in the example results table.

@@
 
+The byte splitting approach error rate is quite small. By splitting at byte
+boundaries rather than converting to a lower precision 4-byte float, we
+preserve the full exponent range and keep the first 11 bits of mantissa
+precision. This leaves a $2^{-12}$ maximum error magnitude for each value--the
+last 40 bits represent less precision together than the immediately preceeding
+bit all by itself.
+
 First, a user will request a particular variable and timesetp. SIRIUS will
@@
 \hline
-1 & $(\frac{1}{4})^3(\frac{3}{8})$A & 10s & $\pm$ 3s & stride, byte-split & 99\% \\
+1 & $(\frac{1}{4})^3(\frac{3}{8})$A & 10s & $\pm$ 3s & stride, byte-split & 98\% \\
 \hline
@@
 \hline
-3 & $(\frac{1}{4})^3(\frac{5}{8})$A & 16s & $\pm$ 5s & stride, byte-split & 0.01\% \\
+3 & $(\frac{1}{4})^3(\frac{5}{8})$A & 16s & $\pm$ 5s & stride, byte-split & n/a \\
 \hline
-4 & $(\frac{3}{4})^3(\frac{5}{8})$A & 120s & $\pm$ 50s & stride, byte-split & 0.01\% \\
+4 & $(\frac{3}{4})^3(\frac{5}{8})$A & 120s & $\pm$ 50s & stride, byte-split & n/a \\
 \hline
@@
 \hline
-7 & $(\frac{3}{8})$A & 1350s & $\pm$ 120s & byte-split & 5\% \\
+7 & $(\frac{3}{8})$A & 1350s & $\pm$ 120s & byte-split & 0.02\% \\
 \hline
-8 & $(\frac{5}{8})$A & 2250s & $\pm$ 120s & byte-split & 0.01\% \\
+8 & $(\frac{5}{8})$A & 2250s & $\pm$ 120s & byte-split & n/a \\
 \hline
@@
 
-This stack has an early prototype implementation intended to test concepts
-rather than performance and scalability. It has focused on examining the
-interaction of the different APIs for each layer to flesh out any detailed
-requirements or concerns that may have been missed in the conceptualization of
-this IO stack. To demonstrate the viability of the IO stack described in this
-paper, we show some very early performance results from the untuned prototype.
+Our initial tests are on the Sith cluster at Oak Ridge. The system contains 40
+compute nodes. Each compute node contains four 2.3 GHz 8 core AMD Opteron
+processors, and 64 GB of memory. The system is configured with a 86 TB Lustre
+file system for scratch space.
+
+We use the XGC1~\cite{ku:2008:xgc1} fusion simulation as our application. We
+run on 5 nodes with 160 processes total. XGC1 is configured to use 819,200,000
+particles evenly spread across the processes.
+
+\begin{table}[tbp]
+\centering
+\caption{Performance For 3/5 byte Split}
+\label{tab:performance}
+\begin{tabular}{|l|l|l|l|l|l|}
+\hline
+Test & Time (s)\\
+\hline
+Write full data & 10.3 \\
+\hline
+Byte split & 2.5 \\
+\hline
+write 3-byte portion & 3.4 \\
+\hline
+write 5-byte portion & 9.3 \\
+\hline
+\hline
+Read 3-byte portion & 10.3 \\
+\hline
+Reassemble after read & 27.9 \\
+\hline
+Read full data & 34.2 \\
+\hline
+\end{tabular}
+\end{table}
+
+Evaluating the error when using just the 3-byte portion compared to the full
+data set yields an $L^2$ norm of 72028.2 and $L_{\inf}$ of 0.00109242.
+
+Writing the 3-byte portion synchronously to fast storage and spooling the
+5-byte portion out to slower storage asynchronously will preserve the full data
+fidelity while improving both writing and reading time even when including the
+disassemble/reassemble times. The introduced errors when using just the 3-byte
+portion, as discussed above, is quite small.
 
@@
 
-%\includegraphics[scale=0.07]{logos/doe_logo}
-%\includegraphics[scale=0.30]{logos/snl_logo}
-%\includegraphics[scale=0.35]{logos/nnsa_logo}
+This project is funded by the following research grants from the Advanced
+Scientific Research Office at the Department of Energy, under contracts SDAV:
+ERKJ200/KJ0403000, RSVP: ERKJU60/KJ0402000, and Sirius: ERKJ311/KJ0402000.
 Sandia National Laboratories is a multi-program laboratory managed and operated

### 2016-03-11 papers/2016-HotStorage/paper.tex
added authors list

@@
 %for single author (just remove % characters)
-\author{
-{\rm Your N.\ Here}\\
-Your Institution
-\and
-{\rm Second Name}\\
-Second Institution
+%\author{
+%{\rm Your N.\ Here}\\
+%Your Institution
+%\and
+%{\rm Second Name}\\
+%Second Institution
 % copy the following lines to add more authors
@@
 %Name Institution
-} % end author
+%} % end author
+%\numberofauthors{1}
+\author{
+Jay Lofstead\footnotemark[1]~~,
+Hasan Abbasi\footnotemark[3]~~,
+Mark Ainsworth\footnotemark[6]~\footnotemark[3]~~,
+Jong Choi\footnotemark[3]~~,
+Matthew Curry\footnotemark[1]~~,\\
+Scott Klasky\footnotemark[3]~\footnotemark[4]~\footnotemark[7]~~,
+Tahsin Kurc\footnotemark[9]~~,
+Qing Liu\footnotemark[3]~~,
+Jeremy Logan\footnotemark[4]~~,
+Carlos Maltzahn\footnotemark[2]~~,
+Manish Parashar\footnotemark[5]~~,\\
+Norbert Podhorzski\footnotemark[3]~~,
+Feiyi Wang\footnotemark[3]~~,
+Matthew Wolf\footnotemark[7]~~,
+C. S. Chang\footnotemark[8]~~,
+Michael Churchill\footnotemark[8]~~,
+Stephane Ethier\footnotemark[8]~~
+\\
+\footnotemark[1]~~Sandia National Laboratories,
+\footnotemark[2]~~University of California Santa Cruz,\\
+\footnotemark[3]~~Oak Ridge National Laboratory,
+\footnotemark[4]~~U. Tenn. Knoxville,
+\footnotemark[5]~~Rutgers University,\\
+\footnotemark[6]~~Brown University,
+\footnotemark[7]~~Georgia Tech,
+\footnotemark[8]~~PPPL,
+\footnotemark[9]~~Stony Brook
+}
 

### 2016-03-11 papers/2016-HotStorage/paper.tex
finished paper.

@@
 
-%for single author (just remove % characters)
-%\author{
-%{\rm Your N.\ Here}\\
-%Your Institution
-%\and
-%{\rm Second Name}\\
-%Second Institution
-% copy the following lines to add more authors
-% \and
-% {\rm Name}\\
-%Name Institution
-%} % end author
-%\numberofauthors{1}
 \author{
@@
 
-%\author{
-%\alignauthor {Jay Lofstead\footnotemark[1], Carlos Maltzahn\footnotemark[2], Scott Klasky\footnotemark[3], Hasan Abbasi\footnotemark[3], Qing Liu\footnotemark[3]\\
-%Matthew Curry\footnotemark[1], Mark Ainsworth\footnotemark[5], Manish Parashar\footnotemark[4]}\\
-%\footnotemark[1]~~Sandia National Laboratories
-%\footnotemark[2]~~University of California\\
-%\footnotemark[3]~~Oak Ridge National Laboratory
-%\footnotemark[4]~~Rutgers University
-%\footnotemark[5]~~Brown University\\
-%}
-
 \maketitle
@@
 The longest existing parallel I/O APIs, HDF5, PnetCDF, and NetCDF, are focused
-on treating the entire output from a paralell application as a single unit.
+on treating the entire output from a parallel application as a single unit.
 These libraries work to gather the data from all processes and arrange it such
@@
 demonstrated good scalability well into the petascale era. Another shift is
-required to address the needs of exascale systems. New system I/O bandwidth is
-not keeping pace with the compute acceleration leading to accute bottlenecks.
-While ``burst buffer'' technology will help address the performance gap, it
-does not address the problem completely. This paper describes a new approach
-to not just an I/O library, but also the underlying storage infrastructure to
-address the needs of exascacle applications cognizant of projected exascale
-platform characteristics and in the context of current industry trends.
+required to address the needs of exascale systems. New supercomputing system
+I/O bandwidth is not keeping pace with the compute acceleration leading to
+accute bottlenecks.  While ``burst buffer'' technology will help address the
+performance gap, it does not address the problem completely. This paper
+describes a new approach to not just an I/O library, but also the underlying
+storage infrastructure to address the needs of exascacle applications cognizant
+of projected exascale platform characteristics and in the context of current
+industry trends.
 
@@
 HDF5~\cite{chilan:2006:hdf5-performance} and the end-of-life NetCDF
-3.x~\cite{netcdf3}. The key feature of these APIs was the logical data model
-that eliminated any parallel data decomoposition artifacts from the actual data
-storage. This offered consistent performance and regular data access patterns.
+3.x~\cite{netcdf3}. The key feature of these APIs is the logical data model
+that eliminates any parallel data decomoposition artifacts from the stored
+data.  This offered consistent performance and regular data access patterns.
 
@@
 This paper presents an overview of the efforts to build such a fourth
-generation storage and IO system. A discussion of the end-user API, storage
-challenges, quality of service features, and metadata challenges are all
-discussed briefly.
+generation storage and IO system. A discussion of the platform features,
+storage challenges, quality of service features, and metadata challenges are
+all discussed briefly.
 
@@
 Ceph~\cite{weil:ceph} is a distributed object store and file system. It offers
-both a POSIX and object interface including features typically found in parallel
-file systems.
-Ceph's unique striping approach uses pseudo-random numbers with a
-known seed eliminating the need for the metadata service to track where each
-stripe in a parallel file is placed.
-PVFS~\cite{carns:pvfs} offers optimizations to reduce metadata server load,
-such as a single process opening a file and sharing the handle.
-It has been
-commercialized in recent years as OrangeFS.
-Lustre~\cite{braam:lustre-arch} has become the de facto standard on most major
-clusters offering scalable performance and fine-grained end-user and
-programmatic control over how data is placed in the storage system.
-GPFS~\cite{schmuck:gpfs} offers a hands-off approach for providing good
-performance for scaling parallel IO tasks and is used extensively by its owner,
-IBM.
-Panasas~\cite{panasas:architecture} seeks to offer a dynamically adaptive
-striping system that detects the need for additional stripes for performance
-and adjusts the file layout as necessary.
+both a POSIX and object interface including features typically found in
+parallel file systems.  Ceph's unique striping approach uses pseudo-random
+numbers with a known seed eliminating the need for the metadata service to
+track where each stripe in a parallel file is placed.  PVFS~\cite{carns:pvfs}
+offers optimizations to reduce metadata server load, such as a single process
+opening a file and sharing the handle.  It has been commercialized in recent
+years as OrangeFS.  Lustre~\cite{braam:lustre-arch} has become the de facto
+standard on most major clusters offering scalable performance and fine-grained
+end-user and programmatic control over how data is placed in the storage
+system.  GPFS~\cite{schmuck:gpfs} offers a hands-off approach for providing
+good performance for scaling parallel IO tasks and is used extensively by its
+owner, IBM.  Panasas~\cite{panasas:architecture} seeks to offer a dynamically
+adaptive striping system that detects the need for additional stripes for
+performance and adjusts the file layout as necessary.
 
@@
 HDFS~\cite{Shvachko:2010:hdfs}, address distributed rather than parallel
-computing and cannot be compared directly.
-The primary difference between
+computing and cannot be compared directly.  The primary difference between
 distributed and parallel file systems is the ability of the file system to
@@
 systems rely on a single client creating a file, but distributing the set of
-files across a wide array of storage devices.
-The other, popular distributed
+files across a wide array of storage devices.  The other, popular distributed
 file system of note is NFS~\cite{powlowski:1994:nfs3} that has been used for
@@
 
-\noindent\textbf{Refactoring} may be a uniform operation across an
-entire variable. For example, splitting a double into three parts--two byes,
-four bytes, and two bytes, yields a two-byte value that is the majority of the
-data precision. The next four bytes would be most of the rest of the data
-precision. The last two bytes are the remainder of the precision. For quick
-overviews, just reading the two-byte portion is sufficient to get a rough idea
-of data contents. If particular data looks interesting, more and more data can
-be read to improve the precision. This can easily be applied uniformly by each
-process without coordinating. From a storage hierarchy perspective, the primary
-value portion could be written to high performance, low capacity storage while
-the rest is written to slower storage. That would reduce the data quantity
-written to NVM, for example, by 75\% for a 2/6 split enabling far more output
-to be written to this fast tier with the rest spooled asynchronously.
+\subsection{Refactoring}
+Refactoring may be a uniform operation across an entire variable. For example,
+splitting a double into three parts--two byes, four bytes, and two bytes,
+yields a two-byte value that is the majority of the data precision. The next
+four bytes would be most of the rest of the data precision. The last two bytes
+are the remainder of the precision. For quick overviews, just reading the
+two-byte portion is sufficient to get a rough idea of data contents. If
+particular data looks interesting, more and more data can be read to improve
+the precision. This can easily be applied uniformly by each process without
+coordinating. From a storage hierarchy perspective, the primary value portion
+could be written to high performance, low capacity storage while the rest is
+written to slower storage. That would reduce the data quantity written to NVM,
+for example, by 75\% for a 2/6 split enabling far more output to be written to
+this fast tier with the rest spooled asynchronously.
+
+A similar approach that maintains full data quality is to use a data stride to
+store only a data subset. For example, by selecting every fourth element in
+three dimensions yields a $\frac{1}{64}$ sized data set to store in fast
+storage--a mere 1.5\% of the original size. The rest can be spooled
+asynchronously. Combining byte-splits and striding can reduce the data sizes
+even more. Doing a 3/5 split and an every fourth element yields 0.9\% of the
+original size.
 
@@
 precision data could be used instead of the full precision in many
-circumstances.  A more dynamic lossy approach would be to apply a wavelet
+circumstances.  For an example auditor, consider a previous generation physics
+model. While it is not as accurate as the state-of-the-art-model, it is very
+close, particularly over short temporal and spatial ranges. Other types of
+auditors are possible and are under investigation. The key idea with the
+auditor is that the reduced quality result can be calculated much more quickly
+than the full precision and will have much smaller data sizes. As a side
+benefit, the auditor can both detect soft errors in the caluclation by the
+drift from the high precision physics diverging too quickly. It can also be
+used to create a statistically representative data set for the high precision
+physics model. Calculating the auditor value and writing its output for
+checkpoint restart or initial analysis exploration can save time and leverage
+otherwise idle excess compute resources.
+
+A more dynamic lossy approach would be to apply a wavelet
 compression~\cite{klappenecker:1995:wavelet} approach that generates different
@@
 
-A middle ground approach between these two refactoring approaches is to use a
+A middle ground approach between these refactoring approaches is to use a
 non-uniform operator across a variable. Instead of, for example, splitting on
@@
 approach may balance precision better than wavelet compression, but at a higher
-analytics cost. We would put data areas with interesting features on the fast
-tier while the rest could be spooled to the slow tier asynchronously.
+analytics cost. SIRIUS would put data areas with interesting features on the
+fast tier while the rest could be spooled to the slow tier asynchronously.
 
-We plan to investigate all three approaches as different applications will have
-different techniques that work best. Our initial tests have demonstrated the
-potential for the auditor approach in an isolated environment. We have also
+We plan to investigate all of these approaches as different applications will
+have different techniques that work best. Our initial tests have demonstrated
+the potential for the auditor approach in an isolated environment. We have also
 built a prototype ADIOS transport method for the byte-split apporach. Any
@@
 
-\noindent\textbf{Quality of Service} deadlines are important for driving how
-much time is spent performing IO. At the same time, data quality must be
-maintained at a sufficient level for the resulting science to be possible.
-The system must offer a way for an end-user to negotiate with the system for
-trading off between time spent in IO and the amount of data written or read.
-Part of that tradeoff will be increased or reduced data quality, depending on
-the system. For example, when data is written, a few different forms may be
-output given sufficient space and time. One may be every fourth element
-across a regular mesh reduces the data size to $1/64^{th}$ the full data set.
-Further dividing that using byte-splitting can reduce it by another 75\%. For
-quick overviews to search data sets, this reduced form can be sufficient
-driving which data sets should be investigated in more detail. The key driving
-factor is time. The applicaiton scientist can read 0.4\% of the full data
-volume with a similar time reduction. SIRIUS will offer a negotiation process
-to read any data.
-
-The byte splitting approach error rate is quite small. By splitting at byte
-boundaries rather than converting to a lower precision 4-byte float, we
-preserve the full exponent range and keep the first 11 bits of mantissa
-precision. This leaves a $2^{-12}$ maximum error magnitude for each value--the
-last 40 bits represent less precision together than the immediately preceeding
-bit all by itself.
-
-First, a user will request a particular variable and timesetp. SIRIUS will
-return a table similar to Table~\ref{tab:results}. For a given variable ``A'',
-the end-user can then select the data quality and approximate reading time for
-the operation. In this case, there are 5 options for obtaining the whole data
-set (1/2/3/4, 5/6, 7/8, 9, 10). The first four represent a spatial and byte
-split, 5/6 represent just strided writing, 7/8 represent just a byte-split, 9
-represents a wavelet compressed version, and 10 is the full precision data.
-The user would specify to read a particular data instance that would proceed
-normally.  Quality of Service soft guarantees specified as part of the request
-will result in SIRIUS allocating soft bandwidth reservations to meet the
-request. While some of the error rates seem to render the data unusable,
-consider a 3000x3000x3000 simulation domain. That would yield a 27 gigapixel
-image should all of the data be used. Visualization will already radically
-reduce the data sizes to make usable visualizations that work on a standard
-display. A 99\% reduction would yield a 270 megapixel image--still quite large.
+\subsection{Quality of Service}
+Quality of service deadlines are important for driving how much time is spent
+performing IO. At the same time, data quality must be maintained at a
+sufficient level for the resulting science to be possible.  The system must
+offer a way for an end-user to negotiate with the system for trading off
+between time spent in IO and the amount of data written or read.  Part of that
+tradeoff will be increased or reduced data quality, depending on the system.
+For example, when data is written, a few different forms may be output given
+sufficient space and time. One may be every fourth element across a regular
+mesh reduces the data size to $1/64^{th}$ the full data set.  Further dividing
+that using byte-splitting can reduce it by another 75\%. For quick overviews to
+search data sets, this reduced form can be sufficient driving which data sets
+should be investigated in more detail. The key driving factor is time. The
+applicaiton scientist can read 0.4\% of the full data volume with a similar
+time reduction. SIRIUS will offer a negotiation process to read any data.
 
@@
 
+The byte splitting approach error rate is quite small. By splitting at byte
+boundaries rather than converting to a lower precision 4-byte float, we
+preserve the full exponent range and keep the first 11 bits of mantissa
+precision. This leaves a $2^{-12}$ maximum error magnitude for each value--the
+last 40 bits represent less precision together than the immediately preceeding
+bit all by itself.
+
+To read these various data organizations, read operations will shift to a
+two-step process. First, an list of options will be requested that includes
+both and error estiates. A second request will retrieve a particular data
+portion from the provided list. Consider Table~\ref{tab:results}.  First, a
+user will request a particular variable and timesetp. SIRIUS will return a
+table similar to Table~\ref{tab:results}. For a given variable ``A'',
+the end-user can then select the data quality and approximate reading time for
+the operation. In this case, there are 5 options for obtaining the whole data
+set (1/2/3/4, 5/6, 7/8, 9, 10). The first four entries represent a spatial and
+byte split, 5/6 represent just strided writing, 7/8 represent just a
+byte-split, 9 represents a wavelet compressed version, and 10 is the full
+precision data.  The user would specify to read a particular data instance that
+would proceed normally.  Quality of Service soft guarantees specified as part
+of the request will result in SIRIUS allocating soft bandwidth reservations to
+meet the request. While some of the error rates seem to render the data
+unusable, consider a 3000x3000x3000 simulation domain. That would yield a 27
+gigapixel image should all of the data be used. Visualization will already
+radically reduce the data sizes to make usable visualizations that work on a
+standard display. A 99\% reduction would yield a 270 megapixel image--still
+quite large.
+
 \section{Storage Challenges and Opportunities}
@@
 the end-user can make decisions about skipping the output or maybe spending a
-little time with a strong data refactoring to reduce the data size.
+little time with a strong data refactoring to reduce the data size.  On
+reading, a user will select from a list of available options based on the time
+and data quality requirements.
 
-On reading, a user may specify that they are willing to wait for an approximate
-time for an operation. This will negotiate with the storage system to determine
-the data quality returned. If the requested data quality cannot be retrieved
-within a given time, the storage array may be able to apply an additional
-data reduction operator to decrease the amount of time spent in IO.
-Alternatively, if multiple different refactored data versions are stored,
-alternative versions can be requested instead.
+Key to making QoS work is to maintain performance metadata to inform the IO
+time estimates and error bounds returned to the user. Without hard
+reservations, no system can offer strict performance guarantees. Instead, by
+monitoring system behavior, SIRIUS will be able to offer reasonable time
+estimates given data placement and the requesting source(s).
 
@@
 both distributed and parallel file systems only have to deal with a single name
-space and storage hierarchy. While hierarichal storage management~\cite{blaze:1992:hsm}
-systems offer a way to show a single name space, the reality is that the tier
-closest to the user is a cache for the other tiers with data migrating as
-necessary to support user interaction. SIRIUS is working in a fundamentally
-different way. Each tier is treated as a first class citizen with direct
-client access possible. Where data is actually retrieved from is where it is
-actually stored rather than forcing data migration and caching. Systems like
-Sirocco~\cite{curry:2015:sirocco} offer a way to support this model directly and are a
-fundamental part of our system.
+space and storage tier. While hierarichal storage
+management~\cite{blaze:1992:hsm} systems offer a way to show a single name
+space, the reality is that the tier closest to the user is a cache for the
+other tiers with data migrating as necessary to support user interaction.
+SIRIUS is working in a fundamentally different way. Each tier is treated as a
+first class citizen with direct client access possible. Where data is actually
+retrieved from is where it is actually stored rather than forcing data
+migration and caching. Systems like Sirocco~\cite{curry:2015:sirocco} offer a
+way to support this model directly and are a fundamental part of our system.
 
@@
 run on 5 nodes with 160 processes total. XGC1 is configured to use 819,200,000
-particles evenly spread across the processes.
+particles evenly spread across the processes. Performance numbers are
+presented in Table~\ref{tab:performance}.
 
@@
 efforts show promise that such a system is both possible and can offer
-attractive performance to better use the supercomputing platform.
+attractive performance to better use the supercomputing platform. By treating
+all storage tiers as first-class resources accessed independently, SIRIUS can
+maximize available IO bandwidth. By refactoring data to extract out important
+parts into separate data set slices, data writing and reading time can be
+greatly improved while still maintaining full data fidelity when necessary.
+The quality of service monitoring will inform the performance estimates giving
+a more predictably responsive system accelerating time to insight.
 

### 2016-03-11 papers/2016-Supercomputing/paper.tex
finished paper.

--- 2016-03-02 papers/2016-Supercomputing/paper.tex
@@
 Qing Liu\footnotemark[3]~~,
+Jeremy Logan\footnotemark[4]~~,
 Carlos Maltzahn\footnotemark[2]~~,

### 2016-03-11 papers/2016-HotStorage/paper.tex
fixed typos

@@
 asynchronously. Combining byte-splits and striding can reduce the data sizes
-even more. Doing a 3/5 split and an every fourth element yields 0.9\% of the
+even more. Doing a 3/5 split and an every fourth element yields 0.5\% of the
 original size.
@@
 both distributed and parallel file systems only have to deal with a single name
-space and storage tier. While hierarichal storage
+space and storage tier. While hierarchical storage
 management~\cite{blaze:1992:hsm} systems offer a way to show a single name

### 2016-03-11 papers/2016-HotStorage/paper.tex
fixed typo
fixed L2 norm.

@@
 I/O bandwidth is not keeping pace with the compute acceleration leading to
-accute bottlenecks.  While ``burst buffer'' technology will help address the
+acute bottlenecks.  While ``burst buffer'' technology will help address the
 performance gap, it does not address the problem completely. This paper
@@
 Evaluating the error when using just the 3-byte portion compared to the full
-data set yields an $L^2$ norm of 72028.2 and $L_{\inf}$ of 0.00109242.
+data set yields an $L^2$ norm of 0.000087925 and $L_{\inf}$ of 0.00109242.
 

### 2016-03-30 papers/2016-Supercomputing/paper.tex
intro start
introduction

@@
 \pdfpageheight=11in
-%\usepackage{subfig}
+% \usepackage{subfig}
 \usepackage{subfigure}
@@
 \usepackage{graphicx}
-%\usepackage{todonotes}
+% \usepackage{todonotes}
 \usepackage{listings}
+\usepackage{comment}
 \lstset {% general command to set parameter(s)
@@
   % commentstyle=\color{black}\emph, % white comments
-     %stringstyle=\ttfamily,          % typewriter type for strings
+  % stringstyle=\ttfamily,          % typewriter type for strings
   % stringstyle=\color{black},       % typewriter type for strings
@@
   showstringspaces=false}
-%can't figure this one out for particles bandwidth
-%\usepackage[caption,label]{subfig}
+% can't figure this one out for particles bandwidth
+% \usepackage[caption,label]{subfig}
 
-%\newcommand{\TITLE}{D\textsuperscript{\huge{2}}T: Doubly Distributed Transactions for High Performance and Distributed Computing}
+% \newcommand{\TITLE}{D\textsuperscript{\huge{2}}T: Doubly Distributed Transactions for High Performance and Distributed Computing}
 
@@
 % in order for balance columns to work, it has to be before begin document...
-%\balancecolumns
+% \balancecolumns
 
@@
 
-%\conferenceinfo{PDSW'15,} {November 17, 2015, Austin, Texas}
-%\CopyrightYear{2015}
-%\crdata{978-1-4503-1103-8/11/11}
-%\clubpenalty=10000
-%\widowpenalty = 10000
+% \conferenceinfo{PDSW'15,} {November 17, 2015, Austin, Texas}
+% \CopyrightYear{2015}
+% \crdata{978-1-4503-1103-8/11/11}
+% \clubpenalty=10000
+% \widowpenalty = 10000
 
@@
 
+\end{abstract}
+
+% \category{D.4}{Software}{Operating Systems}
+% \category{D.4.7}{Operating Systems}{Organization and Design}[hierarchical design]
+
+% \terms{Design, Performance}
+
+\section{Introduction}
+
 \begin{comment}
@@
 \end{comment}
-
-\end{abstract}
-
-%\category{D.4}{Software}{Operating Systems}
-%\category{D.4.7}{Operating Systems}{Organization and Design}[hierarchical design]
-
-%\terms{Design, Performance}
-
-\section{Introduction}
 
@@
 
-%\begin{figure}[htbp]
-%\vspace{-0.10in}
-%\centering
-%\includegraphics[width=\columnwidth]{images/arch-mapping}
-%\vspace{-0.20in}
-%\caption{Architecture and Component Mapping}
-%\label{fig:arch-mapping}
-%\vspace{-0.15in}
-%\end{figure}
+This paper describes our initial foray into hierarchical partitioning techniques
+to match the structure of storage in future systems. We have developed three
+sets of techinques for partitioning scientific data from large scale
+applications, evaluated these techniques on our local cluster to gain deeper
+understanding of the cost-benefit tradeoffs with respect to computational time,
+storage time, storage space, access speeds and power consumption, and utilized
+microbenchmarks to extrapolate our results to future use cases. Our research is
+driven by a combination of application user requirements and data consumption
+use cases, such as scientific visualization. To aid in adoption and
+experimentation, we have integrated our techniques with the ADIOS high
+performance I/O middleware library and the Visit visualization application.
+
+The three broad techniques we study in this paper are bit-based splitting of
+double precision floating point data, spatial dimensional sampling of data, and
+a temporal interpolation based data prediction technique. In
+Section~\ref{details} we provide details on each of these techniques and discuss
+the various tradeoffs each technique emobdies. Our evaluation platform presents
+a two level storage hierarchy with local SSDs and a remote lustre parallel file
+system. Both storage targets have different performance characteristics, and
+more details on the evaluation platform are presented in Section~\ref{hardware}.
+With this background we present our results in Section~\ref{performance} and
+discussion in Section~\ref{res-discussion}. This work is the first step in a
+comprehensive new direction for data management at large scale and we discuss
+the future of this effort and conclusions in Section~\ref{future}.
+
+% \begin{figure}[htbp]
+%   \vspace{-0.10in}
+%   \centering
+%   \includegraphics[width=\columnwidth]{images/arch-mapping}
+%   \vspace{-0.20in}
+%   \caption{Architecture and Component Mapping}
+%   \label{fig:arch-mapping}
+%   \vspace{-0.15in}
+% \end{figure}
 
@@
 
-validation/experiments (XGC1 only for now, 1 & 2 buckets, all other factors):\\
+validation/experiments (XGC1 only for now, 1 \&1 2 buckets, all other factors):\\
 \begin{enumerate}
@@
 \end{document}
+
+%%% Local Variables:
+%%% mode: latex
+%%% TeX-master: t
+%%% End:

### 2016-03-30 papers/2016-Supercomputing/paper.tex
outline

@@
 
+In the next section, Section~\ref{motivation}, we present the application use
+case for our data partitioning approach, and a discussion on the storage
+hierarchy in Section~\ref{storage-h}. 
 The three broad techniques we study in this paper are bit-based splitting of
 double precision floating point data, spatial dimensional sampling of data, and
-a temporal interpolation based data prediction technique. In
+a temporal interpolation based data prediction technique. 
+In
 Section~\ref{details} we provide details on each of these techniques and discuss
-the various tradeoffs each technique emobdies. Our evaluation platform presents
+the various tradeoffs each technique emobdies. Other research efforts and
+directions are discussed in Section~\ref{related-work}. 
+Our evaluation platform presents
 a two level storage hierarchy with local SSDs and a remote lustre parallel file
@@
 
-The rest of the paper is organized as follows. A brief overview of related work
-is presented first in Section~\ref{sec:related}. Section~\ref{sec:end-user}
-discusses the programmatic interface end users will see when interacting with
-the storage array.
-Section~\ref{sec:iof}
-briefly discusses the motivation and proposal for the IO forwarding layer.
-Section~\ref{sec:iod} describes the IO Dispatcher layer and the broad
-functionality it offers.
-Section~\ref{sec:daos} discusses how the DAOS
-layer functions.
-The VOSD layer is discussed in Section~\ref{sec:vosd}.
-Next is an exploration of cross-cutting features like transactions and
-metadata management in Section~\ref{sec:broader}.
-A demonstration of the functionality is presented in
-Section~\ref{sec:evaluation}.
-The paper is concluded in
-Section~\ref{sec:conclusion} with a summary of the broad issues covered in the
-paper.
+\section{Motivation and Application Use Cases}
+\label{motivation}
+
+\section{Storage Hierarchy}
+\label{storage-h}
+
+\section{Partitioning Techniques}
+\label{details}
+
+\section{Evaluation Platform}
+\label{hardware}
+
+\section{Evaluation}
+\label{performance}
+
+\section{Cost-Benefit Discussion}
+\label{discussion}
 
 \section{Related Work}
-\label{sec:related}
+\label{related-work}
 
@@
 
-\section{Demonstration}
-\label{sec:evaluation}
+\section{Evaluation}
+\label{performance}
 
-\section{Conclusions}
-\label{sec:conclusion}
-
-The Fast Forward Storage and IO Stack project has designed a good first pass at
-addressing the requirements for an extreme scale data storage mechanism.  By
-preferring a high level user API like HDF5 rather than using the POSIX
-interface, more advanced functionality can be incorporated with less end-user
-impact. The introduction of the IOD layer with buffering will absorb the
-difference between the compute node IO demands and the available bandwidth in
-the storage array. With DAOS supporting translating the container and object
-model to the underlying storage options, different storage technologies can be
-deployed over time.
-
-With the overall stack design a prototype implementation complete, refinements,
-such as fault detection and recovery, can be designed and tested.  These and
-other activities for phase 2 will ultimately generate what is likely to be the
-next generation storage stack for extreme scale platforms.
+\section{Future Work and Conclusions}
+\label{future}
 

### 2016-04-08 papers/2016-Supercomputing/details.tex
Add interpolation details. Add self as author.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper"
%%% End:

Modern experiments and numerical simulations produce datasets so large that their storage and retrieval strain the capabilities of storage hardware.
Slow writing of the data is problematic on shared machines where memory is periodically purged and transferred to new users; slow reading delays analysis of the data after the experiment is over.
Compression offers a solution to this problem.
A crude but commonly used compression technique is decimation, wherein only every \(\stride\)th datum of a sequence is stored.
In practice, the decimation factor \(\stride\), also known as the stride, is often chosen to be \(10\).
The remaining data is simply discarded.
Clearly, decimation trades a reduction in the size of the dataset for data loss.
The loss can be mitigated by generating approximations to the discarded data from the retained data.
A simple, natural technique for doing so is linear interpolation.
Let \(\Data\) be the original dataset.
To generate an approximation \(\Datum[approx]{j}\) to a discarded datum \(\Datum{j}\), we find stored values \(\Datum{m\stride}\) and \(\Datum{m\stride+\stride}\) that straddle \(\Datum{j}\) (\(m\stride<j<m\stride+\stride\)) and take a weighted average:
\begin{equation*}
\Datum{j} \approx \Datum[approx]{j} = (1-\tfrac{j-m\stride}{\stride})\Datum{m\stride}+\tfrac{m\stride+\stride-j}{\stride}\Datum{m\stride+\stride}
.
\end{equation*}
Observe that decimation paired with interpolation, illustrated in Fig.~\ref{fig:InterpolationReconstructionOfDiscardedValues}, gives a compression ratio of \(\stride\) and provides a full set of values at all original timesteps.
\begin{figure}[ht]
\centering
\includegraphics[width=0.45\textwidth]{images/interpolation/illustration_of_procedure}
\caption{Illustration of decimation procedure for partitioning of data in the case \(\stride=5\).}
\label{fig:InterpolationReconstructionOfDiscardedValues}
\end{figure}
This procedure produces a good compression ratio, and it is computationally attractive: for a given \(j\) only two stored values need to be retrieved in order to reconstruct a value for \(\Datum{j}\).
However, as noted above, there is a direct tradeoff between the compression ratio and the data loss.
Nonetheless, despite these obvious shortcomings, the above approach or variants thereof is in widespread use as a technique for compression large datasets, particularly in the context of computational simulation.
In our implementation, the data is partitioned into the \emph{top level data} \(\DataDecimated\), consisting of the decimated data, and the \emph{bottom level data} \(\InterpolationErrors\), consisting of the differences, or \emph{deltas}, between the original data and the approximated data.

Typically, the deltas resulting from decimation paired with interpolation are simply discarded.
The tacit assumption behind this approach is that the approximated dataset \(\Data[approx]\) is a sufficiently good approximation to the full dataset \(\Data\).
The issue of the accuracy of this process has received little or no attention, but the following result is proved in \cite{AinsworthWhitneyKlasky}:
\begin{equation}
\max_{j\in\NN}\lvert\InterpolationError{j}\rvert \leq \min(
2\MaximumDifferenceGlobal{0},
\tfrac{1}{2}\stride\MaximumDifferenceGlobal{1},
\tfrac{1}{8}\stride^{2}\MaximumDifferenceGlobal{2}
)
\label{eqn:InterpolationMagnitudeOfDeltasBound}
\end{equation}
where \(\MaximumDifferenceGlobal{0}=\max_{j\in\NN}\lvert\Datum{j}\rvert\), \(\MaximumDifferenceGlobal{1}=\max_{j\in\NN}\lvert\Datum{j}-\Datum{j+1}\rvert\), and \(\MaximumDifferenceGlobal{2}=\max_{j\in\NN}\lvert\Datum{j}-2\Datum{j+1}+\Datum{j+2}\rvert\).
These bounds are illustrated for a particular dataset in Fig.~\ref{fig:InterpolationDeltaMagnitudeExample}.
Notice that when \(\stride\) is smaller, the accuracy is high, meaning that the top level data produces an acceptable surrogate for the full original dataset.
However, for larger strides (or, equally well, for higher compression ratios), the accuracy becomes unacceptable.
The bottom level data, containing the deltas, enables lossless reconstruction of the data when the accuracy of the approximations is unacceptable for a particular purpose.
\begin{figure}[ht]
\centering
\includegraphics[width=0.45\textwidth]{images/interpolation/delta_sizes}
\caption{%
Illustration of the bounds in Eq.~\eqref{eqn:InterpolationMagnitudeOfDeltasBound} for a range of strides in the case \(\Datum{j}=\sin(2^{-5}j)\) for \(j=0,\ldots,\lfloor2^{6}\pi\rfloor\).
Observe the transition between different cases of Eq.~\eqref{eqn:InterpolationMagnitudeOfDeltasBound} at \(\stride\approx2^{7}\).%
}
\label{fig:InterpolationDeltaMagnitudeExample}
\end{figure}

One might question whether this partitioning procedure provides any overall compression of the original data.
The amount by which a dataset can be compressed is quantified by its Shannon entropy.
One advantage of the decimation--interpolation process is that for reasonably smooth data the entropy of the deltas \(\InterpolationErrors\) will be smaller than that of the original data stream \(\Data\).
Consequently, the deltas are more amenable to compression using an entropy encoder than were the original data.
The following result is proved in \cite{AinsworthWhitneyKlasky}.

\begin{theorem}
\label{thm:InterpolationDifferentialEntropyBounds}
The average differential entropy per delta in \(\InterpolationErrors\) is bounded by the minimum of
\begin{enumerate}[(1)]
\item \(1+\log_{2}\MaximumDifferenceGlobal{0}+\tfrac{1}{\stride-1}\log_{2}\big[\tfrac{1}{6}(\stride+1)(\stride+2)\big]\)
\item \(1+\log_{2}\MaximumDifferenceGlobal{1}\)
\item \(1+\log_{2}\MaximumDifferenceGlobal{2}-\tfrac{1}{\stride-1}\log_{2}\stride\)
\end{enumerate}
\end{theorem}

These entropy bounds can be translated to bounds on the expected compression ratio achieved by the combination of decimation, interpolation, and compression of the deltas.
See \cite{AinsworthWhitneyKlasky} for details and Fig.~\ref{fig:InterpolationCompressionRatioExample} for an illustration.

\begin{figure}[ht]
\centering
\includegraphics[width=0.45\textwidth]{images/interpolation/compression_ratios}
\caption{%
Illustration of the compression ratio bounds derived in \cite{AinsworthWhitneyKlasky} from Theorem~\ref{thm:InterpolationDifferentialEntropyBounds} for a range of strides in the case \(\Datum{j}=2^{4}(3+\sin(2^{-15}j))\) for \(j=0,\ldots,\lfloor2^{16}\pi\rfloor\).%
}
\label{fig:InterpolationCompressionRatioExample}
\end{figure}

### 2016-04-08 papers/2016-Supercomputing/paper.tex
skeleton
Add interpolation details. Add self as author.

@@
 \usepackage{graphicx}
-% \usepackage{todonotes}
 \usepackage{listings}
 \usepackage{comment}
+
+\usepackage{stys/interpolation}
+\usepackage{enumerate}
+
 \lstset {% general command to set parameter(s)
@@
     Matthew Wolf\footnotemark[7]~~,
+    Ben Whitney\footnotemark[6]~~,
     C. S. Chang\footnotemark[8]~~,
@@
 \label{motivation}
+\input{motivation.tex}
 
@@
 \label{storage-h}
+\input{storage.tex}
 
@@
 \label{details}
+\input{details.tex}
 
@@
 \label{hardware}
+\input{hardware.tex}
 
@@
 \label{performance}
+\input{performance.tex}
 
@@
 \label{discussion}
+\input{discussion.tex}
 

### 2016-04-09 papers/2016-Supercomputing/paper.tex
added a first pass at the storage hierarchy section.
Copied the quality of service section directly from the HotStorage paper as a
placeholder. This may move elsewhere in the paper and definitely needs to be
edited to avoid a plagarism issue.

@@
   \alignauthor {
+    Hasan Abbasi\footnotemark[3]~~,
     Jay Lofstead\footnotemark[1]~~,
-    Hasan Abbasi\footnotemark[3]~~,
     Mark Ainsworth\footnotemark[6]~\footnotemark[3]~~,
@@
 \begin{itemize}
-\item Jay - write draft/drive paper
+\item Hasan - write draft/drive paper
 \item everyone - experiment runs
@@
 \item Tahsin/Jong - small run for interpolation example
-\item Hadan - Real S3D run
+\item Hasan - Real S3D run
 \item Hasan/Norbert - ADIOS writer
@@
 \item Hasan/Ben - ADIOS interpolation pieces
+\item Jay/Carlos - Storage hierarchy information
 \end{itemize}

### 2016-04-09 papers/2016-Supercomputing/storage.tex
added a first pass at the storage hierarchy section.
Copied the quality of service section directly from the HotStorage paper as a
placeholder. This may move elsewhere in the paper and definitely needs to be
edited to avoid a plagarism issue.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "paper"
%%% End:

The storage environment targeted by SIRIUS is complex. Based on industry trends
and our own experiences with hardware projections for future HPC platforms, the
current model of adding only a burst buffer tier to the storage hierarchy will
be supplanted with a much more complex setup.

In addition to the new hardware affecting data storage opportunities, the
storage hierarchy will affect how data is later identified, located, and
retrieved. Important considerations are what metadata must be collected and
how the quality of service estimates are provided.

\subsection{Non-Volatile Memory}

The data refactoring tools SIRIUS incorporates assumes the existence of
multiple different storage technologies, each with different performance and
capacity limitations. The most interesting of these is Non-Volatile Memory
(NVM). We see the potential for NVM to be deployed in at least four distinct
machine locations, each with a motivating use case:

\begin{enumerate}

\item {\em Memory Bus} - By placing NVM on the memory bus, fast, local access
is assured. This is ideal for out of core processing workloads. While this is
not a primary use case targeted by SIRIUS, we would take advantage of this
resource, should it be available. The disadvantage of this location is the
difficulty with remote access. Using RDMA-based techniques can simplify
access, even if it still requires a block interface. High bandwidth, on package
RAM can eliminate much of the contention that would otherwise make RDMA use of
the memory-bus based NVRAM impractical.

\item {\em Node-Local SSD} - This data analytics as well as IBM Burst Buffer
model where each node has a local SSD to eliminate network and centralized
storage access overheads for read intensive workloads has other uses for the
HPC environment. As with the Memory Bus-based NVM, Scalable Checkpoint/Restart
(SCM)~\cite{moody:2010:scr} can make effective use of node-local storage for
both local and remote processes. However, SIRIUS aims for a more interesting
use. By acknowledging the future workflow components connected with the
simulation and where the system either has or will deploy those components,
pre-staging data directly where the analytics needs it can be accomplished.
SIRIUS will aid this scenario by reducing data sizes through data refactoring
or other techniques.

\item {\em In Compute Area} - The Cray Burst Buffer model where, for example,
a single node per rack is deployed as a NVM node with significant non-volatile
storage has dual use. The intent for current generation new machines like
Trinity at Sandia and Los Alamos, is to primarily serve as a fast cache for the
parallel file system. The goal is to accelerate checkpoint/restart operations
by making the writing operations as fast as possible while allowing spooling to
the parallel file system as necessary. The secondary use case, and the one
SIRIUS is most interested in, is using these nodes as data staging for workflow
usage. By moving from the node-local SSD model, potential performance impacts
from remote data access can be eliminated. The intereference effects on
inter-node, intra-application communication can also be avoided. Much like the
Node-local SSD case, SIRIUS will store only the important parts of refactored
data in these resources while asynchronously spooling the less important data
to the parallel file system directly. By ensuring direct access to the parallel
file system without going through these Burst Buffer nodes, two different,
simultaneous paths to storage exists. Noticing this extra bandwidth and seeking
an advantageous way to leverage this to help applications was a major impetus
for SIRIUS.

\item {\em In Parallel File System} - By placing NVM in the parallel file
system, the caching nature desired for the Trinity and similar machines can be
achieved, but at a data center level instead. In addition, the metadata
services can use this faster storage to accelerate metadata operations.

\end{enumerate}

With a system that offers NVM in all of these locations, potential bandwidth to
storage will be maximized only when all of these as well as existing
components, such as the parallel file system, can be used simultaneously. By
focusing on storing only the most important parts of a data set on the
expensive, close NVM resources and asynchrnously spooling the rest to slower,
cheaper, higher capacity storage, SIRIUS achieves the best possible storage
bandwidth while also accelerating time to scientific insight.

\subsection{Metadata}

The metadata required falls into three categories: data characteristics, data
location, and infrastructure characteristics.

\subsubsection{Data Characteristics}

Unlike a traditional POSIX-style file system, SIRIUS is using a richer
approach.  In addition to something related to a file name, information about
individual variables including array dimensions, the data sizes, dimensions of
each piece written, and data features are stored. The full range of this data
must be user defined and will differ from application to application. The
initial model for these characteristics are the data characteristics that
currently exist in ADIOS~\cite{lofstead:2009:adaptable}. The extensible nature
of the characteristics in ADIOS will directly relate to those in SIRIUS. The
indexing structure used in ADIOS will also translate directly. Each chunk
stored will have information about what it belongs to and what portion of that
whole it represents.

\subsubection{Data Location}

By spreading a single data set across the entire storage hierarchy, where each
piece of data resides must either be tracked directly at the cost of
maintaining full consistency, or be found later when requested, at the cost of
the search. Which of these approaches or what hybrid using a partiular balance
of parameters will be explored as part of this project.

\subsubsection{Infrastructure Characteristics}

Almost most importnatly for the metadata are the infrastructure
characteristics. As part of providing quality of service, how the system
performs is critical.  As the system is used, each read or write to different
storage resources must be measured to generate a reasonable estimate for future
accesses under similar conditions. This is further explored in the quality of
service subsection.

\subsubsection{Quality of Service}

Quality of service deadlines are important for driving how much time is spent
performing IO. At the same time, data quality must be maintained at a
sufficient level for the resulting science to be possible.  The system must
offer a way for an end-user to negotiate with the system for trading off
between time spent in IO and the amount of data written or read.  Part of that
tradeoff will be increased or reduced data quality, depending on the system.
For example, when data is written, a few different forms may be output given
sufficient space and time. One may be every fourth element across a regular
mesh reduces the data size to $1/64^{th}$ the full data set.  Further dividing
that using byte-splitting can reduce it by another 75\%. For quick overviews to
search data sets, this reduced form can be sufficient driving which data sets
should be investigated in more detail. The key driving factor is time. The
applicaiton scientist can read 0.4\% of the full data volume with a similar
time reduction. SIRIUS will offer a negotiation process to read any data.

\begin{table}[tbp]
\centering
\caption{Reading Options for Variable ``A''}
\label{tab:results}
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\# & Size & Time & Time Err & Refactoring & Data Err\\
\hline
1 & $(\frac{1}{4})^3(\frac{3}{8})$A & 10s & $\pm$ 3s & stride, byte-split & 98\% \\
\hline
2 & $(\frac{3}{4})^3(\frac{3}{8})$A & 90s & $\pm$ 30s & stride, byte-split & 58\% \\
\hline
3 & $(\frac{1}{4})^3(\frac{5}{8})$A & 16s & $\pm$ 5s & stride, byte-split & n/a \\
\hline
4 & $(\frac{3}{4})^3(\frac{5}{8})$A & 120s & $\pm$ 50s & stride, byte-split & n/a \\
\hline
\hline
5 & $(\frac{1}{4})^3$A & 1200s & $\pm$ 30s & stride & 98\% \\
\hline
6 & $(\frac{3}{4})^3$A & 2400s & $\pm$ 90s & stride & 58\% \\
\hline
\hline
7 & $(\frac{3}{8})$A & 1350s & $\pm$ 120s & byte-split & 0.02\% \\
\hline
8 & $(\frac{5}{8})$A & 2250s & $\pm$ 120s & byte-split & n/a \\
\hline
\hline
9 & A & 36s & $\pm$ 6s & wavelet & 1\% \\
\hline
\hline
10 & A & 3600s & $\pm$ 600s & none & 0\% \\
\hline
\end{tabular}
\end{table}

The byte splitting approach error rate is quite small. By splitting at byte
boundaries rather than converting to a lower precision 4-byte float, we
preserve the full exponent range and keep the first 11 bits of mantissa
precision. This leaves a $2^{-12}$ maximum error magnitude for each value--the
last 40 bits represent less precision together than the immediately preceeding
bit all by itself.

To read these various data organizations, read operations will shift to a
two-step process. First, an list of options will be requested that includes
both and error estiates. A second request will retrieve a particular data
portion from the provided list. Consider Table~\ref{tab:results}.  First, a
user will request a particular variable and timesetp. SIRIUS will return a
table similar to Table~\ref{tab:results}. For a given variable ``A'',
the end-user can then select the data quality and approximate reading time for
the operation. In this case, there are 5 options for obtaining the whole data
set (1/2/3/4, 5/6, 7/8, 9, 10). The first four entries represent a spatial and
byte split, 5/6 represent just strided writing, 7/8 represent just a
byte-split, 9 represents a wavelet compressed version, and 10 is the full
precision data.  The user would specify to read a particular data instance that
would proceed normally.  Quality of Service soft guarantees specified as part
of the request will result in SIRIUS allocating soft bandwidth reservations to
meet the request. While some of the error rates seem to render the data
unusable, consider a 3000x3000x3000 simulation domain. That would yield a 27
gigapixel image should all of the data be used. Visualization will already
radically reduce the data sizes to make usable visualizations that work on a
standard display. A 99\% reduction would yield a 270 megapixel image--still
quite large.

### 2016-04-09 papers/2016-Supercomputing/paper.tex
blinded for review

@@
 \numberofauthors{1}
-\author{
-  \alignauthor {
-    Hasan Abbasi\footnotemark[3]~~,
-    Jay Lofstead\footnotemark[1]~~,
-    Mark Ainsworth\footnotemark[6]~\footnotemark[3]~~,
-    Jong Choi\footnotemark[3]~~,
-    Matthew Curry\footnotemark[1]~~,\\
-    Scott Klasky\footnotemark[3]~\footnotemark[4]~\footnotemark[7]~~,
-    Tahsin Kurc\footnotemark[9]~~,
-    Qing Liu\footnotemark[3]~~,
-    Jeremy Logan\footnotemark[4]~~,
-    Carlos Maltzahn\footnotemark[2]~~,
-    Manish Parashar\footnotemark[5]~~,\\
-    Norbert Podhorzski\footnotemark[3]~~,
-    Feiyi Wang\footnotemark[3]~~,
-    Matthew Wolf\footnotemark[7]~~,
-    Ben Whitney\footnotemark[6]~~,
-    C. S. Chang\footnotemark[8]~~,
-    Michael Churchill\footnotemark[8]~~,
-    Stephane Ethier\footnotemark[8]~~
-  }\\
-  \footnotemark[1]~~Sandia National Laboratories,
-  \footnotemark[2]~~University of California Santa Cruz,\\
-  \footnotemark[3]~~Oak Ridge National Laboratory,
-  \footnotemark[4]~~U. Tenn. Knoxville,
-  \footnotemark[5]~~Rutgers University,\\
-  \footnotemark[6]~~Brown University,
-  \footnotemark[7]~~Georgia Tech,
-  \footnotemark[8]~~PPPL,
-  \footnotemark[9]~~Stony Brook
-}
+\author{blinded for review}
+%\author{
+%  \alignauthor {
+%    Hasan Abbasi\footnotemark[3]~~,
+%    Jay Lofstead\footnotemark[1]~~,
+%    Mark Ainsworth\footnotemark[6]~\footnotemark[3]~~,
+%    Jong Choi\footnotemark[3]~~,
+%    Matthew Curry\footnotemark[1]~~,\\
+%    Scott Klasky\footnotemark[3]~\footnotemark[4]~\footnotemark[7]~~,
+%    Tahsin Kurc\footnotemark[9]~~,
+%    Qing Liu\footnotemark[3]~~,
+%    Jeremy Logan\footnotemark[4]~~,
+%    Carlos Maltzahn\footnotemark[2]~~,
+%    Manish Parashar\footnotemark[5]~~,\\
+%    Norbert Podhorzski\footnotemark[3]~~,
+%    Feiyi Wang\footnotemark[3]~~,
+%    Matthew Wolf\footnotemark[7]~~,
+%    Ben Whitney\footnotemark[6]~~,
+%    C. S. Chang\footnotemark[8]~~,
+%    Michael Churchill\footnotemark[8]~~,
+%    Stephane Ethier\footnotemark[8]~~
+%  }\\
+%  \footnotemark[1]~~Sandia National Laboratories,
+%  \footnotemark[2]~~University of California Santa Cruz,\\
+%  \footnotemark[3]~~Oak Ridge National Laboratory,
+%  \footnotemark[4]~~U. Tenn. Knoxville,
+%  \footnotemark[5]~~Rutgers University,\\
+%  \footnotemark[6]~~Brown University,
+%  \footnotemark[7]~~Georgia Tech,
+%  \footnotemark[8]~~PPPL,
+%  \footnotemark[9]~~Stony Brook
+%}
 \maketitle
@@
 
-Sandia National Laboratories is a multi-program laboratory managed and operated
-by Sandia Corporation, a wholly owned subsidiary of Lockheed Martin
-Corporation, for the U.S. Department of Energy's National Nuclear Security
-Administration under contract DE-AC04-94AL85000.
+blinded for review
+
+%Sandia National Laboratories is a multi-program laboratory managed and operated
+%by Sandia Corporation, a wholly owned subsidiary of Lockheed Martin
+%Corporation, for the U.S. Department of Energy's National Nuclear Security
+%Administration under contract DE-AC04-94AL85000.
 
